### 2012-11-18 drafts/proposal.tex
Added proposal files as is.

\documentclass{article}

\usepackage[ruled,vlined,linesnumbered]{algorithm2e}

\usepackage{ amssymb }
\usepackage{listings}
\lstloadlanguages{Haskell}
\lstnewenvironment{code}
    {\lstset{}%
      \csname lst@SetFirstLabel\endcsname}
    {\csname lst@SaveFirstLabel\endcsname}
    \lstset{
      basicstyle=\small\ttfamily,
      flexiblecolumns=false,
      basewidth={0.5em,0.45em},
      literate={+}{{$+$}}1 {/}{{$/$}}1 {*}{{$*$}}1 {=}{{$=$}}1
               {>}{{$>$}}1 {<}{{$<$}}1 {\\}{{$\lambda$}}1
               {\\\\}{{\char`\\\char`\\}}1
               {->}{{$\rightarrow$}}2 {>=}{{$\geq$}}2 {<-}{{$\leftarrow$}}2
               {<=}{{$\leq$}}2 {=>}{{$\Rightarrow$}}2 
               {\ .}{{$\circ$}}2 {\ .\ }{{$\circ$}}2
               {>>}{{>>}}2 {>>=}{{>>=}}2
               {|}{{$\mid$}}1               
    }

\begin{document}

\newcommand{\learner}{\emph{Learner }}
\newcommand{\evaluator}{\emph{Evaluator }}

\section{The Learning Problem}
We want to characterize the general task of the learner, whom we will
call Linda. At various intervals, Linda is provided with sensory data
$x$. In reponse, she is given the opportunity of submitting a program
$\phi$ to an evaluator, henceforth named Dr. Eval. Dr. Eval produces a
reward $R(\phi, x; \bar{y})$ which he delivers back to Linda, where
$\bar{y}$ simply denotes any other variables Linda is not aware of
that influences Dr. Eval's reward. Linda processes the information,
hopefully learning from her experience with Dr. Eval.

What kinds of scenarios can this learning paradigm (naturally)
contain? Most basically, Linda might be a simple organism: the data
$x$ might be local chemical gradients, the program $\phi$ the motor
program Linda will execute upon sensing $x$, and Dr. Eval might return
the cumulative effect upon Linda of executing that motor program. Note
that in this case $\phi$ does not take $x$ as an input. It has no
input; it is merely a motor program built anew for each sensory datum
that Linda receives.

Of course, we might want Linda, the still humble organism, to be
capable of learning a higher-level planning procedure. Such a
procedure would be a program that, given sensory data, would produce a
motor program. In our framework, this would be represented by letting
$x$ stand for a list of data (maybe several days worth of discrete
food finding missions) and letting $\phi$ be the planing
procedure. Dr. Eval would accordingly take $\phi$, apply it to each
datum in $x$ and return a cumulative reward of some sort. We will
defer for now the problem of deferred reward, in which most learners
do not receive immediate signals about the utility of their actions.

Moving upward somewhat in the hierarchy of intelligent beings,
consider Linda the electrical engineer. She is requested to construct
a circuit $\phi$ to a certain specification $x$. Dr. Eval, her
professor tests the circuit against the requested specification and
gives her a grade. Or perhaps she doubles as the evaluator. As in the
case of Linda the simple organism, we can conceive of $\phi$ as the
general circuit creation procedure rather than the circuit, and create
a higher-level learning problem.

A different perspective on this learning problem is one that is
removed from this supervised reinforcement paradigm: instead of
acheiving a reward of some sort, we can imagine that Linda wants to
find the program $\phi$ which generated the data $x$. Here, Dr. Eval
is the world or some simulator of the world and a comparator which
signals to Linda whether she has successfully simulated the data. This
paradigm corresponds to a deterministic version of the generative
model learning task common these days in probabilistic machine
learning and AI.

TODO: concept learning
TODO: physicist (or scientist in general)

\section{A Universal Basis for Computation}
Our goal is to explain the minimal conditions necessary for concept
learning, and to do so we posit a minimal model of
computation. Various universal models of computation have been
proposed; the most famous is the Turing machine. Though the Turing
machine is an intuitive abstract machine, and though its construction
corresponds to various parts of real-world computing machines,
formally simpler models of computation have been proposed. Most
importantly is the Lambda calculus, proposed by Church, and the basis
of modern functional programming languages. The Lambda calculus is a
rewrite system that is provably equally powerful to Turing
machines. Despite its popularity, the implementation of the Lambda
calculus is complicated by the presence of variables bound by the
abstracting $\lambda$s. The necessary relationships between variables
and their binding $\lambda$s means that the syntax of the Lambda
calculus is not entirely modular. That is, if we represent an
expression in the lambda calculus as a tree, then every subtree of this
expression tree is not itself a valid expression tree in the langauge.

An alternative to the $\lambda$ calculus we use the variable-free
model of computation known as Combinatory Logic (CL). CL was
introduced by Shonfinkel in order to avoid of the mathematical
complexities of $\lambda$-calculus. The CL has a simpler syntax than
$lambda$-calculus: a combinator $C$ is merely the application of two
other combinators $ C = A B$. The semantics of the language is defined
by the primitive combinators chosen for the particular instance of
$CL$. It is well known that a universal basis of two primitive
combinators exists, the $SK$ basis. However, expressions in CL are
much more readible if we extend this primitive basis with a few other
high order combinators. Namely, we include identity $I$, and two
composition combinators $B$ and $C$.

\section{History and related literature}
Angluin 1988~\cite{angluin1988queries}: investigates the complexity of
learning concepts that are subsets of a hypothesis space via a number
of primitive queries such as membership, equivalence, subset, etc. \\

Valiant 1984~\cite{valiant1984theory}: original paper on Valiant's
computational learning theory.

Shapiro 1981~\cite{shapiro1981algorithm}: on the model inference
problem. Focused on learning programs or axioms from a set of positive
and negative examples. Explicitly related to Kuhnian philosophy of
science.

Inferring a program for computing a recursive function: 
Gold 65~\cite{gold1965limiting}, 
Gold 67~\cite{gold1967language},
Gold 78 ~\cite{gold1978complexity},
Blum  Blum 75~\cite{blum1975toward}

Inductive Logic Programming (ILP): ILP has a rich history and
literature, and in some sense its goals are quite close to ours. In
recent years, Genetic Programming style techniques have been applied
to ILP; see~\cite{divina2006evolutionary}.

\section{Type Theory}
In using type theory to facilitate program induction we will mostly be
concerned with using the types of subprograms to determine how
subprograms can be glued together. This focus is nicely stated by Girard~\cite{girard1989proofs}:

\begin{quote}
At a more general level, abstracting away from any peculiar syntactic
choice, one should see a type as an instruction for \emph{plugging}
things together. let us imagine that we program with \emph{modules},
\emph{i.e.} closed units, which we can plug together. A module is
absolutely closed, we have no right to open it. We just have the
ability to use it or not, and to choose the manner of use
(plugging). The type of module is a of course completley determined by
all the possibling \emph{pluggings} it allows without crashing. In
parcticular, one can always susbtitute a module with another of the
same type, in the vent of a breakdown, or for the purpose of
optimisation.
\end{quote}

\section{Domains for concept learning}

\section{Coding Theory}
Consider a distribution $P(\cdot)$ over strings in some set $X \in
\{0, 1\}^*$. Our goal is to find a code $D$ that minimizes the
expected length of a sample from $X$.

\section{Learnable Curriculum}
We would like to formally characterize when a program can be learned
that represents elements in a given dataset. In general such programs
always exist; we can simply iterate over all recursive functions until
we find a set that evaluates to each element in the dataset. However,
such a search will usually take a prohibitively long time, so we wish
to restrict our search to short programs.  Thus, we say that a datum
is learnable if a sufficiently short program exists that evaluates to
it. This is simply the statement that the Kolmogorov complexity
$K_U(x)$ of a datum $x$ (relative to our universal computer $U$) is
less than some fixed number of bits $r$.

Moreover, we should be able to use previously learned programs to help
us search for new programs. This phenomenon, in which a previously
learned datum is used to represent a later learned datum, we will term
\emph{bootstrapping}. We say that $x$ \empth{bootstraps} $y$ if the
conditional Kolmogorov complexity $K(y|p_x) < r$ when $U(p_x) = x$ and
$r$ is again our resource bound.

Let $D= \{d_1, \dots, d_N\}$ be a \emph{dataset}. A \emph{curriculum}
$C$ for $D$ is an ordered tuple $C = (c_1, \dots, c_M)$ of elements in
$D$. Not all data $d_i$ must appear in $C$, and each $d_i$ can appear
more than once.

\section{Model of Computation: Combinatory Logic}
The representational language and related computational model used to
implement our learner is a typed Combinatory Logic
(CL)~\cite{hindley1972introduction}. CL is symbolic model of
computation that is similar to the more popular $\lambda$-calculus,
but, for our purposes, has one very significant difference: it does
not contain variables. $\lambda$-calculus has three basic syntactic
features -- variables, applications, and abstractions -- that
correspond appealingly with intuitions from mathematics and functional
programming. But besides the semantic difficulties that variables
cause (documented elsewhere), variables additionally complicate the
modularity and compositionality of expressions written in
$\lambda$-calculus. Program induction requires the efficient search
over expressions in the chosen model of computation; being able to
treat most subexpressions of well-formed expressions as themselves
well-formed expressions makes things easier. This was first noticed by
Briggs and O'Neill in using combinators for genetic
programming~\cite{briggs2006functioinal}.

Example: Consider the function $f = \lambda y. \lambda x. +\,
(*\, x\, x) (*\, x\, x)$ (note that all functions we are considering
will be \emph{curried}; that is, all functions will be functions of
only one argument, and functions of multiple arguments will be
represented by higher-order functions of a single argument such that
the application of the function to an argument returns a function
whose argument is the next argument in the original series of
arguments.)

\section{Intuitive Psychology}
How can we express the knowledge and relationships of intuitive
psychology as a set of expressions in combinatory logic? To simplify
this task, we will attempt to express these relationships in Haskell,
a sugared and polymorphically typed lambda calculus, knowing that any
expression in this lambda calculus can be systematically translated
into combinatory logic. Most importantly, any modularization and
compositional structure present in the Haskell code will remain
present in its CL translation.

Let's imagine the following situation. You observe several agents
(say, Josh, Max, and Chris) in various scenarios in which they choose,
from among the various options, one of three bars (say, a Zone bar, a
Luna bar and a dark chocolate bar). We will try to represent our
knowledge about these agents, their propensities, and the properties
of these bars in Haskell, as a demonstration that such a
representation is both possible in this langauge, and can be somewhat
natural.

Let's first define the primitive types and combinators at our disposal:

\begin{lstlisting}[mathescape]
-- | Primitive Types
-------------------------
Int -- integers 1, 2, etc.
Bool -- {True, False}
Person $\triangleq$ Int
Object $\triangleq$ Int
Scenario  $\triangleq$ (Person, [Object])

-- | Primitive Combinators
------------------------
-- List Combinators:
map :: (a -> b) -> [b] -> [a]
foldl1 :: (a -> a -> a) -> [a] -> a
elem :: a -> [a] -> Bool
head :: [a] -> a
tail :: [a] -> [a]
case :: [(Bool, a)] -- a set of boolean tests and the result if True
        -> a -- the default result (else clause) 
        -> a
(:) :: a -> [a] -> [a] -- cons
(++) :: [a] -> [a] -> [a] -- concatenate

-- Boolean Combinators
If _ Then _ Else :: Bool -> a -> a
And :: Bool -> Bool -> Bool
Or ::  Bool -> Bool -> Bool
Not ::  Bool -> Bool -> Bool

-- Arithmetic Combinators
(>) :: Int -> Int -> Bool
(<) :: Int -> Int -> Bool
(==) :: Int -> Int -> Bool
(+) ::  Int -> Int -> Bool
(-) :: Int -> Int -> Bool
1, 2, $\dots$, 10 :: Int -- | the integers, 1 through 10

-- Givens
------------

Josh :: Person -- | this is just shorthand, i.e., Josh would be mapped
                  to some integer, say, 1, Max to 2, etc..
Max :: Person
Chris :: Person

ZoneBar :: Object
LunaBar :: Object
ChocolateBar :: Object

-- Learned representation
-------------------------
choose' :: Scenario -> Object
choose' (person, objs) = max' (map (util' person) objs)

max' :: [Int] -> Int
max = foldl1 (>)

util' = Person -> Object -> Int
util' person obj = case (utilityCases' person obj) 0 -- | 0 is the default utility in this case

utilityCases' :: [Person -> Object -> (Bool, Int)]
utilityCases' = \p -> \o -> (
                     \p -> \o -> (isHealthyPerson p, healthyPersonCalculus o)

\end{lstlisting}

\section{Algorithm}
In this section, I describe the program learning algorithm. The
algorithm is parameterized by the following parameters:

\begin{itemize}
\item depth $d$: this bounds the search over expressions at each
  iterations to expressions of depth $\leq d$;
\item log likelihood search bound $l$: this bounds the search over
  expressions at each iteration to expressions whose log likelihood is
  $\geq l$;
\item library prior $\mathcal{L}^0$: this data structure contains the
  primitive combinators and associated pseudo-counts that represent
  the background knowledge of the algorithm. This should also include
  any necessary base combinators (e.g. S, K, I, etc.).
\end{itemize}

The inputs to the algorithm are:
\begin{itemize}
\item a dataset $\mathcal{D} = (D_1, \dots, D_M)$, each of which is an
  expression in the language;
\item a evaluator $\mathcal{E}$ which is a function taking a datum
  $D_i$ and an associated expression $\phi_i$ and returns a reward
  signal $\mathcal{E}(D_i, \phi_i)$.
\end{itemize}

\begin{minipage}{.7\linewidth}
  \begin{algorithm}[H]
    \SetAlgoLined 
    \SetKwFunction{EnumerateCombinators}{EnumerateCombinators}
    \SetKwFunction{IsSuccessful}{IsSuccessful}
    \SetKwFunction{filter}{filter}
    \SetKwFunction{FindBestNewCombinator}{FindBestNewCombinator}
    \SetKwFunction{ReEstimateLibrary}{ReEstimateLibrary}
    \KwData{$\mathcal{D}=(D_1, \dots, D_M)$,
      $\mathcal{E}$, $\mathcal{L}^{t-1}$, $d$, $l$} 
    \KwResult{updated library $\mathcal{L}^t$}
    $E \leftarrow $ \EnumerateCombinators{$d$, $l$, $\mathcal{L}^{t-1}$}\;
    $E' \leftarrow $ 
      \filter{$\lambda e.$ \IsSuccessful{$\mathcal{E}$, $\mathcal{D}$, $e$}, $E$}\;
    $C \leftarrow$ \FindBestNewCombinator{$E'$}\;
    $\mathcal{L}^t \leftarrow$ \ReEstimateLibrary{$\mathcal{L}^{t-1}$, $C$, $E'$}\;
    \caption{OneStep}
  \end{algorithm}
\end{minipage}

\section{Language Implementation}
I've found various resources to guide implementation of typed
languages.  A simple guide to the implementation of dependently typed
programming languages is found in Loh et al.~\cite{lowsimply}.

\bibliographystyle{plain}
\bibliography{bibliography}

\end{document}

### 2012-12-09 drafts/CompressionAndSearch.tex
Redone TI and in good state.

\documentclass{article}

\usepackage{amsmath}
\usepackage{amssymb}

\begin{document}

\section{Bootstrapping Concept Learning by Iterated Search and Compression}

It is a commonly held intuition that an intelligent agent should be
able to extrapolate from multiple solutions of simple problems to
probable solutions of more complex ones in the same class. There have
been several attempts in AI and Machine Learning to put this intuition
into practice, but approaches have been generally limited to small
sets of heuristics (EBL) or restricted to vector-space representations
(transfer learning). The common idea, however, is to alter the
representation of an agent's hypothesis space in such a way that
solutions to new problems are more quickly found. Our goal here is to
propose a framework for such extrapolation based learning over
symbolic programs.

Our proposal is simple: a learning agent possesses some initial notion
of the complexity of various hypotheses in her hypothesis space.  When
presented with a set of problems of varying difficulty, she explores a
fixed volume of her hypothesis space in order of increasing
complexity. Having found a set of solutions to some of the problems in
that space, the learner chooses the most compressible set of these
solutions. The resulting compression induces a new complexity measure
on the hypothesis space. As the learner iterates this procedure, the
complexity landscape of the hypothesis space changes, allowing her to
reach solutions to additional problems while maintaining a fixed
search space size.

The optimal compression for distribution over data will necessarily
assign shorter representations to more probable elements and longer
representations to less probable elements. Conversely, any compression
scheme can be interpreted as a statement of belief regarding the
various probabilties of elements in the data space. The best
compression will be that which faithfully assigns representations of
the appropriate lengths under the true distribution of the data; this
is formalized by the well-known Kraft inequality. If we associate with
each element a complexity, measured according to its representation
length in a given compression scheme, then we are claiming that more
probable elements are less complex.

\end{document}

### 2012-12-09 src/PI/Compress.hs

@@ -1,0 +1,92 @@
+-- Compress.hs
+-- Eyal Dechter
+
+-- | This module contains functions for compressing a set of
+-- combinators.
+
+module Compress where
+
+import Control.Monad.State 
+import Debug.Trace
+
+import CL
+import qualified CombTrie as CT
+
+
+-- | Record all unique combinators that satisfy the following pro
+
+type Index = CT.CombTrie Int
+
+-- | When incrementing a tree:
+-- a ) if it has count 0 set count to 1 and increment children
+-- b) if it has count 1 set count to 2 and decrement children
+-- c) if it has count > 2 set count (+1)
+-- | When decrementing a tree:
+-- a) if it has count 1 set count to 0 and decrement children
+-- b) if it has count 2 or greater, set count (-1) and increment children
+
+
+incr :: Index -> Comb -> Index
+incr index c@(CApp l r _) 
+    = case CT.lookup index c of
+        Nothing -> let index' = CT.insert c 1 index
+                   in incr (incr index' l) r
+        Just i -> let index' = CT.insert c (i + 1) index
+                  in case i  of
+                       0 -> incr (incr index' l) r
+                       1 -> decr (decr index' l) r 
+                       otherwise -> index'
+incr index c@(CNode _ _ _) = case CT.lookup index c of
+                               Nothing -> CT.insert c 1 index
+                               Just i -> CT.insert c (i+1) index
+                  
+
+decr :: Index -> Comb -> Index
+decr index c@(CApp l r _) 
+    = case CT.lookup index c of
+        Nothing -> error $  "Cannot find comb " ++ show c
+        Just i -> let index' = CT.insert c (i - 1) index
+                  in case i of 
+                       1 -> decr ( decr index' l) r
+                       otherwise -> incr (incr index' l) r
+decr index c@(CNode _ _ _) = case CT.lookup index c of
+                               Nothing -> error $  "Cannot find comb " ++ show c
+                               Just i -> CT.insert c (i-1) index
+
+compress1  c = incr CT.empty c
+
+compress :: [Comb] -> Index
+compress (c:cs) = incr (compress cs) c
+compress [] = CT.empty
+
+
+-------------------------------
+-- | Alternative compression scheme 
+-- 1) index and count all the unique
+-- subexpressions in the library 
+-- 2) the complexity of a set of
+-- expressions is the number of unique subexpression.
+
+getUniqueTrees :: Comb -> Index
[…]

### 2012-12-09 src/PI/Enumerate.hs

@@ -5,6 +5,7 @@ module Enumerate where
     
 import qualified Data.Map as Map 
 import Control.Monad
+import Control.Monad.State
 import Debug.Trace
 import Text.CSV
 import Data.Maybe
@@ -19,6 +20,13 @@ import qualified CombTrie as CT
 branchTI :: TI [TI a] -> [ TI a ]
 branchTI m = let ms = runTI m
       in map (\n -> m >> n) ms
+
+--branchTIm :: Monad m => TI (m [TI a]) -> m [ TI a ]
+--branchTIm t = let ms = runTI t
+--              in sequence ms
+
+
+
 
 enumCombsToDepth :: [Comb] -> Int -> Type -> [TI Comb]
 enumCombsToDepth lib 0 t = filterCombinatorsByType lib t
@@ -98,6 +106,76 @@ enumNCombsViaIterativeDeepening :: CT.CombTrie Int -- ^ library
                                 -> Type 
                                 -> [TI (Comb, Double)]
 enumNCombsViaIterativeDeepening = undefined
+
+
+
+enumCombsToProbMemo :: CT.CombTrie Int -- ^ library
+                    -> Double -- ^ log prob minimum
+                    -> Int -- ^ max depth
+                    -> Type 
+                    -> [TI (Comb, Double)]
+enumCombsToProbMemo lib ll d tp 
+    = fst $ runState (memoEnum lib ll d tp) (Map.empty)
+
+memoEnum ::  CT.CombTrie Int -- ^ library
+         -> Double -- ^ log prob minimum
+         -> Int -- ^ max depth
+         -> Type 
+         -> State (Map.Map (Type, Int) [TI (Comb, Double)]) [TI (Comb, Double)] 
+memoEnum lib ll 0 tp
+    = do s <- get
+         case Map.lookup (tp, 0) s of
+           Nothing -> do let cs = [ do {x <- c; return (x, ll)} | c <- xs | ll <- lls]
+                         put $ Map.insert (tp, 0) cs s
+                         return $ cs
+           Just cs -> return cs
+      where xs = filterCombinatorsByType (CT.keys lib) tp
+            counts = [ fromJust $ CT.lookup lib x | x <- map runTI xs]
+            sumcounts = sum counts
+            lls  = [ (log (fromIntegral $ fromJust $ CT.lookup lib x)) 
+                     - log (fromIntegral sumcounts) | x <- map runTI xs]
+
+
+memoEnum lib ll d tp 
+    = do s <- get
[…]

### 2012-12-09 src/PI/Experiment.hs

@@ -28,7 +28,7 @@ rlimit = 1000
 
 taskSet :: TaskSet
 taskSet = map (mkSingleEqualityTask rlimit) xs
-    where xs = [i*i | i <- [1..20]]   ++ [ i | i <- [1..4]]
+    where xs = [i |i <-[1..50]] ++ [i*i | i <- [1..50]]
 
 expSquaredInts 
     = Experiment {
@@ -39,9 +39,9 @@ expSquaredInts
         expPrior = stdlibTrie,
         expInitLib = stdlibTrie,
         expLogLikeBound = (-4),
-        expDepthBound = 3,
+        expDepthBound = 4,
         expDataType = Rtype,
-        expReps = 100}
+        expReps = 5}
 
 ------------------------------
 
@@ -56,21 +56,18 @@ eval ds c = sum $ [diff i | i <- rng]
 
 quad a b c = \i -> a * i^2 + b * i + c
 showQuad a b c = show a ++ " x^2 + " ++ show b ++ " x + " ++ show c
-mkQuadTask a b c = Task (showQuad a b c)
-                   f
+mkQuadTask a b c = Task (showQuad a b c) f (Map Rtype  Rtype)
     where f = eval [num2C $ (quad a b c i)| i<-rng]
 
 squares a b = \i -> (a * i + b)^2
 showSquares a b = "(" ++ show a ++ " x + " ++ show b ++ ")^2"
-mkSquaresTask a b = Task (showSquares a b)
-                   f
+mkSquaresTask a b = Task (showSquares a b) f (Map Rtype  Rtype)
+                   
     where f = eval [num2C $ (squares a b) i| i<-rng]
 
 taskSet' = [mkSquaresTask i j | i <- [0..10] , j <- [0..10]]
            ++ [mkQuadTask i j k | i <- [0..10] , j <- [0..10], k <- [0..10]]
-
-
-
+           ++ (concat $ take 3 $ repeat $  map (mkSingleEqualityTask 100) [0..10])
                                     
 expIntegerSequences 
     = Experiment {
@@ -79,11 +76,38 @@ expIntegerSequences
         expEps = 0,
         expPrior = stdlibTrie,
         expInitLib = stdlibTrie,
-        expLogLikeBound = (-6),
-        expDepthBound = 3,
+        expLogLikeBound = (-4),
+        expDepthBound = 4,
         expDataType = (Map Rtype Rtype),
-        expReps=1
+        expReps=20
       }
-        
+ 
+
+----- List Functions ----------
+-- builtins
[…]

### 2012-12-09 src/PI/MemoType.hs

@@ -7,7 +7,6 @@ import Debug.Trace
 import qualified Data.Map as DMap
 
 import Type
-import CL
 
 
 data TypeTrie v = TypeTrie [(Type, v)] (TypeTrie (TypeTrie v)) deriving Show

### 2012-12-09 src/PI/Search.hs

@@ -9,7 +9,7 @@ module Search where
 import qualified Data.Map as Map
 import qualified Data.List as List
 import Data.List.Extras.Argmax (argmaxWithMaxBy, argmax)
-import Data.List (sortBy, foldl')
+import Data.List (sortBy, foldl', nub)
 import Control.Monad.State
 import Control.Monad.Writer
 import Data.Maybe
@@ -26,25 +26,23 @@ import Task
 import Data
 import ParseCL
 import qualified CombTrie as CT
-import qualified Compress as CP (getUniqueTrees)
+import qualified Compress as CP (getUniqueTrees, incr)
 import Compress (Index)
 import Experiment
 
 data SearchLogEntry = SearchLogEntry { searchIter :: Int,
                                        searchLib :: CT.CombTrie Int,
                                        searchSpace :: [Comb],
-                                       searchDataSet :: DataSet,
                                        searchNumHit :: Int,
-                                       searchExplanation :: [(Datum, Comb)],
+                                       searchExplanation :: [(Task, Comb)],
                                        searchCompression :: CT.CombTrie Int
                                      } deriving Show
 mkEmptySearchLog :: SearchLogEntry
-mkEmptySearchLog  = SearchLogEntry 0 CT.empty [] [] 0 [] CT.empty
+mkEmptySearchLog  = SearchLogEntry 0 CT.empty [] 0 [] CT.empty
 
-type Search a = State SearchLogEntry a 
+type Search a = State [SearchLogEntry] a 
 
-
-runSearch :: Search a -> (a, SearchLogEntry)
+runSearch :: Search a -> (a, [SearchLogEntry])
 runSearch s = runState s mkEmptySearchLog 
 
 -- | Sort data by number of combinators matching each
@@ -53,202 +51,123 @@ sortData xs = sortBy
               (\x1 x2 -> (length . snd $ x1) `compare` (length . snd $ x2))
               xs
 
+gen' (x:xs) cs = join [gen' xs (i:cs) | i <- [0..]]
+gen' [] x = [x]
+
+-- | Generate list of solutions
+gen :: [(Task, [Comb])] -> (Index, [(Task, Comb)]) -> [(Index, [(Task, Comb)])] 
+gen ((d, cs):rest) (index, asc) =  concat vs
+    where vs = [gen rest (index `with` CP.getUniqueTrees c, (d, c):asc) | c <- cs]
+          with = CT.mergeWith (+)
+gen [] x = [x]
[…]

### 2012-12-09 src/PI/StdLib.hs

@@ -103,9 +103,9 @@ stdlib = Map.fromList $
          , ("-", dOp2C "-" (-))
          , ("*", dOp2C "*" (*))
          , ("1", num2C 1)
-         , ("2", num2C 2)
-         , ("3", num2C 3)
-         , ("4", num2C 4)
+--         , ("2", num2C 2)
+--         , ("3", num2C 3)
+--         , ("4", num2C 4)
 --           , ("primRec", cPrim) 
 --           , (":", cCONS)
 --           , ("[]", cEmpty)

### 2012-12-09 src/PI/Task.hs

@@ -9,20 +9,21 @@ import Type
 
 type Reward = Double
 data Task =  Task {taskName :: String, 
-                   task :: (Comb -> Reward)} 
+                   task :: (Comb -> Reward),
+                  taskType :: Type} 
 type TaskSet = [Task]
 
 instance Show Task where
-    show (Task n _) = n
+    show (Task n _ _ ) = n
 
 
 -- | task constructors
 
 mkSingleEqualityTask:: Int -> Double -> Task
 mkSingleEqualityTask rlimit i 
-    = let f c = (-1) * (abs $ a - i)
+    = let f c = (abs $ a - i)
               where Just (R a) = reduceWithLimit rlimit $ comb2Expr' c
-      in Task (show i) f
+      in Task (show i) f Rtype
 
     
 

### 2012-12-09 src/PI/Test.hs

@@ -16,12 +16,14 @@ import qualified CombTrie as CT
 import Experiment
 import qualified Compress as CP
 
-xs = fst 
-     $ runSearch 
-     $ findCombinatorsForEachDatum expIntegerSequences stdlibTrie 
-xs' = filter (\x -> (length $ snd x) > 0)  xs
-(index, rs) = greedyN 5 (sortData xs')
+-- xs = fst 
+--      $ runSearch 
+--      $ findCombinatorsForEachDatum expIntegerSequences stdlibTrie 
+-- xs' = filter (\x -> (length $ snd x) > 0)  xs
+-- (index, rs) = greedyN 5 (sortData xs')
+-- index' = newLibrary $ map snd rs
 
+index = fst $ runSearch $ loop expIntegerSequences
 main = do
 --   let lib = stdlibTrie
 --       c = (map runTI $ enumCombsToProb lib (-10) 3 Rtype)
@@ -35,4 +37,5 @@ main = do
 --   putStrLn $ unlines out1
        
    putStrLn $ unlines $ map (\(c, n) -> show' c ++ ": " ++ show n) (CT.toAscList index)
-   putStrLn $ unlines $ map (\(c, n) -> show' n ++ ": " ++ show c) rs
+--   putStrLn $ unlines $ map (\(c, n) -> show' n ++ ": " ++ show c) rs
+--   putStrLn $ unlines $ map (\(c, n) -> show' c ++ ": " ++ show n) (CT.toAscList index')

### 2012-12-09 src/PI/Tests/TestComb.hs

@@ -1,0 +1,19 @@
+-- TestComb.hs
+
+module TestComb where
+
+import Test.QuickCheck
+import Test.QuickCheck.Gen
+import Control.Monad
+
+import CL
+import StdLib (stdlibTrie)
+import qualified CombTrie as CT
+
+instance Arbitrary Comb where
+    arbitrary = sized combgen
+
+combgen 0 = oneof $ map return (CT.keys stdlibTrie)
+combgen n | n < 10 = liftM3 CApp subcomb subcomb arbitrary
+          where subcomb = combgen (n `div` 2)
+combgen n = combgen 5

### 2012-12-09 src/PI/Tests/TestCompress.hs

@@ -1,0 +1,36 @@
+-- TestCompress.hs
+
+module TestCompress where
+
+import Test.QuickCheck
+import Test.HUnit
+import Data.List 
+import Control.Monad
+
+import ParseCL
+import StdLib (stdlib, stdlibTrie)
+import CL
+import Type
+import qualified CombTrie as CT
+import Compress
+
+import TestComb
+
+
+-- combinators
+cI = runTI $ parseExpr stdlib "I"
+cII = runTI $ parseExpr stdlib "I I"
+cIII = runTI $ parseExpr stdlib "I (I I)"
+c4 = runTI $ parseExpr stdlib "(I I) (I I)"
+
+test1 = TestCase $ assertEqual "test 1" a b
+    where a = CT.fromList [(cI, 2), (cII, 2), (cIII, 1)]
+          b = compress [cI, cII, cIII]
+
+prop_orderInvariant :: [Comb] -> Bool
+prop_orderInvariant cs@(c:(c':xs)) = (compress cs)  == compress (reverse cs)
+prop_orderInvariant _ = True
+
+
+
+

### 2012-12-09 src/PI/Tests/TestType.hs

@@ -1,0 +1,52 @@
+-- TestType.hs
+
+import Test.QuickCheck
+import Test.HUnit
+import Control.Monad
+import qualified Data.Set as Set
+
+import Debug.Trace
+
+import Type
+
+
+instance Arbitrary TyVar where
+    arbitrary = liftM TyVar $ liftM getPositive arbitrary
+
+prop_mkTVar :: Int -> Bool
+prop_mkTVar i = i == k
+    where (TVar (TyVar k)) = mkTVar i
+
+instance Arbitrary Type where
+    arbitrary = sized typegen
+
+typegen 0 = oneof [ return Rtype, 
+                    return Btype,
+                    return TyIntList,
+                    liftM TVar arbitrary
+                    ]
+typegen n | n > 0 = oneof [ liftM2 Map subtype subtype,
+                            liftM2 Prod subtype subtype,
+                            liftM2 Sum subtype subtype]
+          where subtype = typegen (n `div` 2)
+
+    
+                    
+prop_applyNullSubst :: Type -> Bool
+prop_applyNullSubst t = (apply nullSubst t) == t
+
+test_tv = TestCase $ assertEqual "test tv" (Set.fromList ts) (Set.fromList (tv t))
+          where ts = [t1,t2,t3]
+                [t1, t2, t3] = map TyVar $ [1,2,3]
+                t = Map (Map (TVar t3) TyIntList) (Prod (TVar t2) (TVar t1))
+
+prop_mergeSameSubst :: Subst -> Bool
+prop_mergeSameSubst s = merge s s == Just (s ++ s)
+
+test_mgu1 = TestCase $  assertEqual "test mgu" 
+            (mgu Rtype Rtype) (Just nullSubst)                
+prop_mgu2 t = 
+    mgu (Map (TVar t) Rtype) (Map Btype Rtype) == Just [(t, Btype)]
+
+prop_type :: Type -> Bool
+prop_type t = (trace $ show t) $ True

### 2012-12-09 src/PI/Type.hs

@@ -12,15 +12,17 @@ module Type where
 import qualified Data.Set as Set
 import Data.List (find, intersect, union)
 import Control.Monad (foldM)
+import Data.MemoTrie
 
 -- -- | local imports
 import CLError
 
 -- | define a type scheme
 
-data TyVar = TyVar Int deriving (Eq, Ord, Show)
 mkTVar :: Int -> Type
 mkTVar i = TVar (TyVar i)
+
+data TyCon = TyCon Int 
 
 data Type = Map {fromType :: Type,
                  toType :: Type} 
@@ -63,6 +65,7 @@ u --> t = [(u, t)]
 
 --  apply a substitution to a type
 apply :: Subst -> Type -> Type
+{-# INLINE apply #-}
 apply s (TVar u) = case lookup u s of
                      Just t -> t
                      Nothing -> TVar u
@@ -82,6 +85,7 @@ tv _ = []
 -- combine two substitutions into one
 infixr 4 @@
 (@@) :: Subst -> Subst -> Subst
+{-# INLINE (@@) #-}
 s1 @@ s2 = [(u, apply s1 t) | (u, t) <- s2] ++ s1
 
 -- merge two substitutions ensuring that both agree on all variables
@@ -215,7 +219,12 @@ freshInst t = foldM f t ts
           f t s = do (TVar u) <- newTVar
                      let t' = apply [(s, TVar u)] t
                      return t'
-                
+
+
+
+
+
+
                                         
                      
 

### 2012-12-09 src/STLC/Expr.hs

@@ -1,0 +1,63 @@
+-- Expr.hs
+
+module Expr where
+
+type Info = Int
+type Context = [(String, Binding)]
+data Binding = NameBind
+
+data Term = TVar {info :: Info -- ^ aux. info
+                 , index :: Int -- ^  de Bruijn index
+                 , ctxsize :: Int
+                 } 
+          | TAbs {info :: Info
+                 , hint :: String
+                 , term :: Term}
+          | TApp {info :: Info
+                 , terml :: Term
+                 , termr :: Term} deriving (Show, Eq)
+
+printTerm :: Context -> Term -> String
+printTerm ctx (TVar inf x n) = if length ctx == n
+                               then show $ indexToName inf ctx  x
+                               else show "[bad index]"
+
+(!!!)                  :: [a] -> Int -> Maybe a
+xs       !!! n | n < 0 =  Nothing
+[]       !!! _         =  Nothing
+(x : _)  !!! 0         =  Just x
+(_ : xs) !!! n         =  xs !!! (n - 1)
+
+indexToName :: Info -> Context -> Int -> String
+indexToName inf ctx i = case ctx !!! i of
+                          Just (name, binding) -> name
+                          Nothing -> "Variable lookup failure."
+
+-- Shifting --
+-- ^ a d-place shift (see Pierce 6.2.1)
+shift :: Int  -- ^ cutoff
+      -> Int -- ^ shift
+      -> Term
+      -> Term
+shift c d t@(TVar _ k _ ) | k < c     =  t{index = k}
+                          | otherwise   =  t{index = k + d}
+shift c d t@(TAbs _ _ t') = t{term = shift (c+1) d t'}
+shift c d t@(TApp _ tl tr) = t{terml = shift c d tl, termr = shift c d tr}
+
+shift1 = shift 0 1
+
+-- Substitute -- 
+subst :: Int -- ^ de Bruijn index
+      -> Term -- ^ replacement
+      -> Term -- ^ body
+      -> Term
+subst j s t@(TVar _ i _) | i == j    = s
+                         | otherwise = t
+subst j s t@(TAbs _ _ t') = t{term = subst (j+1) (shift1 s) t'}
+subst j s t@(TApp _ tl tr) = t{ terml = subst j s tl,
+                                termr = subst j s tr}
+
+
+
+
+                  

### 2012-12-09 src/STLC/Type.hs

@@ -1,0 +1,1 @@
+-- Type.hs

### 2013-01-03 drafts/ijcai2013/ijcai13.tex
Added IJCAI2013 drafts folder.

\documentclass{article}

\usepackage{ijcai13}
\usepackage{times}
\usepackage{amsmath} 
\usepackage{latexsym} 
\usepackage{ dsfont }
\usepackage{tikz}
\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
\usetikzlibrary{arrows,automata}

\DeclareMathOperator*{\argmax}{arg\,max}

\title{The Exploration-Compression Algorithm: how to learn programs by discovering useful subroutines.}
\author{Eyal Dechter \\
MIT\\
USA \\
edechter@mit.edu
\And
Josh Tenenbaum \\
MIT\\
USA \\
jbt@mit.edu
\And 
Ryan Adams \\
MIT\\
USA \\
rpa@seas.harvard.edu}

\begin{document}

\maketitle

\begin{abstract}
  Every programmer knows that the abstractions provided by a good
  function library are not merely a convenience: they usefully
  constrain the space of implementations and guide the choice of
  programming patterns. Without these abstractions, whether explicit
  or implicit, the designer would wander in a sea of
  detail. Similarly, the learner who wishes to discover the theories
  and concepts that explain the world must use a set of abstractions
  to guide their development. But how do those abstractions arise?
  Here, we show how alternating between a search over program space
  and a compression of discovered programs allows a learner to build
  up abstractions that make the learner's search progressively more
  effective.
\end{abstract}

\section{Introduction.}

\section{Related work.}

\section{A multi-task program induction problem.}
Our goal is to solve a general multi-task learning problem: we are
given set of \emph{tasks} $T=\{t_k\}_{k=1}^K$ where each task is a
function $t_i : \mathcal{L} \rightarrow \mathds{R} > 0$ which returns
a reward value for every element in the solution representation
language $\mathcal{L}$. Our goal is to find a set of solutions
$X = \{x_k\}_{k=1}^k \subset \mathcal{L}$ such that

\[
X = \argmax_{x_1, \dots, x_K \in L} \sum_k t_k(x_k).
\]

\section{A compression scheme expresses a belief about the data.}
There is a deep connection between a distribution over data and the
optimally compressive code for that data. In fact, if we know the
probability distribution for a set of observations, we can obtain an
optimally compressive encoding of those observations via arithmetic
coding. Conversely, the encoding generated by a good compression
scheme can be interpreted as an approximation of the data
distribution.

Take, for example, the Nevill-Manning algorithm for grammar based
compression~\cite{nevill1997identifying}. This algorithm looks for
hierarchical structure in the data and represents that hierarchical
structure as a grammar. The output of the algorithm can be seen as a
grammar and a corresponding parse of the input data. Given that parse,
we can estimate the maximum likelihood probabilistic context-free
grammar for that data~\cite{johnson1998pcfg}.

\section{The E-C Algorithm.}

In this section, we describe the E-C algorithm and discuss its uses,
applicability, and limitations.

As its name suggests, each iteration of the E-C algorithm consists of
an exploration step and a compression step. In the exploration step,
we enumerate the $N$ most probable hypotheses given our currenst
distribution $D$ over hypotheses. We then assign to each task $t_i$
the hypotheses that match it. We say that a task is \emph{hit} if
there is a least one hypothesis that matches it. In the compression
step, our goal is to assign a single hypothesis $h_i^*$ to each task
$t_i$ such that that resulting set of hypotheses $\{t_i\}$ is as
compressible as possible. The compression this process induces is used
to update distribution $D$.

  \begin{algorithm}
    \SetAlgoLined \KwData{Tasks $T=\{t_1, \dots, t_K\}$,
      distribution $\mathcal{D}_j$ over expressions.}

    $\text{types} \leftarrow \cup_k \, type(t_k)$. \\
    For $\tau \in \text{ types}$ , $H_N(\tau) \leftarrow$ enumerate
    $N$ most probable elements in $\mathcal{D}_t$ with type $\tau$.\\

    For each \emph{hit} $s_i$ in 
    $T$, $v_{s_i} \leftarrow \{h \in D_j | s_i(h) == True \}$.\\

    $S = \{(s_i, h^*_{s_i})\} \leftarrow$ assignment of elements to tasks s.t. 
    $S$ is most compressible. \\

    $D_{j+1} \leftarrow$ derive new distribution from compression of $S$. \\

    Repeat.

    \caption{E-C}
  \end{algorithm}

\section{Our implementation.}

\subsection{Representing polymorphically typed programs as binary trees.}

Following~\cite{liang10programs} and~\cite{Briggs:2008}, we use a
combinatory logic -- a variable-free subset of the polymorphic
simply-typed lambda calculus lambda calculus~\cite{Pierce_2002} -- as
our program representation language. In short, the combinatory logic
introduces a \emph{basis} of several primitive \emph{combinators} such
that any function in the lambda calculus can be alternatively written
as applications of the basis combinators. The basis combinators can
themselves been expressed in the lambda calculus. The lambda calculus
has two basic operation -- application and abstraction -- but in using
the combinatory logic we sequester uses of the abstraction operation
inside the combinators, and thus have a representation language that
is effectively abstraction- and, thus, variable-free.

This is very convenient for program synthesis: since every expression
is the application of one expression to another -- with this recursion
bottoming out at the primitive combinators -- each program is a binary
tree. Most importantly, any subtree is itself a well-formed
expression; this is not the case in the lambda calculus, since
abstraction introduces long range dependencies between the $lamdba$-
operator and the variables to which that operator refers. In the
lambda calculus, then, a subtree might have free variables, not bound
to any enclosing $lambda$.

\subsection{Enumerating the most likely programs.}
The simplicity of our program representation suggests a simple
recursive recipe for generating all programs of a given type $\tau$ up
to some maximal depth $d$: such a program is either one of the atomic
programs with type in $\tau$, or it is pairing of a left child program
and a right child program with appropriate types each of which comes
from corresponding sets of programs with depth up to $d-1$. The left
child is any program with type $\sigma' \leftarrow \tau'$ which is a
subset of $\sigma \leftarrow \tau$, where $\sigma$ is a free type
variable. To match, the right child must have type $\sigma'$. The
resulting program will have type $\tau'$ which is a subset of $\tau$.

  \begin{algorithm}
    \SetAlgoLined

    \KwData{Library of primitive/type pairs: $\mathcal{L} = \{(f_1, \tau_1), \dots,
        (f_m, \tau_m)\}$. Target type $\tau^*$. Max depth $d$.}

    Choose $F$ from $\text{ ENUM}(\mathcal{L}, \tau^*, 0)$, or:

    Choose $L$ from $\text{ ENUM}( \mathcal{L}, \sigma \rightarrow
    \tau^*, d-1)$.

    Choose $R$ from $\text{ ENUM}( \mathcal{L}, sourceType(L), d-1)$.

    Return $F = (L\,R)$ with type $targetType(L)$.

    \caption{ENUM}
  \end{algorithm}

\subsection{Finding the most compressive solutions.}

\subsection{Re-estimating the probabilistic grammar over programs.}

\section{Experiments.}
\subsection{Integers: a toy case study.}
\subsection{Symbolic regression.}
\subsection{Learning electric circuits.}
\begin{figure}[h]
\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=2cm,
                    , minimum size=.5cm, thick,main node/.style={circle,draw}]

  \node[main node] (S) {$\{\}$};
  \node[main node] (1) [below left of=S] {$\{a\}$};
  \node[main node] (2) [below right of=S] {$\{b\}$} ;
  \node[main node] (3) [below left of=1] {$\{a, a\}$};
  \node[main node] (4) [below right of=1] {$\{a, b\}$};
  \node[main node] (5) [below right of=2] {$\{b, b\}$};;

  \path
    (S) edge node [left] {a} (1)
        edge node [right] {b} (2)
    (1) edge node [left] {a} (3)
        edge node [right] {b} (4)
    (2) edge node [left] {a} (4)
        edge node [right] {b} (5)

\end{tikzpicture}
\end{figure}

%% The file named.bst is a bibliography style file for BibTeX 0.99c
\bibliographystyle{plain}
\bibliography{ijcai13}
\nocite{*}

\end{document}

### 2013-01-18 drafts/ijcai2013/ijcai13.tex
Changes.

@@
 \usepackage{tikz}
+\usepackage{graphicx}
+\usepackage{subcaption}
 \usepackage[ruled,vlined,linesnumbered]{algorithm2e}
-\usetikzlibrary{arrows,automata}
+\usetikzlibrary{arrows,automata, positioning}
 
+\newtheorem{definition}{Definition}
 \DeclareMathOperator*{\argmax}{arg\,max}
 
-\title{The Exploration-Compression Algorithm: how to learn programs by discovering useful subroutines.}
+\title{The E-C Algorithm: conceptual bootstrapping via grammar-based
+  program compression.}
+
 \author{Eyal Dechter \\
@@
 Ryan Adams \\
-MIT\\
+Harvard University\\
 USA \\
@@
 \begin{abstract}
-  Every programmer knows that the abstractions provided by a good
-  function library are not merely a convenience: they usefully
-  constrain the space of implementations and guide the choice of
-  programming patterns. Without these abstractions, whether explicit
-  or implicit, the designer would wander in a sea of
-  detail. Similarly, the learner who wishes to discover the theories
-  and concepts that explain the world must use a set of abstractions
-  to guide their development. But how do those abstractions arise?
-  Here, we show how alternating between a search over program space
-  and a compression of discovered programs allows a learner to build
-  up abstractions that make the learner's search progressively more
-  effective.
+Suppose a learner is faced with a set of problems about which it knows
+nearly nothing: it doesn't know the distribution of problems, or what
+an appropriate loss function might be. For each hypothesized solution
+to a problem it gets a few bits of positive or negative reinforcement,
+but not enough to steer it effectively in the huge space of possible
+solutions. How can such a learner every get off the ground? The key is
+to find, by brute force, a solution to one or more simple problems in
+that domain and use the information thus gained to pursue subsequent
+searches in an informed manner. Here, we formalize this intuition,
+where the solution space is the space of typed functional programs and
+the gained information is stored as a stochastic grammer over
+programs. Specifically, we propose the E-C algorithm, a two step
+iterative procedure for exploring such spaces: in the E step, the
+learner \emph{explores} a finite subset of the domain for solutions,
+guided by a stochastic grammar; in the C step, the learner
+\emph{compresses} the successful solutions from the E step to estimate
+a new stochastic grammar. As we show in several toy domains, iterating
+these two steps quickly propels the learner from a state of solving
+almost no problems to solving a large number of problems by gaining
+abstract knowledge of the structure of the solution space.
 \end{abstract}
@@
 \section{Introduction.}
+
+Suppose that, having been shipwrecked and saved by a people with no
+knowledge of your language or culture, you find yourself attempting to
+communicate in a language about which you know nothing. How would you
+go about discovering even the most basic mappings of sounds to
+meanings? Overhearing the native speech, you have a sense of what the
+phonetics might be like, so you might try various common combinations
+of sounds while pointing to things you want. Mostly, you receive
+puzzled looks, but if you are lucky, you might pick up on a few key
+words or phrases, like ``that'' or ``I want.'' Naturally, you will
+proceed to use your discoveries to accelerate your learning, mixing
+and matching known phrases to generate new meanings.
 
@@
 
-Following~\cite{liang10programs} and~\cite{Briggs:2008}, we use a
+Following~\cite{liang10programs} and~\cite{Briggs:2008}, we use
 combinatory logic -- a variable-free subset of the polymorphic
@@
 that any function in the lambda calculus can be alternatively written
-as applications of the basis combinators. The basis combinators can
+as applications of the basis combinators. Some common basis
+combinators are defined as follows:
+\begin{align}
+I\; x &\rightarrow  x \\
+S\; f\; g\; x\; &\rightarrow (f\; x)\; (g\; x) \\ 
+B\; f\; g\; x\; &\rightarrow (f\; x)\; g \\ 
+C\; f\; g\; x\; &\rightarrow f\; (g\; x) \\ 
+\end{align}
+
+The basis combinators can
 themselves been expressed in the lambda calculus. The lambda calculus
@@
 expression; this is not the case in the lambda calculus, since
-abstraction introduces long range dependencies between the $lamdba$-
+abstraction introduces long range dependencies between the $\lambda$
 operator and the variables to which that operator refers. In the
 lambda calculus, then, a subtree might have free variables, not bound
-to any enclosing $lambda$.
+to any enclosing $\lambda$.
 
-\subsection{Enumerating the most likely programs.}
-The simplicity of our program representation suggests a simple
-recursive recipe for generating all programs of a given type $\tau$ up
-to some maximal depth $d$: such a program is either one of the atomic
-programs with type in $\tau$, or it is pairing of a left child program
-and a right child program with appropriate types each of which comes
-from corresponding sets of programs with depth up to $d-1$. The left
-child is any program with type $\sigma' \leftarrow \tau'$ which is a
-subset of $\sigma \leftarrow \tau$, where $\sigma$ is a free type
-variable. To match, the right child must have type $\sigma'$. The
-resulting program will have type $\tau'$ which is a subset of $\tau$.
+As a simple example of our representation, consider the squaring
+function $\lamba x. * x x $. Using two of the basis combinators above,
+we can write the squaring function in combinatory logic as $S * I$,
+where application associates to the left, so we drop the
+parentheses. When we apply this combinator to a value $x$, the action
+of the combinator is defined as $S \;*\; I\; x~\rightarrow~(*\; x)\;
+(I\; x)~\rightarrow(*\; x)\; x$. Now, we can extend this
+representation with a simple polymorphic type
+system~\cite{Pierce_2002}. Thus, this combinator can be represented as
+a binary tree~(Figure~\ref{fig:clbintree}) whose leaves are typed
+primitive combinators and whose interior nodes are typed applications of combinators.
 
-  \begin{algorithm}
-    \SetAlgoLined
+\begin{figure}
+  \centering
+  \begin{subfigure}{0.2\linewidth}
+    \label{fig:clbintree}
+    \begin{tikzpicture}[scale=.5, transform shape, level/.style={sibling distance=10mm/#1, font=\scriptsize}]
+      \node [font=\scriptsize] (top) {$I$}
+      child {node [above=.3cm] (a) {$(I \rightarrow I) \rightarrow I \rightarrow I$}
+        child {node [left=.5cm, below=1cm, text width=3cm] 
+          {$S : \\(t_2 \rightarrow t_1 \rightarrow t_0)  \rightarrow (t_2 \rightarrow t_1) 
+            \rightarrow \\ t_2 \rightarrow t_0$}}
+        child {node [right=0cm, text width=2cm] {$*: Int \rightarrow Int \rightarrow Int$}}
+      }
+      child {node [below right=.3cm and 1cm] (b) {$I \rightarrow I$}
+        child {node {$I: t_0 \rightarrow t_0 \rightarrow t_0$}}
+      };
 
-    \KwData{Library of primitive/type pairs: $\mathcal{L} = \{(f_1, \tau_1), \dots,
-        (f_m, \tau_m)\}$. Target type $\tau^*$. Max depth $d$.}
+    \end{tikzpicture}
+  \end{subfigure}
 
-    Choose $F$ from $\text{ ENUM}(\mathcal{L}, \tau^*, 0)$, or:
+  \begin{subfigure}[b]{0.2\linewidth}
+    \label{fig:andor}
+    \begin{tikzpicture}[scale=.5, transform shape, level/.style={sibling distance=10mm/#1}]
+      \node [circle,draw] (a) {}
+      child {node [left=0cm] (b) {$f_1$}}
+      child {node [left=.2cm] (c) {\dots} edge from parent[draw=none] }
+      child {node [left=.4cm] (e) {$f_n$}}
+      child {node [circle,draw, left=1cm] (d) {}
+        child {node [circle, draw, left=.25cm] (l) {} edge from parent[draw=none]
+          child {node [left=.5cm] (b) {$f_1$}} 
+          child {node [left=.3cm] (c) {\dots} edge from parent[draw=none] }
+          child {node [left=.3cm] (e) {$f_n$}}
+          child {node [circle, draw, left=.25cm] (and2) {}
+            child {node [circle, draw, left=.25cm](l2) {} edge from parent[draw=none] 
+              child {node [above=.5cm] {\vdots} edge from parent[draw=none]}
+            }
+            child {node [circle, draw] (r2)  {} edge from parent[draw=none] 
+              child {node [above=.5cm] {\vdots} edge from parent[draw=none]}
+            }
+          }
+        }
+        child {node [circle, draw, right=.25cm] (r) {} edge from parent[draw=none]
+          child {node [circle, draw, right=.25cm] (and3) {}
+            child {node [circle, draw ](l3) {} edge from parent[draw=none] 
+              child {node [above=.5cm] {\vdots} edge from parent[draw=none]}
+            }
+            child {node [circle, draw, right=.2cm ] (r3)  {} edge from parent[draw=none] 
+              child {node [above=.5cm] {\vdots} edge from parent[draw=none]}
+            }
+          }
+          child {node [right=.3cm] (b) {$f_1$}} 
+          child {node [right=.4cm] (c) {\dots} edge from parent[draw=none] }
+          child {node [right=.6cm] (e) {$f_n$}}
+        }
+      };
 
-    Choose $L$ from $\text{ ENUM}( \mathcal{L}, \sigma \rightarrow
-    \tau^*, d-1)$.
+      \path (d)  edge node[inner sep=0mm,pos=.4] (a1) {} node {} (l);
+      \path (d)  edge node[inner sep=0mm,pos=.4] (b1) {} node {} (r);
+      \path (a1) edge[bend right] (b1) ;
 
-    Choose $R$ from $\text{ ENUM}( \mathcal{L}, sourceType(L), d-1)$.
+      \path (and2)  edge node[inner sep=0mm,pos=.4] (t0) {} node {} (l2);
+      \path (and2)  edge node[inner sep=0mm,pos=.4] (k0) {} node {} (r2);
+      \path (t0) edge[bend right] (k0) ;
 
-    Return $F = (L\,R)$ with type $targetType(L)$.
+      \path (and3)  edge node[inner sep=0mm,pos=.4] (t1) {} node {} (l3);
+      \path (and3)  edge node[inner sep=0mm,pos=.4] (k1) {} node {} (r3);
+      \path (t1) edge[bend right] (k1) ;
+    \end{tikzpicture}
 
-    \caption{ENUM}
-  \end{algorithm}
+    \end{subfigure}
+  \caption{A typed binary tree representation of the combinator $S\;*\;I\;$.}
+\end{figure}
 
-\subsection{Finding the most compressive solutions.}
+\subsection{Best-first enumeration of most probably programs.}
+In order to explore the frontier of most probable programs, we need a
+procedure that enumerates the $N$ most probable programs. We can
+formulate this naturally as a best-first exploration of an $AND-OR$
+tree (NEED FIGURE). Complete programs are represented by binary trees
+that are valid subtrees of the $AND-OR$ graph. The tree structure is
+shown in Figure~\ref{fig:andor}. The root is an $OR$ node of the
+requisite program type. It's children are the primitives of that type
+and one $AND$ node with two children. Each child has the same
+structure as the root node, except that the types are different. If
+the type of an $OR$ node is $t$, then the type of the left child of
+its $AND$ node child is $\tau \rightarrow t$ (where $\tau$ is a fresh
+type variable). Since the type of the right child is constrained by
+the subtree rooted at the left child, we always expand the left child
+first, and we use the target type of this left subtree to mark the
+type of the right child.
 
-\subsection{Re-estimating the probabilistic grammar over programs.}
+\begin{figure}
+\label{fig:andor}
+\begin{tikzpicture}[scale=.5, transform shape, level/.style={sibling distance=10mm/#1}]
+\node [circle,draw] (a) {}
+child {node [left=0cm] (b) {$f_1$}}
+child {node [left=.2cm] (c) {\dots} edge from parent[draw=none] }
+child {node [left=.4cm] (e) {$f_n$}}
+child {node [circle,draw, left=1cm] (d) {}
+  child {node [circle, draw, left=.25cm] (l) {} edge from parent[draw=none]
+    child {node [left=.5cm] (b) {$f_1$}} 
+    child {node [left=.3cm] (c) {\dots} edge from parent[draw=none] }
+    child {node [left=.3cm] (e) {$f_n$}}
+    child {node [circle, draw, left=.25cm] (and2) {}
+          child {node [circle, draw, left=.25cm](l2) {} edge from parent[draw=none] 
+            child {node [above=.5cm] {\vdots} edge from parent[draw=none]}
+          }
+          child {node [circle, draw] (r2)  {} edge from parent[draw=none] 
+            child {node [above=.5cm] {\vdots} edge from parent[draw=none]}
+          }
+    }
+  }
+  child {node [circle, draw, right=.25cm] (r) {} edge from parent[draw=none]
+    child {node [circle, draw, right=.25cm] (and3) {}
+          child {node [circle, draw ](l3) {} edge from parent[draw=none] 
+            child {node [above=.5cm] {\vdots} edge from parent[draw=none]}
+          }
+          child {node [circle, draw, right=.2cm ] (r3)  {} edge from parent[draw=none] 
+            child {node [above=.5cm] {\vdots} edge from parent[draw=none]}
+          }
+    }
+    child {node [right=.3cm] (b) {$f_1$}} 
+    child {node [right=.4cm] (c) {\dots} edge from parent[draw=none] }
+    child {node [right=.6cm] (e) {$f_n$}}
+ }
+};
+
+\path (d)  edge node[inner sep=0mm,pos=.4] (a1) {} node {} (l);
+\path (d)  edge node[inner sep=0mm,pos=.4] (b1) {} node {} (r);
+\path (a1) edge[bend right] (b1) ;
+
+\path (and2)  edge node[inner sep=0mm,pos=.4] (t0) {} node {} (l2);
+\path (and2)  edge node[inner sep=0mm,pos=.4] (k0) {} node {} (r2);
+\path (t0) edge[bend right] (k0) ;
+
+\path (and3)  edge node[inner sep=0mm,pos=.4] (t1) {} node {} (l3);
+\path (and3)  edge node[inner sep=0mm,pos=.4] (k1) {} node {} (r3);
+\path (t1) edge[bend right] (k1) ;
+\end{tikzpicture}
+\end{figure}
+
+In accordance with the stochastic grammar over programs, we weight all
+edges from $OR$ nodes according to the negative log likelihood of
+choosing the associated rule from the stochastic grammar, conditioned
+on the requesting type. Expanding an $AND$ node has zero cost. In the
+case that the stochastic grammar is ill condition, we might have an
+infinitely deep solution tree whose cost does not increase as we
+unfold it; this can occur, for example, if there are no primitives of
+the required type since the log probability of expansion to the next
+$AND$ node is 0. . To avoid this, we simply declare a cost for
+expanding the next $AND$ node when it is the only node available.
+
+\subsection{Finding the most compressive set of solution.}
+
+Once we have enumerated the best $N$ programs from our distribution,
+we will find that some tasks have no solutions and some have one or
+more solutions; for those tasks that have solutions, we want to choose
+a solution for each task such that the complete set of such solutions
+for all the tasks is as compressible as possible.
+
+A natural metric for the compressibility of a set of binary trees is
+the number of unique subtrees that this set contains. Thus, we can
+attempt to find the assignment of solutions to tasks that minimizes
+this value. However, an exact optimization for this problem requires
+examining $O(m^k)$ assignments, where $m$ is the maximum number of
+solutions for any one task and $k$ is the number of tasks. Since this
+is prohibitive, we relax the compressibility metric as follows: we
+define the cost of adding each new solution $s_n$ to a partial set of
+solutions $s_1, \dots, s_{n-1}$ to be the number of additional unique
+subtrees in $s_n$ that are not present in $s_{n-1}$. Our solution thus
+becomes approximate and that approximation is order-dependent, but a
+depth-first search on the defined search space goes from exponential
+in the solution degree to quadratic.
+
+\subsection{Re-estimating the stochastic grammar over programs.}
+
+Given a set of solution programs, we want to re-estimate our
+stochastic grammar. Our algorithm for doing this is based broadly on
+the Neville-Manning algorithm for grammar-based
+compression~\cite{nevill1997identifying}. In our algorithm, we simply
+count the numer of times each unique subtree appears as the child of a
+subtree \emph{not yet encountered} in the solution set. That is, we
+traverse the solution set in depth first order, comparing each tree we
+encounter to all the trees we previously encountered; but if we
+encounter a tree we have already seen, we do not descend into it. In
+this traversal, the set of trees we encounter more than once is the
+set of terminal nodes in our new grammar. To estimate the
+probabilities associated with each node production, we again traverse
+the solution set, but these time for each node where a terminal
+production takes place we count which other terminal productions could
+have taken place. We estimate the terminal production weight to be the
+ratio of the number of times that node was used in the first traversal
+divided by the number of times it was used in the second traversal
+(plus 1, to account for the event of an $AND$ node
+expansion). Likewise, we re-estimate the probability of an $AND$ node
+expansion by diving the number of such expansions in the solution set
+by the number of possible expansions in the solution set.
 
 \section{Experiments.}
-\subsection{Integers: a toy case study.}
 \subsection{Symbolic regression.}
-\subsection{Learning electric circuits.}
-\begin{figure}[h]
-\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=2cm,
-                    , minimum size=.5cm, thick,main node/.style={circle,draw}]
 
-  \node[main node] (S) {$\{\}$};
-  \node[main node] (1) [below left of=S] {$\{a\}$};
-  \node[main node] (2) [below right of=S] {$\{b\}$} ;
-  \node[main node] (3) [below left of=1] {$\{a, a\}$};
-  \node[main node] (4) [below right of=1] {$\{a, b\}$};
-  \node[main node] (5) [below right of=2] {$\{b, b\}$};;
+In this section, we walk through the performance of our learning
+algorithm on a symbolic regression problem. In this problem,
+each task is a set of pairs $ (x, f(x))$ where $x \in (0, \dots, 10)$
+and $f$ is a polynomial. The goal is to discover a program for each
+task that outputs the appropriate value $f(x)$ for each input x. We
+assume that the data is noise-free.
 
-  \path
-    (S) edge node [left] {a} (1)
-        edge node [right] {b} (2)
-    (1) edge node [left] {a} (3)
-        edge node [right] {b} (4)
-    (2) edge node [left] {a} (4)
-        edge node [right] {b} (5)
+The choice of an initial grammar for the algorithm will, of course,
+have a big effect on its performance, and so we would like to
+characterise this effect. We can divide primitives into roughly two
+sets; there are low-level primitives, those that operate on concrete
+types like Bool or Double, and these tend to be domain specific. For
+example, conrete primitives include $1, 2$, etc., and $f(x) = 3x$. The
+other class of primitives are abstract primitives, which can be
+characterized by having arguments whose types are primarily function
+types, like $(t \rightarrow s), (t \rightarrow s) \rightarrow \text{
+  Double },$ etc. Of course, we can construct operators that fall in
+between.
 
-\end{tikzpicture}
+The most abstract operators are those that Liang et. al refer to as
+``routers''~\cite{liang10programs}. These operators take the place of
+$\lambda$ abstraction in the general lambda calculus: whereas
+$\lambda$ abstraction allows us to bind arguments to variables nested
+deep inside an expression, without them we need to have operators that
+``route'' arguments to the appropriate functions in a nested
+expression.
+
+\begin{definition}[Router~\cite{liang10programs}] A \emph{router} is a combinator
+represented by a finite sequence of elements from $\{B, C, S\}$. The
+behavior of router $r$ is given by
+\[
+(r \, f \, g \, x_1 \dots x_{|r|}) = (f \, x_{i_1} \dots x_{i_n}) (g
+\, x_{j_1} \dots x_{j_n}),
+\]
+where $i_1 < \dots < i_n$ are indices $i$ such that $r_i \in \{C, S\}$
+and where $j_1 < \dots < j_n$ are indices $j$ such that $r_j \in \{B,
+S\}$. Let $\mathcal{R}_k$ be the set of all routers of length equal to
+$k$. Let $\mathcal{R}_{\leq k}$ be the set of all routers of length
+less than or equal to $k$.
+\end{definition}
+
+To analyze the behavior of our algorithm, we will vary the parameters
+of the problem along two axes while keeping the learning tasks the
+same. Along the first axis, we vary the initial library to which our
+learner has access with respect to the its abstractness (i.e., number
+of routers) and its concreteness (i.e. number of concrete
+primitives). Along the second axis we vary the size of the search
+frontier.
+
+Each task is to predict the value of an unknown polynomial on the
+inputs $0, \dots, 10$. The set of tasks corresponds to the set of
+polynomials $\{ax^2 + bx + c | a, b, c \in \{0, \dots, 10 \}\}$.
+
+In Figure~\ref{fig:symreg}, we show performance results as we vary
+both the routers included and the frontier size. Each individual plot
+shows how changing the frontier size changes the number of tasks the
+algorithm identifies correctly over the course of ten algorithm
+iterations. In addition, we have a baseline for each curve, visits the
+same number of expressions as the corresponding algorithm run
+(i.e. frontier size $\times$ $\#$ of iterations) but all at once; it
+does not do any learning. Thus, this baseline serves as a brute-force
+control for the learning aspect of the algorithm.
+
+\begin{figure}
+\label{fig:symreg}
+\includegraphics[width=\linewidth]{figures/learningCurves.pdf}
+\includegraphics[width=\linewidth]{figures/performanceFrontierSize.pdf}
+\caption{E-C learning curves for the symbolic regression task set as
+  we vary the frontier size.}
 \end{figure}
+
+Why do the learning curves asymptote early? One might object that a
+learning algorithm should continuously improve, if only slowly, as it
+is given more time. But in these runs we observe a finite and
+permanent performance cap. This is expected: once the learner find a
+set of concepts it can explain, if those concepts fill up the
+frontier, then there is nothing to spur the learner to incorporate as
+yet unsolved tasks.
+
+To investigate whether this is in fact the case, we can examine the
+proportion of expressions in the frontier that match tasks in our task
+set; that is, is the learning curve plateauing because the frontier is
+heavily redundant, or is it plateuing because the frontier contains
+many expressions that do not hit the task set at
+all. Figure~\ref{fig:plateau_investigation} shows the proportion of
+the expressions in the frontier that hit a task as the frontier sizeincreases.
+
+\subsection{Digital arithmetic}
+Having demonstrated the behavior of the E-C algorithm on symbolic
+regression, we analyze its performance on tasks involving digital
+arithmetic, a class of problems with known modular structure. The
+general goal is to discover boolean circuits consisting of a limited
+set of gates that perform arithmetic operations on binary number
+representations.
 

### 2013-01-18 src/PI/CL.hs

@@ -40,7 +40,7 @@ isCNode (CNode{}) = True
 isCNode _ = False
 
 type SynthComb = Either String Comb
-
+ 
 app :: SynthComb -> SynthComb -> SynthComb
 app m1 m2 = do c1 <- m1
                c2 <- m2
@@ -65,7 +65,7 @@ infixl 4 <::>
 instance Show Comb where
     show (CApp c1 c2 _ _) = "(" ++ show c1 ++ " " ++ show c2 ++ ")"
     show (CNode n _ _) = n
-    show (CTerminal t) =  "CTerm of type " ++ show t
+    show (CTerminal t) =  "CTerm: " ++ show t
 
 -- | An alternative to show: if combinator is named and evaluates to a
 -- number or bool, show it an an evaluated expressions.
@@ -100,6 +100,7 @@ bool2C c = CNode (show c) (B c) tBool
 getType :: Comb -> Either String Type
 getType (CNode _ _ t) = Right t
 getType c@(CApp c1 c2 _ _) = getAppType c1 c2
+getType c@(CTerminal t) = Right t
 
 getAppType :: Comb -> Comb -> Either String Type
 getAppType c1 c2  

### 2013-01-18 src/PI/Compress.hs

@@ -61,8 +61,8 @@ incr2 index c@(CApp l r _ _) tp
                              l_index <- incr2 index' l t_left
                              r_index <- incr2 l_index r (fromType (cType l))
                              return r_index
-           otherwise -> do let index' = CM.insert c [tp] index 
-                           return index'
+           Just xs -> do let index' = CM.insert c (tp:xs) index 
+                         return index'
 incr2 index c@(CNode _ _ _) tp = return $ CM.insertWith (++) c [tp] index 
 
 compress2 :: [(Type, Comb)] -> CombMap [Type]

### 2013-01-18 src/PI/EnumBF.hs

@@ -47,6 +47,9 @@ isClosed :: CombBase -> Bool
 isClosed cb | path cb == Nothing  = True
             | otherwise           = False
 
+-- hack min grammar val
+hackMinGrammarVal = log (0.5)
+
 
 enumBF :: Grammar 
        -> Int -- max num combinators
@@ -71,9 +74,10 @@ enumBF' gr i bfState@(BFState openPQ closedPQ) =
             closedCBs = filter (isClosed . combBase) cbs
             openCBs= filter (not . isClosed . combBase) cbs
             closedPQ' = (PQ.fromList closedCBs) `PQ.union` closedPQ
-                        
             openPQ'' = PQ.fromList openCBs `PQ.union` openPQ'
-        in enumBF' gr i $ BFState openPQ'' closedPQ'
+        in (trace $ "OPEN: " ++ (show $ PQ.size openPQ'')
+            ++ " CLOSED: " ++ (show $ PQ.size closedPQ'))
+           $ enumBF' gr i $ BFState openPQ'' closedPQ'
         
 
 expandToApp :: Grammar -> CombBase -> StateT Int [] CombBase
@@ -84,7 +88,9 @@ expandToApp gr (CombBase (CTerminal tp) (Just []) v) =
            c_left = (CTerminal t_left) 
            c_right = (CTerminal t_right)
            c = CApp c_left c_right tp 1
-           cb = CombBase c (Just [L]) (exLogProb gr tp )
+           minGrammarVal = hackMinGrammarVal
+--           minGrammarVal = minimum (CM.elems (library gr))
+           cb = CombBase c (Just [L]) (2*minGrammarVal)
        return cb
 
 expand :: Grammar 
@@ -93,17 +99,21 @@ expand :: Grammar
 expand gr cb@(CombBase (CTerminal tp) (Just []) v) = cbs `mplus` cbApp
     where primCombs = CM.keys (library gr)
           cs = filterCombinatorsByType primCombs tp
-          cbs = do c <- cs
-                   let value = (calcLogProb gr tp c)
-                       combBase = CombBase c Nothing (v + value)
-                   return combBase
+--          minGrammarVal = minimum (CM.elems (library gr))
+          minGrammarVal = hackMinGrammarVal
+          cbs =   do c <- cs
+                     let value = (calcLogProb gr tp c)
+                         combBase = CombBase c Nothing value
+                     return combBase
           cbApp = expandToApp gr cb 
[…]

### 2013-01-18 src/PI/Experiment.hs

@@ -24,6 +24,7 @@ data Experiment = Experiment {expName :: String, -- ^ name of experiment
                               expDepthBound   :: Int, 
                               expNumBound :: Int, -- ^ number of combinators to search
                               expReps :: Int } -- ^ number of iterations
+                deriving Show
 
 ----- Squared Integers -------
 rlimit = 100
@@ -49,44 +50,44 @@ expSquaredInts
 ------------------------------
 
 ---- Integer Sequences ----
-rng = [0..10]
-red = (reduceWithLimit 100) . comb2Expr
-eval :: [Comb] -> Comb -> Double
-eval ds c = fromIntegral $ sum $ [(toInteger $ diff i) | i <- rng]
-         where diff i = abs $ dat i - eval i
-               eval i = if y < 0 then maxBound else y 
-                   where y = case red  (CApp c (num2C i) tInt 0) of
-                               Just (N y) ->  y
-                               Nothing -> maxBound
-               dat i = y where Just (N y) = red (ds !! i)
+-- rng = [0..10]
+-- red = (reduceWithLimit 100) . comb2Expr
+-- eval :: [Comb] -> Comb -> Double
+-- eval ds c = fromIntegral $ sum $ [(toInteger $ diff i) | i <- rng]
+--          where diff i = abs $ dat i - eval i
+--                eval i = if y < 0 then maxBound else y 
+--                    where y = case red  (CApp c (num2C i) tInt 0) of
+--                                Just (N y) ->  y
+--                                Nothing -> maxBound
+--                dat i = y where Just (N y) = red (ds !! i)
 
-quad a b c = \i -> a * i^2 + b * i + c
-showQuad a b c = show a ++ " x^2 + " ++ show b ++ " x + " ++ show c
-mkQuadTask a b c = Task (showQuad a b c) f (tInt ->- tInt)
-    where f = eval [num2C $ (quad a b c i)| i<-rng]
+-- quad a b c = \i -> a * i^2 + b * i + c
+-- showQuad a b c = show a ++ " x^2 + " ++ show b ++ " x + " ++ show c
+-- mkQuadTask a b c = Task (showQuad a b c) f (tInt ->- tInt)
+--     where f = eval [num2C $ (quad a b c i)| i<-rng]
 
-squares a b = \i -> (a * i + b)^2
[…]

### 2013-01-18 src/PI/Experiments/ListExperiment.hs

@@ -1,6 +1,6 @@
 -- ListExperiment.hs
 
-module ListExperiment where 
+module Experiments.ListExperiment where 
 
 import Control.Monad
 import Debug.Trace
@@ -10,6 +10,7 @@ import StdLib
 import Data
 import CL
 import CombMap (CombMap)
+import qualified CombMap as CM
 import Task
 import Type
 import Experiment
@@ -28,14 +29,45 @@ maybe2Either Nothing = Left ""
 
 fromRight (Right x) = x
 
-tLast = Task "first" f tp
-    where tp = (TAp tList tInt) ->- tInt
+tLast = Task "last" f tp
+    where tp = (TAp tList tInt) ->- (TAp tList tInt)
           f comb = reward
               where ex = liftM comb2Expr $ Right comb <:> (mkIntList [1,2,3,4]) 
                     result = reduce (fromRight ex)
                     reward = case result of 
                                (N 2) -> 0
                                otherwise -> 1
+listlib = CM.fromList $ 
+          [ ("I", cI)
+          , ("S", cS)
+          , ("B", cB)
+          , ("C", cC)
+        , ("cond", cCond)
+        , ("True", cTrue)
+        , ("False", cFalse)
+        , ("|", cOr)
+        , ("&", cAnd)
+        , ("not", cNot)
+--         , ("<", cLT)
+--         , (">", cGT)
+--         , ("==", cEQ)
+          , ("+", dOp2C "+" (+))
+--          , ("-", dOp2C "-" (-))
+           , ("*", dOp2C "*" (*))
+          ,  ("0", num2C 0)
+          ,  ("1", num2C 1)
+--          , ("2", num2C 2)
+--         , ("3", num2C 3)
+--         , ("4", num2C 4)
+--           , ("primRec", cPrim) 
+          , (":", cCons)
+          , ("[]", cEmpty)
+           , ("head", cHead)
+           , ("tail", cTail)
+           , ("isEmpty", cIsEmpty)
+           ]
+
+
 
 expTList1 = Experiment {
               expName = "List function: first",
@@ -44,7 +76,7 @@ expTList1 = Experiment {
               expPrior = stdgrammar,
               expInitLib = stdgrammar,
               expDepthBound = 2,
-              expNumBound = 100,
+              expNumBound = 10,
               expReps = 1}
                       
                    

### 2013-01-18 src/PI/Expr.hs

@@ -61,7 +61,6 @@ reduce e@(App (Func f) a)  = let z = (reduce a)
                              in case z of 
                                   err@(ExprError _) -> err
                                   otherwise -> z `seq` reduce $ f z
-
 reduce e@(App (Const _) _) =  e
 reduce x@(App a b) | isRedex a =  let a' = reduce a 
                                   in case a' of

### 2013-01-18 src/PI/Grammar.hs

@@ -54,10 +54,10 @@ logsumexp = log . sum . (map exp)
 
 normalizeGrammar :: Grammar -> Grammar 
 normalizeGrammar (Grammar lib ex)
-    = let logTotalMass = logsumexp $ ex:(CM.elems lib)
+    = let logTotalMass = logsumexp $ (CM.elems lib)
           lib' = CM.map (+ (-logTotalMass)) lib
           ex' = ex - logTotalMass
-      in Grammar lib'  ex'
+      in Grammar lib'  0 -- ex'
           
 sum' (a, b) (c, d) = (a + b, c + d)
 
@@ -126,26 +126,27 @@ calcLogProb gr tp c
     = let m = filterCombinatorsByType (CM.keys $ library gr) tp
           altCs = map fst $ runStateT m 0
           combLps = [exp $ (library gr) CM.! x | x <- altCs] 
-          logProbAll = log $ exp (expansions gr) + sum combLps
+          logProbAll = log $ sum combLps
           combLogProb = (library gr) CM.! c - logProbAll
-      in combLogProb
+          out = if length altCs < 2 then log (0.5) else combLogProb
+      in out
 
-exLogProb :: Grammar -> Type -> Double
-exLogProb gr tp  
-    = let m = filterCombinatorsByType (CM.keys $ library gr) tp
-          altCs = map fst $ runStateT m 0
-          combLps = [exp $ (library gr) CM.! x | x <- altCs] 
-          logProbAll = log $ exp (expansions gr) + sum combLps
-          out = if null altCs then log (0.5)  -- ^ this log (0.5) is a
-                                              -- hack. it should be 0,
-                                              -- right? since there
-                                              -- are no alternatives,
-                                              -- the probability is
-                                              -- 1. But this causes
-                                              -- infinite expansions
-                                              -- in the best-first search. 
-                else expansions gr  - logProbAll
-      in  out
+-- exLogProb :: Grammar -> Type -> Double
+-- exLogProb gr tp  
+--     = let m = filterCombinatorsByType (CM.keys $ library gr) tp
+--           altCs = map fst $ runStateT m 0
+--           combLps = [exp $ (library gr) CM.! x | x <- altCs] 
+--           logProbAll = log $ exp (expansions gr) + sum combLps
+--           out = if null altCs then log (0.5)  -- ^ this log (0.5) is a
+--                                               -- hack. it should be 0,
+--                                               -- right? since there
+--                                               -- are no alternatives,
+--                                               -- the probability is
[…]

### 2013-01-18 src/PI/Search.hs

@@ -35,20 +35,15 @@ import qualified Compress as CP (getUniqueTrees, compress2)
 import Compress (Index)
 import Experiment
 
-data SearchLogEntry = SearchLogEntry { searchIter :: Int,
-                                       searchLib :: Grammar,
-                                       searchSpace :: [Comb],
-                                       searchNumHit :: Int,
-                                       searchExplanation :: [(Task, Comb)],
-                                       searchCompression :: CombMap Int
+data SearchLogEntry = SearchLogEntry { searchGrammar :: Grammar,
+                                       searchExplanation :: [(Task, Comb)]
                                      } deriving Show
-mkEmptySearchLog :: SearchLogEntry
-mkEmptySearchLog  = SearchLogEntry 0 nullGrammar [] 0 [] CM.empty
 
-type Search a = State SearchLogEntry a 
 
-runSearch :: Search a -> (a, SearchLogEntry)
-runSearch s = runState s mkEmptySearchLog 
+type Search a = Writer [SearchLogEntry] a 
+
+runSearch :: Search a -> (a, [SearchLogEntry])
+runSearch s = runWriter s 
 
 mkHypothesisSpace :: Experiment 
                   -> Grammar 
@@ -62,16 +57,30 @@ mkHypothesisSpace expr gr
           taskSet = expTaskSet expr
 
 
+
+evalSymReg :: Comb -> [Int]
+evalSymReg c = [ f i | i <- [0..10]]
+    where f i = case reduceComb $ CApp c (num2C i) tInt 0 of 
+                  (N y ) -> y
+                  otherwise -> (maxBound :: Int)
+
+
 -- | For each data point in a dataset list all the expressions that
 -- evaluate to that datapoint. 
 findCombinatorsForEachDatum :: Experiment 
                             -> Grammar -- ^ current grammar
                             -> Search [(Task, [Comb])]
 findCombinatorsForEachDatum expr gr
-    = do s <- get
+    = do 
          db <- mkHypothesisSpace expr gr
-         return $ [(t, [ c | c <- db Map.! tp, 
-                       f c <= eps ]) | t@(Task n f tp) <- taskSet]
+         return $ do 
+           t <- taskSet
+           let cs = case t of 
+                      Task n f tp -> do c <- db Map.! tp
+                                        if f c <= eps then return c else mzero
+                      SymRegTask n vs tp -> do c <- db Map.! tp
[…]

### 2013-01-18 src/PI/StdLib.hs

@@ -12,6 +12,9 @@ import Type
 import CL
 import Expr
 import Grammar
+import Routers (cS, cB, cC, cSS, cSB, cSC, 
+                  cBS, cBB, cBC, cCS, cCB, cCC,
+                  cSSS, cBCB, routers)
 
 type NamedLib = HMap.Map String Comb
 
@@ -29,14 +32,66 @@ t8 = mkTVar 8
  
 -- | define common combinators
 cI = CNode "I" (Func id) ( t0 ->- t0)
-cS = CNode "S" (Func $ \f -> Func $ \g -> Func $ \x -> (App (App f x) (App g x))) typeS 
-    where typeS = (t2 ->- t1 ->- t0) ->- (t2 ->- t1) ->- (t2 ->- t0)
+cK = CNode "K" (Func $ \a -> Func $ \b -> a)   ( t0 ->- t1 ->- t0)
+-- cS = CNode "S" (Func $ \f -> Func $ \g -> Func $ \x -> (App (App f x) (App g x))) typeS 
+--     where typeS = (t2 ->- t1 ->- t0) ->- (t2 ->- t1) ->- (t2 ->- t0)
 
-cB = CNode "B" (Func $ \f -> Func $ \g -> Func $ \x -> (App f (App g x))) typeB 
-    where typeB = (t1 ->- t0) ->- (t2 ->- t1) ->- (t2 ->- t0)
+-- cB = CNode "B" (Func $ \f -> Func $ \g -> Func $ \x -> (App f (App g x))) typeB 
+--     where typeB = (t1 ->- t0) ->- (t2 ->- t1) ->- (t2 ->- t0)
 
-cC = CNode "C" (Func $ \f -> Func $ \g -> Func $ \x -> (App (App f x) g )) typeC 
-    where typeC = (t1 ->- t2 ->- t0) ->- (t2 ->- t1 ->- t0)
+-- cC = CNode "C" (Func $ \f -> Func $ \g -> Func $ \x -> (App (App f x) g )) typeC 
+--     where typeC = (t1 ->- t2 ->- t0) ->- (t2 ->- t1 ->- t0)
+
+-- cBCS = CNode "BCS" (Func $ \f 
+--                 -> Func $ \g 
+--                 -> Func $ \a 
+--                 -> Func $ \b 
+--                 -> Func $ \x 
+--                 -> ((App (App f b) (App (App g a) x)))) tp 
[…]

### 2013-01-18 src/PI/Task.hs

@@ -10,11 +10,15 @@ import Type
 type Reward = Double
 data Task =  Task {taskName :: String, 
                    task :: (Comb -> Reward),
-                  taskType :: Type} 
+                   taskType :: Type} 
+          | SymRegTask {taskName :: String,
+                        taskVals :: [Int],
+                        taskType :: Type}
 type TaskSet = [Task]
 
 instance Show Task where
     show (Task n _ _ ) = n
+    show (SymRegTask n _ _ ) = n
 
 
 -- | task constructors
@@ -22,9 +26,9 @@ instance Show Task where
 mkSingleEqualityTask:: Int -> Int -> Task
 mkSingleEqualityTask rlimit i 
     = let f c = fromIntegral $ (abs $ a - i)
-              where a = case reduceWithLimit rlimit $ comb2Expr c of
-                          (Just (N x)) -> x
-                          Nothing -> 2147483647                    
+              where a  = case  reduceComb c of
+                           (N a) -> a
+                           otherwise -> maxBound
       in Task (show i) f tInt
 
 showTaskCombAssignments :: [(Task, Comb)] -> String

### 2013-01-18 src/PI/Test.hs

@@ -3,8 +3,10 @@
 
 module Main where
 
+import System.IO
 import qualified Data.Map as Map
 import Control.Monad.State
+import Debug.Trace
 
 import Type
 import CL
@@ -16,10 +18,15 @@ import Search
 import qualified CombMap as CM
 import Experiment
 import qualified Compress as CP
-import ListExperiment
+-- import Experiments.ListExperiment
+import Experiments.SymbolicRegression
+import PostProcess 
+import Grammar
 
-grammar = fst $ runSearch $ loop expIntegerSequences
+out =  runSearch $ loop expSymReg
+grammar = (trace $ "OUT: " ++  (show $ snd out)) $ fst out
+searchData = snd out
 
 combs = map comb $ enumBF stdgrammar 1000 (tInt ->- tInt)
 main = do
-   putStrLn $ show grammar
+  saveSearchData "data" expSymReg searchData

### 2013-01-18 src/PI/Type.hs

@@ -205,5 +205,6 @@ tBool = TCon (TyCon "Bool" Star)
 tArrow = TCon (TyCon "(->)" (Kfun Star (Kfun Star Star)))
 tList = TCon $ TyCon "[]" (Kfun Star Star)
 tMaybe = TCon $ TyCon "Maybe" (Kfun Star Star)                                         
+tPair = TCon $ TyCon "(,)" (Kfun Star Star)
                      
 

### 2013-01-18 src/PI/Experiments/DigitalArithmetic.hs
Experiments added.

@@ -1,0 +1,47 @@
+-- DigitalArithmetic
+
+module Experiments.DigitalArithmetic where
+
+import Data.Maybe
+import Debug.Trace
+
+import Experiment
+import Type
+
+import Type
+import CL
+import Expr
+import Task
+import StdLib
+import Data
+import qualified CombMap as CM
+import CombMap (CombMap)
+import Grammar
+import Routers 
+import PostProcess
+import EnumBF
+import ParseCL hiding (eval)
+import Run
+
+numto1dig n = n `mod` 2 
+numto2dig n = let a = n `div` 1 `mod` 2
+                  b = (n-a*1) `div` 2 `mod` 2
+              in (b, a)
+numto3dig n = let a = n `div` 1 `mod` 2
+                  b = (n-a * 1) `div` 2 `mod` 2
+                  c = (n-a * 1 - b*2) `div` 4 `mod` 2
+              in (c, b, a)
+numto4dig n = let a = n `div` 1 `mod` 2
+                  b = (n-a * 1) `div` 2 `mod` 2
+                  c = (n-a * 1 - b*2) `div` 4 `mod` 2
+                  d = (n-a * 1 - b*2 - c*4) `div` 16 `mod` 2
+              in (d, c, b, a)
+numto5dig n = let a = n `div` 1 `mod` 2
+                  b = (n-a * 1) `div` 2 `mod` 2
+                  c = (n-a * 1 - b*2) `div` 4 `mod` 2
+                  d = (n-a * 1 - b*2 - c*4) `div` 16 `mod` 2
+                  e = (n-a * 1 - b*2 - c*4 - d*16) `div` 32 `mod` 2
+              in (e, d, c, b, a)
+
+
+

### 2013-01-18 src/PI/Experiments/SymbolicRegression.hs

@@ -1,0 +1,153 @@
+-- SymbolicRegression.hs
+{-# Language ParallelListComp #-}
+
+module Experiments.SymbolicRegression where
+
+import Data.Maybe
+import Debug.Trace
+
+import Experiment
+import Type
+
+import Type
+import CL
+import Expr
+import Task
+import StdLib
+import Data
+import qualified CombMap as CM
+import CombMap (CombMap)
+import Grammar
+import Routers 
+import PostProcess
+import EnumBF
+import ParseCL hiding (eval)
+import Run
+
+rng = [0..10]
+eval :: [Comb] -> Comb -> Double
+eval ds c = fromIntegral $ sum $ [(toInteger $ diff i) | i <- rng]
+         where diff i = abs $ dat i - eval i
+               eval i = if y < 0 then maxBound else y 
+                   where y = case reduceComb $ (CApp c (num2C i) tInt 0) of
+                               (N y) -> y
+                               otherwise -> maxBound
+               dat i = y where (N y) = reduceComb (ds !! i)
+
+quad a b c = \i -> a * i^2 + b * i + c
+showQuad a b c = show a ++ " x^2 + " ++ show b ++ " x + " ++ show c
+mkQuadTask a b c = Task (showQuad a b c) f (tInt ->- tInt)
+    where f = eval [num2C $ (quad a b c i)| i<-rng]
+
+mkQuadSymRegTask a b c = SymRegTask (showQuad a b c) f (tInt ->- tInt)
+    where f  = [quad a b c i | i <- [0..10]]
+
+squares a b = \i -> (a * i + b)^2
+showSquares a b = "(" ++ show a ++ " x + " ++ show b ++ ")^2"
+mkSquaresTask a b = Task (showSquares a b) f (tInt ->- tInt)
+                   
+    where f = eval [num2C $ (squares a b) i| i<-rng]
+
+taskSet' = --[mkSquaresTask i j | i <- [0..5] , j <- [0..5]]
+           -- ++  (map (mkSingleEqualityTask 100) $ [i*i | i <- [0..20]])
+           [mkQuadTask i j k | i <- [0..10] , j <- [0..10], k <- [0..10]]
+--           ++ [mkQuadTask i j k | i <- [0..0] , j <- [0..5], k <- [0..5]]
+
+symRegTaskSet' = --[mkSquaresTask i j | i <- [0..5] , j <- [0..5]]
[…]

### 2013-01-20 drafts/ijcai2013/ijcai13.tex
Changes.

@@
 As a simple example of our representation, consider the squaring
-function $\lamba x. * x x $. Using two of the basis combinators above,
+function $\lambda x. * x\; x $. Using two of the basis combinators above,
 we can write the squaring function in combinatory logic as $S * I$,
@@
 a binary tree~(Figure~\ref{fig:clbintree}) whose leaves are typed
-primitive combinators and whose interior nodes are typed applications of combinators.
+primitive combinators and whose interior nodes are typed applications
+of combinators.
 
@@
   \centering
-  \begin{subfigure}{0.2\linewidth}
-    \label{fig:clbintree}
-    \begin{tikzpicture}[scale=.5, transform shape, level/.style={sibling distance=10mm/#1, font=\scriptsize}]
+  \begin{subfigure}[Before]{0.5\linewidth}
+    \begin{tikzpicture}[scale=.7,
+        transform shape, level/.style={sibling distance=10mm/#1, font=\scriptsize}
+        ]
       \node [font=\scriptsize] (top) {$I$}
@@
       };
-
     \end{tikzpicture}
+    \caption{The typed combinator $S * I$ ($f(x)=x^2$) represented as binary tree.}
+    \label{fig:clbintree}
   \end{subfigure}
-
-  \begin{subfigure}[b]{0.2\linewidth}
-    \label{fig:andor}
-    \begin{tikzpicture}[scale=.5, transform shape, level/.style={sibling distance=10mm/#1}]
+  \begin{subfigure}[After]{0.4\linewidth}
+    \begin{tikzpicture}[scale=.7, transform shape, 
+        level/.style={sibling distance=10mm/#1},
+        every edge/.style={draw, thin}
+      ]
       \node [circle,draw] (a) {}
@@
       child {node [left=.4cm] (e) {$f_n$}}
-      child {node [circle,draw, left=1cm] (d) {}
+      child {node [circle,draw, left=1cm] (d) {} edge from parent[very thick]
         child {node [circle, draw, left=.25cm] (l) {} edge from parent[draw=none]
@@
           child {node [left=.3cm] (c) {\dots} edge from parent[draw=none] }
-          child {node [left=.3cm] (e) {$f_n$}}
+          child {node [left=.3cm] (e) {$f_n$} edge from parent[very thick]}
           child {node [circle, draw, left=.25cm] (and2) {}
@@
           }
-          child {node [right=.3cm] (b) {$f_1$}} 
+          child {node [right=.3cm] (b) {$f_1$} edge from parent[very thick]} 
           child {node [right=.4cm] (c) {\dots} edge from parent[draw=none] }
@@
 
-      \path (d)  edge node[inner sep=0mm,pos=.4] (a1) {} node {} (l);
-      \path (d)  edge node[inner sep=0mm,pos=.4] (b1) {} node {} (r);
+      \path (d)  edge[very thick]  node[inner sep=0mm,pos=.4] (a1) {} node {} (l);
+      \path (d)  edge[very thick]  node[inner sep=0mm,pos=.4] (b1) {} node {} (r);
       \path (a1) edge[bend right] (b1) ;
@@
     \end{tikzpicture}
-
+    \caption{The space of all typed combinators represented as an
+      infinitely deep And-Or tree. Any specific combinator as a
+      sub-binary tree of this And-Or tree whose leaves consist of only
+      the $f_i$ nodes (i.e., the primitive combinators of the
+      grammar).}
+    \label{fig:andor}
     \end{subfigure}
@@
 \subsection{Best-first enumeration of most probably programs.}
-In order to explore the frontier of most probable programs, we need a
+In order to explore the frontier of most promising programs, we need a
 procedure that enumerates the $N$ most probable programs. We can
-formulate this naturally as a best-first exploration of an $AND-OR$
-tree (NEED FIGURE). Complete programs are represented by binary trees
-that are valid subtrees of the $AND-OR$ graph. The tree structure is
-shown in Figure~\ref{fig:andor}. The root is an $OR$ node of the
-requisite program type. It's children are the primitives of that type
-and one $AND$ node with two children. Each child has the same
-structure as the root node, except that the types are different. If
-the type of an $OR$ node is $t$, then the type of the left child of
-its $AND$ node child is $\tau \rightarrow t$ (where $\tau$ is a fresh
-type variable). Since the type of the right child is constrained by
-the subtree rooted at the left child, we always expand the left child
-first, and we use the target type of this left subtree to mark the
-type of the right child.
-
-\begin{figure}
-\label{fig:andor}
-\begin{tikzpicture}[scale=.5, transform shape, level/.style={sibling distance=10mm/#1}]
-\node [circle,draw] (a) {}
-child {node [left=0cm] (b) {$f_1$}}
-child {node [left=.2cm] (c) {\dots} edge from parent[draw=none] }
-child {node [left=.4cm] (e) {$f_n$}}
-child {node [circle,draw, left=1cm] (d) {}
-  child {node [circle, draw, left=.25cm] (l) {} edge from parent[draw=none]
-    child {node [left=.5cm] (b) {$f_1$}} 
-    child {node [left=.3cm] (c) {\dots} edge from parent[draw=none] }
-    child {node [left=.3cm] (e) {$f_n$}}
-    child {node [circle, draw, left=.25cm] (and2) {}
-          child {node [circle, draw, left=.25cm](l2) {} edge from parent[draw=none] 
-            child {node [above=.5cm] {\vdots} edge from parent[draw=none]}
-          }
-          child {node [circle, draw] (r2)  {} edge from parent[draw=none] 
-            child {node [above=.5cm] {\vdots} edge from parent[draw=none]}
-          }
-    }
-  }
-  child {node [circle, draw, right=.25cm] (r) {} edge from parent[draw=none]
-    child {node [circle, draw, right=.25cm] (and3) {}
-          child {node [circle, draw ](l3) {} edge from parent[draw=none] 
-            child {node [above=.5cm] {\vdots} edge from parent[draw=none]}
-          }
-          child {node [circle, draw, right=.2cm ] (r3)  {} edge from parent[draw=none] 
-            child {node [above=.5cm] {\vdots} edge from parent[draw=none]}
-          }
-    }
-    child {node [right=.3cm] (b) {$f_1$}} 
-    child {node [right=.4cm] (c) {\dots} edge from parent[draw=none] }
-    child {node [right=.6cm] (e) {$f_n$}}
- }
-};
-
-\path (d)  edge node[inner sep=0mm,pos=.4] (a1) {} node {} (l);
-\path (d)  edge node[inner sep=0mm,pos=.4] (b1) {} node {} (r);
-\path (a1) edge[bend right] (b1) ;
-
-\path (and2)  edge node[inner sep=0mm,pos=.4] (t0) {} node {} (l2);
-\path (and2)  edge node[inner sep=0mm,pos=.4] (k0) {} node {} (r2);
-\path (t0) edge[bend right] (k0) ;
-
-\path (and3)  edge node[inner sep=0mm,pos=.4] (t1) {} node {} (l3);
-\path (and3)  edge node[inner sep=0mm,pos=.4] (k1) {} node {} (r3);
-\path (t1) edge[bend right] (k1) ;
-\end{tikzpicture}
-\end{figure}
+formulate this naturally as a best-first exploration of an $AND/OR$
+tree~\cite{nilsson1982principles}, which we denote $\mathcal{G}$. A
+complete program is any binary subtrees of $\mathcal{G}$ whose leaves
+coincide with the leaves of $\mathcal{G}$. The tree structure is shown
+in Figure~\ref{fig:andor}. The root is an $OR$ node of the requisite
+program type. It's children are the primitives of that type $f_1,
+\dots, f_n$, and one $AND$ node with two children. Each child of an
+$AND$ node has the same structure as the root node, with modified
+types: if the type of an $AND$ node is $t$, then the type of its left
+child is $\tau \rightarrow t$ (where $\tau$ is a fresh type
+variable). Since the type of the right child is constrained by the
+subtree rooted at the left child, we always expand the left child
+first and use the target type of this left subtree to mark the type of
+the right child.
 
@@
 Given a set of solution programs, we want to re-estimate our
-stochastic grammar. Our algorithm for doing this is based broadly on
+stochastic grammar. Our inspiration for doing this is based broadly on
 the Neville-Manning algorithm for grammar-based
-compression~\cite{nevill1997identifying}. In our algorithm, we simply
-count the numer of times each unique subtree appears as the child of a
-subtree \emph{not yet encountered} in the solution set. That is, we
-traverse the solution set in depth first order, comparing each tree we
-encounter to all the trees we previously encountered; but if we
-encounter a tree we have already seen, we do not descend into it. In
-this traversal, the set of trees we encounter more than once is the
-set of terminal nodes in our new grammar. To estimate the
-probabilities associated with each node production, we again traverse
-the solution set, but these time for each node where a terminal
-production takes place we count which other terminal productions could
-have taken place. We estimate the terminal production weight to be the
-ratio of the number of times that node was used in the first traversal
-divided by the number of times it was used in the second traversal
-(plus 1, to account for the event of an $AND$ node
-expansion). Likewise, we re-estimate the probability of an $AND$ node
-expansion by diving the number of such expansions in the solution set
-by the number of possible expansions in the solution set.
+compression~\cite{nevill1997identifying}. The idea of that compression
+algorithm is to compress any string by creating a grammar for that
+string such that a) no two symbols appear more than once in the
+grammar and b) every rule is used more than once. From a set of trees
+we generate a grammar according to these criteria (though using a less
+efficient and much simpler algorithm).
+
+In our algorithm, we traverse a set of combinator binary trees in
+depth first order. We count the occurences of every subtree. Every
+time we encounter a subtree for the second time, we designate it as a
+primitive in our new grammar. Furthermore, during our traversal, we do
+not descend into trees that have been added already to the new
+grammar. Thus, we only count an instance of tree if that instance is
+not a subtree of a primitive element of the new grammar.
+
+We want to count the number of
+times every subtree is encountered when it is not a subtree of a tree
+already encountered in our traversal. we simply count the numer of
+times each unique subtree appears as the child of a subtree \emph{not
+  yet encountered} in the solution set. That is, we traverse the
+solution set in depth first order, comparing each tree we encounter to
+all the trees we previously encountered; but if we encounter a tree we
+have already seen, we do not descend into it. In this traversal, the
+set of trees we encounter more than once is the set of terminal nodes
+in our new grammar. To estimate the probabilities associated with each
+node production, we again traverse the solution set, but these time
+for each node where a terminal production takes place we count which
+other terminal productions could have taken place. We estimate the
+terminal production weight to be the ratio of the number of times that
+node was used in the first traversal divided by the number of times it
+was used in the second traversal (plus 1, to account for the event of
+an $AND$ node expansion). Likewise, we re-estimate the probability of
+an $AND$ node expansion by diving the number of such expansions in the
+solution set by the number of possible expansions in the solution set.
 
@@
 \begin{figure}
-\label{fig:symreg}
+
 \includegraphics[width=\linewidth]{figures/learningCurves.pdf}
@@
   we vary the frontier size.}
+\label{fig:symreg} 
 \end{figure}

### 2013-01-20 src/PI/Experiments/SymbolicRegression.hs

@@ -3,6 +3,10 @@
 
 module Experiments.SymbolicRegression where
 
+import Control.Monad.Random
+import Control.Monad.State 
+import qualified  Data.HashMap as HMap
+
 import Data.Maybe
 import Debug.Trace
 
@@ -102,6 +106,12 @@ lib123 = lib `CM.union` routers
 
 grammarFromLib lib = normalizeGrammar $ Grammar l 0
     where l = CM.fromList $ [(c, 0) | c <- CM.elems lib]
+
+jitteredGrammarFromLib :: HMap.Map k Comb -> Rand StdGen Grammar
+jitteredGrammarFromLib lib = do 
+  vs <- getRandomRs (-1.0,2.0)
+  let lib' = CM.fromList [(c, 2 + k) | c <- CM.elems lib | k <- vs]
+  return $ normalizeGrammar $ Grammar lib' 0
                      
 mkSymRegExp name lib frontierSize 
     = Experiment {expName = name,
@@ -111,16 +121,42 @@ mkSymRegExp name lib frontierSize
                   expInitLib = grammarFromLib lib,
                   expDepthBound = 3,
                   expNumBound = frontierSize,
-                  expReps = 25}
+                  expReps = 15}
+
+mkSymRegExpM :: String -> HMap.Map String Comb -> Int -> Rand StdGen Experiment
+mkSymRegExpM name lib frontierSize 
+    = do 
+  lib' <- jitteredGrammarFromLib lib
+  return $ Experiment {expName = name ++ "_wjittermany",
+                  expTaskSet = symRegTaskSet',
+                  expEps = 0,
+                  expPrior = lib',
+                  expInitLib = lib',
+                  expDepthBound = 3,
+                  expNumBound = frontierSize,
+                  expReps = 15}
+
+
 
 libSet = [lib1] -- lib3, lib123]
 libNames = ["R1"] -- , "R3", "R<=3"]
-frontierSizeSet = [9000, 10000]
+frontierSizeSet = join [replicate 10 i | i <- [1000, 2000, 3000, 
+                                          4000, 5000, 6000, 
+                                          7000, 8000, 9000, 
+                                          10000]]
 
 expSet = concat [[mkSymRegExp (n ++ "_" ++ show f) l f | n <- libNames | l <- libSet]
               | f <- frontierSizeSet]
 
[…]

### 2013-01-20 src/PI/PostProcess.hs

@@ -67,11 +67,13 @@ currentExpId exp = do timeStr <- fmap show getZonedTime
                       return expId
 
 
-saveSearchData :: String  -- ^ path to data dir
+saveSearchData :: Show a 
+               => String  -- ^ path to data dir
                -> Experiment 
                -> [SearchLogEntry] 
+               -> a -- ^ Time
                -> IO ()
-saveSearchData datadir exp searchData = 
+saveSearchData datadir exp searchData time = 
     do expDirName <- currentExpId exp
       
        let expPath = joinPath [datadir, expDirName]
@@ -79,19 +81,23 @@ saveSearchData datadir exp searchData =
            numhitFileName = joinPath [expPath, "numhit.csv"]
            explanationsFileName = joinPath [expPath, "explanations.csv"]
            experimentDataFileName = joinPath [expPath, "experimentdata.txt"]
+           timeFileName = joinPath [expPath, "time.txt"]
        createDirectoryIfMissing True expPath
        fgrammar <- openFile grammarFileName WriteMode
        fnumHit <- openFile numhitFileName WriteMode
        fexpl <- openFile explanationsFileName WriteMode
        fexpdata <- openFile experimentDataFileName WriteMode
+       ftime <- openFile timeFileName WriteMode
        hPutStr fgrammar (showLibByIter $ getLibData $ searchData)
        hPutStr fnumHit (showNumHitByIter $ getNumHitByIter $ searchData)
        hPutStr fexpl (showExplanationsByIter $ getExplanationsByIter $ searchData)
        hPutStr fexpdata (show exp)
+       hPutStr ftime (show time)
        hClose fgrammar
        hClose fnumHit
        hClose fexpl
        hClose fexpdata
+       hClose ftime
        
        
        

### 2013-01-20 src/PI/Run.hs

@@ -1,10 +1,12 @@
 -- Runs.hs
-
+{-# Language BangPatterns #-}
 module Run where
 
 import System.IO
 import qualified Data.Map as Map
 import Control.Monad.State
+import Control.Exception (evaluate)
+import Data.Time.Clock
 import Debug.Trace
 
 import Type
@@ -21,8 +23,12 @@ import PostProcess
 import Grammar
 
 runExp :: Experiment -> IO ()
-runExp exp = let (_, searchData) = runSearch $ loop exp
-             in saveSearchData "data" exp searchData
+runExp exp =   do startTime <- getCurrentTime 
+                  let ! (_, !searchData) = runSearch $ loop exp
+                  evaluate (last searchData)
+                  endTime <- getCurrentTime
+                  saveSearchData "data" exp searchData (diffUTCTime  endTime startTime)
+
 
 mkBruteForceExp :: Experiment -> Experiment
 mkBruteForceExp ex = let n = (expNumBound ex) * (expReps ex)
@@ -34,7 +40,7 @@ mkBruteForceExp ex = let n = (expNumBound ex) * (expReps ex)
 runBruteForceExp :: Experiment -> IO ()
 runBruteForceExp exp = let expBrute = mkBruteForceExp exp
                            (_, searchData) = runSearch $ loop expBrute
-                       in saveSearchData "data" expBrute searchData
+                       in saveSearchData "data" expBrute searchData "nan"
 
 
 

### 2013-01-20 src/PI/Search.hs

@@ -35,8 +35,8 @@ import qualified Compress as CP (getUniqueTrees, compress2)
 import Compress (Index)
 import Experiment
 
-data SearchLogEntry = SearchLogEntry { searchGrammar :: Grammar,
-                                       searchExplanation :: [(Task, Comb)]
+data SearchLogEntry = SearchLogEntry { searchGrammar :: ! Grammar,
+                                       searchExplanation :: ! [(Task, Comb)]
                                      } deriving Show
 
 
@@ -73,7 +73,7 @@ findCombinatorsForEachDatum :: Experiment
 findCombinatorsForEachDatum expr gr
     = do 
          db <- mkHypothesisSpace expr gr
-         return $ (trace $ show db) $ do 
+         return $ do 
            t <- taskSet
            let cs = case t of 
                       Task n f tp -> do c <- db Map.! tp
@@ -91,7 +91,7 @@ oneStep :: Experiment
         -> Search Grammar
 oneStep ex gr = do xs <- findCombinatorsForEachDatum ex gr
                    let xs' = filter (not . null . snd) xs
-                       xsShort =  map (\(t, cs) -> (t, take 5 cs)) xs'  
+                       xsShort =  map (\(t, cs) -> (t, take 3 cs)) xs'  
                        rs = (trace $  "Hit: " ++ show (xsShort))
                             $ dfs xsShort
                        ind = CP.compress2 (map (\(tsk, c) -> (taskType tsk, c))  rs)

### 2013-01-20 src/PI/python/symreg.py

@@ -35,6 +35,25 @@ numhitFiles = [
                  "frontier_size" : 1000}
              ]
 
+numhitFiles2 = [ 
+    {"filename" : "../data/R2_1000_2013-01-18_00-18-15.884901_EST/numhit.csv",
+     "frontier_size" : 1000},
+    {"filename" : "../data/R2_2000_2013-01-18_00-21-10.850648_EST/numhit.csv",
+     "frontier_size" : 2000},
+    {"filename" : "../data/R2_3000_2013-01-18_00-27-02.711662_EST/numhit.csv",
+     "frontier_size" : 3000},
+    {"filename" : "../data/R2_4000_2013-01-18_06-05-11.35682_EST/numhit.csv",
+     "frontier_size" : 4000},
+    {"filename" : "../data/R2_5000_2013-01-17_15-27-53.396156_EST/numhit.csv",
+     "frontier_size" : 5000},
+    {"filename" : "../data/R2_6000_2013-01-18_07-35-41.597988_EST/numhit.csv",
+     "frontier_size" : 6000},
+    {"filename" : "../data/R2_7000_2013-01-18_08-35-24.45969_EST/numhit.csv",
+     "frontier_size" : 7000},
+    {"filename" : "../data/R2_8000_2013-01-18_10-23-01.714354_EST/numhit.csv",
+     "frontier_size" : 8000},
+    ]
+
 def loadNumhitFile(filename):
     return pd.read_csv(filename, header=None).T
     

### 2013-01-21 drafts/ijcai2013/ijcai13.tex
commandline arugments for symbolic regressioin experiment.

@@
   \begin{subfigure}[Before]{0.5\linewidth}
-    \begin{tikzpicture}[scale=.7,
+    \begin{tikzpicture}[scale=.9,
         transform shape, level/.style={sibling distance=10mm/#1, font=\scriptsize}
@@
     \end{tikzpicture}
-    \caption{The typed combinator $S * I$ ($f(x)=x^2$) represented as binary tree.}
     \label{fig:clbintree}
@@
     \end{tikzpicture}
-    \caption{The space of all typed combinators represented as an
-      infinitely deep And-Or tree. Any specific combinator as a
-      sub-binary tree of this And-Or tree whose leaves consist of only
-      the $f_i$ nodes (i.e., the primitive combinators of the
-      grammar).}
     \label{fig:andor}
     \end{subfigure}
-  \caption{A typed binary tree representation of the combinator $S\;*\;I\;$.}
+  \caption{(a) The typed combinator $S * I$ ($f(x)=x^2$) represented as
+    binary tree. (b) The space of all typed combinators represented as an
+    infinitely deep AND/OR tree. Any specific combinator as a
+    sub-binary tree of this And-Or tree whose leaves consist of only
+    the $f_i$ nodes (i.e., the primitive combinators of the grammar).}
+
 \end{figure}
@@
 depth first order. We count the occurences of every subtree. Every
-time we encounter a subtree for the second time, we designate it as a
-primitive in our new grammar. Furthermore, during our traversal, we do
-not descend into trees that have been added already to the new
+time we encounter a subtree for the \emph{second} time, we designate
+it as a primitive in our new grammar. This ensures that every
+primitive is used more than once. During our traversal, we do not
+descend into trees that have been added already to the new
 grammar. Thus, we only count an instance of tree if that instance is
@@
 
-We want to count the number of
-times every subtree is encountered when it is not a subtree of a tree
-already encountered in our traversal. we simply count the numer of
-times each unique subtree appears as the child of a subtree \emph{not
-  yet encountered} in the solution set. That is, we traverse the
-solution set in depth first order, comparing each tree we encounter to
-all the trees we previously encountered; but if we encounter a tree we
-have already seen, we do not descend into it. In this traversal, the
-set of trees we encounter more than once is the set of terminal nodes
-in our new grammar. To estimate the probabilities associated with each
-node production, we again traverse the solution set, but these time
-for each node where a terminal production takes place we count which
-other terminal productions could have taken place. We estimate the
-terminal production weight to be the ratio of the number of times that
-node was used in the first traversal divided by the number of times it
-was used in the second traversal (plus 1, to account for the event of
-an $AND$ node expansion). Likewise, we re-estimate the probability of
-an $AND$ node expansion by diving the number of such expansions in the
-solution set by the number of possible expansions in the solution set.
+At the end of this traversal of the solution programs, we have a parse
+of the solution set with a new set of primitives and we know how many
+times each primitive occured in the solutions et. To estimate the
+probabilities associated with each node production, we again traverse
+the solution set, but this time for each node we count which other
+primitive elements from the new grammar could have been used. We
+estimate the terminal production weight to be the ratio of the number
+of times that node was used in the first traversal divided by the
+number of times it was used in the second traversal.
+
+Note that this probability estimation procedure was chosen for its
+simplicity, and it is not the maximum likelihood estimator for the
+multinomial distribution according to which we enumerate the
+data. Because the probability we assign to the use of a primitive
+combinator depends on the type of the node, the ML estimator of the
+probability of that combinator is not a simple ratio of counts. It
+would be interesting to determine whether a more accurate estimator
+can improve this algorithm's performance.
 

### 2013-01-21 src/PI/Experiments/SymbolicRegression.hs

@@ -3,6 +3,7 @@
 
 module Experiments.SymbolicRegression where
 
+import System.Environment (getArgs)
 import Control.Monad.Random
 import Control.Monad.State 
 import qualified  Data.HashMap as HMap
@@ -140,22 +141,20 @@ mkSymRegExpM name lib frontierSize
 
 libSet = [lib1] -- lib3, lib123]
 libNames = ["R1"] -- , "R3", "R<=3"]
-frontierSizeSet = join [replicate 1 i | i <- [8000, 9000,
-                                          10000]]
+frontierSizeSet sizes = join [replicate 1 i | i <- sizes]
+                                          
 
-expSet = concat [[mkSymRegExp (n ++ "_" ++ show f) l f | n <- libNames | l <- libSet]
-              | f <- frontierSizeSet]
+expSet sizes = concat [[mkSymRegExp (n ++ "_" ++ show f) l f | n <- libNames | l <- libSet]
+              | f <- frontierSizeSet sizes]
 
-expSetRand = sequence $ concat [[mkSymRegExpM (n ++ "_" ++ show f) l f | n <- libNames | l <- libSet]
-              | f <- frontierSizeSet]
+expSetRand sizes = sequence $ concat [[mkSymRegExpM (n ++ "_" ++ show f) l f | n <- libNames | l <- libSet]
+              | f <- frontierSizeSet sizes]
 
-runExpSet = sequence $ (map runExp expSet) ++ (map runBruteForceExp expSet)
+runExpSet sizes = sequence $ (map runExp (expSet sizes)) ++ (map runBruteForceExp (expSet sizes))
 
-runExpSetRand = do n <- newStdGen
-                   let (exps, a) = runRand expSetRand n
-                   sequence $ map runExp exps
-
-
+runExpSetRand sizes = do n <- newStdGen
+                         let (exps, a) = runRand (expSetRand sizes) n
+                         sequence $ map runExp exps
 
 expSymReg 
     = Experiment {
@@ -170,7 +169,8 @@ expSymReg
       }
 
 main = do
-  runExpSetRand
+  sizes <- fmap (map read) getArgs
+  runExpSetRand sizes
 
 
 

### 2013-01-22 drafts/ijcai2013/ijcai13.tex
Added digitial arithmetic stuff.

@@
 
-In Figure~\ref{fig:symreg}, we show performance results as we vary
-both the routers included and the frontier size. Each individual plot
-shows how changing the frontier size changes the number of tasks the
-algorithm identifies correctly over the course of ten algorithm
-iterations. In addition, we have a baseline for each curve, visits the
-same number of expressions as the corresponding algorithm run
-(i.e. frontier size $\times$ $\#$ of iterations) but all at once; it
-does not do any learning. Thus, this baseline serves as a brute-force
-control for the learning aspect of the algorithm.
+In Figure~\ref{fig:symreg}, we show performance results as we vary the
+frontier size. Each individual plot shows how changing the frontier
+size changes the number of tasks the algorithm identifies correctly
+over the course of fifteen algorithm iterations. In addition, we have a
+baseline for each curve, visits the same number of expressions as the
+corresponding algorithm run (i.e. frontier size $\times$ $\#$ of
+iterations) but all at once; it does not do any learning. Thus, this
+baseline serves as a brute-force control for the learning aspect of
+the algorithm.
 
 \begin{figure}
-
-\includegraphics[width=\linewidth]{figures/learningCurves.pdf}
-\includegraphics[width=\linewidth]{figures/performanceFrontierSize.pdf}
+\begin{subfigure}[Before]{0.45\linewidth}
+\includegraphics[width=\linewidth]{figures/learningCurves2.pdf}
+\end{subfigure}
+\begin{subfigure}[After]{0.45\linewidth}
+\includegraphics[width=\linewidth]{figures/performanceVsFrontierSize2.pdf}
+\end{subfigure}
 \caption{E-C learning curves for the symbolic regression task set as

### 2013-01-22 src/PI/EnumBF.hs

@@ -96,7 +96,9 @@ expandToApp gr (CombBase (CTerminal tp) (Just []) v) =
 expand :: Grammar 
        -> CombBase 
        -> StateT Int [] CombBase
-expand gr cb@(CombBase (CTerminal tp) (Just []) v) = cbs `mplus` cbApp
+expand gr cb@(CombBase (CTerminal tp) (Just []) v) 
+    = if cs_is_empty then mzero else 
+          cbs `mplus` cbApp
     where primCombs = CM.keys (library gr)
           cs = filterCombinatorsByType primCombs tp
 --          minGrammarVal = minimum (CM.elems (library gr))
@@ -105,6 +107,7 @@ expand gr cb@(CombBase (CTerminal tp) (Just []) v) = cbs `mplus` cbApp
                      let value = (calcLogProb gr tp c)
                          combBase = CombBase c Nothing value
                      return combBase
+          cs_is_empty = False -- (length $  runStateT cs 0) == 0
           cbApp = expandToApp gr cb 
 
 expand gr cb@(CombBase (CApp c_left c_right tp d) (Just (L:rest)) v) = 

### 2013-01-22 src/PI/Experiments/DigitalArithmetic.hs

@@ -1,9 +1,12 @@
 -- DigitalArithmetic
+{-# Language ParallelListComp #-} 
 
 module Experiments.DigitalArithmetic where
 
 import Control.Monad.State
 import Data.Maybe
+import qualified Data.HashMap as HMap
+import Control.Monad.Random
 import Debug.Trace
 
 import Experiment
@@ -84,18 +87,96 @@ taskIncr1 = Task "pair incr 1"
                             (pairIn2, pairOut2)]
                   ismatch c cin cout = pairEq (app' c cin) cout
 
+taskNot = Task "not" f (tBool ->- tBool)
+    where f c = if a &&  b then  0 else 1
+              where
+                a = reduceComb (app' c cTrue) == reduceComb cFalse
+                b = reduceComb (app' c cFalse) == reduceComb cTrue
 
+taskAnd = Task "and" f (tBool ->- tBool ->- tBool)
+    where f c = if a1 && a2 && a3 && a4  then  0 else 1
+              where
+                a1 = reduceComb (app' (app' c cTrue) cTrue) == reduceComb cTrue
+                a2 = reduceComb (app' (app' c cTrue) cFalse) == reduceComb cFalse
+                a3 = reduceComb (app' (app' c cFalse) cTrue) == reduceComb cFalse
+                a4 = reduceComb (app' (app' c cFalse) cFalse) == reduceComb cFalse
+
+taskOr = Task "or" f (tBool ->- tBool ->- tBool)
+    where f c = if a1 && a2 && a3 && a4  then  0 else 1
+              where
+                a1 = reduceComb (app' (app' c cTrue) cTrue) == reduceComb cTrue
+                a2 = reduceComb (app' (app' c cTrue) cFalse) == reduceComb cTrue
+                a3 = reduceComb (app' (app' c cFalse) cTrue) == reduceComb cTrue
+                a4 = reduceComb (app' (app' c cFalse) cFalse) == reduceComb cFalse
+
+taskXor = Task "xor" f (tBool ->- tBool ->- tBool)
+    where f c = if a1 && a2 && a3 && a4  then  0 else 1
+              where
+                a1 = reduceComb (app' (app' c cTrue) cTrue) == reduceComb cFalse
+                a2 = reduceComb (app' (app' c cTrue) cFalse) == reduceComb cTrue
+                a3 = reduceComb (app' (app' c cFalse) cTrue) == reduceComb cTrue
[…]

### 2013-01-22 src/PI/Search.hs

@@ -49,7 +49,7 @@ mkHypothesisSpace :: Experiment
                   -> Grammar 
                   -> Search (Map.Map Type [Comb])
 mkHypothesisSpace expr gr
-    = do return db
+    = do return $ db
     where tps = nub [ taskType t | t <- taskSet]
           cs tp = map comb $ enumBF gr (expNumBound expr) tp
           db =  foldl' (\m (k, a) -> Map.insert k a m) 
@@ -60,6 +60,12 @@ mkHypothesisSpace expr gr
 
 evalSymReg :: Comb -> [Int]
 evalSymReg c = [ f i | i <- [0..10]]
+    where f i = case reduceComb $ CApp c (num2C i) tInt 0 of 
+                  (N y ) -> y
+                  otherwise -> (maxBound :: Int)
+
+evalDigArith :: Comb -> [Bool]
+evalDigArith c = [ f i | i <- [0..10]]
     where f i = case reduceComb $ CApp c (num2C i) tInt 0 of 
                   (N y ) -> y
                   otherwise -> (maxBound :: Int)
@@ -80,6 +86,8 @@ findCombinatorsForEachDatum expr gr
                                         if f c <= eps then return c else mzero
                       SymRegTask n vs tp -> do c <- db Map.! tp
                                                if evalSymReg c == vs then return c else mzero
+                      DigArithTask n vs tp -> do c <- db Map.! tp
+                                                 if evalDigArith c == vs then return c else mzero
            return (t, cs) 
       where 
         taskSet = expTaskSet expr

### 2013-01-22 src/PI/Task.hs

@@ -14,11 +14,16 @@ data Task =  Task {taskName :: String,
           | SymRegTask {taskName :: String,
                         taskVals :: [Int],
                         taskType :: Type}
+          | DigArithTask {taskName :: String,
+                          taskVals :: [[Bool]],
+                          taskType :: Type}
+
 type TaskSet = [Task]
 
 instance Show Task where
     show (Task n _ _ ) = n
     show (SymRegTask n _ _ ) = n
+    show (DigArithTask n _ _ ) = n
 
 
 -- | task constructors

### 2013-01-22 src/PI/python/symreg.py

@@ -99,14 +99,13 @@ def performanceVsFrontierSize(numhitDict, ax=None):
     ax.set_ylabel(ylabel)
     return ax
 
-    
+
+
     
     
                    
                    
 
-
-    
 
 
 if __name__ == "__main__":

### 2013-01-25 drafts/ijcai2013/ijcai13.tex
Changes.

@@
 \usepackage{tikz}
+\usepackage{circuitikz}
 \usepackage{graphicx}
 \usepackage{subcaption}
+\usepackage{alltt}
 \usepackage[ruled,vlined,linesnumbered]{algorithm2e}
-\usetikzlibrary{arrows,automata, positioning}
+\usetikzlibrary{arrows,automata, positioning, patterns}
 
@@
 
-\title{The E-C Algorithm: conceptual bootstrapping via grammar-based
-  program compression.}
+\title{Getting off the ground: bootstrap learning via modular concept discovery.}
 
@@
 \And
+Jon Malmaud \\
+MIT \\ 
+USA \\
+malmaud@mit.edu
+\And 
+Ryan P.  Adams \\
+Harvard University\\
+USA \\
+rpa@seas.harvard.edu
+\And
 Josh Tenenbaum \\
@@
 jbt@mit.edu
-\And 
-Ryan Adams \\
-Harvard University\\
-USA \\
-rpa@seas.harvard.edu}
+}
 
@@
 \begin{abstract}
-Suppose a learner is faced with a set of problems about which it knows
-nearly nothing: it doesn't know the distribution of problems, or what
-an appropriate loss function might be. For each hypothesized solution
-to a problem it gets a few bits of positive or negative reinforcement,
-but not enough to steer it effectively in the huge space of possible
-solutions. How can such a learner every get off the ground? The key is
-to find, by brute force, a solution to one or more simple problems in
-that domain and use the information thus gained to pursue subsequent
-searches in an informed manner. Here, we formalize this intuition,
-where the solution space is the space of typed functional programs and
-the gained information is stored as a stochastic grammer over
-programs. Specifically, we propose the E-C algorithm, a two step
-iterative procedure for exploring such spaces: in the E step, the
-learner \emph{explores} a finite subset of the domain for solutions,
-guided by a stochastic grammar; in the C step, the learner
-\emph{compresses} the successful solutions from the E step to estimate
-a new stochastic grammar. As we show in several toy domains, iterating
+ Suppose a learner is faced with a domain of problems about which she
+ knows nearly nothing. She does not know the distribution of problems,
+ the space of solutions is not smooth, and the reward signal is
+ uninformative, providing perhaps a few bits of information but not
+ enough to steer her effectively. How can such a learner ever get off
+ the ground? A common intuition is that if the solutions to these
+ problems share a common structure, and the learner can solve some
+ simple problems by luck or brute force, she should be able to extract
+ useful modular concepts from these solutions and, by composing them,
+ explore the solution space more efficiently. Here, we formalize this
+ intuition, where the solution space is the space of typed functional
+ programs and the gained information is stored as a stochastic grammar
+ over programs. We propose a two step iterative procedure for
+ exploring such spaces: in the first step, the learner explores a
+ finite subset of the domain, guided by a stochastic grammar; in the
+ second step, the learner compresses the successful solutions from the
+ first step to estimate a new stochastic grammar. As we show in a
+ symbolic regression and a boolean circuit learning example, iterating
  these two steps quickly propels the learner from a state of solving
@@
 
-\section{Introduction.}
+\section{Introduction}
 
-Suppose that, having been shipwrecked and saved by a people with no
-knowledge of your language or culture, you find yourself attempting to
-communicate in a language about which you know nothing. How would you
-go about discovering even the most basic mappings of sounds to
-meanings? Overhearing the native speech, you have a sense of what the
-phonetics might be like, so you might try various common combinations
-of sounds while pointing to things you want. Mostly, you receive
-puzzled looks, but if you are lucky, you might pick up on a few key
-words or phrases, like ``that'' or ``I want.'' Naturally, you will
-proceed to use your discoveries to accelerate your learning, mixing
-and matching known phrases to generate new meanings.
+Imagine you are trying to learn to bake via trial and error. You know
+what kind of ingredients are used in baking (flour, water, butter,
+sugar, yeast, baking soda, etc.) and you know a bit about how to
+combine these ingredient (mix, whisk, etc.) and how to apply heat
+(oven). But baking is a notoriously fragile process: ratios of
+ingredients, the order of mixing, cooking temperatures, resting times,
+all make a huge difference, and the outcome of the baking process is
+mostly all or none. Either the outcome is what you expected or quite
+unlike it. Even if you do create something similar to your intent, it
+is difficult to assign blame for the outcome to any one step or
+decision along the way.
 
-\section{Related work.}
+Many difficult learning problems have this structure -- learning
+language, developing technologies and scientific theories, learning to
+play instruments or execute other complex motor patterns. These are
+problems in which 1) there is a large set of similar tasks; 2) their
+solutions are best described as programs; 3) the solution space is
+not smooth with respect to the outcome; 4) and the outcome is generally
+uninformative unless you are already pretty close to a solution.
 
-\section{A multi-task program induction problem.}
-Our goal is to solve a general multi-task learning problem: we are
+Intuitively, the way to solve such problems is to stumble on solutions
+to simple cases in the domain and extract conceptual building blocks
+that can be composed to scale up to complex cases. To return to the
+baking example, cookies are much simpler than cakes. There are fewer
+steps to get just right. But once you have figured out a few simple
+pastries, the more complex ones are mostly just variations and
+combinations of techniques learned along the way.
+
+Of course, knowledge is not acquired wholly by trial and error. But in
+many domains, there is at least initially such a period of ignorance,
+and the question we are trying to tackle here is how you ever get off
+the ground.
+
+To frame this learning paradigm computationally, we make two modeling
+choices. First, to emphasize the impoverished reward signals in these
+domains, we model the learning task as an all-or-none optimization
+problem: either the learner gets some reinforcement or she does
+not. Second, to emphasize the lack of locally smooth structure in the
+space of possible solutions, we choose as our solution space the space
+of functional programs. These two choices are in contrast to the
+paradigm tackled by most of machine learning, in which a locally
+informative objective function guides search in a space that is
+somewhat smooth.
+
+Formally then, our goal is to solve the following minimal feedback
+multi-task program induction problem: suppose we are
 given set of \emph{tasks} $T=\{t_k\}_{k=1}^K$ where each task is a
-function $t_i : \mathcal{L} \rightarrow \mathds{R} > 0$ which returns
-a reward value for every element in the solution representation
-language $\mathcal{L}$. Our goal is to find a set of solutions
-$X = \{x_k\}_{k=1}^k \subset \mathcal{L}$ such that
+function $t_i : \mathcal{L} \rightarrow \{0, 1\}$, where $\mathcal{L}$
+is the set of programs in our language.  We say that $ p \in
+\mathcal{L}$ \emph{solves} $t_i$ if $ t_i\; p = 1$. Our
+goal is to find a set of solutions $X = \{x_k\}_{k=1}^k \subset
+\mathcal{L}$ such that
 
@@
 
-\section{A compression scheme expresses a belief about the data.}
-There is a deep connection between a distribution over data and the
-optimally compressive code for that data. In fact, if we know the
-probability distribution for a set of observations, we can obtain an
-optimally compressive encoding of those observations via arithmetic
-coding. Conversely, the encoding generated by a good compression
-scheme can be interpreted as an approximation of the data
-distribution.
+Based on the intuition set out above, we present an iterative
+strategy. Given a stochastic grammar $\mathcal{G}$ which defines a
+distribution over the primitives from $\mathcal{L}$ and a integer
+\emph{frontier size} $N$, we perform to steps in each iteration.  In
+the first step, we explore the $N$ highest probability programs
+according $\mathcal{G}$. In the second step, we compress the
+programs that solve any of our tasks, and we use that compression to
+define a new grammar. Repeating these two steps, we effectively
+reparameterize our search space to include programs that are more likely
+to be solutions to our tasks.
 
-Take, for example, the Nevill-Manning algorithm for grammar based
-compression~\cite{nevill1997identifying}. This algorithm looks for
-hierarchical structure in the data and represents that hierarchical
-structure as a grammar. The output of the algorithm can be seen as a
-grammar and a corresponding parse of the input data. Given that parse,
-we can estimate the maximum likelihood probabilistic context-free
-grammar for that data~\cite{johnson1998pcfg}.
+The remainder of this paper will include three major sections. First,
+we will describe the representation language we chose and the goals of
+the exploration and compressions steps. Following that, we will
+discuss the algorithmic details that render our procedure
+feasible. Finally, we present results from two experiments, involving
+the domains of symbolic regression and boolean circuit learning.
 
-\section{The E-C Algorithm.}
+%% \section{}
+%% Let $\mathcal{F}$ be a family of distributions over $\mathcal{L}$ such
+%% that for any $\mathcal{D} \in \mathcal{F}$ we can a) evaluate the
+%% probability of program $p$, $P_D(p)$, and we can b) enumerate elements
+%% from $D$ in descending order.
 
-In this section, we describe the E-C algorithm and discuss its uses,
-applicability, and limitations.
+%%  To motivate a solution to our problem, we begin with the hierarchical
+%% Bayesian formulation of multi-task program induction found in Liang et
+%% al.~\cite{liang10programs}. The general idea there is to estimate
+%% solution programs for a set of related problems by inferring the posterior
+%% distribution over solutions under the assumption that all the solution
+%% programs are generated from a common latent stochastic grammar. By
+%% jointly inferring solutions and the latent grammar, we find a set of
+%% solutions that favors reuse of subcomponents.
 
-As its name suggests, each iteration of the E-C algorithm consists of
-an exploration step and a compression step. In the exploration step,
-we enumerate the $N$ most probable hypotheses given our currenst
-distribution $D$ over hypotheses. We then assign to each task $t_i$
-the hypotheses that match it. We say that a task is \emph{hit} if
-there is a least one hypothesis that matches it. In the compression
-step, our goal is to assign a single hypothesis $h_i^*$ to each task
-$t_i$ such that that resulting set of hypotheses $\{t_i\}$ is as
-compressible as possible. The compression this process induces is used
-to update distribution $D$.
+%% While this framework is appealing in its elegance and Bayesian
+%% interpretation, it cannot be straightforwardly adapted to the
+%% situation we are considering here because their proposed MCMC
+%% algorithm relies on a local log-likelihood signal, the very kind that
+%% we are assuming we do not have access to.
 
-  \begin{algorithm}
-    \SetAlgoLined \KwData{Tasks $T=\{t_1, \dots, t_K\}$,
-      distribution $\mathcal{D}_j$ over expressions.}
+%% To begin, we provide a high level overview of the algorithm and the
+%% problem that it solves. The problem setup requires a representation
+%% language for programs and a family of distributions over expressions
+%% in that language, but we leave these details for the following section
 
-    $\text{types} \leftarrow \cup_k \, type(t_k)$. \\
-    For $\tau \in \text{ types}$ , $H_N(\tau) \leftarrow$ enumerate
-    $N$ most probable elements in $\mathcal{D}_t$ with type $\tau$.\\
+\section{Extracting modular concepts from functional programs}
 
-    For each \emph{hit} $s_i$ in 
-    $T$, $v_{s_i} \leftarrow \{h \in D_j | s_i(h) == True \}$.\\
+In order to solve our stated problem, we need to choose a solution
+representation language and a compression criterion, and we need to
+describe the frontier enumeration and compression procedures.
 
-    $S = \{(s_i, h^*_{s_i})\} \leftarrow$ assignment of elements to tasks s.t. 
-    $S$ is most compressible. \\
-
-    $D_{j+1} \leftarrow$ derive new distribution from compression of $S$. \\
-
-    Repeat.
-
-    \caption{E-C}
-  \end{algorithm}
-
-\section{Our implementation.}
-
-\subsection{Representing polymorphically typed programs as binary trees.}
+\subsection{Representing typed functional programs as binary trees.}
 
@@
 combinatory logic -- a variable-free subset of the polymorphic
-simply-typed lambda calculus lambda calculus~\cite{Pierce_2002} -- as
+simply-typed lambda calculus~\cite{Pierce_2002} -- as
 our program representation language. In short, the combinatory logic
@@
 
-The basis combinators can
-themselves been expressed in the lambda calculus. The lambda calculus
-has two basic operation -- application and abstraction -- but in using
-the combinatory logic we sequester uses of the abstraction operation
-inside the combinators, and thus have a representation language that
-is effectively abstraction- and, thus, variable-free.
+The basis
+combinators are theoretically enough to express any Turing computable
+function, but we will assume that our language has a variety of
+primitives and, together with the basis combinators, we call these the
+primitive combinators.
+
+The basis combinators can themselves been expressed in the lambda
+calculus. The lambda calculus has two basic operation -- application
+and abstraction -- but in using the combinatory logic we sequester
+uses of the abstraction operation inside the combinators, and thus
+have a representation language that is effectively abstraction- and,
+thus, variable-free. In doing so, we have replaced the variable
+binding of the $\lambda$ operator with the variable \emph{routing} of
+these basis combinators. See~\cite{liang10programs} for a more
+detailed discussion of this routing interpretation.
 
@@
 As a simple example of our representation, consider the squaring
-function $\lambda x. * x\; x $. Using two of the basis combinators above,
-we can write the squaring function in combinatory logic as $S * I$,
-where application associates to the left, so we drop the
-parentheses. When we apply this combinator to a value $x$, the action
-of the combinator is defined as $S \;*\; I\; x~\rightarrow~(*\; x)\;
-(I\; x)~\rightarrow(*\; x)\; x$. Now, we can extend this
-representation with a simple polymorphic type
+function $\lambda x. * x\; x $. Using two of the basis combinators
+above, we can write the squaring function in combinatory logic as $S *
+I$ (taking $*$ as a primitive), where application associates to the
+left, so we drop the parentheses. When we apply this combinator to a
+value $x$, the action of the combinator is defined as $S \;*\; I\;
+x~\rightarrow~(*\; x)\; (I\; x)~\rightarrow(*\; x)\; x$.
+
+We can extend this representation with a simple polymorphic type
 system~\cite{Pierce_2002}. Thus, this combinator can be represented as
 a binary tree~(Figure~\ref{fig:clbintree}) whose leaves are typed
-primitive combinators and whose interior nodes are typed applications
+primitive combinators and whose interior nodes represent typed applications
 of combinators.
@@
         ]
-      \node [font=\scriptsize] (top) {$I$}
-      child {node [above=.3cm] (a) {$(I \rightarrow I) \rightarrow I \rightarrow I$}
+      \node [font=\scriptsize] (top) {$Int$}
+      child {node [above=.3cm, text width=2cm] (a) {$(Int \rightarrow Int) \rightarrow Int 
+          \rightarrow Int$}
         child {node [left=.5cm, below=1cm, text width=3cm] 
@@
       }
-      child {node [below right=.3cm and 1cm] (b) {$I \rightarrow I$}
+      child {node [below right=.3cm and 1cm] (b) {$Int \rightarrow Int$}
         child {node {$I: t_0 \rightarrow t_0 \rightarrow t_0$}}
@@
     binary tree. (b) The space of all typed combinators represented as an
-    infinitely deep AND/OR tree. Any specific combinator as a
-    sub-binary tree of this And-Or tree whose leaves consist of only
+    infinitely deep AND/OR tree. Any specific combinator is a
+    sub-binary tree of this AND/OR tree whose leaves consist of only
     the $f_i$ nodes (i.e., the primitive combinators of the grammar).}
@@
 
-\subsection{Best-first enumeration of most probably programs.}
+\subsection{A stochastic grammar over programs.}
+A stochastic grammar over programs represents a distribution over
+expressions in $\mathcal{L}$. We represent our grammar as a
+multinomial distribution over the primitives in our library. Let $p_1,
+\dots, p_m$ be the probabilities associated with primitive combinators
+$c_1, \dots, c_m$. Then we define the probability $p(c_i | t) $ of
+using $c_i$ at a node whose type is $t$ to be $p_i$ normalized by the
+probability mass of all other primitive combinators that could have
+been used at that node. Namely, $p(c_i | t) = \frac{p_i}{\sum_{c_j :
+    type(c_j) \text{ unifies with } t} p_j}$ if type($c_i)$ unifies
+with $t$, 0 otherwise. Note that the distribution thus defined is not
+normalizable if we allow expressions of arbitrary depth, but this is
+not a practicle concern since we are using the distribution mainly as
+a search heuristic.
+
+\subsection{Best-first enumeration of most probable programs.}
 In order to explore the frontier of most promising programs, we need a
-procedure that enumerates the $N$ most probable programs. We can
-formulate this naturally as a best-first exploration of an $AND/OR$
-tree~\cite{nilsson1982principles}, which we denote $\mathcal{G}$. A
-complete program is any binary subtrees of $\mathcal{G}$ whose leaves
-coincide with the leaves of $\mathcal{G}$. The tree structure is shown
-in Figure~\ref{fig:andor}. The root is an $OR$ node of the requisite
+procedure that enumerates the $N$ most probable programs. There has
+been recent interest in this
+problem~\cite{DBLP:conf/sfp/Katayama05}\cite{DBLP:conf/aaip/YakushevJ09}. Our
+approach is to formulate the problem as a best-first exploration of an
+AND/OR tree~\cite{nilsson1982principles}, which we denote
+$\mathcal{G}$. A complete program is any binary subtrees of
+$\mathcal{G}$ whose leaves coincide with the leaves of
+$\mathcal{G}$. The tree structure is shown in
+Figure~\ref{fig:andor}. The root is an OR node of the requisite
 program type. It's children are the primitives of that type $f_1,
-\dots, f_n$, and one $AND$ node with two children. Each child of an
-$AND$ node has the same structure as the root node, with modified
-types: if the type of an $AND$ node is $t$, then the type of its left
-child is $\tau \rightarrow t$ (where $\tau$ is a fresh type
-variable). Since the type of the right child is constrained by the
-subtree rooted at the left child, we always expand the left child
-first and use the target type of this left subtree to mark the type of
-the right child.
+\dots, f_n$, and one AND node with two children. Each child of an AND
+node has the same structure as the root node, with modified types: if
+the type of an AND node is $t$, then the type of its left child is
+$\tau \rightarrow t$ (where $\tau$ is a fresh type variable). Since
+the type of the right child is constrained by the subtree rooted at
+the left child, we always expand the left child first and use the
+target type of this left subtree to mark the type of the right child.
 
 In accordance with the stochastic grammar over programs, we weight all
-edges from $OR$ nodes according to the negative log likelihood of
+edges from OR nodes according to the negative log likelihood of
 choosing the associated rule from the stochastic grammar, conditioned
-on the requesting type. Expanding an $AND$ node has zero cost. In the
-case that the stochastic grammar is ill condition, we might have an
+on the requesting type. Expanding an AND node has zero cost. In the
+case that the stochastic grammar is ill conditioned, we might have an
 infinitely deep solution tree whose cost does not increase as we
@@
 the required type since the log probability of expansion to the next
-$AND$ node is 0. . To avoid this, we simply declare a cost for
-expanding the next $AND$ node when it is the only node available.
+AND node is 0. To avoid this, we simply declare a cost for
+expanding the next AND node when it is the only node available.
 
-\subsection{Finding the most compressive set of solution.}
+\subsection{Finding the most compressive set of solutions.}
 
@@
 In our algorithm, we traverse a set of combinator binary trees in
-depth first order. We count the occurences of every subtree. Every
+depth first order. We count the occurrences of every subtree. Every
 time we encounter a subtree for the \emph{second} time, we designate
@@
 of the solution set with a new set of primitives and we know how many
-times each primitive occured in the solutions et. To estimate the
+times each primitive occurred in the solutions set. To estimate the
 probabilities associated with each node production, we again traverse
@@
 
-The choice of an initial grammar for the algorithm will, of course,
-have a big effect on its performance, and so we would like to
-characterise this effect. We can divide primitives into roughly two
-sets; there are low-level primitives, those that operate on concrete
-types like Bool or Double, and these tend to be domain specific. For
-example, conrete primitives include $1, 2$, etc., and $f(x) = 3x$. The
-other class of primitives are abstract primitives, which can be
-characterized by having arguments whose types are primarily function
-types, like $(t \rightarrow s), (t \rightarrow s) \rightarrow \text{
-  Double },$ etc. Of course, we can construct operators that fall in
-between.
+%% The choice of an initial grammar for the algorithm will, of course,
+%% have a big effect on its performance, and so we would like to
+%% characterize this effect. We can divide primitives into roughly two
+%% sets; there are low-level primitives, those that operate on concrete
+%% types like Bool or Double, and these tend to be domain specific. For
+%% example, concrete primitives include $1, 2$, etc., and $f(x) = 3x$. The
+%% other class of primitives are abstract primitives, which can be
+%% characterized by having arguments whose types are primarily function
+%% types, like $(t \rightarrow s), (t \rightarrow s) \rightarrow \text{
+%%   Double },$ etc. Of course, we can construct operators that fall in
+%% between.
 
-The most abstract operators are those that Liang et. al refer to as
-``routers''~\cite{liang10programs}. These operators take the place of
-$\lambda$ abstraction in the general lambda calculus: whereas
-$\lambda$ abstraction allows us to bind arguments to variables nested
-deep inside an expression, without them we need to have operators that
-``route'' arguments to the appropriate functions in a nested
-expression.
+%% The most abstract operators are those that Liang et. al refer to as
+%% ``routers''~\cite{liang10programs}. These operators take the place of
+%% $\lambda$ abstraction in the general lambda calculus: whereas
+%% $\lambda$ abstraction allows us to bind arguments to variables nested
+%% deep inside an expression, without them we need to have operators that
+%% ``route'' arguments to the appropriate functions in a nested
+%% expression.
 
-\begin{definition}[Router~\cite{liang10programs}] A \emph{router} is a combinator
-represented by a finite sequence of elements from $\{B, C, S\}$. The
-behavior of router $r$ is given by
-\[
-(r \, f \, g \, x_1 \dots x_{|r|}) = (f \, x_{i_1} \dots x_{i_n}) (g
-\, x_{j_1} \dots x_{j_n}),
-\]
-where $i_1 < \dots < i_n$ are indices $i$ such that $r_i \in \{C, S\}$
-and where $j_1 < \dots < j_n$ are indices $j$ such that $r_j \in \{B,
-S\}$. Let $\mathcal{R}_k$ be the set of all routers of length equal to
-$k$. Let $\mathcal{R}_{\leq k}$ be the set of all routers of length
-less than or equal to $k$.
-\end{definition}
+%% \begin{definition}[Router~\cite{liang10programs}] A \emph{router} is a combinator
+%% represented by a finite sequence of elements from $\{B, C, S\}$. The
+%% behavior of router $r$ is given by
+%% \[
+%% (r \, f \, g \, x_1 \dots x_{|r|}) = (f \, x_{i_1} \dots x_{i_n}) (g
+%% \, x_{j_1} \dots x_{j_n}),
+%% \]
+%% where $i_1 < \dots < i_n$ are indices $i$ such that $r_i \in \{C, S\}$
+%% and where $j_1 < \dots < j_n$ are indices $j$ such that $r_j \in \{B,
+%% S\}$. Let $\mathcal{R}_k$ be the set of all routers of length equal to
+%% $k$. Let $\mathcal{R}_{\leq k}$ be the set of all routers of length
+%% less than or equal to $k$.
+%% \end{definition}
 
-To analyze the behavior of our algorithm, we will vary the parameters
-of the problem along two axes while keeping the learning tasks the
-same. Along the first axis, we vary the initial library to which our
-learner has access with respect to the its abstractness (i.e., number
-of routers) and its concreteness (i.e. number of concrete
-primitives). Along the second axis we vary the size of the search
-frontier.
+%% To analyze the behavior of our algorithm, we will vary the parameters
+%% of the problem along two axes while keeping the learning tasks the
+%% same. Along the first axis, we vary the initial library to which our
+%% learner has access with respect to the its abstractness (i.e., number
+%% of routers) and its concreteness (i.e. number of concrete
+%% primitives). Along the second axis we vary the size of the search
+%% frontier.
 
@@
 inputs $0, \dots, 10$. The set of tasks corresponds to the set of
-polynomials $\{ax^2 + bx + c | a, b, c \in \{0, \dots, 10 \}\}$.
+polynomials $\{ax^2 + bx + c | a, b, c \in \{0, \dots, 10 \}\}$. The
+choice of primitives is going to make a big difference in performance;
+we chose to include the basic combinators I, S, B, C and four
+arithmetic primitives, $1$, $0$, $*$, and $+$. The choice of initial
+weights makes a large difference as well, as it determines the
+relative ordering of combinators in the enumeration step. To get a
+general sense of the algorithm's performance, we set the intial
+weights to be slightly different on each run, jittering them around a
+uniform weighting.
 
@@
 size changes the number of tasks the algorithm identifies correctly
-over the course of fifteen algorithm iterations. In addition, we have a
-baseline for each curve, visits the same number of expressions as the
-corresponding algorithm run (i.e. frontier size $\times$ $\#$ of
-iterations) but all at once; it does not do any learning. Thus, this
-baseline serves as a brute-force control for the learning aspect of
-the algorithm.
+over the course of fifteen algorithm iterations. %% In addition, we have a
+%% baseline for each curve, visits the same number of expressions as the
+%% corresponding algorithm run (i.e. frontier size $\times$ $\#$ of
+%% iterations) but all at once; it does not do any learning. Thus, this
+%% baseline serves as a brute-force control for the learning aspect of
+%% the algorithm.
 
@@
 \end{subfigure}
-\caption{E-C learning curves for the symbolic regression task set as
-  we vary the frontier size.}
+\caption{(a) Average learning curves as we vary the frontier size. (b)
+  Percent of tasks solved at iteration 15 as a function of frontier
+  size.}
 \label{fig:symreg} 
@@
 
-Why do the learning curves asymptote early? One might object that a
-learning algorithm should continuously improve, if only slowly, as it
-is given more time. But in these runs we observe a finite and
-permanent performance cap. This is expected: once the learner find a
-set of concepts it can explain, if those concepts fill up the
-frontier, then there is nothing to spur the learner to incorporate as
-yet unsolved tasks.
+%% Why do the learning curves asymptote early? One might object that a
+%% learning algorithm should continuously improve, if only slowly, as it
+%% is given more time. But in these runs we observe a finite and
+%% permanent performance cap. This is expected: once the learner find a
+%% set of concepts it can explain, if those concepts fill up the
+%% frontier, then there is nothing to spur the learner to incorporate as
+%% yet unsolved tasks.
 
-To investigate whether this is in fact the case, we can examine the
-proportion of expressions in the frontier that match tasks in our task
-set; that is, is the learning curve plateauing because the frontier is
-heavily redundant, or is it plateuing because the frontier contains
-many expressions that do not hit the task set at
-all. Figure~\ref{fig:plateau_investigation} shows the proportion of
-the expressions in the frontier that hit a task as the frontier sizeincreases.
+%% To investigate whether this is in fact the case, we can examine the
+%% proportion of expressions in the frontier that match tasks in our task
+%% set; that is, is the learning curve plateauing because the frontier is
+%% heavily redundant, or is it plateuing because the frontier contains
+%% many expressions that do not hit the task set at
+%% all? Figure~\ref{fig:plateau_investigation} shows the proportion of
+%% the expressions in the frontier that hit a task as the frontier size increases.
 
-\subsection{Digital arithmetic}
-Having demonstrated the behavior of the E-C algorithm on symbolic
-regression, we analyze its performance on tasks involving digital
-arithmetic, a class of problems with known modular structure. The
-general goal is to discover boolean circuits consisting of a limited
-set of gates that perform arithmetic operations on binary number
-representations.
+\subsection{Boolean circuits}
+
+As another demonstration of this approaches learning abilities
+investigate its ability to learn boolean functions using only the
+logical primitive NAND and the basic combinators. In
+Figure~\ref{fig:boolLearningCurves},
+
+\begin{figure}
+\begin{subfigure}[Before]{0.45\linewidth}
+\includegraphics[width=\linewidth]{figures/boolLearningCurves.pdf}
+\end{subfigure}
+\caption{(a) Average learning curves as we vary the frontier size. (b)
+  Percent of tasks solved at iteration 15 as a function of frontier
+  size.}
+\label{fig:boolLearningCurves} 
+\end{figure}
+
+What kind of solutions does our algorithm find to common boolean
+circuits? We might hope that the circuits it finds are ``simple.'' One
+way to judge this is to compare the circuits it finds with the
+circuits found in digital electronics textbooks. For example, the XOR
+circuit can be drawn as a circuit with only NAND gates as in
+Figure~\ref{fig:xor_circuit}. Our algorithm returns the solution:
+
+\begin{alltt}
+\small{
+((S (B C (((B (C B nand)) S) 
+(((B (C B nand)) S) (B C 
+((B (B nand)) nand)))))) I)}. 
+\end{alltt}
+This looks more complicated. It should be understood, though, that the
+routing combinators S, B, and C play the role of wiring. Note that
+this solution has four NAND gates, the same number of as in the
+standard circuit. Another interesting feature of this solution is that
+it is not the smallest expression that is equivalent (modulo
+reduction), since we can further reduce it. However, this reduction
+would remove the two uses of \texttt{\small{(B (C B nand) S)}}. This is
+a reused component that appears within the most highly weighted 20
+combinators in the algorithm's induced grammar. We can view it
+schematically with a circuit diagram, but it is not a circuit
+fragment. Rather, it is a function that takes in a circuit of type
+$bool \rightarrow bool \rightarrow bool$ and returns a circuit of type $bool
+\rightarrow bool \rightarrow bool$. Figure~\ref{fig:component} shows
+this component schematically.
+
+\begin{figure}
+\begin{circuitikz} \draw
+%%(1,0) node[nand port, thick] (nor1) {}
+(1,.75) node[nand port, thick] (nor1) {}
+(3,0) node[nand port, thick] (nor2) {}
+(3,1.5) node[nand port, thick] (nor3) {}
+(5,.75) node[nand port, thick] (nor4) {}
+
+(nor1.in 1) node[above left=.5cm] (x) {x}
+(nor1.in 2) node[below left = .5cm](y) {y}
+(x) -| (nor1.in 1)
+(y) -| (nor1.in 2)
+(x) -| (nor3.in 1)
+(y) -| (nor2.in 2)
+(nor1) -- (1.62, .75) -- (nor2.in 1)
+(nor1) -- (1.62, .75) -- (nor3.in 2)
+(nor2) -| (nor4.in 2)
+(nor3) -| (nor4.in 1);
+ \end{circuitikz}
+\caption{A standard XOR circuit built of NAND gates.}
+\label{fig:xor_circuit}
+\end{figure}
+
+\begin{figure}
+\begin{circuitikz} \draw
+
+%%(1,0) node[nand port, thick] (nor1) {}
+(3, 0)  node[ rotate=0, anchor=east,thick, draw, text width=1cm ] (f) {?} 
+(1, 0) node[nand port, thick] (nor) {}
+(nor.in 1) node[above left=.25cm] (x) {x}
+(nor.in 2) node[below left=.25cm] (y) {y}
+(y) -- (1.71,.5cm)
+(nor) -- (1.72,-.5cm)
+(x) -| (nor.in 1) 
+(y) -| (nor.in 2);
+\end{circuitikz}
+\caption{A schematic that represents the reused component
+  \texttt{\small{(B (C B nand) S)}}.}
+\label{fig:component}
+\end{figure}
+
+%% \subsection{Digital arithmetic}
+%% Having demonstrated the behavior of the E-C algorithm on symbolic
+%% regression, we analyze its performance on tasks involving digital
+%% arithmetic, a class of problems with known modular structure. The
+%% general goal is to discover boolean circuits consisting of a limited
+%% set of gates that perform arithmetic operations on binary number
+%% representations.
+
+\subsection{Discussion}
+...
 

### 2013-01-25 src/PI/DigArith.hs

@@ -115,3 +115,14 @@ evalDigArith n c = [ toBool $ f vs | vs <- mkAllAss n]
                    (B True) -> True
                    (B False) -> False
 
+taskTruthTable :: Task -> TruthTable
+taskTruthTable tsk = [(tup, v) | v <- vals | tup <- tups]
+    where card = taskCard tsk
+          vals = taskBoolVals tsk
+          tups = mkAllAss card
+
+showTruthTable :: TruthTable -> String
+showTruthTable tt = unlines $ map showRow tt
+    where showRow (tup, v) = (unwords $ map show tup) ++ 
+                             " --> " ++ show v
+

### 2013-01-25 src/PI/EnumBF.hs

@@ -66,9 +66,10 @@ enumBF' :: Grammar -> Int -> BFState -> BFState
 -- resulting elements are added either to the open list or to the
 -- closed list. This process is repeated until there are at least i
 -- elements in closed list. 
+enumBF' gr i bfState@(BFState openPQ closedPQ) | PQ.size openPQ == 0 = bfState
 enumBF' gr i bfState@(BFState openPQ closedPQ) = 
     if PQ.size closedPQ >= i then bfState else
-        let (cbt, openPQ') = PQ.deleteFindMax openPQ
+        let (cbt, openPQ') = PQ.deleteFindMax openPQ -- get best open solution
             cbs =  map (\(c,i) -> CombBaseTuple c i) 
                   $ runStateT (expand gr (combBase cbt)) (tpVar cbt)
             closedCBs = filter (isClosed . combBase) cbs
@@ -107,7 +108,7 @@ expand gr cb@(CombBase (CTerminal tp) (Just []) v)
                      let value = (calcLogProb gr tp c)
                          combBase = CombBase c Nothing value
                      return combBase
-          cs_is_empty = False -- (length $  runStateT cs 0) == 0
+          cs_is_empty =  False -- (length $  runStateT cs 0) == 0
           cbApp = expandToApp gr cb 
 
 expand gr cb@(CombBase (CApp c_left c_right tp d) (Just (L:rest)) v) = 
@@ -148,9 +149,11 @@ expand gr cb@(CombBase (CApp c_left c_right tp d) (Just (R:rest)) v) =
 
 
 
-
-        
-
-
+doesCombHaveToType :: Comb -> Type -> Bool
+doesCombHaveToType c t = runStateT m tyInt
+    where (TyVar id _) = makeNewTVar [t]
+          tyInt = readId id
+          m = extendTypeOnLeftN 
+          
 
 

### 2013-01-25 src/PI/Experiments/DigitalArithmetic.hs

@@ -8,6 +8,7 @@ import Data.Maybe
 import qualified Data.HashMap as HMap
 import Control.Monad.Random
 import Debug.Trace
+import Data.GraphViz
 
 import Experiment
 import Type
@@ -28,6 +29,7 @@ import EnumBF
 import ParseCL hiding (eval)
 import Run
 import DigArith
+import Visualize
 
 
 taskIncr1 = Task "pair incr 1"
@@ -37,7 +39,7 @@ taskIncr1 = Task "pair incr 1"
                   pairIn0 = intTupleToPair (0, 0)
                   pairOut0 = intTupleToPair (0, 1)
                   pairIn1 = intTupleToPair (0, 1)
-                  pairOut1 = intTupleToPair (0, 0)
+                  pairOut1 = intTupleToPair (1, 0)
                   pairIn2 = intTupleToPair (1, 0)
                   pairOut2 = intTupleToPair (1, 1)
                   points = [(pairIn0, pairOut0),
@@ -85,7 +87,7 @@ taskTrue = Task "True" f tBool
 lib = (CM.fromList $ 
          [
           ("I", cI)
- --        , ("True", cTrue)
+         , ("True", cTrue)
 --        , ("False", cFalse)
          ,("nand", cNand)
 --         , ("&", cAnd)
@@ -141,6 +143,36 @@ mkDigArithM name lib frontierSize
                   expDepthBound = 3,
                   expNumBound = frontierSize,
                   expReps = 15}
+
+
+
+-- Visualization
+fromRight (Right x) = x
+fromRight (Left err) = error err
+
+combstrs = ["((S nand) I)",
+           "B",
+            "(S nand)",
+            "S",
+            "nand",
+            "C",
+            "((S ((B (B nand)) nand)) ((S nand) I))",
+            "(B ((S nand) I))",
+            "(B (B (B ((S nand) I))))",
+            "(B (B ((S nand) I)))",
+            "I"]
+
+
+cs = map (fromRight . parseExpr lib1) combstrs
+combsgr = combsToGraph cs
+
+visualize lib filename = 
+    do combstr <- fmap (filter (not . null)) $ getLibFromFile filename
+       
+       let cs = take 1 $ map (fromRight . parseExpr lib ) combstr
+           gr = combsToGraph cs
+       runGraphviz (graphToDot params gr) Pdf "test.pdf"
+
 
 
 main = do

### 2013-01-25 src/PI/Grammar.hs

@@ -2,7 +2,7 @@
 
 module Grammar where
 
-import Data.List (intersect)
+import Data.List (intersect, sortBy)
 import Control.Monad.State
 import Debug.Trace
 import Data.Maybe
@@ -151,7 +151,13 @@ calcLogProb gr tp c
 
 
 
+truncateGrammar :: Grammar -> Int -> Grammar
+truncateGrammar (Grammar lib _) n = Grammar lib' 0
+    where xs = CM.toList lib
+          xs' = take n $ sortBy (\(_, y) (_, x) -> x `compare` y) xs
+          lib' = CM.fromList xs'
           
+    
                                 
                                        
                       

### 2013-01-25 src/PI/PostProcess.hs

@@ -7,6 +7,10 @@ import System.Directory
 import System.FilePath
 import Data.Time
 import qualified Data.Text as T
+import Data.List.Split
+import Data.List (isInfixOf)
+import Data.Either.Utils
+import Debug.Trace
 
 import Search
 import CL
@@ -16,6 +20,8 @@ import qualified CombMap as CM
 import CombMap (CombMap)
 import Grammar
 import Experiment
+import StdLib
+import ParseCL
 
 allCombs :: [SearchLogEntry] -> [Comb]
 allCombs ss = CM.keys $ foldl1 CM.union $ map (library . searchGrammar) ss 
@@ -99,5 +105,24 @@ saveSearchData datadir exp searchData time =
        hClose fexpdata
        hClose ftime
        
-       
+getLibFromFile filename = do
+  str <- readFile filename
+  let rows = splitOn "\n" str
+      combs = map head $ map (splitOn ",") rows
+  return combs
+
+getGrammarFromFile lib filename = do
+  str <- readFile filename
+  let rows = splitOn "\n" str
+      splitrows' = map (splitOn ",") rows
+      combstrs = map head $ splitrows'
+      valstrs = map last $ splitrows'
+      pairstrs = filter (not . (\x -> isInfixOf "nan" x || null x) . snd) $ zip combstrs valstrs
+      combs = (trace $ show pairstrs) $ map (fromRight .  parseExpr lib) $ map fst pairstrs
+      vals = map (read .  snd  ) pairstrs :: [Double]
+      lib' = (CM.fromList $ zip combs vals)
+      grammar = Grammar lib' 0
+  return grammar
+
+      
        

### 2013-01-25 src/PI/Type.hs

@@ -198,6 +198,22 @@ freshInst t = do let tvs = tv t
           inst (TAp t1 t2) lk= TAp (inst t1 lk) (inst t2 lk)
           inst t _ = t
 
+typeLeftDepth :: Type -> Int 
+-- | Returns the number of possible applications onto an expression of
+-- this type.
+typeLeftDepth (TAp t1 t2) = typeLeftDepth t2 + 1
+typeLeftDepth _ = 1
+
+extendTypeOnLeft :: Monad m => Type -> StateT Int m Type
+extendTypeOnLeft t = do tnew <- newTVar Star
+                        return (tnew ->- t)
+
+extendTypeOnLeftN :: Monad m => Type -> Int -> StateT Int m Type
+extendTypeOnLeftN t 0 = return $ t
+extendTypeOnLeftN t n = do t' <- extendTypeOnLeft t
+                           extendTypeOnLeftN t' (n-1)
+
+
 -- | types 
 tChar = TCon (TyCon "Char" Star)
 tInt = TCon (TyCon "Int" Star)

### 2013-01-26 drafts/ijcai2013/ijcai13.tex
Commits.

@@
 
-\title{Getting off the ground: bootstrap learning via modular concept discovery.}
+\title{Bootstrap learning via modular concept discovery}
 
@@
 \begin{abstract}
- Suppose a learner is faced with a domain of problems about which she
- knows nearly nothing. She does not know the distribution of problems,
- the space of solutions is not smooth, and the reward signal is
- uninformative, providing perhaps a few bits of information but not
- enough to steer her effectively. How can such a learner ever get off
- the ground? A common intuition is that if the solutions to these
- problems share a common structure, and the learner can solve some
- simple problems by luck or brute force, she should be able to extract
- useful modular concepts from these solutions and, by composing them,
- explore the solution space more efficiently. Here, we formalize this
- intuition, where the solution space is the space of typed functional
- programs and the gained information is stored as a stochastic grammar
- over programs. We propose a two step iterative procedure for
- exploring such spaces: in the first step, the learner explores a
- finite subset of the domain, guided by a stochastic grammar; in the
- second step, the learner compresses the successful solutions from the
- first step to estimate a new stochastic grammar. As we show in a
- symbolic regression and a boolean circuit learning example, iterating
- these two steps quickly propels the learner from a state of solving
- almost no problems to solving a large number of problems by gaining
- abstract knowledge of the structure of the solution space.
+ Suppose a learner is faced with a domain of problems about which it knows nearly nothing.
+It does not know the distribution of problems, the space of solutions is not smooth, and the reward signal is uninformative, providing perhaps a few bits of information but not enough to steer it effectively.
+How can such a learner ever get off the ground?
+A common intuition is that if the solutions to these problems share a common structure, and the learner can solve some simple problems by brute force, it should be able to extract useful components from these solutions and, by composing them, explore the solution space more efficiently.
+Here, we formalize this intuition, where the solution space is the space of typed functional programs and the gained information is stored as a stochastic grammar over programs.
+We propose an iterative procedure for exploring such spaces: in the first step of each iteration, the learner explores a finite subset of the domain, guided by a stochastic grammar; in the second step, the learner compresses the successful solutions from the first step to estimate a new stochastic grammar.
+We test this procedure on symbolic regression and boolean circuit learning and show that the learner discovers modular concepts for these domains.
+Whereas the learner is able to solve almost none of the posed problems in the procedure's first iteration, it rapidly becomes able to solve a large number by gaining abstract knowledge of the structure of the solution space.
+
 \end{abstract}
@@
     \end{tikzpicture}
+    \caption{AND/OR expression search space. }
     \label{fig:andor}
@@
 \subsection{A stochastic grammar over programs.}
+\label{sec:stochgrammar}
 A stochastic grammar over programs represents a distribution over
-expressions in $\mathcal{L}$. We represent our grammar as a
-multinomial distribution over the primitives in our library. Let $p_1,
-\dots, p_m$ be the probabilities associated with primitive combinators
-$c_1, \dots, c_m$. Then we define the probability $p(c_i | t) $ of
-using $c_i$ at a node whose type is $t$ to be $p_i$ normalized by the
-probability mass of all other primitive combinators that could have
-been used at that node. Namely, $p(c_i | t) = \frac{p_i}{\sum_{c_j :
-    type(c_j) \text{ unifies with } t} p_j}$ if type($c_i)$ unifies
-with $t$, 0 otherwise. Note that the distribution thus defined is not
-normalizable if we allow expressions of arbitrary depth, but this is
-not a practicle concern since we are using the distribution mainly as
-a search heuristic.
+expressions in $\mathcal{L}$. We associate with each primitive
+combinator $c_i$ in $\mathcal{L}$ a probability $p_i$,
+$0~\leq~p_i~\leq~1, \sum_i~p_i~=~1 $.
+
+We define the probability of an expression $e$ consisting of primitive
+combinators $c_1, \dots, c_N$ to be $p(e)~=~\prod_{n=1}^N p(c_n |
+t(c_n))$ where $t(c_n)$ is the requested type when $c_n$ was used and
+\[
+p(c_i | t(c_n)) =    \frac{1}{4}\frac{p_i}{\sum_{c_j :
+    type(c_j) \text{ unifies with } t(c_n)} p_j}.
+\]
+That is, $p(c_i | t(c_n)$ is the probability of sampling $c_i$ from
+the multinomial distribution defined by the $p_i$'s, conditioned on
+the requesting type, and then divided by 4. The constant factor of
+$1/4$ is introduced as a simple way to ensure that the sum over the
+probabilities of all expressions converges to finite value (so that
+the distribution over expressions is proper).\footnote{The probability
+  mass $M_d$ of all expressions of depth less than or equal to $d$ can
+  be written with the recurrence relation $M_d \leq M_d^2 + M_1$. This
+  has an asymptotic fixed point if $x = x^2 + M_1$ has a solution,
+  which is true if $M_1 \leq 1/4$.}
+
+To estimate the values $p_i$ from set of expressions, we simply divide
+the number of times primitive combinator $c_i$ was used by the number
+of times it could have been used given the requesting types in the
+expression trees. Note that this is not a maximum likelihood estimator
+for this distribution; finding that would require solving a rather
+complex optimization problem because of each choice is conditioned on
+the requesting type. We choose our estimator merely for convenience,
+and find that our empirical results justify this simplicity. It is a
+matter for future research to determine to what extent a more accurate
+estimator would improve performance.
 
@@
 procedure that enumerates the $N$ most probable programs. There has
-been recent interest in this
-problem~\cite{DBLP:conf/sfp/Katayama05}\cite{DBLP:conf/aaip/YakushevJ09}. Our
+been recent interest in this problem, most of it focused on
+enumerating the shortest program satisfying some
+criterion~\cite{DBLP:conf/sfp/Katayama05} \cite{DBLP:conf/aaip/YakushevJ09}. Our
 approach is to formulate the problem as a best-first exploration of an
@@
 
-In accordance with the stochastic grammar over programs, we weight all
-edges from OR nodes according to the negative log likelihood of
-choosing the associated rule from the stochastic grammar, conditioned
-on the requesting type. Expanding an AND node has zero cost. In the
-case that the stochastic grammar is ill conditioned, we might have an
-infinitely deep solution tree whose cost does not increase as we
-unfold it; this can occur, for example, if there are no primitives of
-the required type since the log probability of expansion to the next
-AND node is 0. To avoid this, we simply declare a cost for
-expanding the next AND node when it is the only node available.
+Recall that a valid partial solution is a subtree of $\mathcal{G}$
+satisfying the following properties: its root is the root
+$\mathcal{G}$ and any OR node has at most one child. A complete
+solution is a partial solution whose leaves are leaves of
+$\mathcal{G}$, whose root is the root of $\mathcal{G}$, in which each
+OR node has exactly one child, and in which each AND node has exactly
+two children.
+
+The value of a partial solution $G'$ is calculated by summing the
+value of its leaves. If a leaf of $G'$ is a leaf of $G$ then either it
+is an AND node, to which we assign a value of 0, or it corresponds to
+some primitive combinator $c_i$ in the library and is the child of
+some typed OR node with type $t$. We assign its value of $- \log{p(c_i
+  | t)}$, according to the stochastic grammar specified in
+Section~\ref{sec:stochgrammar}. If a leaf $v$ of $G'$ is an OR node,
+then the value of any extension of it to a complete solution is at
+most $-log(1/4)$. Thus, we have defined a value for partial solutions
+that upper bounds the value of any of its extensions.
+
+Our best-first policy, is to simply expand the next partial solution
+with the highest value. If a partial solution has multiple leaves that
+are not primitive combinators, we expand the left most one first,
+since it is the choice of left combinator that defines the type of its
+right combinator in an application pair. That is, the parent node
+applies its constraint on valid solutions via the left child, not the
+right child. We want to apply constraints as early as possible, so
+expanding left children first prunes more partial solutions.
+
+\textbf{****NEED COMPLEXITY ANALYSIS HERE*******}
 
@@
 
-Once we have enumerated the best $N$ programs from our distribution,
-we will find that some tasks have no solutions and some have one or
-more solutions; for those tasks that have solutions, we want to choose
-a solution for each task such that the complete set of such solutions
-for all the tasks is as compressible as possible.
+We say that a task is \emph{hit} if it has a solution in the
+frontier. Having enumerated the frontier, we want to assign solutions
+to hit tasks such that this set is maximally compressible. This will
+promote the usage of reusable modular subtrees.
 
 A natural metric for the compressibility of a set of binary trees is
-the number of unique subtrees that this set contains. Thus, we can
-attempt to find the assignment of solutions to tasks that minimizes
-this value. However, an exact optimization for this problem requires
+the number of unique subtrees that this set contains. As an example of
+this, suppose we had three tasks whose solutions were integers 16, 25,
+and 36, and suppose our frontier contained the six expressions:
+
+\begin{figure}[h]
+    \begin{tikzpicture}[scale=.9,
+        transform shape, level/.style={sibling distance=10mm/#1, font=\scriptsize}
+        ]
+      \node [font=\scriptsize] (top) {}
+      child {node [above=.3cm, text width=2cm] (a) {$(Int \rightarrow Int) \rightarrow Int 
+          \rightarrow Int$}
+        child {node [left=.5cm, below=1cm, text width=3cm] 
+          {$S : \\(t_2 \rightarrow t_1 \rightarrow t_0)  \rightarrow (t_2 \rightarrow t_1) 
+            \rightarrow \\ t_2 \rightarrow t_0$}}
+        child {node [right=0cm, text width=2cm] {$*: Int \rightarrow Int \rightarrow Int$}}
+      }
+      child {node [below right=.3cm and 1cm] (b) {$Int \rightarrow Int$}
+        child {node {$I: t_0 \rightarrow t_0 \rightarrow t_0$}}
+      };
+    \end{tikzpicture}
+\end{figure}
+
+Let $|e|$ be the
+number of unique binary trees in expression $e$.
+
+we will find that some tasks have no solutions
+and some have one or more solutions; for those tasks that have
+solutions, we want to choose a solution for each task such that the
+complete set of such solutions for all solved tasks is as compressible
+as possible.
+
+Thus, we can attempt
+to find the assignment of solutions to tasks that minimizes this
+value. However, an exact optimization for this problem requires
 examining $O(m^k)$ assignments, where $m$ is the maximum number of
@@
 %%(1,0) node[nand port, thick] (nor1) {}
-(3, 0)  node[ rotate=0, anchor=east,thick, draw, text width=1cm ] (f) {?} 
+(3, 0)  node[ anchor=east, text align=center, thick, draw, text width=1cm, text height=.25cm, align=center] (f) {?} 
+node [right of=f] (out) {}
 (1, 0) node[nand port, thick] (nor) {}
@@
 (nor.in 2) node[below left=.25cm] (y) {y}
-(y) -- (1.71,.5cm)
-(nor) -- (1.72,-.5cm)
+(y) -| (f)
+(nor) -- (f)
+(f) -| (out)
 (x) -| (nor.in 1) 

### 2013-01-26 src/PI/DigIncr.hs

@@ -1,0 +1,63 @@
+-- DigIncr.hs
+
+{-# Language ParallelListComp #-} 
+
+module DigIncr where
+
+import Control.Monad.State
+import Data.Maybe
+import qualified Data.HashMap as HMap
+import Control.Monad.Random
+import Debug.Trace
+import Data.GraphViz
+
+import Experiment
+import Type
+
+import Type
+import CL
+-- import Search
+import Expr
+import Task
+import StdLib
+import Data
+import qualified CombMap as CM
+import CombMap (CombMap)
+import Grammar
+import Routers 
+import PostProcess
+import EnumBF
+import ParseCL hiding (eval)
+import Run
+import Experiments.DigitalArithmetic hiding (main)
+import Visualize
+
+
+filename = "data/digarith_ec2_allbool2and3_overnight_wjittermany_2013-01-22_07-09-56.306334_UTC/grammars.csv"
+frontierSize = 1000
+
+augmentGrammar (Grammar lib _) (Grammar lib' _ ) = normalizeGrammar $ Grammar (lib `CM.union` lib') 0
+
+
+pairLib = CM.fromList $ [
+          ("pair", cPair)
+         , ("fst", cFst)
+         , ("snd", cSnd)]
+
+pairGr = grammarFromLib pairLib
+
+main = do
+  grFromFile <- getGrammarFromFile lib1 filename
+  let lib' = augmentGrammar grFromFile pairGr
+      lib'' = truncateGrammar lib' 10
+  putStr $ show lib''
+  let expDigIncr = Experiment { expName = "dig arith",
+                                expTaskSet = [taskIncr1],
+                                expEps = 0,
+                                expPrior = lib'',
+                                expInitLib = lib'',
+                                expDepthBound = 3,
+                                expNumBound = frontierSize,
+                                expReps = 1}
+  runExp expDigIncr
+

### 2013-01-26 src/PI/EnumBF.hs

@@ -48,7 +48,7 @@ isClosed cb | path cb == Nothing  = True
             | otherwise           = False
 
 -- hack min grammar val
-hackMinGrammarVal = log (0.5)
+hackMinGrammarVal = log (0.25)
 
 
 enumBF :: Grammar 
@@ -105,7 +105,7 @@ expand gr cb@(CombBase (CTerminal tp) (Just []) v)
 --          minGrammarVal = minimum (CM.elems (library gr))
           minGrammarVal = hackMinGrammarVal
           cbs =   do c <- cs
-                     let value = (calcLogProb gr tp c)
+                     let value = (calcLogProb gr tp c) + minGrammarVal
                          combBase = CombBase c Nothing value
                      return combBase
           cs_is_empty =  False -- (length $  runStateT cs 0) == 0
@@ -149,11 +149,11 @@ expand gr cb@(CombBase (CApp c_left c_right tp d) (Just (R:rest)) v) =
 
 
 
-doesCombHaveToType :: Comb -> Type -> Bool
-doesCombHaveToType c t = runStateT m tyInt
-    where (TyVar id _) = makeNewTVar [t]
-          tyInt = readId id
-          m = extendTypeOnLeftN 
+-- doesCombHaveToType :: Comb -> Type -> Bool
+-- doesCombHaveToType c t = runStateT m tyInt
+--     where (TyVar id _) = makeNewTVar [t]
+--           tyInt = readId id
+--           m = extendTypeOnLeftN t
           
 
 

### 2013-01-26 src/PI/python/analyzeDA.py

@@ -1,0 +1,49 @@
+
+
+from os.path import join, exists, dirname
+import pandas as pd
+import matplotlib.pyplot as plt
+import numpy as np
+
+import launchEC2DigArith as DA
+
+grammarFiles = [join(dirname(f), "grammars.csv") for f in DA.files]
+badGrammarFiles = [grammarFiles[i] for i in [0,2,3,4,5,7]]
+goodGrammarFiles = [grammarFiles[i] for i in [1,6,8,9]]
+
+def getGrammar(file):
+    fid = open(file, 'r')
+    rows = []
+    combs = []
+    for row in fid.readlines():
+        combs.append(row.split(",")[0])
+        strvals = row.split(",")[1:]
+        vals = []
+        for sv in strvals:
+            if sv.strip() == 'nan':
+                vals.append(np.nan)
+            else:
+                vals.append(float(sv))
+        rows.append(vals)
+        grammar = np.array(rows)
+    return combs, grammar
+
+
+    
+# grammarFiles = badGrammarFiles + goodGrammarFiles
+# grsColumns = [[getGrammar(f)[1][:, i] for f in grammarFiles] for i in range(15)]
+# # gs =  [[len([v for v in g.tolist() if not np.isnan(v)]) for g in grs] for grs in grsColumns]
+# initLogProbs = [[v for v in g.tolist() if not np.isnan(v)] for g in grsColumns[0]]
+
+
+df = pd.DataFrame([DA.loadFile(f) for f in DA.files]).T/(272.0)
+means_df = df.mean(1)
+std_df = df.std(1)
+
+w, h = plt.figaspect(1.5)
+fig = plt.figure(figsize=(w,h))
+ax = fig.add_subplot(111)
+ax.set_xlabel("iteration")
+ax.set_ylabel("% tasks solved")
+df.plot(linewidth=2, legend=False, ax = ax)
+fig.savefig("/Users/edechter/Dropbox/Projects/ProgramInduction/drafts/ijcai2013/figures/boolLearningCurves.pdf")

### 2013-01-26 src/PI/python/symregmany.py

@@ -1,0 +1,87 @@
+
+from glob import glob
+import pandas as pd
+import numpy as np
+import matplotlib.pyplot as plt
+
+import symreg as SR
+
+files = glob("../data/*wjittermany*21*/numhit.csv")
+
+tuples = [(i, []) for i in range(1000,11000, 1000)]
+
+for i, (n, xs) in enumerate(tuples):
+    for f in files:
+        if f.find(str(n)+"_") >= 0:
+            tuples[i][1].append(f)
+
+def loadFile(file):
+    x = open(file, 'r').read()
+    return [float(i) for i in x.split(",")]
+
+
+vals = {}
+for (n, vs) in tuples:
+    ys = [loadFile(f) for f in vs]
+    if len(ys)>0:
+        vals[n] = ys
+
+def padToArray(length, lists):
+    d = []
+    for list in lists:
+        d.append(list + [np.nan for i in range(0, length - len(list))])
+    return np.array(d)
+
+length = 15
+padded = {}
+means = {}
+stds = {}
+for (k, v) in vals.iteritems():
+    padded[k] = pd.DataFrame(padToArray(length, v)).T/1331.0
+    means[k] = padded[k].mean(1)
+    stds[k] = padded[k].std(1)
+
+means_df = pd.DataFrame(means)
+stds_df = pd.DataFrame(stds)
+
+y = means_df.xs(14)
+yerr = stds_df.xs(14)
+x = means_df.columns
+
+w, h = plt.figaspect(1.5)
+fig = plt.figure(figsize=(w,h))
+ax = fig.add_subplot(111)
+ax.set_xlabel("frontier size")
+ax.set_ylabel("% tasks solved at 15 iterations")
+plt.errorbar(x, y, yerr, linewidth=2)
+labels = ax.get_xticklabels() 
+for label in labels: 
+    label.set_rotation(80) 
+
+fig.savefig("/Users/edechter/Dropbox/Projects/ProgramInduction/drafts/ijcai2013/figures/performanceVsFrontierSize2.pdf")
+
+
+fig_lc = plt.figure(figsize=(w,h))
+ax = fig_lc.add_subplot(111)
+ax.set_xlabel("iteration")
+ax.set_ylabel("% tasks solved")
+ax.set_ylim([0,1])
+
+means_df.plot(linewidth=2, ax=ax, legend=False)
+
+plt.legend(loc=2, prop={'size':8})
+
+
[…]

### 2013-01-26 src/PI/python/timing.py

@@ -1,0 +1,24 @@
+
+
+from os.path import join, exists
+
+import pandas as pd
+import matplotlib as plt
+
+datadir = "../data"
+D = { 1000 : "R1_1000_wjittertiming_2013-01-20_12-45-31.810308_EST/time.txt",
+         500 : "R1_500_wjittertiming_2013-01-20_12-43-23.900711_EST/time.txt",
+         600: "R1_600_wjittertiming_2013-01-20_12-43-41.697841_EST/time.txt",
+         700: "R1_700_wjittertiming_2013-01-20_12-44-01.5959_EST/time.txt",
+         800: "R1_800_wjittertiming_2013-01-20_12-44-25.596314_EST/time.txt",
+         900: "R1_900_wjittertiming_2013-01-20_12-44-56.474792_EST/time.txt"}
+
+
+d = {}
+for k, v in D.iteritems():
+    filename = join(datadir, v)
+    assert(exists(filename))
+    n = float(open(filename, 'r').read()[:-1])
+    d[k] = n
+
+

### 2013-01-28 src/PI/BooleanCircuits.hs
Changes.

@@ -252,8 +252,8 @@ sampleCircuitTasks dI dN dG nT
           
 
 -- DISTRIBUTIONS -- 
-dNumInputs = [(1, 1), (2, 3), (3, 4)] :: [(Int, Rational)]
-dNumGates = [(1, 1), (2, 3), (3,  4), (4, 4), (5, 3)] :: [(Int, Rational)]
+dNumInputs = [(1, 1), (2, 2), (3, 4)] :: [(Int, Rational)]
+dNumGates = [(1, 1), (2, 3), (3,  4), (4, 5), (5, 4)] :: [(Int, Rational)]
 dGateType = [ (MGate notGate, 1), (BGate andGate, 2), (BGate orGate, 2)] :: [(Gate, Rational)]
 
 

### 2013-01-28 drafts/ijcai2013/ijcai13.tex
Changes.
Changes including adding EX3

@@
 \begin{abstract}
- Suppose a learner is faced with a domain of problems about which it knows nearly nothing.
-It does not know the distribution of problems, the space of solutions is not smooth, and the reward signal is uninformative, providing perhaps a few bits of information but not enough to steer it effectively.
-How can such a learner ever get off the ground?
-A common intuition is that if the solutions to these problems share a common structure, and the learner can solve some simple problems by brute force, it should be able to extract useful components from these solutions and, by composing them, explore the solution space more efficiently.
-Here, we formalize this intuition, where the solution space is the space of typed functional programs and the gained information is stored as a stochastic grammar over programs.
-We propose an iterative procedure for exploring such spaces: in the first step of each iteration, the learner explores a finite subset of the domain, guided by a stochastic grammar; in the second step, the learner compresses the successful solutions from the first step to estimate a new stochastic grammar.
-We test this procedure on symbolic regression and boolean circuit learning and show that the learner discovers modular concepts for these domains.
-Whereas the learner is able to solve almost none of the posed problems in the procedure's first iteration, it rapidly becomes able to solve a large number by gaining abstract knowledge of the structure of the solution space.
+ Suppose a learner is faced with a domain of problems about which it
+ knows nearly nothing. It does not know the distribution of problems,
+ the space of solutions is not smooth, and the reward signal is
+ uninformative, providing perhaps a few bits of information but not
+ enough to steer it effectively. How can such a learner ever get off
+ the ground? A common intuition is that if the solutions to these
+ problems share a common structure, and the learner can solve some
+ simple problems by brute force, it should be able to extract useful
+ components from these solutions and, by composing them, explore the
+ solution space more efficiently. Here, we formalize this intuition,
+ where the solution space is the space of typed functional programs
+ and the gained information is stored as a stochastic grammar over
+ programs. We propose an iterative procedure for exploring such
+ spaces: in the first step of each iteration, the learner explores a
+ finite subset of the domain, guided by a stochastic grammar; in the
+ second step, the learner compresses the successful solutions from the
+ first step to estimate a new stochastic grammar. We test this
+ procedure on symbolic regression and boolean circuit learning and
+ show that the learner discovers modular concepts for these
+ domains. Whereas the learner is able to solve almost none of the
+ posed problems in the procedure's first iteration, it rapidly becomes
+ able to solve a large number by gaining abstract knowledge of the
+ structure of the solution space.
 
@@
 somewhat smooth.
+
+Let's assume there is a distribution over tasks in our domain
+$p_{task}(\dot)$ and that these tasks are drawn $i.i.d$ from this
+distribution. We can frame our goal as finding an ordering on programs
+$w : \mathcal{L} \rightarrow (1, \dots)$ such that if we search for programs
+according to $w$ then the expected number of programs we need to
+inspect is minimized. But this ordering is exactly that which is
+defined by $p_{task}$. Namely, let $e_1, e_2 \in \mathcal{L}$ solve
+tasks $t1, t2$ respectively. Then we want $w(e_1) > w(e_2) $ iff $
+p_{task}(t_1) > p_{task}(t_2)$. One way to achieve this is that for every
+$e \in \mathcal{L}$ that solves task $t$, we define $w$ to be ordering
+according to the distribution $p_\mathcal{L}(e) \propto p_{task}(t)$.
 
@@
 \begin{figure}
-\begin{subfigure}[Before]{0.45\linewidth}
-\includegraphics[width=\linewidth]{figures/learningCurves2.pdf}
+\begin{subfigure}[Before]{.6\linewidth}
+%% \includegraphics[width=\linewidth]{figures/learningCurves2.pdf}
+\includegraphics[width=\linewidth]{figures/symreg3frontiers.pdf}
 \end{subfigure}
-\begin{subfigure}[After]{0.45\linewidth}
-\includegraphics[width=\linewidth]{figures/performanceVsFrontierSize2.pdf}
-\end{subfigure}
+%% \begin{subfigure}[After]{0.45\linewidth}
+%% \includegraphics[width=\linewidth]{figures/performanceVsFrontierSize2.pdf}
+%% \end{subfigure}
 \caption{(a) Average learning curves as we vary the frontier size. (b)

### 2013-01-28 src/PI/BooleanCircuits.hs

@@ -189,47 +189,33 @@ isCircuitConnected circ =  (Set.size diff == 1) && (diff == Set.singleton outId)
           diff = Set.difference  allGateIds allInputs
           outId = gateId $ head $ circuitGates circ
 
+sampleConnectedCircuitGivenNumbers :: MonadRandom m 
+                             => Int -- ^ number of inputs
+                          -> Int -- ^ number of gates
+                          -> [(Gate, Rational)]
+                          -> m Circuit
+sampleConnectedCircuitGivenNumbers nI nN dG 
+    = do gates <- sampleGatesFromDistribution dG nN
+         let circ = createInputs nI
+         circ' <- addInstances circ gates
+         if isCircuitConnected circ' then
+             return circ' else
+             sampleConnectedCircuitGivenNumbers nI nN dG
+                             
 
-sampleConnectedCircuitsNGates :: MonadRandom m 
-                                => Int -- ^ number of inputs
-                              -> Int  -- ^ number of gates
-                              -> [(Gate, Rational)]
-                              -> m Circuit -- ^ stream of connected circuits
-sampleConnectedCircuitsNGates nI nG distr  
-    = do circ <- sampleCircuitNGates nI nG distr
-         if isCircuitConnected circ then 
-             return circ else
-             sampleConnectedCircuitsNGates nI nG distr
-                                           
-
-sampleNConnectedCircuits :: MonadRandom m => 
-                            Int -- ^ number of inputs
-                         -> Int -- ^ number of gates
-                         -> [(Gate, Rational)]
-                         -> Int -- ^ number of circuits
-                         -> m [Circuit]
-sampleNConnectedCircuits nI nG distr nC = sequence $ replicate nC$
-                                          sampleConnectedCircuitsNGates nI nG distr
-
-sampleCircuit :: MonadRandom m =>
-                  [(Int, Rational)] -- ^ distribution over number of inputs
-               -> [(Int, Rational)] -- ^ distribution over number of gates
-               -> [(Gate, Rational)] -- ^ distribution over types of gates
-               -> m Circuit
-sampleCircuit dI dN dG = do nI <- fromList dI 
-                            nN <- fromList dN
-                            gates <- sampleGatesFromDistribution dG nN
-                            let circ = createInputs nI
[…]

### 2013-01-29 drafts/ijcai2013/ijcai13.tex
Changes.

--- 2013-01-26 drafts/ijcai2013/ijcai13.tex
@@
 
-Formally then, our goal is to solve the following minimal feedback
-multi-task program induction problem: suppose we are
-given set of \emph{tasks} $T=\{t_k\}_{k=1}^K$ where each task is a
-function $t_i : \mathcal{L} \rightarrow \{0, 1\}$, where $\mathcal{L}$
-is the set of programs in our language.  We say that $ p \in
-\mathcal{L}$ \emph{solves} $t_i$ if $ t_i\; p = 1$. Our
-goal is to find a set of solutions $X = \{x_k\}_{k=1}^k \subset
-\mathcal{L}$ such that
+Let a \emph{task} be a function $t : \mathcal{L} \rightarrow \{0,
+1\}$, where $\mathcal{L}$ is the set of all expressions in our
+representation language. We say that $e \in \mathcal{L}$ \emph{solves}
+task $t$ if $t(e) = 1$. Our goal is to solve as many tasks as
+possible, as efficiently as possible. That is, suppose that there is a
+distribution over tasks $t \sim  p_t(\cdot)$, we would like an ordering $O
+= e_1, \dots $ on expressions in $\mathcal{L}$ such that if $p_t(t_1)
+> p_t(t_2)$ then the first expression in $O$ to solve $t_1$ comes
+before the first expression in $O$ to solve $t_2$.
 
-\[
-X = \argmax_{x_1, \dots, x_K \in L} \sum_k t_k(x_k).
-\]
+We motivate a solution to this problem by considering a heirarchical
+Bayesian formulation of the task distribution. Suppose we believe that
+there is some underlying structure to $p_t(\cdot)$ that is explained by
+the fact that for each task $t$ there is a latent expressions $e \in
+\mathcal{L}$ drawn from a stochastic grammar $\mathcal{G}$ and that
 
-Based on the intuition set out above, we present an iterative
+\begin{align}
+e &\sim p_\mathcal{L}(\cdot) = \mathcal{G}\\
+t | e &\sim \text{U}\{ \text{tasks solved by }e\}.
+\end{align}
+
+Then $p_t(t) = \sum_{e_j \in \mathcal{L} \text{ s.t. } t(e_j) = 1}
+p_\mathcal{L}(e_j)$. If we make the assumption that for the set $E$ of
+expressions in this sum, there is some expression $e^*$ such that
+$p_\mathcal{L}(e^*) >> p_\mathcal{L}(e), e \neq e^*$, then we have, approximately, 
+\[ p_t(t) \propto} p_\mathcal{L}(e^*). \]
+
+This suggests a way to find the ordering $O$: infer a stochastic
+grammar $G$, under the assumption that $G$ favors one solution to a
+given task over any other, and order the elements of $\mathcal{L}$
+according to $G$.
+
+Thus, our basic procedure will be to search for a single ``good''
+solution for as many tasks as we can, and to use those solutions to
+estimate a stochastic grammar. Our search will be performed in the
+order of the most recently inferred grammar. So as our ordering
+becomes more efficient and we can solve more tasks, we learn better
+stochastic grammars.
+
+%% Formally then, our goal is to solve the following minimal feedback
+%% multi-task program induction problem: suppose we are given set of
+%% \emph{tasks} $T=\{t_k\}_{k=1}^K$ where each task is a function $t_i :
+%% \mathcal{L} \rightarrow \{0, 1\}$, where $\mathcal{L}$ is the set of
+%% programs in our language.  We say that $ p \in \mathcal{L}$
+%% \emph{solves} $t_i$ if $ t_i\; p = 1$. Our goal is to find a set of
+%% solutions $X = \{x_k\}_{k=1}^k \subset \mathcal{L}$ such that
+
+%% \[
+%% X = \argmax_{x_1, \dots, x_K \in L} \sum_k t_k(x_k).
+%% \]
+
+Based on the motivation set out above, we present an iterative
 strategy. Given a stochastic grammar $\mathcal{G}$ which defines a
@@
 the first step, we explore the $N$ highest probability programs
-according $\mathcal{G}$. In the second step, we compress the
+according to $\mathcal{G}$. In the second step, we compress the
 programs that solve any of our tasks, and we use that compression to
@@
 
-\subsection{Representing typed functional programs as binary trees.}
+\subsection{Representing typed functional programs as binary trees}
 
@@
 The basis combinators can themselves been expressed in the lambda
-calculus. The lambda calculus has two basic operation -- application
+calculus. The lambda calculus has two basic operations -- application
 and abstraction -- but in using the combinatory logic we sequester
@@
 We can extend this representation with a simple polymorphic type
-system~\cite{Pierce_2002}. Thus, this combinator can be represented as
-a binary tree~(Figure~\ref{fig:clbintree}) whose leaves are typed
-primitive combinators and whose interior nodes represent typed applications
-of combinators.
+system~\cite{Pierce_2002}. In this representation, a type is either a
+type primitive (e.g. reals, integers, booleans, etc.), a type variable
+(e.g. $\sigma$), or a function type $ t_1 \rightarrow t_2$ of
+functions from source type $t_1$ to target type $t_2$. Thus, this
+combinator can be represented as a binary
+tree~(Figure~\ref{fig:clbintree}) whose leaves are typed primitive
+combinators and whose interior nodes represent typed applications of
+combinators.
 
@@
 \subsection{Symbolic regression.}
+\label{sec:symreg}
 
@@
 \subsection{Boolean circuits}
+\label{sec:boolcircuits}
 
-As another demonstration of this approaches learning abilities
-investigate its ability to learn boolean functions using only the
-logical primitive NAND and the basic combinators. In
-Figure~\ref{fig:boolLearningCurves},
+As another demonstration of this approach, we investigate the
+procedure's ability to learn boolean functions using only the logical
+primitive NAND and the basic combinators. It is well known that a
+boolean circuit can be constructed for any boolean function given only
+the NAND gate. Here, the basic combinators take the place of the
+wiring normally found in a circuit, the output of element to multiple
+inputs downstream.
+
+We constructed two task sets. The first of these was constructed
+explicitly to contain familiar module structure. In this set of tasks,
+we sampled $1000$ boolean circuits using AND, OR, and NOT gates. To
+accomplish this, we first sampled either 1, 2, 3 inputs, then between
+1 and 5 gates, randomly wiring the inputs of each new gate to the
+output of one of the existing gates in the circuit. We continued this
+sampling procedure until we had 1000 ``connected'' circuits, i.e.,
+circuits all of whose output are wired to an input except for the last
+one (the output gate). This resulted in $1000$ tasks consisting of
+$82$ unique boolean functions, with a distribution of tasks as shown
+in Figure~\ref{fig:boolhist}. In Figure~\ref{fig:boolLearningCurves1},
+We present learning curves for frontier sizes 100, 500 and 10000,
+using the same procedure as in Section~\ref{sec:symreg}.
+
+This problem is particularly suitable for inspecting the constituents
+of the induced grammar $\mathcal{G}$. We might hope that our algorithm
+recovers the consituent logic gates that were used to build up the
+problems. Table~\ref{table:bool} shows a few of the top ranked
+expressions in the library. These include the basic logic gates; we
+also include one higher-order expression $F$, to stress that the
+representation we are using allows one to build concepts that are more
+expressive that just sub-circuits.
+
+In a second experiment, we include all $272$ boolean truth tables with
+three or fewer inputs as tasks. This is a more difficult problem. Note
+that in the first boolean function problem, there are two kinds of
+learning that can take place: first, many expressions map to a single
+boolean function, so we need to learn primitives that allow us to span
+many different functions instead of allows generating very simple
+functions over and over again. The second kind of learning involves
+the distribution of the functions themselves. In the first boolean
+experiment, that structure is apparent in the distribution in
+Figure~\ref{fig:boolhist}. In this second experiment, we remove the
+second kind of structure. In Figure~\ref{boolLearningCurves}, we show
+10 runs of the algorithm with a frontier size of $2000$: note how
+there are several local minima that majority of the runs get stuck in
+with performance around $50\%$, but several of the runs seem to take
+different trajectories to more successful representations.
 
@@
 \begin{subfigure}[Before]{0.45\linewidth}
-\includegraphics[width=\linewidth]{figures/boolLearningCurves.pdf}
+\includegraphics[width=.9\linewidth]{figures/booleancircuitsBars.pdf}
+\end{subfigure}
+\begin{subfigure}[Before]{0.45\linewidth}
+\includegraphics[width=.9\linewidth]{figures/boolLearningCurves.pdf}
 \end{subfigure}
@@
 \label{fig:boolLearningCurves}
+
 \end{figure}
@@
 
-\begin{figure}
-\begin{circuitikz} \draw
-%%(1,0) node[nand port, thick] (nor1) {}
-(1,.75) node[nand port, thick] (nor1) {}
-(3,0) node[nand port, thick] (nor2) {}
-(3,1.5) node[nand port, thick] (nor3) {}
-(5,.75) node[nand port, thick] (nor4) {}
-
-(nor1.in 1) node[above left=.5cm] (x) {x}
-(nor1.in 2) node[below left = .5cm](y) {y}
-(x) -| (nor1.in 1)
-(y) -| (nor1.in 2)
-(x) -| (nor3.in 1)
-(y) -| (nor2.in 2)
-(nor1) -- (1.62, .75) -- (nor2.in 1)
-(nor1) -- (1.62, .75) -- (nor3.in 2)
-(nor2) -| (nor4.in 2)
-(nor3) -| (nor4.in 1);
- \end{circuitikz}
-\caption{A standard XOR circuit built of NAND gates.}
-\label{fig:xor_circuit}
-\end{figure}
-
-\begin{figure}
-\begin{circuitikz} \draw
-
-%%(1,0) node[nand port, thick] (nor1) {}
-(3, 0)  node[ anchor=east, text align=center, thick, draw, text width=1cm, text height=.25cm, align=center] (f) {?} 
-node [right of=f] (out) {}
-(1, 0) node[nand port, thick] (nor) {}
-(nor.in 1) node[above left=.25cm] (x) {x}
-(nor.in 2) node[below left=.25cm] (y) {y}
-(y) -| (f)
-(nor) -- (f)
-(f) -| (out)
-(x) -| (nor.in 1) 
-(y) -| (nor.in 2);
-\end{circuitikz}
-\caption{A schematic that represents the reused component
-  \texttt{\small{(B (C B nand) S)}}.}
-\label{fig:component}
-\end{figure}
+\begin{table}
+\scriptsize
+  \begin{tabular}{|l|l|l|}
+    function  & CL expressions & circuit  ~ \\ \hline
+        TRUE  & (S NAND) (S NAND I)                   & ~ \\ \hline
+        FALSE & (S B (S NAND)) (S NAND I)             & ~ \\ \hline
+        NOT   & S NAND I                              & \input{figures/NOT.tex} \\ \hline
+        AND   & ((C B NAND) (B (S NAND I)))           & ~ \\ 
+        OR    & ((B (C (B (S NAND) NAND))) (S NAND I) & ~ \\ 
+        XOR   & ~                                     & ~ \\ 
+        F     & S B (S NAND)                          & \input{figures/F.tex}
+    \end{tabular}
+\caption{Expressions in the induced library on sets of tasks
+  corresponding to boolean circuits synthesized from AND, OR, and NOT
+  gates, with boolean circuits of 1, 2, or 3 inputs.}
+\label{table:boolexpr}
+\end{table}
 
@@
 \subsection{Discussion}
-...
 

### 2013-01-29 src/PI/BooleanCircuits.hs

@@ -206,7 +206,7 @@ sampleConnectedCircuitGivenNumbers nI nN dG
 sampleConnectedCircuit dI dN dG 
     = do nI <- fromList dI
          nN <- fromList dN
-         if nI > 2 && nN==1 then
+         if nN < nI then
              sampleConnectedCircuit dI dN dG else
               sampleConnectedCircuitGivenNumbers nI nN dG         
 
@@ -229,8 +229,7 @@ circuitToTask taskname circuit = do tt <- circuitToTruthTable circuit
 
 sampleCircuitTasks dI dN dG nT 
     = do circuits <- sampleConnectedCircuits dI dN dG nT 
-         return (map (\(i, c) -> (c,  fromJust $ circuitToTask ("circuitTask_" ++ show i) c))
-                    $ zip [0..] circuits)
+         return (map (\(i, c) -> (c,  fromJust $ circuitToTask ("circuitTask_" ++ show i) c)) $ zip [0..] circuits)
     where fromJust (Just x) = x
 
 
@@ -252,8 +251,8 @@ mkEquivClasses' (cl:cls) c = if (head cl) `equiv` c then
           
 
 -- DISTRIBUTIONS -- 
-dNumInputs = [(1, 1), (2, 2), (3, 4)] :: [(Int, Rational)]
-dNumGates = [(1, 1), (2, 3), (3,  4), (4, 5), (5, 4)] :: [(Int, Rational)]
+dNumInputs = [(1, 1), (2, 2), (3, 4), (4, 4)] :: [(Int, Rational)]
+dNumGates = [(1, 1), (2, 3), (3,  4), (4, 4), (5, 5), (6, 5)] :: [(Int, Rational)]
 dGateType = [ (MGate notGate, 1), (BGate andGate, 2), (BGate orGate, 2)] :: [(Gate, Rational)]
 
 

### 2013-01-29 src/PI/CL.hs

@@ -134,6 +134,8 @@ filterCombinatorsByType (c:cs) t
 filterCombinatorsByType [] t = lift []
 
 
+
+
                              
 
                                   

### 2013-01-29 src/PI/EnumBF.hs

@@ -4,6 +4,7 @@ module EnumBF where
 
 import Data.Function (on)
 import Control.Monad.State
+import Control.Monad.Identity
 import Data.Maybe
 import Data.PQueue.Max (MaxQueue)
 import qualified Data.PQueue.Max as PQ
@@ -98,7 +99,7 @@ expand :: Grammar
        -> CombBase 
        -> StateT Int [] CombBase
 expand gr cb@(CombBase (CTerminal tp) (Just []) v) 
-    = if cs_is_empty then mzero else 
+    = if cs_is_empty then (trace $ "EMPTY") $ mzero else 
           cbs `mplus` cbApp
     where primCombs = CM.keys (library gr)
           cs = filterCombinatorsByType primCombs tp
@@ -108,7 +109,8 @@ expand gr cb@(CombBase (CTerminal tp) (Just []) v)
                      let value = (calcLogProb gr tp c) + minGrammarVal
                          combBase = CombBase c Nothing value
                      return combBase
-          cs_is_empty =  False -- (length $  runStateT cs 0) == 0
+          cs_is_empty =  if not $ isTAp tp then False else
+                             not $ existCombWithToType primCombs (toType tp)
           cbApp = expandToApp gr cb 
 
 expand gr cb@(CombBase (CApp c_left c_right tp d) (Just (L:rest)) v) = 
@@ -149,11 +151,24 @@ expand gr cb@(CombBase (CApp c_left c_right tp d) (Just (R:rest)) v) =
 
 
 
--- doesCombHaveToType :: Comb -> Type -> Bool
--- doesCombHaveToType c t = runStateT m tyInt
---     where (TyVar id _) = makeNewTVar [t]
---           tyInt = readId id
---           m = extendTypeOnLeftN t
+doTypesUnify :: Type -> Type -> Bool
+doTypesUnify t1 t2 = case mgu t1 t2 of
+                       Just _ -> True
+                       Nothing -> False
+
+doesCombHaveToType :: Comb -> Type -> State Int Bool
+doesCombHaveToType c t = do ct <- freshInst (cType c)
+                            if doTypesUnify ct t then 
+                                return True else
+                                case ct of 
+                                  (TAp ct1 ct2) -> return $ doTypesUnify ct2 t
+                                  _ -> return False
+
+existCombWithToType :: [Comb] -> Type -> Bool
+existCombWithToType cs t = any (==True) $ fst $ runState m 0
+    where m = sequence $ do c <- cs
+                            return $ doesCombHaveToType c t
+
           
 
 

### 2013-01-29 src/PI/PostProcess.hs

@@ -2,15 +2,23 @@
 
 module PostProcess where
 
+
 import System.IO
 import System.Directory 
 import System.FilePath
 import Data.Time
-import qualified Data.Text as T
-import Data.List.Split
+import Data.List.Split hiding (oneOf)
 import Data.List (isInfixOf)
 import Data.Either.Utils
 import Debug.Trace
+
+import Text.ParserCombinators.Parsec 
+import Text.Parsec.Token 
+import Text.ParserCombinators.Parsec.Token
+import Text.ParserCombinators.Parsec ((<|>), (<?>))
+import Text.Parsec.Language (haskellDef)
+import Control.Applicative ((<$>), (<*>), (*>), (<*))
+
 
 import Search
 import CL
@@ -22,6 +30,7 @@ import Grammar
 import Experiment
 import StdLib
 import ParseCL
+
 
 allCombs :: [SearchLogEntry] -> [Comb]
 allCombs ss = CM.keys $ foldl1 CM.union $ map (library . searchGrammar) ss 
@@ -124,5 +133,50 @@ getGrammarFromFile lib filename = do
       grammar = Grammar lib' 0
   return grammar
 
-      
+getExplanationsFromFile lib filename 
+    = do  str <- readFile filename
+          let (Right explStrTuples) = parse explFile "" str
+          let combs = map (map (fromRight . (parseExpr lib))) $ map (map snd) explStrTuples
+          return combs
+
+
+explFile :: GenParser Char st [[(String, String)]]
+explFile = 
+    do result <- many line
+       eof
+       return result
+
+line :: GenParser Char st [(String, String)]
+line = 
+    do oneOf "["
+       result <- cells
+       oneOf "]"
+       eol                       -- end of line
+       return result
+
+cells :: GenParser Char st [(String, String)]
+cells = 
+    do first <- cellContent
+       next <- remainingCells
+       return (first : next)
+
+cellContent :: GenParser Char st (String, String)
+cellContent = do a <- oneOf "(" *> many (noneOf ",")
+                 oneOf ","
+                 b <- many (noneOf ",[]") -- <* oneOf ")"
+                 return (a, init b)
+
+remainingCells = (char ',' >> cells) <|> (return [])
+
+eol :: GenParser Char st String
+eol = string "\n \n" <|> string ""
+                      
+
+
+
+-- getExplanationsFromFile lib filename 
+--     = do str <- readFile filename
+--          rows <- splitOn "\n" str
+--          return ()
+  
        

### 2013-01-29 src/PI/Type.hs

@@ -77,7 +77,9 @@ toType (TAp (TAp tArrow a) b) = b
 mkTVar :: Int -> Type 
 mkTVar i = TVar (TyVar (enumId i) Star)
 
-
+isTAp :: Type -> Bool
+isTAp (TAp t1 t2) = True
+isTAp _ = False
 
 isTVar :: Type -> Bool
 isTVar (TVar _) = True

### 2013-01-29 drafts/ijcai2013/ijcai13.tex
Chnages.

@@
 
-Let a \emph{task} be a function $t : \mathcal{L} \rightarrow \{0,
-1\}$, where $\mathcal{L}$ is the set of all expressions in our
-representation language. We say that $e \in \mathcal{L}$ \emph{solves}
-task $t$ if $t(e) = 1$. Our goal is to solve as many tasks as
-possible, as efficiently as possible. That is, suppose that there is a
-distribution over tasks $t \sim  p_t(\cdot)$, we would like an ordering $O
-= e_1, \dots $ on expressions in $\mathcal{L}$ such that if $p_t(t_1)
-> p_t(t_2)$ then the first expression in $O$ to solve $t_1$ comes
-before the first expression in $O$ to solve $t_2$.
+%% Let a \emph{task} be a function $t : \mathcal{L} \rightarrow \{0,
+%% 1\}$, where $\mathcal{L}$ is the set of all expressions in our
+%% representation language. We say that $e \in \mathcal{L}$ \emph{solves}
+%% task $t$ if $t(e) = 1$. Our goal is to solve as many tasks as
+%% possible, as efficiently as possible. That is, suppose that there is a
+%% distribution over tasks $t \sim  p_t(\cdot)$, we would like an ordering $O
+%% = e_1, \dots $ on expressions in $\mathcal{L}$ such that if $p_t(t_1)
+%% > p_t(t_2)$ then the first expression in $O$ to solve $t_1$ comes
+%% before the first expression in $O$ to solve $t_2$.
 
-We motivate a solution to this problem by considering a heirarchical
-Bayesian formulation of the task distribution. Suppose we believe that
-there is some underlying structure to $p_t(\cdot)$ that is explained by
-the fact that for each task $t$ there is a latent expressions $e \in
-\mathcal{L}$ drawn from a stochastic grammar $\mathcal{G}$ and that
+%% We motivate a solution to this problem by considering a heirarchical
+%% Bayesian formulation of the task distribution. Suppose we believe that
+%% there is some underlying structure to $p_t(\cdot)$ that is explained by
+%% the fact that for each task $t$ there is a latent expressions $e \in
+%% \mathcal{L}$ drawn from a stochastic grammar $\mathcal{G}$ and that
 
-\begin{align}
-e &\sim p_\mathcal{L}(\cdot) = \mathcal{G}\\
-t | e &\sim \text{U}\{ \text{tasks solved by }e\}.
-\end{align}
+%% \begin{align}
+%% e &\sim p_\mathcal{L}(\cdot) = \mathcal{G}\\
+%% t | e &\sim \text{U}\{ \text{tasks solved by }e\}.
+%% \end{align}
 
-Then $p_t(t) = \sum_{e_j \in \mathcal{L} \text{ s.t. } t(e_j) = 1}
-p_\mathcal{L}(e_j)$. If we make the assumption that for the set $E$ of
-expressions in this sum, there is some expression $e^*$ such that
-$p_\mathcal{L}(e^*) >> p_\mathcal{L}(e), e \neq e^*$, then we have, approximately, 
-\[ p_t(t) \propto} p_\mathcal{L}(e^*). \]
+%% Then $p_t(t) = \sum_{e_j \in \mathcal{L} \text{ s.t. } t(e_j) = 1}
+%% p_\mathcal{L}(e_j)$. If we make the assumption that for the set $E$ of
+%% expressions in this sum, there is some expression $e^*$ such that
+%% $p_\mathcal{L}(e^*) >> p_\mathcal{L}(e), e \neq e^*$, then we have, approximately, 
+%% \[ p_t(t) \propto} p_\mathcal{L}(e^*). \]
 
-This suggests a way to find the ordering $O$: infer a stochastic
-grammar $G$, under the assumption that $G$ favors one solution to a
-given task over any other, and order the elements of $\mathcal{L}$
-according to $G$.
+%% This suggests a way to find the ordering $O$: infer a stochastic
+%% grammar $G$, under the assumption that $G$ favors one solution to a
+%% given task over any other, and order the elements of $\mathcal{L}$
+%% according to $G$.
 
-Thus, our basic procedure will be to search for a single ``good''
-solution for as many tasks as we can, and to use those solutions to
-estimate a stochastic grammar. Our search will be performed in the
-order of the most recently inferred grammar. So as our ordering
-becomes more efficient and we can solve more tasks, we learn better
-stochastic grammars.
+%% Thus, our basic procedure will be to search for a single ``good''
+%% solution for as many tasks as we can, and to use those solutions to
+%% estimate a stochastic grammar. Our search will be performed in the
+%% order of the most recently inferred grammar. So as our ordering
+%% becomes more efficient and we can solve more tasks, we learn better
+%% stochastic grammars.
 
@@
 distribution over the primitives from $\mathcal{L}$ and a integer
-\emph{frontier size} $N$, we perform to steps in each iteration.  In
+\emph{frontier size} $N$, we perform two steps in each iteration.  In
 the first step, we explore the $N$ highest probability programs
@@
 
-\begin{figure}[h]
-    \begin{tikzpicture}[scale=.9,
-        transform shape, level/.style={sibling distance=10mm/#1, font=\scriptsize}
-        ]
-      \node [font=\scriptsize] (top) {}
-      child {node [above=.3cm, text width=2cm] (a) {$(Int \rightarrow Int) \rightarrow Int 
-          \rightarrow Int$}
-        child {node [left=.5cm, below=1cm, text width=3cm] 
-          {$S : \\(t_2 \rightarrow t_1 \rightarrow t_0)  \rightarrow (t_2 \rightarrow t_1) 
-            \rightarrow \\ t_2 \rightarrow t_0$}}
-        child {node [right=0cm, text width=2cm] {$*: Int \rightarrow Int \rightarrow Int$}}
-      }
-      child {node [below right=.3cm and 1cm] (b) {$Int \rightarrow Int$}
-        child {node {$I: t_0 \rightarrow t_0 \rightarrow t_0$}}
-      };
-    \end{tikzpicture}
-\end{figure}
+%% \begin{figure}[h]
+%%     \begin{tikzpicture}[scale=.9,
+%%         transform shape, level/.style={sibling distance=10mm/#1, font=\scriptsize}
+%%         ]
+%%       \node [font=\scriptsize] (top) {}
+%%       child {node [above=.3cm, text width=2cm] (a) {$(Int \rightarrow Int) \rightarrow Int 
+%%           \rightarrow Int$}
+%%         child {node [left=.5cm, below=1cm, text width=3cm] 
+%%           {$S : \\(t_2 \rightarrow t_1 \rightarrow t_0)  \rightarrow (t_2 \rightarrow t_1) 
+%%             \rightarrow \\ t_2 \rightarrow t_0$}}
+%%         child {node [right=0cm, text width=2cm] {$*: Int \rightarrow Int \rightarrow Int$}}
+%%       }
+%%       child {node [below right=.3cm and 1cm] (b) {$Int \rightarrow Int$}
+%%         child {node {$I: t_0 \rightarrow t_0 \rightarrow t_0$}}
+%%       };
+%%     \end{tikzpicture}
+%% \end{figure}
 
@@
 $82$ unique boolean functions, with a distribution of tasks as shown
-in Figure~\ref{fig:boolhist}. In Figure~\ref{fig:boolLearningCurves1},
+in Figure~\ref{fig:booldistr}. In Figure~\ref{fig:boolLearningCurves1},
 We present learning curves for frontier sizes 100, 500 and 10000,
 using the same procedure as in Section~\ref{sec:symreg}.
+
+\begin{figure}
+\includegraphics[width=\linewidth]{figures/circuitDistrComp.pdf}
+\label{fig:booldistr}
+\caption{Distribution of boolean functions among (a) the set of 1000
+  circuits randomly constructed from AND, OR and NOT gates; b) the
+  first 1000 expressions from the grammar over programs \emph{before}
+  learning; c) the first 1000 expressions from the grammar over
+  programs \emph{after} learning.}
+\end{figure}
 
@@
 
-What kind of solutions does our algorithm find to common boolean
-circuits? We might hope that the circuits it finds are ``simple.'' One
-way to judge this is to compare the circuits it finds with the
-circuits found in digital electronics textbooks. For example, the XOR
-circuit can be drawn as a circuit with only NAND gates as in
-Figure~\ref{fig:xor_circuit}. Our algorithm returns the solution:
-
-\begin{alltt}
-\small{
-((S (B C (((B (C B nand)) S) 
-(((B (C B nand)) S) (B C 
-((B (B nand)) nand)))))) I)}. 
-\end{alltt}
-This looks more complicated. It should be understood, though, that the
-routing combinators S, B, and C play the role of wiring. Note that
-this solution has four NAND gates, the same number of as in the
-standard circuit. Another interesting feature of this solution is that
-it is not the smallest expression that is equivalent (modulo
-reduction), since we can further reduce it. However, this reduction
-would remove the two uses of \texttt{\small{(B (C B nand) S)}}. This is
-a reused component that appears within the most highly weighted 20
-combinators in the algorithm's induced grammar. We can view it
-schematically with a circuit diagram, but it is not a circuit
-fragment. Rather, it is a function that takes in a circuit of type
-$bool \rightarrow bool \rightarrow bool$ and returns a circuit of type $bool
-\rightarrow bool \rightarrow bool$. Figure~\ref{fig:component} shows
-this component schematically.
-
 \begin{table}
@@
 \subsection{Discussion}
+This work attempts to bring together ideas from various areas of
+research -- hierarchical Bayes, learning to learn, minimum description
+length learning -- to provide a proof of concept that learning
+abstractions from programs via compression can rapidly improve search
+performance. We emphasize this by focusing on domains in which the
+solution space is not smooth and the error signal is all or
+none. Thus, the main source of information for learning is the modular
+and compositional structure of the solution space.
+
+Of course, the representations that people use to learn in the world
+often have a partially smooth or local structure. Techniques in AI and
+machine learning commonly use smoothness or local structure to make
+search tractable. But the difference between this kind of learning and
+conceptual learning is undeniable. The former is what the statistician
+or machine learning researcher implements. The latter is what she
+learns as a student and as a practicioner. As researchers, we try
+different models, watch them fail or succeed, and abstract out the
+elements of these models that capture the relevant differences. We mix
+and match parts of different ideas, and utilize succcessful patterns
+of reasoning that can span many problems and domains.
+
+We believe that the primary characteristic of conceptual learning is
+that it occurs over a modular and compositional computational
+representation that supports higher-order functions. That is why we
+chose a functional programming language as the representational
+language of this work.
+
+The experiments in this work do not contain the richness of real world
+learning, but we believe that the learning algorithm's behavior
+exhibits some of the characteristics we would expect of a conceptual
+learner.
+
+One characteristic of this algorithm's behavior is a rapid initial
+increase in performance followed by a hard plateau. This plateau is
+expected given the circular nature of the algorithm: the learner
+inspects the region of the solution space that is typical of solutions
+already found. We see in this phenomenon a conceptual demarcation of
+the problem space: a coherent set of concepts or problems solving
+techniques can define by their power the extent of problems a problem
+solver is willing to consider. One might hope, however, that an
+intelligent learner would recognize that such a demarcation has
+occured and attempt to find another class of concepts for the
+remaining unsolved tasks.
+
+Another aspect of conceptual learning that we see present in this
+algorithm's behavior is the importance of a curriculum. By this we do
+not mean an explicit ordering of the problems in the task set, but
+rather that a set of sufficiently simple problems must act as stepping
+stones to the acquisition of complex representations.
 

### 2013-01-29 src/PI/BooleanCircuits.hs

@@ -14,6 +14,7 @@ import DigArith
 import Task
 
 
+
 data MonoGate = MonoGate {monoGateName :: String, 
                           monoGateFunc :: Bool -> Bool}
 
@@ -255,12 +256,6 @@ dNumInputs = [(1, 1), (2, 2), (3, 4), (4, 4)] :: [(Int, Rational)]
 dNumGates = [(1, 1), (2, 3), (3,  4), (4, 4), (5, 5), (6, 5)] :: [(Int, Rational)]
 dGateType = [ (MGate notGate, 1), (BGate andGate, 2), (BGate orGate, 2)] :: [(Gate, Rational)]
 
-
-
-
-
-                                               
-                                               
 
 -- not1 = MGateInst 2 notGate 1
 -- x = ConstInst 1 (Const True)

### 2013-01-29 src/PI/Experiments/EX5.hs

@@ -1,8 +1,9 @@
--- EX1
--- Symbolic regression 10 run per frontier size experiment on EC2. 
+-- EX5 
+-- symbolic regression with progressively more impoverished task sets
 {-# Language ParallelListComp #-}
 
-module Experiments.EX1 where
+
+module Experiments.EX5 where
 
 import System.Environment (getArgs)
 import Control.Monad.Random
@@ -54,14 +55,18 @@ mkSquaresTask a b = Task (showSquares a b) f (tInt ->- tInt)
                    
     where f = eval [num2C $ (squares a b) i| i<-rng]
 
--- taskSet' = --[mkSquaresTask i j | i <- [0..5] , j <- [0..5]]
---            -- ++  (map (mkSingleEqualityTask 100) $ [i*i | i <- [0..20]])
---            [mkQuadTask i j k | i <- [0..10] , j <- [0..10], k <- [0..10]]
---           ++ [mkQuadTask i j k | i <- [0..0] , j <- [0..5], k <- [0..5]]
+taskSetNoConstants = [mkQuadSymRegTask i j k | i <- [1..9] , j <- [1..9], k <- [0..9]]
+                     ++ [mkQuadSymRegTask i j k | i <- [1..9] , j <- [0], k <- [0..9]]
+                     ++ [mkQuadSymRegTask i j k | i <- [0] , j <- [1..9], k <- [0..9]]
 
-symRegTaskSet' = --[mkSquaresTask i j | i <- [0..5] , j <- [0..5]]
-           -- ++  (map (mkSingleEqualityTask 100) $ [i*i | i <- [0..20]])
-           [mkQuadSymRegTask i j k | i <- [0..9] , j <- [0..9], k <- [0..9]]
+taskSetNoLinear = [mkQuadSymRegTask i j k | i <- [1..9] , j <- [0..9], k <- [0..9]]
+                  ++ [mkQuadSymRegTask i j k | i <- [0] , j <- [1..9], k <- [0..9]]
+
+taskSetOnlyQuadratic = [mkQuadSymRegTask i j k | i <- [1..9] , j <- [0..9], k <- [0..9]]
+taskSetOnlyComplexQuadratic = [mkQuadSymRegTask i j k | i <- [1..9] , j <- [1..9], k <- [1..9]]
+taskSetUpToQuadratic = [mkQuadSymRegTask i j k | i <- [0..9] , j <- [0..9], k <- [0..9]]
+
+
 
[…]

### 2013-01-29 src/PI/Experiments/EX5.hs
Update.

@@ -147,7 +147,7 @@ main = do
                               4 -> ("OnlyComplexQuadratic", taskSetOnlyComplexQuadratic)
                               5 -> ("UpToQuadratic", taskSetUpToQuadratic)
                               _ -> error $ "No task name " ++ taskname
-  let frontierSize = 10
+  let frontierSize = 10000
       name = taskname  ++  "_frontier"  ++ show frontierSize
   exp <- mkSymRegExp name lib1 frontierSize taskset
   runExp exp "data/EX5"

### 2013-01-31 drafts/ijcai2013/ijcai13.tex
Update.
Changes.

@@
 \usepackage{tikz}
+\usepackage{tikz-qtree}
+\usepackage{float}
 \usepackage{circuitikz}
@@
 \usepackage{alltt}
+\usepackage{dashrule}
+\usepackage{array}
 \usepackage[ruled,vlined,linesnumbered]{algorithm2e}
-\usetikzlibrary{arrows,automata, positioning, patterns}
+\usetikzlibrary{arrows,automata, positioning, patterns,backgrounds}
 
@@
 \DeclareMathOperator*{\argmax}{arg\,max} 
+\DeclareMathOperator*{\argmin}{arg\,min}
 
@@
 \And
-Josh Tenenbaum \\
+Joshua B. Tenenbaum \\
 MIT\\
@@
 \section{Introduction}
-
-Imagine you are trying to learn to bake via trial and error. You know
-what kind of ingredients are used in baking (flour, water, butter,
-sugar, yeast, baking soda, etc.) and you know a bit about how to
-combine these ingredient (mix, whisk, etc.) and how to apply heat
-(oven). But baking is a notoriously fragile process: ratios of
-ingredients, the order of mixing, cooking temperatures, resting times,
-all make a huge difference, and the outcome of the baking process is
-mostly all or none. Either the outcome is what you expected or quite
-unlike it. Even if you do create something similar to your intent, it
-is difficult to assign blame for the outcome to any one step or
-decision along the way.
-
-Many difficult learning problems have this structure -- learning
-language, developing technologies and scientific theories, learning to
-play instruments or execute other complex motor patterns. These are
-problems in which 1) there is a large set of similar tasks; 2) their
-solutions are best described as programs; 3) the solution space is
-not smooth with respect to the outcome; 4) and the outcome is generally
-uninformative unless you are already pretty close to a solution.
-
-Intuitively, the way to solve such problems is to stumble on solutions
-to simple cases in the domain and extract conceptual building blocks
-that can be composed to scale up to complex cases. To return to the
-baking example, cookies are much simpler than cakes. There are fewer
-steps to get just right. But once you have figured out a few simple
-pastries, the more complex ones are mostly just variations and
-combinations of techniques learned along the way.
-
-Of course, knowledge is not acquired wholly by trial and error. But in
-many domains, there is at least initially such a period of ignorance,
-and the question we are trying to tackle here is how you ever get off
-the ground.
-
-To frame this learning paradigm computationally, we make two modeling
-choices. First, to emphasize the impoverished reward signals in these
-domains, we model the learning task as an all-or-none optimization
-problem: either the learner gets some reinforcement or she does
-not. Second, to emphasize the lack of locally smooth structure in the
-space of possible solutions, we choose as our solution space the space
-of functional programs. These two choices are in contrast to the
-paradigm tackled by most of machine learning, in which a locally
-informative objective function guides search in a space that is
-somewhat smooth.
-
-%% Let a \emph{task} be a function $t : \mathcal{L} \rightarrow \{0,
-%% 1\}$, where $\mathcal{L}$ is the set of all expressions in our
-%% representation language. We say that $e \in \mathcal{L}$ \emph{solves}
-%% task $t$ if $t(e) = 1$. Our goal is to solve as many tasks as
-%% possible, as efficiently as possible. That is, suppose that there is a
-%% distribution over tasks $t \sim  p_t(\cdot)$, we would like an ordering $O
-%% = e_1, \dots $ on expressions in $\mathcal{L}$ such that if $p_t(t_1)
-%% > p_t(t_2)$ then the first expression in $O$ to solve $t_1$ comes
-%% before the first expression in $O$ to solve $t_2$.
-
-%% We motivate a solution to this problem by considering a heirarchical
-%% Bayesian formulation of the task distribution. Suppose we believe that
-%% there is some underlying structure to $p_t(\cdot)$ that is explained by
-%% the fact that for each task $t$ there is a latent expressions $e \in
-%% \mathcal{L}$ drawn from a stochastic grammar $\mathcal{G}$ and that
-
-%% \begin{align}
-%% e &\sim p_\mathcal{L}(\cdot) = \mathcal{G}\\
-%% t | e &\sim \text{U}\{ \text{tasks solved by }e\}.
-%% \end{align}
-
-%% Then $p_t(t) = \sum_{e_j \in \mathcal{L} \text{ s.t. } t(e_j) = 1}
-%% p_\mathcal{L}(e_j)$. If we make the assumption that for the set $E$ of
-%% expressions in this sum, there is some expression $e^*$ such that
-%% $p_\mathcal{L}(e^*) >> p_\mathcal{L}(e), e \neq e^*$, then we have, approximately, 
-%% \[ p_t(t) \propto} p_\mathcal{L}(e^*). \]
-
-%% This suggests a way to find the ordering $O$: infer a stochastic
-%% grammar $G$, under the assumption that $G$ favors one solution to a
-%% given task over any other, and order the elements of $\mathcal{L}$
-%% according to $G$.
-
-%% Thus, our basic procedure will be to search for a single ``good''
-%% solution for as many tasks as we can, and to use those solutions to
-%% estimate a stochastic grammar. Our search will be performed in the
-%% order of the most recently inferred grammar. So as our ordering
-%% becomes more efficient and we can solve more tasks, we learn better
-%% stochastic grammars.
-
-%% Formally then, our goal is to solve the following minimal feedback
-%% multi-task program induction problem: suppose we are given set of
-%% \emph{tasks} $T=\{t_k\}_{k=1}^K$ where each task is a function $t_i :
-%% \mathcal{L} \rightarrow \{0, 1\}$, where $\mathcal{L}$ is the set of
-%% programs in our language.  We say that $ p \in \mathcal{L}$
-%% \emph{solves} $t_i$ if $ t_i\; p = 1$. Our goal is to find a set of
-%% solutions $X = \{x_k\}_{k=1}^k \subset \mathcal{L}$ such that
-
-%% \[
-%% X = \argmax_{x_1, \dots, x_K \in L} \sum_k t_k(x_k).
-%% \]
-
-Based on the motivation set out above, we present an iterative
-strategy. Given a stochastic grammar $\mathcal{G}$ which defines a
-distribution over the primitives from $\mathcal{L}$ and a integer
-\emph{frontier size} $N$, we perform two steps in each iteration.  In
-the first step, we explore the $N$ highest probability programs
-according to $\mathcal{G}$. In the second step, we compress the
-programs that solve any of our tasks, and we use that compression to
-define a new grammar. Repeating these two steps, we effectively
-reparameterize our search space to include programs that are more likely
-to be solutions to our tasks.
-
-The remainder of this paper will include three major sections. First,
-we will describe the representation language we chose and the goals of
-the exploration and compressions steps. Following that, we will
-discuss the algorithmic details that render our procedure
-feasible. Finally, we present results from two experiments, involving
-the domains of symbolic regression and boolean circuit learning.
+Many of the problems we want to solve in AI are best cast as search
+problems. There are many search algorithms, but they typically exploit
+local or topological structure in the search space. For example,
+convex optimization and gradient based search methods~\cite{} take
+advantage of local smoothness to move in good directions. Local search
+algorithms move from one candidate solution to a neighboring
+one\cite{}. Heuristic search techniques take advantage of topological
+structure to prune off large parts of the solution space~\cite{}.
+
+However, many of the problems we want solve -- discovering scientific
+theories, developing new technologies, figuring out how to bake -- do
+not come with such a structured search space. This is particularly
+true of problems, as these are, whose solutions can be naturally
+framed as programs. Here we ask how to extract structure from the
+space of programs by attempting to solve multiple related problems in
+a given domain at once.
+
+We take as inspiration the programmer who -- having reused several
+motifs, programming patterns, or subroutines to solve different but
+similar simple tasks -- creates a function library to help manage the
+complexity of larger projects. Our algorithm builds a library
+of reusable program components and a distribution over these,
+effectively learning from simple tasks a search space structure which
+enables it to solve more complex tasks.
+
+In this work, we try to see how far this idea of reusable function
+definition can get us. Our question is, Can an AI system discover the
+structure latent in the solutions to multiple related problems and
+thereby bootstrap effective exploration of the search space? Can
+discovering modular program components transform a problem from one in
+which search is intractable into a problem in which it becomes
+increasingly feasible with experience?
+
+Concretely, our algorithm works as follows. It begins with a small
+library of program primitives and a distribution over programs
+synthesized from this library. In each iteration, it generates
+programs in order of highest probability according to the library
+learned in the previous iteration. From the discovered solutions, it
+finds new reusable functions that compress these solutions, and it
+re-estimates the distribution over programs by weighting these new
+functions according to their frequencies.
+
+We show that this iterative approach allows us to achieve very high
+performance in tasks where just searching according to the initial
+distribution over program primitives clearly fails.
+
+\section{Related Work}
+The idea of discovering and using reusable subcomponents is an old one
+in AI. Rendell's 1985 paper~\cite{DBLP:conf/ijcai/Rendell85},
+``Substantial Constructive Induction Using Layered Information
+Compression: Tractable Feature Formation in Search,'' -- which could
+have been the name of this paper -- did not use programs as its
+solution space representation. Within the Inductive Logic Programming
+literature, researchers have explored ``Constructive Induction,''
+generating new and reusable logic programming
+terms~\cite{DBLP:conf/ijcai/Muggleton87}. We believe that applying
+constructive induction to a compositional functional programming
+paradigm will provide even greater benefits.
+
+Dicovering useful subroutines has been explored in Genetic
+Programming as
+well~\cite{DBLP:conf/foga/Koza92}~\cite{DBLP:conf/ices/KozaBAK96}.
+
+The importance of a curriculum has also been explored previously,
+though generally the curriculum has been explicitly given rather than
+discovered. Quinlan and Cameron-Jones~\cite{Quinlan93foil:a} showed
+that a logic program learning algorithm could learn sequentially more
+difficult list programs if the algorithm was provided with its own
+solutions to previously solved simple problems. More recently, several
+researchers have observed that presenting problems to machine learning
+algorithms in order of difficulty results in better
+performance~\cite{Krueger2009380}~\cite{DBLP:conf/icml/BengioLCW09}.
+
+Our use of a stochastic grammar over combinatory logic is based on
+Liang et al.~\cite{liang10programs}. In that work, a non-parametric
+distribution was placed over combinatory logic expressions, and an
+MCMC algorithm was developed to perform joint inference over these
+expressions and the underlying stochastic grammar. Our approach is
+inspired by this generative model, but avoids the difficulties
+associated with a stochastic local search over programs (the same
+difficulties faced by genetic programming), especially when the
+likelihood function (or the fitness function) are not of much help.
+
+%% AI and machine learning algorithms commonly take advantage of the
+%% relationship between a performance metric (error, reward, fitness,
+%% etc.) and the local or topological structure of the solution
+%% space. Whether in discrete or continuous domains, the key is to design
+%% such a relationship and write a algorithm that either moves in good
+%% directions or prunes off large parts of the solution space. Whether or
+%% not such a relationship exists depends on the representation, and much
+%% work has gone into finding good representations for problem
+%% solving. Recent work, especially in machine learning, has focused on
+%% ``feature learning,'' with the explicit aim of discovered good
+%% representational building blocks in vector-valued solutions spaces.
+
+%% Many of the most natural solution representations for intelligent
+%% agents, however, are not naturally expressed as vector-valued
+%% spaces. Consider the problems of discovering scientific theories,
+%% inventing new technologies, communicating with others, or executing a
+%% complex motor routine (such as dancing or playing an instrument). The
+%% solutions to these problems are most naturally expressed as programs,
+%% that is, as the composition of modular components whose semantics are
+%% determined by the actions of one component on another.
+
+%% How can we learn good solution representations for these kinds of
+%% tasks without building in the kind of domain knowledge that gives rise
+%% to a useful relationship between representation and performance? How
+%% can we get traction on a set of problems when small tweaks to
+%% solutions give rise to wildly varying performances in an unpredictable
+%% manner, or when, for most problems, most of the solutions fail in a
+%% uniformly bad way?
+
+%% A popular idea is that good representations are those that encode the
+%% data space in an efficient way, that is, that enable compression of
+%% the data~\cite{}. For program representations, this most naturally
+%% corresponds to defining reusable components so that the total number
+%% of unique expressions for a set of solutions is minimized (i.e., this
+%% is directly akin to the subroutine libraries programmers use to manage
+%% the complexity of their code).
+
+%% In this work, we try to see how far reusable subroutine definition can
+%% get us on its own. Our question is, Can discovering modular components
+%% transform a problem from one in which brute-force search will not work
+%% to a problem in which it does?
+
+%% To emphasize an impoverished relationship between reward signals and
+%% program representations, we make our tasks all-or-none: a solution
+%% either solves the problem or it does not, and partial solutions do not
+%% get any credit.
+
+%% The intuitive approach to applying this idea is simple: if solutions
+%% to problems in the domain share common structure, and there exist
+%% problems whose solutions are simple enough to find by brute force,
+%% then we should be able to extract a distribution over solution
+%% subcomponents from these simple cases that allows us to run subsequent
+%% searches more efficiently.
+
+%% More concretely, our algorithm works as follows: each iteration, it
+%% first generates programs in order of highest probability according to
+%% a subroutine distribution learned in the previous iteration. From the
+%% discovered solutions, it finds new reusable subroutines that compress
+%% these solutions and estimates their probabilities.
+
+%% We show that this iterative approach allows us to achieve very high
+%% performance in tasks where just searching according to the initial
+%% distribution over program primitives clearly fails.
+
+%%  \hline
+
+%% AI and machine learning algorithms commonly take advantage of the idea
+%% that local structure in the solution space is related to local
+%% structure in the outcome. In continuous domains, learning algorithms
+%% algorithms use error gradients to guide search. In discrete domains,
+%% whether in heuristic or local search paradigms, knowledge about the
+%% solution space topology and its relation to reward and cost metrics
+%% enables pruning, reasoning, and moving in the solution space more
+%% efficiently. However, choosing the representations that have
+%% beneficial structural properties requires a lot of general and domain
+%% knowledge, and this what researchers do and what they cannot leave up
+%% to their algorithms.
+
+%% What is a good strategy for an agent when its representation of the
+%% solution space is not smooth with respect to the outcome? Many hard
+%% learning problems have this property, especially those in which the
+%% solution space is most naturally represented as a set of
+%% programs. Consider the problems of discovering scientific theories,
+%% inventing new technologies, learning to cook, or learning to
+%% communicate. These are problems in which 1) solutions are best
+%% described as programs (i.e. compositions of modular subroutines at
+%% various levels of abstraction), 2) most of the solution space is bad
+%% in a relatively uninformative way (i.e. it just doesn't work), and
+%% thus 3) outcomes are not informative about how you are doing unless
+%% you are already quite close to a solution.
+
+%% Intuitively, the way to solve such problems is to find solutions to
+%% simple cases in the domain and extract conceptual building blocks that
+%% can be composed to scale up to complex cases. With experience, the
+%% agent may be able to discover local structure in the solution space,
+%% and utilize more traditional AI and machine learning techniques. In
+%% this work, we will be focusing on the first phase of this learning
+%% problem.
+
+%% %% Of course, every learner
+
+%% %% Choosing the wrong representation, in
+%% %% which local structure is not apparent, is often the difference between
+%% %% a successful learner and an unsuccessful one. In many respects, this
+%% %% may be the kind of situation that adult humans face when confronting
+%% %% new domains: we have a lot of general knowledge of successful
+%% %% representations, useful strategies, etc.
+
+%% %% But suppose that you have very little knowledge of a domain. Your
+%% %% intial representation of the domain is poor: ``nearby'' points in the
+%% %% solutions space ellicit wildly different outcomes in unpredictable
+%% %% ways. Or perhaps nearby points ellicit the same outcome and its not a
+%% %% good one, so you don't know which way to go. Many learning problems,
+%% %% especially those whose solutions are best described as programs, have
+%% %% these qualities: imagine learning to synthesize new sentences
+
+%% %% teness. We choose the representations our learners use
+%% %% and these representations generally Imagine you are trying to learn to
+%% %% bake via trial and error. You know what kind of ingredients are used
+%% %% in baking (flour, water, butter, sugar, yeast, baking soda, etc.) and
+%% %% you know a bit about how to combine these ingredient (mix, whisk,
+%% %% etc.) and how to apply heat (oven). But baking is a notoriously
+%% %% fragile process: ratios of ingredients, the order of mixing, cooking
+%% %% temperatures, resting times, all make a huge difference, and the
+%% %% outcome of the baking process is mostly all or none. Either the
+%% %% outcome is what you expected or quite unlike it. Even if you do create
+%% %% something similar to your intent, it is difficult to assign blame for
+%% %% the outcome to any one step or decision along the way.
+
+%% %% Many difficult learning problems have this structure -- learning
+%% %% language, developing technologies and scientific theories, learning to
+%% %% play instruments or execute other complex motor patterns. These are
+%% %% problems in which 1) there is a large set of similar tasks; 2) their
+%% %% solutions are best described as programs; 3) the solution space is
+%% %% not smooth with respect to the outcome; 4) and the outcome is generally
+%% %% uninformative unless you are already pretty close to a solution.
+
+%% %% Of course, knowledge is not acquired wholly by trial and error. But in
+%% %% many domains, there is at least initially such a period of ignorance,
+%% %% and the question we are trying to tackle here is how you ever get off
+%% %% the ground.
+
+%% I need help here.
+
+%% %% Let a \emph{task} be a function $t : \mathcal{L} \rightarrow \{0,
+%% %% 1\}$, where $\mathcal{L}$ is the set of all expressions in our
+%% %% representation language. We say that $e \in \mathcal{L}$ \emph{solves}
+%% %% task $t$ if $t(e) = 1$. Our goal is to solve as many tasks as
+%% %% possible, as efficiently as possible. That is, suppose that there is a
+%% %% distribution over tasks $t \sim  p_t(\cdot)$, we would like an ordering $O
+%% %% = e_1, \dots $ on expressions in $\mathcal{L}$ such that if $p_t(t_1)
+%% %% > p_t(t_2)$ then the first expression in $O$ to solve $t_1$ comes
+%% %% before the first expression in $O$ to solve $t_2$.
+
+%% %% We motivate a solution to this problem by considering a heirarchical
+%% %% Bayesian formulation of the task distribution. Suppose we believe that
+%% %% there is some underlying structure to $p_t(\cdot)$ that is explained by
+%% %% the fact that for each task $t$ there is a latent expressions $e \in
+%% %% \mathcal{L}$ drawn from a stochastic grammar $\mathcal{G}$ and that
+
+%% %% \begin{align}
+%% %% e &\sim p_\mathcal{L}(\cdot) = \mathcal{G}\\
+%% %% t | e &\sim \text{U}\{ \text{tasks solved by }e\}.
+%% %% \end{align}
+
+%% %% Then $p_t(t) = \sum_{e_j \in \mathcal{L} \text{ s.t. } t(e_j) = 1}
+%% %% p_\mathcal{L}(e_j)$. If we make the assumption that for the set $E$ of
+%% %% expressions in this sum, there is some expression $e^*$ such that
+%% %% $p_\mathcal{L}(e^*) >> p_\mathcal{L}(e), e \neq e^*$, then we have, approximately, 
+%% %% \[ p_t(t) \propto} p_\mathcal{L}(e^*). \]
+
+%% %% This suggests a way to find the ordering $O$: infer a stochastic
+%% %% grammar $G$, under the assumption that $G$ favors one solution to a
+%% %% given task over any other, and order the elements of $\mathcal{L}$
+%% %% according to $G$.
+
+%% %% Thus, our basic procedure will be to search for a single ``good''
+%% %% solution for as many tasks as we can, and to use those solutions to
+%% %% estimate a stochastic grammar. Our search will be performed in the
+%% %% order of the most recently inferred grammar. So as our ordering
+%% %% becomes more efficient and we can solve more tasks, we learn better
+%% %% stochastic grammars.
+
+%% %% Formally then, our goal is to solve the following minimal feedback
+%% %% multi-task program induction problem: suppose we are given set of
+%% %% \emph{tasks} $T=\{t_k\}_{k=1}^K$ where each task is a function $t_i :
+%% %% \mathcal{L} \rightarrow \{0, 1\}$, where $\mathcal{L}$ is the set of
+%% %% programs in our language.  We say that $ p \in \mathcal{L}$
+%% %% \emph{solves} $t_i$ if $ t_i(p) = 1$. Our goal is to find a set of
+%% %% solutions $X = \{x_k\}_{k=1}^k \subset \mathcal{L}$ such that
+
+%% %% \[
+%% %% X = \argmax_{x_1, \dots, x_K \in L} \sum_k t_k(x_k).
+%% %% \]
+
+%% Based on the motivation set out above, we present an iterative
+%% strategy. Given a stochastic grammar $\mathcal{G}$ which defines a
+%% distribution over the primitives from $\mathcal{L}$ and a integer
+%% \emph{frontier size} $N$, we perform two steps in each iteration.  In
+%% the first step, we explore the $N$ highest probability programs
+%% according to $\mathcal{G}$. In the second step, we compress the
+%% programs that solve any of our tasks, and we use that compression to
+%% define a new grammar. Repeating these two steps, we effectively
+%% reparameterize our search space to include programs that are more likely
+%% to be solutions to our tasks.
+
+The remainder of this paper will include two major sections.  We will
+first describe the representation language we chose and the details of
+the exploration and compression steps of the algorithm. We will then
+present results from two experiments involving the domains of
+symbolic regression and boolean circuit learning.
 
@@
 
+\section{Problem statement and terminology}
+Our goal is to solve the following minimal feedback multi-task program
+induction problem: suppose we are given set of \emph{tasks}
+$T=\{t_k\}_{k=1}^K$ where each task is a function $t_i : \mathcal{L}
+\rightarrow \{0, 1\}$ and where $\mathcal{L}$ is the set of programs
+in our language.  We say that $ p \in \mathcal{L}$ \emph{solves} $t_i$
+if $ t_i(p) = 1$. Our goal is to solve as many of the tasks in $T$ as
+we can.
+
+We will use the following terminology. A \emph{frontier} will be the
+finite set of programs that the algorithm is considering at any one
+time. We will say that a task is \emph{hit} by the frontier if there
+is a program in the frontier that solves it.
+
 \section{Extracting modular concepts from functional programs}
@@
 
-Following~\cite{liang10programs} and~\cite{Briggs:2008}, we use
-combinatory logic -- a variable-free subset of the polymorphic
-simply-typed lambda calculus~\cite{Pierce_2002} -- as
-our program representation language. In short, the combinatory logic
+Following~\cite{liang10programs} and~\cite{Briggs:2008}, we use a
+simply typed combinatory logic -- a variable-free subset of the
+polymorphic simply-typed lambda calculus~\cite{Pierce_2002} -- as our
+program representation language. In short, the combinatory logic
 introduces a \emph{basis} of several primitive \emph{combinators} such
@@
 
-The basis
-combinators are theoretically enough to express any Turing computable
-function, but we will assume that our language has a variety of
-primitives and, together with the basis combinators, we call these the
-primitive combinators.
+The basis combinators are theoretically enough to express any Turing
+computable function, but we will assume that our language has a
+variety of primitives and, together with the basis combinators, we
+call these the primitive combinators. Note that by default we consider
+all primitives to be in their ``curried'' form (e.g. a function like
+``+'' adds two numbers $x$ and $y$ by being first applied to $x$,
+returning the function $(+\;x)$ and then being applied to $y$,
+returning the object $(+\;x)\;y$).
 
 The basis combinators can themselves been expressed in the lambda
-calculus. The lambda calculus has two basic operations -- application
+calculus (with definitions following directly from the equations
+above). The lambda calculus has two basic operations -- application
 and abstraction -- but in using the combinatory logic we sequester
-uses of the abstraction operation inside the combinators, and thus
-have a representation language that is effectively abstraction- and,
-thus, variable-free. In doing so, we have replaced the variable
-binding of the $\lambda$ operator with the variable \emph{routing} of
-these basis combinators. See~\cite{liang10programs} for a more
-detailed discussion of this routing interpretation.
-
-This is very convenient for program synthesis: since every expression
-is the application of one expression to another -- with this recursion
-bottoming out at the primitive combinators -- each program is a binary
-tree. Most importantly, any subtree is itself a well-formed
-expression; this is not the case in the lambda calculus, since
-abstraction introduces long range dependencies between the $\lambda$
-operator and the variables to which that operator refers. In the
-lambda calculus, then, a subtree might have free variables, not bound
-to any enclosing $\lambda$.
+uses of the abstraction operation inside the combinators. In doing so,
+we have replaced the variable binding of the $\lambda$ operator with
+the variable \emph{routing} of these basis combinators. Our
+representation thus becomes variable-free. See Liang et
+al.~\cite{liang10programs} for a more detailed discussion of this
+routing interpretation.
+
+Using combinatory logic is very convenient for program synthesis:
+since every expression is the application of one expression to another
+-- with this recursion bottoming out at the primitive combinators --
+each program is a binary tree. Most importantly, any subtree is itself
+a well-formed expression; this is not the case in the lambda calculus,
+since abstraction introduces long range dependencies between the
+$\lambda$ operator and the variables to which that operator refers. In
+the lambda calculus, then, a subtree might have free variables, not
+bound to any enclosing $\lambda$.
 
@@
 (e.g. $\sigma$), or a function type $ t_1 \rightarrow t_2$ of
-functions from source type $t_1$ to target type $t_2$. Thus, this
-combinator can be represented as a binary
-tree~(Figure~\ref{fig:clbintree}) whose leaves are typed primitive
-combinators and whose interior nodes represent typed applications of
-combinators.
+functions from source type $t_1$ to target type $t_2$. Any combinator
+can be represented as a binary tree~(Figure~\ref{fig:clbintree}) whose
+leaves are typed primitive combinators and whose interior nodes
+represent typed applications of combinators (we annotate interior
+nodes with their types).
 
 \begin{figure}
-  \centering
-  \begin{subfigure}[Before]{0.5\linewidth}
-    \begin{tikzpicture}[scale=.9,
-        transform shape, level/.style={sibling distance=10mm/#1, font=\scriptsize}
+%%  \centering
+  \begin{subfigure}[]{0.45\linewidth}
+    \begin{tikzpicture}[scale=.8,
+        transform shape, level/.style={sibling distance=5mm/#1, font=\scriptsize}
         ]
-      \node [font=\scriptsize] (top) {$Int$}
-      child {node [above=.3cm, text width=2cm] (a) {$(Int \rightarrow Int) \rightarrow Int 
-          \rightarrow Int$}
-        child {node [left=.5cm, below=1cm, text width=3cm] 
-          {$S : \\(t_2 \rightarrow t_1 \rightarrow t_0)  \rightarrow (t_2 \rightarrow t_1) 
+      \node [font=\scriptsize] (top) {$\mathbb{R}$}
+      child {node [above=.3cm, left=.1cm, text width=1.8cm] (a) {$(\mathbb{R} \rightarrow \mathbb{R}) \rightarrow \mathbb{R} 
+          \rightarrow \mathbb{R}$}
+        child {node [left=.5cm, below=1cm, text width=3.0cm] 
+          {$ \mathbf{S} : (t_2 \rightarrow t_1 \rightarrow t_0)  \rightarrow (t_2 \rightarrow t_1) 
             \rightarrow \\ t_2 \rightarrow t_0$}}
-        child {node [right=0cm, text width=2cm] {$*: Int \rightarrow Int \rightarrow Int$}}
+        child {node [right=0cm, text width=2cm] {$ \mathbf{*}: \mathbb{R} \rightarrow \mathbb{R} \rightarrow \mathbb{R}$}}
       }
-      child {node [below right=.3cm and 1cm] (b) {$Int \rightarrow Int$}
-        child {node {$I: t_0 \rightarrow t_0 \rightarrow t_0$}}
-      };
+      child {node [below right=-.5cm and .2cm] (b) {$\mathbb{R} \rightarrow \mathbb{R}$}
+        child {node [above=.5cm] {$ \mathbf{I}: t_0 \rightarrow t_0 \rightarrow t_0$}}
+      }
     \end{tikzpicture}
+    \caption{}
     \label{fig:clbintree}
   \end{subfigure}
-  \begin{subfigure}[After]{0.4\linewidth}
+  \begin{subfigure}[R]{0.5\linewidth}
     \begin{tikzpicture}[scale=.7, transform shape, 
@@
 \label{sec:stochgrammar}
-A stochastic grammar over programs represents a distribution over
-expressions in $\mathcal{L}$. We associate with each primitive
-combinator $c_i$ in $\mathcal{L}$ a probability $p_i$,
-$0~\leq~p_i~\leq~1, \sum_i~p_i~=~1 $.
-
-We define the probability of an expression $e$ consisting of primitive
-combinators $c_1, \dots, c_N$ to be $p(e)~=~\prod_{n=1}^N p(c_n |
-t(c_n))$ where $t(c_n)$ is the requested type when $c_n$ was used and
+We would like a way to describe how likely we believe a program to be
+given the probabilities we ascribe to its components. Recall that
+$\mathcal{L}$ is the set of all programs.  A stochastic grammar over
+programs represents a distribution over expressions in
+$\mathcal{L}$. We associate with each primitive combinator $c_i$ in
+$\mathcal{L}$ a probability $p_i$, $0~\leq~p_i~\leq~1, \sum_i~p_i~=~1
+$.
+
+The probability of using a primitive combinator somewhere in a program
+should depend on which other combinators could have been used in its
+place. This, in turn, depends on the type at that particular location
+in the program tree; we call this type the \emph{requesting} type. For
+example, suppose that we have the primitive $(+1)$ of type $Int
+\rightarrow Int$ and the primitive $NOT$ of type $Bool \rightarrow
+Bool$. If the requesting type is $\sigma \rightarrow Int$, then we can
+only use $(+1)$, but if the requesting type is $\sigma \rightarrow
+\sigma$ then we could use either $(+1)$ or $NOT$. Let $C = c_1, \dots,
+c_n$ be the set of primitive combinators in our grammar; we define
+$C_t$ for any type $t$ to be the set of combinators in $C$ that are
+consistent with $t$, that is, whose types \emph{unify} with t (we use
+unification according to the standard type theory
+presentation~\ref{Pierce_2002}).
+
+We define the probability of an expression $e$ from a grammar with
+primitive $C$ to be $p(e)~=~\prod_{n=1}^N p(c_n | t(c_n))$ where
+$p(c_n | t(c_n))$ is the probability of using primitive $c_n$ when the
+requesting type was $t(c_n)$. This definition means that the
+probability of an expression only depends on the probabilities
+assigned to the library primitives. We define these conditional
+probabilities as
 \[
-p(c_i | t(c_n)) =    \frac{1}{4}\frac{p_i}{\sum_{c_j :
-    type(c_j) \text{ unifies with } t(c_n)} p_j}.
+p(c_i | t(c_n)) =    \frac{1}{\epsilon}\frac{p_i}{\sum_{C_{t(c_n)}} p_j}.
 \]
@@
 the multinomial distribution defined by the $p_i$'s, conditioned on
-the requesting type, and then divided by 4. The constant factor of
-$1/4$ is introduced as a simple way to ensure that the sum over the
-probabilities of all expressions converges to finite value (so that
-the distribution over expressions is proper).\footnote{The probability
-  mass $M_d$ of all expressions of depth less than or equal to $d$ can
-  be written with the recurrence relation $M_d \leq M_d^2 + M_1$. This
-  has an asymptotic fixed point if $x = x^2 + M_1$ has a solution,
-  which is true if $M_1 \leq 1/4$.}
-
-To estimate the values $p_i$ from set of expressions, we simply divide
+the requesting type, with the factor of $\frac{1}{\epsilon}$ added as
+a simple way to ensure that the sum over the probabilities of all
+expressions converges to a finite value (so that the distribution over
+expressions is proper). This is true whenever $\epsilon \geq
+4$\footnote{The probability mass $M_d$ of all expressions of depth
+  less than or equal to $d$ can be written with the recurrence
+  relation $M_d \leq M_{d-1}^2 + M_1$. This has an asymptotic fixed
+  point, as $d \rightarrow \infty$ if $x = x^2 + M_1$ has a solution,
+  which is true if $M_1 \leq 1/4$.}.
+
+To estimate the values $p_i$ from a set of expressions, we simply divide
 the number of times primitive combinator $c_i$ was used by the number
@@
 enumerating the shortest program satisfying some
-criterion~\cite{DBLP:conf/sfp/Katayama05} \cite{DBLP:conf/aaip/YakushevJ09}. Our
-approach is to formulate the problem as a best-first exploration of an
-AND/OR tree~\cite{nilsson1982principles}, which we denote
-$\mathcal{G}$. A complete program is any binary subtrees of
-$\mathcal{G}$ whose leaves coincide with the leaves of
-$\mathcal{G}$. The tree structure is shown in
-Figure~\ref{fig:andor}. The root is an OR node of the requisite
-program type. It's children are the primitives of that type $f_1,
-\dots, f_n$, and one AND node with two children. Each child of an AND
-node has the same structure as the root node, with modified types: if
-the type of an AND node is $t$, then the type of its left child is
-$\tau \rightarrow t$ (where $\tau$ is a fresh type variable). Since
-the type of the right child is constrained by the subtree rooted at
-the left child, we always expand the left child first and use the
-target type of this left subtree to mark the type of the right child.
-
-Recall that a valid partial solution is a subtree of $\mathcal{G}$
-satisfying the following properties: its root is the root
-$\mathcal{G}$ and any OR node has at most one child. A complete
-solution is a partial solution whose leaves are leaves of
-$\mathcal{G}$, whose root is the root of $\mathcal{G}$, in which each
-OR node has exactly one child, and in which each AND node has exactly
-two children.
-
-The value of a partial solution $G'$ is calculated by summing the
-value of its leaves. If a leaf of $G'$ is a leaf of $G$ then either it
-is an AND node, to which we assign a value of 0, or it corresponds to
-some primitive combinator $c_i$ in the library and is the child of
-some typed OR node with type $t$. We assign its value of $- \log{p(c_i
-  | t)}$, according to the stochastic grammar specified in
-Section~\ref{sec:stochgrammar}. If a leaf $v$ of $G'$ is an OR node,
-then the value of any extension of it to a complete solution is at
-most $-log(1/4)$. Thus, we have defined a value for partial solutions
-that upper bounds the value of any of its extensions.
-
-Our best-first policy, is to simply expand the next partial solution
+criterion~\cite{DBLP:conf/sfp/Katayama05}
+\cite{DBLP:conf/aaip/YakushevJ09}. Our approach is to formulate the
+problem as a best-first exploration of an AND/OR tree
+(Figure~\ref{fig:andor})~\cite{nilsson1982principles}~\cite{DBLP:journals/cacm/Hall73},
+which we denote $\mathcal{G}$. In words, the enumeration procedure is
+best described by the following recursion: every program is either a
+primitive combinator or it a left child program applied to a right
+child program.
+
+This translates to the following AND/OR tree. Suppose we want a
+program of type $t$. The root of $\mathcal{G}$ is an OR node of type
+$t$. Its children are elements in $C_t$ (defined in the previous
+section) and one AND node of type $t$ with two children. Each child of
+an AND node has the same structure as the root node, with modified
+types: type of its left child is $\tau \rightarrow t$ (where $\tau$ is
+a fresh type variable). Since the type of the right child is
+constrained by the subtree rooted at the left child, we always expand
+the left child first. Once we have a complete left child program, we
+can use its target type as the type of the right child.
+
+Recall that a valid partial solution $\mathcal{H}$ is a subtree of
+$\mathcal{G}$ satisfying the following properties: the root of
+$\mathcal{H}$ is the root $\mathcal{G}$ and any OR node in
+$\mathcal{H}$ has at most one child. $\mathcal{H}$ is a complete
+solution (and thus a complete program) if it is a partial solution
+whose leaves are leaves of $\mathcal{G}$, if each OR node in
+$\mathcal{H}$ has exactly one child, and if each AND node in
+$\mathcal{H}$ is connected to all of that node's children in
+$\mathcal{G}$
+
+The value of a partial solution $\mathcal{H}$ is calculated by summing
+the value of its leaves, where each such value is the log of the
+probability assigned to that leaf in
+Section~\ref{sec:stochgrammar}. If a leaf of $\mathcal{H}$ is a leaf
+of $\mathcal{G}$ then either it is an AND node, to which we assign a
+value of 0, or it corresponds to some primitive combinator $c_i$ in
+the library and is the child of some typed OR node with type $t$. We
+assign it a value of $\log{p(c_i | t)}$. If a leaf $v$ of
+$\mathcal{H}$ is an OR node, then the value of any extension of it to
+a complete solution is at most $log(1/4)$. Thus, we have defined a
+value for partial solutions that upper bounds the value of any of its
+extensions.
+
+Our best-first policy is to expand the next partial solution
 with the highest value. If a partial solution has multiple leaves that
@@
 
-\textbf{****NEED COMPLEXITY ANALYSIS HERE*******}
-
-\subsection{Finding the most compressive set of solutions.}
-
-We say that a task is \emph{hit} if it has a solution in the
-frontier. Having enumerated the frontier, we want to assign solutions
+\subsection{Finding the most compressive set of solutions}
+
+Having enumerated the frontier, we want to assign solutions
 to hit tasks such that this set is maximally compressible. This will
@@
 A natural metric for the compressibility of a set of binary trees is
-the number of unique subtrees that this set contains. As an example of
-this, suppose we had three tasks whose solutions were integers 16, 25,
-and 36, and suppose our frontier contained the six expressions:
-
-%% \begin{figure}[h]
-%%     \begin{tikzpicture}[scale=.9,
-%%         transform shape, level/.style={sibling distance=10mm/#1, font=\scriptsize}
-%%         ]
-%%       \node [font=\scriptsize] (top) {}
-%%       child {node [above=.3cm, text width=2cm] (a) {$(Int \rightarrow Int) \rightarrow Int 
-%%           \rightarrow Int$}
-%%         child {node [left=.5cm, below=1cm, text width=3cm] 
-%%           {$S : \\(t_2 \rightarrow t_1 \rightarrow t_0)  \rightarrow (t_2 \rightarrow t_1) 
-%%             \rightarrow \\ t_2 \rightarrow t_0$}}
-%%         child {node [right=0cm, text width=2cm] {$*: Int \rightarrow Int \rightarrow Int$}}
-%%       }
-%%       child {node [below right=.3cm and 1cm] (b) {$Int \rightarrow Int$}
-%%         child {node {$I: t_0 \rightarrow t_0 \rightarrow t_0$}}
-%%       };
-%%     \end{tikzpicture}
-%% \end{figure}
-
-Let $|e|$ be the
-number of unique binary trees in expression $e$.
-
-we will find that some tasks have no solutions
-and some have one or more solutions; for those tasks that have
-solutions, we want to choose a solution for each task such that the
-complete set of such solutions for all solved tasks is as compressible
-as possible.
-
-Thus, we can attempt
-to find the assignment of solutions to tasks that minimizes this
-value. However, an exact optimization for this problem requires
-examining $O(m^k)$ assignments, where $m$ is the maximum number of
-solutions for any one task and $k$ is the number of tasks. Since this
-is prohibitive, we relax the compressibility metric as follows: we
-define the cost of adding each new solution $s_n$ to a partial set of
+the number of unique subtrees (counting leaves as subtrees) that this
+set contains. As an example of this, suppose we had six tasks whose
+solutions were integers $4^2, 5^2, 6^2, 7^2, 8^2$ and suppose our
+frontier contained the 10 expressions:
+
+\captionsetup[subfigure]{labelformat=empty} 
+\begin{figure}[h!]
+   \begin{subfigure}{.17\linewidth}
+     \begin{tikzpicture}[scale=.7, background rectangle/.style={fill=lightgray}] 
+       \tikzset{level distance=15pt, sibling distance=10pt}
+       \Tree [[  * 4 ] 4 ]
+     \end{tikzpicture}
+  \end{subfigure}
+  \begin{subfigure}{.17\linewidth}
+     \begin{tikzpicture}[scale=.7]
+       \tikzset{level distance=15pt, sibling distance=10pt}
+       \Tree [[  * 5 ] 5 ]
+     \end{tikzpicture}
+  \end{subfigure}
+   \begin{subfigure}{.17\linewidth}
+     \begin{tikzpicture} [scale=.7]
+       \tikzset{level distance=15pt, sibling distance=10pt}
+       \Tree [[  * 6 ] 6 ]
+     \end{tikzpicture}
+  \end{subfigure}
+   \begin{subfigure}{.17\linewidth}
+     \begin{tikzpicture}[scale=.7] 
+       \tikzset{level distance=15pt, sibling distance=10pt}
+       \Tree [[  * 7 ] 7 ]
+     \end{tikzpicture}
+  \end{subfigure}
+  \begin{subfigure}{.17\linewidth}
+     \begin{tikzpicture}[scale=.7] 
+       \tikzset{level distance=15pt, sibling distance=10pt}
+       \Tree [[  * 8 ] 8 ]
+     \end{tikzpicture}
+  \end{subfigure}
+  \vspace{2mm}
+
+  \hdashrule{.9\linewidth}{1pt}{1pt}
+
+  \vspace{2mm}
+  \begin{subfigure}{.17\linewidth}
+     \begin{tikzpicture}[scale=.6] 
+       \tikzset{level distance=15pt, sibling distance=10pt}
+       \Tree [[[  S * ] I ] 4 ]
+     \end{tikzpicture}
+     \caption{(16)}
+  \end{subfigure}
+  \begin{subfigure}{.17\linewidth}
+     \begin{tikzpicture}[scale=.6]
+       \tikzset{level distance=15pt, sibling distance=10pt}
+       \Tree [[[  S * ] I ] 5 ]
+     \end{tikzpicture}
+     \caption{(25)}
+  \end{subfigure}
+   \begin{subfigure}{.17\linewidth}
+     \begin{tikzpicture} [scale=.6]
+       \tikzset{level distance=15pt, sibling distance=10pt}
+       \Tree [[[  S * ] I ] 6 ]
+     \end{tikzpicture}
+     \caption{(36)}
+  \end{subfigure}
+  \begin{subfigure}{.17\linewidth}
+     \begin{tikzpicture}[scale=.6] 
+       \tikzset{level distance=15pt, sibling distance=10pt}
+       \Tree [[[  S * ] I ] 7 ]
+     \end{tikzpicture}
+     \caption{(49)}
+  \end{subfigure}
+  \begin{subfigure}{.17\linewidth}
+     \begin{tikzpicture}[scale=.6] 
+       \tikzset{level distance=15pt, sibling distance=10pt}
+       \Tree [[[  S * ] I ] 8 ]
+     \end{tikzpicture}
+     \caption{(64)}
+  \end{subfigure}
+\end{figure}
+\captionsetup[subfigure]{labelformat=parens}
+
+The top row consists of one type of representation of an integer
+squared, namely, the integer times itself. The bottom row contains
+another type of representation in which the function $f(x)=x^2$,
+represented in combinatory logic as $S * I$, is applied to the
+integer. The top and bottom row of each expression evaluates to the
+same integer. If we consider only the first column, the top tree has 4
+unique subtrees and the bottom has 7. But each column adds 3 unique
+subtrees to the top row and only 2 to the bottom. So by the fifth
+column the top row has more unique subtrees than the bottom row. The
+increased compressibility of the bottom row is due to the fact that
+the squaring operation has been factored out as an abstraction. This
+is an example, then, of how this compressibility metric favors
+abstraction when multiple expressions are being considered at once.
+
+We want to choose a solution for each hit task such that the complete set
+of such solutions for all solved tasks is as compressible as
+possible. That is, let $t_i$ be a task in the set of hit tasks, and
+let $\{e^k_i\}$ be the expressions in the frontier that solves
+it. Then our goal is to find
+\[
+\{e^*_i\} = \argmin_{\{e^k_i\}} | \cup_i e^k_i |,
+\]
+where $|\cdot|$ denotes the number of unique trees in a set of expressions.
+
+However, an exact optimization of this problem requires examining
+$O(M^K)$ assignments, where $M$ is the maximum number of solutions for
+any one task and $K$ is the number of tasks. Since this is
+prohibitive, we relax the compressibility metric as follows: we define
+the cost of adding each new solution $s_n$ to a partial set of
 solutions $s_1, \dots, s_{n-1}$ to be the number of additional unique
-subtrees in $s_n$ that are not present in $s_{n-1}$. Our solution thus
-becomes approximate and that approximation is order-dependent, but a
+subtrees in $s_n$ that are not present in $s_{n-1}$. That is, the
+penalty of adding a new expression to the set of expressions depends
+only on the last expression we added. Our solution thus becomes
+approximate and that approximation is order-dependent, but a
 depth-first search on the defined search space goes from exponential
-in the solution degree to quadratic.
+in the solution degree to quadratic ($O(KM^2)$).
+
+However, when the number of solutions in the frontier grows, this
+quadratic computation becomes prohibitive. Therefore, in practice, we
+bound the number of potential expressions we are willing to consider
+at any one iteration to any one task to a small number. We find that a
+setting of this bound to 2 is computationally feasible and large
+enough to guide subcomponent discovery in a reliable way. It may be
+necessary to increase this bound as we scale this algorithm to larger
+problems.
 
@@
 Given a set of solution programs, we want to re-estimate our
-stochastic grammar. Our inspiration for doing this is based broadly on
+stochastic grammar. Our inspiration for doing this is based on
 the Neville-Manning algorithm for grammar-based
@@
 grammar and b) every rule is used more than once. From a set of trees
-we generate a grammar according to these criteria (though using a less
-efficient and much simpler algorithm).
-
-In our algorithm, we traverse a set of combinator binary trees in
-depth first order. We count the occurrences of every subtree. Every
-time we encounter a subtree for the \emph{second} time, we designate
-it as a primitive in our new grammar. This ensures that every
-primitive is used more than once. During our traversal, we do not
-descend into trees that have been added already to the new
-grammar. Thus, we only count an instance of tree if that instance is
-not a subtree of a primitive element of the new grammar.
-
-At the end of this traversal of the solution programs, we have a parse
-of the solution set with a new set of primitives and we know how many
-times each primitive occurred in the solutions set. To estimate the
-probabilities associated with each node production, we again traverse
-the solution set, but this time for each node we count which other
-primitive elements from the new grammar could have been used. We
-estimate the terminal production weight to be the ratio of the number
-of times that node was used in the first traversal divided by the
-number of times it was used in the second traversal.
-
-Note that this probability estimation procedure was chosen for its
-simplicity, and it is not the maximum likelihood estimator for the
-multinomial distribution according to which we enumerate the
-data. Because the probability we assign to the use of a primitive
-combinator depends on the type of the node, the ML estimator of the
-probability of that combinator is not a simple ratio of counts. It
-would be interesting to determine whether a more accurate estimator
-can improve this algorithm's performance.
+we generate a grammar according to these criteria.
+
+%% To do this, we traverse a set of combinator binary trees in
+%% depth first order. We count the occurrences of every subtree. Every
+%% time we encounter a subtree for the \emph{second} time, we designate
+%% it as a primitive in our new grammar. This ensures that every
+%% primitive is used more than once. During our traversal, we do not
+%% descend into trees that have been added already to the new
+%% grammar. Thus, we only count an instance of a tree if that instance is
+%% not a subtree of a primitive element of the new grammar.
+
+This procedure generates a parse of the solution set with a new set of
+primitives, and we keep count of how many times each primitive
+occurred in the solutions set. To estimate the probabilities
+associated with each node production, we again traverse the solution
+set, but this time for each node we count which other primitive
+elements from the new grammar could have been used. In accordance with
+Section~\ref{stochgrammar}, we estimate the terminal production weight
+to be the ratio of the number of times that node was used in the first
+traversal divided by the number of times it was used in the second
+traversal.
 
@@
 
-In this section, we walk through the performance of our learning
-algorithm on a symbolic regression problem. In this problem,
-each task is a set of pairs $ (x, f(x))$ where $x \in (0, \dots, 10)$
-and $f$ is a polynomial. The goal is to discover a program for each
-task that outputs the appropriate value $f(x)$ for each input x. We
-assume that the data is noise-free.
+In this section, we discuss the performance of our learning algorithm
+on a symbolic regression problem. In this problem, each task is a set
+of pairs $ (x, f(x))$ where $x \in (0, \dots, 9)$ and $f$ is a
+polynomial. The goal is to discover a program for each task that
+outputs the appropriate value $f(x)$ for each input x.
 
@@
 Each task is to predict the value of an unknown polynomial on the
-inputs $0, \dots, 10$. The set of tasks corresponds to the set of
-polynomials $\{ax^2 + bx + c | a, b, c \in \{0, \dots, 10 \}\}$. The
-choice of primitives is going to make a big difference in performance;
-we chose to include the basic combinators I, S, B, C and four
+inputs $0, \dots, 9$. The set of tasks corresponds to the set of
+polynomials $\{ax^2 + bx + c | a, b, c \in \{0, \dots, 9 \}\}$. In the
+initial library, we include the basic combinators I, S, B, C and four
 arithmetic primitives, $1$, $0$, $*$, and $+$. The choice of initial
-weights makes a large difference as well, as it determines the
-relative ordering of combinators in the enumeration step. To get a
-general sense of the algorithm's performance, we set the intial
-weights to be slightly different on each run, jittering them around a
+weights on these primitives can make a difference, as it determines
+the relative ordering of combinators in the enumeration step. To get a
+general sense of the algorithm's performance, we set the initial
+weights to be slightly different on each run, perturbing them around a
 uniform weighting.
 
-In Figure~\ref{fig:symreg}, we show performance results as we vary the
-frontier size. Each individual plot shows how changing the frontier
-size changes the number of tasks the algorithm identifies correctly
-over the course of fifteen algorithm iterations. %% In addition, we have a
-%% baseline for each curve, visits the same number of expressions as the
-%% corresponding algorithm run (i.e. frontier size $\times$ $\#$ of
-%% iterations) but all at once; it does not do any learning. Thus, this
-%% baseline serves as a brute-force control for the learning aspect of
-%% the algorithm.
-
-\begin{figure}
-\begin{subfigure}[Before]{.6\linewidth}
-%% \includegraphics[width=\linewidth]{figures/learningCurves2.pdf}
-\includegraphics[width=\linewidth]{figures/symreg3frontiers.pdf}
-\end{subfigure}
-%% \begin{subfigure}[After]{0.45\linewidth}
-%% \includegraphics[width=\linewidth]{figures/performanceVsFrontierSize2.pdf}
-%% \end{subfigure}
-\caption{(a) Average learning curves as we vary the frontier size. (b)
-  Percent of tasks solved at iteration 15 as a function of frontier
-  size.}
+\begin{figure}[h]
+\centering
+\begin{subfigure}[Before]{\linewidth}
+\includegraphics[width=.9\linewidth]{figures/symreg3frontiers.pdf}
+\label{fig:symregLearning}
+\caption{Learning curves as a function of frontier size.  As frontier
+  size is increased, it plateaus closer to $100\%$ performance. A
+  control brute-force search over $150000$ only hits $3\%$ of the
+  tasks.}
+\end{subfigure}
+
+\begin{subfigure}[After]{.65\linewidth}
+\includegraphics[width=.9\linewidth]{figures/symRegDiffCurricula.pdf}
+\caption{How do different task sets affect learning curves? Learning
+  curves at frontier size of $10000$ for different task sets. Note
+  that performance is equally good if either the constant functions of
+  the linear functions are missing, by drops dramatically when both
+  are absent, even though these only account for about $10\%$ of the
+  tasks. Performance drops even lower if we only consider ``complex'' quadratics.}
+\label{fig:symregCurricula}
+\end{subfigure}
+\begin{subfigure}[After]{.3\linewidth}\
+\includegraphics[width=.9\linewidth]{figures/symregCircuitDepths.pdf}
+\caption{Circuit depth as a function of iteration on a typical run.}
+\label{fig:symregDepths}
+\end{subfigure}
+
+\caption{Symbolic regression experiment data.}
 \label{fig:symreg} 
@@
 
+In Figure~\ref{fig:symreg}, we show performance results as we vary the
+frontier size. Each individual plot shows how changing the frontier
+size changes the number of tasks the algorithm identifies correctly
+over the course of fifteen algorithm iterations. As a control, we
+enumerated $10000*15=150000$ expressions from the initial
+grammar. This is the total number of expressions that a run of the
+algorithm sees if it has a frontier size of $10000$ and runs for $15$
+iterations. Thus, if the benefit accruing to the runs with larger
+frontier sizes is due to simple an increase in brute-force search
+ability, rather than increased learning, we should see similarly good
+performance from this control run. In fact, however, the control run
+only hit $27$ of the $1000$ asks ($3\%$), whereas our algorithm nears $100\%$.
+
+We would like to assess to what degree ``conceptual bootstrapping'' is
+responsible for the improved performance. In the next experiment we
+will inspect directly the modular components that are learned by this
+algorithm. Here we look at the idea of a ``learnable'' curriculum,
+that is, the notion that conceptual bootstrapping requires that a
+set of tasks contain a subset of tasks that are sufficiently simple so
+that the learner is able to solve those with its initially
+impoverished library of concepts.
+
+In Figure~\ref{fig:symregCurricula}, we present learning curves
+(frontier size 10000) corresponding to various versions of our
+original symbolic regression task set. Recall that the original set
+consisted of all polynomials of degree two or less and with
+coefficients between 0 and 9. In the ``no constant'' task set, we
+remove the 9 constant functions in this set. In the ``no linear'' task
+set, we remove 90 linear functions from the set. We observe that
+performance on those task sets does not decrease. However, when we
+remove both the linear functions and the constant function (``only
+quadratics''), we see a sharp drop in the algorithm's
+performance. When we further restrict the task set to only ``complex''
+quadratics (which we are defining as quadratics all of whose
+coefficients are greater than zero), we observe another comparable
+drop in performance. When we go one step further and restrict the task
+set to quadratics with coefficients greater than 1, performance drops
+to 0.
+
+%% In addition, we have a
+%% baseline for each curve, visits the same number of expressions as the
+%% corresponding algorithm run (i.e. frontier size $\times$ $\#$ of
+%% iterations) but all at once; it does not do any learning. Thus, this
+%% baseline serves as a brute-force control for the learning aspect of
+%% the algorithm.
+
 \subsection{Boolean circuits}
@@
 inputs downstream.
+
+\begin{table}
+\tiny
+\begin{tabular}[width=\linewidth]{|p{1cm}|p{2cm}|p{3.5cm}|}
+\hline \\
+  function  & CL expression & circuit  ~ \\ \hline
+¬†¬†¬†¬†¬†¬†¬†¬†TRUE¬† & (S NAND) (S NAND I)¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬† & ~
+
+%\begin{top}
+\begin{circuitikz}[transform shape] 
+\draw
+(1, 0) node[nand port, scale=.5] (nand1) {}
+(2,.3) node[nand port, scale=.5] (nand2) {}
+(0, 0) node[anchor=east]  (x) {x}
+(x) -- (.3, 0) -| (nand1.in 1)
+(x) -- (.3, 0) -|  (nand1.in 2)
+(x) |- (nand2.in 1)
+(nand1.out) |- (nand2.in 2)
+;
+\end{circuitikz}
+%\end{top}
+\\ \hline
+¬†¬†¬†¬†¬†¬†¬†¬†FALSE & (S B (S NAND)) (S NAND I)¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬† &  ~
+
+%% \begin{circuitikz}[transform shape]
+
+%% ;
+%% \end{circuitikz}
+
+\\ \hline
+¬†¬†¬†¬†¬†¬†¬†¬†NOT¬†¬† & S NAND I¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬† &  ~
+
+\begin{circuitikz}[transform shape]
+\draw
+(1.2,0) node[nand port, scale=.5] (nand1) {}
+(0, 0) node (x) {x}
+(x) -| (nand1.in 1)
+(x) -| (nand1.in 2)
+;
+\end{circuitikz}
+
+\\ \hline
+¬†¬†¬†¬†¬†¬†¬†¬†AND¬†¬† & ((C B NAND) (B (S NAND I)))¬†¬†¬†¬†¬†¬†¬†¬†¬†¬† & ~
+
+\begin{circuitikz}
+\draw
+(1.0,0) node[nand port, scale=.5] (nand1) {}
+(2.0,0) node[nand port, scale=.5] (nand2) {}
+% (0, 0) node (x) {x}
+% (0, 0) node (y) {y}
+node[above left=.2cm and .7cm  of nand1] (x) {x} %-- (nand1.in 1)
+node[below left=.2cm and .7cm  of nand1] (y) {y} %-- (nand1.in 2)
+(x) |- (nand1.in 1)
+(y) |- (nand1.in 2)
+(nand1.out) |- (nand2.in 1)
+(nand1.out) |- (nand2.in 2)
+;
+\end{circuitikz}
+ \\  \hline
+¬†¬†¬†¬†¬†¬†¬†¬†OR¬†¬†¬† & ((B (C (B (S NAND) NAND))) (S NAND I) & ~
+
+\begin{circuitikz}
+\draw
+(0,0) node (x) {x}
+(0,.7) node (y) {y}
+
+node[nand port, scale=.5, right=1cm of x] (nand1) {}
+node[nand port, scale=.5, right=1cm of y] (nand2) {}
+node[nand port, scale=.5, below right=.3cm and 1cm of nand2] (nand3) {}
+(x) -| (nand1.in 1)
+(x) -| (nand1.in 2)
+(y) -| (nand2.in 1)
+(y) -| (nand2.in 2)
+(nand1.out) |- (nand3.in 1)
+(nand2.out) |- (nand3.in 2)
+;
+\end{circuitikz}
+ \\  \hline
+¬†¬†¬†¬†¬†¬†¬†¬†XOR¬†¬† & ((S (B C (((B (C B NAND)) S) (((B (C B NAND)) S) (B C ((B (B NAND)) NAND)))))) I)¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬† & ~
+
+\begin{circuitikz}
+\draw
+node[nand port, scale=.5] (nand1) {}
+node[nand port, scale=.5, above right=.3cm and 1cm of nand1] (nand2) {} 
+node[nand port, scale=.5, below right=.3cm and 1cm of nand1] (nand3) {} 
+node[nand port, scale=.5, right=2cm of nand1] (nand4) {} 
+node[left=1.2cm of nand2.in 1] (x) {x}
+node[left=1.2cm of nand3.in 2] (y) {y}
+node[right=.2cm of nand4] (out) {}
+(x) -| (nand1.in 1)
+(x) -| (nand2.in 1)
+(y) -| (nand1.in 2)
+(y) -| (nand3.in 2)
+(nand1.out) -| (nand2.in 2)
+(nand1.out) -| (nand3.in 1)
+(nand2.out) |- (nand4.in 1)
+(nand3.out) |- (nand4.in 2)
+(nand4.out) -- (out)
+;
+\end{circuitikz}
+
+ \\  \hline
+¬†¬†¬†¬†¬†¬†¬†¬†F¬†¬†¬†  & S B (S NAND)¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬† & ~
+%% \begin{circuitikz}
+%% \draw
+%% (0, 0) node[left=.25) (x) {x}
+
+%% ;
+%% \end{circuitikz}
+
+\\ 
+\hline
+
+\end{tabular}
+\caption{Expressions in the induced library on sets of tasks
+  corresponding to boolean circuits synthesized from AND, OR, and NOT
+  gates, with boolean circuits of 1, 2, or 3 inputs.}
+\label{table:boolexpr}
+
+\end{table}
+
+%% \begin{table}[t]
+%% \scriptsize
+%%    \begin{tabular}{|l|l|l|}
+%%     function  & CL expressions & circuit  ~ \\ \hline
+%%         TRUE  & (S NAND) (S NAND I)                   & ~ \\ \hline
+%%         FALSE & (S B (S NAND)) (S NAND I)             & ~ \\ \hline
+%%         NOT   & S NAND I                              & \input{figures/NOT.tex} \\ \hline
+%%         AND   & ((C B NAND) (B (S NAND I)))           & ~ \\ 
+%%         OR    & ((B (C (B (S NAND) NAND))) (S NAND I) & ~ \\ 
+%%         XOR   & ~                                     & ~ \\ 
+%%         F     & S B (S NAND)                          & \input{figures/F.tex}
+%%     \end{tabular}
+%% \end{table}
 
@@
 $82$ unique boolean functions, with a distribution of tasks as shown
-in Figure~\ref{fig:booldistr}. In Figure~\ref{fig:boolLearningCurves1},
+in Figure~\ref{fig:booldistr}. In Figure~\ref{fig:boolLearningCurvesCircuits},
 We present learning curves for frontier sizes 100, 500 and 10000,
@@
 \includegraphics[width=\linewidth]{figures/circuitDistrComp.pdf}
-\label{fig:booldistr}
+\vspace{-10pt}
 \caption{Distribution of boolean functions among (a) the set of 1000
-  circuits randomly constructed from AND, OR and NOT gates; b) the
-  first 1000 expressions from the grammar over programs \emph{before}
-  learning; c) the first 1000 expressions from the grammar over
-  programs \emph{after} learning.}
+  circuits randomly constructed from AND, OR and NOT gates (see text);
+  b) the first 1000 expressions from the grammar over programs
+  \emph{before} learning; c) the first 1000 expressions from the
+  grammar over programs \emph{after} learning. Gray levels indicate
+  percent of instances (circuits/expressions) which evaluate to that
+  function.\label{fig:booldistr}}
 \end{figure}
@@
 of the induced grammar $\mathcal{G}$. We might hope that our algorithm
-recovers the consituent logic gates that were used to build up the
+recovers the constituent logic gates that were used to build up the
 problems. Table~\ref{table:bool} shows a few of the top ranked
@@
 Figure~\ref{fig:boolhist}. In this second experiment, we remove the
-second kind of structure. In Figure~\ref{boolLearningCurves}, we show
+second kind of structure. In Figure~\ref{fig:boolLearningCurvesUniform}, we show
 10 runs of the algorithm with a frontier size of $2000$: note how
@@
 
-\begin{figure}
-\begin{subfigure}[Before]{0.45\linewidth}
-\includegraphics[width=.9\linewidth]{figures/booleancircuitsBars.pdf}
-\end{subfigure}
-\begin{subfigure}[Before]{0.45\linewidth}
-\includegraphics[width=.9\linewidth]{figures/boolLearningCurves.pdf}
-\end{subfigure}
-\caption{(a) Average learning curves as we vary the frontier size. (b)
-  Percent of tasks solved at iteration 15 as a function of frontier
-  size.}
+\begin{figure}[t]
+\begin{subfigure}[Before]{\linewidth}
+\includegraphics[width=.9\linewidth]{figures/booleancircuits3frontiers.pdf}
+\label{fig:boolLearningCurvesCircuits}
+\caption{Learning curves for various frontier sizes on task set of
+  Boolean functions drawn randomly from a distribution (see text for
+  details) of circuits built up of NOT, AND, OR gates.}
+\end{subfigure}
+
+\begin{subfigure}[Before]{\linewidth}
+\includegraphics[width=.9\linewidth]{figures/booleanfunctionsUniform.pdf}
+\label{fig:boolLearningCurvesUnifrom}
+\caption{10 learning curves for frontier size of 2000 on a task set
+  consisting of all Boolean functions of cardinality 3 or
+  smaller. Note how most of the trajectories get stuck in a local
+  minimum around $50\%$.}
+\end{subfigure}
 \label{fig:boolLearningCurves} 
-
+\caption{Learning curves from Boolean function learning experiment. }
 \end{figure}
-
-\begin{table}
-\scriptsize
-  \begin{tabular}{|l|l|l|}
-    function  & CL expressions & circuit  ~ \\ \hline
-        TRUE  & (S NAND) (S NAND I)                   & ~ \\ \hline
-        FALSE & (S B (S NAND)) (S NAND I)             & ~ \\ \hline
-        NOT   & S NAND I                              & \input{figures/NOT.tex} \\ \hline
-        AND   & ((C B NAND) (B (S NAND I)))           & ~ \\ 
-        OR    & ((B (C (B (S NAND) NAND))) (S NAND I) & ~ \\ 
-        XOR   & ~                                     & ~ \\ 
-        F     & S B (S NAND)                          & \input{figures/F.tex}
-    \end{tabular}
-\caption{Expressions in the induced library on sets of tasks
-  corresponding to boolean circuits synthesized from AND, OR, and NOT
-  gates, with boolean circuits of 1, 2, or 3 inputs.}
-\label{table:boolexpr}
-\end{table}
 
@@
 
-\subsection{Discussion}
+\section{Discussion}
 This work attempts to bring together ideas from various areas of
@@
 \bibliography{ijcai13}
-\nocite{*}
+%% \nocite{*}
 

### 2013-01-31 src/PI/Experiments/EX1.hs

@@ -1,6 +1,6 @@
 -- EX1
 -- Symbolic regression 10 run per frontier size experiment on EC2. 
-{-# Language ParallelListComp #-}
+ {-# Language ParallelListComp #-}
 
 module Experiments.EX1 where
 
@@ -145,33 +145,33 @@ libNames = ["R1"] -- , "R3", "R<=3"]
 frontierSizeSet sizes = join [replicate 1 i | i <- sizes]
                                           
 
-expSet sizes = concat [[mkSymRegExp (n ++ "_" ++ show f) l f | n <- libNames | l <- libSet]
-              | f <- frontierSizeSet sizes]
+-- expSet sizes = concat [[mkSymRegExp (n ++ "_" ++ show f) l f | n <- libNames | l <- libSet]
+--               | f <- frontierSizeSet sizes]
 
-expSetRand sizes = sequence $ concat [[mkSymRegExpM (n ++ "_" ++ show f) l f | n <- libNames | l <- libSet]
-              | f <- frontierSizeSet sizes]
+-- expSetRand sizes = sequence $ concat [[mkSymRegExpM (n ++ "_" ++ show f) l f | n <- libNames | l <- libSet]
+--               | f <- frontierSizeSet sizes]
 
-runExpSet sizes = sequence $ (map runExp (expSet sizes)) ++ (map runBruteForceExp (expSet sizes))
+-- runExpSet sizes = sequence $ (map runExp (expSet sizes)) ++ (map runBruteForceExp (expSet sizes))
 
-runExpSetRand sizes = do n <- newStdGen
-                         let (exps, a) = runRand (expSetRand sizes) n
-                         sequence $ map runExp exps
+-- runExpSetRand sizes = do n <- newStdGen
+--                          let (exps, a) = runRand (expSetRand sizes) n
+--                          sequence $ map runExp exps
 
--- expSymReg 
---     = Experiment {
---         expName = "EX1", 
---         expTaskSet = taskSet',
---         expEps = 0,
---         expPrior = expGrammar,
---         expInitLib = expGrammar,
---         expDepthBound = 3,
---         expNumBound = 10,
---         expReps=10
---       }
+-- -- expSymReg 
+-- --     = Experiment {
+-- --         expName = "EX1", 
+-- --         expTaskSet = taskSet',
+-- --         expEps = 0,
+-- --         expPrior = expGrammar,
+-- --         expInitLib = expGrammar,
[…]

### 2013-01-31 src/PI/Experiments/EX5.hs

@@ -64,7 +64,10 @@ taskSetNoLinear = [mkQuadSymRegTask i j k | i <- [1..9] , j <- [0..9], k <- [0..
 
 taskSetOnlyQuadratic = [mkQuadSymRegTask i j k | i <- [1..9] , j <- [0..9], k <- [0..9]]
 taskSetOnlyComplexQuadratic = [mkQuadSymRegTask i j k | i <- [1..9] , j <- [1..9], k <- [1..9]]
+
 taskSetUpToQuadratic = [mkQuadSymRegTask i j k | i <- [0..9] , j <- [0..9], k <- [0..9]]
+
+taskSetHardestQuadratic = [mkQuadSymRegTask i j k | i <- [2..9] , j <- [2..9], k <- [2..9]]
 
 
 
@@ -146,6 +149,7 @@ main = do
                               3 -> ("OnlyQuadratic", taskSetOnlyQuadratic)
                               4 -> ("OnlyComplexQuadratic", taskSetOnlyComplexQuadratic)
                               5 -> ("UpToQuadratic", taskSetUpToQuadratic)
+                              6 -> ("QuadraticCoef2AndGrater", taskSetHardestQuadratic)
                               _ -> error $ "No task name " ++ taskname
   let frontierSize = 10000
       name = taskname  ++  "_frontier"  ++ show frontierSize

### 2013-01-31 src/PI/python/symregmany.py

@@ -3,6 +3,7 @@ from glob import glob
 import pandas as pd
 import numpy as np
 import matplotlib.pyplot as plt
+import matplotlib
 
 import symreg as SR
 
@@ -61,6 +62,8 @@ for label in labels:
 fig.savefig("/Users/edechter/Dropbox/Projects/ProgramInduction/drafts/ijcai2013/figures/performanceVsFrontierSize2.pdf")
 
 
+matplotlib.rcParams.update({'font.size': 15})
+w,h = plt.figaspect(.5)
 fig_lc = plt.figure(figsize=(w,h))
 ax = fig_lc.add_subplot(111)
 ax.set_xlabel("iteration")

### 2013-02-01 drafts/ijcai2013/ijcai13.tex
Changes.

@@
-\documentclass{article}
+x\documentclass{article}
 
@@
 \usepackage{ dsfont }
+\usepackage[usenames,dvipsnames]{xcolor}
 \usepackage{tikz}
@@
 \usepackage{graphicx}
-\usepackage{subcaption}
+\usepackage[font=small]{subcaption}
+\usepackage[font=small]{caption}
 \usepackage{alltt}
@@
 \usetikzlibrary{arrows,automata, positioning, patterns,backgrounds}
+\usetikzlibrary{arrows,shapes.gates.logic.US,shapes.gates.logic.IEC,calc}
 
@@
 In this work, we try to see how far this idea of reusable function
-definition can get us. Our question is, Can an AI system discover the
-structure latent in the solutions to multiple related problems and
-thereby bootstrap effective exploration of the search space? Can
-discovering modular program components transform a problem from one in
-which search is intractable into a problem in which it becomes
-increasingly feasible with experience?
+definition can get us. We ask the following question: can an AI system
+discover the structure latent in the solutions to multiple related
+problems and thereby bootstrap effective exploration of the search
+space? Can discovering modular program components transform a problem
+from one in which search is intractable into a problem in which it
+becomes increasingly feasible with experience?
 
@@
 re-estimates the distribution over programs by weighting these new
-functions according to their frequencies.
+functions according to their frequencies. In using compression as a
+guide to the choice representation, we instantiate the idea that good
+representations are those which minimize the description length of
+typical elements in the domain.
 
@@
 ``Substantial Constructive Induction Using Layered Information
-Compression: Tractable Feature Formation in Search,'' -- which could
-have been the name of this paper -- did not use programs as its
-solution space representation. Within the Inductive Logic Programming
+Compression: Tractable Feature Formation in Search,'' contained many
+of these ideas. It did not, however, use programs as its solution
+space representation. Within the Inductive Logic Programming
 literature, researchers have explored ``Constructive Induction,''
@@
 
-\section{Problem statement and terminology}
+\section{Extracting modular concepts from functional programs}
+
 Our goal is to solve the following minimal feedback multi-task program
-induction problem: suppose we are given set of \emph{tasks}
+induction problem: suppose we are given set of \textbf{tasks}
 $T=\{t_k\}_{k=1}^K$ where each task is a function $t_i : \mathcal{L}
 \rightarrow \{0, 1\}$ and where $\mathcal{L}$ is the set of programs
-in our language.  We say that $ p \in \mathcal{L}$ \emph{solves} $t_i$
-if $ t_i(p) = 1$. Our goal is to solve as many of the tasks in $T$ as
-we can.
-
-We will use the following terminology. A \emph{frontier} will be the
-finite set of programs that the algorithm is considering at any one
-time. We will say that a task is \emph{hit} by the frontier if there
-is a program in the frontier that solves it.
-
-\section{Extracting modular concepts from functional programs}
-
-In order to solve our stated problem, we need to choose a solution
-representation language and a compression criterion, and we need to
-describe the frontier enumeration and compression procedures.
+in our language.  We say that $ p \in \mathcal{L}$ \textbf{solves}
+$t_i$ if $ t_i(p) = 1$. Our goal is to solve as many of the tasks in
+$T$ as we can.
+
+We propose the \textbf{E-C algorithm}, an iterative procedure to solve this
+problem. It will be useful in presenting the algorithm to define the
+following terminology: a \textbf{frontier} is the finite set of
+programs that the algorithm considers in any one iteration. The
+\textbf{frontier size} $N$ is the prespecified number of elements in
+the frontier. We will say that a task is \textbf{hit} by the frontier
+if there is a program in the frontier that solves it.
+
+The E-C algorithm maintains a distribution $\mathcal{D}$ over
+expressions in $\mathcal{L}$. At each iteration, it
+\begin{enumerate}
+\item \textbf{explores} the frontier -- enumerating the $N$ most probable expressions
+  from the current distribution $\mathcal{D}$ -- and
+\item \textbf{compresses} the hit tasks in the frontier to estimate a new
+  distribution.
+\end{enumerate}
+
+What remains is to specify a representation language $\mathcal{L}$, a
+form for the distribution $\mathcal{D}$, a compression criterion for a
+set of expressions in $\mathcal{L}$. We also need to describe
+procedures for the enumeration of the frontier, compression of the
+expressions, and re-estimations of $\mathcal{D}$.
 
@@
 
-\subsection{A stochastic grammar over programs.}
+\subsection{A stochastic grammar over programs}
 \label{sec:stochgrammar}
-We would like a way to describe how likely we believe a program to be
-given the probabilities we ascribe to its components. Recall that
-$\mathcal{L}$ is the set of all programs.  A stochastic grammar over
-programs represents a distribution over expressions in
-$\mathcal{L}$. We associate with each primitive combinator $c_i$ in
-$\mathcal{L}$ a probability $p_i$, $0~\leq~p_i~\leq~1, \sum_i~p_i~=~1
-$.
-
-The probability of using a primitive combinator somewhere in a program
-should depend on which other combinators could have been used in its
-place. This, in turn, depends on the type at that particular location
-in the program tree; we call this type the \emph{requesting} type. For
-example, suppose that we have the primitive $(+1)$ of type $Int
-\rightarrow Int$ and the primitive $NOT$ of type $Bool \rightarrow
-Bool$. If the requesting type is $\sigma \rightarrow Int$, then we can
-only use $(+1)$, but if the requesting type is $\sigma \rightarrow
-\sigma$ then we could use either $(+1)$ or $NOT$. Let $C = c_1, \dots,
-c_n$ be the set of primitive combinators in our grammar; we define
-$C_t$ for any type $t$ to be the set of combinators in $C$ that are
-consistent with $t$, that is, whose types \emph{unify} with t (we use
+
+We specify a distribution $\mathcal{D}$ over expressions in
+$\mathcal{L}$ as a stochastic
+grammar~\cite{feldman1969grammatical}. Many stochastic grammars have
+the desirable property that the probability of an expression is the
+product of the probabilities of its subcomponents. The distribution we
+specify will be a particularly simple verson of this, in which the
+probability of an expression is the product of the probability of its
+leaves.
+
+$\mathcal{D}$ will associate with each primitive combinator $c_i$ a
+prior probabilities $p_i$, $0~\leq~p_i~\leq~1, \sum_i~p_i~=~1 $. The
+probability of a leaf withr primitive combinator $c_i$ will depend on
+which other primitive combinators could have been chosen in its
+place. This set will depend on the type that was requested by the
+parent of this leaf when the leaf's primitive was chosen; we call this
+type the \emph{requesting} type of the leaf. For example, suppose that
+we have the primitive $(+1)$ of type $Int \rightarrow Int$ and the
+primitive $NOT$ of type $Bool \rightarrow Bool$. If the requesting
+type is $\sigma \rightarrow Int$, then we can only use $(+1)$, but if
+the requesting type is $\sigma \rightarrow \sigma$ then we could use
+either $(+1)$ or $NOT$. Let $C = c_1, \dots, c_n$ be the set of
+primitive combinators in our grammar; we define $C_t$ for any type
+$\tau$ to be the set of combinators in $C$ that are consistent with
+$\tau$, that is, whose types \emph{unify} with $\tau$ (we use
 unification according to the standard type theory
@@
 
-We define the probability of an expression $e$ from a grammar with
-primitive $C$ to be $p(e)~=~\prod_{n=1}^N p(c_n | t(c_n))$ where
-$p(c_n | t(c_n))$ is the probability of using primitive $c_n$ when the
-requesting type was $t(c_n)$. This definition means that the
-probability of an expression only depends on the probabilities
-assigned to the library primitives. We define these conditional
-probabilities as
+Given a distribution $\mathcal{D} = (p_1, \dots, p_n)$, we define the
+probability of an expression $e \in \mathcal{L}$ to be
+$p(e)~=~\prod_{n=1}^N p(c_n | \tau(c_n))$ where $p(c_n | \tau(c_n))$ is the
+probability of using primitive $c_n$ when the requesting type is
+$\tau(c_n)$. In turn, we define the conditional probabilities to be
 \[
-p(c_i | t(c_n)) =    \frac{1}{\epsilon}\frac{p_i}{\sum_{C_{t(c_n)}} p_j}.
+p(c_i | \tau(c_n)) \propto \frac{p_i}{\sum_{C_{\tau(c_n)}} p_j}.
 \]
-That is, $p(c_i | t(c_n)$ is the probability of sampling $c_i$ from
-the multinomial distribution defined by the $p_i$'s, conditioned on
-the requesting type, with the factor of $\frac{1}{\epsilon}$ added as
-a simple way to ensure that the sum over the probabilities of all
-expressions converges to a finite value (so that the distribution over
-expressions is proper). This is true whenever $\epsilon \geq
-4$\footnote{The probability mass $M_d$ of all expressions of depth
-  less than or equal to $d$ can be written with the recurrence
-  relation $M_d \leq M_{d-1}^2 + M_1$. This has an asymptotic fixed
-  point, as $d \rightarrow \infty$ if $x = x^2 + M_1$ has a solution,
-  which is true if $M_1 \leq 1/4$.}.
-
-To estimate the values $p_i$ from a set of expressions, we simply divide
-the number of times primitive combinator $c_i$ was used by the number
-of times it could have been used given the requesting types in the
-expression trees. Note that this is not a maximum likelihood estimator
-for this distribution; finding that would require solving a rather
-complex optimization problem because of each choice is conditioned on
-the requesting type. We choose our estimator merely for convenience,
-and find that our empirical results justify this simplicity. It is a
-matter for future research to determine to what extent a more accurate
-estimator would improve performance.
-
-\subsection{Best-first enumeration of most probable programs.}
+That is, $p(c_i | \tau(c_n)$ is proportional to the probability of
+sampling $c_i$ from the multinomial distribution defined by the
+$p_i$'s, conditioned on the requesting type, with the constant of
+proportionality chosen to ensure that the sum over the probabilities
+of all expressions converges to a finite value (so that the
+distribution over expressions is proper). For binary trees, this is
+true whenever the constant of proportionality is less than
+$\frac{1}{4}$\footnote{The probability mass $M_d$ of all expressions
+  of depth less than or equal to $d$ can be written with the
+  recurrence relation $M_d \leq M_{d-1}^2 + M_1$. This has an
+  asymptotic fixed point, as $d \rightarrow \infty$ if $x = x^2 + M_1$
+  has a solution, which is true if $M_1 \leq 1/4$.}.
+
+Now, suppose that we have a set of expressions $\mathcal{E}$, and we
+wish to estimate the maximum likelihood values of the $p_i$'s. If the
+choice of primitive at each leaf were not conditioned on the
+requesting type, we could just count all occurences of each primitive
+in our expression set, and this would be proportional to the maximum
+likelihood estimate. This is ML estimator for a multinomial
+distribution. However, the fact that we condition on the requesting
+type makes this a considerably more difficult optimization task. One
+of the simplest approximations we can make is to calculate the
+frequency with which each primitive combinator appears in
+$\mathcal{E}$ when the requesting type is such that it could have
+been chosen. This is a particularly straightforward calculuation, and
+we use it for its convenience, finding that our empirical results
+justify its use. Future inquiry will be needed to determine to what
+extent a more accurate estimator can improve these results.
+
+\subsection{Best-first enumeration of programs}
 In order to explore the frontier of most promising programs, we need a
@@
 best described by the following recursion: every program is either a
-primitive combinator or it a left child program applied to a right
+primitive combinator or it is a left child program applied to a right
 child program.
 
-This translates to the following AND/OR tree. Suppose we want a
+This can framed as the following AND/OR tree. Suppose we want a
 program of type $t$. The root of $\mathcal{G}$ is an OR node of type
@@
 to hit tasks such that this set is maximally compressible. This will
-promote the usage of reusable modular subtrees.
+promote reusable modular subtrees.
 
@@
 the squaring operation has been factored out as an abstraction. This
-is an example, then, of how this compressibility metric favors
+is an example of how this compressibility metric favors
 abstraction when multiple expressions are being considered at once.
@@
 Given a set of solution programs, we want to re-estimate our
-stochastic grammar. Our inspiration for doing this is based on
-the Neville-Manning algorithm for grammar-based
+stochastic grammar. Our inspiration for doing this is the
+Neville-Manning algorithm for grammar-based
 compression~\cite{nevill1997identifying}. The idea of that compression
@@
 
-In this section, we discuss the performance of our learning algorithm
-on a symbolic regression problem. In this problem, each task is a set
-of pairs $ (x, f(x))$ where $x \in (0, \dots, 9)$ and $f$ is a
-polynomial. The goal is to discover a program for each task that
-outputs the appropriate value $f(x)$ for each input x.
-
-%% The choice of an initial grammar for the algorithm will, of course,
-%% have a big effect on its performance, and so we would like to
-%% characterize this effect. We can divide primitives into roughly two
-%% sets; there are low-level primitives, those that operate on concrete
-%% types like Bool or Double, and these tend to be domain specific. For
-%% example, concrete primitives include $1, 2$, etc., and $f(x) = 3x$. The
-%% other class of primitives are abstract primitives, which can be
-%% characterized by having arguments whose types are primarily function
-%% types, like $(t \rightarrow s), (t \rightarrow s) \rightarrow \text{
-%%   Double },$ etc. Of course, we can construct operators that fall in
-%% between.
-
-%% The most abstract operators are those that Liang et. al refer to as
-%% ``routers''~\cite{liang10programs}. These operators take the place of
-%% $\lambda$ abstraction in the general lambda calculus: whereas
-%% $\lambda$ abstraction allows us to bind arguments to variables nested
-%% deep inside an expression, without them we need to have operators that
-%% ``route'' arguments to the appropriate functions in a nested
-%% expression.
-
-%% \begin{definition}[Router~\cite{liang10programs}] A \emph{router} is a combinator
-%% represented by a finite sequence of elements from $\{B, C, S\}$. The
-%% behavior of router $r$ is given by
-%% \[
-%% (r \, f \, g \, x_1 \dots x_{|r|}) = (f \, x_{i_1} \dots x_{i_n}) (g
-%% \, x_{j_1} \dots x_{j_n}),
-%% \]
-%% where $i_1 < \dots < i_n$ are indices $i$ such that $r_i \in \{C, S\}$
-%% and where $j_1 < \dots < j_n$ are indices $j$ such that $r_j \in \{B,
-%% S\}$. Let $\mathcal{R}_k$ be the set of all routers of length equal to
-%% $k$. Let $\mathcal{R}_{\leq k}$ be the set of all routers of length
-%% less than or equal to $k$.
-%% \end{definition}
-
-%% To analyze the behavior of our algorithm, we will vary the parameters
-%% of the problem along two axes while keeping the learning tasks the
-%% same. Along the first axis, we vary the initial library to which our
-%% learner has access with respect to the its abstractness (i.e., number
-%% of routers) and its concreteness (i.e. number of concrete
-%% primitives). Along the second axis we vary the size of the search
-%% frontier.
+We first illustrate the performance of our learning algorithm on a
+symbolic regression problem. In task in this problem, we want to find
+an algebraic function, symbolically specified, that maps a set of
+input values to output values. We choose this problem because it has a
+long history in AI, particularly in the genetic programming
+literature~\cite{DBLP:books/daglib/0070933}.
+
+In the formulation we consider, each task $t$ corresponds to a
+polynomial $f$, and an expression $e$ solves $t$ if ${ (e\;
+  i)=f(i) }$ for all ${ i=0,\dots,9 }$.
 
@@
 
-\begin{table}
-\tiny
-\begin{tabular}[width=\linewidth]{|p{1cm}|p{2cm}|p{3.5cm}|}
-\hline \\
-  function  & CL expression & circuit  ~ \\ \hline
-¬†¬†¬†¬†¬†¬†¬†¬†TRUE¬† & (S NAND) (S NAND I)¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬† & ~
-
-%\begin{top}
-\begin{circuitikz}[transform shape] 
-\draw
-(1, 0) node[nand port, scale=.5] (nand1) {}
-(2,.3) node[nand port, scale=.5] (nand2) {}
-(0, 0) node[anchor=east]  (x) {x}
-(x) -- (.3, 0) -| (nand1.in 1)
-(x) -- (.3, 0) -|  (nand1.in 2)
-(x) |- (nand2.in 1)
-(nand1.out) |- (nand2.in 2)
-;
-\end{circuitikz}
-%\end{top}
-\\ \hline
-¬†¬†¬†¬†¬†¬†¬†¬†FALSE & (S B (S NAND)) (S NAND I)¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬† &  ~
+\include{./figures/tablealone}
+%% \begin{table}
+%% \tiny
+%% \begin{tabular}[width=\linewidth]{|p{1cm}|p{2cm}|p{3.5cm}|}
+%% \hline function  & CL expression & circuit  ~ \\ \hline\hline
+%% ¬†¬†¬†¬†¬†¬†¬†¬†NOT¬†¬† & S NAND I¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬† &  ~
 
 %% \begin{circuitikz}[transform shape]
-
+%% \draw
+%% (1.2,0) node[nand port, scale=.3] (nand1) {}
+%% (0, 0) node (x) {x}
+%% (x) -| (nand1.in 1)
+%% (x) -| (nand1.in 2)
 %% ;
@@
 
-\\ \hline
-¬†¬†¬†¬†¬†¬†¬†¬†NOT¬†¬† & S NAND I¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬† &  ~
-
-\begin{circuitikz}[transform shape]
-\draw
-(1.2,0) node[nand port, scale=.5] (nand1) {}
-(0, 0) node (x) {x}
-(x) -| (nand1.in 1)
-(x) -| (nand1.in 2)
-;
-\end{circuitikz}
-
-\\ \hline
-¬†¬†¬†¬†¬†¬†¬†¬†AND¬†¬† & ((C B NAND) (B (S NAND I)))¬†¬†¬†¬†¬†¬†¬†¬†¬†¬† & ~
-
-\begin{circuitikz}
-\draw
-(1.0,0) node[nand port, scale=.5] (nand1) {}
-(2.0,0) node[nand port, scale=.5] (nand2) {}
-% (0, 0) node (x) {x}
-% (0, 0) node (y) {y}
-node[above left=.2cm and .7cm  of nand1] (x) {x} %-- (nand1.in 1)
-node[below left=.2cm and .7cm  of nand1] (y) {y} %-- (nand1.in 2)
-(x) |- (nand1.in 1)
-(y) |- (nand1.in 2)
-(nand1.out) |- (nand2.in 1)
-(nand1.out) |- (nand2.in 2)
-;
-\end{circuitikz}
- \\  \hline
-¬†¬†¬†¬†¬†¬†¬†¬†OR¬†¬†¬† & ((B (C (B (S NAND) NAND))) (S NAND I) & ~
-
-\begin{circuitikz}
-\draw
-(0,0) node (x) {x}
-(0,.7) node (y) {y}
-
-node[nand port, scale=.5, right=1cm of x] (nand1) {}
-node[nand port, scale=.5, right=1cm of y] (nand2) {}
-node[nand port, scale=.5, below right=.3cm and 1cm of nand2] (nand3) {}
-(x) -| (nand1.in 1)
-(x) -| (nand1.in 2)
-(y) -| (nand2.in 1)
-(y) -| (nand2.in 2)
-(nand1.out) |- (nand3.in 1)
-(nand2.out) |- (nand3.in 2)
-;
-\end{circuitikz}
- \\  \hline
-¬†¬†¬†¬†¬†¬†¬†¬†XOR¬†¬† & ((S (B C (((B (C B NAND)) S) (((B (C B NAND)) S) (B C ((B (B NAND)) NAND)))))) I)¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬† & ~
-
-\begin{circuitikz}
-\draw
-node[nand port, scale=.5] (nand1) {}
-node[nand port, scale=.5, above right=.3cm and 1cm of nand1] (nand2) {} 
-node[nand port, scale=.5, below right=.3cm and 1cm of nand1] (nand3) {} 
-node[nand port, scale=.5, right=2cm of nand1] (nand4) {} 
-node[left=1.2cm of nand2.in 1] (x) {x}
-node[left=1.2cm of nand3.in 2] (y) {y}
-node[right=.2cm of nand4] (out) {}
-(x) -| (nand1.in 1)
-(x) -| (nand2.in 1)
-(y) -| (nand1.in 2)
-(y) -| (nand3.in 2)
-(nand1.out) -| (nand2.in 2)
-(nand1.out) -| (nand3.in 1)
-(nand2.out) |- (nand4.in 1)
-(nand3.out) |- (nand4.in 2)
-(nand4.out) -- (out)
-;
-\end{circuitikz}
-
- \\  \hline
-¬†¬†¬†¬†¬†¬†¬†¬†F¬†¬†¬†  & S B (S NAND)¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬† & ~
+%% \\ \hline
+%% ¬†¬†¬†¬†¬†¬†¬†¬†AND¬†¬† & ((C B NAND) (B (S NAND I)))¬†¬†¬†¬†¬†¬†¬†¬†¬†¬† & ~
+
 %% \begin{circuitikz}
 %% \draw
-%% (0, 0) node[left=.25) (x) {x}
-
+%% (1.0,0) node[nand port, scale=.3] (nand1) {}
+%% (2.0,0) node[nand port, scale=.3] (nand2) {}
+%% % (0, 0) node (x) {x}
+%% % (0, 0) node (y) {y}
+%% node[above left=.2cm and .7cm  of nand1] (x) {x} %-- (nand1.in 1)
+%% node[below left=.2cm and .7cm  of nand1] (y) {y} %-- (nand1.in 2)
+%% (x) |- (nand1.in 1)
+%% (y) |- (nand1.in 2)
+%% (nand1.out) |- (nand2.in 1)
+%% (nand1.out) |- (nand2.in 2)
 %% ;
 %% \end{circuitikz}
-
-\\ 
-\hline
-
-\end{tabular}
-\caption{Expressions in the induced library on sets of tasks
-  corresponding to boolean circuits synthesized from AND, OR, and NOT
-  gates, with boolean circuits of 1, 2, or 3 inputs.}
-\label{table:boolexpr}
-
-\end{table}
+%%  \\  \hline
+%% ¬†¬†¬†¬†¬†¬†¬†¬†OR¬†¬†¬† & ((B (C (B (S NAND) NAND))) (S NAND I) & ~
+
+%% \begin{circuitikz}
+%% \draw
+%% (0,0) node (x) {x}
+%% (0,.7) node (y) {y}
+
+%% node[nand port, scale=.3, right=1cm of x] (nand1) {}
+%% node[nand port, scale=.3, right=1cm of y] (nand2) {}
+%% node[nand port, scale=.3, below right=.3cm and 1cm of nand2] (nand3) {}
+%% (x) -| (nand1.in 1)
+%% (x) -| (nand1.in 2)
+%% (y) -| (nand2.in 1)
+%% (y) -| (nand2.in 2)
+%% (nand1.out) |- (nand3.in 1)
+%% (nand2.out) |- (nand3.in 2)
+%% ;
+%% \end{circuitikz}
+%%  \\  \hline\hline
+
+%% ¬†¬†¬†¬†¬†¬†¬†¬†TRUE¬† & (S NAND) (S NAND I)¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬† & ~
+
+%% %\begin{top}
+%% \begin{circuitikz}[transform shape] 
+%% \draw
+%% (1, 0) node[nand port, scale=.3] (nand1) {}
+%% (2,.3) node[nand port, scale=.3] (nand2) {}
+%% (0, 0) node[anchor=east]  (x) {x}
+%% (x) -- (.3, 0) -| (nand1.in 1)
+%% (x) -- (.3, 0) -|  (nand1.in 2)
+%% (x) |- (nand2.in 1)
+%% (nand1.out) |- (nand2.in 2)
+%% ;
+%% \end{circuitikz}
+%% %\end{top}
+%% \\ \hline
+%% ¬†¬†¬†¬†¬†¬†¬†¬†FALSE & (S B (S NAND)) (S NAND I)¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬† &  ~
+
+%% \begin{circuitikz}[transform shape] 
+%% \draw
+%% (1, 0) node[nand port, scale=.3] (nand1) {}
+%% (2,.3) node[nand port, scale=.3] (nand2) {}
+%% node[nand port, scale=.3, right=1cm of nand2] (nand3) {} 
+%% (0, 0) node[anchor=east]  (x) {x}
+%% (x) -- (.3, 0) -| (nand1.in 1)
+%% (x) -- (.3, 0) -|  (nand1.in 2)
+%% (x) |- (nand2.in 1)
+%% (nand1.out) |- (nand2.in 2)
+%% (nand2.out) -| (nand3.in 1)
+%% (nand2.out) -| (nand3.in 2)
+%% ;
+%% \end{circuitikz}
+%% \\ \hline
+%% ¬†¬†¬†¬†¬†¬†¬†¬†XOR¬†¬† & ((S (B C (((B (C B NAND)) S) (((B (C B NAND)) S) (B C ((B (B NAND)) NAND)))))) I)¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬† & ~
+
+%% \begin{circuitikz}
+%% \draw
+%% node[nand port, scale=.3] (nand1) {}
+%% node[nand port, scale=.3, above right=.3cm and 1cm of nand1] (nand2) {} 
+%% node[nand port, scale=.3, below right=.3cm and 1cm of nand1] (nand3) {} 
+%% node[nand port, scale=.3, right=2cm of nand1] (nand4) {} 
+%% node[left=1.2cm of nand2.in 1] (x) {x}
+%% node[left=1.2cm of nand3.in 2] (y) {y}
+%% node[right=.2cm of nand4] (out) {}
+%% (x) -| (nand1.in 1)
+%% (x) -| (nand2.in 1)
+%% (y) -| (nand1.in 2)
+%% (y) -| (nand3.in 2)
+%% (nand1.out) -| (nand2.in 2)
+%% (nand1.out) -| (nand3.in 1)
+%% (nand2.out) |- (nand4.in 1)
+%% (nand3.out) |- (nand4.in 2)
+%% (nand4.out) -- (out)
+%% ;
+%% \end{circuitikz}
+
+%%  \\  \hline\hline
+%% ¬†¬†¬†¬†¬†¬†¬†¬†$E_1$¬†¬†¬†  & S B (S NAND)¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬† & ~
+
+%% \begin{circuitikz}
+%% \draw
+%% node (x) {x}
+%% node[nand port, scale=.3, right=1.5cm of x] (nand) {} 
+%% node[draw=blue!75, fill=blue!20, thick, 
+%%   above right=.3cm and .2cm of x, anchor=west] (f1) {$g(x)$} 
+%% node[draw=blue!75, fill=blue!20, thick, right=.5cm of nand] (f2) {$g(x)$}
+%% node[right=.1cm of f2] (out) {}
+%% (x) -| (f1.west)
+%% (x) -| (nand.in 2)
+%% (f1) -| (nand.in 1)
+%% (nand.out) -- (f2)
+%% (f2) -- (out);
+%% \end{circuitikz}
+
+%% \\ \hline
+%% $E_2$¬†¬†¬†  & B (C B NAND) S & ~
+
+%% \begin{circuitikz}
+%% \draw
+%% node (x) {x} 
+%% node[nand port, scale=.3, right=1.5cm of x] (nand) {} 
+%% node[draw=blue!75, fill=blue!20, thick, 
+%%   above right=.2cm and .3cm of x, anchor=west] (g) {$g(x)$} 
+%% node[draw=blue!75, fill=blue!20, thick, 
+%%   below right=.2cm and .3cm of x, anchor=west] (f) {$f(x)$} 
+%% (x) -- [right=.25] (f.west)
+%% (x) |- (g.west)
+%% (f.east) -| (nand.in 1)
+%% (g.east) -| (nand.in 2)
+%% (nand.out) -- (out)
+%% (x)
+%% ;
+%% \end{circuitikz}
+
+%% \\ \hline
+
+%% \end{tabular}
+%% \caption{Expressions in the induced library on sets of tasks
+%%   corresponding to boolean circuits synthesized from AND, OR, and NOT
+%%   gates, with boolean circuits of 1, 2, or 3 inputs.}
+%% \label{table:boolexpr}
+
+%% \end{table}
 

### 2013-02-01 drafts/ijcai2013/ijcai13.tex
Changes

@@
-x\documentclass{article}
+\documentclass{article}
 
@@
 local or topological structure in the search space. For example,
-convex optimization and gradient based search methods~\cite{} take
+convex optimization and gradient based search methods take
 advantage of local smoothness to move in good directions. Local search
 algorithms move from one candidate solution to a neighboring
-one\cite{}. Heuristic search techniques take advantage of topological
-structure to prune off large parts of the solution space~\cite{}.
+one. Heuristic search techniques take advantage of topological
+structure to prune off large parts of the solution space.
 
@@
 not come with such a structured search space. This is particularly
-true of problems, as these are, whose solutions can be naturally
-framed as programs. Here we ask how to extract structure from the
-space of programs by attempting to solve multiple related problems in
-a given domain at once.
+true of problems, such as those mentioned, whose solutions can be
+naturally framed as programs. Here we ask how to extract structure
+from the space of programs by attempting to solve multiple related
+problems in a given domain at once.
 
 We take as inspiration the programmer who -- having reused several
-motifs, programming patterns, or subroutines to solve different but
-similar simple tasks -- creates a function library to help manage the
-complexity of larger projects. Our algorithm builds a library
-of reusable program components and a distribution over these,
-effectively learning from simple tasks a search space structure which
-enables it to solve more complex tasks.
+subroutines to solve different but similar simple tasks -- creates a
+function library to help manage the complexity of larger projects. The
+EC algorithm, which we introduce here, builds a library of reusable
+program components and places a distribution over these, effectively
+learning from simple tasks a search space structure that enables it
+to solve more complex tasks.
 
@@
 
-Concretely, our algorithm works as follows. It begins with a small
+Concretely, the EC algorithm works as follows. It begins with a small
 library of program primitives and a distribution over programs
@@
 functions according to their frequencies. In using compression as a
-guide to the choice representation, we instantiate the idea that good
-representations are those which minimize the description length of
-typical elements in the domain.
+guide to the choice of representation, we instantiate the idea that
+good representations are those which minimize the description length
+of typical elements in the domain.
 
@@
 ``Substantial Constructive Induction Using Layered Information
-Compression: Tractable Feature Formation in Search,'' contained many
-of these ideas. It did not, however, use programs as its solution
-space representation. Within the Inductive Logic Programming
-literature, researchers have explored ``Constructive Induction,''
-generating new and reusable logic programming
-terms~\cite{DBLP:conf/ijcai/Muggleton87}. We believe that applying
-constructive induction to a compositional functional programming
-paradigm will provide even greater benefits.
-
-Dicovering useful subroutines has been explored in Genetic
-Programming as
-well~\cite{DBLP:conf/foga/Koza92}~\cite{DBLP:conf/ices/KozaBAK96}.
-
-The importance of a curriculum has also been explored previously,
-though generally the curriculum has been explicitly given rather than
-discovered. Quinlan and Cameron-Jones~\cite{Quinlan93foil:a} showed
-that a logic program learning algorithm could learn sequentially more
-difficult list programs if the algorithm was provided with its own
-solutions to previously solved simple problems. More recently, several
-researchers have observed that presenting problems to machine learning
-algorithms in order of difficulty results in better
-performance~\cite{Krueger2009380}~\cite{DBLP:conf/icml/BengioLCW09}.
-
-Our use of a stochastic grammar over combinatory logic is based on
-Liang et al.~\cite{liang10programs}. In that work, a non-parametric
-distribution was placed over combinatory logic expressions, and an
-MCMC algorithm was developed to perform joint inference over these
-expressions and the underlying stochastic grammar. Our approach is
-inspired by this generative model, but avoids the difficulties
-associated with a stochastic local search over programs (the same
-difficulties faced by genetic programming), especially when the
-likelihood function (or the fitness function) are not of much help.
+Compression: Tractable Feature Formation in Search,'' presented within
+a classical search paradigm the idea that compression provides a
+heuristic for constructing good representations. Within the Inductive
+Logic Programming (ILP) literature, researchers have explored ``Constructive
+Induction,'' generating new and reusable logic programming
+terms~\cite{DBLP:conf/ijcai/Muggleton87}. We apply these ideas to
+learning functional programs, and believe that the modularity and
+compositionality afforded by this representation is necessary for
+these ideas to be successful.
+
+Genetic programming
+(GP)~\cite{DBLP:conf/foga/Koza92}~\cite{DBLP:conf/ices/KozaBAK96} has
+tackled the problem of program learning and has explored the notion
+that reusable functions are helpful. This work relies, however, on
+stochastic local search over the space of programs, and automatically
+discovering useful subroutines is seen as a helpful heuristic. More
+recently, Liang et al.~\cite{liang10programs} used a stochastic
+grammar over combinatory logic to induce shared structure in
+multi-task program induction problems. We find this latter
+representation compelling and use it here, but we see both this and
+genetic programming as tackling the question of how to benefit from
+shared structure in situations where local search might be successful
+on its own. Such problems rely on domain knowledge -- whether in the
+form of a fitness function or a likelihood function -- to provide
+local structure to the search space.
+
+But thes problems we want AI systems to solve often lack this kind of
+built in structure. These are the kinds of problems more often
+associated with ILP and classical AI than with GP and machine
+learning. 
+\\
 
@@
 
-The remainder of this paper will include two major sections.  We will
-first describe the representation language we chose and the details of
-the exploration and compression steps of the algorithm. We will then
-present results from two experiments involving the domains of
-symbolic regression and boolean circuit learning.
+The remainder of this paper will include two major sections. In the
+first, we describe in detail the EC algorithm and design choices we
+make in implementing it. We will then present results from two
+experiments in the domains of symbolic regression and Boolean
+functioin learning.
 
@@
 
-\section{Extracting modular concepts from functional programs}
+\section{Extracting concepts from programs}
 
 Our goal is to solve the following minimal feedback multi-task program
-induction problem: suppose we are given set of \textbf{tasks}
-$T=\{t_k\}_{k=1}^K$ where each task is a function $t_i : \mathcal{L}
-\rightarrow \{0, 1\}$ and where $\mathcal{L}$ is the set of programs
-in our language.  We say that $ p \in \mathcal{L}$ \textbf{solves}
-$t_i$ if $ t_i(p) = 1$. Our goal is to solve as many of the tasks in
-$T$ as we can.
+induction problem: suppose we are given a set of \textbf{tasks}
+$T=\{t_k\}_{k=1}^K$ where each task is a function $t_k : \mathcal{L}
+\rightarrow \{0, 1\}$ and where $\mathcal{L}$ is the set of
+expressions in our language.  We say that expression $ e \in
+\mathcal{L}$ \textbf{solves} $t_k$ if $ t_k(p) = 1$. Our goal is to
+solve as many of the tasks in $T$ as we can.
 
@@
 following terminology: a \textbf{frontier} is the finite set of
-programs that the algorithm considers in any one iteration. The
+expressions that the algorithm considers in any one iteration. The
 \textbf{frontier size} $N$ is the prespecified number of elements in
 the frontier. We will say that a task is \textbf{hit} by the frontier
-if there is a program in the frontier that solves it.
+if there is a expression in the frontier that solves it.
 
@@
 
-What remains is to specify a representation language $\mathcal{L}$, a
-form for the distribution $\mathcal{D}$, a compression criterion for a
-set of expressions in $\mathcal{L}$. We also need to describe
+Next, we specify a representation language $\mathcal{L}$, a
+form for the distribution $\mathcal{D}$, and a compression criterion for a
+set of expressions in $\mathcal{L}$. We also describe
 procedures for the enumeration of the frontier, compression of the
-expressions, and re-estimations of $\mathcal{D}$.
+expressions, and re-estimation of $\mathcal{D}$.
 
@@
 introduces a \emph{basis} of several primitive \emph{combinators} such
-that any function in the lambda calculus can be alternatively written
-as applications of the basis combinators. Some common basis
-combinators are defined as follows:
+that any function can be written as applications of the basis
+combinators. Some common basis combinators are defined as follows:
 \begin{align}
-I\; x &\rightarrow  x \\
-S\; f\; g\; x\; &\rightarrow (f\; x)\; (g\; x) \\ 
-B\; f\; g\; x\; &\rightarrow (f\; x)\; g \\ 
-C\; f\; g\; x\; &\rightarrow f\; (g\; x) \\ 
+\mathbf{I}\; x &\rightarrow  x & \text{ (identity) }\\
+\mathbf{S}\; f\; g\; x\; &\rightarrow (f\; x)\; (g\; x)\\ 
+\mathbf{C}\; f\; g\; x\; &\rightarrow (f\; x)\; g  \\ 
+\mathbf{B}\; f\; g\; x\; &\rightarrow f\; (g\; x) & \text{ (composition) }
 \end{align}
@@
 variety of primitives and, together with the basis combinators, we
-call these the primitive combinators. Note that by default we consider
+call these the \textbf{primitive combinators}. Note that by default we consider
 all primitives to be in their ``curried'' form (e.g. a function like
@@
 returning the function $(+\;x)$ and then being applied to $y$,
-returning the object $(+\;x)\;y$).
-
-The basis combinators can themselves been expressed in the lambda
+returning $(+\,x\,y)$\;).
+
+The basis combinators can themselves be expressed in the lambda
 calculus (with definitions following directly from the equations
 above). The lambda calculus has two basic operations -- application
-and abstraction -- but in using the combinatory logic we sequester
-uses of the abstraction operation inside the combinators. In doing so,
-we have replaced the variable binding of the $\lambda$ operator with
-the variable \emph{routing} of these basis combinators. Our
-representation thus becomes variable-free. See Liang et
-al.~\cite{liang10programs} for a more detailed discussion of this
-routing interpretation.
-
-Using combinatory logic is very convenient for program synthesis:
-since every expression is the application of one expression to another
--- with this recursion bottoming out at the primitive combinators --
-each program is a binary tree. Most importantly, any subtree is itself
-a well-formed expression; this is not the case in the lambda calculus,
-since abstraction introduces long range dependencies between the
-$\lambda$ operator and the variables to which that operator refers. In
-the lambda calculus, then, a subtree might have free variables, not
-bound to any enclosing $\lambda$.
+and abstraction (using the $\lambda$ operator -- but in using the
+combinatory logic we sequester uses of the abstraction operation
+inside the combinators. In doing so, we have replaced the variable
+binding of the $\lambda$ operator with the variable \emph{routing} of
+these basis combinators. Our representation thus becomes
+variable-free. See Liang et al.~\cite{liang10programs} for a more
+detailed discussion of this routing interpretation.
+
+Using combinatory logic is very convenient for program
+synthesis~\cite{briggs2006functional}: since every expression is the
+application of one expression to another -- with this recursion
+bottoming out at the primitive combinators -- each program is a binary
+tree. Most importantly, any subtree is itself a well-formed
+expression; this is not the case in the lambda calculus, since
+abstraction introduces long range dependencies between the $\lambda$
+operator and the variables to which that operator refers. In the
+lambda calculus, then, a subtree might have free variables, not bound
+to any enclosing $\lambda$.
 
@@
 function $\lambda x. * x\; x $. Using two of the basis combinators
-above, we can write the squaring function in combinatory logic as $S *
+above, we can write the squaring function in combinatory logic as $S\; *\;
 I$ (taking $*$ as a primitive), where application associates to the
@@
 value $x$, the action of the combinator is defined as $S \;*\; I\;
-x~\rightarrow~(*\; x)\; (I\; x)~\rightarrow(*\; x)\; x$.
+x~\rightarrow~(*\; x)\; (I\; x)~\rightarrow *\; x \; x$.
 
@@
 type primitive (e.g. reals, integers, booleans, etc.), a type variable
-(e.g. $\sigma$), or a function type $ t_1 \rightarrow t_2$ of
-functions from source type $t_1$ to target type $t_2$. Any combinator
+(e.g. $\sigma$), or a function type $ \tau_1 \rightarrow \tau_2$ of
+functions from source type $\tau_1$ to target type $\tau_2$. Any combinator
 can be represented as a binary tree~(Figure~\ref{fig:clbintree}) whose
@@
 %%  \centering
-  \begin{subfigure}[]{0.45\linewidth}
+  \begin{subfigure}[Before]{0.45\linewidth}
     \begin{tikzpicture}[scale=.8,
@@
     \end{tikzpicture}
-    \caption{}
-    \label{fig:clbintree}
-  \end{subfigure}
-  \begin{subfigure}[R]{0.5\linewidth}
+    \caption{    \label{fig:clbintree}}
+  \end{subfigure}
+  \begin{subfigure}[After]{0.5\linewidth}
     \begin{tikzpicture}[scale=.7, transform shape, 
@@
     \end{tikzpicture}
-    \caption{AND/OR expression search space. }
-    \label{fig:andor}
-    \end{subfigure}
-  \caption{(a) The typed combinator $S * I$ ($f(x)=x^2$) represented as
+    \caption{  \label{fig:andor}}
+    \end{subfigure}
+  \caption{(a) The typed combinator $S * I$ ($f(x)=x^2$ over real values ($\mathbb{R}$)) represented as a
     binary tree. (b) The space of all typed combinators represented as an
-    infinitely deep AND/OR tree. Any specific combinator is a
-    sub-binary tree of this AND/OR tree whose leaves consist of only
-    the $f_i$ nodes (i.e., the primitive combinators of the grammar).}
+    infinitely deep AND/OR tree.}
 
@@
 product of the probabilities of its subcomponents. The distribution we
-specify will be a particularly simple verson of this, in which the
+specify will be a simple verson of this, in which the
 probability of an expression is the product of the probability of its
@@
 
-$\mathcal{D}$ will associate with each primitive combinator $c_i$ a
-prior probabilities $p_i$, $0~\leq~p_i~\leq~1, \sum_i~p_i~=~1 $. The
-probability of a leaf withr primitive combinator $c_i$ will depend on
-which other primitive combinators could have been chosen in its
-place. This set will depend on the type that was requested by the
-parent of this leaf when the leaf's primitive was chosen; we call this
-type the \emph{requesting} type of the leaf. For example, suppose that
-we have the primitive $(+1)$ of type $Int \rightarrow Int$ and the
-primitive $NOT$ of type $Bool \rightarrow Bool$. If the requesting
-type is $\sigma \rightarrow Int$, then we can only use $(+1)$, but if
-the requesting type is $\sigma \rightarrow \sigma$ then we could use
-either $(+1)$ or $NOT$. Let $C = c_1, \dots, c_n$ be the set of
-primitive combinators in our grammar; we define $C_t$ for any type
-$\tau$ to be the set of combinators in $C$ that are consistent with
-$\tau$, that is, whose types \emph{unify} with $\tau$ (we use
-unification according to the standard type theory
-presentation~\ref{Pierce_2002}).
-
-Given a distribution $\mathcal{D} = (p_1, \dots, p_n)$, we define the
-probability of an expression $e \in \mathcal{L}$ to be
-$p(e)~=~\prod_{n=1}^N p(c_n | \tau(c_n))$ where $p(c_n | \tau(c_n))$ is the
-probability of using primitive $c_n$ when the requesting type is
-$\tau(c_n)$. In turn, we define the conditional probabilities to be
-\[
-p(c_i | \tau(c_n)) \propto \frac{p_i}{\sum_{C_{\tau(c_n)}} p_j}.
-\]
-That is, $p(c_i | \tau(c_n)$ is proportional to the probability of
-sampling $c_i$ from the multinomial distribution defined by the
-$p_i$'s, conditioned on the requesting type, with the constant of
-proportionality chosen to ensure that the sum over the probabilities
-of all expressions converges to a finite value (so that the
-distribution over expressions is proper). For binary trees, this is
-true whenever the constant of proportionality is less than
-$\frac{1}{4}$\footnote{The probability mass $M_d$ of all expressions
-  of depth less than or equal to $d$ can be written with the
-  recurrence relation $M_d \leq M_{d-1}^2 + M_1$. This has an
+Let ${C = c_1, \dots, c_N }$ be the primitive combinators in
+$\mathcal{L}$. ${ \mathcal{D}=p_1, \dots, p_N }$ will associate with
+each primitive combinator $c_i$ a prior probability $p_i$,
+$0~\leq~p_i~\leq~1, \sum_i~p_i~=~1 $. The probability of a leaf with
+primitive combinator $c_i$ will depend on which other primitive
+combinators could have been chosen in its place. This set will depend
+on the type that was requested by the parent of this leaf when the
+leaf's primitive was chosen; we call this type the \emph{requesting}
+type of the leaf. For example, suppose that we have the primitive
+$(+1)$ of type $Int \rightarrow Int$ and the primitive $NOT$ of type
+$Bool \rightarrow Bool$. If the requesting type is $\sigma \rightarrow
+Int$, then we can only use $(+1)$, but if the requesting type is
+$\sigma \rightarrow \sigma$ then we could use either $(+1)$ or
+$NOT$. We define $C_\tau$ for any type $\tau$ to be the set of
+primitive combinators in $C$ that are consistent with $\tau$, that is,
+whose types \emph{unify} with $\tau$ (we use unification according to
+the standard type theory presentation~\cite{Pierce_2002}).
+
+Let $C_e$ be the set of primitive combinators in the leaves of
+expressions $e$. We define the probability of an expression ${e \in
+  \mathcal{L}}$ to be ${ p(e)~=~\prod_{c \in C_e} p(c | \tau(c)) }$
+where $p(c | \tau(c))$ is the probability of using primitive $c$ when
+the requesting type is $\tau(c)$. In turn, we define the conditional
+probability for each combinator $c_n$ to be $p(c_n| \tau(c_n)) \propto
+\frac{p}{\sum_{c_j \in C_{\tau(c_n)}} p_j}$.  That is, $p(c_n | \tau(c_n)$ is
+proportional to the probability of sampling $c_n$ from the multinomial
+distribution defined by the $p_n$'s, conditioned on the requesting
+type, with the constant of proportionality chosen to ensure that the
+sum over the probabilities of all expressions converges to a finite
+value (so that the distribution over expressions is proper). For
+binary trees, this is true whenever the constant of proportionality is
+less than~$\frac{1}{4}$\footnote{The probability mass $M_d$ of all
+  expressions of depth less than or equal to $d$ can be written with
+  the recurrence relation $M_d \leq M_{d-1}^2 + M_1$. This has an
   asymptotic fixed point, as $d \rightarrow \infty$ if $x = x^2 + M_1$
-  has a solution, which is true if $M_1 \leq 1/4$.}.
-
-Now, suppose that we have a set of expressions $\mathcal{E}$, and we
-wish to estimate the maximum likelihood values of the $p_i$'s. If the
+  has a solution, which is true if $M_1~\leq~1/4$.}.
+
+Now suppose that we have a set of expressions $\mathcal{E}$, and we
+wish to estimate the maximum likelihood values of the $p_n$'s. If the
 choice of primitive at each leaf were not conditioned on the
@@
 in our expression set, and this would be proportional to the maximum
-likelihood estimate. This is ML estimator for a multinomial
+likelihood estimate. This is the ML estimator for a multinomial
 distribution. However, the fact that we condition on the requesting
@@
 $\mathcal{E}$ when the requesting type is such that it could have
-been chosen. This is a particularly straightforward calculuation, and
+been chosen. This is a straightforward calculuation, and
 we use it for its convenience, finding that our empirical results
@@
 problem as a best-first exploration of an AND/OR tree
-(Figure~\ref{fig:andor})~\cite{nilsson1982principles}~\cite{DBLP:journals/cacm/Hall73},
-which we denote $\mathcal{G}$. In words, the enumeration procedure is
-best described by the following recursion: every program is either a
-primitive combinator or it is a left child program applied to a right
-child program.
-
-This can framed as the following AND/OR tree. Suppose we want a
-program of type $t$. The root of $\mathcal{G}$ is an OR node of type
-$t$. Its children are elements in $C_t$ (defined in the previous
-section) and one AND node of type $t$ with two children. Each child of
+(Figure~\ref{fig:andor})~\cite{nilsson1982principles}~\cite{DBLP:journals/cacm/Hall73}. In
+words, the enumeration procedure is best described by the following
+recursion: every program is either a primitive combinator or it is a
+left child program applied to a right child program.
+
+This can framed as the following AND/OR tree $\mathcal{G}$. Suppose we
+want a program of type $\tau$. The root of $\mathcal{G}$ is an OR node of
+type $\tau$. Its children are elements in $C_\tau$ (defined in the previous
+section) and one AND node of type $\tau$ with two children. Each child of
 an AND node has the same structure as the root node, with modified
-types: type of its left child is $\tau \rightarrow t$ (where $\tau$ is
+types: the type of its left child is $\sigma \rightarrow \tau$ (where $\sigma$ is
 a fresh type variable). Since the type of the right child is
@@
 $\mathcal{G}$ satisfying the following properties: the root of
-$\mathcal{H}$ is the root $\mathcal{G}$ and any OR node in
+$\mathcal{H}$ is the root of $\mathcal{G}$ and any OR node in
 $\mathcal{H}$ has at most one child. $\mathcal{H}$ is a complete
-solution (and thus a complete program) if it is a partial solution
+solution (and thus a complete expression) if it is a partial solution
 whose leaves are leaves of $\mathcal{G}$, if each OR node in
@@
 
-\subsection{Re-estimating the stochastic grammar over programs.}
+\subsection{Re-estimating the stochastic grammar}
 
@@
 In the formulation we consider, each task $t$ corresponds to a
-polynomial $f$, and an expression $e$ solves $t$ if ${ (e\;
-  i)=f(i) }$ for all ${ i=0,\dots,9 }$.
-
-Each task is to predict the value of an unknown polynomial on the
-inputs $0, \dots, 9$. The set of tasks corresponds to the set of
-polynomials $\{ax^2 + bx + c | a, b, c \in \{0, \dots, 9 \}\}$. In the
-initial library, we include the basic combinators I, S, B, C and four
-arithmetic primitives, $1$, $0$, $*$, and $+$. The choice of initial
-weights on these primitives can make a difference, as it determines
-the relative ordering of combinators in the enumeration step. To get a
-general sense of the algorithm's performance, we set the initial
-weights to be slightly different on each run, perturbing them around a
-uniform weighting.
-
-\begin{figure}[h]
+polynomial $f$, and an expression $e$ solves $t$ if ${ (e\; i)=f(i) }$
+for all ${ i=0,\dots,9 }$. For this experiment, the set of problems
+corresponds to the set of all polynomials with degree less than or
+equal to 2. In the initial library, we include the basic combinators
+I, S, B, C and four arithmetic primitives, $1$, $0$, $*$, and $+$. The
+choice of initial weights on these primitives can make a difference,
+as it determines the relative ordering of combinators in the
+enumeration step. To get a general sense of the algorithm's
+performance, we set the initial weights to be slightly different on
+each run, perturbing them around a uniform weighting.
+
+\setcounter{subfigure}{0} 
+\begin{figure}[ht]
 \centering
@@
 \includegraphics[width=.9\linewidth]{figures/symreg3frontiers.pdf}
-\label{fig:symregLearning}
 \caption{Learning curves as a function of frontier size.  As frontier
@@
   control brute-force search over $150000$ only hits $3\%$ of the
-  tasks.}
-\end{subfigure}
-
-\begin{subfigure}[After]{.65\linewidth}
+  tasks. \label{fig:symregLearning}}
+\end{subfigure}
+
+\begin{subfigure}[After]{\linewidth}
 \includegraphics[width=.9\linewidth]{figures/symRegDiffCurricula.pdf}
 \caption{How do different task sets affect learning curves? Learning
-  curves at frontier size of $10000$ for different task sets. Note
-  that performance is equally good if either the constant functions of
-  the linear functions are missing, by drops dramatically when both
+  curves at frontier size of $10000$ for different task
+  sets. Performance is equally good if either the constant functions
+  of the linear functions are missing, but drops dramatically when both
   are absent, even though these only account for about $10\%$ of the
-  tasks. Performance drops even lower if we only consider ``complex'' quadratics.}
+  tasks. Performance drops even lower if we only consider ``complex''
+  quadratics.}
 \label{fig:symregCurricula}
 \end{subfigure}
-\begin{subfigure}[After]{.3\linewidth}\
-\includegraphics[width=.9\linewidth]{figures/symregCircuitDepths.pdf}
-\caption{Circuit depth as a function of iteration on a typical run.}
-\label{fig:symregDepths}
-\end{subfigure}
-
-\caption{Symbolic regression experiment data.}
+%% \begin{subfigure}[After]{.3\linewidth}\
+%% \includegraphics[width=.9\linewidth]{figures/symregCircuitDepths.pdf}
+%% \caption{Circuit depth as a function of iteration on a typical run.}
+%% \label{fig:symregDepths}
+%% \end{subfigure}
+\vspace{-.2cm}
+\caption{Symbolic regression experiment results.}
 \label{fig:symreg} 
@@
 
-In Figure~\ref{fig:symreg}, we show performance results as we vary the
+In Figure~\ref{fig:symregLearning}, we show performance results as we vary the
 frontier size. Each individual plot shows how changing the frontier
@@
 
+What kinds of primitives does the E-C algorithm learn in this task?
+Inspecting the top weighted primitives of the final grammars, we find
+many incrementers (e.g. (+1), (+2), etc.), several versions of
+expressions that translate to functions like $x*f(x)$ and $f(x*f(x))$,
+and combinations of these, like $x*(x+3)+3$ and $x*(x+2)$. Informally,
+we see expressions that apply functions to themselves in various ways
+several times, building up complex functions with relatively few
+unique primitives. In the next experiment we will inspect more
+directly the modular components that are learned by this algorithm.
+
 We would like to assess to what degree ``conceptual bootstrapping'' is
-responsible for the improved performance. In the next experiment we
-will inspect directly the modular components that are learned by this
-algorithm. Here we look at the idea of a ``learnable'' curriculum,
-that is, the notion that conceptual bootstrapping requires that a
-set of tasks contain a subset of tasks that are sufficiently simple so
-that the learner is able to solve those with its initially
-impoverished library of concepts.
+responsible for the improved performance, that is, to what extent does
+the presence of simple problems account for ability to learn complex
+functions.  One prediction of this idea is that to succeed on a set of
+tasks, the set must contain a ``learnable'' curriculum, a set of tasks
+that serve as stepping stones from an impoverished representation to a
+rich one. We can test this prediction by varying the set of tasks to
+which the E-C algorithm is exposed.
 
@@
 set to quadratics with coefficients greater than 1, performance drops
-to 0.
+to 0 because no task is hit in the initial frontier.
+
+This data has in it the nonlinearity that we predicted -- that a
+minimal curriculum of simple tasks is sufficient to achieve high
+performance -- but also suggests that, at least in this domain, this
+is not an all-or-none effect. Though the performance drops
+significantly once no linear functions are included in the task set,
+learning does still occur, and the learning curves still have the same
+basic shape.
 
@@
 \label{sec:boolcircuits}
+
+\begin{figure}[ht]
+\includegraphics{./tablealonedoc.pdf}
+\vspace{-.5cm}
+\caption{Some learned primitives in Boolean function learning
+  experiment. We highlight compression using $NOT$, $E_1$ and $E_2$ in
+  some of the expressions they are found. \,(a)The three elementary
+  gates used to generate the task set. (b) Two higher-order primitives
+  in the induced grammer. Note how $E_1$ can be used to define FALSE
+  by taking the NOT function as an
+  argument.  \label{table:booleanCircuits}}
+\vspace{-.4cm}
+\end{figure}
 
@@
 
-\include{./figures/tablealone}
 %% \begin{table}
@@
 \begin{figure}
+\vspace{-.3cm}
 \includegraphics[width=\linewidth]{figures/circuitDistrComp.pdf}
-\vspace{-10pt}
+\vspace{-.6cm}
 \caption{Distribution of boolean functions among (a) the set of 1000
@@
 
-\begin{figure}[t]
-\begin{subfigure}[Before]{\linewidth}
+\begin{figure}[h]
+\begin{subfigure}{.1\linewidth}
+\caption{\label{fig:boolLearningCurvesCircuits} }
+\end{subfigure}
+\begin{subfigure}[Before]{.9\linewidth}
 \includegraphics[width=.9\linewidth]{figures/booleancircuits3frontiers.pdf}
-\label{fig:boolLearningCurvesCircuits}
-\caption{Learning curves for various frontier sizes on task set of
-  Boolean functions drawn randomly from a distribution (see text for
-  details) of circuits built up of NOT, AND, OR gates.}
-\end{subfigure}
-
-\begin{subfigure}[Before]{\linewidth}
+\end{subfigure}
+
+\vspace{-.2cm}
+\begin{subfigure}{.1\linewidth}
+\caption{\label{fig:boolLearningCurvesUnifrom}}
+\end{subfigure}
+\begin{subfigure}[After]{.9\linewidth}
 \includegraphics[width=.9\linewidth]{figures/booleanfunctionsUniform.pdf}
-\label{fig:boolLearningCurvesUnifrom}
-\caption{10 learning curves for frontier size of 2000 on a task set
-  consisting of all Boolean functions of cardinality 3 or
-  smaller. Note how most of the trajectories get stuck in a local
-  minimum around $50\%$.}
-\end{subfigure}
-\label{fig:boolLearningCurves} 
-\caption{Learning curves from Boolean function learning experiment. }
+
+\end{subfigure}
+\vspace{-.3cm}
+\caption{Learning curves from Boolean function learning
+  experiment. (a) Learning curves for various frontier sizes on task
+  set of Boolean functions drawn randomly from a distribution (see
+  text for details) of circuits built up of NOT, AND, OR gates. (b) 10
+  learning curves for frontier size of 2000 on a task set consisting
+  of all Boolean functions of cardinality 3 or smaller. Note how most
+  of the trajectories get stuck in a local minimum around
+  $50\%$.\label{fig:boolLearningCurves}
+
+}
+\vspace{-.5cm}
 \end{figure}

### 2013-02-01 drafts/ijcai2013/ijcai13.tex
Changes.

@@
 
+\usepackage{titling}
+
+\setlength{\droptitle}{0em}   
+\date{}
+
 \newtheorem{definition}{Definition}
@@
 
-\author{Eyal Dechter \\
-MIT\\
-USA \\
-edechter@mit.edu
-\And
-Jon Malmaud \\
-MIT \\ 
-USA \\
-malmaud@mit.edu
-\And 
-Ryan P.  Adams \\
-Harvard University\\
-USA \\
-rpa@seas.harvard.edu
-\And
-Joshua B. Tenenbaum \\
-MIT\\
-USA \\
-jbt@mit.edu
-}
+%% \author{Eyal Dechter \\
+%% MIT\\
+%% USA \\
+%% edechter@mit.edu
+%% \And
+%% Jon Malmaud \\
+%% MIT \\ 
+%% USA \\
+%% malmaud@mit.edu
+%% \And 
+%% Ryan P.  Adams \\
+%% Harvard University\\
+%% USA \\
+%% rpa@seas.harvard.edu
+%% \And
+%% Joshua B. Tenenbaum \\
+%% MIT\\
+%% USA \\
+%% jbt@mit.edu
+%% }
 
@@
 \maketitle
-
 \begin{abstract}
@@
 $\mathcal{H}$ is connected to all of that node's children in
-$\mathcal{G}$
+$\mathcal{G}$.
 
@@
 of $\mathcal{G}$ then either it is an AND node, to which we assign a
-value of 0, or it corresponds to some primitive combinator $c_i$ in
-the library and is the child of some typed OR node with type $t$. We
-assign it a value of $\log{p(c_i | t)}$. If a leaf $v$ of
+value of 0, or it corresponds to some primitive combinator $c_n$ in
+the library and is the child of some typed OR node with type $\tau$. We
+assign it a value of $\log{p(c_n | \tau)}$. If a leaf $v$ of
 $\mathcal{H}$ is an OR node, then the value of any extension of it to
@@
 right child. We want to apply constraints as early as possible, so
-expanding left children first prunes more partial solutions.
+expanding left child first prunes more partial solutions.
 
@@
 A natural metric for the compressibility of a set of binary trees is
-the number of unique subtrees (counting leaves as subtrees) that this
-set contains. As an example of this, suppose we had six tasks whose
-solutions were integers $4^2, 5^2, 6^2, 7^2, 8^2$ and suppose our
-frontier contained the 10 expressions:
-
+the number of unique subtrees (counting leaves) that this
+set contains. As an example of this, suppose we have six tasks whose
+solutions are integers $4^2, 5^2, 6^2, 7^2$, and $8^2$, and suppose our
+frontier contains the $10$ expressions:
+\vspace{-.3cm}
 \captionsetup[subfigure]{labelformat=empty} 
@@
 \captionsetup[subfigure]{labelformat=parens} 
-
+\vspace{-.3cm}
 The top row consists of one type of representation of an integer
@@
 represented in combinatory logic as $S * I$, is applied to the
-integer. The top and bottom row of each expression evaluates to the
+integer. The top and bottom row of each expression evaluate to the
 same integer. If we consider only the first column, the top tree has 4
@@
 possible. That is, let $t_i$ be a task in the set of hit tasks, and
-let $\{e^k_i\}$ be the expressions in the frontier that solves
+let $\{e^k_i\}$ be the expressions in the frontier that solve
 it. Then our goal is to find
@@
 $O(M^K)$ assignments, where $M$ is the maximum number of solutions for
-any one task and $K$ is the number of tasks. Since this is
+any one task, and $K$ is the number of tasks. Since this is
 prohibitive, we relax the compressibility metric as follows: we define
@@
 only on the last expression we added. Our solution thus becomes
-approximate and that approximation is order-dependent, but a
+approximate, and that approximation is order-dependent, but a
 depth-first search on the defined search space goes from exponential
@@
 
-Given a set of solution programs, we want to re-estimate our
-stochastic grammar. Our inspiration for doing this is the
-Neville-Manning algorithm for grammar-based
-compression~\cite{nevill1997identifying}. The idea of that compression
-algorithm is to compress any string by creating a grammar for that
-string such that a) no two symbols appear more than once in the
-grammar and b) every rule is used more than once. From a set of trees
-we generate a grammar according to these criteria.
+Given a set of chosen solution expressions, which we will call the
+\emph{solution set}, want to re-estimate our stochastic grammar. Our
+inspiration for doing this is the Neville-Manning algorithm for
+grammar-based compression~\cite{nevill1997identifying}. The idea of
+that compression algorithm is to compress any string by creating a
+grammar for that string such that a) no two symbols appear more than
+once in the grammar and b) every rule is used more than once. From a
+set of expressions we generate a grammar according to these criteria.
 
@@
 elements from the new grammar could have been used. In accordance with
-Section~\ref{stochgrammar}, we estimate the terminal production weight
+Section~\ref{sec:stochgrammar}, we estimate the terminal production weight
 to be the ratio of the number of times that node was used in the first
@@
 We first illustrate the performance of our learning algorithm on a
-symbolic regression problem. In task in this problem, we want to find
-an algebraic function, symbolically specified, that maps a set of
+symbolic regression problem. For each task in this problem, we want to
+find an algebraic function, symbolically specified, that maps a set of
 input values to output values. We choose this problem because it has a
@@
 In the formulation we consider, each task $t$ corresponds to a
-polynomial $f$, and an expression $e$ solves $t$ if ${ (e\; i)=f(i) }$
-for all ${ i=0,\dots,9 }$. For this experiment, the set of problems
-corresponds to the set of all polynomials with degree less than or
-equal to 2. In the initial library, we include the basic combinators
-I, S, B, C and four arithmetic primitives, $1$, $0$, $*$, and $+$. The
-choice of initial weights on these primitives can make a difference,
-as it determines the relative ordering of combinators in the
-enumeration step. To get a general sense of the algorithm's
-performance, we set the initial weights to be slightly different on
-each run, perturbing them around a uniform weighting.
+polynomial $f$, and an expression $e$ solves $t$ if it returns the
+same value when applied $i \in 0, \dots, 9$ that $f$ does. In the
+experiment shown here, the set of problems corresponds to the set of
+all polynomials with degree less than or equal to 2. In the initial
+library, we include the basic combinators I, S, B, C and four
+arithmetic primitives, $1$, $0$, $*$, and $+$. The choice of initial
+weights on these primitives can make a difference, as it determines
+the relative ordering of combinators in the enumeration step. To get a
+general sense of the algorithm's performance, we set the initial
+weights to be slightly different on each run, perturbing them around a
+uniform weighting.
 
@@
 \caption{Learning curves as a function of frontier size.  As frontier
-  size is increased, it plateaus closer to $100\%$ performance. A
-  control brute-force search over $150000$ only hits $3\%$ of the
+  size is increased, curves plateau closer to $100\%$ performance. A
+  baseline search over $150000$ expressions only hits $3\%$ of the
   tasks. \label{fig:symregLearning}}
@@
 
-In Figure~\ref{fig:symregLearning}, we show performance results as we vary the
-frontier size. Each individual plot shows how changing the frontier
-size changes the number of tasks the algorithm identifies correctly
-over the course of fifteen algorithm iterations. As a control, we
-enumerated $10000*15=150000$ expressions from the initial
-grammar. This is the total number of expressions that a run of the
-algorithm sees if it has a frontier size of $10000$ and runs for $15$
-iterations. Thus, if the benefit accruing to the runs with larger
-frontier sizes is due to simple an increase in brute-force search
-ability, rather than increased learning, we should see similarly good
-performance from this control run. In fact, however, the control run
-only hit $27$ of the $1000$ asks ($3\%$), whereas our algorithm nears $100\%$.
-
-What kinds of primitives does the E-C algorithm learn in this task?
-Inspecting the top weighted primitives of the final grammars, we find
-many incrementers (e.g. (+1), (+2), etc.), several versions of
-expressions that translate to functions like $x*f(x)$ and $f(x*f(x))$,
-and combinations of these, like $x*(x+3)+3$ and $x*(x+2)$. Informally,
-we see expressions that apply functions to themselves in various ways
-several times, building up complex functions with relatively few
-unique primitives. In the next experiment we will inspect more
-directly the modular components that are learned by this algorithm.
-
-We would like to assess to what degree ``conceptual bootstrapping'' is
-responsible for the improved performance, that is, to what extent does
-the presence of simple problems account for ability to learn complex
-functions.  One prediction of this idea is that to succeed on a set of
-tasks, the set must contain a ``learnable'' curriculum, a set of tasks
-that serve as stepping stones from an impoverished representation to a
-rich one. We can test this prediction by varying the set of tasks to
-which the E-C algorithm is exposed.
+In Figure~\ref{fig:symregLearning}, we show performance results as we
+vary the frontier size.Changing the frontier size changes the number
+of tasks the algorithm identifies correctly over the course of 15
+algorithm iterations (all experiments in this work were run for 15
+iterations). As a baseline, we enumerated $10000*15=150000$
+expressions from the initial grammar. This is the total number of
+expressions that a run of the algorithm sees if it has a frontier size
+of $10000$ and runs for $15$ iterations. Thus, if the benefit accruing
+to the runs with larger frontier sizes is due to simply an increase in
+the number of expressions seen, rather than increased learning, we should
+see similarly good performance from this baseline run. In fact,
+however, the baseline run only hits $27$ of the $1000$ tasks ($3\%$),
+whereas our algorithm nears $100\%$ for a frontier of $10000$.
+
+What does the E-C algorithm learn in this task?  Inspecting the top
+weighted primitives of the final grammars, we find many incrementers
+(e.g. (+1), (+2), etc.), several versions of expressions that
+translate to functions like $x*f(x)$ and $f(x*f(x))$, and combinations
+of these, like $x*(x+3)+3$ and $x*(x+2)$. Informally, we see
+expressions that apply functions to themselves, building up complex
+functions with relatively few unique primitives.
+
+To what degree is ``conceptual bootstrapping'' responsible for the
+improved performance. That is, to what extent does the presence of
+simple problems account for the ability to learn complex functions?
+One hypothesis is that to succeed on a set of tasks, the set must
+contain a ``learnable'' curriculum, a set of tasks that serve as
+stepping stones from an impoverished representation to a rich one. We
+can test this hypothesis by varying the set of tasks to which the EC
+algorithm is exposed. If it is true, then we should see a nonlinear
+reponse to reducing the number of easy tasks, as the curriculum
+changes from being learnable to being unlearnable.
 
@@
 performance. When we further restrict the task set to only ``complex''
-quadratics (which we are defining as quadratics all of whose
+quadratics (which we define as quadratics whose
 coefficients are greater than zero), we observe another comparable
@@
 
-\subsection{Boolean circuits}
+\subsection{Boolean function learning}
 \label{sec:boolcircuits}
@@
 \vspace{-.5cm}
-\caption{Some learned primitives in Boolean function learning
-  experiment. We highlight compression using $NOT$, $E_1$ and $E_2$ in
-  some of the expressions they are found. \,(a)The three elementary
-  gates used to generate the task set. (b) Two higher-order primitives
-  in the induced grammer. Note how $E_1$ can be used to define FALSE
-  by taking the NOT function as an
-  argument.  \label{table:booleanCircuits}}
+\caption{Some learned primitives from the circuit based Boolean
+  function learning experiment. We highlight compression using $NOT$,
+  $E_1$ and $E_2$. \,(a)The three elementary gates used to generate
+  the task set. (b) Two higher-order primitives in the induced
+  grammer. Note how $E_1$ can be used to define FALSE by taking the
+  NOT function as an argument. (c) Additional common Boolean functions
+  found in the learned grammar.  \label{table:booleanCircuits}}
 \vspace{-.4cm}
@@
 
-As another demonstration of this approach, we investigate the
-procedure's ability to learn boolean functions using only the logical
+As another demonstration of our approach, we investigate the
+EC algorithm's ability to learn Boolean functions using only the logical
 primitive NAND and the basic combinators. It is well known that a
@@
 the NAND gate. Here, the basic combinators take the place of the
-wiring normally found in a circuit, the output of element to multiple
-inputs downstream.
+wiring normally found in a circuit.
 
@@
 We constructed two task sets. The first of these was constructed
-explicitly to contain familiar module structure. In this set of tasks,
-we sampled $1000$ boolean circuits using AND, OR, and NOT gates. To
-accomplish this, we first sampled either 1, 2, 3 inputs, then between
-1 and 5 gates, randomly wiring the inputs of each new gate to the
-output of one of the existing gates in the circuit. We continued this
-sampling procedure until we had 1000 ``connected'' circuits, i.e.,
-circuits all of whose output are wired to an input except for the last
-one (the output gate). This resulted in $1000$ tasks consisting of
-$82$ unique boolean functions, with a distribution of tasks as shown
-in Figure~\ref{fig:booldistr}. In Figure~\ref{fig:boolLearningCurvesCircuits},
-We present learning curves for frontier sizes 100, 500 and 10000,
-using the same procedure as in Section~\ref{sec:symreg}.
+explicitly to contain familiar modular structure. In this set of
+tasks, we sampled $1000$ Boolean circuits using AND, OR, and NOT
+gates. To accomplish this, we first sampled either 1, 2, 3 inputs,
+then between 1 and 5 gates, randomly wiring the inputs of each new
+gate to the output of one of the existing gates in the circuit. We
+continued this sampling procedure until we had 1000 ``connected''
+circuits, i.e., circuits all of whose outputs are wired to an input
+except for the last one (the output gate). This resulted in $1000$
+tasks consisting of $82$ unique Boolean functions, with a distribution
+of tasks as shown in Figure~\ref{fig:booldistr}~(a). In
+Figure~\ref{fig:boolLearningCurvesCircuits}, We present learning
+curves for frontier sizes 100, 500 and 10000, using the same procedure
+as in Section~\ref{sec:symreg}. In Figure~\ref{fig:booldistr}, we show
+that the distribution over Boolean functions enumerated from the
+grammar over expressions is much more similar to the true function
+distribution after learning than before.
 
@@
   circuits randomly constructed from AND, OR and NOT gates (see text);
-  b) the first 1000 expressions from the grammar over programs
+  b) the first 1000 expressions from the grammar over expressions
   \emph{before} learning; c) the first 1000 expressions from the
-  grammar over programs \emph{after} learning. Gray levels indicate
+  grammar over expressions \emph{after} learning. Gray levels indicate
   percent of instances (circuits/expressions) which evaluate to that
@@
 recovers the constituent logic gates that were used to build up the
-problems. Table~\ref{table:bool} shows a few of the top ranked
+tasks. Table~\ref{table:booleanCircuits} shows a few of the top ranked
 expressions in the library. These include the basic logic gates; we
-also include one higher-order expression $F$, to stress that the
-representation we are using allows one to build concepts that are more
-expressive that just sub-circuits.
-
-In a second experiment, we include all $272$ boolean truth tables with
-three or fewer inputs as tasks. This is a more difficult problem. Note
-that in the first boolean function problem, there are two kinds of
-learning that can take place: first, many expressions map to a single
-boolean function, so we need to learn primitives that allow us to span
-many different functions instead of allows generating very simple
-functions over and over again. The second kind of learning involves
-the distribution of the functions themselves. In the first boolean
+also include two higher-order expression $E_1$ and $E_2$, to stress
+that the representation the EC algorithm is using allows it to build
+concepts that are more expressive that just sub-circuits.
+
+In a second experiment, we include all $272$ Boolean truth tables with
+three or fewer inputs as tasks. This is a more difficult
+problem. There are two kinds of learning that the EC algorithm might
+accomplish: first, many expressions in $\mathcal{L}$ map to a single
+Boolean function, so it needs to learn primitives that allow it to
+span many different functions instead of generating very simple
+functions redundantly. The second kind of learning involves the
+distribution of the functions themselves. In the circuit based Boolean
 experiment, that structure is apparent in the distribution in
-Figure~\ref{fig:boolhist}. In this second experiment, we remove the
-second kind of structure. In Figure~\ref{fig:boolLearningCurvesUniform}, we show
-10 runs of the algorithm with a frontier size of $2000$: note how
-there are several local minima that majority of the runs get stuck in
-with performance around $50\%$, but several of the runs seem to take
-different trajectories to more successful representations.
+Figure~\ref{fig:booldistr}. In this second experiment, we remove the
+second kind of structure. In
+Figure~\ref{fig:boolLearningCurvesUniform}, we show 10 runs of the EC
+algorithm on the second experiment with a frontier size of $2000$: note how there are several
+local minima that the majority of the runs get stuck in with performance
+around $50\%$, but several of the runs seem to take different
+trajectories to more successful representations.
 
@@
 \begin{subfigure}{.1\linewidth}
-\caption{\label{fig:boolLearningCurvesUnifrom}}
+\caption{\label{fig:boolLearningCurvesUniform}}
 \end{subfigure}
@@
 \caption{Learning curves from Boolean function learning
-  experiment. (a) Learning curves for various frontier sizes on task
-  set of Boolean functions drawn randomly from a distribution (see
-  text for details) of circuits built up of NOT, AND, OR gates. (b) 10
+  experiments. (a) Learning curves for various frontier sizes on task
+  set of circuit based Boolean functions. (b) 10
   learning curves for frontier size of 2000 on a task set consisting
@@
 
+\section{Conclusion}
+This work brings together ideas from various areas of research --
+hierarchical Bayes, learning to learn, minimum description length
+learning, and learning programs -- to provide a proof of concept for
+how bootstrap learning of abstract composable concepts can rapidly
+improve search performance.  We show that this approach can learn
+successfully even in domains for which the solution space is not
+smooth and the error signal is all or none, where the only sense of
+locality to guide learning is the modular and compositional structure
+of the solutions that the algorithm itself builds.
+
+Our specific implementation of this approach involved a number of
+choices which should be explored further in future work.  Our claim
+here is not that the EC algorithm is optimal, but that it captures key
+features of how people solve the problem of learning new systems of
+concepts in acquiring deep domain expertise.  Much like a human
+researcher investigating a new area of study, our algorithm tries
+different models, watches them fail or succeed, and abstracts out the
+model elements that capture relevant differences.  It recombines and
+reuses succcessful patterns of reasoning that span many problems in a
+domain.
+
+At their best, human processes of discovery also clearly go beyond
+what our algorithm is capable of.  The algorithm can get stuck in
+plateaus after a rapid initial increase in performance, due to its
+circular nature: it only inspects the region of the solution space
+that is typical of solutions already found.  Humans can become stuck
+in similar ways, but at least sometimes they can realize that they are
+stuck, and attempt to find another class of concepts for the remaining
+unsolved tasks.  Our algorithm requires, as human learners do, a
+sufficiently graded ``curriculum'', or spectrum of problems to solve --
+simple problems provide necessary stepping stones to building complex
+representations.  Also like humans, our algorithm can adaptively
+``self-pace'' its ways through the curriculum, figuring out which
+problems it is ready to tackle when, which problems are appropriate in
+difficulty given what it currently knows.  Yet people can sometimes
+more intelligently compose and modify the set of problems under
+consideration, designing their own stepping stones to expand the set
+of problems they can solve.  Making bootstrap learning systems smarter
+in these more human-like ways is a prime target for future research.
 %% \subsection{Digital arithmetic}
@@
 
-\section{Discussion}
-This work attempts to bring together ideas from various areas of
-research -- hierarchical Bayes, learning to learn, minimum description
-length learning -- to provide a proof of concept that learning
-abstractions from programs via compression can rapidly improve search
-performance. We emphasize this by focusing on domains in which the
-solution space is not smooth and the error signal is all or
-none. Thus, the main source of information for learning is the modular
-and compositional structure of the solution space.
-
-Of course, the representations that people use to learn in the world
-often have a partially smooth or local structure. Techniques in AI and
-machine learning commonly use smoothness or local structure to make
-search tractable. But the difference between this kind of learning and
-conceptual learning is undeniable. The former is what the statistician
-or machine learning researcher implements. The latter is what she
-learns as a student and as a practicioner. As researchers, we try
-different models, watch them fail or succeed, and abstract out the
-elements of these models that capture the relevant differences. We mix
-and match parts of different ideas, and utilize succcessful patterns
-of reasoning that can span many problems and domains.
-
-We believe that the primary characteristic of conceptual learning is
-that it occurs over a modular and compositional computational
-representation that supports higher-order functions. That is why we
-chose a functional programming language as the representational
-language of this work.
-
-The experiments in this work do not contain the richness of real world
-learning, but we believe that the learning algorithm's behavior
-exhibits some of the characteristics we would expect of a conceptual
-learner.
-
-One characteristic of this algorithm's behavior is a rapid initial
-increase in performance followed by a hard plateau. This plateau is
-expected given the circular nature of the algorithm: the learner
-inspects the region of the solution space that is typical of solutions
-already found. We see in this phenomenon a conceptual demarcation of
-the problem space: a coherent set of concepts or problems solving
-techniques can define by their power the extent of problems a problem
-solver is willing to consider. One might hope, however, that an
-intelligent learner would recognize that such a demarcation has
-occured and attempt to find another class of concepts for the
-remaining unsolved tasks.
-
-Another aspect of conceptual learning that we see present in this
-algorithm's behavior is the importance of a curriculum. By this we do
-not mean an explicit ordering of the problems in the task set, but
-rather that a set of sufficiently simple problems must act as stepping
-stones to the acquisition of complex representations.
+%% \section{Discussion}
+%% This work attempts to bring together ideas from various areas of
+%% research -- hierarchical Bayes, learning to learn, minimum description
+%% length learning -- to provide a proof of concept that learning
+%% abstractions from programs via compression can rapidly improve search
+%% performance. We emphasize this by focusing on domains in which the
+%% solution space is not smooth and the error signal is all or
+%% none. Thus, the main source of information for learning is the modular
+%% and compositional structure of the solution space.
+
+%% Of course, the representations that people use to learn in the world
+%% often have a partially smooth or local structure. Techniques in AI and
+%% machine learning commonly use smoothness or local structure to make
+%% search tractable. But the difference between this kind of learning and
+%% conceptual learning is undeniable. The former is what the statistician
+%% or machine learning researcher implements. The latter is what she
+%% learns as a student and as a practicioner. As researchers, we try
+%% different models, watch them fail or succeed, and abstract out the
+%% elements of these models that capture the relevant differences. We mix
+%% and match parts of different ideas, and utilize succcessful patterns
+%% of reasoning that can span many problems and domains.
+
+%% We believe that the primary characteristic of conceptual learning is
+%% that it occurs over a modular and compositional computational
+%% representation that supports higher-order functions. That is why we
+%% chose a functional programming language as the representational
+%% language of this work.
+
+%% The experiments in this work do not contain the richness of real world
+%% learning, but we believe that the learning algorithm's behavior
+%% exhibits some of the characteristics we would expect of a conceptual
+%% learner.
+
+%% One characteristic of this algorithm's behavior is a rapid initial
+%% increase in performance followed by a hard plateau. This plateau is
+%% expected given the circular nature of the algorithm: the learner
+%% inspects the region of the solution space that is typical of solutions
+%% already found. We see in this phenomenon a conceptual demarcation of
+%% the problem space: a coherent set of concepts or problems solving
+%% techniques can define by their power the extent of problems a problem
+%% solver is willing to consider. One might hope, however, that an
+%% intelligent learner would recognize that such a demarcation has
+%% occured and attempt to find another class of concepts for the
+%% remaining unsolved tasks.
+
+%% Another aspect of conceptual learning that we see present in this
+%% algorithm's behavior is the importance of a curriculum. By this we do
+%% not mean an explicit ordering of the problems in the task set, but
+%% rather that a set of sufficiently simple problems must act as stepping
+%% stones to the acquisition of complex representations.
 

### 2013-03-19 drafts/ijcai2013/ijcai13.tex
IJCAI 2013 submission changes before final submission.

@@
 
-\setlength{\droptitle}{0em}   
+\setlength{\droptitle}{-1em}   
 \date{}
@@
 \begin{abstract}
+%\small
  Suppose a learner is faced with a domain of problems about which it
@@
 
-But thes problems we want AI systems to solve often lack this kind of
-built in structure. These are the kinds of problems more often
-associated with ILP and classical AI than with GP and machine
-learning. 
-\\
+But, as mentioned before, the problems we want AI systems to solve
+often lack this kind of built in structure. Here we attempt to learn
+functional programs to solve those kinds of problems.  \\
 
@@
 experiments in the domains of symbolic regression and Boolean
-functioin learning.
+function learning.
 
@@
 
-We propose the \textbf{E-C algorithm}, an iterative procedure to solve this
+We propose the \textbf{EC algorithm}, an iterative procedure to solve this
 problem. It will be useful in presenting the algorithm to define the
@@
 
-The E-C algorithm maintains a distribution $\mathcal{D}$ over
+The EC algorithm maintains a distribution $\mathcal{D}$ over
 expressions in $\mathcal{L}$. At each iteration, it
@@
 
-Next, we specify a representation language $\mathcal{L}$, a
-form for the distribution $\mathcal{D}$, and a compression criterion for a
-set of expressions in $\mathcal{L}$. We also describe
-procedures for the enumeration of the frontier, compression of the
-expressions, and re-estimation of $\mathcal{D}$.
-
 \subsection{Representing typed functional programs as binary trees}
 
-Following~\cite{liang10programs} and~\cite{Briggs:2008}, we use a
+Following~\cite{liang10programs} and~\cite{briggs2006functional}, we use a
 simply typed combinatory logic -- a variable-free subset of the
@@
 above). The lambda calculus has two basic operations -- application
-and abstraction (using the $\lambda$ operator -- but in using the
+and abstraction -- but in using the
 combinatory logic we sequester uses of the abstraction operation
@@
 product of the probabilities of its subcomponents. The distribution we
-specify will be a simple verson of this, in which the
+specify will be a simple version of this, in which the
 probability of an expression is the product of the probability of its
@@
 choice of primitive at each leaf were not conditioned on the
-requesting type, we could just count all occurences of each primitive
+requesting type, we could just count all occurrences of each primitive
 in our expression set, and this would be proportional to the maximum
@@
 $\mathcal{E}$ when the requesting type is such that it could have
-been chosen. This is a straightforward calculuation, and
+been chosen. This is a straightforward calculation, and
 we use it for its convenience, finding that our empirical results
@@
 In order to explore the frontier of most promising programs, we need a
-procedure that enumerates the $N$ most probable programs. There has
+procedure that enumerates the $N$ most probable expressions. There has
 been recent interest in this problem, most of it focused on
@@
 \centering
-\begin{subfigure}[Before]{\linewidth}
-\includegraphics[width=.9\linewidth]{figures/symreg3frontiers.pdf}
+\includegraphics[width=.8\linewidth]{figures/symreg3frontiers.pdf}
+\vspace{-.2cm}
 \caption{Learning curves as a function of frontier size.  As frontier
@@
   tasks. \label{fig:symregLearning}}
-\end{subfigure}
-
-\begin{subfigure}[After]{\linewidth}
-\includegraphics[width=.9\linewidth]{figures/symRegDiffCurricula.pdf}
+\vspace{-.3cm}
+\end{figure}
+\begin{figure}[ht]
+\centering
+\includegraphics[width=.8\linewidth]{figures/symRegDiffCurricula.pdf}
 \caption{How do different task sets affect learning curves? Learning
   curves at frontier size of $10000$ for different task
-  sets. Performance is equally good if either the constant functions
-  of the linear functions are missing, but drops dramatically when both
-  are absent, even though these only account for about $10\%$ of the
-  tasks. Performance drops even lower if we only consider ``complex''
-  quadratics.}
+  sets.} 
 \label{fig:symregCurricula}
-\end{subfigure}
-%% \begin{subfigure}[After]{.3\linewidth}\
-%% \includegraphics[width=.9\linewidth]{figures/symregCircuitDepths.pdf}
-%% \caption{Circuit depth as a function of iteration on a typical run.}
-%% \label{fig:symregDepths}
-%% \end{subfigure}
-\vspace{-.2cm}
-\caption{Symbolic regression experiment results.}
-\label{fig:symreg} 
+\vspace{-.5cm}
 \end{figure}
@@
 
-What does the E-C algorithm learn in this task?  Inspecting the top
+What does the EC algorithm learn in this task?  Inspecting the top
 weighted primitives of the final grammars, we find many incrementers
@@
 algorithm is exposed. If it is true, then we should see a nonlinear
-reponse to reducing the number of easy tasks, as the curriculum
+response to reducing the number of easy tasks, as the curriculum
 changes from being learnable to being unlearnable.
@@
 
-\begin{figure}[ht]
+\begin{figure}[t]
 \includegraphics{./tablealonedoc.pdf}
@@
   the task set. (b) Two higher-order primitives in the induced
-  grammer. Note how $E_1$ can be used to define FALSE by taking the
+  grammar. Note how $E_1$ can be used to define FALSE by taking the
   NOT function as an argument. (c) Additional common Boolean functions
@@
   function.\label{fig:booldistr}}
+\vspace{-.4cm}
 \end{figure}
@@
 
+\vspace{-.3cm}
 \begin{figure}[h]
@@
 
-\vspace{-.2cm}
+\vspace{-.32cm}
 \begin{subfigure}{.1\linewidth}
@@
 model elements that capture relevant differences.  It recombines and
-reuses succcessful patterns of reasoning that span many problems in a
+reuses successful patterns of reasoning that span many problems in a
 domain.
 
-At their best, human processes of discovery also clearly go beyond
+At their best, human processes of discovery clearly go beyond
 what our algorithm is capable of.  The algorithm can get stuck in
@@
 that is typical of solutions already found.  Humans can become stuck
-in similar ways, but at least sometimes they can realize that they are
-stuck, and attempt to find another class of concepts for the remaining
+in similar ways, but at least sometimes they realize that they are
+stuck and attempt to find another class of concepts for the remaining
 unsolved tasks.  Our algorithm requires, as human learners do, a
@@
 ``self-pace'' its ways through the curriculum, figuring out which
-problems it is ready to tackle when, which problems are appropriate in
-difficulty given what it currently knows.  Yet people can sometimes
+problems it is ready to tackle when, which problems are appropriately
+difficult given what it currently knows.  Yet people can sometimes
 more intelligently compose and modify the set of problems under

### 2013-03-19 drafts/general_writeup/writeup.tex
Reorganization and cleanup of drafts directory.

--- 2012-11-18 drafts/proposal.tex
@@
 \usepackage[ruled,vlined,linesnumbered]{algorithm2e}
+\usepackage{amsmath}
 
@@
     }
+
+\DeclareMathOperator*{\argmax}{arg\,max}
 
@@
 
+\section{Algorithm 0}
+In order to motivate our approach, we present the following problem:
+suppose we have a probability model $P(D, E, \mathcal{L}) = P(D|E)
+P(E|L) P(L)$ where $L$ represents the parameters of a latent model,
+$E_i$ are drawn $i.i.d$ from $P(\dot | L)$ and $D_i$ are drawn $i.i.d$
+from $P(\dot|E_i)$. The $E_i$ are hidden, the $D_i$ observed. Suppose
+we are able to 1) draw sampled of $E_i$ given $L$ and 1) we can
+enumerate $E_i$ in order of decreasing probability under $L$. Our goal
+is to determine the most likely $L$ given the data $D$. A rejection
+sampling algorithm approach might involve sampling the $E_i$ from the
+current value of $L$, accepting $E_i$ if  To solve this problem, we propose
+Algorithm~\ref{alg:short}: at each iteration of our algorithm, we
+choose
+
+\begin{minipage}{.7\linewidth}
+  \begin{algorithm}[H]
+    \SetAlgoLined 
+    \KwData{$\mathcal{D}=(D_1, \dots, D_M)$,
+      $\mathcal{L}^{t-1}$ , $0 \geq K \leq 1$} 
+    \KwResult{updated library $\mathcal{L}^t$}
+    $E \leftarrow \{ E_i : p(E_i\, |\, \mathcal{L}^{t-1}) > K\}$\;
+    $S \leftarrow \{ h \in H \cap D \}$\;
+    $\mathcal{L}^t \leftarrow \argmax_L p(L \, | \, S)$
+    \caption{Algorithm 0}
+    \label{alg:short}
+  \end{algorithm}
+\end{minipage}
+
 \section{Algorithm}
@@
 \bibliographystyle{plain}
-\bibliography{bibliography}
+\bibliography{writeup}
 

### 2013-03-19 drafts/grammar/grammar.tex
Reorganization and cleanup of drafts directory.

\documentclass{article}

\usepackage{amsmath}
\usepackage{ amssymb }
\usepackage{graphicx}
\usepackage{syntax}
\usepackage{tikz}
\usetikzlibrary{arrows,automata, positioning, patterns,backgrounds}
\usetikzlibrary{arrows,shapes.gates.logic.US,shapes.gates.logic.IEC,calc}

\newcommand{\funtype}[2]{#1 \rightarrow #2}
\newcommand{\weightedrule}[3]{[#1 \mapsto #2 \, | \, #3]}
\newcommand{\typedexp}[2]{#1 : #2} 
\newcommand{\type}[1]{type(#1)}

\renewcommand{\grammarlabel}[2]{#1 #2}

\begin{document}

In this document, I will describe a stochastic grammar over typed
combinatory logic expressions.

\section{Combinatory expressions as binary trees}
We use a polymorphic combinatory logic. Every combinatory logic
expression $e$ of type $\tau$ is defined to be either a primitive
combinator $c$ of type $\tau$ or the application of a left expression
$e_l$ of type $\funtype{\sigma}{\tau}$ to a right expression $e_r$ of
type $\sigma$ can be represented as a binary tree $t_e$, which
we call a \textbf{combinatory tree}.

We define a stochastic process $\mathcal{D}$ that generates such
trees. A \textbf{terminal rule} $r = \weightedrule{\tau}{c}{w}$ specifies that a
nonterminal node of type $\tau$ can produce a primitive combinator $c$
with associated integer weight parameter $w$. $\mathcal{D}$ is a a set
of

\section{A stochastic grammar for polymorphic combinatory expressions}

A grammar for polymorphically typed combinatory expressions cannot be
context-free. To generate a function of type $t$, for example, I might
decide to apply a function of type $\funtype{s}{t}$ to an argument of
  type $s$. The types of the function and argument need to
  match.

It is natural to describe a stochastic grammar in terms of the process
which generates elements from that grammar. To generate an expression
of type $t$, we begin with a nonterminal node of $t$:

\begin{tikzpicture}
\node[draw]{$t$};
\end{tikzpicture}

We have two choices: 
\begin{enumerate}
\item (terminal production) With probability $p_{t \rightarrow c}$ we
  expand this node to terminal $c$ where $\type{c}$ unifies with $t$,
  and we update the type of the nonterminal to be the most general
  unifier (mgu) of $\type{c}$ and $t$:

\begin{tikzpicture}
\node[draw](top){$mgu(t, \type{c})$};
\node[draw, below=.5cm of top](c){$c$};

\draw (top)--(c);
\end{tikzpicture}

We propogate this type up the expression tree. If we find a
nonterminal node that can be expanded, we procede with that
expansion. Otherwise, we are done.

\item (nonterminal production) With probability $p_{t \rightarrow
  \epsilon}$ we expand this node into an application:

\begin{tikzpicture}
\node[draw](top){$t$};
\node[draw, below left=.5cm of top](left){$\funtype{t'}{t}$};
\node[below right=.5cm of top](right){};

\draw
(top) -- (left)
(top) -- (right);
\end{tikzpicture}
\end{enumerate}

Types:
\begin{grammar}
<$t$> ::= $\tau$, $\sigma$  (type variables)
  \alt $a$, $b$ (concrete types e.g. real, bool, int, etc.)
  \alt $\funtype{t'}{t}$ (function type)
\end{grammar}

Expressions: 
\begin{grammar}
<$\typedexp{e}{t}$> ::= $\typedexp{c}{t'}$ s.t. $t' = mgu(t,
\type{c})$ (substitute a terminal tree that unifies with the requesting type)
  \alt ($\typedexp{e}{\funtype{t'}{t}} \;\;\; \typedexp{e}{t'}$)
\end{grammar}

Let $G(e ; t)$ be the probability of type $t$ yielding expression
$e$. We define $G(\cdot \; ; \cdot)$ according to the following recursion:
\begin{align}
G( (e_1 \;\; e_2 ); t) &\propto \sum_{t' \trianglerighteq t} \phi_{t'} G(e_1 \; ; \; \funtype{s}{t})
   G(e_2 \; ; \;  \text{sourceType}(\type{e_1})\\
G( c \; ; \; t) &\propto \sum_{t' \trianglerighteq t} \phi_{\funtype{t'}{c}}
\end{align}

\section{A nonparametic prior over combinatory expressions}
We would like to describe a distribution over combinatory expressions
such that potential number of ``primitive combinators'' is
unbounded. This will allow us to naturally capture the possibility
that our library of combinator subroutines includes any number of
subroutines (themselves constructed of the primitive combinators). To
do this, we will use a modified version of the Pitman-Yor adaptor
grammar framework~\cite{DBLP:conf/nips/JohnsonGG06}. Our modification
is simply that instead of the base context-free grammar we use the
grammar described in the previous section.

\bibliographystyle{plain}
\bibliography{grammar}

\end{document}

### 2013-03-19 drafts/ijcai2013/ijcai13abstract.tex
More drafts cleanup.

\documentclass{article}

\usepackage{ijcai13}
\usepackage{times}
\usepackage{amsmath} 
\usepackage{latexsym} 
\usepackage{ dsfont }
\usepackage{tikz}
\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
\usetikzlibrary{arrows,automata}

\DeclareMathOperator*{\argmax}{arg\,max}

\title{The E-C Algorithm: conceptual bootstrapping via grammar-based
  program compression.}

%%\title{The Exploration-Compression Algorithm: how to learn programs by discovering useful subroutines.}
\author{Eyal Dechter \\
MIT\\
USA \\
edechter@mit.edu
\And
Josh Tenenbaum \\
MIT\\
USA \\
jbt@mit.edu
\And 
Ryan Adams \\
Harvard University\\
USA \\
rpa@seas.harvard.edu}

\begin{document}

\maketitle

\begin{abstract}
Suppose a learner is faced with a set of problems about which it knows
nearly nothing: it doesn't know the distribution of problems, or what
an appropriate loss function might be. For each hypothesized solution
to a problem it gets one bit: right or wrong. How can such a learner
every get off the ground? The key is to find, by brute force, a
solution to one or more simple problems in that domain and use the
information thus gained to pursue subsequent searches in an informed
manner. Here, we formalize this intuition, where the solution space is
the space of typed functional programs and the gained information is
stored as a stochastic grammer over programs. Specifically, we propose
the E-C algorithm, a two step iterative algorithm for exploring such
spaces: in the E step, the learner \emph{explores} a finite subset of
the domain for solutions, guided by a stochastic grammar; in the C
step, the learner \emph{compresses} the successful solutions from the
E step to estimate a new stochastic grammar. As we show in several toy
domains, iterating these two steps quickly propels the learner from a
state of solving almost no problems to solving a large number of
problems by gaining abstract knowledge of the structure of the
solution space.
\end{abstract}

\end{document}

### 2013-03-19 drafts/ijcai2013/figures/circuitTable.tex
Cleanup and add figures directory.

\documentclass[preview]{standalone}

\usepackage{times}
\usepackage{amsmath} 
\usepackage{latexsym} 
\usepackage{ dsfont }
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{tikz}
\usepackage{tikz-qtree}
\usepackage{float}
\usepackage{circuitikz}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{alltt}
\usepackage{dashrule}
\usepackage{array}
\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
\usetikzlibrary{arrows,automata, positioning, patterns,backgrounds}
\usetikzlibrary{arrows,shapes.gates.logic.US,shapes.gates.logic.IEC,calc}

\usetikzlibrary{shapes.geometric}
\usetikzlibrary{shapes.misc}

\begin{document}

%%%%%%%% E2 %%%%%%%%%%%
\begin{tikzpicture}
\node (x) at (0,0) {$x$};
\node (f)[above right=.1 and .6cm of x, draw=blue!75, fill=blue!25, text width=.6cm] {$f(\cdot)$};
\node (g)[below right=.1 and .6cm of x, draw=red!75, fill=red!25, text width=.6cm] {$g(\cdot)$};
\node (nand)[nand gate US, draw, right=1.8cm of x] {};

\draw (x.east) -| +(.2, 0)  |- (g.west);
\draw (x.east) -| +(.2, 0)  |- (f.west);
\draw (f.east) -| +(.3, 0) |-  (nand.input 1);
\draw (g.east) -| +(.3, 0) |-  (nand.input 2);
\end{tikzpicture}
%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%% NOT %%%%%%%%
\begin{tikzpicture}
\node (x) at (0,0) {$x$};
\node (nand)[nand gate US, draw, right=.6cm of x] {};
\node (equiv)[right=.2cm of nand] {$\equiv$};
\node (not) [not gate US, draw=OliveGreen, fill=OliveGreen!25, right=.2cm of equiv] {};

\draw (x.east) -| +(.2, 0)  |- (nand.input 1);
\draw (x.east) -| +(.2, 0)  |- (nand.input 2);

\end{tikzpicture}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%% AND %%%%%%%%
\begin{tikzpicture}
%%\draw (0,0) -- (0,1cm);
\node (y) at (0,1cm) {$y$};
\node (x) at (0,0cm) {$x$};
\node (nand)[nand gate US, draw, below right=0cm  and .2cm of y] {};
\node (not) [not gate US, draw=OliveGreen, fill=OliveGreen!25, right=.4cm of nand] {};
\node (equiv)[right=.2cm of not] {$\equiv$};
\node (and) [and gate US, draw, right=.2cm of equiv] {};

\draw(nand.output) -- (not.west);
\draw (x.east) -| +(.1, 0)  |- (nand.input 2);
\draw (y.east) -| +(.1, 0)  |- (nand.input 1);

\end{tikzpicture}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%% OR %%%%%%%%
\begin{tikzpicture}
%%\draw (0,0) -- (0,1cm);
\node (y) at (0,1cm) {$y$};
\node (x) at (0,0cm) {$x$};
\node (not1) [not gate US, draw=OliveGreen, fill=OliveGreen!25, right=.4cm of y] {};
\node (not2) [not gate US, draw=OliveGreen, fill=OliveGreen!25, right=.4cm of x] {};
\node (nand)[nand gate US, draw] at (1.75, .5){};

\node (equiv)[right=.2cm of nand] {$\equiv$};
\node (or) [or gate US, draw, , right=.2cm of equiv] {};

\draw (y.east) -| (not1.input);
\draw (x.east) -| (not2.input);
\draw (not1.output) -- +(.1cm, 0) |- (nand.input 1);
\draw (not2.output) -- +(.1cm, 0) |- (nand.input 2);

\end{tikzpicture}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%% E1 %%%%%%%%%%%%%%%%%%%%
\begin{tikzpicture}
\node (x) at (0, 0cm) {$x$};
\node (f)[right=.4cm of x, draw=blue!75, fill=blue!25, text width=.6cm] {$f(\cdot)$};
\node (nand) [nand gate US, draw, above right=0cm and .7cm of f] {};
\node (f2)[right=.4cm of nand, draw=blue!75, fill=blue!25, text width=.6cm] {$f(\cdot)$};

\draw (x.east) -- (f.west);
\draw (x.east) -- +(.2cm, 0) |- (nand.input 1);
\draw (f.east) -- +(.2, 0) |- (nand.input 2);
\draw (nand.output) -- +(.2, 0) |- (f2.west);
\end{tikzpicture}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

 %%%%%%%% TRUE %%%%%%%%%%%%%%%%%%%%
\begin{tikzpicture}
\node (x) at (.2cm, .2cm) (x) {$x$};
\node (not) [not gate US, draw=OliveGreen, fill=OliveGreen!25, right=.4cm of x] {};
\node (nand) [nand gate US, draw, above right=0cm and .7cm of not] {};

\node (equiv)[right=.2cm of nand] {$\equiv$};
\node (true) [draw, right=.2cm of equiv] {TRUE};

\draw (x.east) -- (not.input);
\draw (x.east) -- +(.2cm, 0) |- (nand.input 1);
\draw (not.output) -- +(.2, 0) |- (nand.input 2);
\end{tikzpicture}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

 %%%%%%%% FALSE %%%%%%%%%%%%%%%%%%%%
\begin{tikzpicture}
\node (E1) at (0, 0cm) {$E_1$};
\node (open) [right=-.2cm of E1] {$($};
\node (not0) [not gate US, draw=OliveGreen, fill=OliveGreen!25, right=.1cm of open] {};
\node (close) [right=.1cm of not0] {$)\;\;=$};

\node (x) [right=.2cm of close] {$x$};
\node (not) [not gate US, draw=OliveGreen, fill=OliveGreen!25, right=.4cm of x] {};
\node (nand) [nand gate US, draw, above right=0cm and .7cm of not] {};
\node (not2) [not gate US, draw=OliveGreen, fill=OliveGreen!25, right=.4cm of nand] {};

\node (equiv)[right=.2cm of not2] {$\equiv$};
\node (true) [draw, right=.2cm of equiv] {FALSE};

\draw (x.east) -- (not.input);
\draw (x.east) -- +(.2cm, 0) |- (nand.input 1);
\draw (not.output) -- +(.2, 0) |- (nand.input 2);
\draw (nand.output) -- +(.2, 0) |- (not2.input);
\end{tikzpicture}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%% XOR %%%%%%%%%%%%%%%%%%%%
\begin{tikzpicture}
%%\draw (0,0) -- (0,1cm);
\node (y) at (0,1cm) {$y$};
\node (x) at (0,0cm) {$x$};
\node (nand1)[nand gate US, draw, below right=0cm  and .2cm of y] {};
\node (nand2)[nand gate US, draw, right=1.3cm of y] {};
\node (nand3)[nand gate US, draw, right=1.3cm of x] {};

\node (nand4)[nand gate US, draw, right=1.5cm of nand1] {};

\node (equiv)[right=.2cm of nand4] {$\equiv$};
\node (xor) [xor gate US, draw, right=.2cm of equiv] {};

\draw (x.east) -| +(.1, 0)  |- (nand1.input 2);
\draw (y.east) -| +(.1, 0)  |- (nand1.input 1);
\draw (x.east) -- +(.1cm, 0) |- (nand3.input 2);
\draw (y.east) -- +(.1cm, 0) |- (nand2.input 1);
\draw (nand1.output) -- + (.1cm, 0) |- (nand2.input 2);
\draw (nand1.output) -- +(.1cm, 0) |- (nand3.input 1);
\draw (nand2.output) -- +(.1cm, 0) |- (nand4.input 1);
\draw (nand3.output) -- +(.1cm, 0) |- (nand4.input 2);

\end{tikzpicture}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begingroup\tabcolsep=2pt\def\arraystretch{1}
\begin{table}
\tiny
\begin{tabular}{p{.02\linewidth}p{.95\linewidth}}
 ~ & 
  \begin{tabular}{|m{.5cm}|m{2.5cm}|m{4cm}|}
    \hline
    func. & CL expression & schematic\\
    \hline\hline
  \end{tabular}\\
(a) & 
  \begin{tabular}{|m{.5cm}|m{2.5cm}|m{4cm}|}
    {\color{OliveGreen}NOT} & \color{OliveGreen}{(S NAND I)} & 
    %%%%%%% NOT %%%%%%%%
    \begin{tikzpicture}
      \node (x) at (0,0) {$x$};
      \node (nand)[nand gate US, draw, right=.6cm of x] {};
      \node (equiv)[right=.2cm of nand] {$\equiv$};
      \node (not) [not gate US, draw=OliveGreen, fill=OliveGreen!25, right=.2cm of equiv] {};

      \draw (x.east) -| +(.2, 0)  |- (nand.input 1);
      \draw (x.east) -| +(.2, 0)  |- (nand.input 2);

    \end{tikzpicture}\\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \hline
    %%%%%%% AND %%%%%%%%
    AND & (C B NAND) (B {\color{OliveGreen}(S NAND I)}) \newline
    $\rightarrow$ (C B NAND) (B {\color{OliveGreen}NOT}) & 
    \begin{tikzpicture}
      %%\draw (0,0) -- (0,1cm);
      \node (y) at (0cm,1cm) {$y$};
      \node (x) at (0cm,0cm) {$x$};
      \node (nand) at (1cm, .5cm) [nand gate US, draw] {};
      \node (not) [not gate US, draw=OliveGreen, fill=OliveGreen!25, right=.4cm of nand] {};
      %%\node (equiv)[right=.2cm of not] {$\equiv$};
      %%\node (and) [and gate US, draw, right=.2cm of equiv] {};

      \draw(nand.output) -- (not.west);
      \draw (x.east) -| +(.1, 0)  |- (nand.input 2);
      \draw (y.east) -| +(.1, 0)  |- (nand.input 1);

    \end{tikzpicture}\\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \hline
    %%%%%%% OR %%%%%%%%
    OR & ((B (C (B (S NAND) NAND))) \newline {\color{OliveGreen}(S NAND I)} \newline
    $\rightarrow$ ((B (C (B (S NAND) NAND))) {\color{OliveGreen}NOT} & 
    \begin{tikzpicture}
      %%\draw (0,0) -- (0,1cm);
      \node (y) at (0,1cm) {$y$};
      \node (x) at (0,0cm) {$x$};
      \node (not1) [not gate US, draw=OliveGreen, fill=OliveGreen!25, right=.4cm of y] {};
      \node (not2) [not gate US, draw=OliveGreen, fill=OliveGreen!25, right=.4cm of x] {};
      \node (nand)[nand gate US, draw] at (1.75, .5){};

      %% \node (equiv)[right=.2cm of nand] {$\equiv$};
      %% \node (or) [or gate US, draw, , right=.2cm of equiv] {};

      \draw (y.east) -| (not1.input);
      \draw (x.east) -| (not2.input);
      \draw (not1.output) -- +(.1cm, 0) |- (nand.input 1);
      \draw (not2.output) -- +(.1cm, 0) |- (nand.input 2);

    \end{tikzpicture}\\
    \hline
  \end{tabular}\\
(b) & 
  \begin{tabular}{|m{.5cm}|m{2.5cm}|m{4cm}|}
    \hline
    {\color{red}$E_1$} & {\color{red} S B (S NAND) } & 
    %%%%%%%% E1 %%%%%%%%%%%%%%%%%%%%
    \begin{tikzpicture}
      \node (x) at (0, 0cm) {$x$};
      \node (f)[right=.4cm of x, draw=gray!75, fill=gray!25, text width=.6cm] {$f(\cdot)$};
      \node (nand) [nand gate US, draw, above right=0cm and .7cm of f] {};
      \node (f2)[right=.4cm of nand, draw=gray!75, fill=gray!25, text width=.6cm] {$f(\cdot)$};

      \draw (x.east) -- (f.west);
      \draw (x.east) -- +(.2cm, 0) |- (nand.input 1);
      \draw (f.east) -- +(.2, 0) |- (nand.input 2);
      \draw (nand.output) -- +(.2, 0) |- (f2.west);
    \end{tikzpicture}\\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \hline
    {\color{blue}$E_2$} & {\color{blue}(B (C B NAND) S)} & 
    %%%%%%%% E2 %%%%%%%%%%%
    \begin{tikzpicture}
      \node (x) at (0,0) {$x$};
      \node (f)[above right=.1 and .6cm of x, draw=gray!75, fill=gray!25, text width=.6cm] {$f(\cdot)$};
      \node (g)[below right=.1 and .6cm of x, draw=gray!75, fill=gray!25, text width=.6cm] {$g(\cdot)$};
      \node (nand)[nand gate US, draw, right=1.8cm of x] {};

      \draw (x.east) -| +(.2, 0)  |- (g.west);
      \draw (x.east) -| +(.2, 0)  |- (f.west);
      \draw (f.east) -| +(.3, 0) |-  (nand.input 1);
      \draw (g.east) -| +(.3, 0) |-  (nand.input 2);
    \end{tikzpicture}\\
    %%%%%%%%%%%%%%%%%%%%%%%%%%
    \hline
  \end{tabular}\\
(c) \hspace{-3pt} & 
  \begin{tabular}{|m{.5cm}|m{2.5cm}|m{4cm}|}
    \hline\hline
      TRUE & (S NAND) {\color{OliveGreen}(S NAND I)} \newline
      $\rightarrow$  (S NAND) {\color{OliveGreen}NOT} &
      %%%%%%%% TRUE %%%%%%%%%%%%%%%%%%%%
      \begin{tikzpicture}
        \node (x) at (.2cm, .2cm) (x) {$x$};
        \node (not) [not gate US, draw=OliveGreen, fill=OliveGreen!25, right=.4cm of x] {};
        \node (nand) [nand gate US, draw, above right=0cm and .7cm of not] {};

%%         \node (equiv)[right=.2cm of nand] {$\equiv$};
%%         \node (true) [draw, right=.2cm of equiv] {TRUE};

        \draw (x.east) -- (not.input);
        \draw (x.east) -- +(.2cm, 0) |- (nand.input 1);
        \draw (not.output) -- +(.2, 0) |- (nand.input 2);
      \end{tikzpicture}\\
      %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
      \hline
      %%%%%%%% FALSE %%%%%%%%%%%%%%%%%%%%
      FALSE & {\color{red}(S B (S NAND))} {\color{OliveGreen}(S NAND I)} \newline
      $\rightarrow$ {\color{red}$E_1$}\; {\color{OliveGreen}NOT} 
      & 
      \begin{tikzpicture}
        \node (E1) at (0, 0cm) {{\color{red}$E_1$}};
        \node (open) [right=-.1cm of E1] {$($};
        \node (not0) [not gate US, draw=OliveGreen, fill=OliveGreen!25, right=-.00cm of open] {};
        \node (close) [right=.05cm of not0] {$)$};
        \node (eq) [below=.05cm of open] {$=$};

        \node (x) [right=.2cm of eq] {$x$};
        \node (not) [not gate US, draw=OliveGreen, fill=OliveGreen!25, right=.4cm of x] {};
        \node (nand) [nand gate US, draw, above right=0cm and .7cm of not] {};
        \node (not2) [not gate US, draw=OliveGreen, fill=OliveGreen!25, right=.4cm of nand] {};

%%         \node (equiv)[right=.2cm of not2] {$\equiv$};
%%         \node (true) [draw, right=.2cm of equiv] {FALSE};

        \draw (x.east) -- (not.input);
        \draw (x.east) -- +(.2cm, 0) |- (nand.input 1);
        \draw (not.output) -- +(.2, 0) |- (nand.input 2);
        \draw (nand.output) -- +(.2, 0) |- (not2.input);
      \end{tikzpicture}\\
      %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
      \hline
      %%%%%%%% XOR %%%%%%%%%%%%%%%%%%%%
      XOR & 
      ((S (B C (
      {\color{blue}((B (C B NAND)) S)}
      (
      {\color{blue}((B (C B NAND)) S)}
      (B C ((B (B NAND)) NAND)))))) I) \newline
      $\rightarrow$
      ((S (B C (
      {\color{blue}$E_2$}
      (
      {\color{blue}$E_2$}
      (B C ((B (B NAND)) NAND)))))) I)
      &
      \begin{tikzpicture}
        %%\draw (0,0) -- (0,1cm);
        \node (y) at (0,1cm) {$y$};
        \node (x) at (0,0cm) {$x$};
        \node (nand1)[nand gate US, draw, below right=0cm  and .2cm of y] {};
        \node (nand2)[nand gate US, draw, right=1.3cm of y] {};
        \node (nand3)[nand gate US, draw, right=1.3cm of x] {};

        \node (nand4)[nand gate US, draw, right=1.5cm of nand1] {};

%%         \node (equiv)[right=.2cm of nand4] {$\equiv$};
%%         \node (xor) [xor gate US, draw, right=.2cm of equiv] {};

        \draw (x.east) -| +(.1, 0)  |- (nand1.input 2);
        \draw (y.east) -| +(.1, 0)  |- (nand1.input 1);
        \draw (x.east) -- +(.1cm, 0) |- (nand3.input 2);
        \draw (y.east) -- +(.1cm, 0) |- (nand2.input 1);
        \draw (nand1.output) -- + (.1cm, 0) |- (nand2.input 2);
        \draw (nand1.output) -- +(.1cm, 0) |- (nand3.input 1);
        \draw (nand2.output) -- +(.1cm, 0) |- (nand4.input 1);
        \draw (nand3.output) -- +(.1cm, 0) |- (nand4.input 2);

      \end{tikzpicture}\\
      %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \end{tabular}
  \hline
\end{tabular}
\end{table}
\end{group}

\end{document}

### 2013-03-19 drafts/ijcai2013/ijcai13guidelines/ijcai13.tex
Added more misc files to git.

%%%% ijcai11.tex

\typeout{IJCAI-13 Instructions for Authors}

% These are the instructions for authors for IJCAI-13.
% They are the same as the ones for IJCAI-11 with superficical wording
%   changes only.

\documentclass{article}
% The file ijcai13.sty is the style file for IJCAI-13 (same as ijcai07.sty).
\usepackage{ijcai13}

% Use the postscript times font!
\usepackage{times}

% the following package is optional:
%\usepackage{latexsym}

% Following comment is from ijcai97-submit.tex:
% The preparation of these files was supported by Schlumberger Palo Alto
% Research, AT\&T Bell Laboratories, and Morgan Kaufmann Publishers.
% Shirley Jowell, of Morgan Kaufmann Publishers, and Peter F.
% Patel-Schneider, of AT\&T Bell Laboratories collaborated on their
% preparation.

% These instructions can be modified and used in other conferences as long
% as credit to the authors and supporting agencies is retained, this notice
% is not changed, and further modification or reuse is not restricted.
% Neither Shirley Jowell nor Peter F. Patel-Schneider can be listed as
% contacts for providing assistance without their prior permission.

% To use for other conferences, change references to files and the
% conference appropriate and use other authors, contacts, publishers, and
% organizations.
% Also change the deadline and address for returning papers and the length and
% page charge instructions.
% Put where the files are available in the appropriate places.

\title{IJCAI--13 Formatting Instructions\thanks{These match the formatting instructions of IJCAI-07. The support of IJCAI, Inc. is acknowledged.}}
\author{Francesca Rossi \\
University of Padova\\
Italy \\
pcchair13@ijcai.org}

\begin{document}

\maketitle

\begin{abstract}
  The {\it IJCAI--13 Proceedings} will be printed from electronic
  manuscripts submitted by the authors. The electronic manuscript will
  also be included in the online version of the proceedings. This paper
  provides the style instructions.
\end{abstract}

\section{Introduction}

The {\it IJCAI--13 Proceedings} will be printed from electronic
manuscripts submitted by the authors. These must be PDF ({\em Portable
Document Format}) files formatted for 8-1/2$''$ $\times$ 11$''$ paper.

\subsection{Length of Papers}

Each accepted full paper is allocated six pages in the conference 
proceedings, excluded references. References can take up to one page.
Up to two additional pages may be purchased at a price 
of \$275 per page for any accepted paper. However, all 
{\em submissions} must 
be a maximum of six pages, plus at most one for references, in length.

\subsection{Word Processing Software}

As detailed below, IJCAI has prepared and made available a set of
\LaTeX{} macros and a Microsoft Word template for use in formatting
your paper. If you are using some other word processing software (such
as WordPerfect, etc.), please follow the format instructions given
below and ensure that your final paper looks as much like this sample
as possible.

\section{Style and Format}

\LaTeX{} and Word style files that implement these instructions
can be retrieved electronically. (See Appendix~\ref{stylefiles} for
instructions on how to obtain these files.)

\subsection{Layout}

Print manuscripts two columns to a page, in the manner in which these
instructions are printed. The exact dimensions for pages are:
\begin{itemize}
\item left and right margins: .75$''$
\item column width: 3.375$''$
\item gap between columns: .25$''$
\item top margin---first page: 1.375$''$
\item top margin---other pages: .75$''$
\item bottom margin: 1.25$''$
\item column height---first page: 6.625$''$
\item column height---other pages: 9$''$
\end{itemize}

All measurements assume an 8-1/2$''$ $\times$ 11$''$ page size. For
A4-size paper, use the given top and left margins, column width,
height, and gap, and modify the bottom and right margins as necessary.

\subsection{Format of Electronic Manuscript}

For the production of the electronic manuscript, you must use Adobe's
{\em Portable Document Format} (PDF). A PDF file can be generated, for
instance, on Unix systems using {\tt ps2pdf} or on Windows systems
using Adobe's Distiller. There is also a website with free software
and conversion services: {\tt http://www.ps2pdf.com/}. For reasons of
uniformity, use of Adobe's {\em Times Roman} font is strongly suggested. In
\LaTeX2e{}, this is accomplished by putting
\begin{quote} 
\mbox{\tt $\backslash$usepackage\{times\}}
\end{quote}
in the preamble.\footnote{You may want also to use the package {\tt
latexsym}, which defines all symbols known from the old \LaTeX{}
version.}

Additionally, it is of utmost importance to specify the American {\bf
letter} format (corresponding to 8-1/2$''$ $\times$ 11$''$) when
formatting the paper. When working with {\tt dvips}, for instance, one
should specify {\tt -t letter}.

\subsection{Title and Author Information}

Center the title on the entire width of the page in a 14-point bold
font. Below it, center the author name(s) in a 12-point bold font, and
then center the address(es) in a 12-point regular font. Credit to a
sponsoring agency can appear on the first page as a footnote.

\subsubsection{Blind Review}

In order to make blind reviewing possible, authors must omit their
names and affiliations when submitting the paper for review. In place
of names and affiliations, provide a list of content areas. When
referring to one's own work, use the third person rather than the
first person. For example, say, ``Previously,
Gottlob~\shortcite{gottlob:nonmon} has shown that\ldots'', rather
than, ``In our previous work~\cite{gottlob:nonmon}, we have shown
that\ldots'' Try to avoid including any information in the body of the
paper or references that would identify the authors or their
institutions. Such information can be added to the final camera-ready
version for publication.

\subsection{Abstract}

Place the abstract at the beginning of the first column 3$''$ from the
top of the page, unless that does not leave enough room for the title
and author information. Use a slightly smaller width than in the body
of the paper. Head the abstract with ``Abstract'' centered above the
body of the abstract in a 12-point bold font. The body of the abstract
should be in the same font as the body of the paper.

The abstract should be a concise, one-paragraph summary describing the
general thesis and conclusion of your paper. A reader should be able
to learn the purpose of the paper and the reason for its importance
from the abstract. The abstract should be no more than 200 words long.

\subsection{Text}

The main body of the text immediately follows the abstract. Use
10-point type in a clear, readable font with 1-point leading (10 on
11).

Indent when starting a new paragraph, except after major headings.

\subsection{Headings and Sections}

When necessary, headings should be used to separate major sections of
your paper. (These instructions use many headings to demonstrate their
appearance; your paper should have fewer headings.)

\subsubsection{Section Headings}

Print section headings in 12-point bold type in the style shown in
these instructions. Leave a blank space of approximately 10 points
above and 4 points below section headings.  Number sections with
arabic numerals.

\subsubsection{Subsection Headings}

Print subsection headings in 11-point bold type. Leave a blank space
of approximately 8 points above and 3 points below subsection
headings. Number subsections with the section number and the
subsection number (in arabic numerals) separated by a
period.

\subsubsection{Subsubsection Headings}

Print subsubsection headings in 10-point bold type. Leave a blank
space of approximately 6 points above subsubsection headings. Do not
number subsubsections.

\subsubsection{Special Sections}

You may include an unnumbered acknowledgments section, including
acknowledgments of help from colleagues, financial support, and
permission to publish.

Any appendices directly follow the text and look like sections, except
that they are numbered with capital letters instead of arabic
numerals.

The references section is headed ``References,'' printed in the same
style as a section heading but without a number. A sample list of
references is given at the end of these instructions. Use a consistent
format for references, such as that provided by Bib\TeX{}. The reference
list should not include unpublished work.

\subsection{Citations}

Citations within the text should include the author's last name and
the year of publication, for example~\cite{gottlob:nonmon}.  Append
lowercase letters to the year in cases of ambiguity.  Treat multiple
authors as in the following examples:~\cite{abelson-et-al:scheme}
or~\cite{bgf:Lixto} (for more than two authors) and
\cite{brachman-schmolze:kl-one} (for two authors).  If the author
portion of a citation is obvious, omit it, e.g.,
Nebel~\shortcite{nebel:jair-2000}.  Collapse multiple citations as
follows:~\cite{gls:hypertrees,levesque:functional-foundations}.
\nocite{abelson-et-al:scheme}
\nocite{bgf:Lixto}
\nocite{brachman-schmolze:kl-one}
\nocite{gottlob:nonmon}
\nocite{gls:hypertrees}
\nocite{levesque:functional-foundations}
\nocite{levesque:belief}
\nocite{nebel:jair-2000}

\subsection{Footnotes}

Place footnotes at the bottom of the page in a 9-point font.  Refer to
them with superscript numbers.\footnote{This is how your footnotes
should appear.} Separate them from the text by a short
line.\footnote{Note the line separating these footnotes from the
text.} Avoid footnotes as much as possible; they interrupt the flow of
the text.

\section{Illustrations}

Place all illustrations (figures, drawings, tables, and photographs)
throughout the paper at the places where they are first discussed,
rather than at the end of the paper. If placed at the bottom or top of
a page, illustrations may run across both columns.

Illustrations must be rendered electronically or scanned and placed
directly in your document. All illustrations should be in black and
white, as color illustrations may cause problems. Line weights should
be 1/2-point or thicker. Avoid screens and superimposing type on
patterns as these effects may not reproduce well.

Number illustrations sequentially. Use references of the following
form: Figure 1, Table 2, etc. Place illustration numbers and captions
under illustrations. Leave a margin of 1/4-inch around the area
covered by the illustration and caption.  Use 9-point type for
captions, labels, and other text in illustrations.

\section*{Acknowledgments}

The preparation of these instructions and the \LaTeX{} and Bib\TeX{}
files that implement them was supported by Schlumberger Palo Alto
Research, AT\&T Bell Laboratories, and Morgan Kaufmann Publishers.
Preparation of the Microsoft Word file was supported by IJCAI.  An
early version of this document was created by Shirley Jowell and Peter
F. Patel-Schneider.  It was subsequently modified by Jennifer
Ballentine and Thomas Dean, Bernhard Nebel, and Daniel Pagenstecher.
These instructions are the same as the ones for IJCAI--05, prepared by
Kurt Steinkraus, Massachusetts Institute of Technology, Computer
Science and Artificial Intelligence Lab.

\appendix

\section{\LaTeX{} and Word Style Files}\label{stylefiles}

The \LaTeX{} and Word style files are available on the IJCAI--13
website, {\tt http://www.ijcai-13.org/}.
These style files implement the formatting instructions in this
document.

The \LaTeX{} files are {\tt ijcai13.sty} and {\tt ijcai13.tex}, and
the Bib\TeX{} files are {\tt named.bst} and {\tt ijcai13.bib}. The
\LaTeX{} style file is for version 2e of \LaTeX{}, and the Bib\TeX{}
style file is for version 0.99c of Bib\TeX{} ({\em not} version
0.98i). The {\tt ijcai13.sty} file is the same as the {\tt
ijcai07.sty} file used for IJCAI--07.

The Microsoft Word style file consists of a single file, {\tt
ijcai13.doc}. This template is the same as the one used for
IJCAI--07.

These Microsoft Word and \LaTeX{} files contain the source of the
present document and may serve as a formatting sample.

Further information on using these styles for the preparation of
papers for IJCAI--13 can be obtained by contacting {\tt
pcchair13@ijcai.org}.

%% The file named.bst is a bibliography style file for BibTeX 0.99c
\bibliographystyle{named}
\bibliography{ijcai13}

\end{document}

### 2013-03-19 src/experiments/Timing/EnumWMemo.hs

@@ -1,0 +1,11 @@
+
+import Control.DeepSeq
+import Type
+import StdLib
+import Enumerate
+
+out = map runTI $ enumCombsToProb stdlibTrie (-200) 4 Rtype
+
+main = do
+  putStrLn $ --(any (==(head out)) out) `seq` 
+               "Finished. Total length " ++ show (length out)

### 2013-03-23 drafts/grammar/grammar.tex
Added section about Bayesian intepretation of E-C.

@@
 \usepackage{tikz}
+\usepackage{tikz}
+\usetikzlibrary{bayesnet}
 \usetikzlibrary{arrows,automata, positioning, patterns,backgrounds}
@@
 
+\DeclareMathOperator*{\argmin}{arg\,min}
+\DeclareMathOperator*{\argmax}{arg\,max}
+
+\author{Eyal Dechter}
 \begin{document}
-
-In this document, I will describe a stochastic grammar over typed
-combinatory logic expressions.
+\maketitle
+\date
 
@@
 
-\bibliographystyle{plain}
-\bibliography{grammar}
+\section{E-C as MAP inference in a probabilistic graphical model.}
+The motivation of the E-C algorithm is that a learning agent should
+learn a library of useful program fragments that represents knowledge
+about the kind the programs that are likely to solve problems in a
+given domain. One way to frame this is as a probabilistic generative
+model of the tasks that an agent encounters
+(Figure~\ref{fig:generativeModel}).
+
+\begin{figure}
+\begin{tikzpicture}[->, circle]
+
+\node[latent] (theta) {$\theta$} ;
+\node[latent, below=of theta] (G) {G} ;
+\node[latent, below=of G]     (e) {$e_i$} ;
+\node[obs, below=of e]     (t) {$t_i$} ;
+
+\edge {theta} {G};
+\edge {G} {e};
+\edge {e} {t};
+
+\plate {plate} {
+  (e)(t)}{$i=1, \dots, N$};
+
+\end{tikzpicture}
+\caption{Generative model of program induction
+  tasks. \label{fig:generativeModel}}
+\end{figure}
+
+Here each task $t_i$ is interpreted as a noisy version of the
+underlying program $e_i$, and there is an associated likelihood
+function $p_{t|e}(t|e) \propto f(t(e)) $ where $f$ is a monotonically
+decreasing function. Given a task-specific loss function $t(\cdot)$, a
+natural likelihood is $p_{t|e}(t|e) \propto \exp^{-t(e)}$. A set of
+programs $\{e_i\}$ is drawn from a grammar $G$, which functions as the
+subroutine library. Our prior expectations about $G$ are encoded in
+$\theta$ via the distribution $p_{G|\theta}$.
+
+The goal of the learner is the find the library $G$ and programs
+$\{e_i\}$ that maximize $P(\{e_i\}, G | \{t_i\}, \theta)$. That is,
+
+\begin{align}
+G^*, \{e^*_i\} &= \argmax_{G, \{e_i\} }
+   P(\{e_i\}, G | \{t_i\}, \theta)\\
+&= \argmax_{G, \{e_i\} } P(\{t_i\} | \{e_i\})
+  P(\{e_i\} | G) P(G | \theta)\\
+&= \argmax_{G, \{e_i\} }
+  \sum_i \left( 
+  \log{p_{t|e}(t_i|e_i)}+ 
+  \log{p_{e|G}(e_i|G)}
+  \right ) +
+  \log{p_{G|\theta}(G|\theta)}
+\label{eq:objective}
+\end{align}
+
+The main difficulty in this optimization is that finding any candidate
+program $e_i$ that provides a significant value for the first term is
+difficult. If we new $G$, we could sample programs from it as a way to
+guide searching over the $e_i$. This suggests the following iterative
+heuristic search algorithm. At iteration $j$, let $G^{(j)}$ be the
+current best value of $G$.
+
+\begin{enumerate}
+\item \textbf{Exploration step: } Let $F$ be a set of $T$ samples
+  i.i.d from $p_{e|G}(\cdot | G^{(j)})$.
+\item \textbf{Compression step: } 
+$$G^{(j+1)}, \{e^{(j+1)}_i\} = 
+\argmax_{G, \{e_i | e_i \in F \}}
+\sum_i \left( 
+\log{p_{t|e}(t_i|e_i)}+ 
+\log{p_{e|G}(e_i|G)}
+\right ) +
+\log{p_{G|\theta}(G|\theta)}
+$$
+\end{enumerate}
+
+%% We can break this joint maximization into an iteration of
+%% coordinate-wise maximization steps. First we fix $G$ and maximize the
+%% programs, then we fix the programs and maximize $G$, repeating until
+%% we are no longer improving out objective. This gives us the following
+%% iterative algorithm, roughly corresponding to the indicated steps of
+%% the E-C algorithm:
+
+%% \begin{enumerate}
+%% \item \textbf{Compression step} $$G^{(j+1)} = 
+%%   \argmax_{G} \sum_i \left( 
+%%   \log{p_{e|G}(e^{(j)}_i|G)}
+%%   \right ) +
+%%   \log{p_{G|\theta}(G|\theta)}$$
+%% \item \textbf{Exploration step} $$\{e^{(j+1)}_i\} = 
+%%   \argmax_{\{e_i\}}
+%%   \sum_i \left( 
+%%   \log{p_{t|e}(t_i|e_i)}+ 
+%%   \log{p_{e|G}(e_i|G^{(j)})}
+%%   \right ) $$
+%% \end{enumerate}
+
+\bibliographystyle{plain} \bibliography{grammar}
 

### 2013-03-29 drafts/grammar/grammar.tex
Cleanup.
Added section in grammar.tex on online hierarchical program induction.

@@
 \usepackage{ amssymb }
+\usepackage{algorithm}
+\usepackage{algpseudocode}
 \usepackage{graphicx}
 \usepackage{syntax}
-\usepackage{tikz}
 \usepackage{tikz}
@@
 
-\DeclareMathOperator*{\argmin}{arg\,min}
-\DeclareMathOperator*{\argmax}{arg\,max}
+\DeclareMathOperator*{\argmin}{argmin}
+\DeclareMathOperator*{\argmax}{argmax}
 
@@
 
+\begin{algorithm}
+\caption{Hierarchical program induction (E-C)}\label{alg:basic-progind}
+\begin{algorithmic}[1]
+\State  \textbf{Exploration step: } Let $F$ be a set of $T$ samples
+  i.i.d from $p_{e|G}(\cdot | G^{(j)})$.
+\State \textbf{Compression step: } 
+\begin{align*}
+G^{(j+1)}, \{e^{(j+1)}_i\} &= 
+\argmax_{G, \{e_i | e_i \in F \}}
+\sum_i \left( 
+\log{p_{t|e}(t_i|e_i)}+ 
+\log{p_{e|G}(e_i|G)} \right )\\
+&+
+\log{p_{G|\theta}(G|\theta)}
+\end{align*}
+\end{algorithmic}
+\end{algorithm}
+
 \begin{enumerate}
@@
 \end{enumerate}
+
+\subsection{An online algorithm}
+
+Here we present an online version of the algorithm that only chooses
+an expression for each task once. We can view the problem as MAP
+estimation of the unobserved variables in the following graphical
+model:
+
+\begin{figure}[h]
+\begin{tikzpicture}[->, circle]
+
+\node[obs] (theta) {$\theta$} ;
+\node[latent, below=of theta] (G) {G} 
+   child {node [latent] (e1) {$e_1$} 
+      child {node [obs] (t1) {$t_1$}}}
+   child {node [latent] (e2) {$e_2$} 
+      child {node [obs] (t2) {$t_2$}}}
+   child {node [latent] (e3) {$e_3$} 
+      child {node [obs] (t3) {$t_3$}}}
+   child {node [latent] (e4) {$e_4$} 
+      child {node [obs] (t5) {$t_4$}}}
+   child {node  (edots) {$\dots$} edge from parent[draw=none] 
+      child {node  (tdots) {$\dots$} edge from parent[draw=none]}};
+
+\edge {theta} {G};
+
+\end{tikzpicture}
+\end{figure}
+
+We want to choose $e_i$ and $G$ to maximize our objective function. We
+will perform the following iterative procedure. Given a current
+setting of $G$, set $e_i$ to optimize the weights on the edges to
+$t_i$ and $G$. Update $G$ to optimize the weight on the edge to
+$\theta$ and all the edges to previous set expressions $e_j, j\leq
+i$.
+
+\begin{algorithm}
+\caption{Online hierarchical program induction}\label{alg:online-progind}
+\begin{algorithmic}
+\State Initialize $G_0$.
+\For{$i\gets 1, \dots$}
+\State $e_i^* = \argmax_{e} \log p(t_i|e_i) + \log p(e_i|G_{i-1}) $
+\State $G_i = \argmax_{G} \log p(G|\theta) + \sum_{j \leq i} \log p(e_i|G) $
+\EndFor
+\end{algorithmic}
+\end{algorithm}
 

### 2013-04-02 drafts/grammar/grammar.tex
Added section on combining hierarchical program induction & planning.

@@
 \begin{algorithm}
-\caption{Online hierarchical program induction}\label{alg:online-progind}
+\caption{Online hierarchical program induction (version 1)}\label{alg:online-progind-v1}
 \begin{algorithmic}
@@
 \State $e_i^* = \argmax_{e} \log p(t_i|e_i) + \log p(e_i|G_{i-1}) $
-\State $G_i = \argmax_{G} \log p(G|\theta) + \sum_{j \leq i} \log p(e_i|G) $
+\State $G_i = \argmax_{G} \log p(G|\theta) + \sum_{j \leq i} \log p(e_j|G) $
+\EndFor
+\end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}
+\caption{Online hierarchical program induction (version 2)}\label{alg:online-progind-v2}
+\begin{algorithmic}
+\State Initialize $G_0$.
+\For{$i\gets 1, \dots$}
+\State $F_i \sim \text{ i.i.d } p(\cdot | G)$
+\State $e_i^* = \argmax_{e\in F_i} \log p(t_i|e_i) + \log p(e_i|G_{i-1}) $
+\State $G_i = \argmax_{G} \log p(G|\theta) + \sum_{j \leq i} \log p(e_j|G) $
+\EndFor
+\end{algorithmic}
+\end{algorithm}
+
+\section{Combining planning with hierarchical program induction}
+There is something clearly missing from this learning paradigm?
+Actions, framed as programs acting on the environment, are not
+proposed in a way that is contingent on the environment. But of course
+the actions that we make are contingent on past behavior and sensory
+input. A fundamental question, then, is how we can incorporate sensory
+feedback and past behvarior into this learning paradigm.
+
+One answer to this is to combine planning with hierarchical program
+induction. In this view, a solution to a problem is not a single
+program. It is a series of programs, each of which attempts to get the
+planner closer to the goal. The goal of the learner is to a learn a
+distribution over programs that takes as input a sequence of past
+actions and their effects and produces a new sequence of actions.
+
+Let's consider a concrete example of such a domain: consider the
+classic AI problem of stacking blocks to build a tower of a given
+height. Let's make this problem somewhat more challenging by having
+the agent build this tower on a vibrating platform. We define the
+domain of primitive actions to be placing a block in either horizontal
+or vertical orientation anywhere in a two dimensional vertical
+plane. At any point during an attempted execution, the state of the
+system is the sequence of previous primitive actions, and, perhaps,
+auxiliary sensory information: in this case, an agent may make use of
+the current configuration of blocks (i.e. what the tower looks like)
+or some summary statistic thereof. We further assume that there is
+some heuristic function (e.g. height) which guides the choice of the
+next state.
+
+Let $S$ be a state space and let $h:S \rightarrow \mathbb{R}$ be a
+reward function. The planner's goal is to maximize the expected change
+in reward of any given action $e \in \mathcal{L}$. Of course, the
+change in reward does not only depend on the current action; it also
+depends on the current state. But to simplify matters we will ignore
+that for now, and assume that the state dependence is subsumed in the
+distribution $p(\Delta h | e)$. That is, the planner wants to find
+grammar $G^*$ such that
+\begin{align}
+G^* &= \argmax_G \sum_e \Delta h \;p(\Delta h | e) \, p(e|G) \, p(G|\theta).
+\end{align}
+
+As a first attempt to solve this problem, we can simply splice
+Algorithm~\ref{alg:online-progind-v2} with a greedy actor who always
+picks the best available action at any moment. Here, however, there is
+only one ``task,'' and we are maximizing over only the grammar. This
+is described in Algorithm~\ref{alg:planning-progind-v1}.
+
+\begin{algorithm}
+\caption{Planning with online program induction (version
+  1)}\label{alg:planning-progind-v1}
+\begin{algorithmic}
+\State Initialize $G_0$.
+\For{$i\gets 1, \dots$}
+\State $F_i \sim \text{ i.i.d } p_{e|G}(\cdot | G_i)$
+\State $G_i = \argmax_{G} \log p(G|\theta) 
++ \log \sum_{e_j \in F_0 \cup \dots \cup F_i}
+\prod_{e_j} \Delta h(e_j) \; p(e_j|G) $
+\State $e_i^* = \argmax_{F_i}$
 \EndFor

### 2013-04-08 drafts/general_writeup/writeup.tex
Added brief description of holes to general writeup.

@@
 
+\section{Holes}
+In real-world applications, many of the useful degrees of freedom are
+continuous, and it would be useful to have parameters that could be
+fit in a systematic way. Doing this requires a small extension to the
+framework presented so far: we add a special type of construct as a
+primitive of the language. We call these holes. Each hole is a
+placeholder for values of a specific type, within a specific
+domain. For example, a continuous hole is bounded within a continuous
+integer and has the type of a real value.
+
+A combinator with no holes is filled.
+
 \bibliographystyle{plain}

### 2013-04-18 src/Strappy/Sample.hs
Added Sample.hs and Utils.hs.

@@ -1,0 +1,44 @@
+
+module Strappy.Sample where
+
+import Prelude hiding (flip)
+import Control.Monad.State
+import Control.Monad.Random
+
+import Strappy.CL 
+import Strappy.Type
+import qualified Strappy.CombMap as CM
+import Strappy.CombMap (CombMap)
+import Strappy.Grammar
+import Strappy.Utils 
+
+
+typeOfApp :: Monad m => Comb -> Comb -> StateT Int m Type
+typeOfApp c_left c_right 
+    = do t <- newTVar Star 
+         case mgu (cType c_left) (cType c_right ->- t) of 
+           (Just sub) -> return $ toType (apply sub (cType c_left))
+           Nothing -> error $ "typeOfApp: cannot unify " ++
+                      show c_left ++ " with " ++ show c_right
+             
+sampleFromGrammar :: (MonadPlus m, MonadRandom m) =>
+                     Grammar -- ^ stochastic grammar G
+                  -> Type -- ^ type t
+                  -> StateT Int m Comb
+-- | Samples a combinator of type t from a stochastic grammar G. 
+sampleFromGrammar gr tp = do shouldExpand <- flip $ expansions gr
+                             case shouldExpand of
+                               True -> do t <- newTVar Star
+                                          c_left <- sampleFromGrammar gr (t ->- tp)
+                                          c_right <- sampleFromGrammar gr (fromType (cType c_left))
+                                          return $ c_left `app'` c_right  
+                               False -> do  let  lib = library gr
+                                            let cs = runStateT (filterCombinatorsByType (CM.keys lib) tp) 0
+                                            let dist = [(lib CM.! c, c ) | (c, i) <- cs]
+                                                z = logsumexp . map fst $ dist
+                                                dist' = map (\(x, y) -> (exp (x - z), y)) $ dist 
+                                            guard (not . null $ dist)
+                                            sampleMultinomial  dist'
+                                 
+                                                              
+                                          

### 2013-04-18 src/Strappy/Utils.hs

@@ -1,0 +1,15 @@
+
+module Strappy.Utils where
+
+import Prelude hiding (flip)
+import Control.Monad.Random 
+
+flip :: (Num a, Ord a, Random a, MonadRandom m) => a -> m Bool
+flip p = do r <- getRandomR (0, 1)
+            return $ r < p
+
+sampleMultinomial :: (Num a, Ord a, Random a, MonadRandom m) => [(a, b)] -> m b
+sampleMultinomial dist = do r <- getRandomR (0, 1)
+                            return $ sample r dist
+    where sample r ((p, a):rest) = if r <= p then a else 
+                                       sample (r - p) rest

### 2013-04-18 drafts/grammar/grammar.tex

@@
 \renewcommand{\grammarlabel}[2]{#1 #2}
+\renewcommand{\given}{\, | \,}
 
@@
 underlying program $e_i$, and there is an associated likelihood
-function $p_{t|e}(t|e) \propto f(t(e)) $ where $f$ is a monotonically
+function $p_{t\given e}(t \given e) \propto f(t(e)) $ where $f$ is a monotonically
 decreasing function. Given a task-specific loss function $t(\cdot)$, a
-natural likelihood is $p_{t|e}(t|e) \propto \exp^{-t(e)}$. A set of
+natural likelihood is $p_{t \given e}(t \given e) \propto \exp^{-t(e)}$. A set of
 programs $\{e_i\}$ is drawn from a grammar $G$, which functions as the
 subroutine library. Our prior expectations about $G$ are encoded in
-$\theta$ via the distribution $p_{G|\theta}$.
+$\theta$ via the distribution $p_{G \given \theta}$.
 
 The goal of the learner is the find the library $G$ and programs
-$\{e_i\}$ that maximize $P(\{e_i\}, G | \{t_i\}, \theta)$. That is,
+$\{e_i\}$ that maximize $P(\{e_i\}, G \given  \{t_i\}, \theta)$. That is,
 
@@
 G^*, \{e^*_i\} &= \argmax_{G, \{e_i\} }
-   P(\{e_i\}, G | \{t_i\}, \theta)\\
-&= \argmax_{G, \{e_i\} } P(\{t_i\} | \{e_i\})
-  P(\{e_i\} | G) P(G | \theta)\\
+   P(\{e_i\}, G \given  \{t_i\}, \theta)\\
+&= \argmax_{G, \{e_i\} }  P(\{t_i\} \given \{e_i\}) \;
+  P(\{e_i\} \given G) \; P(G \given \theta)\\
 &= \argmax_{G, \{e_i\} }
   \sum_i \left( 
-  \log{p_{t|e}(t_i|e_i)}+ 
-  \log{p_{e|G}(e_i|G)}
+  \log{p_{t\given e}\;(t_i \given e_i)}+ 
+  \log{p_{e \given G}(e_i \given G)}
   \right ) +
-  \log{p_{G|\theta}(G|\theta)}
+  \log{p_{G \given \theta}\; (G \given \theta)}
 \label{eq:objective}
@@
 program $e_i$ that provides a significant value for the first term is
-difficult. If we new $G$, we could sample programs from it as a way to
-guide searching over the $e_i$. This suggests the following iterative
+difficult. If we knew $G$, we could sample programs from it as a way to
+guide the search over the $e_i$. This suggests the following iterative
 heuristic search algorithm. At iteration $j$, let $G^{(j)}$ be the

### 2013-04-20 drafts/ijcai2013/ijcai13.tex
Cleanup and add figures directory.
Changes to ijcai13 tex file.

@@
  show that the learner discovers modular concepts for these
- domains. Whereas the learner is able to solve almost none of the
- posed problems in the procedure's first iteration, it rapidly becomes
- able to solve a large number by gaining abstract knowledge of the
- structure of the solution space.
+ domains. Whereas the learner is initially unable to solve almost any
+ of the posed problems, it quickly acquires the ability to solve many
+ problems by gaining abstract knowledge of the structure of the
+ solution space.
 
@@
 \section{Introduction}
-Many of the problems we want to solve in AI are best cast as search
-problems. There are many search algorithms, but they typically exploit
+A central question for AI is how to create intelligent agents without
+building in vast amounts of pre-specified knowledge that is tailored
+to one domain or another. Any attempt to solve this problem must
+contend with the fact that the less an agent knows about solving
+problems in a domain, the worse its attempts are, and the less it
+learns about how to solve subsequent problems. For example, imagine
+you know nothing about baking but are supplied with a stocked kitchen:
+your initial attempts will be uniformly disastrous, and you will learn
+little from them, except that you did not create a baked good. A
+little bit of knowledge about baking would go a long way towards
+making your initial attempts more informative.
+
+Like many of the problems in AI, this one can be cast as a search
+problem. There are many search algorithms, but they typically exploit
 local or topological structure in the search space. For example,
@@
 
-However, many of the problems we want solve -- discovering scientific
-theories, developing new technologies, figuring out how to bake -- do
-not come with such a structured search space. This is particularly
-true of problems, such as those mentioned, whose solutions can be
-naturally framed as programs. Here we ask how to extract structure
-from the space of programs by attempting to solve multiple related
-problems in a given domain at once.
-
-We take as inspiration the programmer who -- having reused several
-subroutines to solve different but similar simple tasks -- creates a
-function library to help manage the complexity of larger projects. The
-EC algorithm, which we introduce here, builds a library of reusable
-program components and places a distribution over these, effectively
-learning from simple tasks a search space structure that enables it
-to solve more complex tasks.
+The absense of domain specific knowledge makes exploiting structure
+difficult, especially if the structure of the search space is not
+smooth. This is particularly true of problems -- such as discovering
+scientific theories, developing new technologies, learning how to bake
+-- whose solutions can be naturally framed as programs. Here we ask
+how to bootstrap domain specific knowledge and extract structure from
+the space of programs by attempting to solve multiple related problems
+in a given domain at once.
+
+Out approach is inspired by the programmer who -- having reused
+several subroutines to solve different but similar simple tasks --
+creates a function library to help manage the complexity of larger
+projects. The E.C. algorithm, which we introduce here, builds a library
+of reusable program components and places a distribution over these,
+effectively learning from simple tasks a search space structure that
+enables it to solve more complex tasks.
 
@@
 
-Concretely, the EC algorithm works as follows. It begins with a small
+Concretely, the E.C. algorithm works as follows. It begins with a small
 library of program primitives and a distribution over programs
@@
 The remainder of this paper will include two major sections. In the
-first, we describe in detail the EC algorithm and design choices we
+first, we describe in detail the E.C. algorithm and design choices we
 make in implementing it. We will then present results from two
@@
 
-We propose the \textbf{EC algorithm}, an iterative procedure to solve this
+We propose the \textbf{E.C. algorithm}, an iterative procedure to solve this
 problem. It will be useful in presenting the algorithm to define the
@@
 
-The EC algorithm maintains a distribution $\mathcal{D}$ over
+The E.C. algorithm maintains a distribution $\mathcal{D}$ over
 expressions in $\mathcal{L}$. At each iteration, it
@@
 
-Our best-first policy is to expand the next partial solution
-with the highest value. If a partial solution has multiple leaves that
-are not primitive combinators, we expand the left most one first,
-since it is the choice of left combinator that defines the type of its
-right combinator in an application pair. That is, the parent node
-applies its constraint on valid solutions via the left child, not the
-right child. We want to apply constraints as early as possible, so
-expanding left child first prunes more partial solutions.
+Our best-first policy is to expand the next partial solution with the
+highest value. If a partial solution has multiple leaves that are not
+primitive combinators, we expand the left most one first, since it is
+the choice of left combinator that defines the type of its right
+combinator in an application pair. That is, the parent node applies
+its constraint on valid solutions via the left child, not the right
+child. We want to apply constraints as early as possible, so we expand
+the left child first to prune more partial solutions.
 
@@
 \vspace{-.3cm}
-The top row consists of one type of representation of an integer
+
+\noindent The top row consists of one type of representation of an integer
 squared, namely, the integer times itself. The bottom row contains
@@
 once in the grammar and b) every rule is used more than once. From a
-set of expressions we generate a grammar according to these criteria.
+set of expressions, we generate a grammar according to these criteria.
 
@@
 \centering
-\includegraphics[width=.8\linewidth]{figures/symreg3frontiers.pdf}
+\includegraphics[width=1.0\linewidth]{figures/symreg3frontiers.pdf}
 \vspace{-.2cm}
@@
 \begin{figure}[ht]
-\centering
-\includegraphics[width=.8\linewidth]{figures/symRegDiffCurricula.pdf}
+\includegraphics[width=1.07\linewidth]{figures/symRegDiffCurricula.pdf}
 \caption{How do different task sets affect learning curves? Learning
@@
 \label{fig:symregCurricula}
-\vspace{-.5cm}
+
 \end{figure}
@@
 of $10000$ and runs for $15$ iterations. Thus, if the benefit accruing
-to the runs with larger frontier sizes is due to simply an increase in
-the number of expressions seen, rather than increased learning, we should
-see similarly good performance from this baseline run. In fact,
+to the runs with larger frontier sizes is simply due to an increase in
+the number of expressions seen, rather than increased learning, we
+should see similarly good performance from this baseline run. In fact,
 however, the baseline run only hits $27$ of the $1000$ tasks ($3\%$),
@@
 
-What does the EC algorithm learn in this task?  Inspecting the top
+What does the E.C. algorithm learn in this task?  Inspecting the top
 weighted primitives of the final grammars, we find many incrementers
@@
 stepping stones from an impoverished representation to a rich one. We
-can test this hypothesis by varying the set of tasks to which the EC
+can test this hypothesis by varying the set of tasks to which the E.C.
 algorithm is exposed. If it is true, then we should see a nonlinear
@@
 \begin{figure}[t]
-\includegraphics{./tablealonedoc.pdf}
-\vspace{-.5cm}
+\includegraphics{./figures/tablealonedoc.pdf}
+%% \vspace{-.5cm}
 \caption{Some learned primitives from the circuit based Boolean
@@
   found in the learned grammar.  \label{table:booleanCircuits}}
-\vspace{-.4cm}
+%%\vspace{-.4cm}
 \end{figure}
@@
 As another demonstration of our approach, we investigate the
-EC algorithm's ability to learn Boolean functions using only the logical
+E.C. algorithm's ability to learn Boolean functions using only the logical
 primitive NAND and the basic combinators. It is well known that a
@@
 of tasks as shown in Figure~\ref{fig:booldistr}~(a). In
-Figure~\ref{fig:boolLearningCurvesCircuits}, We present learning
+Figure~\ref{fig:boolLearningCurvesCircuits}, we present learning
 curves for frontier sizes 100, 500 and 10000, using the same procedure
@@
 \begin{figure}
-\vspace{-.3cm}
+%%\vspace{-.3cm}
 \includegraphics[width=\linewidth]{figures/circuitDistrComp.pdf}
-\vspace{-.6cm}
+%%\vspace{-.6cm}
 \caption{Distribution of boolean functions among (a) the set of 1000
@@
   function.\label{fig:booldistr}}
-\vspace{-.4cm}
+%%\vspace{-.4cm}
 \end{figure}
@@
 also include two higher-order expression $E_1$ and $E_2$, to stress
-that the representation the EC algorithm is using allows it to build
+that the representation the E.C. algorithm is using allows it to build
 concepts that are more expressive that just sub-circuits.
@@
 three or fewer inputs as tasks. This is a more difficult
-problem. There are two kinds of learning that the EC algorithm might
+problem. There are two kinds of learning that the E.C. algorithm might
 accomplish: first, many expressions in $\mathcal{L}$ map to a single
@@
 second kind of structure. In
-Figure~\ref{fig:boolLearningCurvesUniform}, we show 10 runs of the EC
-algorithm on the second experiment with a frontier size of $2000$: note how there are several
-local minima that the majority of the runs get stuck in with performance
-around $50\%$, but several of the runs seem to take different
-trajectories to more successful representations.
+Figure~\ref{fig:boolLearningCurvesUniform}, we show 10 runs of the
+E.C.  algorithm on the second experiment with a frontier size of
+$2000$: note how there are several local minima that the majority of
+the runs get stuck in with performance around $50\%$, but several of
+the runs seem to take different trajectories to more successful
+representations.
 
@@
 \end{subfigure}
-\begin{subfigure}[Before]{.9\linewidth}
-\includegraphics[width=.9\linewidth]{figures/booleancircuits3frontiers.pdf}
+\begin{subfigure}[Before]{0.9\linewidth}
+\includegraphics[width=1.0\linewidth]{figures/booleancircuits3frontiers.pdf}
 \end{subfigure}
@@
 \begin{subfigure}[After]{.9\linewidth}
-\includegraphics[width=.9\linewidth]{figures/booleanfunctionsUniform.pdf}
+\includegraphics[width=1.0\linewidth]{figures/booleanfunctionsUniform.pdf}
 
@@
 choices which should be explored further in future work.  Our claim
-here is not that the EC algorithm is optimal, but that it captures key
+here is not that the E.C. algorithm is optimal, but that it captures key
 features of how people solve the problem of learning new systems of

### 2013-04-20 drafts/ijcai2013/ijcai13.tex
Merged changes to ijcai13.tex that were accidentally living in the master branch.

@@
 
-\usepackage{titling}
-
-\setlength{\droptitle}{-1em}   
-\date{}
-
 \newtheorem{definition}{Definition}
@@
 
-%% \author{Eyal Dechter \\
-%% MIT\\
-%% USA \\
-%% edechter@mit.edu
-%% \And
-%% Jon Malmaud \\
-%% MIT \\ 
-%% USA \\
-%% malmaud@mit.edu
-%% \And 
-%% Ryan P.  Adams \\
-%% Harvard University\\
-%% USA \\
-%% rpa@seas.harvard.edu
-%% \And
-%% Joshua B. Tenenbaum \\
-%% MIT\\
-%% USA \\
-%% jbt@mit.edu
-%% }
+\author{Eyal Dechter \\
+MIT\\
+USA \\
+edechter@mit.edu
+\And
+Jon Malmaud \\
+MIT \\ 
+USA \\
+malmaud@mit.edu
+\And 
+Ryan P.  Adams \\
+Harvard University\\
+USA \\
+rpa@seas.harvard.edu
+\And
+Joshua B. Tenenbaum \\
+MIT\\
+USA \\
+jbt@mit.edu
+}
 
@@
  show that the learner discovers modular concepts for these
- domains. Whereas the learner is initially unable to solve almost any
- of the posed problems, it quickly acquires the ability to solve many
- problems by gaining abstract knowledge of the structure of the
- solution space.
+ domains. Whereas the learner is able to solve almost none of the
+ posed problems in the procedure's first iteration, it rapidly becomes
+ able to solve a large number by gaining abstract knowledge of the
+ structure of the solution space.
 
@@
 \section{Related Work}
-The idea of discovering and using reusable subcomponents is an old one
-in AI. Rendell's 1985 paper~\cite{DBLP:conf/ijcai/Rendell85},
-``Substantial Constructive Induction Using Layered Information
-Compression: Tractable Feature Formation in Search,'' presented within
-a classical search paradigm the idea that compression provides a
-heuristic for constructing good representations. Within the Inductive
-Logic Programming (ILP) literature, researchers have explored ``Constructive
-Induction,'' generating new and reusable logic programming
-terms~\cite{DBLP:conf/ijcai/Muggleton87}. We apply these ideas to
+The idea of discovering and using reusable subcomponents is part of an
+old tradition in AI. For instance, it is one of the central themes in
+Herb Simon's 1969 ``The Sciences of the
+Artificial''~\cite{simon1996sciences}. Rendell's 1985
+paper~\cite{DBLP:conf/ijcai/Rendell85}, ``Substantial Constructive
+Induction Using Layered Information Compression: Tractable Feature
+Formation in Search,'' presented within a classical search paradigm
+the idea that compression provides a heuristic for constructing good
+representations. Within the Inductive Logic Programming (ILP)
+literature, researchers have explored ``Constructive Induction,''
+generating new and reusable logic programming
+terms~\cite{DBLP:conf/ijcai/Muggleton87}. Later, predicate invention
+was also applied in the multitask
+setting~\cite{DBLP:conf/ilp/KhanMP98}. We apply these ideas to
 learning functional programs, and believe that the modularity and
@@
 expressions in our language.  We say that expression $ e \in
-\mathcal{L}$ \textbf{solves} $t_k$ if $ t_k(p) = 1$. Our goal is to
+\mathcal{L}$ \textbf{solves} $t_k$ if $ t_k(e) = 1$. Our goal is to
 solve as many of the tasks in $T$ as we can.
@@
 probability for each combinator $c_n$ to be $p(c_n| \tau(c_n)) \propto
-\frac{p}{\sum_{c_j \in C_{\tau(c_n)}} p_j}$.  That is, $p(c_n | \tau(c_n)$ is
+\frac{p_n}{\sum_{c_j \in C_{\tau(c_n)}} p_j}$.  That is, $p(c_n | \tau(c_n)$ is
 proportional to the probability of sampling $c_n$ from the multinomial
@@
 words, the enumeration procedure is best described by the following
-recursion: every program is either a primitive combinator or it is a
-left child program applied to a right child program.
+recursion: every program is either a primitive combinator OR it is a
+left child program applied to a right child program (that is, a left
+child AND a right child).
 
@@
 
-Our best-first policy is to expand the next partial solution with the
-highest value. If a partial solution has multiple leaves that are not
-primitive combinators, we expand the left most one first, since it is
-the choice of left combinator that defines the type of its right
-combinator in an application pair. That is, the parent node applies
-its constraint on valid solutions via the left child, not the right
-child. We want to apply constraints as early as possible, so we expand
-the left child first to prune more partial solutions.
+Our best-first policy is to expand the next partial solution
+with the highest value. If a partial solution has multiple leaves that
+are not primitive combinators, we expand the left most one first,
+since it is the choice of left combinator that defines the type of its
+right combinator in an application pair. That is, the parent node
+applies its constraint on valid solutions via the left child, not the
+right child. We want to apply constraints as early as possible, so
+expanding left child first prunes more partial solutions.
 

### 2013-04-21 drafts/ijcai2013/ijcai13.tex
Edits based on reviewer comments.

@@
 A natural metric for the compressibility of a set of binary trees is
-the number of unique subtrees (counting leaves) that this
-set contains. As an example of this, suppose we have six tasks whose
-solutions are integers $4^2, 5^2, 6^2, 7^2$, and $8^2$, and suppose our
-frontier contains the $10$ expressions:
+the number of unique subtrees (counting leaves) that this set
+contains. More sophisticated metrics exist; we could use the negative
+logarithm of the probability of the solutions under a suitable
+probability distribution~\cite{DBLP:journals/tit/BarronRY98}. Future
+work will explore whether this sophistication is worthwhile and which
+prior on grammars makes sense. For our purposes here, the unique
+subtree counting metric appropriately rewards reuse of useful
+subtrees. As an example of this, suppose we have six tasks whose
+solutions are integers $4^2, 5^2, 6^2, 7^2$, and $8^2$, and suppose
+our frontier contains the $10$ expressions:
 \vspace{-.3cm}
@@
   baseline search over $150000$ expressions only hits $3\%$ of the
-  tasks. \label{fig:symregLearning}}
+  tasks (dashed pink line). \label{fig:symregLearning}}
 \vspace{-.3cm}

### 2013-04-22 drafts/ijcai2013/ijcai13.tex
Edits to abstract and introduction.

@@
  uninformative, providing perhaps a few bits of information but not
- enough to steer it effectively. How can such a learner ever get off
- the ground? A common intuition is that if the solutions to these
- problems share a common structure, and the learner can solve some
- simple problems by brute force, it should be able to extract useful
- components from these solutions and, by composing them, explore the
- solution space more efficiently. Here, we formalize this intuition,
- where the solution space is the space of typed functional programs
- and the gained information is stored as a stochastic grammar over
- programs. We propose an iterative procedure for exploring such
+ enough to steer the learner effectively. How can such a learner ever
+ get off the ground? A common intuition is that if the solutions to
+ these problems share a common structure, and the learner can solve
+ some simple problems by brute force, it should be able to extract
+ useful components from these solutions and, by composing them,
+ explore the solution space more efficiently. Here, we formalize this
+ intuition, where the solution space is the space of typed functional
+ programs and the gained information is stored as a stochastic grammar
+ over programs. We propose an iterative procedure for exploring such
  spaces: in the first step of each iteration, the learner explores a
@@
 contend with the fact that the less an agent knows about solving
-problems in a domain, the worse its attempts are, and the less it
-learns about how to solve subsequent problems. For example, imagine
-you know nothing about baking but are supplied with a stocked kitchen:
-your initial attempts will be uniformly disastrous, and you will learn
-little from them, except that you did not create a baked good. A
-little bit of knowledge about baking would go a long way towards
-making your initial attempts more informative.
+problems in a domain, the worse its attempts are, and, therefore, the
+less informative they are about how to solve subsequent problems. For
+example, imagine you know nothing about baking but are supplied with a
+stocked kitchen: your initial attempts will be uniformly inedible, and
+you will learn little from them. If, on the other hand, you know the
+basics, you are far more likely to generate recipes that are
+somewhat successful and on whose basis you might improve your skills.
 
 Like many of the problems in AI, this one can be cast as a search
-problem. There are many search algorithms, but they typically exploit
-local or topological structure in the search space. For example,
-convex optimization and gradient based search methods take
-advantage of local smoothness to move in good directions. Local search
-algorithms move from one candidate solution to a neighboring
+problem: the goal is to find a baking procedure that maximizes certain
+criter.  There are many search algorithms, but they typically exploit
+local or topological structure in the search space. For example, local
+search algorithms move from one candidate solution to a neighboring
 one. Heuristic search techniques take advantage of topological
@@
 
-The absense of domain specific knowledge makes exploiting structure
-difficult, especially if the structure of the search space is not
-smooth. This is particularly true of problems -- such as discovering
-scientific theories, developing new technologies, learning how to bake
--- whose solutions can be naturally framed as programs. Here we ask
-how to bootstrap domain specific knowledge and extract structure from
-the space of programs by attempting to solve multiple related problems
-in a given domain at once.
-
-Out approach is inspired by the programmer who -- having reused
-several subroutines to solve different but similar simple tasks --
-creates a function library to help manage the complexity of larger
-projects. The E.C. algorithm, which we introduce here, builds a library
-of reusable program components and places a distribution over these,
-effectively learning from simple tasks a search space structure that
-enables it to solve more complex tasks.
+Domain specific knowledge provides this structure. In local search, it
+defines the similarity metric between solutions; in heuristic search
+it provides the heuristic.
+
+Here we ask how to bootstrap domain specific knowledge and extract
+structure from the space of possible solutions by attempting to solve
+multiple related problems in a given domain at once. Our approach is
+motivated by the observation that many interesting problems -- such as
+discovering scientific theories, developing new technologies, learning
+to play a new instrument -- have solutions that are natrually
+specified as programs.
+
+Our proposed solution is thus inspired by the programmer who -- having
+reused subroutines to solve similar simple tasks -- creates a function
+library to help manage the complexity of larger projects. The
+E.C. algorithm, which we introduce here, builds a library of reusable
+program components and places a distribution over these, effectively
+learning from simple tasks a search space structure that enables it to
+solve more complex tasks.
 
@@
 setting~\cite{DBLP:conf/ilp/KhanMP98}. We apply these ideas to
-learning functional programs, and believe that the modularity and
+learning functional programs and believe that the modularity and
 compositionality afforded by this representation is necessary for
@@
 %%\vspace{-.3cm}
-\includegraphics[width=\linewidth]{figures/circuitDistrComp.pdf}
+\includegraphics[width=\linewidth]{figures/circuitDistrComp2.png}
 %%\vspace{-.6cm}

### 2013-04-22 drafts/ijcai2013/ijcai13.tex
More minor edits.

@@
 grammar over combinatory logic to induce shared structure in
-multi-task program induction problems. We find this latter
-representation compelling and use it here, but we see both this and
-genetic programming as tackling the question of how to benefit from
-shared structure in situations where local search might be successful
-on its own. Such problems rely on domain knowledge -- whether in the
-form of a fitness function or a likelihood function -- to provide
-local structure to the search space.
-
-But, as mentioned before, the problems we want AI systems to solve
-often lack this kind of built in structure. Here we attempt to learn
+multi-task program induction problems. We find their representation of
+programs compelling and use it here, but we see both this and genetic
+programming as tackling the question of how to benefit from shared
+structure in situations where local search might be successful on its
+own. Such problems rely on domain knowledge -- whether in the form of
+a fitness function or a likelihood function -- to provide local
+structure to the search space.
+
+But, as mentioned before, our goal is to learn to solve problems in
+solution spaces whose structure is unknown. Here we attempt to learn
 functional programs to solve those kinds of problems.  \\
@@
 the frontier. We will say that a task is \textbf{hit} by the frontier
-if there is a expression in the frontier that solves it.
+if there is an expression in the frontier that solves it.
 
@@
 distribution. However, the fact that we condition on the requesting
-type makes this a considerably more difficult optimization task. One
-of the simplest approximations we can make is to calculate the
-frequency with which each primitive combinator appears in
-$\mathcal{E}$ when the requesting type is such that it could have
-been chosen. This is a straightforward calculation, and
-we use it for its convenience, finding that our empirical results
-justify its use. Future inquiry will be needed to determine to what
-extent a more accurate estimator can improve these results.
+type makes this a considerably more difficult optimization task. A
+simple approximation is to calculate the frequency with which each
+primitive combinator appears in $\mathcal{E}$ when the requesting type
+is such that it could have been chosen. This is a straightforward
+calculation, and we use it for its convenience, finding that our
+empirical results justify its use. Future inquiry will determine to
+what extent a more accurate estimator can improve these results.
 
@@
 
-Our best-first policy is to expand the next partial solution
-with the highest value. If a partial solution has multiple leaves that
-are not primitive combinators, we expand the left most one first,
-since it is the choice of left combinator that defines the type of its
-right combinator in an application pair. That is, the parent node
-applies its constraint on valid solutions via the left child, not the
-right child. We want to apply constraints as early as possible, so
-expanding left child first prunes more partial solutions.
+Our best-first policy is to expand the next partial solution with the
+highest value. If a partial solution has multiple leaves that are not
+primitive combinators, we expand the left most one first, since it is
+the choice of left combinator that defines the type of its right
+combinator in an application pair. That is, the parent node applies
+its constraint on valid solutions via the left child, not the right
+child. We want to apply constraints as early as possible, and thus we
+expand the left child first.
 

### 2013-04-22 drafts/ijcai2013/ijcai13.tex
More edits to paper and added missing citation.

--- 2013-04-22 drafts/ijcai2013/ijcai13.tex
@@
  first step to estimate a new stochastic grammar. We test this
- procedure on symbolic regression and boolean circuit learning and
+ procedure on symbolic regression and Boolean circuit learning and
  show that the learner discovers modular concepts for these
@@
 defines the similarity metric between solutions; in heuristic search
-it provides the heuristic.
-
-Here we ask how to bootstrap domain specific knowledge and extract
-structure from the space of possible solutions by attempting to solve
-multiple related problems in a given domain at once. Our approach is
-motivated by the observation that many interesting problems -- such as
-discovering scientific theories, developing new technologies, learning
-to play a new instrument -- have solutions that are natrually
-specified as programs.
+it provides the heuristic. Here we ask how to bootstrap domain
+specific knowledge and extract structure from the space of possible
+solutions by attempting to solve multiple related problems in a given
+domain at once. Our approach is motivated by the observation that many
+interesting problems -- such as discovering scientific theories,
+developing new technologies, learning to play a new instrument -- have
+solutions that are natrually specified as programs.
 
@@
 system~\cite{Pierce_2002}. In this representation, a type is either a
-type primitive (e.g. reals, integers, booleans, etc.), a type variable
+type primitive (e.g. reals, integers, Booleans, etc.), a type variable
 (e.g. $\sigma$), or a function type $ \tau_1 \rightarrow \tau_2$ of
@@
 work will explore whether this sophistication is worthwhile and which
-prior on grammars makes sense. For our purposes here, the unique
-subtree counting metric appropriately rewards reuse of useful
-subtrees. As an example of this, suppose we have six tasks whose
-solutions are integers $4^2, 5^2, 6^2, 7^2$, and $8^2$, and suppose
-our frontier contains the $10$ expressions:
+prior on grammars makes sense. For our purposes here, penalizing
+according to the number of unique subtrees appropriately rewards reuse
+of useful subtrees. As an example of this, suppose we have six tasks
+whose solutions are integers $4^2, 5^2, 6^2, 7^2$, and $8^2$, and
+suppose our frontier contains the $10$ expressions:
 \vspace{-.3cm}
@@
 polynomial $f$, and an expression $e$ solves $t$ if it returns the
-same value when applied $i \in 0, \dots, 9$ that $f$ does. In the
+same value when applied to $i \in 0, \dots, 9$ that $f$ does. In the
 experiment shown here, the set of problems corresponds to the set of
@@
 library, we include the basic combinators I, S, B, C and four
-arithmetic primitives, $1$, $0$, $*$, and $+$. The choice of initial
+arithmetic primitives $1$, $0$, $*$, and $+$. The choice of initial
 weights on these primitives can make a difference, as it determines
@@
 In Figure~\ref{fig:symregLearning}, we show performance results as we
-vary the frontier size.Changing the frontier size changes the number
+vary the frontier size. Changing the frontier size changes the number
 of tasks the algorithm identifies correctly over the course of 15
@@
 primitive NAND and the basic combinators. It is well known that a
-boolean circuit can be constructed for any boolean function given only
+Boolean circuit can be constructed for any Boolean function given only
 the NAND gate. Here, the basic combinators take the place of the
@@
 tasks, we sampled $1000$ Boolean circuits using AND, OR, and NOT
-gates. To accomplish this, we first sampled either 1, 2, 3 inputs,
+gates. To accomplish this, we first sampled either 1, 2 or 3 inputs,
 then between 1 and 5 gates, randomly wiring the inputs of each new
@@
 tasks consisting of $82$ unique Boolean functions, with a distribution
-of tasks as shown in Figure~\ref{fig:booldistr}~(a). In
+of tasks as shown in Figure~\ref{fig:booldistr}(a). In
 Figure~\ref{fig:boolLearningCurvesCircuits}, we present learning
@@
 %%\vspace{-.6cm}
-\caption{Distribution of boolean functions among (a) the set of 1000
+\caption{Distribution of Boolean functions among (a) the set of 1000
   circuits randomly constructed from AND, OR and NOT gates (see text);

### 2013-04-22 drafts/ijcai2013/ijcai13.tex
Removed unneccessary vspaces.

@@
 \includegraphics[width=1.0\linewidth]{figures/symreg3frontiers.pdf}
-\vspace{-.2cm}
+%\vspace{-.2cm}
 \caption{Learning curves as a function of frontier size.  As frontier
@@
   tasks (dashed pink line). \label{fig:symregLearning}}
-\vspace{-.3cm}
+%\vspace{-.3cm}
 \end{figure}
@@
 \end{subfigure}
-\vspace{-.3cm}
+%\vspace{-.3cm}
 \caption{Learning curves from Boolean function learning
@@
 }
-\vspace{-.5cm}
+%\vspace{-.5cm}
 \end{figure}
@@
 how bootstrap learning of abstract composable concepts can rapidly
-improve search performance.  We show that this approach can learn
-successfully even in domains for which the solution space is not
-smooth and the error signal is all or none, where the only sense of
-locality to guide learning is the modular and compositional structure
-of the solutions that the algorithm itself builds.
+facilitate problem solving in the absence of domain specific
+knowledge. We show that this approach can learn successfully even in
+domains for which the solution space is not smooth and the error
+signal is all or none, where the only sense of locality to guide
+learning is the modular and compositional structure of the solutions
+that the algorithm itself builds.
 

### 2013-04-22 drafts/ijcai2013/ijcai13.tex
Changed bibliography style to named, using the named.bst file.
Changed intro motivating example from baking to circuit building.

@@
 problems in a domain, the worse its attempts are, and, therefore, the
-less informative they are about how to solve subsequent problems. For
-example, imagine you know nothing about baking but are supplied with a
-stocked kitchen: your initial attempts will be uniformly inedible, and
-you will learn little from them. If, on the other hand, you know the
-basics, you are far more likely to generate recipes that are
-somewhat successful and on whose basis you might improve your skills.
+less informative they are about how to solve subsequent problems.
+
+For example, imagine you know nothing about electronics but are
+supplied with a collection of circuit elements and asked to build a
+useful device: your initial attempts will be uniformly
+ineffective. If, on the other hand, you know the basics, you are far
+more likely to generate devices that are somewhat successful and on
+whose basis you might improve your skills.
 
 Like many of the problems in AI, this one can be cast as a search
-problem: the goal is to find a baking procedure that maximizes certain
-criter.  There are many search algorithms, but they typically exploit
-local or topological structure in the search space. For example, local
-search algorithms move from one candidate solution to a neighboring
-one. Heuristic search techniques take advantage of topological
-structure to prune off large parts of the solution space.
+problem: the goal is to find a circuit wiring procedure that maximizes
+certain criteria.  There are many search algorithms, but they typically
+exploit local or topological structure in the search space. For
+example, local search algorithms move from one candidate solution to a
+neighboring one. Heuristic search techniques take advantage of
+topological structure to prune off large parts of the solution space.
 
@@
 %% The file named.bst is a bibliography style file for BibTeX 0.99c
-\bibliographystyle{plain}
+\bibliographystyle{named}
 \bibliography{ijcai13}

### 2013-04-22 drafts/ijcai2013/ijcai13.tex
Got ijcai13.tex to compile.

@@
 \usepackage{amsmath} 
+\usepackage{amsfonts}
 \usepackage{latexsym} 
@@
           {$ \mathbf{S} : (t_2 \rightarrow t_1 \rightarrow t_0)  \rightarrow (t_2 \rightarrow t_1) 
-            \rightarrow \\ t_2 \rightarrow t_0$}}
+            \rightarrow$ \\ $t_2 \rightarrow t_0$}}
         child {node [right=0cm, text width=2cm] {$ \mathbf{*}: \mathbb{R} \rightarrow \mathbb{R} \rightarrow \mathbb{R}$}}
@@
         child {node [above=.5cm] {$ \mathbf{I}: t_0 \rightarrow t_0 \rightarrow t_0$}}
-      }
+      };
     \end{tikzpicture}
@@
 \end{subfigure}
-%\vspace{-.3cm}
 \caption{Learning curves from Boolean function learning
@@
   of the trajectories get stuck in a local minimum around
-  $50\%$.\label{fig:boolLearningCurves}
-
-}
-%\vspace{-.5cm}
+  $50\%$.\label{fig:boolLearningCurves}}
 \end{figure}
@@
 that the algorithm itself builds.
+
+The problem presented here was chosen to highlight the vastness of the
+search space confronting an agent with minimal background knowledge:
+the reward function is binary, it is deterministic, and it cannot
+provide feedback about partial solutions. An important question left
+to be addressed is how to extend this algorithm to take into account
+continuous rewards, noisy data, and partial solutions.
 
@@
 in these more human-like ways is a prime target for future research.
-%% \subsection{Digital arithmetic}
-%% Having demonstrated the behavior of the E-C algorithm on symbolic
-%% regression, we analyze its performance on tasks involving digital
-%% arithmetic, a class of problems with known modular structure. The
-%% general goal is to discover boolean circuits consisting of a limited
-%% set of gates that perform arithmetic operations on binary number
-%% representations.
-
-%% \section{Discussion}
-%% This work attempts to bring together ideas from various areas of
-%% research -- hierarchical Bayes, learning to learn, minimum description
-%% length learning -- to provide a proof of concept that learning
-%% abstractions from programs via compression can rapidly improve search
-%% performance. We emphasize this by focusing on domains in which the
-%% solution space is not smooth and the error signal is all or
-%% none. Thus, the main source of information for learning is the modular
-%% and compositional structure of the solution space.
-
-%% Of course, the representations that people use to learn in the world
-%% often have a partially smooth or local structure. Techniques in AI and
-%% machine learning commonly use smoothness or local structure to make
-%% search tractable. But the difference between this kind of learning and
-%% conceptual learning is undeniable. The former is what the statistician
-%% or machine learning researcher implements. The latter is what she
-%% learns as a student and as a practicioner. As researchers, we try
-%% different models, watch them fail or succeed, and abstract out the
-%% elements of these models that capture the relevant differences. We mix
-%% and match parts of different ideas, and utilize succcessful patterns
-%% of reasoning that can span many problems and domains.
-
-%% We believe that the primary characteristic of conceptual learning is
-%% that it occurs over a modular and compositional computational
-%% representation that supports higher-order functions. That is why we
-%% chose a functional programming language as the representational
-%% language of this work.
-
-%% The experiments in this work do not contain the richness of real world
-%% learning, but we believe that the learning algorithm's behavior
-%% exhibits some of the characteristics we would expect of a conceptual
-%% learner.
-
-%% One characteristic of this algorithm's behavior is a rapid initial
-%% increase in performance followed by a hard plateau. This plateau is
-%% expected given the circular nature of the algorithm: the learner
-%% inspects the region of the solution space that is typical of solutions
-%% already found. We see in this phenomenon a conceptual demarcation of
-%% the problem space: a coherent set of concepts or problems solving
-%% techniques can define by their power the extent of problems a problem
-%% solver is willing to consider. One might hope, however, that an
-%% intelligent learner would recognize that such a demarcation has
-%% occured and attempt to find another class of concepts for the
-%% remaining unsolved tasks.
-
-%% Another aspect of conceptual learning that we see present in this
-%% algorithm's behavior is the importance of a curriculum. By this we do
-%% not mean an explicit ordering of the problems in the task set, but
-%% rather that a set of sufficiently simple problems must act as stepping
-%% stones to the acquisition of complex representations.
 

### 2013-04-22 drafts/ijcai2013/ijcai13.tex
Fixed multiple author citations and added shortcitations where author is obvious.

@@
 Herb Simon's 1969 ``The Sciences of the
-Artificial''~\cite{simon1996sciences}. Rendell's 1985
-paper~\cite{DBLP:conf/ijcai/Rendell85}, ``Substantial Constructive
+Artificial''~\shortcite{simon1996sciences}. Rendell's 1985
+paper~\shortcite{DBLP:conf/ijcai/Rendell85}, ``Substantial Constructive
 Induction Using Layered Information Compression: Tractable Feature
@@
 Genetic programming
-(GP)~\cite{DBLP:conf/foga/Koza92}~\cite{DBLP:conf/ices/KozaBAK96} has
+(GP)~\cite{DBLP:conf/foga/Koza92,DBLP:conf/ices/KozaBAK96} has
 tackled the problem of program learning and has explored the notion
@@
 discovering useful subroutines is seen as a helpful heuristic. More
-recently, Liang et al.~\cite{liang10programs} used a stochastic
+recently, Liang et al.~\shortcite{liang10programs} used a stochastic
 grammar over combinatory logic to induce shared structure in
@@
 these basis combinators. Our representation thus becomes
-variable-free. See Liang et al.~\cite{liang10programs} for a more
+variable-free. See \cite{liang10programs} for a more
 detailed discussion of this routing interpretation.
@@
 enumerating the shortest program satisfying some
-criterion~\cite{DBLP:conf/sfp/Katayama05}
-\cite{DBLP:conf/aaip/YakushevJ09}. Our approach is to formulate the
+criterion~\cite{DBLP:conf/sfp/Katayama05, DBLP:conf/aaip/YakushevJ09}. Our approach is to formulate the
 problem as a best-first exploration of an AND/OR tree
-(Figure~\ref{fig:andor})~\cite{nilsson1982principles}~\cite{DBLP:journals/cacm/Hall73}. In
+(Figure~\ref{fig:andor})~\cite{nilsson1982principles, DBLP:journals/cacm/Hall73}. In
 words, the enumeration procedure is best described by the following

### 2013-04-22 drafts/ijcai2013/ijcai13.tex
Added pdfinfo to preamble.
Minor alterations to citations to get compilation to work.

@@
 
+\usepackage{times}
+\pdfinfo{
+  /Title (Bootstrap learning via modular concept discovery)
+  /Author (Eyal Dechter, Jon Malmaud, Ryan P. Adams, Joshua B. Tenenbaum)
+}
+
 \newtheorem{definition}{Definition}
@@
 \And 
-Ryan P.  Adams \\
+Ryan~P.  Adams \\
 Harvard University\\
@@
 \And
-Joshua B. Tenenbaum \\
+Joshua~B. Tenenbaum \\
 MIT\\
@@
 enumerating the shortest program satisfying some
-criterion~\cite{DBLP:conf/sfp/Katayama05, DBLP:conf/aaip/YakushevJ09}. Our approach is to formulate the
+criterion~\cite{DBLP:conf/sfp/Katayama05,DBLP:conf/aaip/YakushevJ09}. Our approach is to formulate the
 problem as a best-first exploration of an AND/OR tree
-(Figure~\ref{fig:andor})~\cite{nilsson1982principles, DBLP:journals/cacm/Hall73}. In
+(Figure~\ref{fig:andor})~\cite{nilsson1982principles,DBLP:journals/cacm/Hall73}. In
 words, the enumeration procedure is best described by the following
@@
 \includegraphics[width=1.0\linewidth]{figures/symreg3frontiers.pdf}
-%\vspace{-.2cm}
 \caption{Learning curves as a function of frontier size.  As frontier
@@
   tasks (dashed pink line). \label{fig:symregLearning}}
-%\vspace{-.3cm}
 \end{figure}
@@
 \includegraphics{./figures/tablealonedoc.pdf}
-%% \vspace{-.5cm}
 \caption{Some learned primitives from the circuit based Boolean
@@
   found in the learned grammar.  \label{table:booleanCircuits}}
-%%\vspace{-.4cm}
 \end{figure}
@@
 wiring normally found in a circuit.
-
-%% \begin{table}
-%% \tiny
-%% \begin{tabular}[width=\linewidth]{|p{1cm}|p{2cm}|p{3.5cm}|}
-%% \hline function  & CL expression & circuit  ~ \\ \hline\hline
-%% ¬†¬†¬†¬†¬†¬†¬†¬†NOT¬†¬† & S NAND I¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬† &  ~
-
-%% \begin{circuitikz}[transform shape]
-%% \draw
-%% (1.2,0) node[nand port, scale=.3] (nand1) {}
-%% (0, 0) node (x) {x}
-%% (x) -| (nand1.in 1)
-%% (x) -| (nand1.in 2)
-%% ;
-%% \end{circuitikz}
-
-%% \\ \hline
-%% ¬†¬†¬†¬†¬†¬†¬†¬†AND¬†¬† & ((C B NAND) (B (S NAND I)))¬†¬†¬†¬†¬†¬†¬†¬†¬†¬† & ~
-
-%% \begin{circuitikz}
-%% \draw
-%% (1.0,0) node[nand port, scale=.3] (nand1) {}
-%% (2.0,0) node[nand port, scale=.3] (nand2) {}
-%% % (0, 0) node (x) {x}
-%% % (0, 0) node (y) {y}
-%% node[above left=.2cm and .7cm  of nand1] (x) {x} %-- (nand1.in 1)
-%% node[below left=.2cm and .7cm  of nand1] (y) {y} %-- (nand1.in 2)
-%% (x) |- (nand1.in 1)
-%% (y) |- (nand1.in 2)
-%% (nand1.out) |- (nand2.in 1)
-%% (nand1.out) |- (nand2.in 2)
-%% ;
-%% \end{circuitikz}
-%%  \\  \hline
-%% ¬†¬†¬†¬†¬†¬†¬†¬†OR¬†¬†¬† & ((B (C (B (S NAND) NAND))) (S NAND I) & ~
-
-%% \begin{circuitikz}
-%% \draw
-%% (0,0) node (x) {x}
-%% (0,.7) node (y) {y}
-
-%% node[nand port, scale=.3, right=1cm of x] (nand1) {}
-%% node[nand port, scale=.3, right=1cm of y] (nand2) {}
-%% node[nand port, scale=.3, below right=.3cm and 1cm of nand2] (nand3) {}
-%% (x) -| (nand1.in 1)
-%% (x) -| (nand1.in 2)
-%% (y) -| (nand2.in 1)
-%% (y) -| (nand2.in 2)
-%% (nand1.out) |- (nand3.in 1)
-%% (nand2.out) |- (nand3.in 2)
-%% ;
-%% \end{circuitikz}
-%%  \\  \hline\hline
-
-%% ¬†¬†¬†¬†¬†¬†¬†¬†TRUE¬† & (S NAND) (S NAND I)¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬† & ~
-
-%% %\begin{top}
-%% \begin{circuitikz}[transform shape] 
-%% \draw
-%% (1, 0) node[nand port, scale=.3] (nand1) {}
-%% (2,.3) node[nand port, scale=.3] (nand2) {}
-%% (0, 0) node[anchor=east]  (x) {x}
-%% (x) -- (.3, 0) -| (nand1.in 1)
-%% (x) -- (.3, 0) -|  (nand1.in 2)
-%% (x) |- (nand2.in 1)
-%% (nand1.out) |- (nand2.in 2)
-%% ;
-%% \end{circuitikz}
-%% %\end{top}
-%% \\ \hline
-%% ¬†¬†¬†¬†¬†¬†¬†¬†FALSE & (S B (S NAND)) (S NAND I)¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬† &  ~
-
-%% \begin{circuitikz}[transform shape] 
-%% \draw
-%% (1, 0) node[nand port, scale=.3] (nand1) {}
-%% (2,.3) node[nand port, scale=.3] (nand2) {}
-%% node[nand port, scale=.3, right=1cm of nand2] (nand3) {} 
-%% (0, 0) node[anchor=east]  (x) {x}
-%% (x) -- (.3, 0) -| (nand1.in 1)
-%% (x) -- (.3, 0) -|  (nand1.in 2)
-%% (x) |- (nand2.in 1)
-%% (nand1.out) |- (nand2.in 2)
-%% (nand2.out) -| (nand3.in 1)
-%% (nand2.out) -| (nand3.in 2)
-%% ;
-%% \end{circuitikz}
-%% \\ \hline
-%% ¬†¬†¬†¬†¬†¬†¬†¬†XOR¬†¬† & ((S (B C (((B (C B NAND)) S) (((B (C B NAND)) S) (B C ((B (B NAND)) NAND)))))) I)¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬† & ~
-
-%% \begin{circuitikz}
-%% \draw
-%% node[nand port, scale=.3] (nand1) {}
-%% node[nand port, scale=.3, above right=.3cm and 1cm of nand1] (nand2) {} 
-%% node[nand port, scale=.3, below right=.3cm and 1cm of nand1] (nand3) {} 
-%% node[nand port, scale=.3, right=2cm of nand1] (nand4) {} 
-%% node[left=1.2cm of nand2.in 1] (x) {x}
-%% node[left=1.2cm of nand3.in 2] (y) {y}
-%% node[right=.2cm of nand4] (out) {}
-%% (x) -| (nand1.in 1)
-%% (x) -| (nand2.in 1)
-%% (y) -| (nand1.in 2)
-%% (y) -| (nand3.in 2)
-%% (nand1.out) -| (nand2.in 2)
-%% (nand1.out) -| (nand3.in 1)
-%% (nand2.out) |- (nand4.in 1)
-%% (nand3.out) |- (nand4.in 2)
-%% (nand4.out) -- (out)
-%% ;
-%% \end{circuitikz}
-
-%%  \\  \hline\hline
-%% ¬†¬†¬†¬†¬†¬†¬†¬†$E_1$¬†¬†¬†  & S B (S NAND)¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬† & ~
-
-%% \begin{circuitikz}
-%% \draw
-%% node (x) {x}
-%% node[nand port, scale=.3, right=1.5cm of x] (nand) {} 
-%% node[draw=blue!75, fill=blue!20, thick, 
-%%   above right=.3cm and .2cm of x, anchor=west] (f1) {$g(x)$} 
-%% node[draw=blue!75, fill=blue!20, thick, right=.5cm of nand] (f2) {$g(x)$}
-%% node[right=.1cm of f2] (out) {}
-%% (x) -| (f1.west)
-%% (x) -| (nand.in 2)
-%% (f1) -| (nand.in 1)
-%% (nand.out) -- (f2)
-%% (f2) -- (out);
-%% \end{circuitikz}
-
-%% \\ \hline
-%% $E_2$¬†¬†¬†  & B (C B NAND) S & ~
-
-%% \begin{circuitikz}
-%% \draw
-%% node (x) {x} 
-%% node[nand port, scale=.3, right=1.5cm of x] (nand) {} 
-%% node[draw=blue!75, fill=blue!20, thick, 
-%%   above right=.2cm and .3cm of x, anchor=west] (g) {$g(x)$} 
-%% node[draw=blue!75, fill=blue!20, thick, 
-%%   below right=.2cm and .3cm of x, anchor=west] (f) {$f(x)$} 
-%% (x) -- [right=.25] (f.west)
-%% (x) |- (g.west)
-%% (f.east) -| (nand.in 1)
-%% (g.east) -| (nand.in 2)
-%% (nand.out) -- (out)
-%% (x)
-%% ;
-%% \end{circuitikz}
-
-%% \\ \hline
-
-%% \end{tabular}
-%% \caption{Expressions in the induced library on sets of tasks
-%%   corresponding to boolean circuits synthesized from AND, OR, and NOT
-%%   gates, with boolean circuits of 1, 2, or 3 inputs.}
-%% \label{table:boolexpr}
-
-%% \end{table}
-
-%% \begin{table}[t]
-%% \scriptsize
-%%    \begin{tabular}{|l|l|l|}
-%%     function  & CL expressions & circuit  ~ \\ \hline
-%%         TRUE  & (S NAND) (S NAND I)                   & ~ \\ \hline
-%%         FALSE & (S B (S NAND)) (S NAND I)             & ~ \\ \hline
-%%         NOT   & S NAND I                              & \input{figures/NOT.tex} \\ \hline
-%%         AND   & ((C B NAND) (B (S NAND I)))           & ~ \\ 
-%%         OR    & ((B (C (B (S NAND) NAND))) (S NAND I) & ~ \\ 
-%%         XOR   & ~                                     & ~ \\ 
-%%         F     & S B (S NAND)                          & \input{figures/F.tex}
-%%     \end{tabular}
-%% \end{table}
 
@@
 \begin{figure}
-%%\vspace{-.3cm}
 \includegraphics[width=\linewidth]{figures/circuitDistrComp2.png}
-%%\vspace{-.6cm}
 \caption{Distribution of Boolean functions among (a) the set of 1000
@@
   function.\label{fig:booldistr}}
-%%\vspace{-.4cm}
 \end{figure}

### 2013-04-22 drafts/ijcai2013/camera_ready/Dechter.IJCAI13.CameraReady.tex
Removed unused package declarations.
Added info about lexicographic ordering of Boolean functions in caption of the Boolean circuits distribution figure.

--- 2013-04-22 drafts/ijcai2013/ijcai13.tex
@@
 \usepackage{ijcai13}
-\usepackage{times}
 \usepackage{amsmath} 
 \usepackage{amsfonts}
-\usepackage{latexsym} 
-\usepackage{ dsfont }
-\usepackage[usenames,dvipsnames]{xcolor}
-\usepackage{tikz}
+\usepackage{tikz} % don't have to include
 \usepackage{tikz-qtree}
-\usepackage{float}
-\usepackage{circuitikz}
-\usepackage{graphicx}
 \usepackage[font=small]{subcaption}
 \usepackage[font=small]{caption}
-\usepackage{alltt}
 \usepackage{dashrule} 
-\usepackage{array}
-\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
 \usetikzlibrary{arrows,automata, positioning, patterns,backgrounds}
@@
 \begin{abstract}
-%\small
  Suppose a learner is faced with a domain of problems about which it
@@
  structure of the solution space.
-
 \end{abstract}
@@
 \begin{figure}
+\vspace{.5cm}
 \includegraphics[width=\linewidth]{figures/circuitDistrComp2.png}
@@
   percent of instances (circuits/expressions) which evaluate to that
-  function.\label{fig:booldistr}}
+  function. Functions increase in domain size from left to right; each
+  function can be represented as a binary string corresponding to the
+  output column of the truth table, and functions are ordered with
+  respect to these strings. \label{fig:booldistr}}
 \end{figure}
@@
 \bibliographystyle{named}
-\bibliography{ijcai13}
+\bibliography{Dechter.IJCAI13.CameraReady}
 %% \nocite{*}

### 2013-04-22 drafts/ijcai2013/camera_ready/Dechter.IJCAI13.CameraReady.tex
Spellcheck and minor formatting changes.

@@
  explore the solution space more efficiently. Here, we formalize this
- intuition, where the solution space is the space of typed functional
+ intuition, where the solution space is that of typed functional
  programs and the gained information is stored as a stochastic grammar
@@
 developing new technologies, learning to play a new instrument -- have
-solutions that are natrually specified as programs.
+solutions that are naturally specified as programs.
 
@@
 old tradition in AI. For instance, it is one of the central themes in
-Herb Simon's 1969 ``The Sciences of the
-Artificial''~\shortcite{simon1996sciences}. Rendell's 1985
-paper~\shortcite{DBLP:conf/ijcai/Rendell85}, ``Substantial Constructive
-Induction Using Layered Information Compression: Tractable Feature
-Formation in Search,'' presented within a classical search paradigm
-the idea that compression provides a heuristic for constructing good
-representations. Within the Inductive Logic Programming (ILP)
-literature, researchers have explored ``Constructive Induction,''
-generating new and reusable logic programming
-terms~\cite{DBLP:conf/ijcai/Muggleton87}. Later, predicate invention
-was also applied in the multitask
+Herb Simon's ``The Sciences of the
+Artificial''~\shortcite{simon1996sciences}. Rendell's ``Substantial
+Constructive Induction Using Layered Information Compression:
+Tractable Feature Formation in Search'' presented within a classical
+search paradigm the idea that compression provides a heuristic for
+constructing good
+representations~\shortcite{DBLP:conf/ijcai/Rendell85}. Within the
+Inductive Logic Programming (ILP) literature, researchers have
+explored ``Constructive Induction,'' generating new and reusable logic
+programming terms~\cite{DBLP:conf/ijcai/Muggleton87}. Later, predicate
+invention was also applied in the multitask
 setting~\cite{DBLP:conf/ilp/KhanMP98}. We apply these ideas to
@@
 %% %% But suppose that you have very little knowledge of a domain. Your
-%% %% intial representation of the domain is poor: ``nearby'' points in the
-%% %% solutions space ellicit wildly different outcomes in unpredictable
-%% %% ways. Or perhaps nearby points ellicit the same outcome and its not a
+%% %% initial representation of the domain is poor: ``nearby'' points in the
+%% %% solutions space elicit wildly different outcomes in unpredictable
+%% %% ways. Or perhaps nearby points elicit the same outcome and its not a
 %% %% good one, so you don't know which way to go. Many learning problems,
@@
 induction problem: suppose we are given a set of \textbf{tasks}
-$T=\{t_k\}_{k=1}^K$ where each task is a function $t_k : \mathcal{L}
-\rightarrow \{0, 1\}$ and where $\mathcal{L}$ is the set of
+$T=\{t_k\}_{k=1}^K$ where each task is a function \mbox{$t_k : \mathcal{L}
+\rightarrow \{0, 1\}$} and where $\mathcal{L}$ is the set of
 expressions in our language.  We say that expression $ e \in
@@
 expressions that the algorithm considers in any one iteration. The
-\textbf{frontier size} $N$ is the prespecified number of elements in
+\textbf{frontier size} $N$ is the pre-specified number of elements in
 the frontier. We will say that a task is \textbf{hit} by the frontier
@@
 \emph{solution set}, want to re-estimate our stochastic grammar. Our
-inspiration for doing this is the Neville-Manning algorithm for
+inspiration for doing this is the Nevill-Manning algorithm for
 grammar-based compression~\cite{nevill1997identifying}. The idea of
@@
 \caption{How do different task sets affect learning curves? Learning
-  curves at frontier size of $10000$ for different task
+  curves at a frontier size of $10000$ for different task
   sets.} 

### 2013-04-22 drafts/ijcai2013/camera_ready/Dechter.IJCAI13.CameraReady.tex
Changes to first paragraph of introduction.

@@
 building in vast amounts of pre-specified knowledge that is tailored
-to one domain or another. Any attempt to solve this problem must
-contend with the fact that the less an agent knows about solving
-problems in a domain, the worse its attempts are, and, therefore, the
-less informative they are about how to solve subsequent problems.
+to one domain or another. This is generally accomplished by a learning
+algorithm that uses past experience to improve future performance. The
+problem is that if past problem solving attempts entirely missed the
+mark, their failures are unlikely to shed light on will work and what
+will not. Such uninformative solutions, however, vastly outnumber the
+rest, and our uninformed agent has no way to know which solutions will
+be informative and which will not. Thus, our learning algorithms must
+contend with a bootstrapping problem: the less an agent knows to begin
+with, the harder it is to learn from experience.
 
@@
 supplied with a collection of circuit elements and asked to build a
-useful device: your initial attempts will be uniformly
-ineffective. If, on the other hand, you know the basics, you are far
+number of canonical useful devices, such as a digital adder, an
+op-amp,, etc. Your initial attempts will be uniformly ineffective. If,
+on the other hand, you know the basics of circuit design, you are far
 more likely to generate devices that are somewhat successful and on
-whose basis you might improve your skills.
+whose basis you might improve your designs and your design skills.
 
@@
 
-Domain specific knowledge provides this structure. In local search, it
-defines the similarity metric between solutions; in heuristic search
-it provides the heuristic. Here we ask how to bootstrap domain
-specific knowledge and extract structure from the space of possible
-solutions by attempting to solve multiple related problems in a given
-domain at once. Our approach is motivated by the observation that many
-interesting problems -- such as discovering scientific theories,
-developing new technologies, learning to play a new instrument -- have
-solutions that are naturally specified as programs.
+This structure derives from domain specific knowledge. In local
+search, domain knowledge provides the similarity metric between
+solutions; in heuristic search it provides the heuristic. Here we ask
+how to bootstrap domain specific knowledge and extract structure from
+the space of possible solutions by attempting to solve multiple
+related problems in a given domain at once. Our approach is motivated
+by the observation that many interesting problems -- such as
+discovering scientific theories, designing new technologies, learning
+a skill such as playing an instrument or cooking -- have solutions that
+are naturally specified as programs.
 

### 2013-04-22 drafts/ijcai2013/camera_ready/Dechter.IJCAI13.CameraReady.tex
Inserted the rest of Josh's edits.

@@
 discovering scientific theories, designing new technologies, learning
-a skill such as playing an instrument or cooking -- have solutions that
-are naturally specified as programs.
+a new language or learning a skill such as playing an instrument or
+cooking -- have solutions that are naturally specified as programs.
 
@@
 the number of unique subtrees (counting leaves) that this set
-contains. More sophisticated metrics exist; we could use the negative
-logarithm of the probability of the solutions under a suitable
-probability distribution~\cite{DBLP:journals/tit/BarronRY98}. Future
-work will explore whether this sophistication is worthwhile and which
-prior on grammars makes sense. For our purposes here, penalizing
-according to the number of unique subtrees appropriately rewards reuse
-of useful subtrees. As an example of this, suppose we have six tasks
+contains. As an example of this, suppose we have six tasks
 whose solutions are integers $4^2, 5^2, 6^2, 7^2$, and $8^2$, and
@@
 abstraction when multiple expressions are being considered at once.
+
+More sophisticated metrics than the one proposed here exist; we could
+use the negative logarithm of the probability of the solutions under a
+suitable probability
+distribution~\cite{DBLP:journals/tit/BarronRY98}. Future work will
+explore whether this sophistication is worthwhile and which prior on
+grammars makes sense. For our purposes here, penalizing according to
+the number of unique subtrees appropriately and sufficiently rewards
+reuse of useful subtrees.
 
@@
 In Figure~\ref{fig:symregCurricula}, we present learning curves
-(frontier size 10000) corresponding to various versions of our
+(frontier size~10000) corresponding to various versions of our
 original symbolic regression task set. Recall that the original set
@@
 the reward function is binary, it is deterministic, and it cannot
-provide feedback about partial solutions. An important question left
-to be addressed is how to extend this algorithm to take into account
-continuous rewards, noisy data, and partial solutions.
+provide feedback about partial solutions. In ongoing work we are
+exploring a more probabilistic version of the algorithm that can take
+into account continuous rewards, noisy data, and partial solutions.
 

### 2013-04-23 drafts/ijcai2013/camera_ready/Dechter.IJCAI13.CameraReady.tex
Added Josh's version of the introduction.

@@
 \section{Introduction}
-A central question for AI is how to create intelligent agents without
-building in vast amounts of pre-specified knowledge that is tailored
-to one domain or another. This is generally accomplished by a learning
-algorithm that uses past experience to improve future performance. The
-problem is that if past problem solving attempts entirely missed the
-mark, their failures are unlikely to shed light on will work and what
-will not. Such uninformative solutions, however, vastly outnumber the
-rest, and our uninformed agent has no way to know which solutions will
-be informative and which will not. Thus, our learning algorithms must
-contend with a bootstrapping problem: the less an agent knows to begin
-with, the harder it is to learn from experience.
-
-For example, imagine you know nothing about electronics but are
-supplied with a collection of circuit elements and asked to build a
-number of canonical useful devices, such as a digital adder, an
-op-amp,, etc. Your initial attempts will be uniformly ineffective. If,
-on the other hand, you know the basics of circuit design, you are far
-more likely to generate devices that are somewhat successful and on
-whose basis you might improve your designs and your design skills.
-
-Like many of the problems in AI, this one can be cast as a search
-problem: the goal is to find a circuit wiring procedure that maximizes
-certain criteria.  There are many search algorithms, but they typically
-exploit local or topological structure in the search space. For
-example, local search algorithms move from one candidate solution to a
-neighboring one. Heuristic search techniques take advantage of
+
+Imagine that you know nothing about electronics but are supplied with
+a collection of circuit elements and asked to build a number of useful
+digital devices, such as an adder, a flip-flop, a counter, a
+multiplexer, and so on.  Your initial attempts will probably be
+uniformly ineffective.  If your only feedback is based on success or
+failure, you may have trouble learning anything.  If, on the other
+hand, you already know the basics of circuit design, you are far more
+likely to generate some devices that are at least somewhat successful.
+In this case, with feedback you can improve your designs and your
+design skills.
+
+This example illustrates the central ``bootstrapping'' challenge for
+attempts to create intelligent agents without building in vast amounts
+of pre-speciﬁed domain knowledge.  Learning from experience is
+typically most effective in a system that already knows most of what
+it needs to know, because mistakes are most informative.  In a system
+that knows almost nothing, where most problem solving attempts
+entirely miss their mark, failures are unlikely to shed light on will
+work and what will not.  Uninformative solutions vastly outnumber the
+rest, and an uninformed agent has no way to know which solutions will
+be informative and which will not.  If it is harder to learn the less
+an agent knows, how can a learner who knows almost nothing about a domain
+ever get off the ground?
+
+Here we propose an answer to this challenge, based on an approach to
+discovering new and increasingly sophisticated concepts that build
+modularly on previously learned, simpler concepts.  Like many problems
+in AI, the bootstrapping challenge and its solution can be cast in
+terms of search.  In our circuit-building scenario, the goal is to ﬁnd
+circuit wiring procedures that maximize certain criteria.  Search
+algorithm typically aim to exploit local or topological structure in
+the search space -- here the space of all circuit-wiring procedures.
+For example, local search algorithms move from one candidate solution
+to a neighboring one. Heuristic search techniques take advantage of
 topological structure to prune off large parts of the solution space.
-
-This structure derives from domain specific knowledge. In local
+This structure derives from domain-speciﬁc knowledge.  In local
 search, domain knowledge provides the similarity metric between
 solutions; in heuristic search it provides the heuristic.  Here we ask
-how to bootstrap domain specific knowledge and extract structure from
-the space of possible solutions by attempting to solve multiple
-related problems in a given domain at once. Our approach is motivated
-by the observation that many interesting problems -- such as
-discovering scientific theories, designing new technologies, learning
-a new language or learning a skill such as playing an instrument or
-cooking -- have solutions that are naturally specified as programs.
-
-Our proposed solution is thus inspired by the programmer who -- having
-reused subroutines to solve similar simple tasks -- creates a function
-library to help manage the complexity of larger projects. The
-E.C. algorithm, which we introduce here, builds a library of reusable
-program components and places a distribution over these, effectively
-learning from simple tasks a search space structure that enables it to
-solve more complex tasks.
+how to bootstrap domain-speciﬁc knowledge by extracting search
+structure from the space of possible solutions, by attempting to solve
+multiple related problems in a given domain at once.
+
+Our approach is motivated by the observation that many interesting
+problems –- such as discovering scientiﬁc theories, designing new
+technologies, learning a skill such as playing an instrument or
+cooking –- have solutions that are naturally speciﬁed as programs.
+Our proposed solution is inspired by the programmer who, having found
+various subroutines that can be reused to solve similar simple tasks,
+creates a function library to help manage the complexity of larger
+projects. The E.C. algorithm, which we introduce here, builds a
+library of reusable program components and places a distribution over
+these, effectively learning from simple tasks a search space structure
+that enables it to solve more complex tasks.
 

### 2013-04-23 drafts/ijcai2013/camera_ready/Dechter.IJCAI13.CameraReady.tex
Reformatted titles to mixed case.

@@
 
-\title{Bootstrap learning via modular concept discovery}
+\title{Bootstrap Learning via Modular Concept Discovery}
 
@@
 
-\section{Extracting concepts from programs}
+\section{Extracting Concepts from Programs}
 
@@
 
-\subsection{Representing typed functional programs as binary trees}
+\subsection{Representing Typed Functional Programs as Binary Trees}
 
@@
 
-\subsection{A stochastic grammar over programs}
+\subsection{A Stochastic Grammar over Programs}
 \label{sec:stochgrammar}
@@
 
-\subsection{Best-first enumeration of programs}
+\subsection{Best-first Enumeration of Programs}
 In order to explore the frontier of most promising programs, we need a
@@
 
-\subsection{Finding the most compressive set of solutions}
+\subsection{Finding the Most Compressive Set of Solutions}
 
@@
 
-\subsection{Re-estimating the stochastic grammar}
+\subsection{Re-estimating the Stochastic Grammar}
 
@@
 
-\subsection{Boolean function learning}
+\subsection{Boolean Function Learning}
 \label{sec:boolcircuits}

### 2013-04-23 drafts/ijcai2013/camera_ready/Dechter.IJCAI13.CameraReady.tex
Minor edits, and fixed one bibtex record.

@@
 of pre-speciﬁed domain knowledge.  Learning from experience is
-typically most effective in a system that already knows most of what
+typically most effective in a system that already knows much of what
 it needs to know, because mistakes are most informative.  In a system
 that knows almost nothing, where most problem solving attempts
-entirely miss their mark, failures are unlikely to shed light on will
-work and what will not.  Uninformative solutions vastly outnumber the
-rest, and an uninformed agent has no way to know which solutions will
-be informative and which will not.  If it is harder to learn the less
-an agent knows, how can a learner who knows almost nothing about a domain
-ever get off the ground?
+entirely miss their mark, failures are unlikely to shed light on what
+will work and what will not.  Uninformative solutions vastly outnumber
+the rest, and an uninformed agent has no way to know which solutions
+will be informative and which will not.  If it is harder to learn the
+less an agent knows, how can a learner who knows almost nothing about
+a domain ever get off the ground?
 
@@
 to a neighboring one. Heuristic search techniques take advantage of
-topological structure to prune off large parts of the solution space.
+topological structure to prune off large parts of the search space.
 This structure derives from domain-speciﬁc knowledge.  In local
@@
   function learning experiment. We highlight compression using $NOT$,
-  $E_1$ and $E_2$. \,(a)The three elementary gates used to generate
+  $E_1$ and $E_2$. \,(a) The three elementary gates used to generate
   the task set. (b) Two higher-order primitives in the induced
@@
 tasks consisting of $82$ unique Boolean functions, with a distribution
-of tasks as shown in Figure~\ref{fig:booldistr}(a). In
+of tasks as shown in Figure~\ref{fig:booldistr}a. In
 Figure~\ref{fig:boolLearningCurvesCircuits}, we present learning

### 2013-04-23 drafts/ijcai2013/camera_ready/Dechter.IJCAI13.CameraReady.tex
Added required commands to preamble per IJCAI13 submission instructions.
Added acknowledgement footnote.
Minor edits.

@@
-\documentclass{article}
+\documentclass[letter]{article}
+
+\setlength{\pdfpagewidth}{8.5in}
+\setlength{\pdfpageheight}{11in}
 
@@
 
-\title{Bootstrap Learning via Modular Concept Discovery}
+\title{Bootstrap Learning via Modular Concept Discovery\thanks{We acknowledge the support of ARO MURI W911NF-08-1-0242.}}
 
@@
 likely to generate some devices that are at least somewhat successful.
-In this case, with feedback you can improve your designs and your
-design skills.
+In this case, you get feedback that helps you improve your designs and
+your design skills.
 
@@
 typically most effective in a system that already knows much of what
-it needs to know, because mistakes are most informative.  In a system
-that knows almost nothing, where most problem solving attempts
-entirely miss their mark, failures are unlikely to shed light on what
-will work and what will not.  Uninformative solutions vastly outnumber
-the rest, and an uninformed agent has no way to know which solutions
-will be informative and which will not.  If it is harder to learn the
-less an agent knows, how can a learner who knows almost nothing about
-a domain ever get off the ground?
+it needs to know, because that is when mistakes are most informative.
+In a system that knows almost nothing, where most problem solving
+attempts entirely miss their mark, failures are unlikely to shed light
+on what will work and what will not.  Uninformative solutions vastly
+outnumber the rest, and an uninformed agent has no way to know which
+solutions will be informative and which will not.  If it is harder to
+learn the less an agent knows, how can a learner who knows almost
+nothing about a domain ever get off the ground?
 
@@
 in AI, the bootstrapping challenge and its solution can be cast in
-terms of search.  In our circuit-building scenario, the goal is to ﬁnd
-circuit wiring procedures that maximize certain criteria.  Search
-algorithm typically aim to exploit local or topological structure in
+terms of search.  In our circuit-building scenario, the goal is to
+find circuit wiring procedures that maximize certain criteria.  Search
+algorithms typically aim to exploit local or topological structure in
 the search space -- here the space of all circuit-wiring procedures.
@@
 how to bootstrap domain-speciﬁc knowledge by extracting search
-structure from the space of possible solutions, by attempting to solve
+structure from the space of possible solutions by attempting to solve
 multiple related problems in a given domain at once.
@@
 
-But, as mentioned before, the problems we want AI systems to solve
-often lack this kind of built in structure. Here we attempt to learn
-functional programs to solve those kinds of problems.  \\
-
-%% AI and machine learning algorithms commonly take advantage of the
-%% relationship between a performance metric (error, reward, fitness,
-%% etc.) and the local or topological structure of the solution
-%% space. Whether in discrete or continuous domains, the key is to design
-%% such a relationship and write a algorithm that either moves in good
-%% directions or prunes off large parts of the solution space. Whether or
-%% not such a relationship exists depends on the representation, and much
-%% work has gone into finding good representations for problem
-%% solving. Recent work, especially in machine learning, has focused on
-%% ``feature learning,'' with the explicit aim of discovered good
-%% representational building blocks in vector-valued solutions spaces.
-
-%% Many of the most natural solution representations for intelligent
-%% agents, however, are not naturally expressed as vector-valued
-%% spaces. Consider the problems of discovering scientific theories,
-%% inventing new technologies, communicating with others, or executing a
-%% complex motor routine (such as dancing or playing an instrument). The
-%% solutions to these problems are most naturally expressed as programs,
-%% that is, as the composition of modular components whose semantics are
-%% determined by the actions of one component on another.
-
-%% How can we learn good solution representations for these kinds of
-%% tasks without building in the kind of domain knowledge that gives rise
-%% to a useful relationship between representation and performance? How
-%% can we get traction on a set of problems when small tweaks to
-%% solutions give rise to wildly varying performances in an unpredictable
-%% manner, or when, for most problems, most of the solutions fail in a
-%% uniformly bad way?
-
-%% A popular idea is that good representations are those that encode the
-%% data space in an efficient way, that is, that enable compression of
-%% the data~\cite{}. For program representations, this most naturally
-%% corresponds to defining reusable components so that the total number
-%% of unique expressions for a set of solutions is minimized (i.e., this
-%% is directly akin to the subroutine libraries programmers use to manage
-%% the complexity of their code).
-
-%% In this work, we try to see how far reusable subroutine definition can
-%% get us on its own. Our question is, Can discovering modular components
-%% transform a problem from one in which brute-force search will not work
-%% to a problem in which it does?
-
-%% To emphasize an impoverished relationship between reward signals and
-%% program representations, we make our tasks all-or-none: a solution
-%% either solves the problem or it does not, and partial solutions do not
-%% get any credit.
-
-%% The intuitive approach to applying this idea is simple: if solutions
-%% to problems in the domain share common structure, and there exist
-%% problems whose solutions are simple enough to find by brute force,
-%% then we should be able to extract a distribution over solution
-%% subcomponents from these simple cases that allows us to run subsequent
-%% searches more efficiently.
-
-%% More concretely, our algorithm works as follows: each iteration, it
-%% first generates programs in order of highest probability according to
-%% a subroutine distribution learned in the previous iteration. From the
-%% discovered solutions, it finds new reusable subroutines that compress
-%% these solutions and estimates their probabilities.
-
-%% We show that this iterative approach allows us to achieve very high
-%% performance in tasks where just searching according to the initial
-%% distribution over program primitives clearly fails.
-
-%%  \hline
-
-%% AI and machine learning algorithms commonly take advantage of the idea
-%% that local structure in the solution space is related to local
-%% structure in the outcome. In continuous domains, learning algorithms
-%% algorithms use error gradients to guide search. In discrete domains,
-%% whether in heuristic or local search paradigms, knowledge about the
-%% solution space topology and its relation to reward and cost metrics
-%% enables pruning, reasoning, and moving in the solution space more
-%% efficiently. However, choosing the representations that have
-%% beneficial structural properties requires a lot of general and domain
-%% knowledge, and this what researchers do and what they cannot leave up
-%% to their algorithms.
-
-%% What is a good strategy for an agent when its representation of the
-%% solution space is not smooth with respect to the outcome? Many hard
-%% learning problems have this property, especially those in which the
-%% solution space is most naturally represented as a set of
-%% programs. Consider the problems of discovering scientific theories,
-%% inventing new technologies, learning to cook, or learning to
-%% communicate. These are problems in which 1) solutions are best
-%% described as programs (i.e. compositions of modular subroutines at
-%% various levels of abstraction), 2) most of the solution space is bad
-%% in a relatively uninformative way (i.e. it just doesn't work), and
-%% thus 3) outcomes are not informative about how you are doing unless
-%% you are already quite close to a solution.
-
-%% Intuitively, the way to solve such problems is to find solutions to
-%% simple cases in the domain and extract conceptual building blocks that
-%% can be composed to scale up to complex cases. With experience, the
-%% agent may be able to discover local structure in the solution space,
-%% and utilize more traditional AI and machine learning techniques. In
-%% this work, we will be focusing on the first phase of this learning
-%% problem.
-
-%% %% Of course, every learner
-
-%% %% Choosing the wrong representation, in
-%% %% which local structure is not apparent, is often the difference between
-%% %% a successful learner and an unsuccessful one. In many respects, this
-%% %% may be the kind of situation that adult humans face when confronting
-%% %% new domains: we have a lot of general knowledge of successful
-%% %% representations, useful strategies, etc.
-
-%% %% But suppose that you have very little knowledge of a domain. Your
-%% %% initial representation of the domain is poor: ``nearby'' points in the
-%% %% solutions space elicit wildly different outcomes in unpredictable
-%% %% ways. Or perhaps nearby points elicit the same outcome and its not a
-%% %% good one, so you don't know which way to go. Many learning problems,
-%% %% especially those whose solutions are best described as programs, have
-%% %% these qualities: imagine learning to synthesize new sentences
-
-%% %% teness. We choose the representations our learners use
-%% %% and these representations generally Imagine you are trying to learn to
-%% %% bake via trial and error. You know what kind of ingredients are used
-%% %% in baking (flour, water, butter, sugar, yeast, baking soda, etc.) and
-%% %% you know a bit about how to combine these ingredient (mix, whisk,
-%% %% etc.) and how to apply heat (oven). But baking is a notoriously
-%% %% fragile process: ratios of ingredients, the order of mixing, cooking
-%% %% temperatures, resting times, all make a huge difference, and the
-%% %% outcome of the baking process is mostly all or none. Either the
-%% %% outcome is what you expected or quite unlike it. Even if you do create
-%% %% something similar to your intent, it is difficult to assign blame for
-%% %% the outcome to any one step or decision along the way.
-
-%% %% Many difficult learning problems have this structure -- learning
-%% %% language, developing technologies and scientific theories, learning to
-%% %% play instruments or execute other complex motor patterns. These are
-%% %% problems in which 1) there is a large set of similar tasks; 2) their
-%% %% solutions are best described as programs; 3) the solution space is
-%% %% not smooth with respect to the outcome; 4) and the outcome is generally
-%% %% uninformative unless you are already pretty close to a solution.
-
-%% %% Of course, knowledge is not acquired wholly by trial and error. But in
-%% %% many domains, there is at least initially such a period of ignorance,
-%% %% and the question we are trying to tackle here is how you ever get off
-%% %% the ground.
-
-%% I need help here.
-
-%% %% Let a \emph{task} be a function $t : \mathcal{L} \rightarrow \{0,
-%% %% 1\}$, where $\mathcal{L}$ is the set of all expressions in our
-%% %% representation language. We say that $e \in \mathcal{L}$ \emph{solves}
-%% %% task $t$ if $t(e) = 1$. Our goal is to solve as many tasks as
-%% %% possible, as efficiently as possible. That is, suppose that there is a
-%% %% distribution over tasks $t \sim  p_t(\cdot)$, we would like an ordering $O
-%% %% = e_1, \dots $ on expressions in $\mathcal{L}$ such that if $p_t(t_1)
-%% %% > p_t(t_2)$ then the first expression in $O$ to solve $t_1$ comes
-%% %% before the first expression in $O$ to solve $t_2$.
-
-%% %% We motivate a solution to this problem by considering a heirarchical
-%% %% Bayesian formulation of the task distribution. Suppose we believe that
-%% %% there is some underlying structure to $p_t(\cdot)$ that is explained by
-%% %% the fact that for each task $t$ there is a latent expressions $e \in
-%% %% \mathcal{L}$ drawn from a stochastic grammar $\mathcal{G}$ and that
-
-%% %% \begin{align}
-%% %% e &\sim p_\mathcal{L}(\cdot) = \mathcal{G}\\
-%% %% t | e &\sim \text{U}\{ \text{tasks solved by }e\}.
-%% %% \end{align}
-
-%% %% Then $p_t(t) = \sum_{e_j \in \mathcal{L} \text{ s.t. } t(e_j) = 1}
-%% %% p_\mathcal{L}(e_j)$. If we make the assumption that for the set $E$ of
-%% %% expressions in this sum, there is some expression $e^*$ such that
-%% %% $p_\mathcal{L}(e^*) >> p_\mathcal{L}(e), e \neq e^*$, then we have, approximately, 
-%% %% \[ p_t(t) \propto} p_\mathcal{L}(e^*). \]
-
-%% %% This suggests a way to find the ordering $O$: infer a stochastic
-%% %% grammar $G$, under the assumption that $G$ favors one solution to a
-%% %% given task over any other, and order the elements of $\mathcal{L}$
-%% %% according to $G$.
-
-%% %% Thus, our basic procedure will be to search for a single ``good''
-%% %% solution for as many tasks as we can, and to use those solutions to
-%% %% estimate a stochastic grammar. Our search will be performed in the
-%% %% order of the most recently inferred grammar. So as our ordering
-%% %% becomes more efficient and we can solve more tasks, we learn better
-%% %% stochastic grammars.
-
-%% %% Formally then, our goal is to solve the following minimal feedback
-%% %% multi-task program induction problem: suppose we are given set of
-%% %% \emph{tasks} $T=\{t_k\}_{k=1}^K$ where each task is a function $t_i :
-%% %% \mathcal{L} \rightarrow \{0, 1\}$, where $\mathcal{L}$ is the set of
-%% %% programs in our language.  We say that $ p \in \mathcal{L}$
-%% %% \emph{solves} $t_i$ if $ t_i(p) = 1$. Our goal is to find a set of
-%% %% solutions $X = \{x_k\}_{k=1}^k \subset \mathcal{L}$ such that
-
-%% %% \[
-%% %% X = \argmax_{x_1, \dots, x_K \in L} \sum_k t_k(x_k).
-%% %% \]
-
-%% Based on the motivation set out above, we present an iterative
-%% strategy. Given a stochastic grammar $\mathcal{G}$ which defines a
-%% distribution over the primitives from $\mathcal{L}$ and a integer
-%% \emph{frontier size} $N$, we perform two steps in each iteration.  In
-%% the first step, we explore the $N$ highest probability programs
-%% according to $\mathcal{G}$. In the second step, we compress the
-%% programs that solve any of our tasks, and we use that compression to
-%% define a new grammar. Repeating these two steps, we effectively
-%% reparameterize our search space to include programs that are more likely
-%% to be solutions to our tasks.
+But, as mentioned before, our question is how to solve these problems
+when we lack knowledge of this kind of built in structure. \\
 
@@
 
-%% \section{}
-%% Let $\mathcal{F}$ be a family of distributions over $\mathcal{L}$ such
-%% that for any $\mathcal{D} \in \mathcal{F}$ we can a) evaluate the
-%% probability of program $p$, $P_D(p)$, and we can b) enumerate elements
-%% from $D$ in descending order.
-
-%%  To motivate a solution to our problem, we begin with the hierarchical
-%% Bayesian formulation of multi-task program induction found in Liang et
-%% al.~\cite{liang10programs}. The general idea there is to estimate
-%% solution programs for a set of related problems by inferring the posterior
-%% distribution over solutions under the assumption that all the solution
-%% programs are generated from a common latent stochastic grammar. By
-%% jointly inferring solutions and the latent grammar, we find a set of
-%% solutions that favors reuse of subcomponents.
-
-%% While this framework is appealing in its elegance and Bayesian
-%% interpretation, it cannot be straightforwardly adapted to the
-%% situation we are considering here because their proposed MCMC
-%% algorithm relies on a local log-likelihood signal, the very kind that
-%% we are assuming we do not have access to.
-
-%% To begin, we provide a high level overview of the algorithm and the
-%% problem that it solves. The problem setup requires a representation
-%% language for programs and a family of distributions over expressions
-%% in that language, but we leave these details for the following section
-
 \section{Extracting Concepts from Programs}
 
-Our goal is to solve the following minimal feedback multi-task program
+Our goal is to solve the following multi-task program
 induction problem: suppose we are given a set of \textbf{tasks}
@@
 
-Our best-first policy is to expand the next partial solution
-with the highest value. If a partial solution has multiple leaves that
-are not primitive combinators, we expand the left most one first,
-since it is the choice of left combinator that defines the type of its
-right combinator in an application pair. That is, the parent node
-applies its constraint on valid solutions via the left child, not the
-right child. We want to apply constraints as early as possible, so
-expanding left child first prunes more partial solutions.
+Our best-first policy is to expand the next partial solution with the
+highest value. If a partial solution has multiple leaves that are not
+primitive combinators, we expand the left most one first, since it is
+the choice of left combinator that defines the type of its right
+combinator in an application pair. That is, the parent node applies
+its constraint on valid solutions via the left child, not the right
+child. We want to apply constraints as early as possible, so we expand
+the left child first.
 
@@
 the reward function is binary, it is deterministic, and it cannot
-provide feedback about partial solutions. In ongoing work we are
+provide feedback about partial solutions. In ongoing work, we are
 exploring a more probabilistic version of the algorithm that can take
-into account continuous rewards, noisy data, and partial solutions.
+into account continuous rewards, noisy data and partial solutions.
 
@@
 unsolved tasks.  Our algorithm requires, as human learners do, a
-sufficiently graded ``curriculum'', or spectrum of problems to solve --
+sufficiently graded ``curriculum,'' or spectrum of problems to solve --
 simple problems provide necessary stepping stones to building complex

### 2013-04-24 drafts/ijcai2013/camera_ready/Dechter.IJCAI13.CameraReady.tex
Removed acknowledgements from title and put them in a separate acknowledgements section at the end of the paper. Added acknowledgement to Google as well.
Minor edits to acknowledgements.
Added Ryan's acknowledgment.

@@
 
-\title{Bootstrap Learning via Modular Concept Discovery\thanks{We acknowledge the support of ARO MURI W911NF-08-1-0242.}}
+\title{Bootstrap Learning via Modular Concept Discovery}
 
@@
 
+\section{Acknowledgements}
+This work was supported by ARO MURI contract W911NF-08-1-0242 and by a
+grant from Google. It was also funded in part by DARPA Young Faculty
+Award N66001-12-1-4219.
+
 %% The file named.bst is a bibliography style file for BibTeX 0.99c

### 2013-05-21 drafts/general_writeup/writeup.tex
Minor changes to writeup.
Started section on Genetic Programming in the general writeup.

@@
 learned datum is used to represent a later learned datum, we will term
-\emph{bootstrapping}. We say that $x$ \empth{bootstraps} $y$ if the
-conditional Kolmogorov complexity $K(y|p_x) < r$ when $U(p_x) = x$ and
+\emph{bootstrapping}. We say that $x$ \emph{bootstraps} $y$ if the
+conditional Kolmogorov complexity $K(y | p_x) < r$ when $U(p_x) = x$ and
 $r$ is again our resource bound.
@@
 Briggs and O'Neill in using combinators for genetic
-programming~\cite{briggs2006functioinal}.
+programming~\cite{briggs2006functional}.
 
@@
 
+\section{Genetic Programming}
+
+The most directly related work to ours is that on Genetic Programming.
+Therefore, we need to declare the differences between our work and genetic
+programming in as clear a way as possible. This will consist of two steps: 1)
+we need to review the literature and understand what its achievements are; 2)
+we need to implement a GP system and show how our system differs from a GP
+system in practice.
+
+An implementation of a symbolic regression genetic programming system is
+TinyGP, Java source code for which can be found in Appendix B
+of~\cite{poli08:fieldguide}.
+
+There are five basic things that need to be specified in order to create a GP
+algorithm. We need to determine the terminal set, the function set, the fitness
+function, the GP parameter values, and the terminal solution designation. The
+terminal set consists of the external inputs (i.e.  variables of the form x, y,
+etc.), functions with no arguments (e.g., rand and other functions with side
+effects), and constants.
+
+The function set is a set of functions and their arities. Function sets must
+satisfy \emph{closure}, which consists of \emph{type consistency} and
+\emph{evaluation safety}. Type consistency means that all functions return
+values of the same type. Evaluation safety means that functions do not throw
+errors.
+
+The fitness function is the objective function.
+
+The GP parameters are various: population size, probabilities of performing
+genetic operations (crossover and mutation), selection mechanism, etc.
+
+Finally, we need to specify termination condition. These include the number of
+runs or other predicates based on which the program should terminate.
+
+There are several papers by O'Reilly that analyze GP that are worth looking at:
+\cite{OReilly:thesis} \cite{OReilly:1995:tabbGP}
+
+\subsection{Modulearity and Hierarchy in GP}
+
+Koza's Automatically defined functions (ADFs) are the most popular way of
+evolving reusable components.
+In automatically defined functions, we explicitly add auxiliary functional genetic material that the main function can call.
+Typically, the ADFS need different terminal and function sets (for example, ADFs need argument terminals).
+In this scheme, each function only has access to its own ADFs.
+See Olsson's ADATE system where subfunctions as anonymous lambda expressions are extracted~\cite{DBLP:conf/eurogp/Olsson9}.
+
 \bibliographystyle{plain}

### 2013-05-24 drafts/general_writeup/writeup.tex
Modifications to GP section of general writeup.
Started section on density estimation with program planning.

@@
 Koza's Automatically defined functions (ADFs) are the most popular way of
-evolving reusable components.
-In automatically defined functions, we explicitly add auxiliary functional genetic material that the main function can call.
-Typically, the ADFS need different terminal and function sets (for example, ADFs need argument terminals).
-In this scheme, each function only has access to its own ADFs.
-See Olsson's ADATE system where subfunctions as anonymous lambda expressions are extracted~\cite{DBLP:conf/eurogp/Olsson9}.
+evolving reusable components. In automatically defined functions, we explicitly
+add auxiliary functional genetic material that the main function can call.
+Typically, the ADFS need different terminal and function sets (for example,
+ADFs need argument terminals). In this scheme, each function only has access to
+its own ADFs. See Olsson's ADATE system where subfunctions as anonymous lambda
+expressions are extracted~\cite{DBLP:conf/eurogp/Olsson9}.
+
+\subsection{Software}
+from page 148 of \cite{poli08:fieldguide}:
+\begin{itemize}
+\item Lil-GP (Punch and Zongker, 1998)
+\item ECJ (Luke, Panait, Balan, Paus, Skolicki, Popvici, Harrison, Bassett,
+Hubley, and Chircop, 2000-2007)
+\item Open Beagle (Gahge and Parizeau, 2002)
+\item GPC++ (Fraser and Weinbrenner)
+\item commercial: Discipulus (RML Technologies)
+\item review: Foster, 2001 
+\end{itemize}
 
@@
 
+\section{Density estimation with program planning} 
+Density estimation -- an unsupervised learning task whose importance is
+increasing in modern machine learning -- is the process of learning the
+probability density function associated with some set of data. The utility of
+doing this comes from the fact that a low dimensional representation of the
+learned density function is useful in other tasks.
+
+Suppose, for example, that we wanted to learn the distribution over real-world
+images from a dataset of images that we download from the web. The space of
+possible images is extremely large, but space of images that are probable in
+the real world is vanishingly small within the space of all possible images. To
+learn the distribution of images means that a new real world image will be
+assigned a relatively high likelihood under our learned distribution and noise,
+for example, would be assigned a low likelihood.
+
 \end{document}

### 2013-05-31 drafts/general_writeup/writeup.tex
Started new introduction to the general writeup document.

@@
 \usepackage{amsmath} 
-
+\usepackage{algpseudocode}
 \usepackage{ amssymb }
 \usepackage{listings}
+\usepackage{syntax}
+
+\newcommand{\funtype}[2]{#1 \rightarrow #2}
+\newcommand{\weightedrule}[3]{[#1 \mapsto #2 \, | \, #3]}
+\newcommand{\typedexp}[2]{#1 : #2} 
+\newcommand{\type}[1]{type(#1)}
+
+\renewcommand{\grammarlabel}[2]{#1 #2}
+\newcommand{\given}{\, | \,}
+
+\DeclareMathOperator*{\argmin}{argmin}
+\DeclareMathOperator*{\argmax}{argmax}
+
 \lstloadlanguages{Haskell}
@@
 \newcommand{\evaluator}{\emph{Evaluator }}
+\section{The problem}
+In all but the simplest problems, problem solving agents suffer from vast
+search spaces. Much of the research in AI, machine learning, and optimization
+focuses on how to overcome this search problem. Local information can help find
+better nearby solutions but make few gaurantees about global optima. Heuristic
+search techniques make use of the structure of the search space and information
+about the utility of partial solutions to hopefully minimize the numeber of
+candidate solutions that need to be explored. Other frameworks attempt to
+approximate the target search space with a more tractable, convex, surrogate
+search space.
 
-\section{The Learning Problem}
-We want to characterize the general task of the learner, whom we will
-call Linda. At various intervals, Linda is provided with sensory data
-$x$. In reponse, she is given the opportunity of submitting a program
-$\phi$ to an evaluator, henceforth named Dr. Eval. Dr. Eval produces a
-reward $R(\phi, x; \bar{y})$ which he delivers back to Linda, where
-$\bar{y}$ simply denotes any other variables Linda is not aware of
-that influences Dr. Eval's reward. Linda processes the information,
-hopefully learning from her experience with Dr. Eval.
+But since the begining of AI, a central idea -- at center stage at times and
+lurking in the shadows at others -- is that as problem solving agents solve
+more problems they should learn to solve problems better.
 
-What kinds of scenarios can this learning paradigm (naturally)
-contain? Most basically, Linda might be a simple organism: the data
-$x$ might be local chemical gradients, the program $\phi$ the motor
-program Linda will execute upon sensing $x$, and Dr. Eval might return
-the cumulative effect upon Linda of executing that motor program. Note
-that in this case $\phi$ does not take $x$ as an input. It has no
-input; it is merely a motor program built anew for each sensory datum
-that Linda receives.
+Here, we present a unified perspective on learning to solve problems, grounding the
+problem as one of Bayesian inference over a grammar that generates candidate
+problem solutions.
 
-Of course, we might want Linda, the still humble organism, to be
-capable of learning a higher-level planning procedure. Such a
-procedure would be a program that, given sensory data, would produce a
-motor program. In our framework, this would be represented by letting
-$x$ stand for a list of data (maybe several days worth of discrete
-food finding missions) and letting $\phi$ be the planing
-procedure. Dr. Eval would accordingly take $\phi$, apply it to each
-datum in $x$ and return a cumulative reward of some sort. We will
-defer for now the problem of deferred reward, in which most learners
-do not receive immediate signals about the utility of their actions.
+Types:
 
-Moving upward somewhat in the hierarchy of intelligent beings,
-consider Linda the electrical engineer. She is requested to construct
-a circuit $\phi$ to a certain specification $x$. Dr. Eval, her
-professor tests the circuit against the requested specification and
-gives her a grade. Or perhaps she doubles as the evaluator. As in the
-case of Linda the simple organism, we can conceive of $\phi$ as the
-general circuit creation procedure rather than the circuit, and create
-a higher-level learning problem.
+\begin{grammar}
+<$t$> ::= $\tau$, $\sigma$  (type variables)
+  \alt $a$, $b$ (concrete types e.g. real, bool, int, etc.)
+  \alt $\funtype{t'}{t}$ (function type)
+\end{grammar}
 
-A different perspective on this learning problem is one that is
-removed from this supervised reinforcement paradigm: instead of
-acheiving a reward of some sort, we can imagine that Linda wants to
-find the program $\phi$ which generated the data $x$. Here, Dr. Eval
-is the world or some simulator of the world and a comparator which
-signals to Linda whether she has successfully simulated the data. This
-paradigm corresponds to a deterministic version of the generative
-model learning task common these days in probabilistic machine
-learning and AI.
+Expressions: 
+\begin{grammar}
+<$\typedexp{e}{t}$> ::= $\typedexp{c}{t'}$ s.t. $t' = mgu(t,
+\type{c})$ (substitute a terminal tree that unifies with the requesting type)
+  \alt ($\typedexp{e}{\funtype{t'}{t}} \;\;\; \typedexp{e}{t'}$)
+\end{grammar}
 
-TODO: concept learning
-TODO: physicist (or scientist in general)
+First we will present the model over solutions in which our problem solving
+agent will be doing inference. Second, we will present a hybrid inference and
+problem solving algorithm.
 
-\section{A Universal Basis for Computation}
-Our goal is to explain the minimal conditions necessary for concept
-learning, and to do so we posit a minimal model of
-computation. Various universal models of computation have been
-proposed; the most famous is the Turing machine. Though the Turing
-machine is an intuitive abstract machine, and though its construction
-corresponds to various parts of real-world computing machines,
-formally simpler models of computation have been proposed. Most
-importantly is the Lambda calculus, proposed by Church, and the basis
-of modern functional programming languages. The Lambda calculus is a
-rewrite system that is provably equally powerful to Turing
-machines. Despite its popularity, the implementation of the Lambda
-calculus is complicated by the presence of variables bound by the
-abstracting $\lambda$s. The necessary relationships between variables
-and their binding $\lambda$s means that the syntax of the Lambda
-calculus is not entirely modular. That is, if we represent an
-expression in the lambda calculus as a tree, then every subtree of this
-expression tree is not itself a valid expression tree in the langauge.
+\section{The model}
+Our goal in this section is to describe a generic probability distribution over
+solutions spaces. Learning to search will be implemented by learning a
+distribution over the solution space which describes the probability of a 
+candidate solution being the solution for a given problem.
 
-An alternative to the $\lambda$ calculus we use the variable-free
-model of computation known as Combinatory Logic (CL). CL was
-introduced by Shonfinkel in order to avoid of the mathematical
-complexities of $\lambda$-calculus. The CL has a simpler syntax than
-$lambda$-calculus: a combinator $C$ is merely the application of two
-other combinators $ C = A B$. The semantics of the language is defined
-by the primitive combinators chosen for the particular instance of
-$CL$. It is well known that a universal basis of two primitive
-combinators exists, the $SK$ basis. However, expressions in CL are
-much more readible if we extend this primitive basis with a few other
-high order combinators. Namely, we include identity $I$, and two
-composition combinators $B$ and $C$.
+Problems vary considerably in their solution spaces. Some are numeric and
+continuous, others are integral and discrete, and still others consist of
+structured data like trees. Some solutions have a finite number of elements, or
+are bounded in some fashion. Others are infinite and unbounded. To present a
+framework for general problem solving, we need a representation of a solution
+space that subsumes all of these possibilities.
 
+As has long been noted by
+Genetic Programming researchers, any solution space can be modeled as a set of
+programs that generate space. As an example, if the solution space to a problem
+is the set of all boolean circuits constructed from some set of gates, then the
+corresponding space of programs is the set of programs that wire up boolean
+circuits from those gates.
+
+How does this realization help us? It means that if we can formulate a single
+model that places a distribution over solution distributions by describing a
+distribution over program generating processes.
+
+We will first describe a simplified version of the hierarchical model over programs that we will be using. We start with a base grammar $\mathcal{G}$ of the form 
+\section{The model}
+The elements of any solution space can be represented as programs. u
  \section{History and related literature}
@@
 \end{quote}
-
-\section{Domains for concept learning}
-
-\section{Coding Theory}
-Consider a distribution $P(\cdot)$ over strings in some set $X \in
-\{0, 1\}^*$. Our goal is to find a code $D$ that minimizes the
-expected length of a sample from $X$.
 
@@
 arguments.)
-
-\section{Intuitive Psychology}
-How can we express the knowledge and relationships of intuitive
-psychology as a set of expressions in combinatory logic? To simplify
-this task, we will attempt to express these relationships in Haskell,
-a sugared and polymorphically typed lambda calculus, knowing that any
-expression in this lambda calculus can be systematically translated
-into combinatory logic. Most importantly, any modularization and
-compositional structure present in the Haskell code will remain
-present in its CL translation.
-
-Let's imagine the following situation. You observe several agents
-(say, Josh, Max, and Chris) in various scenarios in which they choose,
-from among the various options, one of three bars (say, a Zone bar, a
-Luna bar and a dark chocolate bar). We will try to represent our
-knowledge about these agents, their propensities, and the properties
-of these bars in Haskell, as a demonstration that such a
-representation is both possible in this langauge, and can be somewhat
-natural.
-
-Let's first define the primitive types and combinators at our disposal:
-
-\begin{lstlisting}[mathescape]
--- | Primitive Types
--------------------------
-Int -- integers 1, 2, etc.
-Bool -- {True, False}
-Person $\triangleq$ Int
-Object $\triangleq$ Int
-Scenario  $\triangleq$ (Person, [Object])
-
--- | Primitive Combinators
-------------------------
--- List Combinators:
-map :: (a -> b) -> [b] -> [a]
-foldl1 :: (a -> a -> a) -> [a] -> a
-elem :: a -> [a] -> Bool
-head :: [a] -> a
-tail :: [a] -> [a]
-case :: [(Bool, a)] -- a set of boolean tests and the result if True
-        -> a -- the default result (else clause) 
-        -> a
-(:) :: a -> [a] -> [a] -- cons
-(++) :: [a] -> [a] -> [a] -- concatenate
-
--- Boolean Combinators
-If _ Then _ Else :: Bool -> a -> a
-And :: Bool -> Bool -> Bool
-Or ::  Bool -> Bool -> Bool
-Not ::  Bool -> Bool -> Bool
-
--- Arithmetic Combinators
-(>) :: Int -> Int -> Bool
-(<) :: Int -> Int -> Bool
-(==) :: Int -> Int -> Bool
-(+) ::  Int -> Int -> Bool
-(-) :: Int -> Int -> Bool
-1, 2, $\dots$, 10 :: Int -- | the integers, 1 through 10
-
--- Givens
-------------
-
-Josh :: Person -- | this is just shorthand, i.e., Josh would be mapped
-                  to some integer, say, 1, Max to 2, etc..
-Max :: Person
-Chris :: Person
-
-ZoneBar :: Object
-LunaBar :: Object
-ChocolateBar :: Object
-
--- Learned representation
--------------------------
-choose' :: Scenario -> Object
-choose' (person, objs) = max' (map (util' person) objs)
-
-max' :: [Int] -> Int
-max = foldl1 (>)
-
-util' = Person -> Object -> Int
-util' person obj = case (utilityCases' person obj) 0 -- | 0 is the default utility in this case
-
-utilityCases' :: [Person -> Object -> (Bool, Int)]
-utilityCases' = \p -> \o -> (
-                     \p -> \o -> (isHealthyPerson p, healthyPersonCalculus o)
-
-\end{lstlisting}
 

### 2013-06-05 drafts/general_writeup/writeup.tex
Added to first section of writeup.

@@
 \usepackage{syntax}
+\usepackage{mdframed}
 
@@
 
-\DeclareMathOperator*{\argmax}{arg\,max}
-
 \begin{document}
@@
 focuses on how to overcome this search problem. Local information can help find
-better nearby solutions but make few gaurantees about global optima. Heuristic
+better nearby solutions but makes few gaurantees about global optima. Heuristic
 search techniques make use of the structure of the search space and information
@@
 problem solutions.
-
-Types:
-
-\begin{grammar}
-<$t$> ::= $\tau$, $\sigma$  (type variables)
-  \alt $a$, $b$ (concrete types e.g. real, bool, int, etc.)
-  \alt $\funtype{t'}{t}$ (function type)
-\end{grammar}
-
-Expressions: 
-\begin{grammar}
-<$\typedexp{e}{t}$> ::= $\typedexp{c}{t'}$ s.t. $t' = mgu(t,
-\type{c})$ (substitute a terminal tree that unifies with the requesting type)
-  \alt ($\typedexp{e}{\funtype{t'}{t}} \;\;\; \typedexp{e}{t'}$)
-\end{grammar}
 
@@
 
-As has long been noted by
-Genetic Programming researchers, any solution space can be modeled as a set of
-programs that generate space. As an example, if the solution space to a problem
-is the set of all boolean circuits constructed from some set of gates, then the
-corresponding space of programs is the set of programs that wire up boolean
-circuits from those gates.
+As has long been noted by Genetic Programming researchers, any solution space
+can be modeled as a set of programs. As an example, if the solution space to a
+problem is the set of all boolean circuits constructed from some set of gates,
+then the orresponding space of programs is the set of programs that wire up
+boolean circuits from those gates.
 
-How does this realization help us? It means that if we can formulate a single
+How does this realization help us? It means that we can formulate a single
 model that places a distribution over solution distributions by describing a
@@
 
-We will first describe a simplified version of the hierarchical model over programs that we will be using. We start with a base grammar $\mathcal{G}$ of the form 
-\section{The model}
-The elements of any solution space can be represented as programs. u
+We will first describe a simplified version of the hierarchical model over
+programs that we will be using. We start with a base grammar $\mathcal{G}$ of
+the form shown in Figure~\ref{fig:grammar}.
+
+\begin{figure}
+  \begin{grammar}
+    <$e$> ::= $p_1$ | $p_2$ | \dots | $p_N$   \text{ (primitive combinators)}
+              \alt (e e) \text{ (application) }
+  \end{grammar}
+ \caption{A grammar for a typed combinatory programming language.\label{fig:grammar}}
+\end{figure}
+
+Our probabilistic model is motivated by the fact that while $\mathcal{G}$
+represents the language in which solutions are written, in any particular
+domain there is a distribution $\mathcal{H}$ over solutions that is itself drawn
+according to an adaptor grammar~\cite{} based on $\mathcal{G}$ and some
+parameters $\theta$.
+
  \section{History and related literature}

### 2013-06-07 drafts/general_writeup/writeup.tex
Started section on learning physics.

@@
 \usepackage{mdframed}
+\usepackage{graphicx}
 
@@
 domain there is a distribution $\mathcal{H}$ over solutions that is itself drawn
-according to an adaptor grammar~\cite{} based on $\mathcal{G}$ and some
-parameters $\theta$.
+according to an adaptor grammar prior~\cite{} based on $\mathcal{G}$ and some
+parameters $\theta$. $\mathcal{H}$ is a probabilistic context-free grammar
+(PCFG), and this PCFG generates programs $\{e_n\}$ that solve problems in the
+domain. Finally, any candidate solution $s_n$ is scored according to its
+likelihood given the latent $e_i$. Thus we have the generative model pictured in
+Figure~\ref{fig:generativeModel}.
+
+\begin{figure}[h] 
+  \includegraphics[width=.3\linewidth]{./figures/basicGM.pdf}  
+  \caption{Generative model for expressions.\label{fig:generativeModel}}  
+\end{figure}
+
+The likelihood function $p_{s|e}(\cdot | e)$ takes the place in our framework that the fitness function would take in genetic programming.
+
+How does the specification of the generative model relate to the specification
+of the problems our agent needs to solve? The idea is that for any objective
+function $ r:S \rightarrow \mathcal{D}$, where $S$ is the set of solutions and
+$\mathcal{D}$ is the domain, we can map $\mathcal{D}$ to the interval $[0,1]$,
+thereby generating a likelihood function that scores candidate solutions
+relative to the latent program $e_n$.
+
+How can we learn in this model? Our goal is to learn the distribution
+$\mathcal{H}$.
+
+\section{Potential application}
+\begin{itemize}
+  \item Learning GP kernels
+  \item Learning control circuits (signals \& systems book)
+  \item Gen audio
+    Ryan suggested this. MAX MSP and its open source version (Pd or
+    Pd-extended) allow composers to generate sounds by using a graphical
+    programming language. 
+  \item Gen vis
+  \item Compositional object detectors
+  \item Constructive solid geometry
+\end{itemize}
+
+\subsection{Discovering the laws of physics}
+One of the main challenges that scientists face when discovering theories that
+explain the world's phenomena is that they must perform inference over the
+latent properties of objects in the world and the way in which these properties
+relate to each other and the observable world. For example, Newton's first law
+of motion postulates that things called forces exist in the world, and that
+they obey the property that their vector sum is equal to the mass times the
+acceleration of the object on which they act.
+
+Here we consider simpler example of this joint inference problem: suppose you
+observe various sets of charged particles interacting, but you do not know that
+they are charged or that the property of electromagnetic charge exists. How can
+you learn both that each particle has a charge, which is a positive or negative
+scalar property, and that Coulomb's law dictates the force between those
+particles.
+
+Understanding Coulomb's law consists of holding that the following are true: 1)
+every particle has a unique charge and, under normal circumstances, the charge
+of a particle does not change. 2) Charge is a scalar quantity on the real line
+that can take values on in the reals. 3) The force on any object $i$ with charge $q_i$ at location $r_i$ due to charged object j with charge $q_j$ at location $r_j$ is
+
+\begin{align}
+  F_i \propto \frac{q_i q_j}{|r_i - r_j|^2}
+\end{align}
+
+To simplify matters, and remove the complexities of vector arithmetic, we can
+suppose that this is a one-dimensional problem:
+
+\begin{align}
+  F_i \propto \frac{q_i q_j}{(r_i - r_j)^2} \text{sign}(r_i - r_j)
+\end{align}
+
+Conceptually, this force law has two components. The first is that the force is
+inversely propotional to the distance between the objects squared. The second
+is that the force is proportional to the product of the charges. The scientist
+can learn these two facts separately. By comparing the force between two
+objects at different distances, she can discover the inverse squared distance
+behavior. By noting that the force differs by the same multiplicative factor
+when you swap object a for object c regardless of the paired object b.
+
+In order to do this though, the scientist needs to be able to focus her
+attention an different scenarios.
 

### 2013-06-10 drafts/planner/planner.tex
Added a writeup of a possible generative model for the planner

\documentclass{article}
\usepackage{clrscode4e}
\usepackage{tikz}
\usetikzlibrary{fit,positioning}
\usepackage{textcomp}
\usepackage{amsmath}
\usepackage{amssymb}

\begin{document}

\section{Overview}
We want to learn how to plan using program induction.
The general idea is that we want a set of compact programs that enable us to sample plans that have a high probability of solving the given tasks.

A \emph{plan} is a list of \emph{actions}.
Each action performs some change of \emph{state}.
Our objective is to find a plan whose actions transform an initial state in to a goal state; there may be many different goal states, or only one.

The learned programs will be either \emph{operators} or \emph{heuristics}.
Operators extend an existing plan by proposing new action sequences.
Heuristics guide the search by rewarding/penalizing certain states.
Given a set of operators and heuristics, a plan may be sampled as follows:

Extend the existing plan by applying all available operators, producing a set of new plans.
Sum the heuristic values for each plan to produce an energy for each plan;
sampling the new plan with probability proportional to the minus exponential of its energy multiplied by an inverse ``temperature'' $\beta=1/T$.
The temperature controls the strength of the heuristic functions:
in the limit as $T\to\infty$, all actions become equally likely, and in the limit as $T\to 0$, the action with the lowest heuristic score is always chosen.

If $\{h_j\}$ are the heuristic functions and $\{(p_i,s_i)\}$ the successor plans and final states to choose from, then
\begin{eqnarray}
\hat{H} &=& \sum_j h_j\mbox{, the ``Heuristic'' or ``Hamiltonian''}\\
Z &=& \sum_i \exp\left( -\frac{\hat{H} s_i}{T} \right)\\
\mbox{P}[(p_i,s_i)] &=& \frac{1}{Z} \exp\left( -\frac{\hat{H} s_i}{T} \right)
\end{eqnarray}

The types of operators and heuristics are:
\begin{eqnarray}
\hat{o}_i \; &:& \; [\mbox{Action}]\times \mbox{State} \to [\mbox{Action}]\\
h_j \; &:& \; \mbox{State} \to \mathbb{R}\cup\{\infty\}
\end{eqnarray}

This model draws from Simon and Newell's models.
According to their models, human planning is sequential, heuristic guided, and doesn't involve (much) backtracking, features which are shared by this sampling procedure.

\section{Generative Model}
The heuristics and actions are drawn from a grammar, $G$, which is drawn from a prior over grammars controlled by the hyperparameter $\lambda$.
Something like $\mbox{P}[G|\lambda]\propto \exp\left(-\lambda |G|\right)$ seems reasonable.
There should also be a prior over the number of heuristics and the number of actions.

We're given $N$ tasks.
For each task a plan $p_i$ is sampled (see section \ref{sec:sample})
which is used to solve task $t_i$.
There is some problem-dependent $P[t_i | p_i]$ which should be higher for plans that do a better job of solving the task, and zero for plans that don't solve the task.
One possibilty is 
\begin{equation}
P[t_i | p_i] \propto \exp\left( -\alpha \mbox{Loss}(t_i,p_i)\right)
\end{equation}
For tower building the loss function is the number of blocks used.
For constructive solid geometry, one possibility is the sum of the $\mbox{L}_2$ norms of the pixel differences between the target image and the rendered image when projected on to various 2D planes.
Or, we could look at the number of differing voxels.

\begin{figure}
\centering
\begin{tikzpicture}
\tikzstyle{main}=[circle, minimum size = 10mm, thick, draw =black!80, node distance = 16mm]
\tikzstyle{connect}=[-latex, thick]
\tikzstyle{box}=[rectangle, draw=black!100]

\node[main, fill = white!100] at (0,2) (lambda)  {$\lambda$ };
\node[main, fill = white!100] at (0, 0) (grammar)  {$G$ };
\node[main, fill = white!100] at (2,-2) (actions) {$\{\hat{o}_i\}$ };
\node[main, fill = white!100] at (-2,-2) (heuristic) {$\{h_j\}$ };
\node[main, fill = white!100] at (0,-4) (psubk) {$p_k$ };
\node[main, fill = white!100] at (0,-6) (tsubk) {$t_k$ };
\path (lambda) edge [connect] (grammar);
\path (grammar) edge [connect] (heuristic);
\path (grammar) edge [connect] (actions);
\path (heuristic) edge [connect] (psubk);
\path (actions) edge [connect] (psubk);
\path (psubk) edge [connect] (tsubk);
\node[rectangle, inner sep=0mm, fit= (psubk) (tsubk),label=below right:N, xshift=-1mm, yshift=-8mm] {};
\node[rectangle, inner sep=4.4mm,draw=black!100, fit= (psubk) (tsubk)] {};
\end{tikzpicture}
\end{figure}

\section{Sampling}\label{sec:sample}
\begin{codebox}
\Procname{$\proc{Sample-Plan}(\{\hat{o}_i\}, \{h_j\}, depth)$}
\li \id{plan} = [ ]
\li \id{state} = \const{initial state}
\li \id{T} = \const{temperature}
\li \textbf{repeat} \id{depth} \textbf{times}\Do
\li \Comment{Get new plans and new states:}
\li \id{p'_i} = \id{\hat{o}_i}(\id{p}, \id{s})
\li \id{s'_i} = \func{Run-Actions}(\id{p'_i}, \id{s})
\li
\li \Comment{Compute energies and partition function:}
\li \id{E_i} = $\sum_j$ \id{h_j}(\id{s'_i})
\li \id{Z} = $\sum_i$ $\exp \left( -E_i / T \right)$
\li
\li \Comment{Sample next action, and then apply it:}
\li \id{i^*} $\sim$ \func{Multinomial}$\left(\frac{1}{Z}\exp(-E_1/T), \frac{1}{Z}\exp(-E_2/T), ...\right)$
\li \id{plan} = \id{plan} $\cdot$ \id{p_{i^*}}
\li \id{state} = \id{s'_{i^*}}
\li 
\li \Comment{Check if goal state is reached:}
\li \If \func{Is-Goal}(\id{state}) \Then
\li \Return \id{plan}, \id{state}
\li \End\End
\li \Comment{Failure to find a plan within depth \emph{depth}:}
\li \Return $\bot$
\end{codebox}

\end{document}

### 2013-06-10 drafts/inference/inference.tex
Added a writeup of how EM and variational inference can be used to perform inference in our generative model

\documentclass[12pt]{article}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{amsfonts}

\begin{document}
\section{EM Inference}
E Step: Compute $q(\{e_i\})=P(\{e_i\} | \{t_i\},G^{old})\propto P(\{t_i\}|\{e_i\})P(\{e_i\}|G^{old})$.
Observe that $q(\{e_1,e_2,\cdots\})=q(\{e_1\})q(\{e_2\})\cdots$

M Step: Find $G$ by
\begin{eqnarray*}
&\max_G& \mathbb{E}_q \left[ \log P(\{t_i|e_i\}) P(\{e_i\}|G)P(G|\theta) \right]\\
&\max_G& \log P(G|\theta) + \mathbb{E}_q \left[ \log P(\{e_i\}|G)\right] \\ 
&\max_G& \log P(G|\theta) + \sum_i \mathbb{E}_q \left[ \log P(e_i | G) \right]
\end{eqnarray*}
Consider each term in the sum on the r.h.s.:
\begin{eqnarray*}
\mathbb{E}_q \left[ \log P(e_i | G) \right] &=& \sum_{e_1}\cdots\sum_{e_i}\cdots \sum_{e_n} q(\{e_1,\cdots\})\log P(e_i|G)\\
&=& \sum_{\{e_{j\not=i}\}} q(\{e_{j\not= i}\})\sum_{e_i}q(e_i)\log P(e_i|G)\\
&=& \sum_{\{e_{j\not=i}\}} P(\{e_{j\not= i}\}|G^{old}, \{t_{j\not= i}\})\sum_{e_i}q(e_i)\log P(e_i|G)\\
&=& \sum_{e_i}  P(t_i|e_i)P(e_i|G^{old})/P(t_i|G^{old}) \log P(e_i|G)\\
&=& \mathbb{E}_{q_i} \left[ \log P(e_i|G) \right]
\end{eqnarray*}
Back to what we're maxing,
\begin{eqnarray*}
&\max_G& \sum_i \mathbb{E}_{q_i} \left[ \log P(e_i|G) \right]
\end{eqnarray*}
Draw from $q_i$ using importance sampling: draw from $P(\cdot | G^{old})$, and then weight by $P(t_i | \cdot)$.
Let the draws be $e_i^k$ and the corresponding weights be $w_i^k$.
Approximate the expectation by using this sample; now the maximized quantity is
\begin{eqnarray*}
&\max_G& \sum_i \sum_k w_i^k \log P(e_i|G)\\
&\max_G& \prod_{i,k}  \left( P(e_i|G) \right)^{w_i^k}
\end{eqnarray*}
This is a nice form: what we're maximizing is the probability of generating a set of parse trees, each of which is observed in proportion to $w_i^k$, which is proportional to $P(t_i|e_i)$.
We can perform this max using hill climbing or some sort of stochastic search, completing the ``M'' step of EM.

\section{Variational Inference}
It would be nice to represent a distribution over grammars.
This would, for example, allow generalization from a single task, permitting one-shot program induction.
It also would make our inference algorithm closer to the actual generative model.

As a first attempt, I'll approximation the posterior $P(\{e_i\},G | \{t_i\})$ using a factorized distribution (mean-field approximation):
$$
Q(\{e_i\},G) = q(G) \times \prod_i q(e_i) 
$$
where $Q$ is our approximation to the posterior.
Following the standard derivation of variational Bayes, we want to solve the equations
\begin{eqnarray*}
\log q(e_i) &\stackrel{+}{=}& \sum_G q(G)\sum_{e_{j\not= i}} q(\{e_j\}) \log P(e_i, t_i, \{e_{j\not= i}\}, \{t_{j\not= i}\}, G) \\
\log q(G) &\stackrel{+}{=}& \sum_{\{e_i\}} q(\{e_i\}) \log P(\{e_i\}, \{t_i\}, G )
\end{eqnarray*}
The second equation, of course, gives the same distribution over $G$ as EM:
\begin{eqnarray*}
\log q(G) &\stackrel{+}{=}& \sum_{\{e_i\}} q(\{e_i\}) \log P(\{e_i\}, \{t_i\}, G )\\
 &\stackrel{+}{=}& \sum_{\{e_i\}} q(\{e_i\}) \log P(\{e_i\} | G )P(G|\theta)\\
 &=& \log P(G|\theta) + \sum_{e_1} q(e_1)\log P(e_1 | G) + \sum_{e_2} q(e_2)\log P(e_2 | G)+...
\end{eqnarray*}
The first gives a \emph{different} distribution over $\{e_i\}$ than EM:
\begin{eqnarray*}
\log q(e_i) &\stackrel{+}{=}& \sum_G q(G)\sum_{e_{j\not= i}} q(\{e_j\}) \log P(e_i, t_i, \{e_{j\not= i}\}, \{t_{j\not= i}\}, G) \\
&\stackrel{+}{=}& \sum_G q(G) q(e_i) \log P(e_i, t_i, G) \\
&\stackrel{+}{=}& \log P(t_i | e_i) + \sum_G q(G) \log P(e_i | G) \\
q(e_i) &\propto& P(t_i | e_i) \prod_G \left[ P(e_i | E)\right]^{q(G)}
\end{eqnarray*}
This can be approximated by weighting samples from the obvious proposal distribution
$$
r(e_i) = \sum_G P(e_i | G) q(G)
$$
A reasonable form for $q(G)$ is (yet another) factored distribution:
\begin{eqnarray*}
q(G) &=& \prod_k q(z_k)\\
q(z_k) &=& \mbox{Beta}(\alpha_k, \beta_k)
\end{eqnarray*}
where each $z_k$ is the $k$th production probability.
So, we could do gradient descent on  $\{\alpha_k\}\cup\{\beta_k\}$.
For computing description length, we could say that there is a Bernoulli RV for each production controlling whether it is even in the grammar, with a Beta prior.
Then, we can easily find the expected description length of a grammar drawn from $q(G)$.
\end{document}

### 2013-06-10 drafts/inference/inference.tex
Added writeup of the linear program

@@
 Then, we can easily find the expected description length of a grammar drawn from $q(G)$.
+\section{Grammar induction via linear programming}
+In EM, we need to find the grammar that most compresses our corpus.
+We can approximate our likelihood and prior by saying that the negative log likelihood is equal to (up to an additive normalizing constant) the number of symbols in the corpus plus the number of symbols in the grammar.
+
+Taking this approximation as the true likelihood and prior, let $\lambda$ be the regularization coefficient for the grammar (eg, how much the negative log likelihood increases for each symbol in the grammar).
+For the sake of simplicity, assume that, if a tree is put in to the grammar, then all of its subtrees are also in the grammar.
+This condition also follows from the requirement that the grammar be in Chomsky Normal Form.
+Then, we want to find the grammar $G$ minimizing
+$$
+\lambda \times \left( \mbox{\# symbols in $G$}\right)+\sum_i w_i \mbox{\# symbols in $e_i$ when encoded using $G$}
+$$
+$\lambda$ should be in $(1,2)$, because adding a tree to the grammar should cost at least as much as having the tree in the corpus (otherwise we'll throw everything in the grammar) and should be $< 2$ so that repeated subtrees are taken up in to the grammar.
+
+Now I'll write down a linear program that corresponds to this minimization problem.
+Define the variable $T_j\in \{0,1\}$, which is 1 iff the $j$th subtree of the corpus is in the grammar.
+\begin{eqnarray*}
+&\min& \lambda \sum_j T_j - \sum_i w_i \sum_{j\in e_i} T_j\\
+&\mbox{subject to}& T_j - T_k \geq 0\mbox{, whenever $j$ is a subtree of $k$}
+\end{eqnarray*}
+The constraint matrix for this linear program is totally unimodular, meaning that every non-singular submatrix has determinant $\pm 1$.
+When this condition holds, the integer linear program has the same solution as the relaxed linear program.
+So, we can solve this using Simplex (or any other linear programming algorithm), subject to the constraint $T_j\in [0,1]$.
+
 \end{document}

### 2013-06-12 src/Strappy/Sample.hs
Changes for generating expressions along with their contexts, so that we can apply substitutions whenever we want.

@@ -64,8 +64,8 @@ sampleExprM gr@Grammar{grApp=p, grExprDistr=exprDistr} tp
                           annotateRequestedM tp e
 
 -- | Wrapper over sampleExpr that keeps trying to sample when it fails
-safeSample :: MonadRandom m => Grammar -> Type -> m Expr
-safeSample gr tp = do
+safeSampleWithContext :: MonadRandom m => Grammar -> Type -> m (Expr, (Int, M.Map Int 
+safeSampleWithContext gr tp = do
   maybeSample <- runMaybeT $ sampleExpr gr tp
   case maybeSample of
     Nothing -> safeSample gr tp
@@ -78,6 +78,13 @@ sampleExprs n library tp =
   where accSample acc _ = do
           expr <- safeSample library tp
           return $ Map.insertWith (+) expr 1 acc
+{-
+sampleExprsWithContext n library tp =
+  liftM (Map.map fromIntegral) $ foldM accSample Map.empty [1..n]
+  where accSample acc _ = do
+          expr <- safeSample library tp
+          return $ Map.insertWith (+) expr 1 acc
+-}
 
 -- | Uses breadth-first enumeration to "sample" a grammar
 -- This allows us to get many more programs

### 2013-06-12 src/Strappy/Type.hs

@@ -41,6 +41,7 @@ instance Show Type where
                       t2 = ts !! 1
 
 -- | type inference monad
+type Sub = Sub
 type TypeInference = StateT (Int, -- ^ next type var
                              M.Map Int Type) -- ^ Union-Find substitution
 

### 2013-06-12 drafts/inference/inference.tex

@@
 \begin{eqnarray*}
-&\max_G& \sum_i \mathbb{E}_{q_i} \left[ \log P(e_i|G) \right]
+&\max_G& \log P(G|\theta) + \sum_i \mathbb{E}_{q_i} \left[ \log P(e_i|G) \right]
 \end{eqnarray*}
@@
 \begin{eqnarray*}
-&\max_G& \sum_i \sum_k w_i^k \log P(e_i|G)\\
-&\max_G& \prod_{i,k}  \left( P(e_i|G) \right)^{w_i^k}
+&\max_G& \log P(G|\theta) + \sum_i \sum_k w_i^k \log P(e_i|G)\\
+&\max_G& P(G|\theta) \prod_{i,k}  \left( P(e_i|G) \right)^{w_i^k}
 \end{eqnarray*}
@@
 &\stackrel{+}{=}& \log P(t_i | e_i) + \sum_G q(G) \log P(e_i | G) \\
-q(e_i) &\propto& P(t_i | e_i) \prod_G \left[ P(e_i | E)\right]^{q(G)}
+q(e_i) &\propto& P(t_i | e_i) \prod_G \left[ P(e_i | G)\right]^{q(G)}
 \end{eqnarray*}
-This can be approximated by weighting samples from the obvious proposal distribution
+This can be approximated by weighting samples from the proposal distribution
 $$
@@
 For computing description length, we could say that there is a Bernoulli RV for each production controlling whether it is even in the grammar, with a Beta prior.
-Then, we can easily find the expected description length of a grammar drawn from $q(G)$.
+Then, we can find the expected description length of a grammar drawn from $q(G)$.
 \section{Grammar induction via linear programming}
@@
 
-Taking this approximation as the true likelihood and prior, let $\lambda$ be the regularization coefficient for the grammar (eg, how much the negative log likelihood increases for each symbol in the grammar).
+Taking this approximation as the true likelihood and prior, let $\lambda$ be the regularization coefficient for the grammar (how much the negative log likelihood increases for each symbol in the grammar).
 For the sake of simplicity, assume that, if a tree is put in to the grammar, then all of its subtrees are also in the grammar.
@@
 So, we can solve this using Simplex (or any other linear programming algorithm), subject to the constraint $T_j\in [0,1]$.
+\section{Drawing samples from a grammar}
+Our current sampling procedure, as described in the IJCAI paper, is a reasonable and natural way of sampling typed programs from a stochastic grammar.
+However, we don't know how to efficiently find ML estimates of the parameters of the grammar (solving it would require iteratively finding the zeros of very high-degree polynomials).
+If we instead sample from the PCFG, given that the program is typed, we can use the inside-out algorithm to estimate the parameters.
 
+We can efficiently sample from the distribution
+$$P(e | e\mbox{ is typed}, G)\propto \mathbb{I}[e\mbox{ is typed}] P(e | G)$$
+by sampling (or enumerating) using the distribution described in the IJCAI paper, and then reweighting the samples appropriately.
+We don't currently know how to sample directly from $P(e | e\mbox{ is typed}, G)$.
+Empirically, this proposal for importance sampling seems to work well on polynomial regression.
 \end{document}

### 2013-06-25 drafts/planner/planner_inference.tex
Added a writeup describing how EM and MCMC could be used to learn how to plan

\documentclass{article}
\usepackage{clrscode4e}
\usepackage{tikz}
\usetikzlibrary{fit,positioning}
\usepackage{textcomp}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{verbatim}

\begin{document}

\section{Overview}
This document describes one method of learning to plan via program induction.
It uses a simpler version of the generative model described in the other planner document.

A \emph{plan} is a sequence of programs $e_1, e_2, ...$ that are composed to produce $e_1\circ e_2 \circ ...$ and applied to an initial empty plan (an empty list, the identity function, a constant function, etc, depending on the specific domain).
Our domain has a large number of planning tasks.
We wish to learn a grammar such that, when programs are sampled from this grammar and composed to produce plans, they have a high probability of solving one of the planning tasks.
We'll do this using either Monte Carlo EM or AIP.
This document will describe the generative model and detail one means of performing posterior inference.

\section{Generative Model}
We're given $N$ tasks.
For each task $t$ a plan $\rho$ and a plan length $|\rho|$ is sampled, with the length of the plan chosen uniformly from $\{1,2,...,\ell\}$.
There is some problem-dependent $P(t | \rho)\in[0,1]$ which should be higher for plans that do a better job of solving the task, and zero for plans that don't solve the task.

The grammar produces plan $\rho = e_1\circ e_2\circ \cdots \circ e_k$ with probability
$$
P\left( e_1\circ e_2\circ \cdots \circ e_k | G \right) =
\begin{cases}
    \frac{1}{\ell}\prod_i P(e_i | G),& \text{if } 0<k\leq \ell\\
    0,              & \text{otherwise}
\end{cases}
$$
The generative model is illustrated in Fig. 1.

\begin{figure}
\centering
\begin{tikzpicture}
\tikzstyle{main}=[circle, minimum size = 10mm, thick, draw =black!80, node distance = 16mm]
\tikzstyle{connect}=[-latex, thick]
\tikzstyle{box}=[rectangle, draw=black!100]

\node[main, fill = white!100] at (0,2) (lambda)  {$\theta$ };
\node[main, fill = white!100] at (0, 0) (grammar)  {$G$ };
\node[main, fill = white!100] at (0,-2) (psubk) {$\rho_n$ };
\node[main, fill = white!100] at (0,-4) (tsubk) {$t_n$ };
\path (lambda) edge [connect] (grammar);
\path (grammar) edge [connect] (psubk);
\path (psubk) edge [connect] (tsubk);
\node[rectangle, inner sep=0mm, fit= (psubk) (tsubk),label=below right:N, xshift=-1mm, yshift=-8mm] {};
\node[rectangle, inner sep=4.4mm,draw=black!100, fit= (psubk) (tsubk)] {};
\end{tikzpicture}
\caption{Generative model used for the planner}
\end{figure}

\section{Inference}
Following the Monte Carlo EM derivation described in the ``inference'' document, we want to find the grammar maximizing
$$
\log P(G|\theta)+\sum_n \mathbb{E}_{q_n}\left[ \log P(\rho_n | G) \right]
$$
where
$$
\tilde{q}_n(e_1\cdots e_k) = P(t_n | e_1\cdots e_k)\prod_i P(e_i | G)
$$
Now, if we can sample from $\tilde{q}$, we're done.
We could just sample/enumerate random plans from $G$ and then do likelihood weighting, but if our frontier is of order $10^4$ or so, this will quickly become impractical for $\ell$ of order $10^1$.
Instead, we'll use MCMC to get samples from a proposal distribution, which we'll then reweight.
\subsection{Planning via MCMC}
We'll draw samples (plans) from a distribution $\tilde{g}(\cdot)$ that will serve as a proposal distribution for the expectation w.r.t. $q_n$.
This MCMC algorithm is not Metropolis-Hastings, but it has a stationary distribution which is similar to $q_n$.
Given a plan $\rho$ with $|\rho|<\ell$, the proposal distribution $h(\cdot | \cdot)$ is
$$
h(\rho e | \rho) \propto P(t|\rho e) P(e|G)
$$
A plan $|\rho|=\ell$ has $h(\emptyset | \rho)=1$.
Intuitively, the sampler is growing out a plan, starting from $\emptyset$, sampling the next move according to the posterior distribution upon the next program (greedy one-step lookahead).
The stationary distribution of this Markov chain is
$$
\tilde{g} (e_1 e_2 \cdots e_k) = P(e_1 | G) P(e_2 | G) \cdots P(e_k | G) P(t | e_1) P(t | e_1e_2) \cdots P(t | e_1e_2\cdots e_k)
$$
as can be verified by observing that $\tilde{g}(\rho e)=h(\rho e | \rho) \tilde{g}(\rho)$.
So, we reweight a plan $\rho=e_1e_2\cdots e_k$ by the quantity
$$
\frac{\tilde{q}}{\tilde{g}} = \frac{1}{P(t|e_1)P(t|e_1e_2)\cdots P(t|e_1e_2\cdots e_{k-1})}
$$
Each plan of length $\ell$ sampled contributes $\ell$ total samples (all prefixes of the plan).
Each total plan $\rho=e_1\cdots e_\ell$ contributes the following quantity to the maximization problem in $G$ (up to a normalizing constant):
\begin{eqnarray*}
&&\sum_{k\leq\ell} \frac{\tilde{q}(e_1\cdots e_k)}{\tilde{g}(e_1\cdots e_k)} \sum_{j\leq k} \log P(e_j | G)\\
&=&\sum_{k\leq\ell} \frac{1}{P(t|e_1)P(t|e_1e_2)\cdots P(t|e_1e_2\cdots e_{k-1})} \sum_{j\leq k} \log P(e_j | G)\\
&=&\sum_{j\leq \ell} w_j\log P(e_j | G)
\end{eqnarray*}
where I've introduced the quantities
$$
w_j=\sum_{m\geq j} \frac{1}{P(t|e_1)P(t|e_1e_2)\cdots P(t|e_1\cdots e_{j-1})}
$$
The $w_j$'s factor in a nice way that allows them to be computed in linear time using a dynamic program.
The following procedure samples a plan and returns the weights upon each program in that plan:
\begin{codebox}
\Procname{$\proc{Sample-Plan}(\rho)$}
\li \Comment{Draw the next program in the plan}
\li \id{e} $\sim$ $P(\rho\cdot | t, G)$
\li \Comment{If the plan should be extended, recurse, creating the suffix of the plan}
\li \If $|\rho e|<\ell$ \Then
\li \id{suffixWeight}, \id{suffixRewards} = \func{Sample-Plan}$(\rho e)$
\li \Else
\li \id{suffixWeight}, \id{suffixRewards} = $0$, $\emptyset$
\li \End
\li \Comment{To satisfy the boundary conditions of the recurrence, $P(t|\emptyset)$ is defined to be 1.}
\li \id{eReward} = $\frac{1+\mbox{\id{suffixRewards}}}{P(t|\rho)}$
\li \Return \id{eReward}, $\{(e, \mbox{\id{eReward}})\}\cup \mbox{\id{suffixRewards}}$
\end{codebox}

It may seem strange that, in the generative model, we produce plans by doing a random walk without any bias towards the target.
However, this is actually to be expected, because the target has not yet been generated.
When performing inference, we condition upon the random walk hitting the target; this conditioning has the same role as planning would have.
Rather than use rejection sampling, we use a stochastic, greedy search (the MCMC procedure) to bias us towards plans that hit the targets.
The $w_j$'s serve to correct for the biases introduced by this stochastic search.
They may also be thought of as ``rewards'' for each program used in a plan.
Note that their dependence upon likelihood at each point in the plan is non-trivial; initial programs get some of the credit for subsequent increases in likelihood, as is intuitively correct.

The likelihood isn't the only heuristic that could be used in the stochastic greedy search.
If we had some other heuristic, $\hat{\mbox{H}}$, one reasonable proposal for the MCMC walk would be the Boltzman distribution:
$$
h(\rho e | \rho) \propto \exp\left(-\hat{\mbox{H}}(\rho e)\right + \log P(e|G))
$$

\end{document}

### 2013-06-27 drafts/planner/planner_inference.tex
Updates to writeup and addition of clrs latex template

@@
 The $w_j$'s factor in a nice way that allows them to be computed in linear time using a dynamic program.
-The following procedure samples a plan and returns the weights upon each program in that plan:
+\footnote{The sums for $w_j$ end up being the partition function if $j=1$ and is a sort of ``truncated partition function'' for $j>1$.
+For the $j>1$ case, the weights $w_j$ are what the partition function \emph{would have been} if the plan had started with programs $e_{k<j}$ for free.}
+The following procedure samples a plan and returns the weights upon each program in that plan, along with the partition function:
 \begin{codebox}

### 2013-06-27 drafts/planner/planner_inference.tex
Updated writeup on planner

@@
 \begin{codebox}
-\Procname{$\proc{Sample-Plan}(\rho)$}
+\Procname{$\proc{MCMC-Plan}(\rho, \mbox{\id{reciprocalLikelihoods}})$}
 \li \Comment{Draw the next program in the plan}
@@
 \li \If $|\rho e|<\ell$ \Then
-\li \id{suffixWeight}, \id{suffixRewards} = \func{Sample-Plan}$(\rho e)$
+\li \id{suffixZ}, \id{suffixWeights} = \func{MCMC-Plan}$(\rho e, \mbox{\id{reciprocalLikelihoods}}/P(t|\rho e))$
 \li \Else
-\li \id{suffixWeight}, \id{suffixRewards} = $0$, $\emptyset$
+\li \id{suffixZ}, \id{suffixWeights} = $0$, $\emptyset$
 \li \End
-\li \Comment{To satisfy the boundary conditions of the recurrence, $P(t|\emptyset)$ is defined to be 1.}
-\li \id{eReward} = $\frac{1+\mbox{\id{suffixRewards}}}{P(t|\rho)}$
-\li \Return \id{eReward}, $\{(e, \mbox{\id{eReward}})\}\cup \mbox{\id{suffixRewards}}$
+\li \id{Z} = \id{suffixZ} + \id{reciprocalLikelihoods}
+\li \Return \id{Z}, $\{(e, \mbox{\id{Z}})\}\cup \mbox{\id{suffixWeights}}$
 \end{codebox}
+where \emph{reciprocalLikelihoods} is set to $1$ initially.
 
@@
 $$
-h(\rho e | \rho) \propto \exp\left(-\hat{\mbox{H}}(\rho e)\right + \log P(e|G))
+h(\rho e | \rho) \propto \exp\left(-\hat{\mbox{H}}(\rho e) + \log P(e|G)\right)
 $$

### 2013-08-13 drafts/enumeration/enumeration.tex
Added writeup describing memory-efficient enumeration

\documentclass{article}

\begin{document}

The EC algorithm requires some means of obtaining many unique programs that have high likelihood under a given grammar.
We could do this by sampling from the grammar, but doing so can often fail to reveal shared substructures until a very large number of programs have been sampled.
A breadth-first enumeration tends to reveal more common substructure, but has memory consumption that grows exponentially with the number of enumerated programs.
This document describes an algorithm for performing memory-efficient (approximate) enumeration.

The general idea is to use the provided grammar to build a binary prefix-free code for parse trees, satisfying
$$
- \log P(z|G) \approx \ell (z|G)
$$
where $z$ is the parse tree, $G$ is the grammar, and $\ell(\cdot|\cdot)$ is conditional description length.
Then, we enumerate binary strings in lexicographic order and attempt to decode these strings in to parse trees.
Upon successful decoding of a binary string, the corresponding program is added to the frontier.
So, the space complexity of this algorithm is very low: we need to maintain one binary string (whose length grows logarithmically with the size of the frontier) as well as the set of programs we've already enumerated.
This algorithm can be seen as a form of iterative deepening, and so should have the same time complexity as the breadth-first enumeration algorithm.

The encoding of a parse tree is as follows.
First, a Huffman encoding tree is constructed from all available productions at the given choice point; that is, all productions whose type unifies with the current requested type.
Then, the chosen production is encoded using this tree.
Finally, the encodings for all non-terminals in this production are sequentially appended.

This encoding is best illustrated using an example.
Say that our library has only one terminal, $I$, having type $\alpha\to\alpha$.
The requested type for a given program is $\alpha\to\alpha$; suppose that the chosen program is $(I\; I)$ and that the grammar has probability 1/2 of choosing either $I$ or an application.

There's only one possible parse of $(I I)$.
First, an application is chosen.
So, construct a Huffman encoding tree for the two productions $S\to S S$ and $S\to I$; say that the code word for $S\to SS$ is 0 and the code for $S\to I$ is 1.
Then, the first bit is a 0 because the first production used in the parse is that of an application.
The encoding of $I$ is a 1, so the next two bits are both 1's, giving an encoding of $011$ for $(I\; I)$.

\end{document}

### 2013-09-23 src/experiments/TowerPlan/run.sh
Added larger frontier test case to Tower. Added fixed-length vector types.

@@ -3,6 +3,7 @@
 
 # Make data directory
 mkdir data
+mkdir data2000
 
 # CD to src/ directory
 cd ../../
@@ -11,4 +12,5 @@ make clean
 make tower
 
 # Run script with default parameters
-./Strappy/Tower 0 p 0.015 0.1 500 60 5 'experiments/TowerPlan/data' | tee experiments/TowerPlan/data/log
+#./Strappy/Tower 0 p 0.015 0.1 500 60 5 'experiments/TowerPlan/data' | tee experiments/TowerPlan/data/log
+./Strappy/Tower 0 p 0.015 0.1 2000 60 5 'experiments/TowerPlan/data2000' | tee experiments/TowerPlan/data2000/log

### 2013-09-23 drafts/tower/tower.tex
Added tower writeup

\documentclass{article}
\usepackage{verbatim}
\begin{document}
\section{Overview}
This document summarizes experiments in tower-building via EC.
In these experiments, a tower is a list of tuples of real numbers and booleans.
The list corresponds to a sequence of block placements in a 2D world.
The real number component of the $i$th tuple is the $x$ coordinate of the $i$th block.
The boolean component of the $i$th tuple is the orientation of the $i$th block (horizontal or vertical).
The blocks are placed in the sequence specified by the list.

Whether the tower is stable when pushed with varying strengths is measured.
The ``log likelihood'' of a tower is an increasing function of its stability and its height.
So, the best towers are those that are tall, stable, and have small description length.

It's interesting to note that we don't have a curriculum of tasks to solve.
Instead, we hope that the ability of EC to handle partial rewards will serve as an adequate curriculum.

\section{Experiments}
EC with planning was used to solve the tower task. The parameters were: $\lambda=0.015$, pseudocounts=$0.1$, frontier size = 2000, plans per task = 60, plan length = 5.
These results may be reproduced using the run.sh script in the TowerPlan directory.
Decreasing the frontier size to 500 produced qualitatively similar results.

EC built multilevel, arch-like towers.
After one iteration of the E and C steps, the grammar did not change significantly.
The final grammar was:
\begin{verbatim}
      p:   0.21
  (I I):  -2.14
(map I):  -2.10
      I:  -2.09
(: ((pair -1.0) False)):  -1.80
(: ((pair ((*. -1.0) 1.0)) False)):  -1.78
(: (I ((pair 1.0) False))):  -1.75
(: (I ((pair 0.0) False))):  -1.72
(++ (single ((pair 1.0) False))):  -1.71
(++ (single ((pair -1.0) False))):  -1.67
(: ((pair -1.0) True)):  -1.56
(: ((pair 1.0) True)):  -1.47
(((C (C I)) :) ((pair 0.0) True)):  -1.42
    fst:  -0.85
    snd:  -0.82
     ++:  -0.68
      ::  -0.64
    map:  -0.64
  foldl:  -0.14
      C:  -0.13
      S:  -0.12
 single:  -0.08
      B:  -0.08
  False:  -0.06
   True:  -0.05
   pair:  -0.02
    rep:  -0.02
     +.:  -0.02
     -.:  -0.02
     *.:  -0.02
     /.:  -0.02
    0.0:  -0.02
    1.0:  -0.02
   -1.0:  -0.02
\end{verbatim}

\section{Analysis}
Observations about the above grammar:
\begin{itemize}
\item{(a)} The lowest-description length productions are all primitives. This could be due to at least three different effects:
\begin{itemize}
\item{(i)} The other combinators in the library are very shallow.
Thus, when estimating the grammar's parameters, a non-neglible fraction of the credit for any of the non-primitive combinators is distributed to the primitive combinators as well.
This is especially apparent in the \verb=(map I)= case; when \verb=(map I)= occurs in the corpus to be compressed, it is not significantly more compressive to parse it as a single production than as three (application, map, I).
\item{(ii)} Plans are stochastically sampled, which decreases the potential for sharing as compared to some sort of deterministic enumeration.
As a result, more complex combinators are penalized because they belong to a very large class of equivalent combinators.
\item{(iii)} The EM version of EC might be stuck in a local minimum. The most compressive grammar is, unintuitively, a function of the previous grammar; thus, if an early grammar (such as the first grammar) penalizes non-primitive combinators, EC will continue to weight those combinators too highly. (However, this didn't occur for polynomial regression. Why? Fortunately, it is easy to test empirically if this effect is present, as detailed in ``Future Experiments.'')
\end{itemize}
\item{(b)} Missed opportunities for compression. Many combinators in the library differ only in one or two leaves (eg, \verb=(: ((pair -1.0) True))= and \verb=(: ((pair 1.0) True)))=.
Although something like antiunification would resolve these specific instances, it would not solve the more general problem of abstracting out subtrees that perform the same computation.
If the more compressive combinators are never sampled/enumerated, then we can't solve this problem.
However, if they are, they still might not be put in to the grammar, for the same reasons given in observation a.iii.
\item{(c)} Nearly complete lack of higher-order planning primitives, excepting \verb=(map I)=. This is due to the small frontier size.
I tried having EC learn plan manipulation functions, such as \verb=reverse=, \verb=(map (\(x, o). (-x,o)))= (reflect over $x$ axis), etc, using the same initial grammar as the tower building program had.
Both with and without planning, EC failed to learn these primitives.
It is also possible that there is little pressure to learn higher-order primitives because the planning mechanism is able to find high-scoring towers without them.
However, disabling planning results in qualitatively similar towers and grammars.
\end{itemize}

\section{Possible Future Experiments}
 We can test if it is the stochastic nature of our search algorithm that discourages sharing of subtrees by implementing a deterministic planner.
 We don't want to do best-first enumeratation of plans - otherwise, we might as well just enumerate more programs.
Instead we might do something like beam search.
A determinstic planner should be tested both on Towers and Polynomial Regression.

Are we doing planning the right way? Perhaps we're introducing the wrong inductive bias by forcing our programs to be broken up in to segments of type $\alpha\to\alpha$, where $\alpha$  is the target type of the task. This factoring of our programs could be increasing the description length of the programs of interest.\footnote{Although, looking at reasonable higher-order plan operators for the towers task seems to suggest that this is not the case; if anything, we might be introducing the correct inductive bias for this domain.}
One way of testing this hypothesis is to experiment with means of planning that do not factor the program in any one way.
There are at least three different ways of incrementally building programs that do not put restrictions on the factoring of the program:
\begin{itemize}
\item{Monte Carl Tree Search:} Writing a program can be viewed as playing a game, where each move in the game correspondings to choosing the next production.
We can therefore adapt Monte Carlo Tree Search to the problem of incremental program writing, with the payout being either the likelihood or the (unnormed) posterior, depending on how strongly we want to regularize the search for programs.
This has been implemented but not tested.
\item{MCMC:} We could use MCMC; our proposal distribution would be something along the lines of, ``pick a subtree and redraw it from the grammar.'' I don't like this - I'm afraid it would take forever to mix.
\item{Genetic Algorithms:} We could use genetic programming techniques instead of the enumeration phase of EC. We'd probably want to do Evolutionary MCMC. I'm hesitant about this idea for the same reason I'm hesitant about MCMC.
\end{itemize}

As discussed in the previous section, in the EM version of EC, the next grammar is a function of the previous grammar, and not just the programs and their likelihoods.
This seems intuitively wrong. The original version of EC got this right.
The previous section explains why this might lead to the observed behavior on Towers.
If we've observed the programs in the set $\Omega$, then one reasonable way of picking the next grammar is to maximize the lower bound on the log posterior of the grammar:
\begin{eqnarray*}
\ln P(G|t) &=& \ln P(G) + \sum_n \ln \sum_e P(t_n|e)P(e|G)\\
&\geq& \ln P(G) + \sum_n \ln \sum_{e\in \Omega} P(t_n|e)P(e|G)
\end{eqnarray*}
One could approximately maximize this quantity by greedily adding productions to the grammar, or by an EM-like procedure.
If this was the cause of the lack of abstraction in the Tower grammar, then we would see bigger, but more compressive, combinators in the grammar.

\end{document}

### 2013-10-20 drafts/nips/nips2013.tex
Added NIPS template

\documentclass{article} % For LaTeX2e
\usepackage{nips13submit_e,times}
\usepackage{hyperref}
\usepackage{url}
\usepackage{verbatim}
%\documentstyle[nips13submit_09,times,art10]{article} % For LaTeX 2.09

\title{Learning Programs through Bayesian Inference and Planning}

\begin{comment}
\author{
David S.~Hippocampus\thanks{ Use footnote for providing further information
about author (webpage, alternative address)---\emph{not} for acknowledging
funding agencies.} \\
Department of Computer Science\\
Cranberry-Lemon University\\
Pittsburgh, PA 15213 \\
\texttt{hippo@cs.cranberry-lemon.edu} \\
\And
Coauthor \\
Affiliation \\
Address \\
\texttt{email} \\
\AND
Coauthor \\
Affiliation \\
Address \\
\texttt{email} \\
\And
Coauthor \\
Affiliation \\
Address \\
\texttt{email} \\
\And
Coauthor \\
Affiliation \\
Address \\
\texttt{email} \\
(if needed)\\
}
\end{comment}
% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to \LaTeX{} to determine where to break
% the lines. Using \AND forces a linebreak at that point. So, if \LaTeX{}
% puts 3 of 4 authors names on the first line, and the last on the second
% line, try using \AND instead of \And before the third author name.

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

%\nipsfinalcopy % Uncomment for camera-ready version

\begin{document}

\maketitle

\begin{abstract}
The abstract paragraph should be indented 1/2~inch (3~picas) on both left and
right-hand margins. Use 10~point type, with a vertical spacing of 11~points.
The word \textbf{Abstract} must be centered, bold, and in point size 12. Two
line spaces precede the abstract. The abstract must be limited to one
paragraph.
\end{abstract}

\section{Submission of papers to NIPS 2013}

NIPS requires electronic submissions.  The electronic submission site is  
\begin{center}
   \url{http://papers.nips.cc}
\end{center}

Please read carefully the
instructions below, and follow them faithfully.
\subsection{Style}

Papers to be submitted to NIPS 2013 must be prepared according to the
instructions presented here. Papers may be only up to eight pages long,
including figures. Since 2009 an additional ninth page \textit{containing only
cited references} is allowed. Papers that exceed nine pages will not be
reviewed, or in any other way considered for presentation at the conference.
%This is a strict upper bound.

Please note that this year we have introduced automatic line number generation
into the style file (for \LaTeXe and Word versions). This is to help reviewers
refer to specific lines of the paper when they make their comments. Please do
NOT refer to these line numbers in your paper as they will be removed from the
style file for the final version of accepted papers.

The margins in 2013 are the same as since 2007, which allow for $\approx 15\%$
more words in the paper compared to earlier years. We are also again using 
double-blind reviewing. Both of these require the use of new style files.

Authors are required to use the NIPS \LaTeX{} style files obtainable at the
NIPS website as indicated below. Please make sure you use the current files and
not previous versions. Tweaking the style files may be grounds for rejection.

%% \subsection{Double-blind reviewing}

%% This year we are doing double-blind reviewing: the reviewers will not know 
%% who the authors of the paper are. For submission, the NIPS style file will 
%% automatically anonymize the author list at the beginning of the paper.

%% Please write your paper in such a way to preserve anonymity. Refer to
%% previous work by the author(s) in the third person, rather than first
%% person. Do not provide Web links to supporting material at an identifiable
%% web site.

%%\subsection{Electronic submission}
%%
%% \textbf{THE SUBMISSION DEADLINE IS MAY 31st, 2013. SUBMISSIONS MUST BE LOGGED BY
%% 23:00, MAY 31st, 2013, UNIVERSAL TIME}

%% You must enter your submission in the electronic submission form available at
%% the NIPS website listed above. You will be asked to enter paper title, name of
%% all authors, keyword(s), and data about the contact
%% author (name, full address, telephone, fax, and email). You will need to
%% upload an electronic (postscript or pdf) version of your paper.

%% You can upload more than one version of your paper, until the
%% submission deadline. We strongly recommended uploading your paper in
%% advance of the deadline, so you can avoid last-minute server congestion.
%%
%% Note that your submission is only valid if you get an e-mail
%% confirmation from the server. If you do not get such an e-mail, please
%% try uploading again.

\subsection{Retrieval of style files}

The style files for NIPS and other conference information are available on the World Wide Web at
\begin{center}
   \url{http://www.nips.cc/}
\end{center}
The file \verb+nips2013.pdf+ contains these 
instructions and illustrates the
various formatting requirements your NIPS paper must satisfy. \LaTeX{}
users can choose between two style files:
\verb+nips11submit_09.sty+ (to be used with \LaTeX{} version 2.09) and
\verb+nips11submit_e.sty+ (to be used with \LaTeX{}2e). The file
\verb+nips2013.tex+ may be used as a ``shell'' for writing your paper. All you
have to do is replace the author, title, abstract, and text of the paper with
your own. The file
\verb+nips2013.rtf+ is provided as a shell for MS Word users.

The formatting instructions contained in these style files are summarized in
sections \ref{gen_inst}, \ref{headings}, and \ref{others} below.

%% \subsection{Keywords for paper submission}
%% Your NIPS paper can be submitted with any of the following keywords (more than one keyword is possible for each paper):

%% \begin{verbatim}
%% Bioinformatics
%% Biological Vision
%% Brain Imaging and Brain Computer Interfacing
%% Clustering
%% Cognitive Science
%% Control and Reinforcement Learning
%% Dimensionality Reduction and Manifolds
%% Feature Selection
%% Gaussian Processes
%% Graphical Models
%% Hardware Technologies
%% Kernels
%% Learning Theory
%% Machine Vision
%% Margins and Boosting
%% Neural Networks
%% Neuroscience
%% Other Algorithms and Architectures
%% Other Applications
%% Semi-supervised Learning
%% Speech and Signal Processing
%% Text and Language Applications

%% \end{verbatim}

\section{General formatting instructions}
\label{gen_inst}

The text must be confined within a rectangle 5.5~inches (33~picas) wide and
9~inches (54~picas) long. The left margin is 1.5~inch (9~picas).
Use 10~point type with a vertical spacing of 11~points. Times New Roman is the
preferred typeface throughout. Paragraphs are separated by 1/2~line space,
with no indentation.

Paper title is 17~point, initial caps/lower case, bold, centered between
2~horizontal rules. Top rule is 4~points thick and bottom rule is 1~point
thick. Allow 1/4~inch space above and below title to rules. All pages should
start at 1~inch (6~picas) from the top of the page.

%The version of the paper submitted for review should have ``Anonymous Author(s)'' as the author of the paper.

For the final version, authors' names are
set in boldface, and each name is centered above the corresponding
address. The lead author's name is to be listed first (left-most), and
the co-authors' names (if different address) are set to follow. If
there is only one co-author, list both author and co-author side by side.

Please pay special attention to the instructions in section \ref{others}
regarding figures, tables, acknowledgments, and references.

\section{Headings: first level}
\label{headings}

First level headings are lower case (except for first word and proper nouns),
flush left, bold and in point size 12. One line space before the first level
heading and 1/2~line space after the first level heading.

\subsection{Headings: second level}

Second level headings are lower case (except for first word and proper nouns),
flush left, bold and in point size 10. One line space before the second level
heading and 1/2~line space after the second level heading.

\subsubsection{Headings: third level}

Third level headings are lower case (except for first word and proper nouns),
flush left, bold and in point size 10. One line space before the third level
heading and 1/2~line space after the third level heading.

\section{Citations, figures, tables, references}
\label{others}

These instructions apply to everyone, regardless of the formatter being used.

\subsection{Citations within the text}

Citations within the text should be numbered consecutively. The corresponding
number is to appear enclosed in square brackets, such as [1] or [2]-[5]. The
corresponding references are to be listed in the same order at the end of the
paper, in the \textbf{References} section. (Note: the standard
\textsc{Bib\TeX} style \texttt{unsrt} produces this.) As to the format of the
references themselves, any style is acceptable as long as it is used
consistently.

As submission is double blind, refer to your own published work in the 
third person. That is, use ``In the previous work of Jones et al.\ [4]'',
not ``In our previous work [4]''. If you cite your other papers that
are not widely available (e.g.\ a journal paper under review), use
anonymous author names in the citation, e.g.\ an author of the
form ``A.\ Anonymous''.

\subsection{Footnotes}

Indicate footnotes with a number\footnote{Sample of the first footnote} in the
text. Place the footnotes at the bottom of the page on which they appear.
Precede the footnote with a horizontal rule of 2~inches
(12~picas).\footnote{Sample of the second footnote}

\subsection{Figures}

All artwork must be neat, clean, and legible. Lines should be dark
enough for purposes of reproduction; art work should not be
hand-drawn. The figure number and caption always appear after the
figure. Place one line space before the figure caption, and one line
space after the figure. The figure caption is lower case (except for
first word and proper nouns); figures are numbered consecutively.

Make sure the figure caption does not get separated from the figure.
Leave sufficient space to avoid splitting the figure and figure caption.

You may use color figures. 
However, it is best for the
figure captions and the paper body to make sense if the paper is printed
either in black/white or in color.
\begin{figure}[h]
\begin{center}
%\framebox[4.0in]{$\;$}
\fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
\end{center}
\caption{Sample figure caption.}
\end{figure}

\subsection{Tables}

All tables must be centered, neat, clean and legible. Do not use hand-drawn
tables. The table number and title always appear before the table. See
Table~\ref{sample-table}.

Place one line space before the table title, one line space after the table
title, and one line space after the table. The table title must be lower case
(except for first word and proper nouns); tables are numbered consecutively.

\begin{table}[t]
\caption{Sample table title}
\label{sample-table}
\begin{center}
\begin{tabular}{ll}
\multicolumn{1}{c}{\bf PART}  &\multicolumn{1}{c}{\bf DESCRIPTION}
\\ \hline \\
Dendrite         &Input terminal \\
Axon             &Output terminal \\
Soma             &Cell body (contains cell nucleus) \\
\end{tabular}
\end{center}
\end{table}

\section{Final instructions}
Do not change any aspects of the formatting parameters in the style files.
In particular, do not modify the width or length of the rectangle the text
should fit into, and do not change font sizes (except perhaps in the
\textbf{References} section; see below). Please note that pages should be
numbered.

\section{Preparing PostScript or PDF files}

Please prepare PostScript or PDF files with paper size ``US Letter'', and
not, for example, ``A4''. The -t
letter option on dvips will produce US Letter files.

Fonts were the main cause of problems in the past years. Your PDF file must
only contain Type 1 or Embedded TrueType fonts. Here are a few instructions
to achieve this.

\begin{itemize}

\item You can check which fonts a PDF files uses.  In Acrobat Reader,
select the menu Files$>$Document Properties$>$Fonts and select Show All Fonts. You can
also use the program \verb+pdffonts+ which comes with \verb+xpdf+ and is
available out-of-the-box on most Linux machines.

\item The IEEE has recommendations for generating PDF files whose fonts
are also acceptable for NIPS. Please see
\url{http://www.emfield.org/icuwb2010/downloads/IEEE-PDF-SpecV32.pdf}

\item LaTeX users:

\begin{itemize}

\item Consider directly generating PDF files using \verb+pdflatex+
(especially if you are a MiKTeX user). 
PDF figures must be substituted for EPS figures, however.

\item Otherwise, please generate your PostScript and PDF files with the following commands:
\begin{verbatim} 
dvips mypaper.dvi -t letter -Ppdf -G0 -o mypaper.ps
ps2pdf mypaper.ps mypaper.pdf
\end{verbatim}

Check that the PDF files only contains Type 1 fonts. 
%For the final version, please send us both the Postscript file and
%the PDF file.

\item xfig "patterned" shapes are implemented with 
bitmap fonts.  Use "solid" shapes instead. 
\item The \verb+\bbold+ package almost always uses bitmap
fonts.  You can try the equivalent AMS Fonts with command
\begin{verbatim}
\usepackage[psamsfonts]{amssymb}
\end{verbatim}
 or use the following workaround for reals, natural and complex: 
\begin{verbatim}
\newcommand{\RR}{I\!\!R} %real numbers
\newcommand{\Nat}{I\!\!N} %natural numbers 
\newcommand{\CC}{I\!\!\!\!C} %complex numbers
\end{verbatim}

\item Sometimes the problematic fonts are used in figures
included in LaTeX files. The ghostscript program \verb+eps2eps+ is the simplest
way to clean such figures. For black and white figures, slightly better
results can be achieved with program \verb+potrace+.
\end{itemize}
\item MSWord and Windows users (via PDF file):
\begin{itemize}
\item Install the Microsoft Save as PDF Office 2007 Add-in from
\url{http://www.microsoft.com/downloads/details.aspx?displaylang=en\&familyid=4d951911-3e7e-4ae6-b059-a2e79ed87041}
\item Select ``Save or Publish to PDF'' from the Office or File menu
\end{itemize}
\item MSWord and Mac OS X users (via PDF file):
\begin{itemize}
\item From the print menu, click the PDF drop-down box, and select ``Save
as PDF...''
\end{itemize}
\item MSWord and Windows users (via PS file):
\begin{itemize}
\item To create a new printer
on your computer, install the AdobePS printer driver and the Adobe Distiller PPD file from
\url{http://www.adobe.com/support/downloads/detail.jsp?ftpID=204} {\it Note:} You must reboot your PC after installing the
AdobePS driver for it to take effect.
\item To produce the ps file, select ``Print'' from the MS app, choose
the installed AdobePS printer, click on ``Properties'', click on ``Advanced.''
\item Set ``TrueType Font'' to be ``Download as Softfont''
\item Open the ``PostScript Options'' folder
\item Select ``PostScript Output Option'' to be ``Optimize for Portability''
\item Select ``TrueType Font Download Option'' to be ``Outline''
\item Select ``Send PostScript Error Handler'' to be ``No''
\item Click ``OK'' three times, print your file.
\item Now, use Adobe Acrobat Distiller or ps2pdf to create a PDF file from
the PS file. In Acrobat, check the option ``Embed all fonts'' if
applicable.
\end{itemize}

\end{itemize}
If your file contains Type 3 fonts or non embedded TrueType fonts, we will
ask you to fix it.

\subsection{Margins in LaTeX}

Most of the margin problems come from figures positioned by hand using
\verb+\special+ or other commands. We suggest using the command
\verb+\includegraphics+
from the graphicx package. Always specify the figure width as a multiple of
the line width as in the example below using .eps graphics
\begin{verbatim}
   \usepackage[dvips]{graphicx} ... 
   \includegraphics[width=0.8\linewidth]{myfile.eps} 
\end{verbatim}
or % Apr 2009 addition
\begin{verbatim}
   \usepackage[pdftex]{graphicx} ... 
   \includegraphics[width=0.8\linewidth]{myfile.pdf} 
\end{verbatim}
for .pdf graphics. 
See section 4.4 in the graphics bundle documentation (\url{http://www.ctan.org/tex-archive/macros/latex/required/graphics/grfguide.ps})

A number of width problems arise when LaTeX cannot properly hyphenate a
line. Please give LaTeX hyphenation hints using the \verb+\-+ command.

\subsubsection*{Acknowledgments}

Use unnumbered third level headings for the acknowledgments. All
acknowledgments go at the end of the paper. Do not include 
acknowledgments in the anonymized submission, only in the 
final paper.

\subsubsection*{References}

References follow the acknowledgments. Use unnumbered third level heading for
the references. Any choice of citation style is acceptable as long as you are
consistent. It is permissible to reduce the font size to `small' (9-point) 
when listing the references. {\bf Remember that this year you can use
a ninth page as long as it contains \emph{only} cited references.}

\small{
[1] Alexander, J.A. \& Mozer, M.C. (1995) Template-based algorithms
for connectionist rule extraction. In G. Tesauro, D. S. Touretzky
and T.K. Leen (eds.), {\it Advances in Neural Information Processing
Systems 7}, pp. 609-616. Cambridge, MA: MIT Press.

[2] Bower, J.M. \& Beeman, D. (1995) {\it The Book of GENESIS: Exploring
Realistic Neural Models with the GEneral NEural SImulation System.}
New York: TELOS/Springer-Verlag.

[3] Hasselmo, M.E., Schnell, E. \& Barkai, E. (1995) Dynamics of learning
and recall at excitatory recurrent synapses and cholinergic modulation
in rat hippocampal region CA3. {\it Journal of Neuroscience}
{\bf 15}(7):5249-5262.
}

\end{document}

### 2013-10-21 drafts/nips/nips2013.tex
Progress on NIPS draft

@@
 \end{abstract}
+
+\section{Introduction}
+
+Motivation: Many approaches to constructive machine learning use structured representations, such as graphs, grammars, Horn clauses, relations, ..., ???.
+All of these may be thought of as a program - programs are at least as expressive as these other representations.
+
+% TODO: Sequential EC? SEC? What is a good name for this?
+The space of programs is enormous and inherently unpredictable \cite{PACunpredictability}.
+We introduce a new algorithm, SEC, for searching the space of programs.
+Two main ideas underly SEC:
+\begin{itemize}
+\item \emph{Learning to search}: Programs admit compositionality and reuse of subprograms.
+SEC exploits this compositionality to learn a library of program fragments, which it then reuses when performing further searches for programs.
+In a multitask setting, this permits bootstrap learning of more complicated programs after the synthesis of simpler programs.
+\item \emph{Incremental Program Synthesis}: In many search problems, there is a natural means of assigning partial credit for partially-correct solutions.
+SEC uses this partial credit to select promising programs and modify them, creating complicated programs step-by-step.
+Rather than using a mutation operator, as in genetic programming, or Metropolis-Hastings proposals, as in \cite{MHprogramInduction}, SEC uses function composition to combine smaller programs in to larger ones.
+\end{itemize}
+
+The SEC algorithm comes from a generative model over compositions of functions drawn from a stochastic grammar over programs.
+We describe this model, as well as how we perform inference, in ...
+
+SEC treats planning as an inference problem.
+
+\section{The SEC Algorithm}
+\subsection{Generative Model}
+\subsection{Inference}
+\subsection{Program Representation}
+
+\section{Application to Structure Learning}
+
+\subsection{Graph Combinators}
+
+\subsection{Learning Graphical Models}
+
+\section{Building Towers in Blocks World}
+
+\subsection{The Blocks World Domain}
+
+\subsection{Learned Tower programs}
+
+\section{Discussion}
+
+\subsection{Relation to Prior Work}
+
+\subsection{Limitations of SEC}
 

### 2013-10-21 src/Strappy/Library.hs

@@ -322,6 +322,7 @@ cSingle = mkTerm "single" (t ->- tList t) $
 cRep = mkTerm "rep" (tInt ->- t ->- tList t) $
        replicate
 cConcat = mkTerm "concat" (tList (tList t) ->- tList t) concat
+cReverse = mkTerm "reverse" (tList t ->- tList t) reverse
 cFoldl = mkTerm "foldl" ((t ->- t1 ->- t) ->- t ->- tList t1 ->- t) $ 
          List.foldl'
 cFoldr = mkTerm "foldr" ((t1 ->- t2 ->- t2) ->- t2 ->- tList t1 ->- t2) $ 
@@ -380,7 +381,8 @@ towerExprs :: [Expr]
 towerExprs = [cI, 
               cS, 
               cB, 
-              cC, 
+              cC,
+              cK,
               cFPlus,
               cFMinus,
               cFTimes,
@@ -388,7 +390,7 @@ towerExprs = [cI,
               cCons,
               cAppend,
               cMap,
-              cFoldl,
+              cReverse,
               cSingle,
               cOnFst, cOnSnd,
               cNand,

### 2013-10-22 drafts/nips/nips2013.tex
Outline of NIPS paper, and fixed really bad bug in beam search that was causing it to almost never learn new productions

@@
 
-\title{Learning Programs through Bayesian Inference and Planning}
+\title{Incremental Construction of Programs in a Multitask Setting}
 
@@
 
-Motivation: Many approaches to constructive machine learning use structured representations, such as graphs, grammars, Horn clauses, relations, ..., ???.
+Motivation: Many approaches to constructive machine learning use structured representations, such as graphs, grammars, Horn clauses.
 All of these may be thought of as a program - programs are at least as expressive as these other representations.
 
-% TODO: Sequential EC? SEC? What is a good name for this?
 The space of programs is enormous and inherently unpredictable \cite{PACunpredictability}.
-We introduce a new algorithm, SEC, for searching the space of programs.
+We introduce a new algorithm, SEC, (TODO: Sequential EC? SEC? What is a good name for this?) for searching the space of programs.
 Two main ideas underly SEC:
@@
 
-The SEC algorithm comes from a generative model over compositions of functions drawn from a stochastic grammar over programs.
-We describe this model, as well as how we perform inference, in ...
+Applications: graphical model structure, towers
 
-SEC treats planning as an inference problem.
+The SEC algorithm is motivated by a generative model over compositions of functions drawn from a stochastic grammar over programs.
+
+SEC treats inference as a planning problem
 
@@
 \subsection{Generative Model}
+Write down the generative model. Connect to EC and Percy Liang's work. Discuss ``programs as plans.''
 \subsection{Inference}
+Write down EM updates. Interpret them as a form of compression solved via a weighted Neville-Manning variant.
+
+Discuss bootstrap learning: the right curriculum allows us to build more complicated programs.
+Make some analogy involving how humans learn to paint, write poetry, bake, compose music, play with blocks, or design statistical models.
+
 \subsection{Program Representation}
+Discuss combinatory logic and the adaptor-grammar-esque distribution we use.
 
 \section{Application to Structure Learning}
+Learning the structure of graphical models is a difficult problem - not only for machines, but also for humans, where expert knowledge is often required.
+In practice, the sorts of graphical models that humans design, such as Hidden Markov Models, phylogenetic trees, topic models, and Ising models, exhibit certain symmetries and recursive structures that are amenable to synthesis by programs.
+Could an algorithm, such as SEC, automatically learn these structures?
 
@@
 
+Diagrams illustrating the graph combinators. An example program that makes an Ising model.
+
 \subsection{Learning Graphical Models}
+
+Show the actual results for our test cases: ``density estimation of the space of interesting graphical models.''
+In future work, could be used as high-level moves in the search for a good structure.
+Connect to, eg, Charles Kemp's work.
 
@@
 
+How could an agent, endowed with little prior knowledge, learn to build tall, robust towers out of toy blocks?
+Although the space of possible towers is large enough to make exhaustive search intractible, the useful towers lie within a much smaller space.
+Most useful towers use compositions of stereotyped structures, such as arches.
+
 \subsection{The Blocks World Domain}
 
+Tower representation, objective function
+
 \subsection{Learned Tower programs}
+
+Picture of tower. Grammar that produced that tower. Note lack of abstraction in grammar; speculate as to reasons.
 
@@
 
+MH sampling of programs, ILP, Kitzelman, graph grammars
+
 \subsection{Limitations of SEC}
+
+Does not look at data. Does everything top-down. Blind to the utility of trying one program versus another (no active learning)
 

### 2013-10-22 src/Strappy/BeamSearch.hs

@@ -50,7 +50,7 @@ beamPlan pt beamSize seedDist dist planLen = do
                     foldl (\beam (ll, (e, pLL)) -> insertIntoBeam beamSize beam (ll+pLL, pLL) ([e],pLL)) MinQueue.empty seedRewards
     let seedRewards' = filter (not . isInvalidNum . fst) seedRewards
     let seedRewards'' = Map.fromList $ map (\ (ll, (e, pLL)) -> (e, ll+pLL)) seedRewards'
-    let seedLogZ = logSumExpList $ map (snd . snd) seedRewards
+    let seedLogZ = logSumExpList $ map (\ (ll, (_, pLL)) -> ll + pLL) seedRewards
     let seedHits = Set.fromList [([e], pLL) | (ll, (e, pLL)) <- seedRewards, ll > -0.1 ]
     (logRewards, logZ, hits) <- bs seedPlans planLen seedRewards'' seedLogZ seedHits
     return (Map.map (\x -> x-logZ) logRewards, hits)
@@ -77,7 +77,7 @@ beamPlan pt beamSize seedDist dist planLen = do
                     concatMap (\(logReward, (p, _)) -> map (,logReward) p) scoredPlans
             -- Calculate new beam
             let newBeam = map snd $ MinQueue.toList $
-                          foldl (\beam (planScore, plan@(p, pLL)) -> insertIntoBeam beamSize beam (planScore, pLL) plan) MinQueue.empty scoredPlans
+                          foldl (\beam (planScore, plan@(_, pLL)) -> insertIntoBeam beamSize beam (planScore, pLL) plan) MinQueue.empty scoredPlans
             -- Merge old with new
             let newLogZ = logSumExp myLogZ oldLogZ
             let newRewards = Map.unionWith logSumExp myRewards oldRewards

### 2013-10-22 src/Strappy/UGM.hs

@@ -52,7 +52,12 @@ mkUEdge v1 v2 | v1 < v2 = UEdge v1 v2
 mkUEdge v1 v2 = UEdge v2 v1
 
 data UGM = UGM [Vertex] [UEdge]
-    deriving(Eq, Ord, Show)
+    deriving(Show)
+
+instance Eq UGM where
+  (UGM v e) == (UGM v' e') = sort v == sort v' && sort e == sort e'
+instance Ord UGM where
+  compare (UGM v e) (UGM v' e') = compare (sort v, sort e) (sort v', sort e')
 
 nullUGM :: UGM
 nullUGM = UGM [] []
@@ -111,6 +116,8 @@ ugmRename2Disjoint u1@(UGM vs1 es1) u2@(UGM vs2 es2) =
     in
         (u1, buildSub (map getVertexID $ filter isLatent vs2) u2)
 
+ugmExplode :: UGM -> [UGM]
+ugmExplode (UGM vs _) = map (\v -> UGM [v] []) (sort vs)
 
 -- | Test cases
 ugmLatentify :: UGM -> UGM
@@ -132,6 +139,12 @@ ugmIsing vss = foldr1 ugmUnionCorrespondence (map ugmChain vss)
 ugmHMM :: [UGM] -> UGM
 ugmHMM vs = ugmUnionCorrespondence (foldl1 ugmUnion vs) (ugmChain $ map ugmLatentify vs)
 
+ugmrHMM :: [UGM] -> UGM
+ugmrHMM vs = ugmUnionCorrespondence (ugmChain vs) (foldl1 ugmUnion $ map ugmLatentify vs)
+
+ugmCylinder :: [UGM] -> UGM
+ugmCylinder vs = ugmUnionCorrespondence (ugmRing vs) (ugmRing $ map ugmLatentify vs)
+
 ugmMRF :: [[UGM]] -> UGM
 ugmMRF vss = ugmUnionCorrespondence (foldr1 ugmUnion $ concat vss) $
              ugmIsing $ map (map ugmLatentify) vss
@@ -139,15 +152,14 @@ ugmMRF vss = ugmUnionCorrespondence (foldr1 ugmUnion $ concat vss) $
 ugmLatentCommon :: [UGM] -> UGM
 ugmLatentCommon = foldl ugmUnionHeadTail (singletonUGM $ Latent 1)
 
+-- Just for fun: the graphical model behind EC
+ugmEC :: [UGM] -> UGM
[…]

### 2013-10-22 drafts/nips/nips2013.tex
Wrote first draft of introduction and abstract.

--- 2013-10-21 drafts/nips/nips2013.tex
@@
 \usepackage{verbatim}
+\usepackage{graphicx}
+\usepackage{subcaption}
+\usepackage{listings}
 %\documentstyle[nips13submit_09,times,art10]{article} % For LaTeX 2.09
@@
 \begin{abstract}
-The abstract paragraph should be indented 1/2~inch (3~picas) on both left and
-right-hand margins. Use 10~point type, with a vertical spacing of 11~points.
-The word \textbf{Abstract} must be centered, bold, and in point size 12. Two
-line spaces precede the abstract. The abstract must be limited to one
-paragraph.
+How can machine learning techniques be used to solve problems whose solutions are best represented as computer programs?
+For example, suppose a researcher wants to design a probabilistic graphical model for a novel domain.
+On the one hand, designing a good model requires a great deal of domain-specific knowledge and expertise, which may be hard to come by.
+On the other hand, searching the space of probabilistic models automatically is notoriously difficult.
+However, when researchers do set out to design graphical models, they draw on abstractions such as trees, chains, grids and plates to constrain and direct the kinds of models they produce.
+Thus, if we want to produce machine learning tools that aid in this task, we want tools that can learn these kinds of relevant abstractions.
+One natural way to think of these abstractions is as programs that take sets of random variables and produce graphical models that relate them.
+In this work, we describe the SEC algorithm, which attempts to learn distributions over such programs by incrementally finding program components that commonly help solve problems in a given domain.
+We discuss the behavior of the SEC algorithm for the task of graphical model structure learning.
+We also describe SEC's behavior on the dissimilar problem of automatically constructing tall and physically stable towers of blocks.
 \end{abstract}
@@
 \section{Introduction}
+
+The vast majority of research in machine learning is focused on utilizing large amounts of noisy data living in high-dimensional, real-valued vector spaces.
+Such research has been immensely successful, and this has enabled automatic and accurate modelling on a scale that was previously infeasible.
+However, when we want to build a car, draw a painting, or design the machine learning models to which we apply our new algorithms and techniques, we face problems that current machine learning tools are unable to handle.
+The highly structured objects that humans produce seems to require a different representation than the ones which machine learning algorithms have most successfully employed.
+
+To address this problem, many structured representations have been proposed in the machine learning and AI literatures: learning over graphs \cite{}, grammars\cite{}, Horn clauses\cite{}, relations{}, etc., has been investigated.
+These are representations that attempt to bridge the gap between the tractablity of learning and reasoning in smooth continuous spaces and the expressivity of general purpose programming languages.
+
+However, many of the most natural representations of structured domains are best described as programs, are difficult to fit into more restrictive representations.
+To take an example directly from the machine learning literature, consider the plate notation~\cite{DBLP:journals/jair/Buntine94} commonly used to describe graphical models over collections of variables.
+A typical example is shown in Figure~\ref{fig:lda_plate}~\cite{DBLP:journals/jmlr/BleiNJ03}.
+The goal of plate notation is precisely to express the notion of a ``for loop'' within the graphical model representation.
+The reason this is a convenient tool is that researchs onceptualize their models as generative programs, not graphs; see the equivalent of the LDA graphical model in the pseudo-code in Figure~\ref{fig:lda_code}.
+In fact, the field of probabilistic programming, which has gained recent popularity, is spurred on by exactly the motivation that our models are most naturally expressed as programs.
+
+In this paper, we describe our approach to learning over structured representations as a search over the space of programs in a functional programming language.
+We characterize learning in this space as incrementally developing a weighted library of program abstractions, i.e.
+a collection of subroutines, that are commonly used to solve problems in a given domain.
+The SEC algorithm (for Sequential Exploration-Compression), which we present here, takes a set of related tasks and produces a distribution over programs that solve those tasks; as it refines this distribution, SEC is able to search the space of programs more efficiently and solve more tasks.
+
+We will present here two domains of interest.
+First, continuing with the example above, we will describe how SEC can learn to efficiently explore the space of relevant graphical models for a given problem.
+Second, we will examine SEC's performance on a completely different type of task, in which we want to an algorithm to build tall and physically stable block towers.
+
+\lstset{
+	basicstyle=\scriptsize\ttfamily,
+    commentstyle=\color{white!40!black},
+    escapechar=+,
+    escapebegin=\color{white!40!black},
+    escapeend={},
+    keywordstyle=\bfseries\color{Periwinkle},
+    numbers=left,
+    numberstyle=\sffamily\tiny\color{gray},
+}
+
+\begin{figure}[h]
+\centering
+\begin{subfigure}{.5\textwidth}
+  \centering
+\includegraphics[width=.7\linewidth]{./figures/lda_plate_blei.pdf}
+  \caption{An example of a typical plate notation use. \label{fig:lda_plate}}
+
+\end{subfigure}%
+\begin{subfigure}{.4\textwidth}
+	\begin{lstlisting}[frame=single, numbers=none, xleftmargin=0pt]
+	for +$k$+ = 1:NumTopics
+	  +$\beta$+[k] +$\sim$+ Dirichlet(+$\eta$+)
+	for m = 1:NumDocuments
+	  +$\theta$+[m] +$\sim$+ Dirichlet(+$\alpha$+) 
+	  for n = 1:NumWords 
+	    z +$\sim$+ Categorical(+$\theta$+[m])
+	   w +$\sim$+ Categorical(+$\beta$+[z])
+	\end{lstlisting}
+  \caption{The same model as a program.}
+  \label{fig:lda_code}
+\end{subfigure}
+\caption{The popular Latent Dirichlet Allocation model in two forms.}
+\label{fig:lda}
+\end{figure}
+
+% \begin{figure}[h]
+% \begin{center}
+
+% \end{center}
+% \caption{A typical example of a graphical model in plate notation. }
+% \end{figure}
+
+% \begin{figure}[h]
+% \begin{verbatim}
+
+% for k = 1:NumTopics
+% 	beta[k]  ~ Dirichlet(eta)
+% for m = 1:NumDocuments
+% 	theta[m] ~ Dirichlet(alpha) 
+% 	for n = 1:NumWords
+% 		z ~ Categorical(theta[m])
+% 		w[n, m] ~ Categorical(beta[z])
+
+% \end{verbatim}
+% \end{figure}
+
+Here, we present an approach to learning
 
@@
 
-\subsection{Tables}
+% \subsection{Tables}
 
-All tables must be centered, neat, clean and legible. Do not use hand-drawn
-tables. The table number and title always appear before the table. See
-Table~\ref{sample-table}.
+% All tables must be centered, neat, clean and legible. Do not use hand-drawn
+% tables. The table number and title always appear before the table. See
+% Table~\ref{sample-table}.
 
-Place one line space before the table title, one line space after the table
-title, and one line space after the table. The table title must be lower case
-(except for first word and proper nouns); tables are numbered consecutively.
+% Place one line space before the table title, one line space after the table
+% title, and one line space after the table. The table title must be lower case
+% (except for first word and proper nouns); tables are numbered consecutively.
 
-\begin{table}[t]
-\caption{Sample table title}
-\label{sample-table}
-\begin{center}
-\begin{tabular}{ll}
-\multicolumn{1}{c}{\bf PART}  &\multicolumn{1}{c}{\bf DESCRIPTION}
-\\ \hline \\
-Dendrite         &Input terminal \\
-Axon             &Output terminal \\
-Soma             &Cell body (contains cell nucleus) \\
-\end{tabular}
-\end{center}
-\end{table}
+% \begin{table}[t]
+% \caption{Sample table title}
+% \label{sample-table}
+% \begin{center}
+% \begin{tabular}{ll}
+% \multicolumn{1}{c}{\bf PART}  &\multicolumn{1}{c}{\bf DESCRIPTION}
+% \\ \hline \\
+% Dendrite         &Input terminal \\
+% Axon             &Output terminal \\
+% Soma             &Cell body (contains cell nucleus) \\
+% \end{tabular}
+% \end{center}
+% \end{table}
 
-\section{Final instructions}
-Do not change any aspects of the formatting parameters in the style files.
-In particular, do not modify the width or length of the rectangle the text
-should fit into, and do not change font sizes (except perhaps in the
-\textbf{References} section; see below). Please note that pages should be
-numbered.
+% \section{Final instructions}
+% Do not change any aspects of the formatting parameters in the style files.
+% In particular, do not modify the width or length of the rectangle the text
+% should fit into, and do not change font sizes (except perhaps in the
+% \textbf{References} section; see below). Please note that pages should be
+% numbered.
 
-\section{Preparing PostScript or PDF files}
+% \section{Preparing PostScript or PDF files}
 
-Please prepare PostScript or PDF files with paper size ``US Letter'', and
-not, for example, ``A4''. The -t
-letter option on dvips will produce US Letter files.
+% Please prepare PostScript or PDF files with paper size ``US Letter'', and
+% not, for example, ``A4''. The -t
+% letter option on dvips will produce US Letter files.
 
-Fonts were the main cause of problems in the past years. Your PDF file must
-only contain Type 1 or Embedded TrueType fonts. Here are a few instructions
-to achieve this.
+% Fonts were the main cause of problems in the past years. Your PDF file must
+% only contain Type 1 or Embedded TrueType fonts. Here are a few instructions
+% to achieve this.
 
-\begin{itemize}
+% \begin{itemize}
 
-\item You can check which fonts a PDF files uses.  In Acrobat Reader,
-select the menu Files$>$Document Properties$>$Fonts and select Show All Fonts. You can
-also use the program \verb+pdffonts+ which comes with \verb+xpdf+ and is
-available out-of-the-box on most Linux machines.
+% \item You can check which fonts a PDF files uses.  In Acrobat Reader,
+% select the menu Files$>$Document Properties$>$Fonts and select Show All Fonts. You can
+% also use the program \verb+pdffonts+ which comes with \verb+xpdf+ and is
+% available out-of-the-box on most Linux machines.
 
-\item The IEEE has recommendations for generating PDF files whose fonts
-are also acceptable for NIPS. Please see
-\url{http://www.emfield.org/icuwb2010/downloads/IEEE-PDF-SpecV32.pdf}
+% \item The IEEE has recommendations for generating PDF files whose fonts
+% are also acceptable for NIPS. Please see
+% \url{http://www.emfield.org/icuwb2010/downloads/IEEE-PDF-SpecV32.pdf}
 
-\item LaTeX users:
+% \item LaTeX users:
 
-\begin{itemize}
+% \begin{itemize}
 
-\item Consider directly generating PDF files using \verb+pdflatex+
-(especially if you are a MiKTeX user). 
-PDF figures must be substituted for EPS figures, however.
+% \item Consider directly generating PDF files using \verb+pdflatex+
+% (especially if you are a MiKTeX user). 
+% PDF figures must be substituted for EPS figures, however.
 
-\item Otherwise, please generate your PostScript and PDF files with the following commands:
-\begin{verbatim} 
-dvips mypaper.dvi -t letter -Ppdf -G0 -o mypaper.ps
-ps2pdf mypaper.ps mypaper.pdf
-\end{verbatim}
+% \item Otherwise, please generate your PostScript and PDF files with the following commands:
+% \begin{verbatim} 
+% dvips mypaper.dvi -t letter -Ppdf -G0 -o mypaper.ps
+% ps2pdf mypaper.ps mypaper.pdf
+% \end{verbatim}
 
-Check that the PDF files only contains Type 1 fonts. 
-%For the final version, please send us both the Postscript file and
-%the PDF file.
+% Check that the PDF files only contains Type 1 fonts. 
+% %For the final version, please send us both the Postscript file and
+% %the PDF file.
 
-\item xfig "patterned" shapes are implemented with 
-bitmap fonts.  Use "solid" shapes instead. 
-\item The \verb+\bbold+ package almost always uses bitmap
-fonts.  You can try the equivalent AMS Fonts with command
-\begin{verbatim}
-\usepackage[psamsfonts]{amssymb}
-\end{verbatim}
- or use the following workaround for reals, natural and complex: 
-\begin{verbatim}
-\newcommand{\RR}{I\!\!R} %real numbers
-\newcommand{\Nat}{I\!\!N} %natural numbers 
-\newcommand{\CC}{I\!\!\!\!C} %complex numbers
-\end{verbatim}
+% \item xfig "patterned" shapes are implemented with 
+% bitmap fonts.  Use "solid" shapes instead. 
+% \item The \verb+\bbold+ package almost always uses bitmap
+% fonts.  You can try the equivalent AMS Fonts with command
+% \begin{verbatim}
+% \usepackage[psamsfonts]{amssymb}
+% \end{verbatim}
+%  or use the following workaround for reals, natural and complex: 
+% \begin{verbatim}
+% \newcommand{\RR}{I\!\!R} %real numbers
+% \newcommand{\Nat}{I\!\!N} %natural numbers 
+% \newcommand{\CC}{I\!\!\!\!C} %complex numbers
+% \end{verbatim}
 
-\item Sometimes the problematic fonts are used in figures
-included in LaTeX files. The ghostscript program \verb+eps2eps+ is the simplest
-way to clean such figures. For black and white figures, slightly better
-results can be achieved with program \verb+potrace+.
-\end{itemize}
-\item MSWord and Windows users (via PDF file):
-\begin{itemize}
-\item Install the Microsoft Save as PDF Office 2007 Add-in from
-\url{http://www.microsoft.com/downloads/details.aspx?displaylang=en\&familyid=4d951911-3e7e-4ae6-b059-a2e79ed87041}
-\item Select ``Save or Publish to PDF'' from the Office or File menu
-\end{itemize}
-\item MSWord and Mac OS X users (via PDF file):
-\begin{itemize}
-\item From the print menu, click the PDF drop-down box, and select ``Save
-as PDF...''
-\end{itemize}
-\item MSWord and Windows users (via PS file):
-\begin{itemize}
-\item To create a new printer
-on your computer, install the AdobePS printer driver and the Adobe Distiller PPD file from
-\url{http://www.adobe.com/support/downloads/detail.jsp?ftpID=204} {\it Note:} You must reboot your PC after installing the
-AdobePS driver for it to take effect.
-\item To produce the ps file, select ``Print'' from the MS app, choose
-the installed AdobePS printer, click on ``Properties'', click on ``Advanced.''
-\item Set ``TrueType Font'' to be ``Download as Softfont''
-\item Open the ``PostScript Options'' folder
-\item Select ``PostScript Output Option'' to be ``Optimize for Portability''
-\item Select ``TrueType Font Download Option'' to be ``Outline''
-\item Select ``Send PostScript Error Handler'' to be ``No''
-\item Click ``OK'' three times, print your file.
-\item Now, use Adobe Acrobat Distiller or ps2pdf to create a PDF file from
-the PS file. In Acrobat, check the option ``Embed all fonts'' if
-applicable.
-\end{itemize}
+% \item Sometimes the problematic fonts are used in figures
+% included in LaTeX files. The ghostscript program \verb+eps2eps+ is the simplest
+% way to clean such figures. For black and white figures, slightly better
+% results can be achieved with program \verb+potrace+.
+% \end{itemize}
+% \item MSWord and Windows users (via PDF file):
+% \begin{itemize}
+% \item Install the Microsoft Save as PDF Office 2007 Add-in from
+% \url{http://www.microsoft.com/downloads/details.aspx?displaylang=en\&familyid=4d951911-3e7e-4ae6-b059-a2e79ed87041}
+% \item Select ``Save or Publish to PDF'' from the Office or File menu
+% \end{itemize}
+% \item MSWord and Mac OS X users (via PDF file):
+% \begin{itemize}
+% \item From the print menu, click the PDF drop-down box, and select ``Save
+% as PDF...''
+% \end{itemize}
+% \item MSWord and Windows users (via PS file):
+% \begin{itemize}
+% \item To create a new printer
+% on your computer, install the AdobePS printer driver and the Adobe Distiller PPD file from
+% \url{http://www.adobe.com/support/downloads/detail.jsp?ftpID=204} {\it Note:} You must reboot your PC after installing the
+% AdobePS driver for it to take effect.
+% \item To produce the ps file, select ``Print'' from the MS app, choose
+% the installed AdobePS printer, click on ``Properties'', click on ``Advanced.''
+% \item Set ``TrueType Font'' to be ``Download as Softfont''
+% \item Open the ``PostScript Options'' folder
+% \item Select ``PostScript Output Option'' to be ``Optimize for Portability''
+% \item Select ``TrueType Font Download Option'' to be ``Outline''
+% \item Select ``Send PostScript Error Handler'' to be ``No''
+% \item Click ``OK'' three times, print your file.
+% \item Now, use Adobe Acrobat Distiller or ps2pdf to create a PDF file from
+% the PS file. In Acrobat, check the option ``Embed all fonts'' if
+% applicable.
+% \end{itemize}
 
-\end{itemize}
-If your file contains Type 3 fonts or non embedded TrueType fonts, we will
-ask you to fix it.
+% \end{itemize}
+% If your file contains Type 3 fonts or non embedded TrueType fonts, we will
+% ask you to fix it.
 
-\subsection{Margins in LaTeX}
+% \subsection{Margins in LaTeX}
 
-Most of the margin problems come from figures positioned by hand using
-\verb+\special+ or other commands. We suggest using the command
-\verb+\includegraphics+
-from the graphicx package. Always specify the figure width as a multiple of
-the line width as in the example below using .eps graphics
-\begin{verbatim}
-   \usepackage[dvips]{graphicx} ... 
-   \includegraphics[width=0.8\linewidth]{myfile.eps} 
-\end{verbatim}
-or % Apr 2009 addition
-\begin{verbatim}
-   \usepackage[pdftex]{graphicx} ... 
-   \includegraphics[width=0.8\linewidth]{myfile.pdf} 
-\end{verbatim}
-for .pdf graphics. 
-See section 4.4 in the graphics bundle documentation (\url{http://www.ctan.org/tex-archive/macros/latex/required/graphics/grfguide.ps})
+% Most of the margin problems come from figures positioned by hand using
+% \verb+\special+ or other commands. We suggest using the command
+% \verb+\includegraphics+
+% from the graphicx package. Always specify the figure width as a multiple of
+% the line width as in the example below using .eps graphics
+% \begin{verbatim}
+%    \usepackage[dvips]{graphicx} ... 
+%    \includegraphics[width=0.8\linewidth]{myfile.eps} 
+% \end{verbatim}
+% or % Apr 2009 addition
+% \begin{verbatim}
+%    \usepackage[pdftex]{graphicx} ... 
+%    \includegraphics[width=0.8\linewidth]{myfile.pdf} 
+% \end{verbatim}
+% for .pdf graphics. 
+% See section 4.4 in the graphics bundle documentation (\url{http://www.ctan.org/tex-archive/macros/latex/required/graphics/grfguide.ps})
 
-A number of width problems arise when LaTeX cannot properly hyphenate a
-line. Please give LaTeX hyphenation hints using the \verb+\-+ command.
+% A number of width problems arise when LaTeX cannot properly hyphenate a
+% line. Please give LaTeX hyphenation hints using the \verb+\-+ command.
 
-\subsubsection*{Acknowledgments}
+% \subsubsection*{Acknowledgments}
 
-Use unnumbered third level headings for the acknowledgments. All
-acknowledgments go at the end of the paper. Do not include 
-acknowledgments in the anonymized submission, only in the 
-final paper.
+% Use unnumbered third level headings for the acknowledgments. All
+% acknowledgments go at the end of the paper. Do not include 
+% acknowledgments in the anonymized submission, only in the 
+% final paper.
 
-\subsubsection*{References}
+% \subsubsection*{References}
 
-References follow the acknowledgments. Use unnumbered third level heading for
-the references. Any choice of citation style is acceptable as long as you are
-consistent. It is permissible to reduce the font size to `small' (9-point) 
-when listing the references. {\bf Remember that this year you can use
-a ninth page as long as it contains \emph{only} cited references.}
+% References follow the acknowledgments. Use unnumbered third level heading for
+% the references. Any choice of citation style is acceptable as long as you are
+% consistent. It is permissible to reduce the font size to `small' (9-point) 
+% when listing the references. {\bf Remember that this year you can use
+% a ninth page as long as it contains \emph{only} cited references.}
 
-\small{
-[1] Alexander, J.A. \& Mozer, M.C. (1995) Template-based algorithms
-for connectionist rule extraction. In G. Tesauro, D. S. Touretzky
-and T.K. Leen (eds.), {\it Advances in Neural Information Processing
-Systems 7}, pp. 609-616. Cambridge, MA: MIT Press.
+% \small{
+% [1] Alexander, J.A. \& Mozer, M.C. (1995) Template-based algorithms
+% for connectionist rule extraction. In G. Tesauro, D. S. Touretzky
+% and T.K. Leen (eds.), {\it Advances in Neural Information Processing
+% Systems 7}, pp. 609-616. Cambridge, MA: MIT Press.
 
-[2] Bower, J.M. \& Beeman, D. (1995) {\it The Book of GENESIS: Exploring
-Realistic Neural Models with the GEneral NEural SImulation System.}
-New York: TELOS/Springer-Verlag.
+% [2] Bower, J.M. \& Beeman, D. (1995) {\it The Book of GENESIS: Exploring
+% Realistic Neural Models with the GEneral NEural SImulation System.}
+% New York: TELOS/Springer-Verlag.
 
-[3] Hasselmo, M.E., Schnell, E. \& Barkai, E. (1995) Dynamics of learning
-and recall at excitatory recurrent synapses and cholinergic modulation
-in rat hippocampal region CA3. {\it Journal of Neuroscience}
-{\bf 15}(7):5249-5262.
-}
+% [3] Hasselmo, M.E., Schnell, E. \& Barkai, E. (1995) Dynamics of learning
+% and recall at excitatory recurrent synapses and cholinergic modulation
+% in rat hippocampal region CA3. {\it Journal of Neuroscience}
+% {\bf 15}(7):5249-5262.
+% }
+
+\bibliographystyle{plain}
+\bibliography{nips2013}
 

### 2013-10-22 drafts/nips/nips2013.tex
Edits to intro.

@@
 The vast majority of research in machine learning is focused on utilizing large amounts of noisy data living in high-dimensional, real-valued vector spaces.
-Such research has been immensely successful, and this has enabled automatic and accurate modelling on a scale that was previously infeasible.
+Such research has been immensely successful and has enabled automatic and accurate modelling on a scale that was previously infeasible.
 However, when we want to build a car, draw a painting, or design the machine learning models to which we apply our new algorithms and techniques, we face problems that current machine learning tools are unable to handle.
@@
 
-However, many of the most natural representations of structured domains are best described as programs, are difficult to fit into more restrictive representations.
+However, many of the most natural representations of structured domains are best described as programs and are difficult to fit into more restrictive representations.
 To take an example directly from the machine learning literature, consider the plate notation~\cite{DBLP:journals/jair/Buntine94} commonly used to describe graphical models over collections of variables.
@@
 The goal of plate notation is precisely to express the notion of a ``for loop'' within the graphical model representation.
-The reason this is a convenient tool is that researchs onceptualize their models as generative programs, not graphs; see the equivalent of the LDA graphical model in the pseudo-code in Figure~\ref{fig:lda_code}.
+One reason this is a convenient tool is that researchers conceptualize their models as sampling procedures -- easily expressed as programs (see Figure~\ref{fig:lda_code}, but not as static graphs -- and can use plate notation to express a commonly occuring control structure.
 In fact, the field of probabilistic programming, which has gained recent popularity, is spurred on by exactly the motivation that our models are most naturally expressed as programs.
+
+Suppose, then, that we want to create machine learning algorithms that can help discover good models for data in novel domains; that is, we want algorithms that discovers the structure of the latent graphical model underlying the data.
+This is a field of active research \
 
@@
 	    z +$\sim$+ Categorical(+$\theta$+[m])
-	   w +$\sim$+ Categorical(+$\beta$+[z])
+	    w+$\sim$+ Categorical(+$\beta$+[z])
 	\end{lstlisting}
@@
 \end{figure}
-
-% \begin{figure}[h]
-% \begin{center}
-
-% \end{center}
-% \caption{A typical example of a graphical model in plate notation. }
-% \end{figure}
-
-% \begin{figure}[h]
-% \begin{verbatim}
-
-% for k = 1:NumTopics
-% 	beta[k]  ~ Dirichlet(eta)
-% for m = 1:NumDocuments
-% 	theta[m] ~ Dirichlet(alpha) 
-% 	for n = 1:NumWords
-% 		z ~ Categorical(theta[m])
-% 		w[n, m] ~ Categorical(beta[z])
-
-% \end{verbatim}
-% \end{figure}
 
@@
 
-\section{Submission of papers to NIPS 2013}
-
-NIPS requires electronic submissions.  The electronic submission site is  
-\begin{center}
-   \url{http://papers.nips.cc}
-\end{center}
-
-Please read carefully the
-instructions below, and follow them faithfully.
-\subsection{Style}
-
-Papers to be submitted to NIPS 2013 must be prepared according to the
-instructions presented here. Papers may be only up to eight pages long,
-including figures. Since 2009 an additional ninth page \textit{containing only
-cited references} is allowed. Papers that exceed nine pages will not be
-reviewed, or in any other way considered for presentation at the conference.
-%This is a strict upper bound.
-
-Please note that this year we have introduced automatic line number generation
-into the style file (for \LaTeXe and Word versions). This is to help reviewers
-refer to specific lines of the paper when they make their comments. Please do
-NOT refer to these line numbers in your paper as they will be removed from the
-style file for the final version of accepted papers.
-
-The margins in 2013 are the same as since 2007, which allow for $\approx 15\%$
-more words in the paper compared to earlier years. We are also again using 
-double-blind reviewing. Both of these require the use of new style files.
-
-Authors are required to use the NIPS \LaTeX{} style files obtainable at the
-NIPS website as indicated below. Please make sure you use the current files and
-not previous versions. Tweaking the style files may be grounds for rejection.
-
-%% \subsection{Double-blind reviewing}
-
-%% This year we are doing double-blind reviewing: the reviewers will not know 
-%% who the authors of the paper are. For submission, the NIPS style file will 
-%% automatically anonymize the author list at the beginning of the paper.
-
-%% Please write your paper in such a way to preserve anonymity. Refer to
-%% previous work by the author(s) in the third person, rather than first
-%% person. Do not provide Web links to supporting material at an identifiable
-%% web site.
-
-%%\subsection{Electronic submission}
-%%
-%% \textbf{THE SUBMISSION DEADLINE IS MAY 31st, 2013. SUBMISSIONS MUST BE LOGGED BY
-%% 23:00, MAY 31st, 2013, UNIVERSAL TIME}
-
-%% You must enter your submission in the electronic submission form available at
-%% the NIPS website listed above. You will be asked to enter paper title, name of
-%% all authors, keyword(s), and data about the contact
-%% author (name, full address, telephone, fax, and email). You will need to
-%% upload an electronic (postscript or pdf) version of your paper.
-
-%% You can upload more than one version of your paper, until the
-%% submission deadline. We strongly recommended uploading your paper in
-%% advance of the deadline, so you can avoid last-minute server congestion.
-%%
-%% Note that your submission is only valid if you get an e-mail
-%% confirmation from the server. If you do not get such an e-mail, please
-%% try uploading again.
-
-\subsection{Retrieval of style files}
-
-The style files for NIPS and other conference information are available on the World Wide Web at
-\begin{center}
-   \url{http://www.nips.cc/}
-\end{center}
-The file \verb+nips2013.pdf+ contains these 
-instructions and illustrates the
-various formatting requirements your NIPS paper must satisfy. \LaTeX{}
-users can choose between two style files:
-\verb+nips11submit_09.sty+ (to be used with \LaTeX{} version 2.09) and
-\verb+nips11submit_e.sty+ (to be used with \LaTeX{}2e). The file
-\verb+nips2013.tex+ may be used as a ``shell'' for writing your paper. All you
-have to do is replace the author, title, abstract, and text of the paper with
-your own. The file
-\verb+nips2013.rtf+ is provided as a shell for MS Word users.
-
-The formatting instructions contained in these style files are summarized in
-sections \ref{gen_inst}, \ref{headings}, and \ref{others} below.
-
-%% \subsection{Keywords for paper submission}
-%% Your NIPS paper can be submitted with any of the following keywords (more than one keyword is possible for each paper):
-
-%% \begin{verbatim}
-%% Bioinformatics
-%% Biological Vision
-%% Brain Imaging and Brain Computer Interfacing
-%% Clustering
-%% Cognitive Science
-%% Control and Reinforcement Learning
-%% Dimensionality Reduction and Manifolds
-%% Feature Selection
-%% Gaussian Processes
-%% Graphical Models
-%% Hardware Technologies
-%% Kernels
-%% Learning Theory
-%% Machine Vision
-%% Margins and Boosting
-%% Neural Networks
-%% Neuroscience
-%% Other Algorithms and Architectures
-%% Other Applications
-%% Semi-supervised Learning
-%% Speech and Signal Processing
-%% Text and Language Applications
-
-%% \end{verbatim}
-
-\section{General formatting instructions}
-\label{gen_inst}
-
-The text must be confined within a rectangle 5.5~inches (33~picas) wide and
-9~inches (54~picas) long. The left margin is 1.5~inch (9~picas).
-Use 10~point type with a vertical spacing of 11~points. Times New Roman is the
-preferred typeface throughout. Paragraphs are separated by 1/2~line space,
-with no indentation.
-
-Paper title is 17~point, initial caps/lower case, bold, centered between
-2~horizontal rules. Top rule is 4~points thick and bottom rule is 1~point
-thick. Allow 1/4~inch space above and below title to rules. All pages should
-start at 1~inch (6~picas) from the top of the page.
-
-%The version of the paper submitted for review should have ``Anonymous Author(s)'' as the author of the paper.
-
-For the final version, authors' names are
-set in boldface, and each name is centered above the corresponding
-address. The lead author's name is to be listed first (left-most), and
-the co-authors' names (if different address) are set to follow. If
-there is only one co-author, list both author and co-author side by side.
-
-Please pay special attention to the instructions in section \ref{others}
-regarding figures, tables, acknowledgments, and references.
-
-\section{Headings: first level}
-\label{headings}
-
-First level headings are lower case (except for first word and proper nouns),
-flush left, bold and in point size 12. One line space before the first level
-heading and 1/2~line space after the first level heading.
-
-\subsection{Headings: second level}
-
-Second level headings are lower case (except for first word and proper nouns),
-flush left, bold and in point size 10. One line space before the second level
-heading and 1/2~line space after the second level heading.
-
-\subsubsection{Headings: third level}
-
-Third level headings are lower case (except for first word and proper nouns),
-flush left, bold and in point size 10. One line space before the third level
-heading and 1/2~line space after the third level heading.
-
-\section{Citations, figures, tables, references}
-\label{others}
-
-These instructions apply to everyone, regardless of the formatter being used.
-
-\subsection{Citations within the text}
-
-Citations within the text should be numbered consecutively. The corresponding
-number is to appear enclosed in square brackets, such as [1] or [2]-[5]. The
-corresponding references are to be listed in the same order at the end of the
-paper, in the \textbf{References} section. (Note: the standard
-\textsc{Bib\TeX} style \texttt{unsrt} produces this.) As to the format of the
-references themselves, any style is acceptable as long as it is used
-consistently.
-
-As submission is double blind, refer to your own published work in the 
-third person. That is, use ``In the previous work of Jones et al.\ [4]'',
-not ``In our previous work [4]''. If you cite your other papers that
-are not widely available (e.g.\ a journal paper under review), use
-anonymous author names in the citation, e.g.\ an author of the
-form ``A.\ Anonymous''.
-
-\subsection{Footnotes}
-
-Indicate footnotes with a number\footnote{Sample of the first footnote} in the
-text. Place the footnotes at the bottom of the page on which they appear.
-Precede the footnote with a horizontal rule of 2~inches
-(12~picas).\footnote{Sample of the second footnote}
-
-\subsection{Figures}
-
-All artwork must be neat, clean, and legible. Lines should be dark
-enough for purposes of reproduction; art work should not be
-hand-drawn. The figure number and caption always appear after the
-figure. Place one line space before the figure caption, and one line
-space after the figure. The figure caption is lower case (except for
-first word and proper nouns); figures are numbered consecutively.
-
-Make sure the figure caption does not get separated from the figure.
-Leave sufficient space to avoid splitting the figure and figure caption.
-
-You may use color figures. 
-However, it is best for the
-figure captions and the paper body to make sense if the paper is printed
-either in black/white or in color.
-\begin{figure}[h]
-\begin{center}
-%\framebox[4.0in]{$\;$}
-\fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
-\end{center}
-\caption{Sample figure caption.}
-\end{figure}
-
-% \subsection{Tables}
-
-% All tables must be centered, neat, clean and legible. Do not use hand-drawn
-% tables. The table number and title always appear before the table. See
-% Table~\ref{sample-table}.
-
-% Place one line space before the table title, one line space after the table
-% title, and one line space after the table. The table title must be lower case
-% (except for first word and proper nouns); tables are numbered consecutively.
-
-% \begin{table}[t]
-% \caption{Sample table title}
-% \label{sample-table}
-% \begin{center}
-% \begin{tabular}{ll}
-% \multicolumn{1}{c}{\bf PART}  &\multicolumn{1}{c}{\bf DESCRIPTION}
-% \\ \hline \\
-% Dendrite         &Input terminal \\
-% Axon             &Output terminal \\
-% Soma             &Cell body (contains cell nucleus) \\
-% \end{tabular}
-% \end{center}
-% \end{table}
-
-% \section{Final instructions}
-% Do not change any aspects of the formatting parameters in the style files.
-% In particular, do not modify the width or length of the rectangle the text
-% should fit into, and do not change font sizes (except perhaps in the
-% \textbf{References} section; see below). Please note that pages should be
-% numbered.
-
-% \section{Preparing PostScript or PDF files}
-
-% Please prepare PostScript or PDF files with paper size ``US Letter'', and
-% not, for example, ``A4''. The -t
-% letter option on dvips will produce US Letter files.
-
-% Fonts were the main cause of problems in the past years. Your PDF file must
-% only contain Type 1 or Embedded TrueType fonts. Here are a few instructions
-% to achieve this.
-
-% \begin{itemize}
-
-% \item You can check which fonts a PDF files uses.  In Acrobat Reader,
-% select the menu Files$>$Document Properties$>$Fonts and select Show All Fonts. You can
-% also use the program \verb+pdffonts+ which comes with \verb+xpdf+ and is
-% available out-of-the-box on most Linux machines.
-
-% \item The IEEE has recommendations for generating PDF files whose fonts
-% are also acceptable for NIPS. Please see
-% \url{http://www.emfield.org/icuwb2010/downloads/IEEE-PDF-SpecV32.pdf}
-
-% \item LaTeX users:
-
-% \begin{itemize}
-
-% \item Consider directly generating PDF files using \verb+pdflatex+
-% (especially if you are a MiKTeX user). 
-% PDF figures must be substituted for EPS figures, however.
-
-% \item Otherwise, please generate your PostScript and PDF files with the following commands:
-% \begin{verbatim} 
-% dvips mypaper.dvi -t letter -Ppdf -G0 -o mypaper.ps
-% ps2pdf mypaper.ps mypaper.pdf
-% \end{verbatim}
-
-% Check that the PDF files only contains Type 1 fonts. 
-% %For the final version, please send us both the Postscript file and
-% %the PDF file.
-
-% \item xfig "patterned" shapes are implemented with 
-% bitmap fonts.  Use "solid" shapes instead. 
-% \item The \verb+\bbold+ package almost always uses bitmap
-% fonts.  You can try the equivalent AMS Fonts with command
-% \begin{verbatim}
-% \usepackage[psamsfonts]{amssymb}
-% \end{verbatim}
-%  or use the following workaround for reals, natural and complex: 
-% \begin{verbatim}
-% \newcommand{\RR}{I\!\!R} %real numbers
-% \newcommand{\Nat}{I\!\!N} %natural numbers 
-% \newcommand{\CC}{I\!\!\!\!C} %complex numbers
-% \end{verbatim}
-
-% \item Sometimes the problematic fonts are used in figures
-% included in LaTeX files. The ghostscript program \verb+eps2eps+ is the simplest
-% way to clean such figures. For black and white figures, slightly better
-% results can be achieved with program \verb+potrace+.
-% \end{itemize}
-% \item MSWord and Windows users (via PDF file):
-% \begin{itemize}
-% \item Install the Microsoft Save as PDF Office 2007 Add-in from
-% \url{http://www.microsoft.com/downloads/details.aspx?displaylang=en\&familyid=4d951911-3e7e-4ae6-b059-a2e79ed87041}
-% \item Select ``Save or Publish to PDF'' from the Office or File menu
-% \end{itemize}
-% \item MSWord and Mac OS X users (via PDF file):
-% \begin{itemize}
-% \item From the print menu, click the PDF drop-down box, and select ``Save
-% as PDF...''
-% \end{itemize}
-% \item MSWord and Windows users (via PS file):
-% \begin{itemize}
-% \item To create a new printer
-% on your computer, install the AdobePS printer driver and the Adobe Distiller PPD file from
-% \url{http://www.adobe.com/support/downloads/detail.jsp?ftpID=204} {\it Note:} You must reboot your PC after installing the
-% AdobePS driver for it to take effect.
-% \item To produce the ps file, select ``Print'' from the MS app, choose
-% the installed AdobePS printer, click on ``Properties'', click on ``Advanced.''
-% \item Set ``TrueType Font'' to be ``Download as Softfont''
-% \item Open the ``PostScript Options'' folder
-% \item Select ``PostScript Output Option'' to be ``Optimize for Portability''
-% \item Select ``TrueType Font Download Option'' to be ``Outline''
-% \item Select ``Send PostScript Error Handler'' to be ``No''
-% \item Click ``OK'' three times, print your file.
-% \item Now, use Adobe Acrobat Distiller or ps2pdf to create a PDF file from
-% the PS file. In Acrobat, check the option ``Embed all fonts'' if
-% applicable.
-% \end{itemize}
-
-% \end{itemize}
-% If your file contains Type 3 fonts or non embedded TrueType fonts, we will
-% ask you to fix it.
-
-% \subsection{Margins in LaTeX}
-
-% Most of the margin problems come from figures positioned by hand using
-% \verb+\special+ or other commands. We suggest using the command
-% \verb+\includegraphics+
-% from the graphicx package. Always specify the figure width as a multiple of
-% the line width as in the example below using .eps graphics
-% \begin{verbatim}
-%    \usepackage[dvips]{graphicx} ... 
-%    \includegraphics[width=0.8\linewidth]{myfile.eps} 
-% \end{verbatim}
-% or % Apr 2009 addition
-% \begin{verbatim}
-%    \usepackage[pdftex]{graphicx} ... 
-%    \includegraphics[width=0.8\linewidth]{myfile.pdf} 
-% \end{verbatim}
-% for .pdf graphics. 
-% See section 4.4 in the graphics bundle documentation (\url{http://www.ctan.org/tex-archive/macros/latex/required/graphics/grfguide.ps})
-
-% A number of width problems arise when LaTeX cannot properly hyphenate a
-% line. Please give LaTeX hyphenation hints using the \verb+\-+ command.
-
-% \subsubsection*{Acknowledgments}
-
-% Use unnumbered third level headings for the acknowledgments. All
-% acknowledgments go at the end of the paper. Do not include 
-% acknowledgments in the anonymized submission, only in the 
-% final paper.
-
-% \subsubsection*{References}
-
-% References follow the acknowledgments. Use unnumbered third level heading for
-% the references. Any choice of citation style is acceptable as long as you are
-% consistent. It is permissible to reduce the font size to `small' (9-point) 
-% when listing the references. {\bf Remember that this year you can use
-% a ninth page as long as it contains \emph{only} cited references.}
-
-% \small{
-% [1] Alexander, J.A. \& Mozer, M.C. (1995) Template-based algorithms
-% for connectionist rule extraction. In G. Tesauro, D. S. Touretzky
-% and T.K. Leen (eds.), {\it Advances in Neural Information Processing
-% Systems 7}, pp. 609-616. Cambridge, MA: MIT Press.
-
-% [2] Bower, J.M. \& Beeman, D. (1995) {\it The Book of GENESIS: Exploring
-% Realistic Neural Models with the GEneral NEural SImulation System.}
-% New York: TELOS/Springer-Verlag.
-
-% [3] Hasselmo, M.E., Schnell, E. \& Barkai, E. (1995) Dynamics of learning
-% and recall at excitatory recurrent synapses and cholinergic modulation
-% in rat hippocampal region CA3. {\it Journal of Neuroscience}
-% {\bf 15}(7):5249-5262.
-% }
-
 \bibliographystyle{plain}

### 2013-10-22 drafts/nips/nips2013.tex

--- 2013-10-22 drafts/nips/nips2013.tex
@@
 
-Motivation: Many approaches to constructive machine learning use structured representations, such as graphs, grammars, Horn clauses.
-All of these may be thought of as a program - programs are at least as expressive as these other representations.
+The vast majority of research in machine learning is focused on utilizing large amounts of noisy data living in high-dimensional, real-valued vector spaces.
+Such research has been immensely successful and has enabled automatic and accurate modelling on a scale that was previously infeasible.
+However, when we want to build a car, draw a painting, or design the machine learning models to which we apply our new algorithms and techniques, we face problems that current machine learning tools are unable to handle.
+The highly structured objects that humans produce seems to require a different representation than the ones which machine learning algorithms have most successfully employed.
 
-The space of programs is enormous and inherently unpredictable \cite{PACunpredictability}.
-We introduce a new algorithm, SEC, (TODO: Sequential EC? SEC? What is a good name for this?) for searching the space of programs.
-Two main ideas underly SEC:
-\begin{itemize}
-\item \emph{Learning to search}: Programs admit compositionality and reuse of subprograms.
-SEC exploits this compositionality to learn a library of program fragments, which it then reuses when performing further searches for programs.
-In a multitask setting, this permits bootstrap learning of more complicated programs after the synthesis of simpler programs.
-\item \emph{Incremental Program Synthesis}: In many search problems, there is a natural means of assigning partial credit for partially-correct solutions.
-SEC uses this partial credit to select promising programs and modify them, creating complicated programs step-by-step.
-Rather than using a mutation operator, as in genetic programming, or Metropolis-Hastings proposals, as in \cite{MHprogramInduction}, SEC uses function composition to combine smaller programs in to larger ones.
-\end{itemize}
+To address this problem, many structured representations have been proposed in the machine learning and AI literatures: learning over graphs \cite{}, grammars\cite{}, Horn clauses\cite{}, relations{}, etc., has been investigated.
+These are representations that attempt to bridge the gap between the tractablity of learning and reasoning in smooth continuous spaces and the expressivity of general purpose programming languages.
 
-Applications: graphical model structure, towers
+However, many of the most natural representations of structured domains are best described as programs and are difficult to fit into more restrictive representations.
+To take an example directly from the machine learning literature, consider the plate notation~\cite{DBLP:journals/jair/Buntine94} commonly used to describe graphical models over collections of variables.
+A typical example is shown in Figure~\ref{fig:lda_plate}~\cite{DBLP:journals/jmlr/BleiNJ03}.
+The goal of plate notation is precisely to express the notion of a ``for loop'' within the graphical model representation.
+One reason this is a convenient tool is that researchers conceptualize their models as sampling procedures -- easily expressed as programs (see Figure~\ref{fig:lda_code}, but not as static graphs -- and can use plate notation to express a commonly occuring control structure.
+In fact, the field of probabilistic programming, which has gained recent popularity, is spurred on by exactly the motivation that our models are most naturally expressed as programs.
 
-The SEC algorithm is motivated by a generative model over compositions of functions drawn from a stochastic grammar over programs.
+Suppose, then, that we want to create machine learning algorithms that can help discover good models for data in novel domains; that is, we want algorithms that discovers the structure of the latent graphical model underlying the data.
+This is a field of active research \
 
-SEC treats inference as a planning problem
+In this paper, we describe our approach to learning over structured representations as a search over the space of programs in a functional programming language.
+We characterize learning in this space as incrementally developing a weighted library of program abstractions, i.e.
+a collection of subroutines, that are commonly used to solve problems in a given domain.
+The SEC algorithm (for Sequential Exploration-Compression), which we present here, takes a set of related tasks and produces a distribution over programs that solve those tasks; as it refines this distribution, SEC is able to search the space of programs more efficiently and solve more tasks.
+
+We will present here two domains of interest.
+First, continuing with the example above, we will describe how SEC can learn to efficiently explore the space of relevant graphical models for a given problem.
+Second, we will examine SEC's performance on a completely different type of task, in which we want to an algorithm to build tall and physically stable block towers.
+
+\lstset{
+	basicstyle=\scriptsize\ttfamily,
+    commentstyle=\color{white!40!black},
+    escapechar=+,
+    escapebegin=\color{white!40!black},
+    escapeend={},
+    keywordstyle=\bfseries\color{Periwinkle},
+    numbers=left,
+    numberstyle=\sffamily\tiny\color{gray},
+}
+
+\begin{figure}[h]
+\centering
+\begin{subfigure}{.5\textwidth}
+  \centering
+\includegraphics[width=.7\linewidth]{./figures/lda_plate_blei.pdf}
+  \caption{An example of a typical plate notation use. \label{fig:lda_plate}}
+
+\end{subfigure}%
+\begin{subfigure}{.4\textwidth}
+	\begin{lstlisting}[frame=single, numbers=none, xleftmargin=0pt]
+	for +$k$+ = 1:NumTopics
+	  +$\beta$+[k] +$\sim$+ Dirichlet(+$\eta$+)
+	for m = 1:NumDocuments
+	  +$\theta$+[m] +$\sim$+ Dirichlet(+$\alpha$+) 
+	  for n = 1:NumWords 
+	    z +$\sim$+ Categorical(+$\theta$+[m])
+	    w+$\sim$+ Categorical(+$\beta$+[z])
+	\end{lstlisting}
+  \caption{The same model as a program.}
+  \label{fig:lda_code}
+\end{subfigure}
+\caption{The popular Latent Dirichlet Allocation model in two forms.}
+\label{fig:lda}
+\end{figure}
 
@@
 to achieve this.
+=======
+>>>>>>> 0c3d0be073ba9aaf201c99bcf3dc620f5bb9f7cb
 
-\begin{itemize}
-
-\item You can check which fonts a PDF files uses.  In Acrobat Reader,
-select the menu Files$>$Document Properties$>$Fonts and select Show All Fonts. You can
-also use the program \verb+pdffonts+ which comes with \verb+xpdf+ and is
-available out-of-the-box on most Linux machines.
-
-\item The IEEE has recommendations for generating PDF files whose fonts
-are also acceptable for NIPS. Please see
-\url{http://www.emfield.org/icuwb2010/downloads/IEEE-PDF-SpecV32.pdf}
-
-\item LaTeX users:
-
-\begin{itemize}
-
-\item Consider directly generating PDF files using \verb+pdflatex+
-(especially if you are a MiKTeX user). 
-PDF figures must be substituted for EPS figures, however.
-
-\item Otherwise, please generate your PostScript and PDF files with the following commands:
-\begin{verbatim} 
-dvips mypaper.dvi -t letter -Ppdf -G0 -o mypaper.ps
-ps2pdf mypaper.ps mypaper.pdf
-\end{verbatim}
-
-Check that the PDF files only contains Type 1 fonts. 
-%For the final version, please send us both the Postscript file and
-%the PDF file.
-
-\item xfig "patterned" shapes are implemented with 
-bitmap fonts.  Use "solid" shapes instead. 
-\item The \verb+\bbold+ package almost always uses bitmap
-fonts.  You can try the equivalent AMS Fonts with command
-\begin{verbatim}
-\usepackage[psamsfonts]{amssymb}
-\end{verbatim}
- or use the following workaround for reals, natural and complex: 
-\begin{verbatim}
-\newcommand{\RR}{I\!\!R} %real numbers
-\newcommand{\Nat}{I\!\!N} %natural numbers 
-\newcommand{\CC}{I\!\!\!\!C} %complex numbers
-\end{verbatim}
-
-\item Sometimes the problematic fonts are used in figures
-included in LaTeX files. The ghostscript program \verb+eps2eps+ is the simplest
-way to clean such figures. For black and white figures, slightly better
-results can be achieved with program \verb+potrace+.
-\end{itemize}
-\item MSWord and Windows users (via PDF file):
-\begin{itemize}
-\item Install the Microsoft Save as PDF Office 2007 Add-in from
-\url{http://www.microsoft.com/downloads/details.aspx?displaylang=en\&familyid=4d951911-3e7e-4ae6-b059-a2e79ed87041}
-\item Select ``Save or Publish to PDF'' from the Office or File menu
-\end{itemize}
-\item MSWord and Mac OS X users (via PDF file):
-\begin{itemize}
-\item From the print menu, click the PDF drop-down box, and select ``Save
-as PDF...''
-\end{itemize}
-\item MSWord and Windows users (via PS file):
-\begin{itemize}
-\item To create a new printer
-on your computer, install the AdobePS printer driver and the Adobe Distiller PPD file from
-\url{http://www.adobe.com/support/downloads/detail.jsp?ftpID=204} {\it Note:} You must reboot your PC after installing the
-AdobePS driver for it to take effect.
-\item To produce the ps file, select ``Print'' from the MS app, choose
-the installed AdobePS printer, click on ``Properties'', click on ``Advanced.''
-\item Set ``TrueType Font'' to be ``Download as Softfont''
-\item Open the ``PostScript Options'' folder
-\item Select ``PostScript Output Option'' to be ``Optimize for Portability''
-\item Select ``TrueType Font Download Option'' to be ``Outline''
-\item Select ``Send PostScript Error Handler'' to be ``No''
-\item Click ``OK'' three times, print your file.
-\item Now, use Adobe Acrobat Distiller or ps2pdf to create a PDF file from
-the PS file. In Acrobat, check the option ``Embed all fonts'' if
-applicable.
-\end{itemize}
-
-\end{itemize}
-If your file contains Type 3 fonts or non embedded TrueType fonts, we will
-ask you to fix it.
-
-\subsection{Margins in LaTeX}
-
-Most of the margin problems come from figures positioned by hand using
-\verb+\special+ or other commands. We suggest using the command
-\verb+\includegraphics+
-from the graphicx package. Always specify the figure width as a multiple of
-the line width as in the example below using .eps graphics
-\begin{verbatim}
-   \usepackage[dvips]{graphicx} ... 
-   \includegraphics[width=0.8\linewidth]{myfile.eps} 
-\end{verbatim}
-or % Apr 2009 addition
-\begin{verbatim}
-   \usepackage[pdftex]{graphicx} ... 
-   \includegraphics[width=0.8\linewidth]{myfile.pdf} 
-\end{verbatim}
-for .pdf graphics. 
-See section 4.4 in the graphics bundle documentation (\url{http://www.ctan.org/tex-archive/macros/latex/required/graphics/grfguide.ps})
-
-A number of width problems arise when LaTeX cannot properly hyphenate a
-line. Please give LaTeX hyphenation hints using the \verb+\-+ command.
-
-\subsubsection*{Acknowledgments}
-
-Use unnumbered third level headings for the acknowledgments. All
-acknowledgments go at the end of the paper. Do not include 
-acknowledgments in the anonymized submission, only in the 
-final paper.
-
-\subsubsection*{References}
-
-References follow the acknowledgments. Use unnumbered third level heading for
-the references. Any choice of citation style is acceptable as long as you are
-consistent. It is permissible to reduce the font size to `small' (9-point) 
-when listing the references. {\bf Remember that this year you can use
-a ninth page as long as it contains \emph{only} cited references.}
-
-\small{
-[1] Alexander, J.A. \& Mozer, M.C. (1995) Template-based algorithms
-for connectionist rule extraction. In G. Tesauro, D. S. Touretzky
-and T.K. Leen (eds.), {\it Advances in Neural Information Processing
-Systems 7}, pp. 609-616. Cambridge, MA: MIT Press.
-
-[2] Bower, J.M. \& Beeman, D. (1995) {\it The Book of GENESIS: Exploring
-Realistic Neural Models with the GEneral NEural SImulation System.}
-New York: TELOS/Springer-Verlag.
-
-[3] Hasselmo, M.E., Schnell, E. \& Barkai, E. (1995) Dynamics of learning
-and recall at excitatory recurrent synapses and cholinergic modulation
-in rat hippocampal region CA3. {\it Journal of Neuroscience}
-{\bf 15}(7):5249-5262.
-}
+\bibliographystyle{plain}
+\bibliography{nips2013}
 

### 2013-10-23 drafts/nips/nips2013.tex
Added new bibliography file. Added some figures.

--- 2013-10-22 drafts/nips/nips2013.tex
@@
 
-\section{Application to Structure Learning}
+\section{Application to Graphical Model Structure Learning}
+
+Learning the structure of graphical models that capture the independencies underlying a dataset is an active area of research\cite{adams-wallach-ghahramani-2010a}\cite{ISI:000240797500002}\cite{ISI:000178037200004}\cite{ISI:A1995RX35400001}.
+Graphical model structures enforce the conditional independences a modeler believes exists among latent and observed random variables\cite{DBLP:books/daglib/0066829}.
 
@@
 
+\begin{figure}
+\begin{minipage}[c][8cm][t]{.4\textwidth}
+  \vspace*{\fill}
+  \includegraphics[width=\linewidth]{./figures/tasks.pdf}
+  \caption{test figure one}
+  \label{fig:test1}
+\end{minipage}%
+\begin{minipage}[c][8cm][t]{.3\textwidth}
+  \vspace*{\fill}
+  \fbox{\includegraphics[width=\linewidth]{./figures/GraphCombinators.pdf}}
+  \caption{test figure two}
+  \label{fig:test2}\par\vfill
+  % \includegraphics[width=5cm,height=4.5cm]{im
+  \caption{test figure three}
+  \label{fig:test3}
+\end{minipage}
+\end{figure}
 \section{Building Towers in Blocks World}
@@
 \bibliographystyle{plain}
-\bibliography{nips2013}
+\bibliography{eyal_bib}
 

### 2013-10-23 drafts/nips/nips2013.tex
more writeup

--- 2013-10-22 drafts/nips/nips2013.tex
@@
 \usepackage{listings}
+\usepackage{clrscode4e}
+\usepackage{tikz}
+\usepackage{amsmath}
+
 %\documentstyle[nips13submit_09,times,art10]{article} % For LaTeX 2.09
@@
 \includegraphics[width=.7\linewidth]{./figures/lda_plate_blei.pdf}
-  \caption{An example of a typical plate notation use. \label{fig:lda_plate}}
+  \caption{An example of typical plate notation use. \label{fig:lda_plate}}
 
@@
 \subsection{Generative Model}
+We propose a generative model over programs that captures two intuitions:
+\begin{itemize}
+\item{Useful programs are composed of useful subparts.}
+\item{Short programs are a prior more likely.}
+\end{itemize}
+\begin{codebox}
+\li \id{G} $\sim$ $P_G(\cdot)$ \Comment{Draw grammar from description-length prior}
+\li \id{\ell_n} $\sim\mbox{Uniform}(1, L)$ \Comment{Draw number of subprograms in $n^{th}$ program}
+\li \Comment{Draw $\rho_n$, the $n^{th}$ program, by composing its subprograms, $e_n^{i}$}
+\li \id{e_n^{i}} $\sim P_{e|G}(\cdot | G)$
+\li \id{\rho_n} $= e_n^{\ell_n} (e_n^{\ell_n-1} ( \cdots e_n^{1}))$
+\li \Comment{Draw $t_n$ by adding noise $\epsilon_n$ to the output of $\rho_n$}
+\li \id{\epsilon_n} $\sim P_{\mbox{noise}}(\cdot)$
+\li \id{t_n} $= \rho_n + \epsilon_n$
+\end{codebox}
+
 Write down the generative model. Connect to EC and Percy Liang's work. Discuss ``programs as plans.''
 \subsection{Inference}
-Write down EM updates. Interpret them as a form of compression solved via a weighted Neville-Manning variant.
+We implement an iterative Expectation-Maximization algorithm that performs MAP estimation of the libraray, $G$, given the tasks, $\{t_n\}$.
+The joint distribution factorizes as
+\begin{equation}
+P(G,\{\rho_n\},\{t_n\}) = \frac{1}{L^N} P(G) \prod_n P(t_n | \rho_n) \prod_i P(e^i_n | G)
+\label{eq:joint}
+\end{equation}
+From equation (\ref{eq:joint}), the EM updates are
+\begin{eqnarray}
+q(\rho_n) &\propto& P(t_n | \rho_n) \prod_i P(e^i_n | G^{old})\\
+\label{eq:qdist}
+G^{new} &=& \operatorname{argmin}_G \left( -\ln P(G) -
+\sum_n 
+E_{q_n}
+ \left[ \ln P(\rho_n | G) \right] \right)
+ \label{eq:gmax}
+\end{eqnarray}
+Exact computation of $q(\cdot)$ in equation (\ref{eq:qdist}) would involve marginalizing over the space of all programs.
+Instead, we perform a heuristic search over the space of programs, with the goal of finding those programs for which $q(\cdot)$ is high.
+Rather than use mutation and crossover operators, as done in genetic programming, or random resampling of subprograms, as done in recent work in applying Metropolis-Hastings to program learning, we propose that the search moves should themselves be programs drawn from the grammar $G$.
+Concretely, in order to modify a program $e$ to produce new candidate programs, we draw a new function from our grammar, $G$, and apply that new function to $e$.
+Thus, we use the grammar not only as a distribution over programs of interest, but also as a distribution over programs that take as input other programs.
+This incremental construction of programs by repeatedly applying draws from the grammar mirrors the construction of programs in the generative model.
 
-Discuss bootstrap learning: the right curriculum allows us to build more complicated programs.
-Make some analogy involving how humans learn to paint, write poetry, bake, compose music, play with blocks, or design statistical models.
+In our experiments, we used beam search. The objective function was the unnormalized posterior $P_{t_n|\rho_n}(t_n | \cdot )P_{\rho_n | G}(\cdot | G)$.
+The beam was initialized to be the programs with the highest prior probability according to the grammar $G$.
+Similarly, the search moves considered were those functions with the most highest prior probability under $G$.
+
+Equation (\ref{eq:gmax}) has a natural interpretation as a form of compression.
+Interpreting negative log probabilities as description lengths, the update (\ref{eq:gmax}) picks the grammar minimizing the sum of the description length of the grammar, plus the expected description length of the programs found to solve the tasks.
+
+Performing the minimization operation in equation (\ref{eq:gmax}) is generally difficult.
+We approximate the description length of the grammar by the number of unique program fragments in the grammar, multiplied by a regularization constant.
+The description length of a program is approximated by the number of (not necessarily unique) program fragments in its smallest parse tree given the grammar.
+A slight generalization of the Neville-Manning algorithm then permits tractible minimization of (\ref{eq:gmax}).
+Estimation of the production probabilities of $G$ is performed as in \cite{ijcai}.
 

### 2013-10-23 src/Strappy/BeamSearch.hs

@@ -42,31 +42,42 @@ beamPlan :: PlanTask ->
             [(Expr, Double)] -> -- ^ (Log) Distribution over seed expressions drawn from the grammar
             [(Expr, Double)] -> -- ^ (Log) Distribution over transition expressions drawn from the grammar
             Int -> -- ^ length of the plan
-            IO (Map.Map Expr Double, Set.Set ([Expr], Double)) -- ^ Log rewards for each expression, programs hitting task
+            IO (Map.Map Expr Double, -- ^ Log rewards for each expression
+                Set.Set ([Expr], Double), -- ^ programs hitting task
+                Maybe ([Expr], Double, Double)) -- ^ Best partial credit
 beamPlan pt beamSize seedDist dist planLen = do
     seedRewards <- forM seedDist $ \(e, pLL) -> ptLogLikelihood pt e >>=
                                                 \ll -> return (ll, (e, pLL))
+    let seedPartial = foldl mergePartial Nothing $ flip map seedRewards $ \(ll, (e, pLL)) ->
+                                                                          if isInvalidNum ll
+                                                                          then Nothing
+                                                                          else Just ([e], ll, pLL)
     let seedPlans = map snd $ MinQueue.toList $
                     foldl (\beam (ll, (e, pLL)) -> insertIntoBeam beamSize beam (ll+pLL, pLL) ([e],pLL)) MinQueue.empty seedRewards
     let seedRewards' = filter (not . isInvalidNum . fst) seedRewards
     let seedRewards'' = Map.fromList $ map (\ (ll, (e, pLL)) -> (e, ll+pLL)) seedRewards'
     let seedLogZ = logSumExpList $ map (\ (ll, (_, pLL)) -> ll + pLL) seedRewards
     let seedHits = Set.fromList [([e], pLL+ll) | (ll, (e, pLL)) <- seedRewards, ll > -0.1 ]
-    (logRewards, logZ, hits) <- bs seedPlans planLen seedRewards'' seedLogZ seedHits
-    return (Map.map (\x -> x-logZ) logRewards, hits)
+    (logRewards, logZ, hits, part) <- bs seedPlans planLen seedRewards'' seedLogZ seedHits seedPartial
+    return (Map.map (\x -> x-logZ) logRewards, hits, part)
     where bs :: [([Expr], Double)] -> -- ^ Incoming plans to be extended
                 Int -> -- ^ Remaining plan length
[…]

### 2013-10-23 src/Strappy/UGM.hs

@@ -1,6 +1,6 @@
 
---module Strappy.UGM where
-module Main where
+module Strappy.UGM where
+--module Main where
 
 
 import Strappy.EM
@@ -87,7 +87,7 @@ ugmUnionCorrespondence ugm ugm' =
 -- | Connects first vertex of a UGM to the last vertex of another UGM
 ugmUnionHeadTail :: UGM -> UGM -> UGM
 ugmUnionHeadTail ugm ugm' =
-    let (UGM vs1 es1, UGM vs2 es2) = ugmRename2Disjoint ugm ugm'
+    let (UGM vs1 es1, UGM vs2 es2) = (ugm, ugm') --ugmRename2Disjoint ugm ugm'
         headV = minimum vs2
         tailV = maximum vs1
         newEdge = if null vs1 || null vs2 || headV == tailV
@@ -145,12 +145,15 @@ ugmrHMM vs = ugmUnionCorrespondence (ugmChain vs) (foldl1 ugmUnion $ map ugmLate
 ugmCylinder :: [UGM] -> UGM
 ugmCylinder vs = ugmUnionCorrespondence (ugmRing vs) (ugmRing $ map ugmLatentify vs)
 
+ugmDeepCylinder :: [[UGM]] -> UGM
+ugmDeepCylinder vss = foldr1 ugmUnionCorrespondence (map ugmRing vss)
+
 ugmMRF :: [[UGM]] -> UGM
 ugmMRF vss = ugmUnionCorrespondence (foldr1 ugmUnion $ concat vss) $
              ugmIsing $ map (map ugmLatentify) vss
 
 ugmLatentCommon :: [UGM] -> UGM
-ugmLatentCommon = foldl ugmUnionHeadTail (singletonUGM $ Latent 1)
+ugmLatentCommon = foldl ugmUnionHeadTail (singletonUGM $ Latent 0)
 
 -- Just for fun: the graphical model behind EC
 ugmEC :: [UGM] -> UGM
@@ -169,7 +172,7 @@ partialCredit (UGM vs es) (UGM vs' es') =
   in if vs1 == vs2
      then if es1 == es2
           then 1.0
-          else 0.5
+          else 0.15
       else 0.0
 
 makeUGMTask nm tp tests =
@@ -238,6 +241,12 @@ cylTask = makeUGMTask "cylinder" (tList tUGM ->- tUGM)
              (isingRow 1 2, ugmCylinder (isingRow 1 2)),
              (isingRow 1 3, ugmCylinder (isingRow 1 3)),
[…]

### 2013-10-23 drafts/nips/nips2013.tex

@@
 
+\section{Application to Graphical Model Structure Learning}
+
+Learning the structure of graphical models that capture the independencies underlying a dataset is an active area of research\cite{adams-wallach-ghahramani-2010a}\cite{ISI:000240797500002}\cite{ISI:000178037200004}\cite{ISI:A1995RX35400001}.
+Graphical model structures enforce the conditional independences a modeler believes exists among latent and observed random variables\cite{DBLP:books/daglib/0066829}.
+
 \subsection{Graph Combinators}

### 2013-10-23 drafts/nips/nips2013.tex
Edited abstract.

--- 2013-10-23 drafts/nips/nips2013.tex
@@
 For example, suppose a researcher wants to design a probabilistic graphical model for a novel domain.
-On the one hand, designing a good model requires a great deal of domain-specific knowledge and expertise, which may be hard to come by.
-On the other hand, searching the space of probabilistic models automatically is notoriously difficult.
-However, when researchers do set out to design graphical models, they draw on abstractions such as trees, chains, grids and plates to constrain and direct the kinds of models they produce.
-Thus, if we want to produce machine learning tools that aid in this task, we want tools that can learn these kinds of relevant abstractions.
-One natural way to think of these abstractions is as programs that take sets of random variables and produce graphical models that relate them.
-In this work, we describe the SEC algorithm, which attempts to learn distributions over such programs by incrementally finding program components that commonly help solve problems in a given domain.
-We discuss the behavior of the SEC algorithm for the task of graphical model structure learning.
-We also describe SEC's behavior on the dissimilar problem of automatically constructing tall and physically stable towers of blocks.
+Searching the space of probabilistic models automatically is notoriously difficult and is especially difficult when the latent variables are possible.
+However, when researchers seem able to relatively easily adapt commonly used modeling motifs to new domains.
+In doing so, they draw on abstractions such as trees, chains, grids and plates to constrain and direct the kinds of models they produce.
+This suggests that before we ask machine learning algorithms to discover parsimonious models of new domains, we should develop techniques that enable our algorithms to automatically learn these ``graphical conepts'' in much the same way that researchers themselves do, by seeing examples in the literature.
+One natural way to think of these graphical concepts is as programs that take sets of random variables and produce graphical models that relate them.
+In this work, we describe the SEC algorithm, which attempts to learn a distribution over programs by incrementally finding program components that commonly help to solve problems in a given domain, and we show prelimarily that SEC is able to discover the graphical concepts that underlie many of the common graphical model structures.
 \end{abstract}

### 2013-10-23 drafts/nips/nips2013.tex
Added a description of the combinatory logic.

--- 2013-10-23 drafts/nips/nips2013.tex
@@
 \usepackage{listings}
+\usepackage{clrscode4e}
+\usepackage{tikz}
+\usepackage{amsmath}
+
 %\documentstyle[nips13submit_09,times,art10]{article} % For LaTeX 2.09
 
-\title{Learning Programs through Bayesian Inference and Planning}
+\title{Incremental Construction of Programs in a Multitask Setting}
 
@@
 
-Here, we present an approach to learning
-
-Motivation: Many approaches to constructive machine learning use structured representations, such as graphs, grammars, Horn clauses, relations, ..., ???.
-All of these may be thought of as a program - programs are at least as expressive as these other representations.
-
-% TODO: Sequential EC? SEC? What is a good name for this?
-The space of programs is enormous and inherently unpredictable \cite{PACunpredictability}.
-We introduce a new algorithm, SEC, for searching the space of programs.
-Two main ideas underly SEC:
-\begin{itemize}
-\item \emph{Learning to search}: Programs admit compositionality and reuse of subprograms.
-SEC exploits this compositionality to learn a library of program fragments, which it then reuses when performing further searches for programs.
-In a multitask setting, this permits bootstrap learning of more complicated programs after the synthesis of simpler programs.
-\item \emph{Incremental Program Synthesis}: In many search problems, there is a natural means of assigning partial credit for partially-correct solutions.
-SEC uses this partial credit to select promising programs and modify them, creating complicated programs step-by-step.
-Rather than using a mutation operator, as in genetic programming, or Metropolis-Hastings proposals, as in \cite{MHprogramInduction}, SEC uses function composition to combine smaller programs in to larger ones.
-\end{itemize}
-
-The SEC algorithm comes from a generative model over compositions of functions drawn from a stochastic grammar over programs.
-We describe this model, as well as how we perform inference, in ...
-
-SEC treats planning as an inference problem.
-
 \section{The SEC Algorithm}
 \subsection{Generative Model}
+We propose a generative model over programs that captures two intuitions:
+\begin{itemize}
+\item{Useful programs are composed of useful subparts.}
+\item{Short programs are a prior more likely.}
+\end{itemize}
+\begin{codebox}
+\li \id{G} $\sim$ $P_G(\cdot)$ \Comment{Draw grammar from description-length prior}
+\li \id{\ell_n} $\sim\mbox{Uniform}(1, L)$ \Comment{Draw number of subprograms in $n^{th}$ program}
+\li \Comment{Draw $\rho_n$, the $n^{th}$ program, by composing its subprograms, $e_n^{i}$}
+\li \id{e_n^{i}} $\sim P_{e|G}(\cdot | G)$
+\li \id{\rho_n} $= e_n^{\ell_n} (e_n^{\ell_n-1} ( \cdots e_n^{1}))$
+\li \Comment{Draw $t_n$ by adding noise $\epsilon_n$ to the output of $\rho_n$}
+\li \id{\epsilon_n} $\sim P_{\mbox{noise}}(\cdot)$
+\li \id{t_n} $= \rho_n + \epsilon_n$
+\end{codebox}
+
+Write down the generative model. Connect to EC and Percy Liang's work. Discuss ``programs as plans.''
 \subsection{Inference}
+We implement an iterative Expectation-Maximization algorithm that performs MAP estimation of the libraray, $G$, given the tasks, $\{t_n\}$.
+The joint distribution factorizes as
+\begin{equation}
+P(G,\{\rho_n\},\{t_n\}) = \frac{1}{L^N} P(G) \prod_n P(t_n | \rho_n) \prod_i P(e^i_n | G)
+\label{eq:joint}
+\end{equation}
+From equation (\ref{eq:joint}), the EM updates are
+\begin{eqnarray}
+q(\rho_n) &\propto& P(t_n | \rho_n) \prod_i P(e^i_n | G^{old})\\
+\label{eq:qdist}
+G^{new} &=& \operatorname{argmin}_G \left( -\ln P(G) -
+\sum_n 
+E_{q_n}
+ \left[ \ln P(\rho_n | G) \right] \right)
+ \label{eq:gmax}
+\end{eqnarray}
+Exact computation of $q(\cdot)$ in equation (\ref{eq:qdist}) would involve marginalizing over the space of all programs.
+Instead, we perform a heuristic search over the space of programs, with the goal of finding those programs for which $q(\cdot)$ is high.
+Rather than use mutation and crossover operators, as done in genetic programming, or random resampling of subprograms, as done in recent work in applying Metropolis-Hastings to program learning, we propose that the search moves should themselves be programs drawn from the grammar $G$.
+Concretely, in order to modify a program $e$ to produce new candidate programs, we draw a new function from our grammar, $G$, and apply that new function to $e$.
+Thus, we use the grammar not only as a distribution over programs of interest, but also as a distribution over programs that take as input other programs.
+This incremental construction of programs by repeatedly applying draws from the grammar mirrors the construction of programs in the generative model.
+
+In our experiments, we used beam search. The objective function was the unnormalized posterior $P_{t_n|\rho_n}(t_n | \cdot )P_{\rho_n | G}(\cdot | G)$.
+The beam was initialized to be the programs with the highest prior probability according to the grammar $G$.
+Similarly, the search moves considered were those functions with the most highest prior probability under $G$.
+
+Equation (\ref{eq:gmax}) has a natural interpretation as a form of compression.
+Interpreting negative log probabilities as description lengths, the update (\ref{eq:gmax}) picks the grammar minimizing the sum of the description length of the grammar, plus the expected description length of the programs found to solve the tasks.
+
+Performing the minimization operation in equation (\ref{eq:gmax}) is generally difficult.
+We approximate the description length of the grammar by the number of unique program fragments in the grammar, multiplied by a regularization constant.
+The description length of a program is approximated by the number of (not necessarily unique) program fragments in its smallest parse tree given the grammar.
+A slight generalization of the Neville-Manning algorithm then permits tractible minimization of (\ref{eq:gmax}).
+Estimation of the production probabilities of $G$ is performed as in \cite{ijcai}.
+
 \subsection{Program Representation}
+
+Following~\cite{DBLP:conf/ijcai/DechterMAT13} and~\cite{DBLP:conf/icml/LiangJK10}, we use a
+simply typed combinatory calculus -- a variable-free subset of the
+polymorphic simply-typed lambda calculus~\cite{DBLP:books/daglib/0005958} -- as our
+program representation language.
+The simplicity the combinatory calculus makes describing a distribution over its expressions particularly simple.
+A more detailed explanation and discussion of this program representation can be found in the papers cited above.
+
+\section{Application to Structure Learning}
+Learning the structure of graphical models is a difficult problem - not only for machines, but also for humans, where expert knowledge is often required.
+In practice, the sorts of graphical models that humans design, such as Hidden Markov Models, phylogenetic trees, topic models, and Ising models, exhibit certain symmetries and recursive structures that are amenable to synthesis by programs.
+Could an algorithm, such as SEC, automatically learn these structures?
 
@@
 \end{figure}
-\section{Building Towers in Blocks World}
-
-\subsection{The Blocks World Domain}
-
-\subsection{Learned Tower programs}
 

### 2013-10-23 drafts/nips/nips2013.tex
Edited SEC algorithm seciton.

@@
 \section{The SEC Algorithm}
+
+The SEC algorithm takes a library of primitives and a set of related tasks to solve; its goal is to produce a set of programs that solve these tasks.
+Like the EC algorithm~\cite{DBLP:conf/ijcai/DechterMAT13} (of which it an extension), the SEC algorithm maintains a distribution $\mathcal{D}$ over programs and this distribution helps guide its search to relevant parts of of the space of programs.
+In each iteration, SEC does two things:
+\begin{itemize}
+	\item first, it uses $\mathcal{D}$ to explore the space of programs that are likely to be relevant for the tasks it has been given;
+	\item second, it compresses the solutions it found and uses this compression generate a new weighted library of primitives that defines an updated $\mathcal{D}$.
+\end{itemize}
+
 \subsection{Generative Model}
-We propose a generative model over programs that captures two intuitions:
+The SEC algorithm can be expressed as MAP inference in a probabilistic generative model over the space of programs that captures two intuitions:
 \begin{itemize}
-\item{Useful programs are composed of useful subparts.}
-\item{Short programs are a prior more likely.}
+\item{For a set of related tasks, useful programs are composed of useful subparts.}
+\item{Short programs are a priori more likely than long programs.}
 \end{itemize}

### 2013-10-23 drafts/nips/nips2013.tex
updated figures and edited section on algorithm.

@@
 
-Write down the generative model. Connect to EC and Percy Liang's work. Discuss ``programs as plans.''
+In short, this generative model assumes that the programs that solve a set of related tasks are themselves constructing by composing program fragments that are drawn from a common stochastic grammar.
+This grammar is itself drawn from a distribution over grammars that is biased towards grammars with smaller descriptions lengths.
+
 \subsection{Inference}
-We implement an iterative Expectation-Maximization algorithm that performs MAP estimation of the libraray, $G$, given the tasks, $\{t_n\}$.
+We implement an iterative Expectation-Maximization algorithm that performs MAP estimation of the grammar, $G$, given the tasks, $\{t_n\}$.
 The joint distribution factorizes as

### 2013-10-23 drafts/nips/nips2013.tex
writeup

--- 2013-10-23 drafts/nips/nips2013.tex
@@
 For example, suppose a researcher wants to design a probabilistic graphical model for a novel domain.
-Searching the space of probabilistic models automatically is notoriously difficult and is especially difficult when the latent variables are possible.
-However, when researchers seem able to relatively easily adapt commonly used modeling motifs to new domains.
-In doing so, they draw on abstractions such as trees, chains, grids and plates to constrain and direct the kinds of models they produce.
-This suggests that before we ask machine learning algorithms to discover parsimonious models of new domains, we should develop techniques that enable our algorithms to automatically learn these ``graphical conepts'' in much the same way that researchers themselves do, by seeing examples in the literature.
-One natural way to think of these graphical concepts is as programs that take sets of random variables and produce graphical models that relate them.
-In this work, we describe the SEC algorithm, which attempts to learn a distribution over programs by incrementally finding program components that commonly help to solve problems in a given domain, and we show prelimarily that SEC is able to discover the graphical concepts that underlie many of the common graphical model structures.
+On the one hand, designing a good model requires a great deal of domain-specific knowledge and expertise, which may be hard to come by.
+On the other hand, searching the space of probabilistic models automatically is notoriously difficult.
+However, when researchers do set out to design graphical models, they draw on abstractions such as trees, chains, grids and plates to constrain and direct the kinds of models they produce.
+Thus, if we want to produce machine learning tools that aid in this task, we want tools that can learn these kinds of relevant abstractions.
+One natural way to think of these abstractions is as programs that take sets of random variables and produce graphical models that relate them.
+In this work, we describe the SEC algorithm, which attempts to learn distributions over such programs by incrementally finding program components that commonly help solve problems in a given domain.
+We discuss the behavior of the SEC algorithm for the task of graphical model structure learning.
+We also describe SEC's behavior on the dissimilar problem of automatically constructing tall and physically stable towers of blocks.
 \end{abstract}
@@
 \subsection{Program Representation}
+Discuss combinatory logic and the adaptor-grammar-esque distribution we use.
 
-Following~\cite{DBLP:conf/ijcai/DechterMAT13} and~\cite{DBLP:conf/icml/LiangJK10}, we use a
-simply typed combinatory calculus -- a variable-free subset of the
-polymorphic simply-typed lambda calculus~\cite{DBLP:books/daglib/0005958} -- as our
-program representation language.
-The simplicity the combinatory calculus makes describing a distribution over its expressions particularly simple.
-A more detailed explanation and discussion of this program representation can be found in the papers cited above.
-
-\section{Application to Structure Learning}
-Learning the structure of graphical models is a difficult problem - not only for machines, but also for humans, where expert knowledge is often required.
-In practice, the sorts of graphical models that humans design, such as Hidden Markov Models, phylogenetic trees, topic models, and Ising models, exhibit certain symmetries and recursive structures that are amenable to synthesis by programs.
-Could an algorithm, such as SEC, automatically learn these structures?
-
-\section{Application to Graphical Model Structure Learning}
+\section{Learning to Construct Graphical Models}
 
@@
 
-\subsection{Graph Combinators}
+In practice, the sorts of graphical models that humans design, such as Hidden Markov Models, phylogenetic trees, topic models, and Ising models, exhibit certain symmetries and recursive structures that might be amenable to synthesis by programs.
+We use our algorithm to learn the motifs and design patterns that underly some popular graphical models.
+Concretely, our algorithm induces programs that take as input the visible nodes of an undirected graphical model and introduces new latent nodes and edges.
 
-Diagrams illustrating the graph combinators. An example program that makes an Ising model.
+\subsection{Experiments}
+
+The SEC algorithm was provided a set of related tasks, each of which involved the construction of a program that built undirected graphical models.
+Each program synthesis task included a set of input/output test cases; the likelihood of a task, given a program, was defined as the fraction of test cases passed.
+Example input/output pairs of the target programs are shown in figure (\ref{fig:tasks}).
+
+The initial library of primitive combinators contains four binary operations upon undirected graphical models.
+Each of these operations we introduce, called an \emph{undirected union}, serves to combine two smaller models while also introducing new edges.
+%Using this representation, many models may be expressed in a compositional fashion when combined with higher-order combinators taken from functional programming languages, such as \emph{map} and \emph{fold}.
+The operation of these four undirected unions is diagrammed in figure (\ref{fig:graphcomb}).
+%In this representation of graphs, the vertexes are given unique indexes.
+%Some of the undirected unions exploit these indexes by adding edges as a function of index.
+
+\begin{figure}
+\begin{minipage}[c][8cm][t]{.4\textwidth}
+  \vspace*{\fill}
+  \includegraphics[width=\linewidth]{./figures/tasks.pdf}
+  \caption{Example input/output pairs of learned programs. Each induced program takes as input a list of visible nodes and produces an undirected graphical model.}
+  \label{fig:tasks}
+\end{minipage}%
+\begin{minipage}[c][8cm][t]{.3\textwidth}
+  \vspace*{\fill}
+  \fbox{\includegraphics[width=\linewidth]{./figures/GraphCombinators.pdf}}
+  \caption{The operation of the undirected unions.
+Union: Computes union of vertexes and edges.
+UnionC: Adds edges between corresponding vertexes.
+UnionHT: Adds edge between first (head) vertex and last (tail) vertex.
+UnionD: Preserves uniqueness of distinctly indexed vertexes.}
+  \label{fig:graphcomb}\par\vfill
+\end{minipage}
+\end{figure}
+
+\subsection{Results}
 
@@
 
-\begin{figure}
-\begin{minipage}[c][8cm][t]{.4\textwidth}
-  \vspace*{\fill}
-  \includegraphics[width=\linewidth]{./figures/tasks.pdf}
-  \caption{test figure one}
-  \label{fig:test1}
-\end{minipage}%
-\begin{minipage}[c][8cm][t]{.3\textwidth}
-  \vspace*{\fill}
-  \fbox{\includegraphics[width=\linewidth]{./figures/GraphCombinators.pdf}}
-  \caption{test figure two}
-  \label{fig:test2}\par\vfill
-  % \includegraphics[width=5cm,height=4.5cm]{im
-  \caption{test figure three}
-  \label{fig:test3}
-\end{minipage}
-\end{figure}
+\section{Building Towers in Blocks World}
+
+How could an agent, endowed with little prior knowledge, learn to build tall, robust towers out of toy blocks?
+Although the space of possible towers is large enough to make exhaustive search intractible, the useful towers lie within a much smaller space.
+Most useful towers use compositions of stereotyped structures, such as arches.
+
+\subsection{The Blocks World Domain}
+
+Tower representation, objective function
+
+\subsection{Learned Tower programs}
+
+Picture of tower. Grammar that produced that tower. Note lack of abstraction in grammar; speculate as to reasons.
 
@@
 
+\section{Submission of papers to NIPS 2013}
+
+NIPS requires electronic submissions.  The electronic submission site is  
+\begin{center}
+   \url{http://papers.nips.cc}
+\end{center}
+
+Please read carefully the
+instructions below, and follow them faithfully.
+\subsection{Style}
+
+Papers to be submitted to NIPS 2013 must be prepared according to the
+instructions presented here. Papers may be only up to eight pages long,
+including figures. Since 2009 an additional ninth page \textit{containing only
+cited references} is allowed. Papers that exceed nine pages will not be
+reviewed, or in any other way considered for presentation at the conference.
+%This is a strict upper bound.
+
+Please note that this year we have introduced automatic line number generation
+into the style file (for \LaTeXe and Word versions). This is to help reviewers
+refer to specific lines of the paper when they make their comments. Please do
+NOT refer to these line numbers in your paper as they will be removed from the
+style file for the final version of accepted papers.
+
+The margins in 2013 are the same as since 2007, which allow for $\approx 15\%$
+more words in the paper compared to earlier years. We are also again using 
+double-blind reviewing. Both of these require the use of new style files.
+
+Authors are required to use the NIPS \LaTeX{} style files obtainable at the
+NIPS website as indicated below. Please make sure you use the current files and
+not previous versions. Tweaking the style files may be grounds for rejection.
+
+%% \subsection{Double-blind reviewing}
+
+%% This year we are doing double-blind reviewing: the reviewers will not know 
+%% who the authors of the paper are. For submission, the NIPS style file will 
+%% automatically anonymize the author list at the beginning of the paper.
+
+%% Please write your paper in such a way to preserve anonymity. Refer to
+%% previous work by the author(s) in the third person, rather than first
+%% person. Do not provide Web links to supporting material at an identifiable
+%% web site.
+
+%%\subsection{Electronic submission}
+%%
+%% \textbf{THE SUBMISSION DEADLINE IS MAY 31st, 2013. SUBMISSIONS MUST BE LOGGED BY
+%% 23:00, MAY 31st, 2013, UNIVERSAL TIME}
+
+%% You must enter your submission in the electronic submission form available at
+%% the NIPS website listed above. You will be asked to enter paper title, name of
+%% all authors, keyword(s), and data about the contact
+%% author (name, full address, telephone, fax, and email). You will need to
+%% upload an electronic (postscript or pdf) version of your paper.
+
+%% You can upload more than one version of your paper, until the
+%% submission deadline. We strongly recommended uploading your paper in
+%% advance of the deadline, so you can avoid last-minute server congestion.
+%%
+%% Note that your submission is only valid if you get an e-mail
+%% confirmation from the server. If you do not get such an e-mail, please
+%% try uploading again.
+
+\subsection{Retrieval of style files}
+
+The style files for NIPS and other conference information are available on the World Wide Web at
+\begin{center}
+   \url{http://www.nips.cc/}
+\end{center}
+The file \verb+nips2013.pdf+ contains these 
+instructions and illustrates the
+various formatting requirements your NIPS paper must satisfy. \LaTeX{}
+users can choose between two style files:
+\verb+nips11submit_09.sty+ (to be used with \LaTeX{} version 2.09) and
+\verb+nips11submit_e.sty+ (to be used with \LaTeX{}2e). The file
+\verb+nips2013.tex+ may be used as a ``shell'' for writing your paper. All you
+have to do is replace the author, title, abstract, and text of the paper with
+your own. The file
+\verb+nips2013.rtf+ is provided as a shell for MS Word users.
+
+The formatting instructions contained in these style files are summarized in
+sections \ref{gen_inst}, \ref{headings}, and \ref{others} below.
+
+%% \subsection{Keywords for paper submission}
+%% Your NIPS paper can be submitted with any of the following keywords (more than one keyword is possible for each paper):
+
+%% \begin{verbatim}
+%% Bioinformatics
+%% Biological Vision
+%% Brain Imaging and Brain Computer Interfacing
+%% Clustering
+%% Cognitive Science
+%% Control and Reinforcement Learning
+%% Dimensionality Reduction and Manifolds
+%% Feature Selection
+%% Gaussian Processes
+%% Graphical Models
+%% Hardware Technologies
+%% Kernels
+%% Learning Theory
+%% Machine Vision
+%% Margins and Boosting
+%% Neural Networks
+%% Neuroscience
+%% Other Algorithms and Architectures
+%% Other Applications
+%% Semi-supervised Learning
+%% Speech and Signal Processing
+%% Text and Language Applications
+
+%% \end{verbatim}
+
+\section{General formatting instructions}
+\label{gen_inst}
+
+The text must be confined within a rectangle 5.5~inches (33~picas) wide and
+9~inches (54~picas) long. The left margin is 1.5~inch (9~picas).
+Use 10~point type with a vertical spacing of 11~points. Times New Roman is the
+preferred typeface throughout. Paragraphs are separated by 1/2~line space,
+with no indentation.
+
+Paper title is 17~point, initial caps/lower case, bold, centered between
+2~horizontal rules. Top rule is 4~points thick and bottom rule is 1~point
+thick. Allow 1/4~inch space above and below title to rules. All pages should
+start at 1~inch (6~picas) from the top of the page.
+
+%The version of the paper submitted for review should have ``Anonymous Author(s)'' as the author of the paper.
+
+For the final version, authors' names are
+set in boldface, and each name is centered above the corresponding
+address. The lead author's name is to be listed first (left-most), and
+the co-authors' names (if different address) are set to follow. If
+there is only one co-author, list both author and co-author side by side.
+
+Please pay special attention to the instructions in section \ref{others}
+regarding figures, tables, acknowledgments, and references.
+
+\section{Headings: first level}
+\label{headings}
+
+First level headings are lower case (except for first word and proper nouns),
+flush left, bold and in point size 12. One line space before the first level
+heading and 1/2~line space after the first level heading.
+
+\subsection{Headings: second level}
+
+Second level headings are lower case (except for first word and proper nouns),
+flush left, bold and in point size 10. One line space before the second level
+heading and 1/2~line space after the second level heading.
+
+\subsubsection{Headings: third level}
+
+Third level headings are lower case (except for first word and proper nouns),
+flush left, bold and in point size 10. One line space before the third level
+heading and 1/2~line space after the third level heading.
+
+\section{Citations, figures, tables, references}
+\label{others}
+
+These instructions apply to everyone, regardless of the formatter being used.
+
+\subsection{Citations within the text}
+
+Citations within the text should be numbered consecutively. The corresponding
+number is to appear enclosed in square brackets, such as [1] or [2]-[5]. The
+corresponding references are to be listed in the same order at the end of the
+paper, in the \textbf{References} section. (Note: the standard
+\textsc{Bib\TeX} style \texttt{unsrt} produces this.) As to the format of the
+references themselves, any style is acceptable as long as it is used
+consistently.
+
+As submission is double blind, refer to your own published work in the 
+third person. That is, use ``In the previous work of Jones et al.\ [4]'',
+not ``In our previous work [4]''. If you cite your other papers that
+are not widely available (e.g.\ a journal paper under review), use
+anonymous author names in the citation, e.g.\ an author of the
+form ``A.\ Anonymous''.
+
+\subsection{Footnotes}
+
+Indicate footnotes with a number\footnote{Sample of the first footnote} in the
+text. Place the footnotes at the bottom of the page on which they appear.
+Precede the footnote with a horizontal rule of 2~inches
+(12~picas).\footnote{Sample of the second footnote}
+
+\subsection{Figures}
+
+All artwork must be neat, clean, and legible. Lines should be dark
+enough for purposes of reproduction; art work should not be
+hand-drawn. The figure number and caption always appear after the
+figure. Place one line space before the figure caption, and one line
+space after the figure. The figure caption is lower case (except for
+first word and proper nouns); figures are numbered consecutively.
+
+Make sure the figure caption does not get separated from the figure.
+Leave sufficient space to avoid splitting the figure and figure caption.
+
+You may use color figures. 
+However, it is best for the
+figure captions and the paper body to make sense if the paper is printed
+either in black/white or in color.
+\begin{figure}[h]
+\begin{center}
+%\framebox[4.0in]{$\;$}
+\fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
+\end{center}
+\caption{Sample figure caption.}
+\end{figure}
+
+\subsection{Tables}
+
+All tables must be centered, neat, clean and legible. Do not use hand-drawn
+tables. The table number and title always appear before the table. See
+Table~\ref{sample-table}.
+
+Place one line space before the table title, one line space after the table
+title, and one line space after the table. The table title must be lower case
+(except for first word and proper nouns); tables are numbered consecutively.
+
+\begin{table}[t]
+\caption{Sample table title}
+\label{sample-table}
+\begin{center}
+\begin{tabular}{ll}
+\multicolumn{1}{c}{\bf PART}  &\multicolumn{1}{c}{\bf DESCRIPTION}
+\\ \hline \\
+Dendrite         &Input terminal \\
+Axon             &Output terminal \\
+Soma             &Cell body (contains cell nucleus) \\
+\end{tabular}
+\end{center}
+\end{table}
+
+\section{Final instructions}
+Do not change any aspects of the formatting parameters in the style files.
+In particular, do not modify the width or length of the rectangle the text
+should fit into, and do not change font sizes (except perhaps in the
+\textbf{References} section; see below). Please note that pages should be
+numbered.
+
+\section{Preparing PostScript or PDF files}
+
+Please prepare PostScript or PDF files with paper size ``US Letter'', and
+not, for example, ``A4''. The -t
+letter option on dvips will produce US Letter files.
+
+Fonts were the main cause of problems in the past years. Your PDF file must
+only contain Type 1 or Embedded TrueType fonts. Here are a few instructions
+to achieve this.
+=======
+>>>>>>> 0c3d0be073ba9aaf201c99bcf3dc620f5bb9f7cb
+
 \bibliographystyle{plain}

### 2013-10-23 src/Strappy/UGM.hs

@@ -1,6 +1,6 @@
 
-module Strappy.UGM where
---module Main where
+--module Strappy.UGM where
+module Main where
 
 
 import Strappy.EM
@@ -116,8 +116,8 @@ ugmRename2Disjoint u1@(UGM vs1 es1) u2@(UGM vs2 es2) =
     in
         (u1, buildSub (map getVertexID $ filter isLatent vs2) u2)
 
-ugmExplode :: UGM -> [UGM]
-ugmExplode (UGM vs _) = map (\v -> UGM [v] []) (sort vs)
+ugmExplode :: UGM -> UGM
+ugmExplode (UGM vs _) = UGM vs []
 
 -- | Test cases
 ugmLatentify :: UGM -> UGM
@@ -193,7 +193,7 @@ cUnion = mkTerm "union" (tUGM ->- tUGM ->- tUGM) $ ugmUnion
 cUnionC = mkTerm "unionC" (tUGM ->- tUGM ->- tUGM) $ ugmUnionCorrespondence
 cUnionHT = mkTerm "unionHT" (tUGM ->- tUGM ->- tUGM) $ ugmUnionHeadTail
 cUnionD = mkTerm "unionD" (tUGM ->- tUGM ->- tUGM) $ ugmUnionDisjoint
-cExplode = mkTerm "explode" (tUGM ->- tList tUGM) $ ugmExplode
+cExplode = mkTerm "explode" (tUGM ->- tUGM) $ ugmExplode
 cLatentify = mkTerm "latentify" (tUGM ->- tUGM) $ ugmLatentify
 cLatent = mkTerm "latent" tUGM $ singletonUGM $ Latent 0
 cNull = mkTerm "null" tUGM $ nullUGM
@@ -243,10 +243,10 @@ cylTask = makeUGMTask "cylinder" (tList tUGM ->- tUGM)
              (isingRow 2 6, ugmCylinder (isingRow 2 6))]
 deepCylTask :: EMTask
 deepCylTask = makeUGMTask "deepCylinder" (tList (tList tUGM) ->- tUGM)
-            [(isingRow 1 1, ugmDeepCylinder (isingRow 1 1)),
-             (isingRow 1 2, ugmDeepCylinder (isingRow 1 2)),
-             (isingRow 1 3, ugmDeepCylinder (isingRow 1 3)),
-             (isingRow 2 6, ugmDeepCylinder (isingRow 2 6))]
[…]

### 2013-10-23 drafts/nips/nips2013.tex

--- 2013-10-23 drafts/nips/nips2013.tex
@@
 In doing so, they draw on abstractions such as trees, chains, grids and plates to constrain and direct the kinds of models they produce.
-This suggests that before we ask machine learning algorithms to discover parsimonious models of new domains, we should develop techniques that enable our algorithms to automatically learn these ``graphical conepts'' in much the same way that researchers themselves do, by seeing examples in the literature.
+This suggests that before we ask machine learning algorithms to discover parsimonious models of new domains, we should develop techniques that enable our algorithms to automatically learn these ``graphical concepts'' in much the same way that researchers themselves do, by seeing examples in the literature.
 One natural way to think of these graphical concepts is as programs that take sets of random variables and produce graphical models that relate them.
@@
 The SEC algorithm takes a library of primitives and a set of related tasks to solve; its goal is to produce a set of programs that solve these tasks.
-Like the EC algorithm~\cite{DBLP:conf/ijcai/DechterMAT13} (of which it an extension), the SEC algorithm maintains a distribution $\mathcal{D}$ over programs and this distribution helps guide its search to relevant parts of of the space of programs.
+Like the EC algorithm~\cite{DBLP:conf/ijcai/DechterMAT13}, of which it an extension, the SEC algorithm maintains a distribution $\mathcal{D}$ over programs and this distribution helps guide its search to relevant parts of of the space of programs.
 In each iteration, SEC does two things:
@@
 
-\section{Application to Structure Learning}
-Learning the structure of graphical models is a difficult problem - not only for machines, but also for humans, where expert knowledge is often required.
-In practice, the sorts of graphical models that humans design, such as Hidden Markov Models, phylogenetic trees, topic models, and Ising models, exhibit certain symmetries and recursive structures that are amenable to synthesis by programs.
-Could an algorithm, such as SEC, automatically learn these structures?
-
-\section{Application to Graphical Model Structure Learning}
+\section{Learning to Construct Graphical Models}
 
@@
 
-\subsection{Graph Combinators}
+In practice, the sorts of graphical models that humans design, such as Hidden Markov Models, phylogenetic trees, topic models, and Ising models, exhibit certain symmetries and recursive structures that might be amenable to synthesis by programs.
+We use our algorithm to learn the motifs and design patterns that underly certain popular graphical models.
+Concretely, our algorithm induces programs that take as input the visible nodes of an undirected graphical model and introduces new latent nodes and edges.
 
-Diagrams illustrating the graph combinators. An example program that makes an Ising model.
+\subsection{Experiments}
 
-\subsection{Learning Graphical Models}
+The SEC algorithm was provided a set of related tasks, each of which involved the construction of a program that built undirected graphical models.
+Each program synthesis task included a set of input/output test cases; the likelihood of a task, given a program, was defined as the fraction of test cases passed.
+Example input/output pairs of the target programs are shown in figure (\ref{fig:tasks}).
 
-Show the actual results for our test cases: ``density estimation of the space of interesting graphical models.''
-In future work, could be used as high-level moves in the search for a good structure.
-Connect to, eg, Charles Kemp's work.
+The initial library of primitive combinators contains four binary operations upon undirected graphical models.
+Each of these operations we introduce, called an \emph{undirected union}, serves to combine two smaller models while also introducing new edges.
+%Using this representation, many models may be expressed in a compositional fashion when combined with higher-order combinators taken from functional programming languages, such as \emph{map} and \emph{fold}.
+The operation of these four undirected unions is diagrammed in figure (\ref{fig:graphcomb}).
+%In this representation of graphs, the vertexes are given unique indexes.
+%Some of the undirected unions exploit these indexes by adding edges as a function of index.
 
@@
   \includegraphics[width=\linewidth]{./figures/tasks.pdf}
-  \caption{test figure one}
-  \label{fig:test1}
+  \caption{Example input/output pairs of learned programs. Each induced program takes as input a list of visible nodes and produces an undirected graphical model.}
+  \label{fig:tasks}
 \end{minipage}%
@@
   \fbox{\includegraphics[width=\linewidth]{./figures/GraphCombinators.pdf}}
-  \caption{test figure two}
-  \label{fig:test2}\par\vfill
-  % \includegraphics[width=5cm,height=4.5cm]{im
-  \caption{test figure three}
-  \label{fig:test3}
+  \caption{The operation of the undirected unions.
+Union: Computes union of vertexes and edges.
+UnionC: Adds edges between corresponding vertexes.
+UnionHT: Adds edge between first (head) vertex and last (tail) vertex.
+UnionD: Preserves uniqueness of distinctly indexed vertexes.}
+  \label{fig:graphcomb}\par\vfill
 \end{minipage}
 \end{figure}
+
+\subsection{Results}
+
+In future work, could be used as high-level moves in the search for a good structure.
+Connect to, eg, Charles Kemp's work.
 

### 2013-10-23 drafts/nips/nips2013.tex
summary data and writeup

@@
 In this paper, we describe our approach to learning over structured representations as a search over the space of programs in a functional programming language.
-We characterize learning in this space as incrementally developing a weighted library of program abstractions, i.e.
+We characterize learning in this space as incrementally building a weighted library of program abstractions, i.e.
 a collection of subroutines, that are commonly used to solve problems in a given domain.
@@
 The SEC algorithm takes a library of primitives and a set of related tasks to solve; its goal is to produce a set of programs that solve these tasks.
-Like the EC algorithm~\cite{DBLP:conf/ijcai/DechterMAT13}, of which it an extension, the SEC algorithm maintains a distribution $\mathcal{D}$ over programs and this distribution helps guide its search to relevant parts of of the space of programs.
+Like the EC algorithm~\cite{DBLP:conf/ijcai/DechterMAT13}, of which it an extension, the SEC algorithm maintains a distribution $\mathcal{D}$ over programs.
+This distribution helps guide its search to relevant parts of of the space of programs.
 In each iteration, SEC does two things:
 \begin{itemize}
-	\item first, it uses $\mathcal{D}$ to explore the space of programs that are likely to be relevant for the tasks it has been given;
-	\item second, it compresses the solutions it found and uses this compression generate a new weighted library of primitives that defines an updated $\mathcal{D}$.
+	\item First, it uses $\mathcal{D}$ to explore the space of programs that are likely to be relevant for the tasks it has been given;
+	\item Second, it compresses the solutions it found and uses this compression generate a new weighted library of primitives that defines an updated $\mathcal{D}$.
 \end{itemize}
@@
 \subsection{Results}
+Initially, SEC is unable to find any solutions to the complicated graph structure tasks, such as the Ising Model or the Hidden Markov Model, even after enumerating 4 million programs through its heuristic search.
+However, it is able to find programs that generate simpler graphical models, such as Markov chains.
+By compressing the solutions to these simpler problems, it modifies its search procedure and, in subsequent iterations, learns more and more successful programs.
+Shown in figure (\ref{TODO}) are the learning curves for SEC, represented both as the number of solved tasks per iteration and as the size of the grammar per iteration.
 
-In future work, could be used as high-level moves in the search for a good structure.
-Connect to, eg, Charles Kemp's work.
+What grammars does SEC learn, and how do they enable the bootstrap learning of more complicated programs?
+The simplest example is that of how the Markov chain bootstraps the learning of the Ising model.
+Most of the easy tasks, such as the Markov chain, can be solved with the right choice of one higher-order function and one undirected union.
+After one iteration, the program fragment \verb|(foldr unionC)|, used in the Markov chain program, is incorporated in to the grammar.
+On the second iteration, the Ising model is learned using this fragment, producing the program \verb|(((S B) map) ((foldr unionC) null))|.
+
+%In future work, could be used as high-level moves in the search for a good structure.
 

### 2013-10-23 src/Strappy/UGM.hs

@@ -1,6 +1,6 @@
 
---module Strappy.UGM where
-module Main where
+module Strappy.UGM where
+--module Main where
 
 
 import Strappy.EM
@@ -123,7 +123,7 @@ ugmExplode (UGM vs _) = UGM vs []
 ugmLatentify :: UGM -> UGM
 ugmLatentify (UGM vs es) =
     let l (Visible x) = Latent x
-        l x = x
+        l (Latent x) = Visible x
         e (UEdge v v') = mkUEdge (l v) (l v')
     in UGM (nub $ map l vs) (nub $ map e es)
 

### 2013-10-23 drafts/nips/nips2013.tex
More edits.

--- 2013-10-23 drafts/nips/nips2013.tex
@@
 In doing so, they draw on abstractions such as trees, chains, grids and plates to constrain and direct the kinds of models they produce.
-This suggests that before we ask machine learning algorithms to discover parsimonious models of new domains, we should develop techniques that enable our algorithms to automatically learn these ``graphical concepts'' in much the same way that researchers themselves do, by seeing examples in the literature.
+This suggests that before we ask machine learning algorithms to discover parsimonious models of new domains, we should develop techniques that enable our algorithms to automatically learn these ``graphical conepts'' in much the same way that researchers themselves do, by seeing examples in the literature.
 One natural way to think of these graphical concepts is as programs that take sets of random variables and produce graphical models that relate them.
@@
 The SEC algorithm takes a library of primitives and a set of related tasks to solve; its goal is to produce a set of programs that solve these tasks.
-Like the EC algorithm~\cite{DBLP:conf/ijcai/DechterMAT13}, of which it an extension, the SEC algorithm maintains a distribution $\mathcal{D}$ over programs and this distribution helps guide its search to relevant parts of of the space of programs.
+Like the EC algorithm~\cite{DBLP:conf/ijcai/DechterMAT13} (of which it an extension), the SEC algorithm maintains a distribution $\mathcal{D}$ over programs and this distribution helps guide its search to relevant parts of of the space of programs.
 In each iteration, SEC does two things:
@@
 \begin{equation}
-P(G,\{\rho_n\},\{t_n\}) = \frac{1}{L^N} P(G) \prod_n P(t_n | \rho_n) \prod_i P(e^i_n | G)
+P(G,\{\rho_n\},\{t_n\}) = \frac{1}{L^N} P(G) \prod_n P(t_n | \rho_n) P(\rho_n | G),
 \label{eq:joint}
 \end{equation}
+where $P(\rho_n | G) = \prod_i P(e^i_n | G)$.
 From equation (\ref{eq:joint}), the EM updates are
 \begin{eqnarray}
-q(\rho_n) &\propto& P(t_n | \rho_n) \prod_i P(e^i_n | G^{old})\\
+q(\rho_n) &\propto& P(t_n | \rho_n) P(\rho_n | G^{\text{old}})\\
 \label{eq:qdist}
-G^{new} &=& \operatorname{argmin}_G \left( -\ln P(G) -
+G^{\text{new}} &=& \operatorname{argmin}_G \left( -\ln P(G) -
 \sum_n 
@@
 \end{eqnarray}
-Exact computation of $q(\cdot)$ in equation (\ref{eq:qdist}) would involve marginalizing over the space of all programs.
+Exact computation of these updates requires summing over the infinite space of programs that can be generated from the grammar. To approximate this, we can attempt to evaluate 
 Instead, we perform a heuristic search over the space of programs, with the goal of finding those programs for which $q(\cdot)$ is high.
@@
 
-\section{Learning to Construct Graphical Models}
+\section{Application to Structure Learning}
+Learning the structure of graphical models is a difficult problem - not only for machines, but also for humans, where expert knowledge is often required.
+In practice, the sorts of graphical models that humans design, such as Hidden Markov Models, phylogenetic trees, topic models, and Ising models, exhibit certain symmetries and recursive structures that are amenable to synthesis by programs.
+Could an algorithm, such as SEC, automatically learn these structures?
+
+\section{Application to Graphical Model Structure Learning}
 
@@
 
-In practice, the sorts of graphical models that humans design, such as Hidden Markov Models, phylogenetic trees, topic models, and Ising models, exhibit certain symmetries and recursive structures that might be amenable to synthesis by programs.
-We use our algorithm to learn the motifs and design patterns that underly certain popular graphical models.
-Concretely, our algorithm induces programs that take as input the visible nodes of an undirected graphical model and introduces new latent nodes and edges.
+\subsection{Graph Combinators}
 
-\subsection{Experiments}
+Diagrams illustrating the graph combinators. An example program that makes an Ising model.
 
-The SEC algorithm was provided a set of related tasks, each of which involved the construction of a program that built undirected graphical models.
-Each program synthesis task included a set of input/output test cases; the likelihood of a task, given a program, was defined as the fraction of test cases passed.
-Example input/output pairs of the target programs are shown in figure (\ref{fig:tasks}).
+\subsection{Learning Graphical Models}
 
-The initial library of primitive combinators contains four binary operations upon undirected graphical models.
-Each of these operations we introduce, called an \emph{undirected union}, serves to combine two smaller models while also introducing new edges.
-%Using this representation, many models may be expressed in a compositional fashion when combined with higher-order combinators taken from functional programming languages, such as \emph{map} and \emph{fold}.
-The operation of these four undirected unions is diagrammed in figure (\ref{fig:graphcomb}).
-%In this representation of graphs, the vertexes are given unique indexes.
-%Some of the undirected unions exploit these indexes by adding edges as a function of index.
+Show the actual results for our test cases: ``density estimation of the space of interesting graphical models.''
+In future work, could be used as high-level moves in the search for a good structure.
+Connect to, eg, Charles Kemp's work.
 
@@
   \includegraphics[width=\linewidth]{./figures/tasks.pdf}
-  \caption{Example input/output pairs of learned programs. Each induced program takes as input a list of visible nodes and produces an undirected graphical model.}
-  \label{fig:tasks}
+  \caption{test figure one}
+  \label{fig:test1}
 \end{minipage}%
@@
   \fbox{\includegraphics[width=\linewidth]{./figures/GraphCombinators.pdf}}
-  \caption{The operation of the undirected unions.
-Union: Computes union of vertexes and edges.
-UnionC: Adds edges between corresponding vertexes.
-UnionHT: Adds edge between first (head) vertex and last (tail) vertex.
-UnionD: Preserves uniqueness of distinctly indexed vertexes.}
-  \label{fig:graphcomb}\par\vfill
+  \caption{test figure two}
+  \label{fig:test2}\par\vfill
+  % \includegraphics[width=5cm,height=4.5cm]{im
+  \caption{test figure three}
+  \label{fig:test3}
 \end{minipage}
 \end{figure}
-
-\subsection{Results}
-
-In future work, could be used as high-level moves in the search for a good structure.
-Connect to, eg, Charles Kemp's work.
 

### 2013-10-23 drafts/nips/nips2013.tex
Added new figure and lots of edits.

@@
 
-\begin{comment}
+% \begin{comment}
 \author{
-David S.~Hippocampus\thanks{ Use footnote for providing further information
-about author (webpage, alternative address)---\emph{not} for acknowledging
-funding agencies.} \\
-Department of Computer Science\\
-Cranberry-Lemon University\\
-Pittsburgh, PA 15213 \\
-\texttt{hippo@cs.cranberry-lemon.edu} \\
+Kevin Ellis\thanks{Questions and correspondence should be sent to 
+\href{mailto:ellisk@mit.edu}{ellisk@mit.edu}.} \\
+Department of Brain \& Cognitive Sciences\\
+MIT\\
 \And
-Coauthor \\
-Affiliation \\
-Address \\
-\texttt{email} \\
+Eyal Dechter \\
+Department of Brain \& Cognitive Sciences\\
+MIT\\
 \AND
-Coauthor \\
-Affiliation \\
-Address \\
-\texttt{email} \\
+Ryan P. Adams \\
+School of Engineering \& Applied Sciences \\
+Harvard University\\
 \And
-Coauthor \\
-Affiliation \\
-Address \\
-\texttt{email} \\
-\And
-Coauthor \\
-Affiliation \\
-Address \\
-\texttt{email} \\
-(if needed)\\
+Joshua B. Tenenebaum \\
+Department of Brain \& Cognitive Sciences\\
+MIT\\
 }
-\end{comment}
+% \end{comment}
 % The \author macro works with any number of authors. There are two commands
@@
 
-%\nipsfinalcopy % Uncomment for camera-ready version
+\nipsfinalcopy % Uncomment for camera-ready version
 
@@
 
-To address this problem, many structured representations have been proposed in the machine learning and AI literatures: learning over graphs \cite{}, grammars\cite{}, Horn clauses\cite{}, relations{}, etc., has been investigated.
+One example of a structured representation we would like to be able to learn is the probabilistic graphical model.
+Although graphical models are widely used in machine learning to represent the factorizes structure of probability distributions, they are generally designed by researchers rather than by algorithms.
+This is especially true when the goal is to learn about latent variables underlying the observed data, because it is difficult to capture the knowledge which allows researchers to make reasonable assumptions about which latent variables exists and what they depend on.
+
+In particular, machine learning researchers make frequent use of graphical motifs and abstractions such as trees, chains, rings, grids, mixtures, and plates to constrain the space of graphical models they consider.
+But how do they learn that these are good motifs?
+How do they learn how to use them and combine them to design new models?
+
+One hypothesis is that these motifs and abstractions are best represented as programs that manipulate and compose graph operations.
+In this work, we discuss how one could learn these motifs and abstracts by finding programs that generate commonly occured graphical model types.
+In particular, we present the SEC algorithm, a general purpose multitask program induction algorithm, that attempts to solve many related tasks by generating a weighted library of useful subroutines that induces a distribution over the space of programs.
+We discuss how SEC can be used to extract from a set of training tasks the kind of abstract knowledge that machine learning researchers bring to the task of designing graphical models, and we speculate that learning such representations can assist in the difficult tasks of automatically finding good latent-variable graphical models for novel datasets.
+
+% To address this problem, many structured representations have been proposed in the machine learning and AI literatures: learning over graphs \cite{}, grammars\cite{}, Horn clauses\cite{}, relations{}, etc., has been investigated.
 These are representations that attempt to bridge the gap between the tractablity of learning and reasoning in smooth continuous spaces and the expressivity of general purpose programming languages.
 
-However, many of the most natural representations of structured domains are best described as programs and are difficult to fit into more restrictive representations.
+% However, many of the most natural representations of structured domains are best described as programs and are difficult to fit into more restrictive representations.
 To take an example directly from the machine learning literature, consider the plate notation~\cite{DBLP:journals/jair/Buntine94} commonly used to describe graphical models over collections of variables.
@@
 
-Suppose, then, that we want to create machine learning algorithms that can help discover good models for data in novel domains; that is, we want algorithms that discovers the structure of the latent graphical model underlying the data.
+% Suppose, then, that we want to create machine learning algorithms that can help discover good models for data in novel domains; that is, we want algorithms that discovers the structure of the latent graphical model underlying the data.
 This is a field of active research \
 
-In this paper, we describe our approach to learning over structured representations as a search over the space of programs in a functional programming language.
+% In this paper, we describe our approach to learning over structured representations as a search over the space of programs in a functional programming language.
 We characterize learning in this space as incrementally building a weighted library of program abstractions, i.e.
@@
 
-We will present here two domains of interest.
+% We will present here two domains of interest.
 First, continuing with the example above, we will describe how SEC can learn to efficiently explore the space of relevant graphical models for a given problem.
@@
 
-\lstset{
-	basicstyle=\scriptsize\ttfamily,
-    commentstyle=\color{white!40!black},
-    escapechar=+,
-    escapebegin=\color{white!40!black},
-    escapeend={},
-    keywordstyle=\bfseries\color{Periwinkle},
-    numbers=left,
-    numberstyle=\sffamily\tiny\color{gray},
-}
+% \lstset{
+% 	basicstyle=\scriptsize\ttfamily,
+%     commentstyle=\color{white!40!black},
+%     escapechar=+,
+%     escapebegin=\color{white!40!black},
+%     escapeend={},
+%     keywordstyle=\bfseries\color{Periwinkle},
+%     numbers=left,
+%     numberstyle=\sffamily\tiny\color{gray},
+% }
 
-\begin{figure}[h]
-\centering
-\begin{subfigure}{.5\textwidth}
-  \centering
-\includegraphics[width=.7\linewidth]{./figures/lda_plate_blei.pdf}
-  \caption{An example of typical plate notation use. \label{fig:lda_plate}}
+% \begin{figure}[h]
+% \centering
+% \begin{subfigure}{.5\textwidth}
+%   \centering
+% \includegraphics[width=.7\linewidth]{./figures/lda_plate_blei.pdf}
+%   \caption{An example of typical plate notation use. \label{fig:lda_plate}}
 
-\end{subfigure}%
-\begin{subfigure}{.4\textwidth}
-	\begin{lstlisting}[frame=single, numbers=none, xleftmargin=0pt]
-	for +$k$+ = 1:NumTopics
-	  +$\beta$+[k] +$\sim$+ Dirichlet(+$\eta$+)
-	for m = 1:NumDocuments
-	  +$\theta$+[m] +$\sim$+ Dirichlet(+$\alpha$+) 
-	  for n = 1:NumWords 
-	    z +$\sim$+ Categorical(+$\theta$+[m])
-	    w+$\sim$+ Categorical(+$\beta$+[z])
-	\end{lstlisting}
-  \caption{The same model as a program.}
-  \label{fig:lda_code}
-\end{subfigure}
-\caption{The popular Latent Dirichlet Allocation model in two forms.}
-\label{fig:lda}
-\end{figure}
+% \end{subfigure}%
+% \begin{subfigure}{.4\textwidth}
+% 	\begin{lstlisting}[frame=single, numbers=none, xleftmargin=0pt]
+% 	for +$k$+ = 1:NumTopics
+% 	  +$\beta$+[k] +$\sim$+ Dirichlet(+$\eta$+)
+% 	for m = 1:NumDocuments
+% 	  +$\theta$+[m] +$\sim$+ Dirichlet(+$\alpha$+) 
+% 	  for n = 1:NumWords 
+% 	    z +$\sim$+ Categorical(+$\theta$+[m])
+% 	    w+$\sim$+ Categorical(+$\beta$+[z])
+% 	\end{lstlisting}
+%   \caption{The same model as a program.}
+%   \label{fig:lda_code}
+% \end{subfigure}
+% \caption{The popular Latent Dirichlet Allocation model in two forms.}
+% \label{fig:lda}
+% \end{figure}
 
@@
 \subsection{Generative Model}
-The SEC algorithm can be expressed as MAP inference in a probabilistic generative model over the space of programs that captures two intuitions:
-\begin{itemize}
-\item{For a set of related tasks, useful programs are composed of useful subparts.}
-\item{Short programs are a priori more likely than long programs.}
-\end{itemize}
+The SEC algorithm can be expressed as MAP inference in a probabilistic generative model over the space of programs that captures two intuitions: 1) for a set of related tasks, useful programs are composed of useful subparts, and 2) short programs are a priori more likely than long programs.
+
+Our generative model assumes that the programs that solve a set of related tasks are themselves constructing by composing program fragments that are drawn from a common stochastic grammar.
+This grammar is itself drawn from a distribution over grammars that is biased towards grammars with smaller descriptions lengths.
+This generative model is shown in Figure~\ref{ref:gm}.
+
+\begin{figure}
 \begin{codebox}
@@
 \end{codebox}
-
-In short, this generative model assumes that the programs that solve a set of related tasks are themselves constructing by composing program fragments that are drawn from a common stochastic grammar.
-This grammar is itself drawn from a distribution over grammars that is biased towards grammars with smaller descriptions lengths.
+\caption{The generative model underlying the SEC algorithm. \label{ref:gm}}
+\end{figure}
 
 \subsection{Inference}
-We implement an iterative Expectation-Maximization algorithm that performs MAP estimation of the grammar, $G$, given the tasks, $\{t_n\}$.
+We implement an version of the EM algorithm~\cite{dempster:1977} for MAP estimation of the grammar, $G$, given the tasks, $\{t_n\}$.
+The hidden data in this case is the latent program that solves each task.
+
 The joint distribution factorizes as
@@
 \end{eqnarray}
+
 Exact computation of these updates requires summing over the infinite space of programs that can be generated from the grammar. To approximate this, we can attempt to evaluate 
@@
 In practice, the sorts of graphical models that humans design, such as Hidden Markov Models, phylogenetic trees, topic models, and Ising models, exhibit certain symmetries and recursive structures that might be amenable to synthesis by programs.
-We use our algorithm to learn the motifs and design patterns that underly certain popular graphical models.
-Concretely, our algorithm induces programs that take as input the visible nodes of an undirected graphical model and introduces new latent nodes and edges.
 
-\subsection{Experiments}
+\begin{figure}[h]
+% \centering
+\begin{minipage}[t]{.4\textwidth}
+% \vspace{0pt}
+% \begin{subfigure}{.45\textwidth}
+  % \vspace*{\fill}
+  % \centering
+  \includegraphics[width=\linewidth]{./figures/tasks.pdf}
+  \caption{Example input/output pairs of learned programs.
+Each induced program takes as input a list of observed nodes and produces an undirected graphical model.
+Here, we show each program with a small number of input nodes, but the learned programs transfer arbitrary length lists of visible nodes into the corresponding graphs.
+Labeled nodes correspond to observed variables; unlabeled nodes correspond to latent variables.}
+  \label{fig:tasks}
+\end{minipage}
+\begin{minipage}[t]{.5\textwidth}
+% \vspace{\fill}
+  \includegraphics[width=\linewidth]{./figures/GraphCombinators.pdf}
+  \caption{The operation of the undirected unions.
+Union: computes union of vertexes and edges.
+UnionC: adds edges between corresponding vertexes.
+UnionHT: adds edge between first (head) vertex and last (tail) vertex.
+UnionD: relabels input graph vertices so that they do not overlap and then computes union.}
+  \label{fig:graphcomb}\par\vfill
+% \end{subfigure}
+\end{minipage}
+\end{figure}
 
-The SEC algorithm was provided a set of related tasks, each of which involved the construction of a program that built undirected graphical models.
+\subsection{Experiment}
+
+We wanted to use the SEC algorithm to learn the common abstractions and motifs that underlie the graphical models that are commonly found in the machine learning literature.
+To that end, we supplied SEC with a set of twelve tasks, each of which corresponds to learning how to build a specific type of graphical model, with potentially latent vertices, given that it is provided with a list of observed verticies.
+A partial list of the tasks is shown in Figure~\ref{fig:tasks}.
+For example, the \textsc{Cylinder} task requires EC to build a program that takes as input at list of observed vertices (of any length) and produces a graph where one ring consists of the observed vertices and the other consists of corresponding latent variables.
+
+Tasks were specified via a set of input/output test cases, and we defined the likelihood of a program for a task to be the fraction of test cases it passed.
+
+The SEC algorithm was provided with an initial library of primitive  combinators, which included the basis combinators of the combinatory calculus, function list operations such as \textsc{Map} and \textsc{Foldr}/\textsc{Foldl}.
+We also included four binary operations on undirected graphical models, each of which computes some kind of union between two graphs.
+These graph operations are shown in Figure~\ref{fig:graphcomb}.
+
 Each program synthesis task included a set of input/output test cases; the likelihood of a task, given a program, was defined as the fraction of test cases passed.
@@
 
+\subsection{Results}
+
+We ran the SEC algorithm on the set of twelve tasks.
+The algorithm is parameterized by the frontier size (the number of unique subprograms $\{e_i\}$ enumerated from the library on each iteration).
+We ran the algorithm both for frontier sizes of 1000 and 5000.
+Note that because proposed solutions are compositions of these programs, the algorithm looks at many more potential solutions than the number of programs in the frontier.
+
+With a frontier size of 1000, the algorithm solves 7/12 tasks.
+With a frontier size of 5000, the algorithm solves 8/12 tasks.
+In each case, the algorithm returns solutions that partially satisfy the remaining tasks.
+As an example, we show in Figure~\ref{ref:cylinder} an instance of SEC's solution to the \textsc{Cylinder} task.
+
 \begin{figure}
-\begin{minipage}[c][8cm][t]{.4\textwidth}
-  \vspace*{\fill}
-  \includegraphics[width=\linewidth]{./figures/tasks.pdf}
-  \caption{Example input/output pairs of learned programs. Each induced program takes as input a list of visible nodes and produces an undirected graphical model.}
-  \label{fig:tasks}
-\end{minipage}%
-\begin{minipage}[c][8cm][t]{.3\textwidth}
-  \vspace*{\fill}
-  \fbox{\includegraphics[width=\linewidth]{./figures/GraphCombinators.pdf}}
-  \caption{The operation of the undirected unions.
-Union: Computes union of vertexes and edges.
-UnionC: Adds edges between corresponding vertexes.
-UnionHT: Adds edge between first (head) vertex and last (tail) vertex.
-UnionD: Preserves uniqueness of distinctly indexed vertexes.}
-  \label{fig:graphcomb}\par\vfill
-\end{minipage}
+\includegraphics[width=.5\linewidth]{./figures/cylinder.pdf}
+\caption{An example of how the \textsc{Cylinder} task is solved by SEC, with the basis combinators replaced by control flow arrows.
+The full program is written in the combinatory calculus below the image.
+\label{ref:cylinder}}
 \end{figure}
 
-\subsection{Results}
-Initially, SEC is unable to find any solutions to the complicated graph structure tasks, such as the Ising Model or the Hidden Markov Model, even after enumerating 4 million programs through its heuristic search.
-However, it is able to find programs that generate simpler graphical models, such as Markov chains.
-By compressing the solutions to these simpler problems, it modifies its search procedure and, in subsequent iterations, learns more and more successful programs.
-Shown in figure (\ref{TODO}) are the learning curves for SEC, represented both as the number of solved tasks per iteration and as the size of the grammar per iteration.
+% Initially, SEC is unable to find any solutions to the complicated graph structure tasks, such as the Ising Model or the Hidden Markov Model, even after enumerating 4 million programs through its heuristic search.
+% However, it is able to find programs that generate simpler graphical models, such as Markov chains.
+% By compressing the solutions to these simpler problems, it modifies its search procedure and, in subsequent iterations, learns more and more successful programs.
+% Shown in figure (\ref{TODO}) are the learning curves for SEC, represented both as the number of solved tasks per iteration and as the size of the grammar per iteration.
 
-What grammars does SEC learn, and how do they enable the bootstrap learning of more complicated programs?
-The simplest example is that of how the Markov chain bootstraps the learning of the Ising model.
-Most of the easy tasks, such as the Markov chain, can be solved with the right choice of one higher-order function and one undirected union.
-After one iteration, the program fragment \verb|(foldr unionC)|, used in the Markov chain program, is incorporated in to the grammar.
-On the second iteration, the Ising model is learned using this fragment, producing the program \verb|(((S B) map) ((foldr unionC) null))|.
+% What grammars does SEC learn, and how do they enable the bootstrap learning of more complicated programs?
+% The simplest example is that of how the Markov chain bootstraps the learning of the Ising model.
+% Most of the easy tasks, such as the Markov chain, can be solved with the right choice of one higher-order function and one undirected union.
+% After one iteration, the program fragment \verb|(foldr unionC)|, used in the Markov chain program, is incorporated in to the grammar.
+% On the second iteration, the Ising model is learned using this fragment, producing the program \verb|(((S B) map) ((foldr unionC) null))|.
 

### 2013-10-23 drafts/nips/nips2013.tex
Edited figures. Edited citations. Wrote discussion.

@@
 \usepackage{amsmath}
+\usepackage{natbib}
 
@@
 The SEC algorithm takes a library of primitives and a set of related tasks to solve; its goal is to produce a set of programs that solve these tasks.
-Like the EC algorithm~\cite{DBLP:conf/ijcai/DechterMAT13}, of which it an extension, the SEC algorithm maintains a distribution $\mathcal{D}$ over programs.
+Like the EC algorithm~\citep{DBLP:conf/ijcai/DechterMAT13}, of which it an extension, the SEC algorithm maintains a distribution $\mathcal{D}$ over programs.
 This distribution helps guide its search to relevant parts of of the space of programs.
@@
 \begin{codebox}
-\li \id{G} $\sim$ $P_G(\cdot)$ \Comment{Draw grammar from description-length prior}
+\id{G} $\sim$ $P_G(\cdot)$ \Comment{Draw grammar from description-length prior}
 \li \id{\ell_n} $\sim\mbox{Uniform}(1, L)$ \Comment{Draw number of subprograms in $n^{th}$ program}
@@
 
-Exact computation of these updates requires summing over the infinite space of programs that can be generated from the grammar. To approximate this, we can attempt to evaluate 
+Exact computation of These updates requires summing over the infinite space of programs that can be generated from the grammar. To approximate this, we can attempt to evaluate 
 Instead, we perform a heuristic search over the space of programs, with the goal of finding those programs for which $q(\cdot)$ is high.
@@
 A slight generalization of the Neville-Manning algorithm then permits tractible minimization of (\ref{eq:gmax}).
-Estimation of the production probabilities of $G$ is performed as in \cite{ijcai}.
+Estimation of the production probabilities of $G$ is performed as in~\citet{DBLP:conf/ijcai/DechterMAT13}.
 
@@
 
-Following~\cite{DBLP:conf/ijcai/DechterMAT13} and~\cite{DBLP:conf/icml/LiangJK10}, we use a
+Following~\citet{DBLP:conf/ijcai/DechterMAT13} and~\citet{DBLP:conf/icml/LiangJK10}, we use a
 simply typed combinatory calculus -- a variable-free subset of the
-polymorphic simply-typed lambda calculus~\cite{DBLP:books/daglib/0005958} -- as our
+polymorphic simply-typed lambda calculus~\citep{DBLP:books/daglib/0005958} -- as our
 program representation language.
@@
 
-Learning the structure of graphical models that capture the independencies underlying a dataset is an active area of research\cite{adams-wallach-ghahramani-2010a}\cite{ISI:000240797500002}\cite{ISI:000178037200004}\cite{ISI:A1995RX35400001}.
+Learning the structure of graphical models that capture the independencies underlying a dataset is an active area of research\cite{adams-wallach-ghahramani-2010a}\cite{ISI:000240797500002}\cite{ISI:000178037200004}\cite{DBLP:journals/corr/abs-1302-6815}.
 Graphical model structures enforce the conditional independences a modeler believes exists among latent and observed random variables\cite{DBLP:books/daglib/0066829}.
@@
 \begin{minipage}[t]{.4\textwidth}
-% \vspace{0pt}
-% \begin{subfigure}{.45\textwidth}
-  % \vspace*{\fill}
-  % \centering
+
   \includegraphics[width=\linewidth]{./figures/tasks.pdf}
@@
 \end{minipage}
+\hspace{0.2in}
 \begin{minipage}[t]{.5\textwidth}
-% \vspace{\fill}
   \includegraphics[width=\linewidth]{./figures/GraphCombinators.pdf}
@@
 \begin{figure}
-\includegraphics[width=.5\linewidth]{./figures/cylinder.pdf}
-\caption{An example of how the \textsc{Cylinder} task is solved by SEC, with the basis combinators replaced by control flow arrows.
-The full program is written in the combinatory calculus below the image.
-\label{ref:cylinder}}
+% \begin{minipage}[b]{.45\textwidth}
+\begin{subfigure}[b]{.45\textwidth}
+\includegraphics[width=\linewidth]{./figures/cylinder.pdf}
+\caption{The program generates a cylinder from nodes by generating two ring -- one latent, one visible -- and connecting them.}
+% \end{minipage}
+\end{subfigure}
+\hspace{0.2in}
+\begin{minipage}[b]{.5\textwidth}
+\begin{subfigure}[t]{\textwidth}
+\includegraphics[width=.5\linewidth]{./figures/cylinderCombinator.pdf}
+\caption{In the combinatory calculus. }
+\vspace{.3in}
+\end{subfigure}
+\begin{subfigure}[b]{\textwidth}
+\begin{lstlisting}[language=haskell, basicstyle=\small, escapechar=+]
++$\lambda$+vs. let vs' = foldr unionC null vs
+          vs'' = unionHT vs' (I vs')   
+          vs''' = flip vs''
+       in unionC vs'' vs'''
+\end{lstlisting}
+\vfill
+\caption{The combinatory calculus expression tranlated into \textsc{Hakell}-like syntax for readability.}
+% \end{minipage}
+\end{subfigure}
+\end{minipage}
+\caption{An example of solution that SEC finds to solve. \textsc{Cylinder}\label{ref:cylinder}}
 \end{figure}
@@
 
-\subsection{Relation to Prior Work}
+To be clear, what we have shown so far does not provide a way a way to learn the structure of a graphical model directly from observed data.
+Rather, this work suggests that we may be able to learn automatically an inductive bias that will subsequently make learning graphical model structures more efficient.
+Some of the programs that SEC includes in its library correspond to graph transformations, and we think that using this transformations as search operators will make structure learning of graphical models with latent variables a more tractable proposition than it currently is.
 
-MH sampling of programs, ILP, Kitzelman, graph grammars
+The notion that an appropriate inductive bias might help in the structure learning of graphical models is not new: \citep{DBLP:conf/icml/HonorioS10} shows that the inductive bias induced via multitask learning helps a search algorithm find accurate graphical model structures.
+But while work in this field has focused on how to evaluate a given graph, we hope to show that SEC can learn appropriate search operators directly from a set of training examples.
 
-\subsection{Limitations of SEC}
+We have presented the SEC algorithm in the context of learning graph structures, but we note that SEC is a general purpose multitask program induction algorithm system.
+For an excellent survey of the current state of inductive programming systems, see~\citep{DBLP:conf/aaip/Kitzelmann09}.
 
-Does not look at data. Does everything top-down. Blind to the utility of trying one program versus another (no active learning)
-
-\bibliographystyle{plain}
-\bibliography{eyal_bib}
+\bibliographystyle{plainnat}
+\bibliography{main}
 

### 2013-10-23 drafts/nips/nips2013.tex
writeup

--- 2013-10-23 drafts/nips/nips2013.tex
@@
 This distribution helps guide its search to relevant parts of of the space of programs.
-In each iteration, SEC does two things:
-\begin{itemize}
-	\item First, it uses $\mathcal{D}$ to explore the space of programs that are likely to be relevant for the tasks it has been given;
-	\item Second, it compresses the solutions it found and uses this compression generate a new weighted library of primitives that defines an updated $\mathcal{D}$.
-\end{itemize}
+In each iteration, SEC does two things: first, it uses $\mathcal{D}$ to explore the space of programs that are likely to be relevant for the tasks it has been given; and second, compresses the solutions it found and uses this compression generate a new weighted library of primitives that defines an updated $\mathcal{D}$.
 
@@
 
-Exact computation of these updates requires summing over the infinite space of programs that can be generated from the grammar. To approximate this, we can attempt to evaluate 
-Instead, we perform a heuristic search over the space of programs, with the goal of finding those programs for which $q(\cdot)$ is high.
-Rather than use mutation and crossover operators, as done in genetic programming, or random resampling of subprograms, as done in recent work in applying Metropolis-Hastings to program learning, we propose that the search moves should themselves be programs drawn from the grammar $G$.
-Concretely, in order to modify a program $e$ to produce new candidate programs, we draw a new function from our grammar, $G$, and apply that new function to $e$.
+Exact computation of the normalizing constant in equation (\ref{eq:qdist}), or the expectation in equation (\ref{eqn:gmax}), would require summing over the infinite space of programs that can be generated from the grammar.
+To approximate this expectation, we perform a heuristic search over the space of programs, with the goal of finding those programs for which $q(\cdot)$ is high.
+%Rather than use mutation and crossover operators, as done in genetic programming, or random resampling of subprograms, as done in recent work in applying Metropolis-Hastings to program learning, we propose that the search moves should themselves be programs drawn from the grammar $G$.
+Concretely, in order to modify a program $e$ to produce new candidate programs, we draw a new function from our grammar, $e'$, and apply that new function to produce the program $e'(e)$.
 Thus, we use the grammar not only as a distribution over programs of interest, but also as a distribution over programs that take as input other programs.
-This incremental construction of programs by repeatedly applying draws from the grammar mirrors the construction of programs in the generative model.
 
@@
 The description length of a program is approximated by the number of (not necessarily unique) program fragments in its smallest parse tree given the grammar.
-A slight generalization of the Neville-Manning algorithm then permits tractible minimization of (\ref{eq:gmax}).
+A slight generalization of the Neville-Manning algorithm then permits tractible approximate maximization of (\ref{eq:gmax}).
 Estimation of the production probabilities of $G$ is performed as in \cite{ijcai}.
@@
 Graphical model structures enforce the conditional independences a modeler believes exists among latent and observed random variables\cite{DBLP:books/daglib/0066829}.
-
 In practice, the sorts of graphical models that humans design, such as Hidden Markov Models, phylogenetic trees, topic models, and Ising models, exhibit certain symmetries and recursive structures that might be amenable to synthesis by programs.
@@
 \end{minipage}
+\hspace{1cm}
 \begin{minipage}[t]{.5\textwidth}
@@
 We wanted to use the SEC algorithm to learn the common abstractions and motifs that underlie the graphical models that are commonly found in the machine learning literature.
-To that end, we supplied SEC with a set of twelve tasks, each of which corresponds to learning how to build a specific type of graphical model, with potentially latent vertices, given that it is provided with a list of observed verticies.
+To that end, we supplied SEC with a set of twelve tasks, each of which corresponds to learning how to build a specific type of graphical model, potentially with latent vertices, given that it is provided with a list of observed verticies.
 A partial list of the tasks is shown in Figure~\ref{fig:tasks}.
-For example, the \textsc{Cylinder} task requires EC to build a program that takes as input at list of observed vertices (of any length) and produces a graph where one ring consists of the observed vertices and the other consists of corresponding latent variables.
-
+For example, the \textsc{Cylinder} task requires EC to build a program that takes as input a list of observed vertices (of any length) and produces a graph where one ring consists of the observed vertices and the other consists of corresponding latent variables.
 Tasks were specified via a set of input/output test cases, and we defined the likelihood of a program for a task to be the fraction of test cases it passed.
@@
 We also included four binary operations on undirected graphical models, each of which computes some kind of union between two graphs.
-These graph operations are shown in Figure~\ref{fig:graphcomb}.
+These graph operations are diagrammed in Figure~\ref{fig:graphcomb}.
 
-Each program synthesis task included a set of input/output test cases; the likelihood of a task, given a program, was defined as the fraction of test cases passed.
-Example input/output pairs of the target programs are shown in figure (\ref{fig:tasks}).
-
-The initial library of primitive combinators contains four binary operations upon undirected graphical models.
-Each of these operations we introduce, called an \emph{undirected union}, serves to combine two smaller models while also introducing new edges.
+%The initial library of primitive combinators contains four binary operations upon undirected graphical models.
+%Each of these operations we introduce, called an \emph{undirected union}, serves to combine two smaller models while also introducing new edges.
 %Using this representation, many models may be expressed in a compositional fashion when combined with higher-order combinators taken from functional programming languages, such as \emph{map} and \emph{fold}.
-The operation of these four undirected unions is diagrammed in figure (\ref{fig:graphcomb}).
+%The operation of these four undirected unions is diagrammed in figure (\ref{fig:graphcomb}).
 %In this representation of graphs, the vertexes are given unique indexes.
@@
 We ran the SEC algorithm on the set of twelve tasks.
-The algorithm is parameterized by the frontier size (the number of unique subprograms $\{e_i\}$ enumerated from the library on each iteration).
+The algorithm is parameterized by the frontier size, the number of unique subprograms $\{e_i\}$ enumerated from the library on each iteration.
 We ran the algorithm both for frontier sizes of 1000 and 5000.
@@
 
+Initially, SEC is unable to find solutions to the complicated graph structure tasks, such as the Ising Model or the Hidden Markov Model, even after enumerating 4 million programs through its heuristic search.
+However, it is able to find programs that generate simpler graphical models, such as Markov chains.
+By compressing the solutions to these simpler problems, it modifies its search procedure and, in subsequent iterations, learns more successful programs.
+
 \begin{figure}
@@
 
-% Initially, SEC is unable to find any solutions to the complicated graph structure tasks, such as the Ising Model or the Hidden Markov Model, even after enumerating 4 million programs through its heuristic search.
-% However, it is able to find programs that generate simpler graphical models, such as Markov chains.
-% By compressing the solutions to these simpler problems, it modifies its search procedure and, in subsequent iterations, learns more and more successful programs.
-% Shown in figure (\ref{TODO}) are the learning curves for SEC, represented both as the number of solved tasks per iteration and as the size of the grammar per iteration.
-
-% What grammars does SEC learn, and how do they enable the bootstrap learning of more complicated programs?
-% The simplest example is that of how the Markov chain bootstraps the learning of the Ising model.
-% Most of the easy tasks, such as the Markov chain, can be solved with the right choice of one higher-order function and one undirected union.
-% After one iteration, the program fragment \verb|(foldr unionC)|, used in the Markov chain program, is incorporated in to the grammar.
-% On the second iteration, the Ising model is learned using this fragment, producing the program \verb|(((S B) map) ((foldr unionC) null))|.
-
-%In future work, could be used as high-level moves in the search for a good structure.
+What grammars does SEC learn, and how do they enable the bootstrap learning of more complicated programs?
+The simplest example is that of how the Markov chain bootstraps the learning of the Ising model.
+%Most of the easy tasks, such as the Markov chain, can be solved with the right choice of one higher-order function and one undirected union.
+After one iteration, the program fragment \verb|(foldr unionC)|, used in the Markov chain program, is incorporated in to the grammar.
+On the second iteration, the Ising model is learned using this fragment, producing the program \verb|(((S B) map) ((foldr unionC) null))|.
 

### 2013-10-23 drafts/nips/nips2013.tex
merged

@@
 The description length of a program is approximated by the number of (not necessarily unique) program fragments in its smallest parse tree given the grammar.
-A slight generalization of the Neville-Manning algorithm then permits tractible approximate maximization of (\ref{eq:gmax}).
-Estimation of the production probabilities of $G$ is performed as in \cite{ijcai}.
+A slight generalization of the Neville-Manning algorithm then permits tractible approximate minimization of (\ref{eq:gmax}).
+Estimation of the production probabilities of $G$ is performed as in \cite{DBLP:conf/ijcai/DechterMAT13}.
 
@@
 \end{minipage}
-\hspace{1cm}
+\hspace{0.2in}
 \begin{minipage}[t]{.5\textwidth}
-% \vspace{\fill}
   \includegraphics[width=\linewidth]{./figures/GraphCombinators.pdf}
@@
 
-\subsection{Relation to Prior Work}
+To be clear, what we have shown so far does not provide a way to learn the structure of a graphical model directly from observed data.
+Rather, this work suggests that we may be able to learn automatically an inductive bias that will subsequently make learning graphical model structures more efficient.
+Some of the programs that SEC includes in its library correspond to graph transformations, and we think that using this transformations as search operators will make structure learning of graphical models with latent variables a more tractable proposition than it currently is.
 
-MH sampling of programs, ILP, Kitzelman, graph grammars
+The notion that an appropriate inductive bias might help in the structure learning of graphical models is not new: \citep{DBLP:conf/icml/HonorioS10} shows that the inductive bias induced via multitask learning helps a search algorithm find accurate graphical model structures.
+But while work in this field has focused on how to evaluate a given graph, we hope to show that SEC can learn appropriate search operators directly from a set of training examples.
 
-\subsection{Limitations of SEC}
+We have presented the SEC algorithm in the context of learning graph structures, but we note that SEC is a general purpose multitask program induction algorithm.
+For an excellent survey of the current state of inductive programming systems, see~\citep{DBLP:conf/aaip/Kitzelmann09}.
 
-Does not look at data. Does everything top-down. Blind to the utility of trying one program versus another (no active learning)
-
-\bibliographystyle{plain}
-\bibliography{eyal_bib}
+\bibliographystyle{plainnat}
+\bibliography{main}
 

### 2013-10-23 drafts/nips/nips2013.tex
writeup

@@
 One example of a structured representation we would like to be able to learn is the probabilistic graphical model.
-Although graphical models are widely used in machine learning to represent the factorizes structure of probability distributions, they are generally designed by researchers rather than by algorithms.
+Although graphical models are widely used in machine learning to represent the factorized structure of probability distributions, they are generally designed by researchers rather than by algorithms.
 This is especially true when the goal is to learn about latent variables underlying the observed data, because it is difficult to capture the knowledge which allows researchers to make reasonable assumptions about which latent variables exists and what they depend on.
@@
 One hypothesis is that these motifs and abstractions are best represented as programs that manipulate and compose graph operations.
-In this work, we discuss how one could learn these motifs and abstracts by finding programs that generate commonly occured graphical model types.
+In this work, we discuss how one could learn these motifs and abstractions by finding programs that generate common graphical model types.
 In particular, we present the SEC algorithm, a general purpose multitask program induction algorithm, that attempts to solve many related tasks by generating a weighted library of useful subroutines that induces a distribution over the space of programs.
@@
 Like the EC algorithm~\cite{DBLP:conf/ijcai/DechterMAT13}, of which it an extension, the SEC algorithm maintains a distribution $\mathcal{D}$ over programs.
-This distribution helps guide its search to relevant parts of of the space of programs.
+%This distribution helps guide its search to relevant parts of of the space of programs.
 In each iteration, SEC does two things: first, it uses $\mathcal{D}$ to explore the space of programs that are likely to be relevant for the tasks it has been given; and second, compresses the solutions it found and uses this compression generate a new weighted library of primitives that defines an updated $\mathcal{D}$.
@@
   \caption{Example input/output pairs of learned programs.
-Each induced program takes as input a list of observed nodes and produces an undirected graphical model.
-Here, we show each program with a small number of input nodes, but the learned programs transfer arbitrary length lists of visible nodes into the corresponding graphs.
-Labeled nodes correspond to observed variables; unlabeled nodes correspond to latent variables.}
+Each induced program takes as input an arbitrarily large list of observed nodes and produces an undirected graphical model.
+Labeled/Unlabeled nodes correspond to observed/latent variables.}
   \label{fig:tasks}
@@
   \includegraphics[width=\linewidth]{./figures/GraphCombinators.pdf}
-  \caption{The operation of the undirected unions.
-Union: computes union of vertexes and edges.
+  \caption{Union: computes union of vertexes and edges.
 UnionC: adds edges between corresponding vertexes.
@@
 We ran the SEC algorithm on the set of twelve tasks.
-The algorithm is parameterized by the frontier size, the number of unique subprograms $\{e_i\}$ enumerated from the library on each iteration.
+The algorithm is parameterized by the frontier size, the number of unique programs drawn from the library on each iteration.
 We ran the algorithm both for frontier sizes of 1000 and 5000.
@@
 In each case, the algorithm returns solutions that partially satisfy the remaining tasks.
-As an example, we show in Figure~\ref{ref:cylinder} an instance of SEC's solution to the \textsc{Cylinder} task.
+As an example, we show in Figure~\ref{ref:cylinder} SEC's solution to the \textsc{Cylinder} task.
 
-Initially, SEC is unable to find solutions to the complicated graph structure tasks, such as the Ising Model or the Hidden Markov Model, even after enumerating 4 million programs through its heuristic search.
+Initially, SEC is unable to find solutions to the complicated graph structure tasks, such as the Ising Model or the Hidden Markov Model, even after enumerating $\approx 4$ million programs through its heuristic search.
 However, it is able to find programs that generate simpler graphical models, such as Markov chains.
@@
 % \begin{minipage}[b]{.45\textwidth}
-\begin{subfigure}[b]{.45\textwidth}
-\includegraphics[width=\linewidth]{./figures/cylinder.pdf}
-\caption{The program generates a cylinder from nodes by generating two ring -- one latent, one visible -- and connecting them.}
+%\begin{subfigure}[b]{.45\textwidth}
+\includegraphics[width=0.45\linewidth]{./figures/cylinder.pdf}
+\caption{The program generates a cylinder from nodes by generating two rings -- one latent, one visible -- and connecting them.}
 % \end{minipage}
-\end{subfigure}
+%\end{subfigure}
+\begin{comment}
 \hspace{0.2in}
@@
 \vfill
-\caption{The combinatory calculus expression tranlated into \textsc{Hakell}-like syntax for readability.}
+\caption{The combinatory calculus expression tranlated into \textsc{Haskell}-like syntax for readability.}
 % \end{minipage}
@@
 \end{minipage}
-\caption{An example of solution that SEC finds to solve. \textsc{Cylinder}\label{ref:cylinder}}
+\end{comment}
+%\caption{An example of solution SEC finds to solve \textsc{Cylinder}\label{ref:cylinder}}
 \end{figure}

### 2013-10-23 drafts/nips/nips2013.tex
writeup

@@
 \usepackage{natbib}
+\usepackage{wrapfig}
 
@@
 
-\title{Incremental Construction of Programs in a Multitask Setting}
+\title{Learning Graphical Concepts}
 
@@
 
-One example of a structured representation we would like to be able to learn is the probabilistic graphical model.
-Although graphical models are widely used in machine learning to represent the factorized structure of probability distributions, they are generally designed by researchers rather than by algorithms.
-This is especially true when the goal is to learn about latent variables underlying the observed data, because it is difficult to capture the knowledge which allows researchers to make reasonable assumptions about which latent variables exists and what they depend on.
-
-In particular, machine learning researchers make frequent use of graphical motifs and abstractions such as trees, chains, rings, grids, mixtures, and plates to constrain the space of graphical models they consider.
+One example of a structured representation we would like to be able to learn is the probabilistic graphical model. %Although graphical models are widely used in machine learning
+% to represent the factorized structure of probability distributions
+% , they are generally designed by researchers rather than by algorithms.
+%This is especially true when the goal is to learn about latent variables underlying the observed data, because it is difficult to capture the knowledge which allows researchers to make reasonable assumptions about which latent variables exists and what they depend on.
+Machine learning researchers make frequent use of graphical motifs and abstractions such as trees, chains, rings, grids, mixtures, and plates to constrain the space of graphical models they consider.
 But how do they learn that these are good motifs?
@@
 One hypothesis is that these motifs and abstractions are best represented as programs that manipulate and compose graph operations.
-In this work, we discuss how one could learn these motifs and abstractions by finding programs that generate common graphical model types.
-In particular, we present the SEC algorithm, a general purpose multitask program induction algorithm, that attempts to solve many related tasks by generating a weighted library of useful subroutines that induces a distribution over the space of programs.
-We discuss how SEC can be used to extract from a set of training tasks the kind of abstract knowledge that machine learning researchers bring to the task of designing graphical models, and we speculate that learning such representations can assist in the difficult tasks of automatically finding good latent-variable graphical models for novel datasets.
+In this work, we discuss how one could learn these motifs and abstractions by finding programs that generate common graphical models.
+In particular, we present the SEC algorithm, a general purpose multitask program induction algorithm.
+%, that attempts to solve related tasks by generating a weighted library of useful subroutines that induces a distribution over the space of programs.
+We discuss how SEC can be used to extract from a set of examples the kind of abstract knowledge that machine learning researchers bring to the task of designing graphical models, and we speculate that learning such representations can assist in the difficult tasks of automatically finding good latent-variable graphical models for novel datasets.
 
@@
 %This distribution helps guide its search to relevant parts of of the space of programs.
-In each iteration, SEC does two things: first, it uses $\mathcal{D}$ to explore the space of programs that are likely to be relevant for the tasks it has been given; and second, compresses the solutions it found and uses this compression generate a new weighted library of primitives that defines an updated $\mathcal{D}$.
+In each iteration, SEC does two things: first, it uses $\mathcal{D}$ to explore the space of programs
+%that are likely to be relevant for the tasks it has been given
+; and second, compresses the solutions it found to generate a new weighted library of primitives that define an updated $\mathcal{D}$.
+
+Following~\citet{DBLP:conf/ijcai/DechterMAT13} and~\citet{DBLP:conf/icml/LiangJK10}, SEC learns programs expressed in
+simply typed combinatory calculus 
+%-- a variable-free subset of the
+%polymorphic simply-typed lambda calculus
+~\citep{DBLP:books/daglib/0005958} %-- as our program representation language.
+ The simplicity the combinatory calculus makes describing a distribution over its expressions particularly simple.
+A more detailed explanation and discussion of this program representation can be found in the papers cited above.
 
@@
 
-Our generative model assumes that the programs that solve a set of related tasks are themselves constructing by composing program fragments that are drawn from a common stochastic grammar.
-This grammar is itself drawn from a distribution over grammars that is biased towards grammars with smaller descriptions lengths.
-This generative model is shown in Figure~\ref{ref:gm}.
-
-\begin{figure}
+\begin{wrapfigure}{R}{0.6\textwidth}
 \begin{codebox}
-\li \id{G} $\sim$ $P_G(\cdot)$ \Comment{Draw grammar from description-length prior}
+\li \id{G} $\sim$ $P_G(\cdot)$ \Comment{Description-length prior}
 \li \id{\ell_n} $\sim\mbox{Uniform}(1, L)$ \Comment{Draw number of subprograms in $n^{th}$ program}
@@
 \li \id{\epsilon_n} $\sim P_{\mbox{noise}}(\cdot)$
-\li \id{t_n} $= \rho_n + \epsilon_n$
+\li \id{t_n} $= \textsc{Eval}(\rho_n) + \epsilon_n$
 \end{codebox}
 \caption{The generative model underlying the SEC algorithm. \label{ref:gm}}
-\end{figure}
+\end{wrapfigure}
+
+Our generative model assumes that the programs that solve a set of related tasks are themselves constructing by composing program fragments that are drawn from a common stochastic grammar.
+This grammar is itself drawn from a distribution over grammars that is biased towards grammars with smaller descriptions lengths.
+This generative model is shown in Figure~\ref{ref:gm}.
 
@@
 %Rather than use mutation and crossover operators, as done in genetic programming, or random resampling of subprograms, as done in recent work in applying Metropolis-Hastings to program learning, we propose that the search moves should themselves be programs drawn from the grammar $G$.
-Concretely, in order to modify a program $e$ to produce new candidate programs, we draw a new function from our grammar, $e'$, and apply that new function to produce the program $e'(e)$.
-Thus, we use the grammar not only as a distribution over programs of interest, but also as a distribution over programs that take as input other programs.
+%Concretely, in order to modify a program $e$ to produce new candidate programs, we draw a new function from our grammar, $e'$, and apply that new function to produce the program $e'(e)$.
+%Thus, we use the grammar not only as a distribution over programs of interest, but also as a distribution over programs that take as input other programs.
 
-In our experiments, we used beam search. The objective function was the unnormalized posterior $P_{t_n|\rho_n}(t_n | \cdot )P_{\rho_n | G}(\cdot | G)$.
+In our experiments, we used beam search. The objective function was equation (\ref{eq:qdist}).
 The beam was initialized to be the programs with the highest prior probability according to the grammar $G$.
@@
 Interpreting negative log probabilities as description lengths, the update (\ref{eq:gmax}) picks the grammar minimizing the sum of the description length of the grammar, plus the expected description length of the programs found to solve the tasks.
+A slight generalization of the Neville-Manning algorithm permits tractible approximate minimization of (\ref{eq:gmax}).
 
+\begin{comment}
 Performing the minimization operation in equation (\ref{eq:gmax}) is generally difficult.
@@
 Estimation of the production probabilities of $G$ is performed as in \cite{DBLP:conf/ijcai/DechterMAT13}.
-
-\subsection{Program Representation}
-
-Following~\citet{DBLP:conf/ijcai/DechterMAT13} and~\citet{DBLP:conf/icml/LiangJK10}, we use a
-simply typed combinatory calculus -- a variable-free subset of the
-polymorphic simply-typed lambda calculus~\citep{DBLP:books/daglib/0005958} -- as our
-program representation language.
-The simplicity the combinatory calculus makes describing a distribution over its expressions particularly simple.
-A more detailed explanation and discussion of this program representation can be found in the papers cited above.
+\end{comment}
 
@@
 \end{minipage}
+\vspace{-0.5cm}
 \end{figure}
 
-\subsection{Experiment}
+\subsection{Experimental Setup}
 
@@
 The SEC algorithm was provided with an initial library of primitive  combinators, which included the basis combinators of the combinatory calculus, function list operations such as \textsc{Map} and \textsc{Foldr}/\textsc{Foldl}.
-We also included four binary operations on undirected graphical models, each of which computes some kind of union between two graphs.
+We also included four binary operations on undirected graphical models, each of which computes a union between two graphs.
 These graph operations are diagrammed in Figure~\ref{fig:graphcomb}.
@@
 The algorithm is parameterized by the frontier size, the number of unique programs drawn from the library on each iteration.
-We ran the algorithm both for frontier sizes of 1000 and 5000.
-Note that because proposed solutions are compositions of these programs, the algorithm looks at many more potential solutions than the number of programs in the frontier.
-
+%We ran the algorithm both for frontier sizes of 1000 and 5000.
+%Note that because proposed solutions are compositions of these programs, the algorithm looks at many more potential solutions than the number of programs in the frontier.
 With a frontier size of 1000, the algorithm solves 7/12 tasks.
@@
 In each case, the algorithm returns solutions that partially satisfy the remaining tasks.
-As an example, we show in Figure~\ref{ref:cylinder} SEC's solution to the \textsc{Cylinder} task.
-
+Figure~\ref{ref:cylinder} shows SEC's solution to the \textsc{Cylinder} task.
+\begin{wrapfigure}{R}{0.6\textwidth}
+\includegraphics[width=\linewidth]{./figures/cylinder.pdf}
+\caption{The program generates a cylinder from nodes by generating two rings -- one latent, one visible -- and connecting them.\label{ref:cylinder}}
+\end{wrapfigure}
 Initially, SEC is unable to find solutions to the complicated graph structure tasks, such as the Ising Model or the Hidden Markov Model, even after enumerating $\approx 4$ million programs through its heuristic search.
@@
 
-\begin{figure}
-% \begin{minipage}[b]{.45\textwidth}
-%\begin{subfigure}[b]{.45\textwidth}
-\includegraphics[width=0.45\linewidth]{./figures/cylinder.pdf}
-\caption{The program generates a cylinder from nodes by generating two rings -- one latent, one visible -- and connecting them.}
-% \end{minipage}
-%\end{subfigure}
-\begin{comment}
-\hspace{0.2in}
-\begin{minipage}[b]{.5\textwidth}
-\begin{subfigure}[t]{\textwidth}
-\includegraphics[width=.5\linewidth]{./figures/cylinderCombinator.pdf}
-\caption{In the combinatory calculus. }
-\vspace{.3in}
-\end{subfigure}
-\begin{subfigure}[b]{\textwidth}
-\begin{lstlisting}[language=haskell, basicstyle=\small, escapechar=+]
-+$\lambda$+vs. let vs' = foldr unionC null vs
-          vs'' = unionHT vs' (I vs')   
-          vs''' = flip vs''
-       in unionC vs'' vs'''
-\end{lstlisting}
-\vfill
-\caption{The combinatory calculus expression tranlated into \textsc{Haskell}-like syntax for readability.}
-% \end{minipage}
-\end{subfigure}
-\end{minipage}
-\end{comment}
-%\caption{An example of solution SEC finds to solve \textsc{Cylinder}\label{ref:cylinder}}
-\end{figure}
-
 What grammars does SEC learn, and how do they enable the bootstrap learning of more complicated programs?
@@
 %Most of the easy tasks, such as the Markov chain, can be solved with the right choice of one higher-order function and one undirected union.
-After one iteration, the program fragment \verb|(foldr unionC)|, used in the Markov chain program, is incorporated in to the grammar.
-On the second iteration, the Ising model is learned using this fragment, producing the program \verb|(((S B) map) ((foldr unionC) null))|.
+After one iteration, the program fragment \textsc{(foldr unionC)}, used in the Markov chain program, is incorporated in to the grammar.
+On the second iteration, the Ising model is learned using this fragment, producing the program
+\textsc{(((S B) map) ((foldr unionC) null))}.
 
@@
 
-We have presented the SEC algorithm in the context of learning graph structures, but we note that SEC is a general purpose multitask program induction algorithm.
+%We have presented the SEC algorithm in the context of learning graph structures, but we note that SEC is a general purpose multitask program induction algorithm.
 For an excellent survey of the current state of inductive programming systems, see~\citep{DBLP:conf/aaip/Kitzelmann09}.

### 2013-10-24 drafts/nips/nips2013.tex
final writeup

@@
 For example, suppose a researcher wants to design a probabilistic graphical model for a novel domain.
-Searching the space of probabilistic models automatically is notoriously difficult and is especially difficult when the latent variables are possible.
-However, when researchers seem able to relatively easily adapt commonly used modeling motifs to new domains.
+Searching the space of probabilistic models automatically is notoriously difficult, especially difficult when latent variables are involved.
+However, researchers seem able to easily adapt commonly used modeling motifs to new domains.
 In doing so, they draw on abstractions such as trees, chains, grids and plates to constrain and direct the kinds of models they produce.
@@
 One natural way to think of these graphical concepts is as programs that take sets of random variables and produce graphical models that relate them.
-In this work, we describe the SEC algorithm, which attempts to learn a distribution over programs by incrementally finding program components that commonly help to solve problems in a given domain, and we show prelimarily that SEC is able to discover the graphical concepts that underlie many of the common graphical model structures.
+In this work, we describe the CEC algorithm, which attempts to learn a distribution over programs by incrementally finding program components that commonly help to solve problems in a given domain, and we show preliminary results indicating that CEC is able to discover the graphical concepts that underlie many of the common graphical model structures.
 \end{abstract}
@@
 But how do they learn that these are good motifs?
-How do they learn how to use them and combine them to design new models?
+How do they learn how to use and combine them to design new models?
 
@@
 In this work, we discuss how one could learn these motifs and abstractions by finding programs that generate common graphical models.
-In particular, we present the SEC algorithm, a general purpose multitask program induction algorithm.
+In particular, we present the CEC algorithm, a general purpose multitask program induction algorithm.
 %, that attempts to solve related tasks by generating a weighted library of useful subroutines that induces a distribution over the space of programs.
-We discuss how SEC can be used to extract from a set of examples the kind of abstract knowledge that machine learning researchers bring to the task of designing graphical models, and we speculate that learning such representations can assist in the difficult tasks of automatically finding good latent-variable graphical models for novel datasets.
+We discuss how CEC can be used to extract from a set of examples the kind of abstract knowledge that machine learning researchers bring to the task of designing graphical models, and we speculate that learning such representations can assist in the difficult tasks of automatically finding good latent-variable graphical models for novel datasets.
 
@@
 
-\section{The SEC Algorithm}
+\section{The CEC Algorithm}
 
-The SEC algorithm takes a library of primitives and a set of related tasks to solve; its goal is to produce a set of programs that solve these tasks.
-Like the EC algorithm~\cite{DBLP:conf/ijcai/DechterMAT13}, of which it an extension, the SEC algorithm maintains a distribution $\mathcal{D}$ over programs.
+The CEC algorithm takes a library of primitives and a set of related tasks to solve; its goal is to produce a set of programs that solve these tasks.
+Like the EC algorithm~\cite{DBLP:conf/ijcai/DechterMAT13}, of which it an extension, the CEC algorithm maintains a distribution $\mathcal{D}$ over programs.
 %This distribution helps guide its search to relevant parts of of the space of programs.
-In each iteration, SEC does two things: first, it uses $\mathcal{D}$ to explore the space of programs
+In each iteration, CEC does two things: first, it uses $\mathcal{D}$ to explore the space of programs;
 %that are likely to be relevant for the tasks it has been given
-; and second, compresses the solutions it found to generate a new weighted library of primitives that define an updated $\mathcal{D}$.
+ and second, it compresses the solutions it found to generate a new weighted library of primitives that define an updated distribution.
 
-Following~\citet{DBLP:conf/ijcai/DechterMAT13} and~\citet{DBLP:conf/icml/LiangJK10}, SEC learns programs expressed in
+Following~\cite{DBLP:conf/ijcai/DechterMAT13} and~\cite{DBLP:conf/icml/LiangJK10}, CEC learns programs expressed in
 simply typed combinatory calculus 
@@
 ~\citep{DBLP:books/daglib/0005958} %-- as our program representation language.
- The simplicity the combinatory calculus makes describing a distribution over its expressions particularly simple.
-A more detailed explanation and discussion of this program representation can be found in the papers cited above.
+ The simplicity of the combinatory calculus makes describing a distribution over its expressions particularly simple.
+A more detailed explanation and discussion of this program representation can be found in~\cite{DBLP:conf/ijcai/DechterMAT13,DBLP:conf/icml/LiangJK10}.
 
 \subsection{Generative Model}
-The SEC algorithm can be expressed as MAP inference in a probabilistic generative model over the space of programs that captures two intuitions: 1) for a set of related tasks, useful programs are composed of useful subparts, and 2) short programs are a priori more likely than long programs.
+The CEC algorithm can be expressed as MAP inference in a probabilistic generative model.
+This generative model captures two intuitions: 1) for a set of related tasks, useful programs are composed of useful subparts, and 2) short programs are a priori more likely than long programs.
 
@@
 
-Our generative model assumes that the programs that solve a set of related tasks are themselves constructing by composing program fragments that are drawn from a common stochastic grammar.
-This grammar is itself drawn from a distribution over grammars that is biased towards grammars with smaller descriptions lengths.
+Our generative model assumes that the programs are themselves constructed by composing program fragments that are drawn from a common stochastic grammar.
+We write $G$ to indicate this grammar, which is drawn from a distribution over grammars that is biased towards grammars with smaller descriptions lengths.
+We write $\{\rho_n\}_{n=1}^N$ to indicate the $N$ programs, each of which solves one of the $N$ tasks.
+We write $\{e_n^i\}_{i=1}^{\ell_n}$ to indicate the $\ell_n$ program fragments that are composed to create $\rho_n$, where $\ell_n$ is drawn uniformly from 1 to $L$, the maximum program length.
+The set $\{t_n\}_{n=1}^N$ refers to the tasks we want to write programs to solve.
+We use how well a program $\rho_n$ solves task $t_n$ as a surrogate for the likelihood function $P(t_n | \rho_n)$.
 This generative model is shown in Figure~\ref{ref:gm}.
@@
 \subsection{Inference}
-We implement an version of the EM algorithm~\cite{dempster:1977} for MAP estimation of the grammar, $G$, given the tasks, $\{t_n\}$.
+We implement a version of the EM algorithm~\cite{dempster:1977} for MAP estimation of the grammar, $G$, given the tasks, $\{t_n\}$.
 The hidden data in this case is the latent program that solves each task.
@@
 \begin{eqnarray}
+\label{eq:qdist}
 q(\rho_n) &\propto& P(t_n | \rho_n) P(\rho_n | G^{\text{old}})\\
-\label{eq:qdist}
+ \label{eq:gmax}
 G^{\text{new}} &=& \operatorname{argmin}_G \left( -\ln P(G) -
@@
  \left[ \ln P(\rho_n | G) \right] \right)
- \label{eq:gmax}
 \end{eqnarray}
 
-Exact computation of the normalizing constant in equation (\ref{eq:qdist}), or the expectation in equation (\ref{eqn:gmax}), would require summing over the infinite space of programs that can be generated from the grammar.
+Exact computation of the normalizing constant in equation (\ref{eq:qdist}), or the expectation in equation (\ref{eq:gmax}), would require summing over the infinite space of programs that can be generated from the grammar.
 To approximate this expectation, we perform a heuristic search over the space of programs, with the goal of finding those programs for which $q(\cdot)$ is high.
@@
 
-In our experiments, we used beam search. The objective function was equation (\ref{eq:qdist}).
-The beam was initialized to be the programs with the highest prior probability according to the grammar $G$.
-Similarly, the search moves considered were those functions with the most highest prior probability under $G$.
+The CEC algorithm uses beam search \cite{Russell:2003:AIM:773294} to maximize $q(\cdot )$.
+The beam is initialized with the programs with the highest prior probability according to the grammar $G$.
+Similarly, the search moves considered are those functions with the highest prior probability under $G$.
 
@@
 Interpreting negative log probabilities as description lengths, the update (\ref{eq:gmax}) picks the grammar minimizing the sum of the description length of the grammar, plus the expected description length of the programs found to solve the tasks.
-A slight generalization of the Neville-Manning algorithm permits tractible approximate minimization of (\ref{eq:gmax}).
+A slight generalization of the Neville-Manning algorithm~\cite{DBLP:journals/corr/cs-AI-9709102} permits tractible approximate minimization of (\ref{eq:gmax}).
 
@@
 Graphical model structures enforce the conditional independences a modeler believes exists among latent and observed random variables\cite{DBLP:books/daglib/0066829}.
-In practice, the sorts of graphical models that humans design, such as Hidden Markov Models, phylogenetic trees, topic models, and Ising models, exhibit certain symmetries and recursive structures that might be amenable to synthesis by programs.
+In practice, the sorts of graphical models that humans design -- Hidden Markov Models, phylogenetic trees, topic models, Ising models -- exhibit certain symmetries and recursive structures that might be amenable to synthesis by programs.
 
@@
 Each induced program takes as input an arbitrarily large list of observed nodes and produces an undirected graphical model.
-Labeled/Unlabeled nodes correspond to observed/latent variables.}
+Labeled/unlabeled nodes correspond to observed/latent variables.}
   \label{fig:tasks}
@@
   \includegraphics[width=\linewidth]{./figures/GraphCombinators.pdf}
-  \caption{Union: computes union of vertexes and edges.
+  \caption{Graph combinators in initial library.
+Union: computes union of vertexes and edges.
 UnionC: adds edges between corresponding vertexes.
@@
 
-We wanted to use the SEC algorithm to learn the common abstractions and motifs that underlie the graphical models that are commonly found in the machine learning literature.
-To that end, we supplied SEC with a set of twelve tasks, each of which corresponds to learning how to build a specific type of graphical model, potentially with latent vertices, given that it is provided with a list of observed verticies.
+We wanted to use the CEC algorithm to learn the common abstractions and motifs that underlie the graphical models that are commonly found in the machine learning literature.
+To that end, we supplied CEC with a set of twelve tasks, each of which corresponds to learning how to build a specific type of graphical model, potentially with latent vertexes, given that it is provided with a list of observed vertexes.
 A partial list of the tasks is shown in Figure~\ref{fig:tasks}.
-For example, the \textsc{Cylinder} task requires EC to build a program that takes as input a list of observed vertices (of any length) and produces a graph where one ring consists of the observed vertices and the other consists of corresponding latent variables.
-Tasks were specified via a set of input/output test cases, and we defined the likelihood of a program for a task to be the fraction of test cases it passed.
+For example, the \textsc{Cylinder} task requires CEC to build a program that takes as input a list of observed vertices (of any length) and produces a graph with one ring consisting of the observed vertexes and the other consists of corresponding latent variables.
+Tasks were specified via a set of input/output test cases, and we defined the score of a program on a task to be the fraction of test cases it passed.
 
-The SEC algorithm was provided with an initial library of primitive  combinators, which included the basis combinators of the combinatory calculus, function list operations such as \textsc{Map} and \textsc{Foldr}/\textsc{Foldl}.
+The CEC algorithm was provided with an initial library of primitive  combinators, which included the basis combinators of the combinatory calculus and functional list operations such as \textsc{Map} and \textsc{Foldr}/\textsc{Foldl}.
 We also included four binary operations on undirected graphical models, each of which computes a union between two graphs.
@@
 
-We ran the SEC algorithm on the set of twelve tasks.
-The algorithm is parameterized by the frontier size, the number of unique programs drawn from the library on each iteration.
+We ran the CEC algorithm on the set of twelve tasks.
+The algorithm is parameterized by the frontier size (the number of unique programs drawn from the library on each iteration).
 %We ran the algorithm both for frontier sizes of 1000 and 5000.
@@
 In each case, the algorithm returns solutions that partially satisfy the remaining tasks.
-Figure~\ref{ref:cylinder} shows SEC's solution to the \textsc{Cylinder} task.
+Figure~\ref{ref:cylinder} shows CEC's solution to the \textsc{Cylinder} task.
 \begin{wrapfigure}{R}{0.6\textwidth}
@@
 \end{wrapfigure}
-Initially, SEC is unable to find solutions to the complicated graph structure tasks, such as the Ising Model or the Hidden Markov Model, even after enumerating $\approx 4$ million programs through its heuristic search.
+Initially, CEC is unable to find solutions to the complicated graph structure tasks, such as the Ising Model or the Hidden Markov Model, even after enumerating $\approx 4$ million programs through its heuristic search.
 However, it is able to find programs that generate simpler graphical models, such as Markov chains.
@@
 
-What grammars does SEC learn, and how do they enable the bootstrap learning of more complicated programs?
-The simplest example is that of how the Markov chain bootstraps the learning of the Ising model.
+What grammars does CEC learn, and how do they enable the bootstrap learning of more complicated programs?
+The simplest example is how the Markov chain bootstraps the learning of the Ising model.
 %Most of the easy tasks, such as the Markov chain, can be solved with the right choice of one higher-order function and one undirected union.
@@
 Rather, this work suggests that we may be able to learn automatically an inductive bias that will subsequently make learning graphical model structures more efficient.
-Some of the programs that SEC includes in its library correspond to graph transformations, and we think that using this transformations as search operators will make structure learning of graphical models with latent variables a more tractable proposition than it currently is.
+Some of the programs that CEC includes in its library correspond to graph transformations, and we think that using these transformations as search operators will make structure learning of graphical models with latent variables a more tractable proposition than it currently is.
 
-The notion that an appropriate inductive bias might help in the structure learning of graphical models is not new: \citep{DBLP:conf/icml/HonorioS10} shows that the inductive bias induced via multitask learning helps a search algorithm find accurate graphical model structures.
-But while work in this field has focused on how to evaluate a given graph, we hope to show that SEC can learn appropriate search operators directly from a set of training examples.
+The notion that an appropriate inductive bias might help in the structure learning of graphical models is not new: \cite{DBLP:conf/icml/HonorioS10} shows that the inductive bias induced via multitask learning helps a search algorithm find accurate graphical model structures.
+But while work in this field has focused on how to evaluate a given graph, we hope to show that CEC can learn appropriate search operators directly from a set of training examples.
 
@@
 
-\bibliographystyle{plainnat}
+\bibliographystyle{plain}
 \bibliography{main}

### 2014-06-08 old/drafts/ijcai2013/ResponsesToReviewers.txt
Moved everything into an a new 'old' directory.

Draft responses to IJCAI 2013 Review

overall common criticisms that should be addressed: 
1) is the work original (1, 4)?
2) how is the work related to ILP and predicate invention (1)
3) how is the work related to Chaitin's Algorithmic Complexity Theory (4)
4)

############################## Response to reviewer 1 ##############################
Main issue: What is the relationship to ILP?

Response: The reviewer is right to raise a connection between our work
and predicate invention in ILP, and we will revise the paper to note
this. However, while predicate invention in ILP is motivated by some
of the same considerations which motivate our work, the
learning framework in ILP is quite different from ours. ILP concerns
learning FOL theories from positive and negative
examples; our framework learns functional programs with respect to a
blackbox evaluator. This results in two main
differences. First, our framework is easily applicable to
both traditional ILP problems (like the boolean circuit learning
example) and problems more commonly found in statistical learning or
genetic programming (like the symbolic regression example). The second
and more crucial difference is that our choice of a functional
programming language makes it natural to capture reusable concepts as
higher-order functions (see, for example, the learned concept $E_1$ or
$E_2$ in Figure 4).

############################## Response to reviewer 2 ##############################
Main issue: Why not do a proper bayesian thing? Statistical discovery?

Response: The reviewer wonders about the algorithm's scalability and
points out that a sufficient number of solutions need to be found by
brute force search for the algorithm to work: we agree with this and
wish to emphasize that it is precisely the presence of some easy
problems that enables the algorithm to initially bootstrap an
effective grammar.

We agree that our approach can be framed as a statistical discovery
technique in which we iteratively re-estimate a stochastic grammar. We
are currently investigating this and will revise the paper to note
this interpretation. However, generating directly from this grammar,
as the reviewer suggests, instead of enumerating in a best-first
manner exacerbates the "needle-in-a-haystack" problem that our
algorithm faces; by enumerating programs, we are able to exhaustively
cover a large part of the prior mass of our stochastic grammar. We are
working to determine the advantages and disadvantages of
these two approaches.

############################## Response to reviewer 3 ##############################
Main issue: Why not use mdl measure like -log(p) instead of counting nodes.

Response: We would like to respond particularly to the reviewer's
question of why an mdl measure such as -log p was not used instead of
counting the nodes of the trees in Section 3.4. We agree with the
reviewer that this approach would be more elegant, and we will revise
the paper to make note of this. In order to take into account the
possibility of new rules (i.e. concept learning), however, we would
need to place a prior distribution over grammars (see, also, the
response to reviewer 2). This would likely require an expensive and
potentially inaccurate Monte Carlo approximation. We found the proxy
calculation of counting the number of unique subtrees in a set of
solutions to be a simple, intuitive, and elegant computation that gets
at the compressibility of the data.

############################## Response to reviewer 4 ##############################
Main issues: 1) Does this extend to noisy data. 2) What about Chaitin? (refer to response to reviewer 1).

Response: We understand the reviewer's concerns that a) this work has
similarities to Chaitin's Algorithmic Information Theory framework and
that b) that it is not clear in the paper how to extend this work to
noisy data. The reviewer is right that our approach is similar in
spirit to some of Chaitin's theoretical work, and his work will be
cited prominently, but our emphasis here is a novel search algorithm
that allows us to quickly bootstrap hundreds of programs for many
related tasks.

Dealing with noisy data requires one to choose a function that scores
candidate solutions to a task. We emphasize the aspects of our
approach that are independent of this, but there are straightforward
ways to extend our algorithm to noisy data. The most way natural is to
weight the subtree counting in section 3.4 by the chosen scoring
function. The paper will be revised to reflect this.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
TODO: implement josh's notes on this:

Notes from Josh: "We will revise the paper to note the The reviewers
are right to raise a connection to ILP and we will revise the paper to
note this, and we will revise the paper to note this, but we wills
tress the following differents.

Some of the same considerations not all
First order logic vs FOL.

We agree with the reviewer and we are currently investigating
(stochastic model thing) maybe move the we are going to do this part
to the start.

We will focus on this initially.

Typo "we would like to respond"

Can we argue about how the number of subtrees relating to the
nonparametric distribution. We will revise the paper to discuss this.
We do see what we are doing as an approximation to all that stuff, and
we will revise the paper to reflect that.

"It was not clear in our paper" instead of "It is not clear." and we
will revise the paper to reflect this. For the noisy data part.

Chaitin, should be cited prominently in the paper and we will.

"We believe there are straightforward ways" --> there are
straightforward ways and we are currently exploring them, and we will
revise the paper to reflect this.

### 2014-06-08 old/drafts/CommitteeMeetingProposal/committeeMeetingProposal.tex
Moved everything into an a new 'old' directory.

\documentclass{article}

\usepackage{times}
\usepackage{amsmath} 
\usepackage{latexsym} 
\usepackage{ dsfont }
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{tikz}
\usepackage{tikz-qtree}
\usetikzlibrary{bayesnet}
\usetikzlibrary{arrows,automata, positioning, patterns,backgrounds}

\newtheorem{definition}{Definition}
\DeclareMathOperator*{\argmax}{argmax} 
\DeclareMathOperator*{\argmin}{argmin} 
\newcommand{\E}{\mathrm{E}}

\title{Second Year Academic Committee Meeting Proposal}

\author{Eyal Dechter}

\begin{document}
\maketitle
\date

\section{Introduction}

Many of the problems an intelligent should be able to solve have
solutions that look like programs. They are hierarchical and
compositional, they reuse modular subcomponents, and they involve
higher order transformations. But the space of programs in any
langauge is vast and does not have the kind of smooth topology of
which many learning algorithms take advantage. How can intelligent
agents learn to induce programs to solve problems from experience?

In recent work, we have proposed an algorithm for bootstrap learning
of modular concepts to facilitate program induction. This algorithm,
which we call E.C. (for Exploration-Compression), iteratively learns a
library of program primtives that enable the algorithm to explore the
space of relevant solutions more efficiently. In the first step of
each iteration, E.C. searchs (\textbf{explores}) the space of ``short''
programs given the current library for programs that solve the tasks
at hand. In the second step, E.C finds common subroutines in the
successful programs and uses these to construct a new library of
program primitives (\textbf{compression}).

In our work on this algorithm, we found this simple procedure can
quickly make search over the space of programs much more
efficient. Whereas the algorithm finds solutions to almost no programs
in the initial iterations, it quickly finds a set of program
primitives that allows it to traverse the set of relevant programs
much more effectively.

I propose to continue on the research program initiated by this
work. In the remainder of this document, I will detail my ideas for
current and future research in this area.

\section{The E.C.E.M algorithm}

The E.C. algorithm, as presented so far, draws on many ideas in
machine learning; these include Learning To Learn, Minimum Description
Length, and Predicate Invention in Inductive Logic
Programming. However, we aim to show that the E.C. algorithm, or
something very similar to it, can be derived as the application of E.M
to MAP inference for a generative model in which a latent stochastic
grammar (nature) generates programs. Individual tasks are the ``data''
of this generative process, for the rewards they provide convey
how close a particular solution is to the task the nature had in
mind. This generative process is depicted in Figure~\ref{fig:ECGM}.
\begin{figure}
\begin{tikzpicture}[->, circle]

\node[latent] (theta) {$\theta$} ;
\node[latent, below=of theta] (G) {G} ;
\node[latent, below=of G]     (e) {$e_i$} ;
\node[obs, below=of e]     (t) {$t_i$} ;

\edge {theta} {G};
\edge {G} {e};
\edge {e} {t};

\plate {plate} {
  (e)(t)}{$i=1, \dots, N$};

\end{tikzpicture}
\caption{Generative model of program induction
  tasks. \label{fig:ECGM}}
\end{figure}

of E.C as finding the grammar $G$ that maximizes the maximizes the
posterior probability of $G$ given the tasks and the prior on the
grammar. We can use E.M. to solve
this procedure. We can derive the E.M update rule as follows:

\begin{align}
G_{new} &= \argmax_{G'} 
     \E [\log \prod_i p(t_i, e_i | G') | G, t_1, \dots, t_N]\\
        =& \argmax_{G'} \E 
           [\sum_i^N \log p(t_i, e_i | G') | G, t_1, \dots, t_N]\\
  =& \argmax_{G'} \sum_i^N \E [\log p(t_i |  e_i) | G, t_1, \dots, t_N] + 
  \sum_i^N \E [\log p(e_i |  G') | G, t_1, \dots, t_N]\\
  =& \argmax_{G'}  
  \sum_i^N \E [\log p(e_i |  G') | G, t_1, \dots, t_N]\\
  =& \argmax_{G'} \sum_i^N \E [\log p(e_i |  G') | G, t_1, \dots, t_N]\\
  =& \argmax_{G'} \sum_i^N \sum_{e_j \sim p(e|G)}
                            p(t_i | e_j, G) 
                            \log p(e_i |  G')
\end{align}

Part of the E.M. algorithm applied to this problem
requires taking the expectation of a function with respect to the
probability of a program given the current grammar and the tasks. This
expectation can be approximated by by likelihood reweighted sampling,
in which we sample programs from the prior and weight by their fit to
the task. This is very similar to the exploration step of the
E.C. algorithm.

Once we estimated this expectation using likelihood reweighted
sampling, we still need to perform the optimization of the objective
over grammars. Any prior distribution over grammars which favors a
short description length is going to support populating the grammar
with production rules corresponding to frequently occuring
subroutines. Thus, the maximization of the object involves a
compression-like step.

Thus, we see that we arrive at an iterative procedure that
successively re-estimate the latent grammar in much the same way that
the E.C. algorithm does.

\section{E.C. with partial solutions}

One of key limitations of the E.C. algorithm as presented so far is
that it does not enable the problem solving agent to utilize
information about partial solutions. But humans, especially when they
are learning to perform a task, rely on incremental feedback from the
world. Thus, if there is a way to partially evaluate solutions, we
would like to find a way to incorporate this into our learning
algorithm.

We are currently developing an algorithm that would do just this. To
make our approach concrete, suppose your task is to build a structure
of blocks that reaches a specified height but is both stable and uses
as few blocks as possible. Moreover, in every task the space has
different obstacles present and perhaps different building blocks
available.

A solution to this problem can be represented as a list of block
placing actions. We could ask our learning agent to produce a program
that generates a list of such actions. Instead, however, we will ask
the agent to generate a program which transform one list of actions
to another. These can be arbitrary transformations. For example, a
transformation can append a new action to the list of actions, it can
repeat the current list of actions a fixed number of times, it can
reverse the list of actions, and it can shift every action in the list
over by some vector.

For each task, the agent starts with an empty list of actions and
searches for a sequences of such transformations whose composition
solve the given task. Every time the agent performs such a
transformation, it can calculate the difference between the objective
function before the transformation and the object function before;
this provides some information about the utility of that particular
transformation.

We are currently working on implementing this approach both in the
domain of tower building and in the domain of constructive solid
geometry. In the latter domain, the task is to discover the set of 3D
object transformations that can reconstruct an input solid.

\section{Looking head}
Not only can humans score partial solutions to estimate their
efficacies, but they can also plan future actions by taking the
current state into account. That is, humans can look at the world,
analyze it, and use that analysis to choose their next step. The
algorithm we proposed in the previous section does not do this. It
only takes into account the previous actions, not the resulting world
state.

Eventually, then, we would like to take sensory information as an
input this program induction system. The major unanswered question is
what the form of the sensory information should be? If its visual
information, can we really expect our program induction system to
learn to extract useful information from an image via trial and error?
If not, then what is the alternative?

\end{document}

### 2014-06-08 old/drafts/ObjectDetection/ObjectDetection.tex
Moved everything into an a new 'old' directory.

\documentclass{article}

\title{Learning compositional object detectors}
\author{Eyal Dechter}

\begin{document}
\maketitle
\date
\section{Introduction}

At that heart of modern AI and machine learning is a tension between the
structured approaches to knowledge representation that dominate in classical
AI and bottom up data driven statistical techniques that deal flexibly with
large amounts of noisy and incomplete data. The field of computer vision is an
excellent example of this tension: on the one hand, visual data is extremely
high-dimensional, ambiguous, and noisy; on the other hand, the knowledge that
computer vision systems attempt to extract from images is highly structured.
The world behind the images is composed of objects and environments in
relations to one another, these objects are composed of other objects, and
both objects and their environments must obey the physical laws.

Recently, several attempts to bridge this gap have been proposed in the
subfield of computer vision that is concerned with object detection. One of
these approaches, due to Fleizen... and McAllester, works by constructing a
grammar over object detectors and learning from data the object detector that
best suits a given classification problem over object types. A detector drawn
from this grammar has at its leaves primitive feature detectors
(Fleizensomething et al. specifically used HOG detectors). The score of a
detector on an image is calculated recursively given the primitive feature
detectors.

Object detectors, which are drawn from a grammar, are a natural way to
describe the ways in which parts give rise to wholes in images of objects.

\end{document}

### 2014-06-08 old/drafts/aistats2013/DechterEllisAISTATS2013_DRAFT.tex
Moved everything into an a new 'old' directory.

\documentclass[twoside]{article}
\usepackage{aistats2014}

% If your paper is accepted, change the options for the package
% aistats2014 as follows:
%
%\usepackage[accepted]{aistats2014}
%
% This option will print headings for the title of your paper and
% headings for the authors names, plus a copyright note at the end of
% the first column of the first page.

\begin{document}

% If your paper is accepted and the title of your paper is very long,
% the style will print as headings an error message. Use the following
% command to supply a shorter title of your paper so that it can be
% used as headings.
%
%\runningtitle{I use this title instead because the last one was very long}

% If your paper is accepted and the number of authors is large, the
% style will print as headings an error message. Use the following
% command to supply a shorter version of the authors names so that
% they can be used as headings (for example, use only the surnames)
%
%\runningauthor{Surname 1, Surname 2, Surname 3, ...., Surname n}

\twocolumn[

\aistatstitle{Symbolic Density Estimation: Uncovering the code.}

\aistatsauthor{ Anonymous Author 1 \And Anonymous Author 2 \And Anonymous Author 3 }

\aistatsaddress{ Unknown Institution 1 \And Unknown Institution 2 \And Unknown Institution 3 } ]

\begin{abstract}
  The Abstract paragraph should be indented 0.25 inch (1.5 picas) on
  both left and right-hand margins. Use 10~point type, with a vertical
  spacing of 11~points.  The {\bf Abstract} heading must be centered,
  bold, and in point size 12. Two line spaces precede the
  Abstract. The Abstract must be limited to one paragraph.
\end{abstract}

\section{GENERAL FORMATTING INSTRUCTIONS}

Both submitted and camera-ready versions of the paper are 8 pages,
plus any additional pages needed for references.

Papers are in 2 columns with the overall line width of 6.75~inches (41~picas). Each column is 3.25~inches wide (19.5~picas).  The space
between the columns is .25~inches wide (1.5~picas).  The left margin is 1~inch (6~picas).  Use 10~point type with a vertical spacing of
11~points. Times Roman is the preferred typeface throughout.

Paper title is 16~point, caps/lc, bold, centered between 2~horizontal rules.  Top rule is 4~points thick and bottom rule is 1~point thick.
Allow 1/4~inch space above and below title to rules.

Author descriptions are center-justified, initial caps.  The lead
author is to be listed first (left-most), and the Co-authors are set
to follow.  If up to three authors, use a single row of author
descriptions, each one center-justified, and all set side by side;
with more authors or unusually long names or institutions, use more
rows.  (But, do not include author names in the initial double-blind
submission!  Instead leave a row of ``Anonymous Author'' descriptions
as above.)

One-half line space between paragraphs, with no indent.

\section{FIRST LEVEL HEADINGS}

First level headings are all caps, flush left, bold, and in point size
12. One line space before the first level heading and 1/2~line space
after the first level heading.

\subsection{Second Level Heading}

Second level headings are initial caps, flush left, bold, and in point
size 10. One line space before the second level heading and 1/2~line
space after the second level heading.

\subsubsection{Third Level Heading}

Third level headings are flush left, initial caps, bold, and in point
size 10. One line space before the third level heading and 1/2~line
space after the third level heading.

\paragraph{Fourth Level Heading}

Fourth level headings must be flush left, initial caps, bold, and
Roman type.  One line space before the fourth level heading, and
place the section text immediately after the heading with, no line
break, but an 11 point horizontal space.

\subsection{CITATIONS, FIGURES, REFERENCES}

\subsubsection{Citations in Text}

Citations within the text should include the author's last name and
year, e.g., (Cheesman, 1985). References should follow any style that
you are used to using, as long as their style is consistent throughout
the paper.  Be sure that the sentence reads correctly if the citation
is deleted: e.g., instead of ``As described by (Cheesman, 1985), we
first frobulate the widgets,'' write ``As described by Cheesman
(1985), we first frobulate the widgets.''  Be sure to avoid
accidentally disclosing author identities through citations.

\subsubsection{Footnotes}

Indicate footnotes with a number\footnote{Sample of the first
  footnote.} in the text. Use 8 point type for footnotes. Place the
footnotes at the bottom of the column in which their markers appear,
continuing to the next column if required. Precede the footnote
section of a column with a 0.5 point horizontal rule 1~inch (6~picas)
long.\footnote{Sample of the second footnote.}

\subsubsection{Figures}

All artwork must be centered, neat, clean, and legible.  All lines
should be very dark for purposes of reproduction, and art work should
not be hand-drawn.  Figures may appear at the top of a column, at the
top of a page spanning multiple columns, inline within a column, or
with text wrapped around them, but the figure number and caption
always appear immediately below the figure.  Leave 2 line spaces
between the figure and the caption. The figure caption is initial caps
and each figure should be numbered consecutively.

Make sure that the figure caption does not get separated from the
figure. Leave extra white space at the bottom of the page rather than
splitting the figure and figure caption.
\begin{figure}[h]
\vspace{.3in}
\centerline{\fbox{This figure intentionally left non-blank}}
\vspace{.3in}
\caption{Sample Figure Caption}
\end{figure}

\subsubsection{Tables}

All tables must be centered, neat, clean, and legible. Do not use hand-drawn tables. Table number and title always appear above the table.
See Table~\ref{sample-table}.

One line space before the table title, one line space after the table title, and one line space after the table. The table title must be
initial caps and each table numbered consecutively.

\begin{table}[h]
\caption{Sample Table Title} \label{sample-table}
\begin{center}
\begin{tabular}{ll}
{\bf PART}  &{\bf DESCRIPTION} \\
\hline \\
Dendrite         &Input terminal \\
Axon             &Output terminal \\
Soma             &Cell body (contains cell nucleus) \\
\end{tabular}
\end{center}
\end{table}

\section{SUPPLEMENTARY MATERIAL}

If you need to include additional appendices during submission, you
can include them in the supplementary material file.

\newpage

\section{INSTRUCTIONS FOR CAMERA-READY PAPERS}

For the camera-ready paper, if you are using \LaTeX, please make sure
that you follow these instructions.  (If you are not using \LaTeX,
please make sure to achieve the same effect using your chosen
typesetting package.)

\begin{enumerate}
    \item Install the package \texttt{fancyhdr.sty}. The
    \texttt{aistats2014.sty} file will make use of it.
    \item Begin your document with
    \begin{flushleft}
    \texttt{\textbackslash documentclass[twoside]\{article\}}\\
    \texttt{\textbackslash usepackage[accepted]\{aistats2014\}}
    \end{flushleft}
    The \texttt{twoside} option for the class article allows the
    package \texttt{fancyhdr.sty} to include headings for even and odd
    numbered pages. The option \texttt{accepted} for the package
    \texttt{aistats2014.sty} will write a copyright notice at the end of
    the first column of the first page. This option will also print
    headings for the paper.  For the \emph{even} pages, the title of
    the paper will be used as heading and for \emph{odd} pages the
    author names will be used as heading.  If the title of the paper
    is too long or the number of authors is too large, the style will
    print a warning message as heading. If this happens additional
    commands can be used to place as headings shorter versions of the
    title and the author names. This is explained in the next point.
    \item  If you get warning messages as described above, then
    immediately after $\texttt{\textbackslash
    begin\{document\}}$, write
    \begin{flushleft}
    \texttt{\textbackslash runningtitle\{Provide here an alternative shorter version of the title of your
    paper\}}\\
    \texttt{\textbackslash runningauthor\{Provide here the surnames of the authors of your paper, all separated by
    commas\}}
    \end{flushleft}
    The text that appears as argument in \texttt{\textbackslash
      runningtitle} will be printed as a heading in the \emph{even}
    pages. The text that appears as argument in \texttt{\textbackslash
      runningauthor} will be printed as a heading in the \emph{odd}
    pages.  If even the author surnames do not fit, it is acceptable
    to give a subset of author names followed by ``et al.''

    \item Use the file sample\_paper.tex as an example.

    \item Both submitted and camera-ready versions of the paper are 8
      pages, plus any additional pages needed for references.

    \item If you need to include additional appendices,
      you can include them in the supplementary
      material file.

    \item Please, don't change the layout given by the above
      instructions and by the style file.

\end{enumerate}

\subsubsection*{Acknowledgements}

Use unnumbered third level headings for the acknowledgements.  All
acknowledgements go at the end of the paper.  Be sure to omit any
identifying information in the initial double-blind submission!

\subsubsection*{References}

References follow the acknowledgements.  Use an unnumbered third level
heading for the references section.  Any choice of citation style is
acceptable as long as you are consistent.  Please use the same font
size for references as for the body of the paper---remember that
references do not count against your page length total.

J.~Alspector, B.~Gupta, and R.~B.~Allen (1989). Performance of a
stochastic learning microchip.  In D. S. Touretzky (ed.), {\it
  Advances in Neural Information Processing Systems 1}, 748-760.  San
Mateo, Calif.: Morgan Kaufmann.

F.~Rosenblatt (1962). {\it Principles of Neurodynamics.} Washington,
D.C.: Spartan Books.

G.~Tesauro (1989). Neurogammon wins computer Olympiad.  {\it Neural
  Computation} {\bf 1}(3):321-323.

\end{document}

### 2014-06-08 old/drafts/aistats2013/abstracts.tex
Moved everything into an a new 'old' directory.

--- 2014-06-08 old/drafts/aistats2013/DechterEllisAISTATS2013_DRAFT.tex
@@
 
-% If your paper is accepted, change the options for the package
-% aistats2014 as follows:
-%
-%\usepackage[accepted]{aistats2014}
-%
-% This option will print headings for the title of your paper and
-% headings for the authors names, plus a copyright note at the end of
-% the first column of the first page.
-
 \begin{document}
-
-% If your paper is accepted and the title of your paper is very long,
-% the style will print as headings an error message. Use the following
-% command to supply a shorter title of your paper so that it can be
-% used as headings.
-%
-%\runningtitle{I use this title instead because the last one was very long}
-
-% If your paper is accepted and the number of authors is large, the
-% style will print as headings an error message. Use the following
-% command to supply a shorter version of the authors names so that
-% they can be used as headings (for example, use only the surnames)
-%
-%\runningauthor{Surname 1, Surname 2, Surname 3, ...., Surname n}
 
@@
 
-\aistatstitle{Symbolic Density Estimation: Uncovering the code.}
+\aistatstitle{Symbolic Density Estimation: Uncovering the Code.}
 
@@
 \begin{abstract}
-  The Abstract paragraph should be indented 0.25 inch (1.5 picas) on
-  both left and right-hand margins. Use 10~point type, with a vertical
-  spacing of 11~points.  The {\bf Abstract} heading must be centered,
-  bold, and in point size 12. Two line spaces precede the
-  Abstract. The Abstract must be limited to one paragraph.
+One prominent view in Machine Learning is that to make accurate predictions and generalizations from data, we need to learn the latent probabilistic generative models that give rise to the data we observe.
+Graphical models and neural networks
+
 \end{abstract}
 
-\section{GENERAL FORMATTING INSTRUCTIONS}
+\begin{abstract}
 
-Both submitted and camera-ready versions of the paper are 8 pages,
-plus any additional pages needed for references.
-
-Papers are in 2 columns with the overall line width of 6.75~inches (41~picas). Each column is 3.25~inches wide (19.5~picas).  The space
-between the columns is .25~inches wide (1.5~picas).  The left margin is 1~inch (6~picas).  Use 10~point type with a vertical spacing of
-11~points. Times Roman is the preferred typeface throughout.
-
-Paper title is 16~point, caps/lc, bold, centered between 2~horizontal rules.  Top rule is 4~points thick and bottom rule is 1~point thick.
-Allow 1/4~inch space above and below title to rules.
-
-Author descriptions are center-justified, initial caps.  The lead
-author is to be listed first (left-most), and the Co-authors are set
-to follow.  If up to three authors, use a single row of author
-descriptions, each one center-justified, and all set side by side;
-with more authors or unusually long names or institutions, use more
-rows.  (But, do not include author names in the initial double-blind
-submission!  Instead leave a row of ``Anonymous Author'' descriptions
-as above.)
-
-One-half line space between paragraphs, with no indent.
-
-\section{FIRST LEVEL HEADINGS}
-
-First level headings are all caps, flush left, bold, and in point size
-12. One line space before the first level heading and 1/2~line space
-after the first level heading.
-
-\subsection{Second Level Heading}
-
-Second level headings are initial caps, flush left, bold, and in point
-size 10. One line space before the second level heading and 1/2~line
-space after the second level heading.
-
-\subsubsection{Third Level Heading}
-
-Third level headings are flush left, initial caps, bold, and in point
-size 10. One line space before the third level heading and 1/2~line
-space after the third level heading.
-
-\paragraph{Fourth Level Heading}
-
-Fourth level headings must be flush left, initial caps, bold, and
-Roman type.  One line space before the fourth level heading, and
-place the section text immediately after the heading with, no line
-break, but an 11 point horizontal space.
-
-\subsection{CITATIONS, FIGURES, REFERENCES}
-
-\subsubsection{Citations in Text}
-
-Citations within the text should include the author's last name and
-year, e.g., (Cheesman, 1985). References should follow any style that
-you are used to using, as long as their style is consistent throughout
-the paper.  Be sure that the sentence reads correctly if the citation
-is deleted: e.g., instead of ``As described by (Cheesman, 1985), we
-first frobulate the widgets,'' write ``As described by Cheesman
-(1985), we first frobulate the widgets.''  Be sure to avoid
-accidentally disclosing author identities through citations.
-
-\subsubsection{Footnotes}
-
-Indicate footnotes with a number\footnote{Sample of the first
-  footnote.} in the text. Use 8 point type for footnotes. Place the
-footnotes at the bottom of the column in which their markers appear,
-continuing to the next column if required. Precede the footnote
-section of a column with a 0.5 point horizontal rule 1~inch (6~picas)
-long.\footnote{Sample of the second footnote.}
-
-\subsubsection{Figures}
-
-All artwork must be centered, neat, clean, and legible.  All lines
-should be very dark for purposes of reproduction, and art work should
-not be hand-drawn.  Figures may appear at the top of a column, at the
-top of a page spanning multiple columns, inline within a column, or
-with text wrapped around them, but the figure number and caption
-always appear immediately below the figure.  Leave 2 line spaces
-between the figure and the caption. The figure caption is initial caps
-and each figure should be numbered consecutively.
-
-Make sure that the figure caption does not get separated from the
-figure. Leave extra white space at the bottom of the page rather than
-splitting the figure and figure caption.
-\begin{figure}[h]
-\vspace{.3in}
-\centerline{\fbox{This figure intentionally left non-blank}}
-\vspace{.3in}
-\caption{Sample Figure Caption}
-\end{figure}
-
-\subsubsection{Tables}
-
-All tables must be centered, neat, clean, and legible. Do not use hand-drawn tables. Table number and title always appear above the table.
-See Table~\ref{sample-table}.
-
-One line space before the table title, one line space after the table title, and one line space after the table. The table title must be
-initial caps and each table numbered consecutively.
-
-\begin{table}[h]
-\caption{Sample Table Title} \label{sample-table}
-\begin{center}
-\begin{tabular}{ll}
-{\bf PART}  &{\bf DESCRIPTION} \\
-\hline \\
-Dendrite         &Input terminal \\
-Axon             &Output terminal \\
-Soma             &Cell body (contains cell nucleus) \\
-\end{tabular}
-\end{center}
-\end{table}
-
-\section{SUPPLEMENTARY MATERIAL}
-
-If you need to include additional appendices during submission, you
-can include them in the supplementary material file.
-
-\newpage
-
-\section{INSTRUCTIONS FOR CAMERA-READY PAPERS}
-
-For the camera-ready paper, if you are using \LaTeX, please make sure
-that you follow these instructions.  (If you are not using \LaTeX,
-please make sure to achieve the same effect using your chosen
-typesetting package.)
-
-\begin{enumerate}
-    \item Install the package \texttt{fancyhdr.sty}. The
-    \texttt{aistats2014.sty} file will make use of it.
-    \item Begin your document with
-    \begin{flushleft}
-    \texttt{\textbackslash documentclass[twoside]\{article\}}\\
-    \texttt{\textbackslash usepackage[accepted]\{aistats2014\}}
-    \end{flushleft}
-    The \texttt{twoside} option for the class article allows the
-    package \texttt{fancyhdr.sty} to include headings for even and odd
-    numbered pages. The option \texttt{accepted} for the package
-    \texttt{aistats2014.sty} will write a copyright notice at the end of
-    the first column of the first page. This option will also print
-    headings for the paper.  For the \emph{even} pages, the title of
-    the paper will be used as heading and for \emph{odd} pages the
-    author names will be used as heading.  If the title of the paper
-    is too long or the number of authors is too large, the style will
-    print a warning message as heading. If this happens additional
-    commands can be used to place as headings shorter versions of the
-    title and the author names. This is explained in the next point.
-    \item  If you get warning messages as described above, then
-    immediately after $\texttt{\textbackslash
-    begin\{document\}}$, write
-    \begin{flushleft}
-    \texttt{\textbackslash runningtitle\{Provide here an alternative shorter version of the title of your
-    paper\}}\\
-    \texttt{\textbackslash runningauthor\{Provide here the surnames of the authors of your paper, all separated by
-    commas\}}
-    \end{flushleft}
-    The text that appears as argument in \texttt{\textbackslash
-      runningtitle} will be printed as a heading in the \emph{even}
-    pages. The text that appears as argument in \texttt{\textbackslash
-      runningauthor} will be printed as a heading in the \emph{odd}
-    pages.  If even the author surnames do not fit, it is acceptable
-    to give a subset of author names followed by ``et al.''
-
-    \item Use the file sample\_paper.tex as an example.
-
-    \item Both submitted and camera-ready versions of the paper are 8
-      pages, plus any additional pages needed for references.
-
-    \item If you need to include additional appendices,
-      you can include them in the supplementary
-      material file.
-
-    \item Please, don't change the layout given by the above
-      instructions and by the style file.
-
-\end{enumerate}
-
-\subsubsection*{Acknowledgements}
-
-Use unnumbered third level headings for the acknowledgements.  All
-acknowledgements go at the end of the paper.  Be sure to omit any
-identifying information in the initial double-blind submission!
-
-\subsubsection*{References}
-
-References follow the acknowledgements.  Use an unnumbered third level
-heading for the references section.  Any choice of citation style is
-acceptable as long as you are consistent.  Please use the same font
-size for references as for the body of the paper---remember that
-references do not count against your page length total.
-
-J.~Alspector, B.~Gupta, and R.~B.~Allen (1989). Performance of a
-stochastic learning microchip.  In D. S. Touretzky (ed.), {\it
-  Advances in Neural Information Processing Systems 1}, 748-760.  San
-Mateo, Calif.: Morgan Kaufmann.
-
-F.~Rosenblatt (1962). {\it Principles of Neurodynamics.} Washington,
-D.C.: Spartan Books.
-
-G.~Tesauro (1989). Neurogammon wins computer Olympiad.  {\it Neural
-  Computation} {\bf 1}(3):321-323.
+\end{abstract}
 

### 2014-06-08 old/drafts/ijcai2013/figures/4sqtree.tex
Moved everything into an a new 'old' directory.

--- 2013-04-22 drafts/ijcai2013/ijcai13.tex
@@
-\documentclass{article}
-
-\usepackage{ijcai13}
-\usepackage{times}
-\usepackage{amsmath} 
-\usepackage{amsfonts}
-\usepackage{latexsym} 
-\usepackage{ dsfont }
-\usepackage[usenames,dvipsnames]{xcolor}
+\documentclass[preview]{standalone}
 \usepackage{tikz}
 \usepackage{tikz-qtree}
-\usepackage{float}
-\usepackage{circuitikz}
-\usepackage{graphicx}
-\usepackage[font=small]{subcaption}
-\usepackage[font=small]{caption}
-\usepackage{alltt}
-\usepackage{dashrule}
-\usepackage{array}
-\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
-\usetikzlibrary{arrows,automata, positioning, patterns,backgrounds}
-\usetikzlibrary{arrows,shapes.gates.logic.US,shapes.gates.logic.IEC,calc}
-
-\usepackage{times}
-\pdfinfo{
-  /Title (Bootstrap learning via modular concept discovery)
-  /Author (Eyal Dechter, Jon Malmaud, Ryan P. Adams, Joshua B. Tenenbaum)
-}
-
-\newtheorem{definition}{Definition}
-\DeclareMathOperator*{\argmax}{arg\,max} 
-\DeclareMathOperator*{\argmin}{arg\,min}
-
-\title{Bootstrap learning via modular concept discovery}
-
-\author{Eyal Dechter \\
-MIT\\
-USA \\
-edechter@mit.edu
-\And
-Jon Malmaud \\
-MIT \\ 
-USA \\
-malmaud@mit.edu
-\And 
-Ryan~P.  Adams \\
-Harvard University\\
-USA \\
-rpa@seas.harvard.edu
-\And
-Joshua~B. Tenenbaum \\
-MIT\\
-USA \\
-jbt@mit.edu
-}
+\usepackage{subcaption}
+\usetikzlibrary{arrows,automata, positioning, patterns}
 
 \begin{document}
-
-\maketitle
-\begin{abstract}
-%\small
- Suppose a learner is faced with a domain of problems about which it
- knows nearly nothing. It does not know the distribution of problems,
- the space of solutions is not smooth, and the reward signal is
- uninformative, providing perhaps a few bits of information but not
- enough to steer the learner effectively. How can such a learner ever
- get off the ground? A common intuition is that if the solutions to
- these problems share a common structure, and the learner can solve
- some simple problems by brute force, it should be able to extract
- useful components from these solutions and, by composing them,
- explore the solution space more efficiently. Here, we formalize this
- intuition, where the solution space is the space of typed functional
- programs and the gained information is stored as a stochastic grammar
- over programs. We propose an iterative procedure for exploring such
- spaces: in the first step of each iteration, the learner explores a
- finite subset of the domain, guided by a stochastic grammar; in the
- second step, the learner compresses the successful solutions from the
- first step to estimate a new stochastic grammar. We test this
- procedure on symbolic regression and Boolean circuit learning and
- show that the learner discovers modular concepts for these
- domains. Whereas the learner is able to solve almost none of the
- posed problems in the procedure's first iteration, it rapidly becomes
- able to solve a large number by gaining abstract knowledge of the
- structure of the solution space.
-
-\end{abstract}
-
-\section{Introduction}
-A central question for AI is how to create intelligent agents without
-building in vast amounts of pre-specified knowledge that is tailored
-to one domain or another. Any attempt to solve this problem must
-contend with the fact that the less an agent knows about solving
-problems in a domain, the worse its attempts are, and, therefore, the
-less informative they are about how to solve subsequent problems.
-
-For example, imagine you know nothing about electronics but are
-supplied with a collection of circuit elements and asked to build a
-useful device: your initial attempts will be uniformly
-ineffective. If, on the other hand, you know the basics, you are far
-more likely to generate devices that are somewhat successful and on
-whose basis you might improve your skills.
-
-Like many of the problems in AI, this one can be cast as a search
-problem: the goal is to find a circuit wiring procedure that maximizes
-certain criteria.  There are many search algorithms, but they typically
-exploit local or topological structure in the search space. For
-example, local search algorithms move from one candidate solution to a
-neighboring one. Heuristic search techniques take advantage of
-topological structure to prune off large parts of the solution space.
-
-Domain specific knowledge provides this structure. In local search, it
-defines the similarity metric between solutions; in heuristic search
-it provides the heuristic. Here we ask how to bootstrap domain
-specific knowledge and extract structure from the space of possible
-solutions by attempting to solve multiple related problems in a given
-domain at once. Our approach is motivated by the observation that many
-interesting problems -- such as discovering scientific theories,
-developing new technologies, learning to play a new instrument -- have
-solutions that are natrually specified as programs.
-
-Our proposed solution is thus inspired by the programmer who -- having
-reused subroutines to solve similar simple tasks -- creates a function
-library to help manage the complexity of larger projects. The
-E.C. algorithm, which we introduce here, builds a library of reusable
-program components and places a distribution over these, effectively
-learning from simple tasks a search space structure that enables it to
-solve more complex tasks.
-
-In this work, we try to see how far this idea of reusable function
-definition can get us. We ask the following question: can an AI system
-discover the structure latent in the solutions to multiple related
-problems and thereby bootstrap effective exploration of the search
-space? Can discovering modular program components transform a problem
-from one in which search is intractable into a problem in which it
-becomes increasingly feasible with experience?
-
-Concretely, the E.C. algorithm works as follows. It begins with a small
-library of program primitives and a distribution over programs
-synthesized from this library. In each iteration, it generates
-programs in order of highest probability according to the library
-learned in the previous iteration. From the discovered solutions, it
-finds new reusable functions that compress these solutions, and it
-re-estimates the distribution over programs by weighting these new
-functions according to their frequencies. In using compression as a
-guide to the choice of representation, we instantiate the idea that
-good representations are those which minimize the description length
-of typical elements in the domain.
-
-We show that this iterative approach allows us to achieve very high
-performance in tasks where just searching according to the initial
-distribution over program primitives clearly fails.
-
-\section{Related Work}
-The idea of discovering and using reusable subcomponents is part of an
-old tradition in AI. For instance, it is one of the central themes in
-Herb Simon's 1969 ``The Sciences of the
-Artificial''~\shortcite{simon1996sciences}. Rendell's 1985
-paper~\shortcite{DBLP:conf/ijcai/Rendell85}, ``Substantial Constructive
-Induction Using Layered Information Compression: Tractable Feature
-Formation in Search,'' presented within a classical search paradigm
-the idea that compression provides a heuristic for constructing good
-representations. Within the Inductive Logic Programming (ILP)
-literature, researchers have explored ``Constructive Induction,''
-generating new and reusable logic programming
-terms~\cite{DBLP:conf/ijcai/Muggleton87}. Later, predicate invention
-was also applied in the multitask
-setting~\cite{DBLP:conf/ilp/KhanMP98}. We apply these ideas to
-learning functional programs and believe that the modularity and
-compositionality afforded by this representation is necessary for
-these ideas to be successful.
-
-Genetic programming
-(GP)~\cite{DBLP:conf/foga/Koza92,DBLP:conf/ices/KozaBAK96} has
-tackled the problem of program learning and has explored the notion
-that reusable functions are helpful. This work relies, however, on
-stochastic local search over the space of programs, and automatically
-discovering useful subroutines is seen as a helpful heuristic. More
-recently, Liang et al.~\shortcite{liang10programs} used a stochastic
-grammar over combinatory logic to induce shared structure in
-multi-task program induction problems. We find this latter
-representation compelling and use it here, but we see both this and
-genetic programming as tackling the question of how to benefit from
-shared structure in situations where local search might be successful
-on its own. Such problems rely on domain knowledge -- whether in the
-form of a fitness function or a likelihood function -- to provide
-local structure to the search space.
-
-But, as mentioned before, the problems we want AI systems to solve
-often lack this kind of built in structure. Here we attempt to learn
-functional programs to solve those kinds of problems.  \\
-
-%% AI and machine learning algorithms commonly take advantage of the
-%% relationship between a performance metric (error, reward, fitness,
-%% etc.) and the local or topological structure of the solution
-%% space. Whether in discrete or continuous domains, the key is to design
-%% such a relationship and write a algorithm that either moves in good
-%% directions or prunes off large parts of the solution space. Whether or
-%% not such a relationship exists depends on the representation, and much
-%% work has gone into finding good representations for problem
-%% solving. Recent work, especially in machine learning, has focused on
-%% ``feature learning,'' with the explicit aim of discovered good
-%% representational building blocks in vector-valued solutions spaces.
-
-%% Many of the most natural solution representations for intelligent
-%% agents, however, are not naturally expressed as vector-valued
-%% spaces. Consider the problems of discovering scientific theories,
-%% inventing new technologies, communicating with others, or executing a
-%% complex motor routine (such as dancing or playing an instrument). The
-%% solutions to these problems are most naturally expressed as programs,
-%% that is, as the composition of modular components whose semantics are
-%% determined by the actions of one component on another.
-
-%% How can we learn good solution representations for these kinds of
-%% tasks without building in the kind of domain knowledge that gives rise
-%% to a useful relationship between representation and performance? How
-%% can we get traction on a set of problems when small tweaks to
-%% solutions give rise to wildly varying performances in an unpredictable
-%% manner, or when, for most problems, most of the solutions fail in a
-%% uniformly bad way?
-
-%% A popular idea is that good representations are those that encode the
-%% data space in an efficient way, that is, that enable compression of
-%% the data~\cite{}. For program representations, this most naturally
-%% corresponds to defining reusable components so that the total number
-%% of unique expressions for a set of solutions is minimized (i.e., this
-%% is directly akin to the subroutine libraries programmers use to manage
-%% the complexity of their code).
-
-%% In this work, we try to see how far reusable subroutine definition can
-%% get us on its own. Our question is, Can discovering modular components
-%% transform a problem from one in which brute-force search will not work
-%% to a problem in which it does?
-
-%% To emphasize an impoverished relationship between reward signals and
-%% program representations, we make our tasks all-or-none: a solution
-%% either solves the problem or it does not, and partial solutions do not
-%% get any credit.
-
-%% The intuitive approach to applying this idea is simple: if solutions
-%% to problems in the domain share common structure, and there exist
-%% problems whose solutions are simple enough to find by brute force,
-%% then we should be able to extract a distribution over solution
-%% subcomponents from these simple cases that allows us to run subsequent
-%% searches more efficiently.
-
-%% More concretely, our algorithm works as follows: each iteration, it
-%% first generates programs in order of highest probability according to
-%% a subroutine distribution learned in the previous iteration. From the
-%% discovered solutions, it finds new reusable subroutines that compress
-%% these solutions and estimates their probabilities.
-
-%% We show that this iterative approach allows us to achieve very high
-%% performance in tasks where just searching according to the initial
-%% distribution over program primitives clearly fails.
-
-%%  \hline
-
-%% AI and machine learning algorithms commonly take advantage of the idea
-%% that local structure in the solution space is related to local
-%% structure in the outcome. In continuous domains, learning algorithms
-%% algorithms use error gradients to guide search. In discrete domains,
-%% whether in heuristic or local search paradigms, knowledge about the
-%% solution space topology and its relation to reward and cost metrics
-%% enables pruning, reasoning, and moving in the solution space more
-%% efficiently. However, choosing the representations that have
-%% beneficial structural properties requires a lot of general and domain
-%% knowledge, and this what researchers do and what they cannot leave up
-%% to their algorithms.
-
-%% What is a good strategy for an agent when its representation of the
-%% solution space is not smooth with respect to the outcome? Many hard
-%% learning problems have this property, especially those in which the
-%% solution space is most naturally represented as a set of
-%% programs. Consider the problems of discovering scientific theories,
-%% inventing new technologies, learning to cook, or learning to
-%% communicate. These are problems in which 1) solutions are best
-%% described as programs (i.e. compositions of modular subroutines at
-%% various levels of abstraction), 2) most of the solution space is bad
-%% in a relatively uninformative way (i.e. it just doesn't work), and
-%% thus 3) outcomes are not informative about how you are doing unless
-%% you are already quite close to a solution.
-
-%% Intuitively, the way to solve such problems is to find solutions to
-%% simple cases in the domain and extract conceptual building blocks that
-%% can be composed to scale up to complex cases. With experience, the
-%% agent may be able to discover local structure in the solution space,
-%% and utilize more traditional AI and machine learning techniques. In
-%% this work, we will be focusing on the first phase of this learning
-%% problem.
-
-%% %% Of course, every learner
-
-%% %% Choosing the wrong representation, in
-%% %% which local structure is not apparent, is often the difference between
-%% %% a successful learner and an unsuccessful one. In many respects, this
-%% %% may be the kind of situation that adult humans face when confronting
-%% %% new domains: we have a lot of general knowledge of successful
-%% %% representations, useful strategies, etc.
-
-%% %% But suppose that you have very little knowledge of a domain. Your
-%% %% intial representation of the domain is poor: ``nearby'' points in the
-%% %% solutions space ellicit wildly different outcomes in unpredictable
-%% %% ways. Or perhaps nearby points ellicit the same outcome and its not a
-%% %% good one, so you don't know which way to go. Many learning problems,
-%% %% especially those whose solutions are best described as programs, have
-%% %% these qualities: imagine learning to synthesize new sentences
-
-%% %% teness. We choose the representations our learners use
-%% %% and these representations generally Imagine you are trying to learn to
-%% %% bake via trial and error. You know what kind of ingredients are used
-%% %% in baking (flour, water, butter, sugar, yeast, baking soda, etc.) and
-%% %% you know a bit about how to combine these ingredient (mix, whisk,
-%% %% etc.) and how to apply heat (oven). But baking is a notoriously
-%% %% fragile process: ratios of ingredients, the order of mixing, cooking
-%% %% temperatures, resting times, all make a huge difference, and the
-%% %% outcome of the baking process is mostly all or none. Either the
-%% %% outcome is what you expected or quite unlike it. Even if you do create
-%% %% something similar to your intent, it is difficult to assign blame for
-%% %% the outcome to any one step or decision along the way.
-
-%% %% Many difficult learning problems have this structure -- learning
-%% %% language, developing technologies and scientific theories, learning to
-%% %% play instruments or execute other complex motor patterns. These are
-%% %% problems in which 1) there is a large set of similar tasks; 2) their
-%% %% solutions are best described as programs; 3) the solution space is
-%% %% not smooth with respect to the outcome; 4) and the outcome is generally
-%% %% uninformative unless you are already pretty close to a solution.
-
-%% %% Of course, knowledge is not acquired wholly by trial and error. But in
-%% %% many domains, there is at least initially such a period of ignorance,
-%% %% and the question we are trying to tackle here is how you ever get off
-%% %% the ground.
-
-%% I need help here.
-
-%% %% Let a \emph{task} be a function $t : \mathcal{L} \rightarrow \{0,
-%% %% 1\}$, where $\mathcal{L}$ is the set of all expressions in our
-%% %% representation language. We say that $e \in \mathcal{L}$ \emph{solves}
-%% %% task $t$ if $t(e) = 1$. Our goal is to solve as many tasks as
-%% %% possible, as efficiently as possible. That is, suppose that there is a
-%% %% distribution over tasks $t \sim  p_t(\cdot)$, we would like an ordering $O
-%% %% = e_1, \dots $ on expressions in $\mathcal{L}$ such that if $p_t(t_1)
-%% %% > p_t(t_2)$ then the first expression in $O$ to solve $t_1$ comes
-%% %% before the first expression in $O$ to solve $t_2$.
-
-%% %% We motivate a solution to this problem by considering a heirarchical
-%% %% Bayesian formulation of the task distribution. Suppose we believe that
-%% %% there is some underlying structure to $p_t(\cdot)$ that is explained by
-%% %% the fact that for each task $t$ there is a latent expressions $e \in
-%% %% \mathcal{L}$ drawn from a stochastic grammar $\mathcal{G}$ and that
-
-%% %% \begin{align}
-%% %% e &\sim p_\mathcal{L}(\cdot) = \mathcal{G}\\
-%% %% t | e &\sim \text{U}\{ \text{tasks solved by }e\}.
-%% %% \end{align}
-
-%% %% Then $p_t(t) = \sum_{e_j \in \mathcal{L} \text{ s.t. } t(e_j) = 1}
-%% %% p_\mathcal{L}(e_j)$. If we make the assumption that for the set $E$ of
-%% %% expressions in this sum, there is some expression $e^*$ such that
-%% %% $p_\mathcal{L}(e^*) >> p_\mathcal{L}(e), e \neq e^*$, then we have, approximately, 
-%% %% \[ p_t(t) \propto} p_\mathcal{L}(e^*). \]
-
-%% %% This suggests a way to find the ordering $O$: infer a stochastic
-%% %% grammar $G$, under the assumption that $G$ favors one solution to a
-%% %% given task over any other, and order the elements of $\mathcal{L}$
-%% %% according to $G$.
-
-%% %% Thus, our basic procedure will be to search for a single ``good''
-%% %% solution for as many tasks as we can, and to use those solutions to
-%% %% estimate a stochastic grammar. Our search will be performed in the
-%% %% order of the most recently inferred grammar. So as our ordering
-%% %% becomes more efficient and we can solve more tasks, we learn better
-%% %% stochastic grammars.
-
-%% %% Formally then, our goal is to solve the following minimal feedback
-%% %% multi-task program induction problem: suppose we are given set of
-%% %% \emph{tasks} $T=\{t_k\}_{k=1}^K$ where each task is a function $t_i :
-%% %% \mathcal{L} \rightarrow \{0, 1\}$, where $\mathcal{L}$ is the set of
-%% %% programs in our language.  We say that $ p \in \mathcal{L}$
-%% %% \emph{solves} $t_i$ if $ t_i(p) = 1$. Our goal is to find a set of
-%% %% solutions $X = \{x_k\}_{k=1}^k \subset \mathcal{L}$ such that
-
-%% %% \[
-%% %% X = \argmax_{x_1, \dots, x_K \in L} \sum_k t_k(x_k).
-%% %% \]
-
-%% Based on the motivation set out above, we present an iterative
-%% strategy. Given a stochastic grammar $\mathcal{G}$ which defines a
-%% distribution over the primitives from $\mathcal{L}$ and a integer
-%% \emph{frontier size} $N$, we perform two steps in each iteration.  In
-%% the first step, we explore the $N$ highest probability programs
-%% according to $\mathcal{G}$. In the second step, we compress the
-%% programs that solve any of our tasks, and we use that compression to
-%% define a new grammar. Repeating these two steps, we effectively
-%% reparameterize our search space to include programs that are more likely
-%% to be solutions to our tasks.
-
-The remainder of this paper will include two major sections. In the
-first, we describe in detail the E.C. algorithm and design choices we
-make in implementing it. We will then present results from two
-experiments in the domains of symbolic regression and Boolean
-function learning.
-
-%% \section{}
-%% Let $\mathcal{F}$ be a family of distributions over $\mathcal{L}$ such
-%% that for any $\mathcal{D} \in \mathcal{F}$ we can a) evaluate the
-%% probability of program $p$, $P_D(p)$, and we can b) enumerate elements
-%% from $D$ in descending order.
-
-%%  To motivate a solution to our problem, we begin with the hierarchical
-%% Bayesian formulation of multi-task program induction found in Liang et
-%% al.~\cite{liang10programs}. The general idea there is to estimate
-%% solution programs for a set of related problems by inferring the posterior
-%% distribution over solutions under the assumption that all the solution
-%% programs are generated from a common latent stochastic grammar. By
-%% jointly inferring solutions and the latent grammar, we find a set of
-%% solutions that favors reuse of subcomponents.
-
-%% While this framework is appealing in its elegance and Bayesian
-%% interpretation, it cannot be straightforwardly adapted to the
-%% situation we are considering here because their proposed MCMC
-%% algorithm relies on a local log-likelihood signal, the very kind that
-%% we are assuming we do not have access to.
-
-%% To begin, we provide a high level overview of the algorithm and the
-%% problem that it solves. The problem setup requires a representation
-%% language for programs and a family of distributions over expressions
-%% in that language, but we leave these details for the following section
-
-\section{Extracting concepts from programs}
-
-Our goal is to solve the following minimal feedback multi-task program
-induction problem: suppose we are given a set of \textbf{tasks}
-$T=\{t_k\}_{k=1}^K$ where each task is a function $t_k : \mathcal{L}
-\rightarrow \{0, 1\}$ and where $\mathcal{L}$ is the set of
-expressions in our language.  We say that expression $ e \in
-\mathcal{L}$ \textbf{solves} $t_k$ if $ t_k(e) = 1$. Our goal is to
-solve as many of the tasks in $T$ as we can.
-
-We propose the \textbf{E.C. algorithm}, an iterative procedure to solve this
-problem. It will be useful in presenting the algorithm to define the
-following terminology: a \textbf{frontier} is the finite set of
-expressions that the algorithm considers in any one iteration. The
-\textbf{frontier size} $N$ is the prespecified number of elements in
-the frontier. We will say that a task is \textbf{hit} by the frontier
-if there is a expression in the frontier that solves it.
-
-The E.C. algorithm maintains a distribution $\mathcal{D}$ over
-expressions in $\mathcal{L}$. At each iteration, it
-\begin{enumerate}
-\item \textbf{explores} the frontier -- enumerating the $N$ most probable expressions
-  from the current distribution $\mathcal{D}$ -- and
-\item \textbf{compresses} the hit tasks in the frontier to estimate a new
-  distribution.
-\end{enumerate}
-
-\subsection{Representing typed functional programs as binary trees}
-
-Following~\cite{liang10programs} and~\cite{briggs2006functional}, we use a
-simply typed combinatory logic -- a variable-free subset of the
-polymorphic simply-typed lambda calculus~\cite{Pierce_2002} -- as our
-program representation language. In short, the combinatory logic
-introduces a \emph{basis} of several primitive \emph{combinators} such
-that any function can be written as applications of the basis
-combinators. Some common basis combinators are defined as follows:
-\begin{align}
-\mathbf{I}\; x &\rightarrow  x & \text{ (identity) }\\
-\mathbf{S}\; f\; g\; x\; &\rightarrow (f\; x)\; (g\; x)\\ 
-\mathbf{C}\; f\; g\; x\; &\rightarrow (f\; x)\; g  \\ 
-\mathbf{B}\; f\; g\; x\; &\rightarrow f\; (g\; x) & \text{ (composition) }
-\end{align}
-
-The basis combinators are theoretically enough to express any Turing
-computable function, but we will assume that our language has a
-variety of primitives and, together with the basis combinators, we
-call these the \textbf{primitive combinators}. Note that by default we consider
-all primitives to be in their ``curried'' form (e.g. a function like
-``+'' adds two numbers $x$ and $y$ by being first applied to $x$,
-returning the function $(+\;x)$ and then being applied to $y$,
-returning $(+\,x\,y)$\;).
-
-The basis combinators can themselves be expressed in the lambda
-calculus (with definitions following directly from the equations
-above). The lambda calculus has two basic operations -- application
-and abstraction -- but in using the
-combinatory logic we sequester uses of the abstraction operation
-inside the combinators. In doing so, we have replaced the variable
-binding of the $\lambda$ operator with the variable \emph{routing} of
-these basis combinators. Our representation thus becomes
-variable-free. See \cite{liang10programs} for a more
-detailed discussion of this routing interpretation.
-
-Using combinatory logic is very convenient for program
-synthesis~\cite{briggs2006functional}: since every expression is the
-application of one expression to another -- with this recursion
-bottoming out at the primitive combinators -- each program is a binary
-tree. Most importantly, any subtree is itself a well-formed
-expression; this is not the case in the lambda calculus, since
-abstraction introduces long range dependencies between the $\lambda$
-operator and the variables to which that operator refers. In the
-lambda calculus, then, a subtree might have free variables, not bound
-to any enclosing $\lambda$.
-
-As a simple example of our representation, consider the squaring
-function $\lambda x. * x\; x $. Using two of the basis combinators
-above, we can write the squaring function in combinatory logic as $S\; *\;
-I$ (taking $*$ as a primitive), where application associates to the
-left, so we drop the parentheses. When we apply this combinator to a
-value $x$, the action of the combinator is defined as $S \;*\; I\;
-x~\rightarrow~(*\; x)\; (I\; x)~\rightarrow *\; x \; x$.
-
-We can extend this representation with a simple polymorphic type
-system~\cite{Pierce_2002}. In this representation, a type is either a
-type primitive (e.g. reals, integers, Booleans, etc.), a type variable
-(e.g. $\sigma$), or a function type $ \tau_1 \rightarrow \tau_2$ of
-functions from source type $\tau_1$ to target type $\tau_2$. Any combinator
-can be represented as a binary tree~(Figure~\ref{fig:clbintree}) whose
-leaves are typed primitive combinators and whose interior nodes
-represent typed applications of combinators (we annotate interior
-nodes with their types).
-
 \begin{figure}
-%%  \centering
-  \begin{subfigure}[Before]{0.45\linewidth}
-    \begin{tikzpicture}[scale=.8,
-        transform shape, level/.style={sibling distance=5mm/#1, font=\scriptsize}
-        ]
-      \node [font=\scriptsize] (top) {$\mathbb{R}$}
-      child {node [above=.3cm, left=.1cm, text width=1.8cm] (a) {$(\mathbb{R} \rightarrow \mathbb{R}) \rightarrow \mathbb{R} 
-          \rightarrow \mathbb{R}$}
-        child {node [left=.5cm, below=1cm, text width=3.0cm] 
-          {$ \mathbf{S} : (t_2 \rightarrow t_1 \rightarrow t_0)  \rightarrow (t_2 \rightarrow t_1) 
-            \rightarrow$ \\ $t_2 \rightarrow t_0$}}
-        child {node [right=0cm, text width=2cm] {$ \mathbf{*}: \mathbb{R} \rightarrow \mathbb{R} \rightarrow \mathbb{R}$}}
-      }
-      child {node [below right=-.5cm and .2cm] (b) {$\mathbb{R} \rightarrow \mathbb{R}$}
-        child {node [above=.5cm] {$ \mathbf{I}: t_0 \rightarrow t_0 \rightarrow t_0$}}
-      };
-    \end{tikzpicture}
-    \caption{    \label{fig:clbintree}}
-  \end{subfigure}
-  \begin{subfigure}[After]{0.5\linewidth}
-    \begin{tikzpicture}[scale=.7, transform shape, 
-        level/.style={sibling distance=10mm/#1},
-        every edge/.style={draw, thin}
-      ]
-      \node [circle,draw] (a) {}
-      child {node [left=0cm] (b) {$f_1$}}
-      child {node [left=.2cm] (c) {\dots} edge from parent[draw=none] }
-      child {node [left=.4cm] (e) {$f_n$}}
-      child {node [circle,draw, left=1cm] (d) {} edge from parent[very thick]
-        child {node [circle, draw, left=.25cm] (l) {} edge from parent[draw=none]
-          child {node [left=.5cm] (b) {$f_1$}} 
-          child {node [left=.3cm] (c) {\dots} edge from parent[draw=none] }
-          child {node [left=.3cm] (e) {$f_n$} edge from parent[very thick]}
-          child {node [circle, draw, left=.25cm] (and2) {}
-            child {node [circle, draw, left=.25cm](l2) {} edge from parent[draw=none] 
-              child {node [above=.5cm] {\vdots} edge from parent[draw=none]}
-            }
-            child {node [circle, draw] (r2)  {} edge from parent[draw=none] 
-              child {node [above=.5cm] {\vdots} edge from parent[draw=none]}
-            }
-          }
-        }
-        child {node [circle, draw, right=.25cm] (r) {} edge from parent[draw=none]
-          child {node [circle, draw, right=.25cm] (and3) {}
-            child {node [circle, draw ](l3) {} edge from parent[draw=none] 
-              child {node [above=.5cm] {\vdots} edge from parent[draw=none]}
-            }
-            child {node [circle, draw, right=.2cm ] (r3)  {} edge from parent[draw=none] 
-              child {node [above=.5cm] {\vdots} edge from parent[draw=none]}
-            }
-          }
-          child {node [right=.3cm] (b) {$f_1$} edge from parent[very thick]} 
-          child {node [right=.4cm] (c) {\dots} edge from parent[draw=none] }
-          child {node [right=.6cm] (e) {$f_n$}}
-        }
-      };
-
-      \path (d)  edge[very thick]  node[inner sep=0mm,pos=.4] (a1) {} node {} (l);
-      \path (d)  edge[very thick]  node[inner sep=0mm,pos=.4] (b1) {} node {} (r);
-      \path (a1) edge[bend right] (b1) ;
-
-      \path (and2)  edge node[inner sep=0mm,pos=.4] (t0) {} node {} (l2);
-      \path (and2)  edge node[inner sep=0mm,pos=.4] (k0) {} node {} (r2);
-      \path (t0) edge[bend right] (k0) ;
-
-      \path (and3)  edge node[inner sep=0mm,pos=.4] (t1) {} node {} (l3);
-      \path (and3)  edge node[inner sep=0mm,pos=.4] (k1) {} node {} (r3);
-      \path (t1) edge[bend right] (k1) ;
-    \end{tikzpicture}
-    \caption{  \label{fig:andor}}
-    \end{subfigure}
-  \caption{(a) The typed combinator $S * I$ ($f(x)=x^2$ over real values ($\mathbb{R}$)) represented as a
-    binary tree. (b) The space of all typed combinators represented as an
-    infinitely deep AND/OR tree.}
-
-\end{figure}
-
-\subsection{A stochastic grammar over programs}
-\label{sec:stochgrammar}
-
-We specify a distribution $\mathcal{D}$ over expressions in
-$\mathcal{L}$ as a stochastic
-grammar~\cite{feldman1969grammatical}. Many stochastic grammars have
-the desirable property that the probability of an expression is the
-product of the probabilities of its subcomponents. The distribution we
-specify will be a simple version of this, in which the
-probability of an expression is the product of the probability of its
-leaves.
-
-Let ${C = c_1, \dots, c_N }$ be the primitive combinators in
-$\mathcal{L}$. ${ \mathcal{D}=p_1, \dots, p_N }$ will associate with
-each primitive combinator $c_i$ a prior probability $p_i$,
-$0~\leq~p_i~\leq~1, \sum_i~p_i~=~1 $. The probability of a leaf with
-primitive combinator $c_i$ will depend on which other primitive
-combinators could have been chosen in its place. This set will depend
-on the type that was requested by the parent of this leaf when the
-leaf's primitive was chosen; we call this type the \emph{requesting}
-type of the leaf. For example, suppose that we have the primitive
-$(+1)$ of type $Int \rightarrow Int$ and the primitive $NOT$ of type
-$Bool \rightarrow Bool$. If the requesting type is $\sigma \rightarrow
-Int$, then we can only use $(+1)$, but if the requesting type is
-$\sigma \rightarrow \sigma$ then we could use either $(+1)$ or
-$NOT$. We define $C_\tau$ for any type $\tau$ to be the set of
-primitive combinators in $C$ that are consistent with $\tau$, that is,
-whose types \emph{unify} with $\tau$ (we use unification according to
-the standard type theory presentation~\cite{Pierce_2002}).
-
-Let $C_e$ be the set of primitive combinators in the leaves of
-expressions $e$. We define the probability of an expression ${e \in
-  \mathcal{L}}$ to be ${ p(e)~=~\prod_{c \in C_e} p(c | \tau(c)) }$
-where $p(c | \tau(c))$ is the probability of using primitive $c$ when
-the requesting type is $\tau(c)$. In turn, we define the conditional
-probability for each combinator $c_n$ to be $p(c_n| \tau(c_n)) \propto
-\frac{p_n}{\sum_{c_j \in C_{\tau(c_n)}} p_j}$.  That is, $p(c_n | \tau(c_n)$ is
-proportional to the probability of sampling $c_n$ from the multinomial
-distribution defined by the $p_n$'s, conditioned on the requesting
-type, with the constant of proportionality chosen to ensure that the
-sum over the probabilities of all expressions converges to a finite
-value (so that the distribution over expressions is proper). For
-binary trees, this is true whenever the constant of proportionality is
-less than~$\frac{1}{4}$\footnote{The probability mass $M_d$ of all
-  expressions of depth less than or equal to $d$ can be written with
-  the recurrence relation $M_d \leq M_{d-1}^2 + M_1$. This has an
-  asymptotic fixed point, as $d \rightarrow \infty$ if $x = x^2 + M_1$
-  has a solution, which is true if $M_1~\leq~1/4$.}.
-
-Now suppose that we have a set of expressions $\mathcal{E}$, and we
-wish to estimate the maximum likelihood values of the $p_n$'s. If the
-choice of primitive at each leaf were not conditioned on the
-requesting type, we could just count all occurrences of each primitive
-in our expression set, and this would be proportional to the maximum
-likelihood estimate. This is the ML estimator for a multinomial
-distribution. However, the fact that we condition on the requesting
-type makes this a considerably more difficult optimization task. One
-of the simplest approximations we can make is to calculate the
-frequency with which each primitive combinator appears in
-$\mathcal{E}$ when the requesting type is such that it could have
-been chosen. This is a straightforward calculation, and
-we use it for its convenience, finding that our empirical results
-justify its use. Future inquiry will be needed to determine to what
-extent a more accurate estimator can improve these results.
-
-\subsection{Best-first enumeration of programs}
-In order to explore the frontier of most promising programs, we need a
-procedure that enumerates the $N$ most probable expressions. There has
-been recent interest in this problem, most of it focused on
-enumerating the shortest program satisfying some
-criterion~\cite{DBLP:conf/sfp/Katayama05,DBLP:conf/aaip/YakushevJ09}. Our approach is to formulate the
-problem as a best-first exploration of an AND/OR tree
-(Figure~\ref{fig:andor})~\cite{nilsson1982principles,DBLP:journals/cacm/Hall73}. In
-words, the enumeration procedure is best described by the following
-recursion: every program is either a primitive combinator OR it is a
-left child program applied to a right child program (that is, a left
-child AND a right child).
-
-This can framed as the following AND/OR tree $\mathcal{G}$. Suppose we
-want a program of type $\tau$. The root of $\mathcal{G}$ is an OR node of
-type $\tau$. Its children are elements in $C_\tau$ (defined in the previous
-section) and one AND node of type $\tau$ with two children. Each child of
-an AND node has the same structure as the root node, with modified
-types: the type of its left child is $\sigma \rightarrow \tau$ (where $\sigma$ is
-a fresh type variable). Since the type of the right child is
-constrained by the subtree rooted at the left child, we always expand
-the left child first. Once we have a complete left child program, we
-can use its target type as the type of the right child.
-
-Recall that a valid partial solution $\mathcal{H}$ is a subtree of
-$\mathcal{G}$ satisfying the following properties: the root of
-$\mathcal{H}$ is the root of $\mathcal{G}$ and any OR node in
-$\mathcal{H}$ has at most one child. $\mathcal{H}$ is a complete
-solution (and thus a complete expression) if it is a partial solution
-whose leaves are leaves of $\mathcal{G}$, if each OR node in
-$\mathcal{H}$ has exactly one child, and if each AND node in
-$\mathcal{H}$ is connected to all of that node's children in
-$\mathcal{G}$.
-
-The value of a partial solution $\mathcal{H}$ is calculated by summing
-the value of its leaves, where each such value is the log of the
-probability assigned to that leaf in
-Section~\ref{sec:stochgrammar}. If a leaf of $\mathcal{H}$ is a leaf
-of $\mathcal{G}$ then either it is an AND node, to which we assign a
-value of 0, or it corresponds to some primitive combinator $c_n$ in
-the library and is the child of some typed OR node with type $\tau$. We
-assign it a value of $\log{p(c_n | \tau)}$. If a leaf $v$ of
-$\mathcal{H}$ is an OR node, then the value of any extension of it to
-a complete solution is at most $log(1/4)$. Thus, we have defined a
-value for partial solutions that upper bounds the value of any of its
-extensions.
-
-Our best-first policy is to expand the next partial solution
-with the highest value. If a partial solution has multiple leaves that
-are not primitive combinators, we expand the left most one first,
-since it is the choice of left combinator that defines the type of its
-right combinator in an application pair. That is, the parent node
-applies its constraint on valid solutions via the left child, not the
-right child. We want to apply constraints as early as possible, so
-expanding left child first prunes more partial solutions.
-
-\subsection{Finding the most compressive set of solutions}
-
-Having enumerated the frontier, we want to assign solutions
-to hit tasks such that this set is maximally compressible. This will
-promote reusable modular subtrees.
-
-A natural metric for the compressibility of a set of binary trees is
-the number of unique subtrees (counting leaves) that this set
-contains. More sophisticated metrics exist; we could use the negative
-logarithm of the probability of the solutions under a suitable
-probability distribution~\cite{DBLP:journals/tit/BarronRY98}. Future
-work will explore whether this sophistication is worthwhile and which
-prior on grammars makes sense. For our purposes here, penalizing
-according to the number of unique subtrees appropriately rewards reuse
-of useful subtrees. As an example of this, suppose we have six tasks
-whose solutions are integers $4^2, 5^2, 6^2, 7^2$, and $8^2$, and
-suppose our frontier contains the $10$ expressions:
-\vspace{-.3cm}
-\captionsetup[subfigure]{labelformat=empty} 
-\begin{figure}[h!]
-   \begin{subfigure}{.17\linewidth}
-     \begin{tikzpicture}[scale=.7, background rectangle/.style={fill=lightgray}] 
+  \begin{subfigure}{.3\linewidth}
+     \begin{tikzpicture} 
        \tikzset{level distance=15pt, sibling distance=10pt}
@@
   \end{subfigure}
-  \begin{subfigure}{.17\linewidth}
-     \begin{tikzpicture}[scale=.7]
+  \begin{subfigure}{.3\linewidth}
+     \begin{tikzpicture} 
        \tikzset{level distance=15pt, sibling distance=10pt}
@@
   \end{subfigure}
-   \begin{subfigure}{.17\linewidth}
-     \begin{tikzpicture} [scale=.7]
+  \begin{subfigure}{.3\linewidth}
+     \begin{tikzpicture} 
        \tikzset{level distance=15pt, sibling distance=10pt}
@@
   \end{subfigure}
-   \begin{subfigure}{.17\linewidth}
-     \begin{tikzpicture}[scale=.7] 
-       \tikzset{level distance=15pt, sibling distance=10pt}
-       \Tree [[  * 7 ] 7 ]
-     \end{tikzpicture}
-  \end{subfigure}
-  \begin{subfigure}{.17\linewidth}
-     \begin{tikzpicture}[scale=.7] 
-       \tikzset{level distance=15pt, sibling distance=10pt}
-       \Tree [[  * 8 ] 8 ]
-     \end{tikzpicture}
-  \end{subfigure}
-  \vspace{2mm}
 
-  \hdashrule{.9\linewidth}{1pt}{1pt}
-
-  \vspace{2mm}
-  \begin{subfigure}{.17\linewidth}
-     \begin{tikzpicture}[scale=.6] 
+  \begin{subfigure}{.3\linewidth}
+     \begin{tikzpicture} 
        \tikzset{level distance=15pt, sibling distance=10pt}
@@
      \end{tikzpicture}
-     \caption{(16)}
   \end{subfigure}
-  \begin{subfigure}{.17\linewidth}
-     \begin{tikzpicture}[scale=.6]
+   \begin{subfigure}{.3\linewidth}
+     \begin{tikzpicture} 
        \tikzset{level distance=15pt, sibling distance=10pt}
@@
      \end{tikzpicture}
-     \caption{(25)}
   \end{subfigure}
-   \begin{subfigure}{.17\linewidth}
-     \begin{tikzpicture} [scale=.6]
+  \begin{subfigure}{.3\linewidth}
+     \begin{tikzpicture} 
        \tikzset{level distance=15pt, sibling distance=10pt}
@@
      \end{tikzpicture}
-     \caption{(36)}
-  \end{subfigure}
-  \begin{subfigure}{.17\linewidth}
-     \begin{tikzpicture}[scale=.6] 
-       \tikzset{level distance=15pt, sibling distance=10pt}
-       \Tree [[[  S * ] I ] 7 ]
-     \end{tikzpicture}
-     \caption{(49)}
-  \end{subfigure}
-  \begin{subfigure}{.17\linewidth}
-     \begin{tikzpicture}[scale=.6] 
-       \tikzset{level distance=15pt, sibling distance=10pt}
-       \Tree [[[  S * ] I ] 8 ]
-     \end{tikzpicture}
-     \caption{(64)}
   \end{subfigure}
 \end{figure}
-\captionsetup[subfigure]{labelformat=parens} 
-\vspace{-.3cm}
-
-\noindent The top row consists of one type of representation of an integer
-squared, namely, the integer times itself. The bottom row contains
-another type of representation in which the function $f(x)=x^2$,
-represented in combinatory logic as $S * I$, is applied to the
-integer. The top and bottom row of each expression evaluate to the
-same integer. If we consider only the first column, the top tree has 4
-unique subtrees and the bottom has 7. But each column adds 3 unique
-subtrees to the top row and only 2 to the bottom. So by the fifth
-column the top row has more unique subtrees than the bottom row. The
-increased compressibility of the bottom row is due to the fact that
-the squaring operation has been factored out as an abstraction. This
-is an example of how this compressibility metric favors
-abstraction when multiple expressions are being considered at once.
-
-We want to choose a solution for each hit task such that the complete set
-of such solutions for all solved tasks is as compressible as
-possible. That is, let $t_i$ be a task in the set of hit tasks, and
-let $\{e^k_i\}$ be the expressions in the frontier that solve
-it. Then our goal is to find
-\[
-\{e^*_i\} = \argmin_{\{e^k_i\}} | \cup_i e^k_i |,
-\]
-where $|\cdot|$ denotes the number of unique trees in a set of expressions.
-
-However, an exact optimization of this problem requires examining
-$O(M^K)$ assignments, where $M$ is the maximum number of solutions for
-any one task, and $K$ is the number of tasks. Since this is
-prohibitive, we relax the compressibility metric as follows: we define
-the cost of adding each new solution $s_n$ to a partial set of
-solutions $s_1, \dots, s_{n-1}$ to be the number of additional unique
-subtrees in $s_n$ that are not present in $s_{n-1}$. That is, the
-penalty of adding a new expression to the set of expressions depends
-only on the last expression we added. Our solution thus becomes
-approximate, and that approximation is order-dependent, but a
-depth-first search on the defined search space goes from exponential
-in the solution degree to quadratic ($O(KM^2)$).
-
-However, when the number of solutions in the frontier grows, this
-quadratic computation becomes prohibitive. Therefore, in practice, we
-bound the number of potential expressions we are willing to consider
-at any one iteration to any one task to a small number. We find that a
-setting of this bound to 2 is computationally feasible and large
-enough to guide subcomponent discovery in a reliable way. It may be
-necessary to increase this bound as we scale this algorithm to larger
-problems.
-
-\subsection{Re-estimating the stochastic grammar}
-
-Given a set of chosen solution expressions, which we will call the
-\emph{solution set}, want to re-estimate our stochastic grammar. Our
-inspiration for doing this is the Neville-Manning algorithm for
-grammar-based compression~\cite{nevill1997identifying}. The idea of
-that compression algorithm is to compress any string by creating a
-grammar for that string such that a) no two symbols appear more than
-once in the grammar and b) every rule is used more than once. From a
-set of expressions, we generate a grammar according to these criteria.
-
-%% To do this, we traverse a set of combinator binary trees in
-%% depth first order. We count the occurrences of every subtree. Every
-%% time we encounter a subtree for the \emph{second} time, we designate
-%% it as a primitive in our new grammar. This ensures that every
-%% primitive is used more than once. During our traversal, we do not
-%% descend into trees that have been added already to the new
-%% grammar. Thus, we only count an instance of a tree if that instance is
-%% not a subtree of a primitive element of the new grammar.
-
-This procedure generates a parse of the solution set with a new set of
-primitives, and we keep count of how many times each primitive
-occurred in the solutions set. To estimate the probabilities
-associated with each node production, we again traverse the solution
-set, but this time for each node we count which other primitive
-elements from the new grammar could have been used. In accordance with
-Section~\ref{sec:stochgrammar}, we estimate the terminal production weight
-to be the ratio of the number of times that node was used in the first
-traversal divided by the number of times it was used in the second
-traversal.
-
-\section{Experiments.}
-\subsection{Symbolic regression.}
-\label{sec:symreg}
-
-We first illustrate the performance of our learning algorithm on a
-symbolic regression problem. For each task in this problem, we want to
-find an algebraic function, symbolically specified, that maps a set of
-input values to output values. We choose this problem because it has a
-long history in AI, particularly in the genetic programming
-literature~\cite{DBLP:books/daglib/0070933}.
-
-In the formulation we consider, each task $t$ corresponds to a
-polynomial $f$, and an expression $e$ solves $t$ if it returns the
-same value when applied to $i \in 0, \dots, 9$ that $f$ does. In the
-experiment shown here, the set of problems corresponds to the set of
-all polynomials with degree less than or equal to 2. In the initial
-library, we include the basic combinators I, S, B, C and four
-arithmetic primitives $1$, $0$, $*$, and $+$. The choice of initial
-weights on these primitives can make a difference, as it determines
-the relative ordering of combinators in the enumeration step. To get a
-general sense of the algorithm's performance, we set the initial
-weights to be slightly different on each run, perturbing them around a
-uniform weighting.
-
-\setcounter{subfigure}{0} 
-\begin{figure}[ht]
-\centering
-\includegraphics[width=1.0\linewidth]{figures/symreg3frontiers.pdf}
-\caption{Learning curves as a function of frontier size.  As frontier
-  size is increased, curves plateau closer to $100\%$ performance. A
-  baseline search over $150000$ expressions only hits $3\%$ of the
-  tasks (dashed pink line). \label{fig:symregLearning}}
-\end{figure}
-\begin{figure}[ht]
-\includegraphics[width=1.07\linewidth]{figures/symRegDiffCurricula.pdf}
-\caption{How do different task sets affect learning curves? Learning
-  curves at frontier size of $10000$ for different task
-  sets.} 
-\label{fig:symregCurricula}
-
-\end{figure}
-
-%% Why do the learning curves asymptote early? One might object that a
-%% learning algorithm should continuously improve, if only slowly, as it
-%% is given more time. But in these runs we observe a finite and
-%% permanent performance cap. This is expected: once the learner find a
-%% set of concepts it can explain, if those concepts fill up the
-%% frontier, then there is nothing to spur the learner to incorporate as
-%% yet unsolved tasks.
-
-%% To investigate whether this is in fact the case, we can examine the
-%% proportion of expressions in the frontier that match tasks in our task
-%% set; that is, is the learning curve plateauing because the frontier is
-%% heavily redundant, or is it plateuing because the frontier contains
-%% many expressions that do not hit the task set at
-%% all? Figure~\ref{fig:plateau_investigation} shows the proportion of
-%% the expressions in the frontier that hit a task as the frontier size increases.
-
-In Figure~\ref{fig:symregLearning}, we show performance results as we
-vary the frontier size. Changing the frontier size changes the number
-of tasks the algorithm identifies correctly over the course of 15
-algorithm iterations (all experiments in this work were run for 15
-iterations). As a baseline, we enumerated $10000*15=150000$
-expressions from the initial grammar. This is the total number of
-expressions that a run of the algorithm sees if it has a frontier size
-of $10000$ and runs for $15$ iterations. Thus, if the benefit accruing
-to the runs with larger frontier sizes is simply due to an increase in
-the number of expressions seen, rather than increased learning, we
-should see similarly good performance from this baseline run. In fact,
-however, the baseline run only hits $27$ of the $1000$ tasks ($3\%$),
-whereas our algorithm nears $100\%$ for a frontier of $10000$.
-
-What does the E.C. algorithm learn in this task?  Inspecting the top
-weighted primitives of the final grammars, we find many incrementers
-(e.g. (+1), (+2), etc.), several versions of expressions that
-translate to functions like $x*f(x)$ and $f(x*f(x))$, and combinations
-of these, like $x*(x+3)+3$ and $x*(x+2)$. Informally, we see
-expressions that apply functions to themselves, building up complex
-functions with relatively few unique primitives.
-
-To what degree is ``conceptual bootstrapping'' responsible for the
-improved performance. That is, to what extent does the presence of
-simple problems account for the ability to learn complex functions?
-One hypothesis is that to succeed on a set of tasks, the set must
-contain a ``learnable'' curriculum, a set of tasks that serve as
-stepping stones from an impoverished representation to a rich one. We
-can test this hypothesis by varying the set of tasks to which the E.C.
-algorithm is exposed. If it is true, then we should see a nonlinear
-response to reducing the number of easy tasks, as the curriculum
-changes from being learnable to being unlearnable.
-
-In Figure~\ref{fig:symregCurricula}, we present learning curves
-(frontier size 10000) corresponding to various versions of our
-original symbolic regression task set. Recall that the original set
-consisted of all polynomials of degree two or less and with
-coefficients between 0 and 9. In the ``no constant'' task set, we
-remove the 9 constant functions in this set. In the ``no linear'' task
-set, we remove 90 linear functions from the set. We observe that
-performance on those task sets does not decrease. However, when we
-remove both the linear functions and the constant function (``only
-quadratics''), we see a sharp drop in the algorithm's
-performance. When we further restrict the task set to only ``complex''
-quadratics (which we define as quadratics whose
-coefficients are greater than zero), we observe another comparable
-drop in performance. When we go one step further and restrict the task
-set to quadratics with coefficients greater than 1, performance drops
-to 0 because no task is hit in the initial frontier.
-
-This data has in it the nonlinearity that we predicted -- that a
-minimal curriculum of simple tasks is sufficient to achieve high
-performance -- but also suggests that, at least in this domain, this
-is not an all-or-none effect. Though the performance drops
-significantly once no linear functions are included in the task set,
-learning does still occur, and the learning curves still have the same
-basic shape.
-
-%% In addition, we have a
-%% baseline for each curve, visits the same number of expressions as the
-%% corresponding algorithm run (i.e. frontier size $\times$ $\#$ of
-%% iterations) but all at once; it does not do any learning. Thus, this
-%% baseline serves as a brute-force control for the learning aspect of
-%% the algorithm.
-
-\subsection{Boolean function learning}
-\label{sec:boolcircuits}
-
-\begin{figure}[t]
-\includegraphics{./figures/tablealonedoc.pdf}
-\caption{Some learned primitives from the circuit based Boolean
-  function learning experiment. We highlight compression using $NOT$,
-  $E_1$ and $E_2$. \,(a)The three elementary gates used to generate
-  the task set. (b) Two higher-order primitives in the induced
-  grammar. Note how $E_1$ can be used to define FALSE by taking the
-  NOT function as an argument. (c) Additional common Boolean functions
-  found in the learned grammar.  \label{table:booleanCircuits}}
-\end{figure}
-
-As another demonstration of our approach, we investigate the
-E.C. algorithm's ability to learn Boolean functions using only the logical
-primitive NAND and the basic combinators. It is well known that a
-Boolean circuit can be constructed for any Boolean function given only
-the NAND gate. Here, the basic combinators take the place of the
-wiring normally found in a circuit.
-
-We constructed two task sets. The first of these was constructed
-explicitly to contain familiar modular structure. In this set of
-tasks, we sampled $1000$ Boolean circuits using AND, OR, and NOT
-gates. To accomplish this, we first sampled either 1, 2 or 3 inputs,
-then between 1 and 5 gates, randomly wiring the inputs of each new
-gate to the output of one of the existing gates in the circuit. We
-continued this sampling procedure until we had 1000 ``connected''
-circuits, i.e., circuits all of whose outputs are wired to an input
-except for the last one (the output gate). This resulted in $1000$
-tasks consisting of $82$ unique Boolean functions, with a distribution
-of tasks as shown in Figure~\ref{fig:booldistr}(a). In
-Figure~\ref{fig:boolLearningCurvesCircuits}, we present learning
-curves for frontier sizes 100, 500 and 10000, using the same procedure
-as in Section~\ref{sec:symreg}. In Figure~\ref{fig:booldistr}, we show
-that the distribution over Boolean functions enumerated from the
-grammar over expressions is much more similar to the true function
-distribution after learning than before.
-
-\begin{figure}
-\includegraphics[width=\linewidth]{figures/circuitDistrComp2.png}
-\caption{Distribution of Boolean functions among (a) the set of 1000
-  circuits randomly constructed from AND, OR and NOT gates (see text);
-  b) the first 1000 expressions from the grammar over expressions
-  \emph{before} learning; c) the first 1000 expressions from the
-  grammar over expressions \emph{after} learning. Gray levels indicate
-  percent of instances (circuits/expressions) which evaluate to that
-  function.\label{fig:booldistr}}
-\end{figure}
-
-This problem is particularly suitable for inspecting the constituents
-of the induced grammar $\mathcal{G}$. We might hope that our algorithm
-recovers the constituent logic gates that were used to build up the
-tasks. Table~\ref{table:booleanCircuits} shows a few of the top ranked
-expressions in the library. These include the basic logic gates; we
-also include two higher-order expression $E_1$ and $E_2$, to stress
-that the representation the E.C. algorithm is using allows it to build
-concepts that are more expressive that just sub-circuits.
-
-In a second experiment, we include all $272$ Boolean truth tables with
-three or fewer inputs as tasks. This is a more difficult
-problem. There are two kinds of learning that the E.C. algorithm might
-accomplish: first, many expressions in $\mathcal{L}$ map to a single
-Boolean function, so it needs to learn primitives that allow it to
-span many different functions instead of generating very simple
-functions redundantly. The second kind of learning involves the
-distribution of the functions themselves. In the circuit based Boolean
-experiment, that structure is apparent in the distribution in
-Figure~\ref{fig:booldistr}. In this second experiment, we remove the
-second kind of structure. In
-Figure~\ref{fig:boolLearningCurvesUniform}, we show 10 runs of the
-E.C.  algorithm on the second experiment with a frontier size of
-$2000$: note how there are several local minima that the majority of
-the runs get stuck in with performance around $50\%$, but several of
-the runs seem to take different trajectories to more successful
-representations.
-
-\vspace{-.3cm}
-\begin{figure}[h]
-\begin{subfigure}{.1\linewidth}
-\caption{\label{fig:boolLearningCurvesCircuits} }
-\end{subfigure}
-\begin{subfigure}[Before]{0.9\linewidth}
-\includegraphics[width=1.0\linewidth]{figures/booleancircuits3frontiers.pdf}
-\end{subfigure}
-
-\vspace{-.32cm}
-\begin{subfigure}{.1\linewidth}
-\caption{\label{fig:boolLearningCurvesUniform}}
-\end{subfigure}
-\begin{subfigure}[After]{.9\linewidth}
-\includegraphics[width=1.0\linewidth]{figures/booleanfunctionsUniform.pdf}
-
-\end{subfigure}
-\caption{Learning curves from Boolean function learning
-  experiments. (a) Learning curves for various frontier sizes on task
-  set of circuit based Boolean functions. (b) 10
-  learning curves for frontier size of 2000 on a task set consisting
-  of all Boolean functions of cardinality 3 or smaller. Note how most
-  of the trajectories get stuck in a local minimum around
-  $50\%$.\label{fig:boolLearningCurves}}
-\end{figure}
-
-\section{Conclusion}
-This work brings together ideas from various areas of research --
-hierarchical Bayes, learning to learn, minimum description length
-learning, and learning programs -- to provide a proof of concept for
-how bootstrap learning of abstract composable concepts can rapidly
-facilitate problem solving in the absence of domain specific
-knowledge. We show that this approach can learn successfully even in
-domains for which the solution space is not smooth and the error
-signal is all or none, where the only sense of locality to guide
-learning is the modular and compositional structure of the solutions
-that the algorithm itself builds.
-
-The problem presented here was chosen to highlight the vastness of the
-search space confronting an agent with minimal background knowledge:
-the reward function is binary, it is deterministic, and it cannot
-provide feedback about partial solutions. An important question left
-to be addressed is how to extend this algorithm to take into account
-continuous rewards, noisy data, and partial solutions.
-
-Our specific implementation of this approach involved a number of
-choices which should be explored further in future work.  Our claim
-here is not that the E.C. algorithm is optimal, but that it captures key
-features of how people solve the problem of learning new systems of
-concepts in acquiring deep domain expertise.  Much like a human
-researcher investigating a new area of study, our algorithm tries
-different models, watches them fail or succeed, and abstracts out the
-model elements that capture relevant differences.  It recombines and
-reuses successful patterns of reasoning that span many problems in a
-domain.
-
-At their best, human processes of discovery clearly go beyond
-what our algorithm is capable of.  The algorithm can get stuck in
-plateaus after a rapid initial increase in performance, due to its
-circular nature: it only inspects the region of the solution space
-that is typical of solutions already found.  Humans can become stuck
-in similar ways, but at least sometimes they realize that they are
-stuck and attempt to find another class of concepts for the remaining
-unsolved tasks.  Our algorithm requires, as human learners do, a
-sufficiently graded ``curriculum'', or spectrum of problems to solve --
-simple problems provide necessary stepping stones to building complex
-representations.  Also like humans, our algorithm can adaptively
-``self-pace'' its ways through the curriculum, figuring out which
-problems it is ready to tackle when, which problems are appropriately
-difficult given what it currently knows.  Yet people can sometimes
-more intelligently compose and modify the set of problems under
-consideration, designing their own stepping stones to expand the set
-of problems they can solve.  Making bootstrap learning systems smarter
-in these more human-like ways is a prime target for future research.
-
-%% The file named.bst is a bibliography style file for BibTeX 0.99c
-\bibliographystyle{named}
-\bibliography{ijcai13}
-%% \nocite{*}
 

### 2014-06-08 old/drafts/ijcai2013/ijcai13.tex
Moved everything into an a new 'old' directory.

--- 2013-04-22 drafts/ijcai2013/ijcai13.tex
@@
 \usepackage{amsmath} 
-\usepackage{amsfonts}
 \usepackage{latexsym} 
@@
  uninformative, providing perhaps a few bits of information but not
- enough to steer the learner effectively. How can such a learner ever
- get off the ground? A common intuition is that if the solutions to
- these problems share a common structure, and the learner can solve
- some simple problems by brute force, it should be able to extract
- useful components from these solutions and, by composing them,
- explore the solution space more efficiently. Here, we formalize this
- intuition, where the solution space is the space of typed functional
- programs and the gained information is stored as a stochastic grammar
- over programs. We propose an iterative procedure for exploring such
+ enough to steer it effectively. How can such a learner ever get off
+ the ground? A common intuition is that if the solutions to these
+ problems share a common structure, and the learner can solve some
+ simple problems by brute force, it should be able to extract useful
+ components from these solutions and, by composing them, explore the
+ solution space more efficiently. Here, we formalize this intuition,
+ where the solution space is the space of typed functional programs
+ and the gained information is stored as a stochastic grammar over
+ programs. We propose an iterative procedure for exploring such
  spaces: in the first step of each iteration, the learner explores a
@@
  first step to estimate a new stochastic grammar. We test this
- procedure on symbolic regression and Boolean circuit learning and
+ procedure on symbolic regression and boolean circuit learning and
  show that the learner discovers modular concepts for these
@@
 \section{Introduction}
-A central question for AI is how to create intelligent agents without
-building in vast amounts of pre-specified knowledge that is tailored
-to one domain or another. Any attempt to solve this problem must
-contend with the fact that the less an agent knows about solving
-problems in a domain, the worse its attempts are, and, therefore, the
-less informative they are about how to solve subsequent problems.
-
-For example, imagine you know nothing about electronics but are
-supplied with a collection of circuit elements and asked to build a
-useful device: your initial attempts will be uniformly
-ineffective. If, on the other hand, you know the basics, you are far
-more likely to generate devices that are somewhat successful and on
-whose basis you might improve your skills.
-
-Like many of the problems in AI, this one can be cast as a search
-problem: the goal is to find a circuit wiring procedure that maximizes
-certain criteria.  There are many search algorithms, but they typically
-exploit local or topological structure in the search space. For
-example, local search algorithms move from one candidate solution to a
-neighboring one. Heuristic search techniques take advantage of
-topological structure to prune off large parts of the solution space.
-
-Domain specific knowledge provides this structure. In local search, it
-defines the similarity metric between solutions; in heuristic search
-it provides the heuristic. Here we ask how to bootstrap domain
-specific knowledge and extract structure from the space of possible
-solutions by attempting to solve multiple related problems in a given
-domain at once. Our approach is motivated by the observation that many
-interesting problems -- such as discovering scientific theories,
-developing new technologies, learning to play a new instrument -- have
-solutions that are natrually specified as programs.
-
-Our proposed solution is thus inspired by the programmer who -- having
-reused subroutines to solve similar simple tasks -- creates a function
-library to help manage the complexity of larger projects. The
-E.C. algorithm, which we introduce here, builds a library of reusable
+Many of the problems we want to solve in AI are best cast as search
+problems. There are many search algorithms, but they typically exploit
+local or topological structure in the search space. For example,
+convex optimization and gradient based search methods take
+advantage of local smoothness to move in good directions. Local search
+algorithms move from one candidate solution to a neighboring
+one. Heuristic search techniques take advantage of topological
+structure to prune off large parts of the solution space.
+
+However, many of the problems we want solve -- discovering scientific
+theories, developing new technologies, figuring out how to bake -- do
+not come with such a structured search space. This is particularly
+true of problems, such as those mentioned, whose solutions can be
+naturally framed as programs. Here we ask how to extract structure
+from the space of programs by attempting to solve multiple related
+problems in a given domain at once.
+
+We take as inspiration the programmer who -- having reused several
+subroutines to solve different but similar simple tasks -- creates a
+function library to help manage the complexity of larger projects. The
+EC algorithm, which we introduce here, builds a library of reusable
 program components and places a distribution over these, effectively
-learning from simple tasks a search space structure that enables it to
-solve more complex tasks.
+learning from simple tasks a search space structure that enables it
+to solve more complex tasks.
 
@@
 
-Concretely, the E.C. algorithm works as follows. It begins with a small
+Concretely, the EC algorithm works as follows. It begins with a small
 library of program primitives and a distribution over programs
@@
 setting~\cite{DBLP:conf/ilp/KhanMP98}. We apply these ideas to
-learning functional programs and believe that the modularity and
+learning functional programs, and believe that the modularity and
 compositionality afforded by this representation is necessary for
@@
 The remainder of this paper will include two major sections. In the
-first, we describe in detail the E.C. algorithm and design choices we
+first, we describe in detail the EC algorithm and design choices we
 make in implementing it. We will then present results from two
@@
 
-We propose the \textbf{E.C. algorithm}, an iterative procedure to solve this
+We propose the \textbf{EC algorithm}, an iterative procedure to solve this
 problem. It will be useful in presenting the algorithm to define the
@@
 
-The E.C. algorithm maintains a distribution $\mathcal{D}$ over
+The EC algorithm maintains a distribution $\mathcal{D}$ over
 expressions in $\mathcal{L}$. At each iteration, it
@@
 system~\cite{Pierce_2002}. In this representation, a type is either a
-type primitive (e.g. reals, integers, Booleans, etc.), a type variable
+type primitive (e.g. reals, integers, booleans, etc.), a type variable
 (e.g. $\sigma$), or a function type $ \tau_1 \rightarrow \tau_2$ of
@@
           {$ \mathbf{S} : (t_2 \rightarrow t_1 \rightarrow t_0)  \rightarrow (t_2 \rightarrow t_1) 
-            \rightarrow$ \\ $t_2 \rightarrow t_0$}}
+            \rightarrow \\ t_2 \rightarrow t_0$}}
         child {node [right=0cm, text width=2cm] {$ \mathbf{*}: \mathbb{R} \rightarrow \mathbb{R} \rightarrow \mathbb{R}$}}
@@
         child {node [above=.5cm] {$ \mathbf{I}: t_0 \rightarrow t_0 \rightarrow t_0$}}
-      };
+      }
     \end{tikzpicture}
@@
 A natural metric for the compressibility of a set of binary trees is
-the number of unique subtrees (counting leaves) that this set
-contains. More sophisticated metrics exist; we could use the negative
-logarithm of the probability of the solutions under a suitable
-probability distribution~\cite{DBLP:journals/tit/BarronRY98}. Future
-work will explore whether this sophistication is worthwhile and which
-prior on grammars makes sense. For our purposes here, penalizing
-according to the number of unique subtrees appropriately rewards reuse
-of useful subtrees. As an example of this, suppose we have six tasks
-whose solutions are integers $4^2, 5^2, 6^2, 7^2$, and $8^2$, and
-suppose our frontier contains the $10$ expressions:
+the number of unique subtrees (counting leaves) that this
+set contains. As an example of this, suppose we have six tasks whose
+solutions are integers $4^2, 5^2, 6^2, 7^2$, and $8^2$, and suppose our
+frontier contains the $10$ expressions:
 \vspace{-.3cm}
@@
 \vspace{-.3cm}
-
-\noindent The top row consists of one type of representation of an integer
+The top row consists of one type of representation of an integer
 squared, namely, the integer times itself. The bottom row contains
@@
 once in the grammar and b) every rule is used more than once. From a
-set of expressions, we generate a grammar according to these criteria.
+set of expressions we generate a grammar according to these criteria.
 
@@
 polynomial $f$, and an expression $e$ solves $t$ if it returns the
-same value when applied to $i \in 0, \dots, 9$ that $f$ does. In the
+same value when applied $i \in 0, \dots, 9$ that $f$ does. In the
 experiment shown here, the set of problems corresponds to the set of
@@
 library, we include the basic combinators I, S, B, C and four
-arithmetic primitives $1$, $0$, $*$, and $+$. The choice of initial
+arithmetic primitives, $1$, $0$, $*$, and $+$. The choice of initial
 weights on these primitives can make a difference, as it determines
@@
 \centering
-\includegraphics[width=1.0\linewidth]{figures/symreg3frontiers.pdf}
-%\vspace{-.2cm}
+\includegraphics[width=.8\linewidth]{figures/symreg3frontiers.pdf}
+\vspace{-.2cm}
 \caption{Learning curves as a function of frontier size.  As frontier
@@
   baseline search over $150000$ expressions only hits $3\%$ of the
-  tasks (dashed pink line). \label{fig:symregLearning}}
-%\vspace{-.3cm}
+  tasks. \label{fig:symregLearning}}
+\vspace{-.3cm}
 \end{figure}
 \begin{figure}[ht]
-\includegraphics[width=1.07\linewidth]{figures/symRegDiffCurricula.pdf}
+\centering
+\includegraphics[width=.8\linewidth]{figures/symRegDiffCurricula.pdf}
 \caption{How do different task sets affect learning curves? Learning
@@
 \label{fig:symregCurricula}
-
+\vspace{-.5cm}
 \end{figure}
@@
 In Figure~\ref{fig:symregLearning}, we show performance results as we
-vary the frontier size. Changing the frontier size changes the number
+vary the frontier size.Changing the frontier size changes the number
 of tasks the algorithm identifies correctly over the course of 15
@@
 of $10000$ and runs for $15$ iterations. Thus, if the benefit accruing
-to the runs with larger frontier sizes is simply due to an increase in
-the number of expressions seen, rather than increased learning, we
-should see similarly good performance from this baseline run. In fact,
+to the runs with larger frontier sizes is due to simply an increase in
+the number of expressions seen, rather than increased learning, we should
+see similarly good performance from this baseline run. In fact,
 however, the baseline run only hits $27$ of the $1000$ tasks ($3\%$),
@@
 
-What does the E.C. algorithm learn in this task?  Inspecting the top
+What does the EC algorithm learn in this task?  Inspecting the top
 weighted primitives of the final grammars, we find many incrementers
@@
 stepping stones from an impoverished representation to a rich one. We
-can test this hypothesis by varying the set of tasks to which the E.C.
+can test this hypothesis by varying the set of tasks to which the EC
 algorithm is exposed. If it is true, then we should see a nonlinear
@@
 \includegraphics{./figures/tablealonedoc.pdf}
-%% \vspace{-.5cm}
+\vspace{-.5cm}
 \caption{Some learned primitives from the circuit based Boolean
@@
   found in the learned grammar.  \label{table:booleanCircuits}}
-%%\vspace{-.4cm}
+\vspace{-.4cm}
 \end{figure}
@@
 As another demonstration of our approach, we investigate the
-E.C. algorithm's ability to learn Boolean functions using only the logical
+EC algorithm's ability to learn Boolean functions using only the logical
 primitive NAND and the basic combinators. It is well known that a
-Boolean circuit can be constructed for any Boolean function given only
+boolean circuit can be constructed for any boolean function given only
 the NAND gate. Here, the basic combinators take the place of the
@@
 tasks, we sampled $1000$ Boolean circuits using AND, OR, and NOT
-gates. To accomplish this, we first sampled either 1, 2 or 3 inputs,
+gates. To accomplish this, we first sampled either 1, 2, 3 inputs,
 then between 1 and 5 gates, randomly wiring the inputs of each new
@@
 tasks consisting of $82$ unique Boolean functions, with a distribution
-of tasks as shown in Figure~\ref{fig:booldistr}(a). In
-Figure~\ref{fig:boolLearningCurvesCircuits}, we present learning
+of tasks as shown in Figure~\ref{fig:booldistr}~(a). In
+Figure~\ref{fig:boolLearningCurvesCircuits}, We present learning
 curves for frontier sizes 100, 500 and 10000, using the same procedure
@@
 \begin{figure}
-%%\vspace{-.3cm}
-\includegraphics[width=\linewidth]{figures/circuitDistrComp2.png}
-%%\vspace{-.6cm}
-\caption{Distribution of Boolean functions among (a) the set of 1000
+\vspace{-.3cm}
+\includegraphics[width=\linewidth]{figures/circuitDistrComp.pdf}
+\vspace{-.6cm}
+\caption{Distribution of boolean functions among (a) the set of 1000
   circuits randomly constructed from AND, OR and NOT gates (see text);
@@
   function.\label{fig:booldistr}}
-%%\vspace{-.4cm}
+\vspace{-.4cm}
 \end{figure}
@@
 also include two higher-order expression $E_1$ and $E_2$, to stress
-that the representation the E.C. algorithm is using allows it to build
+that the representation the EC algorithm is using allows it to build
 concepts that are more expressive that just sub-circuits.
@@
 three or fewer inputs as tasks. This is a more difficult
-problem. There are two kinds of learning that the E.C. algorithm might
+problem. There are two kinds of learning that the EC algorithm might
 accomplish: first, many expressions in $\mathcal{L}$ map to a single
@@
 second kind of structure. In
-Figure~\ref{fig:boolLearningCurvesUniform}, we show 10 runs of the
-E.C.  algorithm on the second experiment with a frontier size of
-$2000$: note how there are several local minima that the majority of
-the runs get stuck in with performance around $50\%$, but several of
-the runs seem to take different trajectories to more successful
-representations.
+Figure~\ref{fig:boolLearningCurvesUniform}, we show 10 runs of the EC
+algorithm on the second experiment with a frontier size of $2000$: note how there are several
+local minima that the majority of the runs get stuck in with performance
+around $50\%$, but several of the runs seem to take different
+trajectories to more successful representations.
 
@@
 \end{subfigure}
-\begin{subfigure}[Before]{0.9\linewidth}
-\includegraphics[width=1.0\linewidth]{figures/booleancircuits3frontiers.pdf}
+\begin{subfigure}[Before]{.9\linewidth}
+\includegraphics[width=.9\linewidth]{figures/booleancircuits3frontiers.pdf}
 \end{subfigure}
@@
 \begin{subfigure}[After]{.9\linewidth}
-\includegraphics[width=1.0\linewidth]{figures/booleanfunctionsUniform.pdf}
-
-\end{subfigure}
+\includegraphics[width=.9\linewidth]{figures/booleanfunctionsUniform.pdf}
+
+\end{subfigure}
+\vspace{-.3cm}
 \caption{Learning curves from Boolean function learning
@@
   of the trajectories get stuck in a local minimum around
-  $50\%$.\label{fig:boolLearningCurves}}
+  $50\%$.\label{fig:boolLearningCurves}
+
+}
+\vspace{-.5cm}
 \end{figure}
@@
 how bootstrap learning of abstract composable concepts can rapidly
-facilitate problem solving in the absence of domain specific
-knowledge. We show that this approach can learn successfully even in
-domains for which the solution space is not smooth and the error
-signal is all or none, where the only sense of locality to guide
-learning is the modular and compositional structure of the solutions
-that the algorithm itself builds.
-
-The problem presented here was chosen to highlight the vastness of the
-search space confronting an agent with minimal background knowledge:
-the reward function is binary, it is deterministic, and it cannot
-provide feedback about partial solutions. An important question left
-to be addressed is how to extend this algorithm to take into account
-continuous rewards, noisy data, and partial solutions.
+improve search performance.  We show that this approach can learn
+successfully even in domains for which the solution space is not
+smooth and the error signal is all or none, where the only sense of
+locality to guide learning is the modular and compositional structure
+of the solutions that the algorithm itself builds.
 
@@
 choices which should be explored further in future work.  Our claim
-here is not that the E.C. algorithm is optimal, but that it captures key
+here is not that the EC algorithm is optimal, but that it captures key
 features of how people solve the problem of learning new systems of
@@
 in these more human-like ways is a prime target for future research.
+%% \subsection{Digital arithmetic}
+%% Having demonstrated the behavior of the E-C algorithm on symbolic
+%% regression, we analyze its performance on tasks involving digital
+%% arithmetic, a class of problems with known modular structure. The
+%% general goal is to discover boolean circuits consisting of a limited
+%% set of gates that perform arithmetic operations on binary number
+%% representations.
+
+%% \section{Discussion}
+%% This work attempts to bring together ideas from various areas of
+%% research -- hierarchical Bayes, learning to learn, minimum description
+%% length learning -- to provide a proof of concept that learning
+%% abstractions from programs via compression can rapidly improve search
+%% performance. We emphasize this by focusing on domains in which the
+%% solution space is not smooth and the error signal is all or
+%% none. Thus, the main source of information for learning is the modular
+%% and compositional structure of the solution space.
+
+%% Of course, the representations that people use to learn in the world
+%% often have a partially smooth or local structure. Techniques in AI and
+%% machine learning commonly use smoothness or local structure to make
+%% search tractable. But the difference between this kind of learning and
+%% conceptual learning is undeniable. The former is what the statistician
+%% or machine learning researcher implements. The latter is what she
+%% learns as a student and as a practicioner. As researchers, we try
+%% different models, watch them fail or succeed, and abstract out the
+%% elements of these models that capture the relevant differences. We mix
+%% and match parts of different ideas, and utilize succcessful patterns
+%% of reasoning that can span many problems and domains.
+
+%% We believe that the primary characteristic of conceptual learning is
+%% that it occurs over a modular and compositional computational
+%% representation that supports higher-order functions. That is why we
+%% chose a functional programming language as the representational
+%% language of this work.
+
+%% The experiments in this work do not contain the richness of real world
+%% learning, but we believe that the learning algorithm's behavior
+%% exhibits some of the characteristics we would expect of a conceptual
+%% learner.
+
+%% One characteristic of this algorithm's behavior is a rapid initial
+%% increase in performance followed by a hard plateau. This plateau is
+%% expected given the circular nature of the algorithm: the learner
+%% inspects the region of the solution space that is typical of solutions
+%% already found. We see in this phenomenon a conceptual demarcation of
+%% the problem space: a coherent set of concepts or problems solving
+%% techniques can define by their power the extent of problems a problem
+%% solver is willing to consider. One might hope, however, that an
+%% intelligent learner would recognize that such a demarcation has
+%% occured and attempt to find another class of concepts for the
+%% remaining unsolved tasks.
+
+%% Another aspect of conceptual learning that we see present in this
+%% algorithm's behavior is the importance of a curriculum. By this we do
+%% not mean an explicit ordering of the problems in the task set, but
+%% rather that a set of sufficiently simple problems must act as stepping
+%% stones to the acquisition of complex representations.
 
 %% The file named.bst is a bibliography style file for BibTeX 0.99c
-\bibliographystyle{named}
+\bibliographystyle{plain}
 \bibliography{ijcai13}

### 2014-06-08 old/drafts/mca/ModularityCompositionalityAbstraction.tex
Moved everything into an a new 'old' directory.

\documentclass{article}

\begin{document}
Much has been made of the importance of notions of modularity,
compositionality, and abstraction in representations of knowledge and
computation, especially as it relates to human cognition.

Modularity is a particular type of relationship between the function
of a system and its structure. A modular system is one in which
various functional requirement of a system are structurally
separated. A paradigmatic example of a modular system is the human
body. The body has various subsystems that perform individual
important tasks: there is a cardiovascular system, an immune system, a
liver, a kidney, a brain, etc. Each one of these systems performs an
individual functions. Small changes to the immune system results in
changes in immune function and small changes in the cardiovascular
system results in changes in cardiovascular structure. And, of course,
it is not the case that changes to one system can't change other
functional aspects of the system; but it seems necessary that the
change be relatively small compared to the change in the associated
functional sytem. In this sense, modularity seems to be analagous to a
linear transformation. Take the structural description of a system to
be a point in some high-dimensional space, where each dimension
corresponds to the value of some structural component. And let the
functional description of a system be another point in a different
high-dimensional space. Then there exists some function between any
two corresponding points. The claim that the system is modular is akin
to saying that the transformation between the structural and
functional descriptions is locally a linear transformation.

The fact that whether a system is modular depends on the mapping
between its structural description and its functional description
means that a system is modular with respect to the granularity of the
functional description. For example, if a car is system that produces
speed given some inputs, then on inspecting the system, we have no
idea whether it is modular or not.

\end{document}

### 2014-06-08 old/drafts/theory/theory.tex
Moved everything into an a new 'old' directory.

\documentclass{article}
\usepackage{amsmath}
\begin{document}

Consider the problem of learning boolean functions of $N$ variables.
Suppose we are given an oracle that, when handed such a function, tells us if it is correct or not.
Then, the size of the space of possible functions is $2^{2^N}$.

Supppose there exists a factorization of the target function, called a \emph{curriculum}.
The \emph{curriculum} is a set of $C$ boolean functions of arity $d$ such that the target function may be written only in terms of up to $d$ members of the set consisting of the union of the $N$ variables and the $C$ curriculum functions.

Now we'll calculate an upper bound on the work required to identify any such factorizations.
First, there must be at least one curriculum function that is only a set of the $N$ variables.
There are $\binom{N}{d}$ ways of choosing the input variables and $2^{2^d}$ ways of constructing functions from those variables, so there are $\binom{N}{d}2^{2^d}$ different functions to consider.
If we are given corresponding oracles for all $C$ members of the curriculum, then we will need at most $C\binom{N}{d}2^{2^d}$ queries to such oracles.
Similarly, the curriculum function which is a function of $d$ members of the $N$ variables and the previous curriculum function requires at most $(C-1)\binom{N+1}{d}2^{2^d}$ queries.
Continuing this reasoning, we get the following upper bound on the number of oracle queries required to identify the function of interest given the curriculum:
$$
\binom{N+C}{d}2^{2^d} + \sum_{j=0}^{C-1} (C-j) \binom{N+j}{d} 2^{2^d}
$$
where the first term comes from the oracle queries required to reconstruct the taraget function after learning the curriculum.

This analysis applies in the case where we don't know what order to learn the curriculum in. This might be unrealistic for human learners.
However, it is the setting EC uses.

One problem with the previous analysis is that it does not reflect the fact that there are multiple factorizations, and for each of them, the value of $d$ is a function of $C$.
One nice property of the analysis is that it gives us an upper bound on what $d$ should be given $C$: it should be less than the value which makes the bound equal to $2^{2^N}$, because otherwise, the curriculum doesn't help us.

We'd like to show that the reduction in the search space, for certain classes of factorizations, is large (exponential?) in $C$.
One approach would be to analyze the change in $d$ when $C$ is decreased by one.
If $C'=C-1$ then $d\leq d'\leq 2d-1$.
But, most of the nodes don't have their degree changed, so it seems that we ought to change out analysis to reflect that fact.

Another (probably more difficult) problem to analyze is that of discovering the curriculum as one learns.
This is closer to what EC does.
\end{document}

### 2014-06-08 old/drafts/nips/figures/curves.py

@@ -1,0 +1,17 @@
+
+import os, sys
+from matplotlib import pyplot as plt
+
+
+# frontier 5000
+ntasks = 12
+xs = range(1, 6)
+solved = [5,7,8,8,8]
+productions = [0, 10, 13, 13, 14]
+
+fig = plt.figure()
+ax = fig.add_subplot(111)
+ax.set_title("Learning over iterations.")
+ax.plot(xs, solved)
+ax.set_ylim(0, 12)
+

### 2014-06-08 old/src/IterativeDeepening.hs

@@ -1,0 +1,1 @@
+kdjlksjf

### 2014-06-08 old/src/Strappy/Alternatives/GeneticProgramming.hs

@@ -1,0 +1,97 @@
+
+
+module Strappy.Alternatives.GeneticProgramming where
+
+-- We implement a typed genetic programming search over our expressions. 
+-- The steps we need are 1) initialization of population, 2) selecting best
+-- n% of the population for reproduction, 3) generation of a new population
+-- via crossover and mutation. 
+--
+import Prelude hiding (flip)
+import Control.Monad.Random
+import Control.Monad
+
+import Strappy.Core
+
+data GPconfig = GPconfig {gpPopulatonSize :: Integer,
+                          gpPercentMutation :: Double,
+                          gpMaxMutationDepth :: Integer,
+                          gpInitDepth :: Integer} deriving Show
+
+initPopulation :: (MonadRandom m, MonadPlus m) =>
+                [Expr] -- primitive set 
+                -> Integer -- population size
+                -> Type 
+                -> m (ExprMap (Double, Sub)) -- population
+-- | Initialize population 
+initPopulation primitives populationSize tp 
+    = sampleExprsWithContexts (fromIntegral populationSize) gr tp
+    where gr = normalizeGrammar $ Grammar 3 (mkExprDistr primitives)
+   
+
+chooseRandomNode :: (MonadPlus m, MonadRandom m) => Expr -> m (Expr, Path)
+chooseRandomNode e@App{eLeft=l, eRight=r} 
+    = do shouldChooseItself <- flip (0.9 :: Double)
+         if not shouldChooseItself   
+            then do shouldChooseLeft <- flip (0.5 :: Double)
+                    if shouldChooseLeft 
+                      then do (l', path) <- chooseRandomNode l
+                              return (l' <> r, L : path) 
+                      else do (r', path) <- chooseRandomNode r
+                              return (l <> r', R : path) 
+            else return (cBottom, [])
+chooseRandomNode e@Term{} = return (cBottom, [])
+
+data Dir = L | R deriving (Show, Eq)
+type Path =  [Dir] 
+mutateExpr :: (MonadPlus m, MonadRandom m) 
+            => [Expr] -- ^ primitive set
+            -> Expr -- ^ expr to mutate
+            -> m Expr
+mutateExpr primitives expr = do (expr', path) <- chooseRandomNode expr
+                                replaceEmptyNode expr' path 
+
+      where 
+            gr = normalizeGrammar $ Grammar 3 (mkExprDistr primitives)
+    
[…]

### 2014-06-08 old/src/Strappy/Experiments/ObjectDetection/CV.hs

@@ -1,0 +1,9 @@
+
+module Strappy.Experiments.ObjectDectection.CV where
+
+-- global imports
+
+
+-- local imports
+import Strappy.Core
+

### 2014-06-08 old/src/Strappy/Experiments/Words/Corpus.hs

@@ -1,0 +1,14 @@
+
+module Strappy.Experiments.Words.Corpus where
+
+import Data.Char
+
+
+tokenize :: String -> [String]
+-- | basic tokenizer: removes punctuation except for apostrophes and makes everything lowercase
+tokenize txt = let ls = lines txt
+                   clean = map (\x -> if x `elem` "!.,-\"\'!@#$%^&*()_+=?<>{}[];/" then ' ' else x) 
+               in (concatMap words . map (map toLower) . map clean) ls
+
+--main = do txt <- readFile "cedrick.txt"
+--          putStrLn $ show (tokenize txt)

### 2014-06-08 old/src/Strappy/Experiments/Words/Likelihood.hs

@@ -1,0 +1,12 @@
+
+module Strappy.Experiments.Words.Likelihood where
+
+import Text.EditDistance
+
+levenshteinLogLikelihood :: Double -> String -> String -> Double
+-- | Return a log likelihood score that is based on the negative edit 
+-- | distance between the input strings. This value is then scaled by 
+-- | the factor beta.
+levenshteinLogLikelihood beta string1 string2 = 
+	negate . (* beta) . fromIntegral $ levenshteinDistance defaultEditCosts string1 string2
+

### 2014-06-08 old/src/Strappy/Experiments/Words/Run.hs

@@ -1,0 +1,46 @@
+module Main where
+
+import qualified Data.Map as Map
+
+import Strappy.Config
+import Strappy.Planner
+import Strappy.Utils
+import Strappy.Library
+import Strappy.Core
+import Strappy.Sample
+import Strappy.Type
+import Strappy.Expr
+
+import Strappy.Experiments.Words.Grammar
+import Strappy.Experiments.Words.Tasks
+import Strappy.Experiments.Words.Corpus
+
+
+{-doEMPlan :: Eq a, Ord a, Show a =>
+            [PlanTask a] -- ^ Tasks
+            -> (PlanTask a -> PlanTask a -> Bool) -- ^ Partial order among tasks: is the first task harder than the second task?
+            -> Double -- ^ Lambda
+            -> Double -- ^ pseudocounts
+            -> Int -- ^ frontier size
+            -> Int -- ^ number of plans sampled per task
+            -> Int -- ^ max length of each plan
+            -> Grammar -- ^ Initial grammar
+            -> IO Grammar -- ^ Improved grammar
+-}
+
+beta = 3
+
+runWordsEM :: [String] -> IO ()
+runWordsEM corpus = do let tasks = map (string2PlanTask beta) corpus 
+                           filename = "out_test.txt"
+                       loopM wordsGr [0..5] $ \gr step -> do
+                        putStrLn $ "EM Planning Iteration: " ++ show step
+                        (gr', numHitTasks) <- doEMPlan (Just filename) tasks 2.0 2.0 frontierSize numberOfPlansPerTask maximumPlanLength gr 
+                        return gr'
+                       return ()
+
+main = 
+       do txt <- readFile "alice.txt"
+          let tokens = take 100 $ tokenize txt 
+          print tokens
+          runWordsEM tokens

### 2014-06-08 old/src/Strappy/Experiments/Words/Tasks.hs

@@ -1,0 +1,25 @@
+{-# Language ScopedTypeVariables #-}
+
+module Strappy.Experiments.Words.Tasks where
+
+import Control.Exception
+import Debug.Trace
+
+import Strappy.Expr
+import Strappy.Type
+import Strappy.Planner
+import Strappy.Library
+
+import Strappy.Experiments.Words.Grammar
+import Strappy.Experiments.Words.Likelihood
+
+string2PlanTask :: Double    -- beta 
+                -> String   
+                -> PlanTask 
+-- | Returns a PlanTask whose target string is the input string. *beta* scales
+--the loglikelihood scoring function. Larger *beta* results in a steeper
+--scoring function; lower *beta* results in a shallower one. 
+string2PlanTask beta string = PlanTask string loglikelihood (tList tChar) cEmpty
+    where loglikelihood' expr = levenshteinLogLikelihood beta string (eval expr :: String)
+          loglikelihood expr = let x = loglikelihood' expr in 
+          						catch (x `seq` return x) $ \(err :: SomeException) -> return (-10000) 

### 2014-06-08 old/src/Strappy/Experiments/Words/watchRunHeap.sh

@@ -1,0 +1,11 @@
+ #!/bin/sh
+  head -`fgrep -n END_SAMPLE Run.hp | tail -1 | cut -d : -f 1` Run.hp \
+    | hp2ps > FOO.ps
+  gv Run.ps &
+  gvpsnum=$!
+  while [ 1 ] ; do
+    sleep 10
+    head -`fgrep -n END_SAMPLE Run.hp | tail -1 | cut -d : -f 1` Run.hp \
+      | hp2ps > Run.ps
+    kill -HUP $gvpsnum
+  done

### 2014-06-08 old/src/Strappy/TestNumber.hs

@@ -1,0 +1,26 @@
+
+
+-- Author: Eyal Dechter
+-- Date : 5/29/2014
+
+--module Strappy.TestNumber where 
+
+import Strappy.Number hiding (main)
+import Strappy.EnumBF
+import Strappy.EnumBits
+import Strappy.Type
+import Strappy.Library
+import System.Environment
+
+main = do
+    args <- getArgs
+    let enumType = head args
+    let frontierSize = read $ args !! 1
+    if enumType == "enumBF" then do
+        putStrLn "Enumeration type: enumBF"
+        print $ enumBF seedGrammar frontierSize tInt
+                            else do 
+        putStrLn "Enumeration type: enumBits"
+        print $ enumBits seedGrammar frontierSize tInt
+
+    putStrLn "Testing enumeration from seed grammar used in Number experiment."

### 2014-06-08 old/src/Strappy/Tests/Expr.hs

@@ -1,0 +1,11 @@
+
+module Strappy.Tests.Expr where
+
+--import Test.Framework (defaultMain, testGroup)
+--import Test.Framework.Providers.HUnit
+--import Test.Framework.Providers.QuickCheck (testProperty)
+
+import Test.QuickCheck
+--import Test.HUnit
+
+

### 2014-06-08 old/src/Strappy/Tests/LPCompress.hs

@@ -1,0 +1,11 @@
+
+module Strappy.Tests.LPCompress where
+
+import Strappy.ParseExpr
+import Strappy.Expr
+import Strappy.LPCompress
+import Strappy.Library
+
+e = annotateRequested $ parseExpr basicExprs "S * I"
+weightedCorpus = [(e, 1.0), (e, 1.0)]
+out = compressWeightedCorpus 1.0 1.0 basicGrammar weightedCorpus

### 2014-06-08 old/src/experiments/Experiments/CircuitsExperiment.hs

@@ -1,0 +1,81 @@
+-- BooleanCircuits.hs
+{-# Language ParallelListComp #-} 
+
+module Experiments.CircuitsExperiment where
+
+import Control.Monad.State
+import Data.Maybe
+import qualified Data.HashMap as HMap
+import Control.Monad.Random
+
+import BooleanCircuits
+import Type
+import CL
+import Expr
+import Task
+import StdLib
+import Data
+import qualified CombMap as CM
+import CombMap (CombMap)
+import Grammar
+import Routers 
+import PostProcess
+import EnumBF
+import ParseCL hiding (eval)
+import Run
+import DigArith
+import Experiment
+
+-- import Visualize
+
+
+
+lib = (CM.fromList $ 
+         [
+          ("I", cI)
+--         , ("True", cTrue)
+--        , ("False", cFalse)
+         ,("nand", cNand)
+--         , ("&", cAnd)
+--         , ("not", cNot)
+--         , ("pair", cPair)
+--         , ("fst", cFst)
+--         , ("snd", cSnd)
+--         , ("triple", cTriple)
+--         , ("fst3", cFst3)
+--         , ("snd3", cSnd3)
+--         , ("thrd3", cThrd3)
+           ]) 
+
+lib1 = lib `CM.union` one_routers
+
+jitteredGrammarFromLib :: MonadRandom m => HMap.Map k Comb -> m Grammar
+jitteredGrammarFromLib lib = do 
+  vs <- getRandomRs (-1.0,1.0)
+  let lib' = CM.fromList [(c, 2 + k) | c <- CM.elems lib | k <- vs]
+  return $ normalizeGrammar $ Grammar lib' 0
+
+mkBoolCircuitsExperiment :: MonadRandom m => String 
+                          -> HMap.Map String Comb -- ^ lib
+                          -> Int -- ^ frontier size
+                          -> [Task]
+                          -> m Experiment
+mkBoolCircuitsExperiment name lib frontierSize taskSet
+    = do lib' <- jitteredGrammarFromLib lib
+         return $ Experiment {
+                      expName = name,
+                      expTaskSet = taskSet,
+                      expEps = 0,
+                      expPrior = lib',
+                      expInitLib = lib',
+                      expDepthBound=3,
+                      expNumBound = frontierSize,
+                      expReps=15}
+
+
+-- main :: IO ()
+-- main = do
[…]

### 2014-06-08 old/src/experiments/Experiments/DigitalArithmetic.hs

@@ -1,0 +1,188 @@
+-- DigitalArithmetic
+{-# Language ParallelListComp #-} 
+
+module Experiments.DigitalArithmetic where
+
+import Control.Monad.State
+import Data.Maybe
+import qualified Data.HashMap as HMap
+import Control.Monad.Random
+import Debug.Trace
+-- import Data.GraphViz
+
+import Experiment
+import Type
+
+import Type
+import CL
+-- import Search
+import Expr
+import Task
+import StdLib
+import Data
+import qualified CombMap as CM
+import CombMap (CombMap)
+import Grammar
+import Routers 
+import PostProcess
+import EnumBF
+import ParseCL hiding (eval)
+import Run
+import DigArith
+-- import Visualize
+
+
+taskIncr1 = Task "pair incr 1"
+            f
+            ((TAp tPair tBool) ->- (TAp tPair tBool))
+            where f c = if all (==True) [ismatch c i j | (i, j) <- points] then 0 else 1
+                  pairIn0 = intTupleToPair (0, 0)
+                  pairOut0 = intTupleToPair (0, 1)
+                  pairIn1 = intTupleToPair (0, 1)
+                  pairOut1 = intTupleToPair (1, 0)
+                  pairIn2 = intTupleToPair (1, 0)
+                  pairOut2 = intTupleToPair (1, 1)
+                  points = [(pairIn0, pairOut0),
+                            (pairIn1, pairOut1),
+                            (pairIn2, pairOut2)]
+                  ismatch c cin cout = pairEq (app' c cin) cout
+
+taskNot = Task "not" f (tBool ->- tBool)
+    where f c = if a &&  b then  0 else 1
+              where
+                a = reduceComb (app' c cTrue) == reduceComb cFalse
+                b = reduceComb (app' c cFalse) == reduceComb cTrue
+
+taskAnd = Task "and" f (tBool ->- tBool ->- tBool)
+    where f c = if a1 && a2 && a3 && a4  then  0 else 1
+              where
+                a1 = reduceComb (app' (app' c cTrue) cTrue) == reduceComb cTrue
+                a2 = reduceComb (app' (app' c cTrue) cFalse) == reduceComb cFalse
+                a3 = reduceComb (app' (app' c cFalse) cTrue) == reduceComb cFalse
+                a4 = reduceComb (app' (app' c cFalse) cFalse) == reduceComb cFalse
+
[…]

### 2014-06-08 old/src/experiments/Experiments/EX1.hs

@@ -1,0 +1,188 @@
+-- EX1
+-- Symbolic regression 10 run per frontier size experiment on EC2. 
+ {-# Language ParallelListComp #-}
+
+module Experiments.EX1 where
+
+import System.Environment (getArgs)
+import Control.Monad.Random
+import Control.Monad.State 
+import qualified  Data.HashMap as HMap
+
+import Data.Maybe
+import Debug.Trace
+
+import Experiment
+import Type
+
+import Type
+import CL
+import Expr
+import Task
+import StdLib
+import Data
+import qualified CombMap as CM
+import CombMap (CombMap)
+import Grammar
+import Routers 
+import PostProcess
+import EnumBF
+import ParseCL hiding (eval)
+import Run
+
+rng = [0..10]
+eval :: [Comb] -> Comb -> Double
+eval ds c = fromIntegral $ sum $ [(toInteger $ diff i) | i <- rng]
+         where diff i = abs $ dat i - eval i
+               eval i = if y < 0 then maxBound else y 
+                   where y = case reduceComb $ (CApp c (num2C i) tInt 0) of
+                               (N y) -> y
+                               otherwise -> maxBound
+               dat i = y where (N y) = reduceComb (ds !! i)
+
+quad a b c = \i -> a * i^2 + b * i + c
+showQuad a b c = show a ++ " x^2 + " ++ show b ++ " x + " ++ show c
+mkQuadTask a b c = Task (showQuad a b c) f (tInt ->- tInt)
+    where f = eval [num2C $ (quad a b c i)| i<-rng]
+
+mkQuadSymRegTask a b c = SymRegTask (showQuad a b c) f (tInt ->- tInt)
+    where f  = [quad a b c i | i <- [0..10]]
+
+squares a b = \i -> (a * i + b)^2
+showSquares a b = "(" ++ show a ++ " x + " ++ show b ++ ")^2"
+mkSquaresTask a b = Task (showSquares a b) f (tInt ->- tInt)
+                   
+    where f = eval [num2C $ (squares a b) i| i<-rng]
+
+-- taskSet' = --[mkSquaresTask i j | i <- [0..5] , j <- [0..5]]
+--            -- ++  (map (mkSingleEqualityTask 100) $ [i*i | i <- [0..20]])
+--            [mkQuadTask i j k | i <- [0..10] , j <- [0..10], k <- [0..10]]
[…]

### 2014-06-08 old/src/experiments/Experiments/EX2.hs

@@ -1,0 +1,26 @@
+-- EX2.hs
+
+module Experiments.EX2 where
+
+import Run
+import Experiments.CircuitsExperiment 
+import BooleanCircuits
+
+
+main :: IO ()
+main = do 
+  tasks <- readTasksFromCircuits "data/EX2/circuitTasks.txt"
+  exp1000 <- mkBoolCircuitsExperiment "ex2_circuitlearning_1000"
+               lib1
+                1000
+                tasks
+  exp5000 <- mkBoolCircuitsExperiment "ex2_circuitlearning_5000"
+               lib1
+                5000
+                tasks
+  exp10000 <-mkBoolCircuitsExperiment "ex2_circuitlearning_10000"
+               lib1
+                10000
+                tasks
+  sequence $ map (\x -> runExp x "data/EX2") [exp1000, exp5000, exp10000]
+  return ()

### 2014-06-08 old/src/experiments/Experiments/EX3.hs

@@ -1,0 +1,30 @@
+-- EX3.hs boolean circuits (snd try) difference from EX2: in EX2 the
+-- circuit picking algorithm was biased towards very simple circuits.
+-- This has been fixed. 
+
+-- EX3.hs
+
+module Experiments.EX3 where
+
+import Run
+import Experiments.CircuitsExperiment 
+import BooleanCircuits
+
+
+main :: IO ()
+main = do 
+  tasks <- readTasksFromCircuits "data/EX3/circuitTasks.txt"
+  exp100 <- mkBoolCircuitsExperiment "ex3_circuitlearning_100"
+               lib1
+                100
+                tasks
+  exp500 <- mkBoolCircuitsExperiment "ex3_circuitlearning_500"
+               lib1
+                500
+                tasks
+  exp1000 <-mkBoolCircuitsExperiment "ex3_circuitlearning_1000"
+               lib1
+                1000
+                tasks
+  sequence $ map (\x -> runExp x "data/EX2") [exp100, exp500, exp1000]
+  return ()

### 2014-06-08 old/src/experiments/Experiments/EX5.hs

@@ -1,0 +1,171 @@
+-- EX5 
+-- symbolic regression with progressively more impoverished task sets
+{-# Language ParallelListComp #-}
+
+
+module Experiments.EX5 where
+
+import System.Environment (getArgs)
+import Control.Monad.Random
+import Control.Monad.State 
+import qualified  Data.HashMap as HMap
+
+import Data.Maybe
+import Debug.Trace
+
+import Experiment
+import Type
+
+import Type
+import CL
+import Expr
+import Task
+import StdLib
+import Data
+import qualified CombMap as CM
+import CombMap (CombMap)
+import Grammar
+import Routers 
+import PostProcess
+import EnumBF
+import ParseCL hiding (eval)
+import Run
+
+rng = [0..10]
+eval :: [Comb] -> Comb -> Double
+eval ds c = fromIntegral $ sum $ [(toInteger $ diff i) | i <- rng]
+         where diff i = abs $ dat i - eval i
+               eval i = if y < 0 then maxBound else y 
+                   where y = case reduceComb $ (CApp c (num2C i) tInt 0) of
+                               (N y) -> y
+                               otherwise -> maxBound
+               dat i = y where (N y) = reduceComb (ds !! i)
+
+quad a b c = \i -> a * i^2 + b * i + c
+showQuad a b c = show a ++ " x^2 + " ++ show b ++ " x + " ++ show c
+mkQuadTask a b c = Task (showQuad a b c) f (tInt ->- tInt)
+    where f = eval [num2C $ (quad a b c i)| i<-rng]
+
+mkQuadSymRegTask a b c = SymRegTask (showQuad a b c) f (tInt ->- tInt)
+    where f  = [quad a b c i | i <- [0..10]]
+
+squares a b = \i -> (a * i + b)^2
+showSquares a b = "(" ++ show a ++ " x + " ++ show b ++ ")^2"
+mkSquaresTask a b = Task (showSquares a b) f (tInt ->- tInt)
+                   
+    where f = eval [num2C $ (squares a b) i| i<-rng]
+
+taskSetNoConstants = [mkQuadSymRegTask i j k | i <- [1..9] , j <- [1..9], k <- [0..9]]
+                     ++ [mkQuadSymRegTask i j k | i <- [1..9] , j <- [0], k <- [0..9]]
[…]

### 2014-06-08 old/src/experiments/Experiments/ListExperiment.hs

@@ -1,0 +1,82 @@
+-- ListExperiment.hs
+
+module Experiments.ListExperiment where 
+
+import Control.Monad
+import Debug.Trace
+
+import Expr
+import StdLib
+import Data
+import CL
+import CombMap (CombMap)
+import qualified CombMap as CM
+import Task
+import Type
+import Experiment
+
+
+-- Basic List Experiment --
+
+
+mkIntList :: [Int] -> SynthComb
+mkIntList (x:xs) = Right cCons <:> Right (num2C x) <:> mkIntList xs
+mkIntList [] = Right cEmpty
+
+maybe2Either :: Maybe a -> Either String a
+maybe2Either (Just a) = Right a
+maybe2Either Nothing = Left ""
+
+fromRight (Right x) = x
+
+tLast = Task "last" f tp
+    where tp = (TAp tList tInt) ->- (TAp tList tInt)
+          f comb = reward
+              where ex = liftM comb2Expr $ Right comb <:> (mkIntList [1,2,3,4]) 
+                    result = reduce (fromRight ex)
+                    reward = case result of 
+                               (N 2) -> 0
+                               otherwise -> 1
+listlib = CM.fromList $ 
+          [ ("I", cI)
+          , ("S", cS)
+          , ("B", cB)
+          , ("C", cC)
+        , ("cond", cCond)
+        , ("True", cTrue)
+        , ("False", cFalse)
+        , ("|", cOr)
+        , ("&", cAnd)
+        , ("not", cNot)
+--         , ("<", cLT)
+--         , (">", cGT)
+--         , ("==", cEQ)
+          , ("+", dOp2C "+" (+))
+--          , ("-", dOp2C "-" (-))
+           , ("*", dOp2C "*" (*))
+          ,  ("0", num2C 0)
+          ,  ("1", num2C 1)
+--          , ("2", num2C 2)
+--         , ("3", num2C 3)
+--         , ("4", num2C 4)
+--           , ("primRec", cPrim) 
+          , (":", cCons)
+          , ("[]", cEmpty)
+           , ("head", cHead)
+           , ("tail", cTail)
+           , ("isEmpty", cIsEmpty)
+           ]
+
+
+
+expTList1 = Experiment {
+              expName = "List function: first",
+              expTaskSet = [tLast],
+              expEps = 0,
+              expPrior = stdgrammar,
+              expInitLib = stdgrammar,
+              expDepthBound = 2,
+              expNumBound = 10,
+              expReps = 1}
+                      
+                   

### 2014-06-08 old/src/experiments/Experiments/SymbolicRegression.hs

@@ -1,0 +1,187 @@
+-- SymbolicRegression.hs
+{-# Language ParallelListComp #-}
+
+module Experiments.SymbolicRegression where
+
+import System.Environment (getArgs)
+import Control.Monad.Random
+import Control.Monad.State 
+import qualified  Data.HashMap as HMap
+
+import Data.Maybe
+import Debug.Trace
+
+import Experiment
+import Type
+
+import Type
+import CL
+import Expr
+import Task
+import StdLib
+import Data
+import qualified CombMap as CM
+import CombMap (CombMap)
+import Grammar
+import Routers 
+import PostProcess
+import EnumBF
+import ParseCL hiding (eval)
+import Run
+
+rng = [0..10]
+eval :: [Comb] -> Comb -> Double
+eval ds c = fromIntegral $ sum $ [(toInteger $ diff i) | i <- rng]
+         where diff i = abs $ dat i - eval i
+               eval i = if y < 0 then maxBound else y 
+                   where y = case reduceComb $ (CApp c (num2C i) tInt 0) of
+                               (N y) -> y
+                               otherwise -> maxBound
+               dat i = y where (N y) = reduceComb (ds !! i)
+
+quad a b c = \i -> a * i^2 + b * i + c
+showQuad a b c = show a ++ " x^2 + " ++ show b ++ " x + " ++ show c
+mkQuadTask a b c = Task (showQuad a b c) f (tInt ->- tInt)
+    where f = eval [num2C $ (quad a b c i)| i<-rng]
+
+mkQuadSymRegTask a b c = SymRegTask (showQuad a b c) f (tInt ->- tInt)
+    where f  = [quad a b c i | i <- [0..10]]
+
+squares a b = \i -> (a * i + b)^2
+showSquares a b = "(" ++ show a ++ " x + " ++ show b ++ ")^2"
+mkSquaresTask a b = Task (showSquares a b) f (tInt ->- tInt)
+                   
+    where f = eval [num2C $ (squares a b) i| i<-rng]
+
+taskSet' = --[mkSquaresTask i j | i <- [0..5] , j <- [0..5]]
+           -- ++  (map (mkSingleEqualityTask 100) $ [i*i | i <- [0..20]])
+           [mkQuadTask i j k | i <- [0..10] , j <- [0..10], k <- [0..10]]
+--           ++ [mkQuadTask i j k | i <- [0..0] , j <- [0..5], k <- [0..5]]
+
[…]

### 2014-06-08 old/src/experiments/UGM/parse_xbif.py

@@ -1,0 +1,56 @@
+
+import os
+from os.path import join, exists
+import os.path
+
+import xml.etree.ElementTree as ET
+import pydot 
+import itertools
+
+def parse_xbif(filename):
+	assert(exists(filename))
+	tree = ET.parse(filename)
+	root = tree.getroot()
+
+	variable_nodes = root[0].findall("./VARIABLE")
+	prob_nodes = root[0].findall("./PROBABILITY")
+
+	variables = [v.find("./NAME").text for v in variable_nodes]
+
+	edges = []
+	for p in prob_nodes:
+		v = p.find("./FOR").text
+		pars = p.findall("./GIVEN")
+		edges.extend([(v, par.text) for par in pars])
+
+	graph = pydot.graph_from_edges(edges, directed=True)
+	fout = open(os.path.splitext(os.path.basename(filename))[0] +".pdf", "w")
+	print >>fout, graph.create_pdf()
+	fout.close()
+
+	return variables, edges
+
+def parse_uai(fname):
+	assert(exists(filename))
+
+	lines = open(filename, "r").readlines()
+	nvars = int(lines[1].strip())
+	ncliques = int(lines[3].strip())
+
+	edges = []
+	for line in lines[4:(ncliques+4)]:
+		line = line.strip()
+		vs = [int(s) for s in line.split() if s.isdigit()]
+		edges.extend(list(itertools.combinations(vs, 2)))
+
+	graph = pydot.graph_from_edges(edges, directed=False)
+	fout = open(os.path.splitext(os.path.basename(fname))[0] +".pdf", "w")
+	print >>fout, graph.create_pdf()
+	fout.close()
+	return graph
+
+
+if __name__ == "__main__":
+	filename="uai/rbm_20.uai"
+
+	graph = parse_uai(filename)

### 2014-06-08 old/src/experiments/matlab/plotNumHits.m

@@ -1,0 +1,8 @@
+function plotNumHits(files)
+
+f = figure; hold on;
+for file = files(:)
+    file
+    d = load(file);
+    plot d;
+end

### 2014-06-08 old/src/Strappy/Number.hs
Deleted files and Mmoved a bunch of stuff around.

@@ -1,129 +1,0 @@
-module Strappy.Number where
-
-import Strappy.Expr
-import Strappy.Type
-import Strappy.Utils
-import Strappy.Library
-import Strappy.BeamSearch
-import Strappy.Planner
-import Strappy.EM
-import Strappy.ProdFeat
-import Strappy.Sample
-import Strappy.EnumBits
-import Strappy.Response
-
-import qualified Data.Map as Map
-import System.Environment
-import System.Random
-import Data.Maybe
-import Control.Monad
-import Debug.Trace
-import Data.List
-import Data.Array
-
--- | TASK SPECIFIC UTILITIES | --
-
-xHack = 1
-oHack = 2
-
-logInt = log . fromIntegral
-
-world2Expr :: [Int] -> Expr
-world2Expr = intListToExpr
-
-makeBag :: Int -> Int -> [Int]
-makeBag x o = (replicate x xHack) ++ (replicate o oHack)
-
-checkTask :: String      -- ^ determiner string
-          -> String      -- ^ noun string
-          -> Int         -- ^ number of x's
-          -> Int         -- ^ number of o's
-          -> Maybe [Int] -- ^ output of the task
-          -> Bool        -- ^ whether task output is correct
-checkTask _ _ _ _ Nothing = False
-checkTask "all" noun x o (Just result) =
-    case noun of
-        "X"     -> result == replicate x xHack
-        "O"     -> result == replicate o oHack
-        "thing" -> (length $ filter (== xHack) result) == x &&
-                   (length $ filter (== oHack) result) == o &&
-                   (length result) == (x+o)
-        _       -> False
-checkTask "one" noun x o (Just result) =
-    case noun of
-        "X"     -> result == [xHack]
-        "O"     -> result == [oHack]
-        "thing" -> result == [xHack] || result == [oHack]
-        _       -> False
-checkTask "two" noun x o (Just result) =
-    case noun of
-        "X"     -> result == [xHack,xHack]
-        "O"     -> result == [oHack,oHack]
-        "thing" -> result == [xHack,xHack] || result == [oHack,oHack] ||
-                   result == [xHack,oHack] || result == [oHack,xHack]
-        _       -> False
-
-checkSuperTask :: [(String, String, Int, Int, Maybe [Int])] -> Int
-checkSuperTask result = sum $ map (\(det,noun,x,o,eval) -> bool2Binary $ checkTask det noun x o eval) result
-
--- | TASKS | --
[…]

### 2014-06-08 old/src/Strappy/TestNumber.hs

@@ -1,26 +1,0 @@
-
-
--- Author: Eyal Dechter
--- Date : 5/29/2014
-
---module Strappy.TestNumber where 
-
-import Strappy.Number hiding (main)
-import Strappy.EnumBF
-import Strappy.EnumBits
-import Strappy.Type
-import Strappy.Library
-import System.Environment
-
-main = do
-    args <- getArgs
-    let enumType = head args
-    let frontierSize = read $ args !! 1
-    if enumType == "enumBF" then do
-        putStrLn "Enumeration type: enumBF"
-        print $ enumBF seedGrammar frontierSize tInt
-                            else do 
-        putStrLn "Enumeration type: enumBits"
-        print $ enumBits seedGrammar frontierSize tInt
-
-    putStrLn "Testing enumeration from seed grammar used in Number experiment."
